### 2020-06-21 pabm-notes.Rmd
first go at typing up results

---
title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph"
output: pdf_document
# output: html_document
# geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')

import::from(magrittr, `%>%`)
library(plot.matrix)

set.seed(314159)
```

```{r include = FALSE}
embedding <- function(A, p = NULL, q = NULL,
                      eps = 1e-6) {
  eigen.A <- eigen(A, symmetric = TRUE)
  if (is.null(p) | is.null(q)) {
    keep <- (abs(eigen.A$values) > eps)
  } else {
    keep <- c(seq(p), seq(n, n - q + 1))
  }

  U <- eigen.A$vectors[, keep]
  S <- diag(sqrt(abs(eigen.A$values[keep])))
  return(U %*% S)
}

draw.graph <- function(P) {
  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
  A[lower.tri(A)] <- 0
  diag(A) <- 0
  A <- A + t(A)
  return(A)
}

rot.mat.2 <- function(angle, axis = 1) {
  # rotation matrix for K=2

  if (axis %in% c(1, 3)) {
    rotation <- matrix(c(cos(angle), -sin(angle), sin(angle), cos(angle)),
                       nrow = 2, ncol = 2, byrow = TRUE)
    if (axis == 1) {
      return(as.matrix(Matrix::bdiag(1, rotation, -1)))
    } else {
      return(as.matrix(Matrix::bdiag(rotation, 1, -1)))
    }
  } else if (axis == 2) {
    return(matrix(c(cos(angle), 0, sin(angle), 0,
                    0, 1, 0, 0,
                    -sin(angle), 0, cos(angle), 0,
                    0, 0, 0, -1),
                  nrow = 4, ncol = 4,
                  byrow = TRUE))
  } else {
    stop(simpleError('axis must be 1, 2, or 3'))
  }
}

hyp.rot.mat.2 <- function(angle, axis = 1) {
  # hyperbolic rotation matrix for $K=2

  if (axis == 1) {
    return(matrix(c(1, 0, 0, 0,
                    0, 1, 0, 0, 
                    0, 0, cosh(angle), sinh(angle),
                    0, 0, sinh(angle), cosh(angle)), 
                  byrow = TRUE,
                  nrow = 4, ncol = 4))
  } else if (axis == 2) {
    return(matrix(c(1, 0, 0, 0,
                    0, cosh(angle), 0, sinh(angle),
                    0, 0, 1, 0,
                    0, sinh(angle), 0, cosh(angle)),
                  byrow = TRUE,
                  nrow = 4, ncol = 4))
  } else if (axis == 3) {
    return(matrix(c(cosh(angle), 0, 0, sinh(angle),
                    0, 1, 0, 0, 
                    0, 0, 1, 0, 
                    sinh(angle), 0, 0, cosh(angle)),
                  byrow = TRUE,
                  nrow = 4, ncol = 4))
  } else {
    stop(simpleError('axis must be 1, 2, or 3'))
  }
}

construct.Q <- function(angles) {
  rot.angles <- angles[1:3]
  hyp.angles <- c(angles[1], angles[4], -angles[4])
  rot.matrices <- lapply(seq(3), function(i) rot.mat.2(rot.angles[i], i))
  hyp.matrices <- lapply(seq(3), function(i) hyp.rot.mat.2(hyp.angles[i], i))
  matrices <- c(rot.matrices, hyp.matrices)
  return(Reduce(`%*%`, matrices))
}

obj.fun <- function(angles, Z.hat) {
  R <- construct.Q(angles)

  Z.candidate <- Z.hat %*% R
  B <- Z.candidate %*% t(Z.candidate)
  return(sum(apply(Z.candidate[, 1:2] ** 2, 1, min)) + 
           sum(apply(cbind((Z.candidate[, 3] - Z.candidate[, 4]) ** 2,
                           (Z.candidate[, 3] + Z.candidate[, 4]) ** 2), 
                     1, min)))
}

transform.embedding <- function(Z.hat) {
  init.angles <- runif(4, -pi, pi)
  angles.1 <- optim(init.angles, function(x) obj.fun(x, Z.hat), 
                    method = 'SANN')$par
  angles.2 <- optim(angles.1, function(x) obj.fun(x, Z.hat),
                    method = 'Nelder-Mead',
                    control = list(abstol = 1e-10, maxit = 1e6))$par
  R <- construct.Q(angles.2)
  return(Z.hat %*% R)
}
```

\newcommand{\diag}{\text{diag}}

### The popularity adjusted block model (PABM)[^pabm]

[^pabm]: https://arxiv.org/abs/1910.01931

Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
affinity matrix $A \in \{0, 1\}^{n \times n}$. Then $A$ is a random matrix 
with corresponding edge probability matrix $P$ such that 
$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ 
($A_{ji} = A_{ij}$ and $A_{ii} = 0$). Let there exist $K$ underlying 
communities in $G$, and let $n_{k}$ be the size of the $k$^th^ community in 
$G$ such that $\sum_{k=1}^K n_k = n$.

If $A$ and $P$ are organized such that $n_k \times n_l$ blocks $A^{(kl)}$ and 
$P^{(kl)}$ describe the edges and edge probabilities between communities $k$ 
and $l$, then $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$ for a set of 
fixed vectors $\{\lambda^{(st)}\}_{s, t = 1, ..., K}$. Each $\lambda^{(st)}$ 
for $s, t = 1, ..., K$ is a column vector of length $n_s$ (i.e., the community 
corresponding to the first index provides the vector length).

We will use the notation $A \sim PABM(\{\lambda^{(kl)}\}_K)$ to denote a random 
affinity matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
consisting of $K$ underlying clusters/communities.

### The generalized random dot product graph (GRDPG)[^grdpg]

[^grdpg]: https://arxiv.org/abs/1709.05506

Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
graph $G$. $X$ consists of row vectors $x_i^\top$. Let 
$A \in \{0, 1\}^{n \times n}$ be the corresponding affinity matrix.

Fix $p$, $q$ such that $p + q = d$ and define 
$I_{pq} = \begin{bmatrix} I_p & 0 \\ 0 & -I_q \end{bmatrix}$.

Then $G = (V, E)$ is a generalized random dot product graph with signature 
$(p, q)$ and patent positions $X$ iff its random affinity matrix can be 
described as $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ where 
$P_{ij} = x_i^\top I_{pq} x_j$.

We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random affinity 
matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.

### Connecting the PABM to the GRDPG for $K = 2$

Let $X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
\end{bmatrix}$ and 
$Y = \begin{bmatrix} \lambda^{(11)} & 0 & \lambda^{(12)} & 0 \\
0 & \lambda^{(21)} & 0 & \lambda^{(22)} \end{bmatrix}$. 
Then $P = X Y^\top$.

We can note that $Y = X \Pi$ where $\Pi$ is the permutation matrix 
$\Pi = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0 & 0 & 1 & 0 \\ 
0 & 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}$. 
Therefore, $P = X \Pi X^\top$.

Taking the spectral decomposition of $\Pi = U D U^\top$, we can see that 
$P = (X U) D (X U)^\top$. We can then denote $\Sigma = |D|^{1/2}$, the 
square root of the absolute values of the (diagonal) entries of $D$ and 
obtain $P = (X U \Sigma) I_{pq} (X U \Sigma)^\top$ where $p$ and $q$ correspond 
to the number of positive and negative eigenvalues of $\Pi$, respectively. 
Therefore, the PABM with $K = 2$ is a special case of the GRDPG. We can however 
expand upon this a bit further.

The permutation described by $\Pi$ has two fixed points and one cycle of order 
2. The two fixed points are at positions $1$ and $4$, so $\Pi$ has two 
eigenvalues equal to $1$ and corresponding eigenvectors $e_1$ and $e_4$. The 
cycle of order 2 switching positions $2$ and $3$ corresponds to eigenvalues 
$1$ and $-1$ with corresponding eigenvalues $(e_2 + e_3) / \sqrt{2}$ and 
$(e_2 - e_3) / \sqrt{2}$ respectively. Therefore, 
$D = \begin{bmatrix} 
1 & 0 & 0 & 0 \\ 
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & -1
\end{bmatrix} = I_{3, 1}$ and 
$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix}$.

Putting it all together, we get $P = (X U) I_{3, 1} (X U)^\top$. Therefore, 
the PABM with $K = 2$ is a GRDPG with $p = 3$, $q = 1$, $d = K^2 = 4$, and 
latent positions 
$X U = \begin{bmatrix} 
\lambda^{(11)} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(12)} / \sqrt{2} \\
0 & \lambda^{(22)} & \lambda^{(21)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2}
\end{bmatrix}$.

### Generalization to $K > 2$

Let $\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in [0, 1]^{n_k \times K}$.  
Let $X$ be a block diagonal matrix 
$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in [0, 1]^{n \times K^2}$.

Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
$l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
[0, 1]^{n \times K}$.  
Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
[0, 1]^{n \times K^2}$.

Then $P = X Y^\top$.

Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
$\Pi$, so $P = X \Pi X^\top$.

The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$[^pairs].

[^pairs]: TODO: describe pairs $(s, t)$ in more elegant/succinct closed form

So $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
eigenvalues equal to $-1$. $\Pi$ has the decomposed form 
$\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top$, and we can describe the 
PABM with $K$ communities as a GRDPG with latent positions $X U$ with signature 
$\Big( K (K + 1) / 2, K (K - 1) / 2 \Big)$.

#### Example: $K = 3$

Using the same notation as before:

$X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & \lambda^{(21)} & \lambda^{(22)} & \lambda^{(23)} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
\end{bmatrix}$

$Y = \begin{bmatrix} 
\lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
0 & \lambda^{(21)} & 0 & 0 & \lambda^{(22)} & 0 & 0 & \lambda^{(23)} & 0 \\
0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
\end{bmatrix}$

Then $Y = X \Pi$ where 
$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}$

Perhaps a simpler way to look at this is:

* Positions 1, 5, 9 are fixed
* The cycles of order 2 are
    * $(2, 4)$
    * $(3, 7)$
    * $(6, 8)$

Therefore, we can decompose $\Pi = U I_{6, 3} U^\top$ where the first three 
columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the fixed 
positions $1$, $5$, and $9$, the next three columns consist of eigenvectors 
$(e_k + e_l) / \sqrt{2}$, and the last three columns consist of eigenvectors 
$(e_k - e_l) / \sqrt{2}$, where pairs $(k, l)$ correspond to the cycles of 
order 2 described above.

The latent positions are the rows of  
$XU = \begin{bmatrix}
\lambda^{(11)} & 0 & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
0 & \lambda^{(22)} & 0 & \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
0 & 0 & \lambda^{(33)} & 0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} & 0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
\end{bmatrix}$.

#### Additional notes and observations

* $XU$ is a convenient but not necessarily unique characterization of the 
latent positions. This comes into play when we try to estimate the latent 
positions---the estimate is only unique up to matrix multiplication by an 
arbitrary matrix in $\mathbb{O}(p, q)$.
    * i.e., let $Q \in \mathbb{O}(p, q)$, and let $P = (XU) I_{pq} (XU)^\top$. 
    Then we can also decompose $P = (XUQ) I_{pq} (XUQ)^\top$
    $= (XU) (Q I_{pq} Q^\top) (XU)^\top$
    $= (XU) I_{pq} (XU)^\top = P$
    * $\mathbb{O}(p, q) = 
    \{Q \in \mathbb{R}^{(p+q) \times (p+q)} : Q I_{pq} Q^\top = I_{pq}\}$
* If we use the latent positions described by $XU$, the dot product of two 
points in different clusters equals 0. Points that are not within the same 
cluster are orthogonal to each other (again, if we take $XU$ as the true 
latent positions).
* Even though the latent positions $XU$ follow these conditions, we can 
multiply by any matrix $Q \in \mathbb{O}(p, q)$ and have these conditions 
fail.

### Recovery/Estimation/Inference

The main goals are:

1. Determining the underlying cluster/community memberships/structure
2. Estimating the latent positions $XU$
3. Estimating $\lambda^{(kl)}$ for $k, l = 1, ..., K$

#### Previous results

Theorem 5 from the GRDPG paper[^grdpg] says the following:

* Let 
    * $X$ be an latent position matrix with latent positions as $n$ rows 
    $x_i^\top$ in $d$ dimensions
    * $A \sim GRDPG_{p, q}(X)$
    * $\hat{x}_i$'s be the adjacency embedding of $A$
* Then 
    * $\max\limits_{i = 1, ..., n} || Q \hat{x}_i - x_i || = 
    O_P(\frac{(\log n)^c}{n^{1/2}})$
    * $Q \in \mathbb{O}(p, q)$ (unidentifiable)
    * $c > 0$

Under some sparsity conditions.

So we can only identify the embedding up to matrix multiplication.

Both the GRDPG paper and the PABM paper[^pabm] starting with the spectral 
embedding (adjacency or Laplacian) of $A$ and performing a clustering step 
($K$-means, GMM, etc.). Unfortunately, there aren't any theoretical guarantees 
for the clustering step.

The PABM paper suggests subspace clustering on the embedding, and we can 
clearly see the justification by looking at $XU$. The rows of $XU$ that are 
within the same cluster lie in a $K$-dimensional subspace. We can further 
see that the $K$ subspaces defined by the $K$ clusters are all orthogonal to 
each other. However, in practice, the unidentifiable $Q$ matrix breaks this 
property when we decompose $A$.

***TODO: Investigate CLT***

#### Proposed Methods

Let $z_i^\top$ denote the $i$^th^ row of $XU$.  
Let $\gamma(x, y)$ denote the cosine similarity between $x$ and $y$.  
Then $\gamma(z_i, z_j) = 0$ if vertices $i$ and $j$ belong to different 
clusters, and $\gamma(z_i, z_j) > 0$ almost surely if $i$ and $j$ belong to the 
same cluster.

This then gives us the result:

* Let $\hat{Z}$ denote the adjacency embedding of 
$A \sim PABM(\{\lambda^{(kl)}\}_K)$ with rows $\hat{z}_i^\top$.
* Then $\exists Q \in \mathbb{O}(p, q)$ s.t. 
$\max\limits_{i, j = 1, ...,n} | \gamma(Q \hat{z}_i, Q \hat{z}_j) | =
O_P(\frac{(\log n)^c}{n^{1/2}})$ if vertices $i$ and $j$ are in different 
clusters.[^convergence]
* We can state a similar result using the Laplacian embedding of $A$.

[^convergence]: TODO: Check that the convergence rate is true (I'm pretty sure 
it is since we are taking the dot product of two vectors and then dividing 
by their norms, so we should get the same convergence rate as in the GRDPG 
result). Or alternatively, 
$|(Q \hat{z}_i)^\top (Q \hat{z}_j)| = O_P(\frac{(\log n)^{2c}}{n})$.

***TODO: CLT results***

Thus if we can recover $Q$, we can assign cluster memberships based on the behavior of the cosine similarity matrix.[^clt] The problem here is that we cannot get rid of the intractable matrix $Q$.

Also note that $Q$ is not merely a rotation matrix. Multiplying by $Q$ changes 
inter-point distances (and angles).

[^clt]: TODO: Use CLT to describe deviation from 1 or 0, e.g., if $i$, $j$ in 
different clusters then what is the distribution/density of 
$\gamma(\hat{z}_i, \hat{z}_j)$, which should be centered at 0.

However, perhaps we can use the properties of $XU$ to estimate $Q$. $XU$ 
consists of points on $K$ $K$-dimensional orthogonal hyperplanes. The PABM 
paper uses the hyperplanes part to justify subspace clustering but does not 
use the fact that they are orthogonal hyperplanes---the orthogonality is 
broken by the decomposition of $P$. If $Z$ is the adjacency embedding of $P$, 
then $Z = X U Q^{-1}$ for some unknown $Q \in \mathbb{O}(p, q)$. However, 
the multiplication by $Q^{-1}$ does not affect the fact that the points lie 
along (not necessarily orthogonal) hyperplanes, which is a justification for 
using subspace clustering to identify the community memberships.

For now, we will focus on he case $K = 2$. In this scenario, we have the 
following: We wish to find $Q \in \mathbb{O}(3, 1)$ such that $Z Q$ "looks like" 
$XU$. $X U = \begin{bmatrix} 
\lambda^{(11)} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(12)} / \sqrt{2} \\
0 & \lambda^{(22)} & \lambda^{(21)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2}
\end{bmatrix}$, and we can observe the following properties:

1. If two rows of $XU$ are not in the same cluster, they are orthogonal.
2. If vertex $i$ is in the first cluster, its second entry is $0$. Likewise,
if vertex $j$ is in the second cluster, its first entry is $0$. 
3. If vertex $i$ is in the first cluster, its third and fourth entries are 
identical. Similarly, if vertex $j$ is in the second cluster, the sum of its 
third and fourth entries is equal to $0$.

It is straightforward to generalize these properties to $K > 2$.

So we need to find $Q \in \mathbb{O}(3, 1)$ such that $\hat{Z} Q$ meets these 
criteria as closely as possible. One optimization problem that could arise 
from this is:

$$\begin{split}
  \arg\min_Q & ||\gamma^{(1)}||^2 + ||\gamma^{(2)}||^2 \\
  \text{s.t. } & \Xi = \hat{Z} Q \text{ with rows } \xi_i^\top \\
  & \gamma^{(1)}_i = \min (|\xi_{i1}|, |\xi_{i2}|) \\
  & \gamma^{(2)}_i = \min(|\xi_{i3} + \xi_{i4}|, |\xi_{i3} - \xi_{i4}|) \\
  & Q \in \mathbb{O}(3, 1)
\end{split}$$

Again, this part is easily generalizable to $K > 2$.

If we embed $P$ instead of $A$, the optimal value of this objective is $0$.

A possible pseudo-algorithm for this is as follows:

1. Start with affinity matrix $A$ and its adjacency embedding $\hat{Z}$
2. Pick some $\tilde{Q} \in \mathbb{O}(3, 1)$.
3. Compute $\Xi = \hat{Z} \tilde{Q}$ and vectors $\gamma^{(1)}$, $\gamma^{(2)}$.
4. Compute the objective function.
5. Pick a new $\tilde{Q} \in \mathbb{O}(3, 1)$ that lowers the objective 
function, and repeat steps 3-5.

Note that this doesn't include orthogonality between points in different 
clusters, but the other two criteria imply orthogonality.

The advantage of this approach over others is that we can directly estimate 
$X U$ and therefore $X$ (since we know what $U$ is exactly for any given $K$) 
instead of relying on subspace clustering. Once we have the embedding 
$\hat{Z}$ and we find $\hat{Q}$, we can set $\hat{X} = \hat{Z} \hat{Q} U^\top$, 
and the elements of $\hat{X}$ will correspond to the PABM vectors 
$\lambda^{(kl)}$ (exactly if the embedding is of $P$, approximately if the 
embedding is of $A$).

The difficulty is then in choosing $Q$ from $\mathbb{O}(3, 1)$. It turns out 
that $\mathbb{O}(3, 1)$ is the Lorentz group[^lorentz] and matrices of this 
group are equivalent to Lorentz transformations, which describe the warping of 
spacetime in special relativity[^convention]. Any matrix 
$Q \in \mathbb{O}(3, 1)$ can be expressed as a product of six matrices that 
depend on four parameters[^generator], i.e., 
$Q = R_1(\theta) R_2(\phi) R_3(\psi) S_1(\theta) S_2(\tau) S_3(-\tau)$.The six 
matrices are analogous to three rotation matrices and six squeeze mappings.

$R_1(\theta) = \begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & \cos \theta & \sin \theta & 0 \\
0 & -\sin \theta & \cos \theta & 0 \\
0 & 0 & 0 & -1
\end{bmatrix}$

$R_2(\phi) = \begin{bmatrix}
\cos \phi & 0 & \sin \phi & 0 \\
0 & 1 & 0 & 0 \\
-\sin \phi & 0 & \cos \phi & 0 \\
0 & 0 & 0 & -1
\end{bmatrix}$

$R_3(\psi) = \begin{bmatrix}
\cos \psi & \sin \psi & 0 & 0 \\
-\sin \psi & \cos \psi & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1
\end{bmatrix}$

$S_1(\theta) = \begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & \cosh \theta & \sinh \theta \\
0 & 0 & \sinh \theta & \cosh \theta
\end{bmatrix}$

$S_2(\tau) = \begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & \cosh \tau & 0 & \sinh \tau \\
0 & 0 & 1 & 0 \\
0 & \sinh \tau & 0 & \cosh \tau
\end{bmatrix}$

$S_3(-\tau) = \begin{bmatrix}
\cosh \tau & 0 & 0 & \sinh(-\tau) \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
\sinh(-\tau) & 0 & 0 & \cosh \tau
\end{bmatrix}$

We can write $Q = Q(\theta, \phi, \psi, \tau)$ and tweak the four angles until 
convergence.

**TODO: Show that for arbitrary $K$, we need $K^2$ angles and $K (K+1)$ 
matrices.**

**TODO: Compute gradients(?)**

[^lorentz]: https://en.wikipedia.org/wiki/Lorentz_group

[^convention]: There is actually a multiplication by $-1$ to account for here.

[^generator]: https://iopscience.iop.org/book/978-1-6817-4254-0/chapter/bk978-1-6817-4254-0ch1

#### Alternative approaches

1. If we know the cluster memberships (or if we estimate them using subspace 
clustering), then we can write a perhaps easier or more robust optimization 
problem:  
Rearrange $\hat{Z}$ such that the first $n_1$ rows are in cluster 1 and the 
next $n_2$ rows are in cluster 2. Then solve the following.
$$\begin{split}
\arg\min_Q & ||\Xi_{1:n_1, 1}||^2 + ||\Xi_{n_1+1 : n, 2}||^2 + 
||\Xi_{1:n_1, 3} - \Xi_{1:n1, 4}||^2 + 
||\xi_{n_1+1 : n, 3} + \xi_{n_1 + 1, 4}||^2 \\
\text{s.t. } & \Xi = \hat{Z} Q \\
& Q \in \mathbb{O}(3, 1)
\end{split}$$  
Numerical expeirments thus far suggest that this method does not perform any 
better than estimating $X U$ directly without a clustering step (although 
more investigation is required to say anything for sure).

2. $Q \in \mathbb{O}(3, 1)$ can be written as:  
$Q = \begin{bmatrix} M & -a \\ -b^\top & \gamma \end{bmatrix}$[^decomp]. This 
might produce a better parameterization of the $Q$ matrices, although so far I 
haven't had much luck here. We can see that if $Q I_{3, 1} Q^\top = I_{3, 1}$, 
then we get the following constraints:
    * $M M^\top - a a^\top = I$
    * $\gamma a = M b$
    * $\gamma^2 = b^\top b + 1$

[^decomp]: https://en.wikipedia.org/wiki/Lorentz_transformation

### Numerical example

#### PABM with $K = 2$

We will draw the elements of $\lambda^{(11)}$ and $\lambda^{(22)}$ 
independently  from a beta distribution with shape parameters $a$ and $b$ such 
that $a > b$. 
The elements of $\lambda^{(12)}$ and $\lambda^{(21)}$ will also be 
independently drawn from a beta distribution, this time with shape parameters 
$c$ and $d$ such that $c < d$. This ensures that it is more likely for a pair 
of vertices from the same cluster to be connected than it is for a pair of 
vertices from different clusters.

```{r}
# parameters
n1 <- 2 ** 5
n2 <- n1
n <- n1 + n2
Ipq <- diag(c(1, 1, 1, -1))

# cluster assignments
z <- c(rep(1, n1), 
       rep(2, n2))

# construct P
lambda11 <- rbeta(n1, 3, 1)
lambda22 <- rbeta(n2, 3, 1)
lambda12 <- rbeta(n1, 1, 3)
lambda21 <- rbeta(n2, 1, 3)
X <- cbind(c(lambda11, rep(0, n2)),
           c(lambda12, rep(0, n2)),
           c(rep(0, n1), lambda21),
           c(rep(0, n1), lambda22))
Y <- cbind(c(lambda11, rep(0, n2)),
           c(rep(0, n1), lambda21),
           c(lambda12, rep(0, n2)),
           c(rep(0, n1), lambda22))
P <- X %*% t(Y)

# draw affinity matrix
A <- draw.graph(P)

# U for K=2
U <- cbind(c(1, 0, 0, 0),
           c(0, 0, 0, 1),
           c(0, 1 / sqrt(2), 1 / sqrt(2), 0),
           c(0, 1 / sqrt(2), -1 / sqrt(2), 0))

# latent positions
XU <- X %*% U

# print(XU)

Z <- embedding(P)
Zhat <- embedding(A, 3, 1)
```

Probability matrix:

```{r, fig.width = 3}
plot(P, border = NA)
```

One realization $A$ of $P$:

```{r}
qgraph::qgraph(A, groups = factor(z))
```

```{r fig.width = 3}
plot(A, border = NA)
```

The "true" latent positions $X U$:

```{r fig.width = 4}
pairs(XU, col = z, asp = 1)
```

$Z$, the adjacency embedding of $P$:

```{r fig.width = 4}
pairs(Z, asp = 1, col = z)
```

$\hat{Z}$, the adjacency embedding of $A$:

```{r fig.width = 4}
pairs(Zhat, asp = 1, col = z)
```

Recovering $X U$ from $Z$:

```{r cache = FALSE, fig.width = 4}
XU.prime <- transform.embedding(Z)
pairs(XU.prime, asp = 1, col = z)

# XU.prime
```

Recovering $X U$ from $\hat{Z}$:

```{r cache = TRUE, fig.width = 4}
XU.hat <- transform.embedding(Zhat)
pairs(XU.hat, asp = 1, col = z)

# XU.hat
```

Some brief checking shows that we are able to recover the $XU$ and the 
$\lambda^{(kl)}$'s from $P$ exactly, which suggests that this method should 
work asymptotically, but it is not as well behaved when we use $A$.

### Additional TODOs

* We are currently treating the optimization step as a black box (just using 
Nelder-Mead or simulated annealing at the moment). Since 
$Q = Q(\theta, \phi, \psi, \tau)$ and is smooth with respect to each of those 
four parameters, we should be able to use a more disciplined 
gradient/Hessian-based approach.
* Doing the above should provide us a good way to describe the convergence 
behavior of estimating $Q$.
* Is the connection between PABM and special relativity a happy coincidence, or 
is there something more to explore here?
* Generalize optimization step to $K > 2$.
* The estimate $\hat{Z} \hat{Q}$ doesn't necessarily obey GRDPG rules. How 
can we force it to?
* What happens when we let $\lambda^{(kl)}$'s be random variables instead of 
fixed vectors, e.g., $\lambda^{(kl)}_i \sim Beta(a_{kl}, b_{kl})$ (or some 
Dirichlet equivalent)?

### 2020-07-11 pabm-notes.Rmd
uploading scratch file to resume work on server

@@
 title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph"
-output: pdf_document
-# output: html_document
+# output: pdf_document
+output: html_document
 # geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
@@
 
-  Z.candidate <- Z.hat %*% R
-  B <- Z.candidate %*% t(Z.candidate)
-  return(sum(apply(Z.candidate[, 1:2] ** 2, 1, min)) + 
-           sum(apply(cbind((Z.candidate[, 3] - Z.candidate[, 4]) ** 2,
-                           (Z.candidate[, 3] + Z.candidate[, 4]) ** 2), 
-                     1, min)))
+  XU.c <- Z.hat %*% R
+  # clusters <- apply(XU.c[, 1:2] ** 2, 1, which.min)
+  clusters <- c(rep(1, 32), rep(2, 32))
+  in.prods <- XU.c[clusters == 1, ] %*% t(XU.c[clusters == 2, , drop = FALSE])
+  return(
+    sum(apply(XU.c[, 1:2] ** 2, 1, min)) +
+      sum(apply(cbind((XU.c[, 3] - XU.c[, 4]) ** 2,
+                      (XU.c[, 3] + XU.c[, 4]) ** 2),
+                1, min))
+    )
+  # return(
+  #   sum(apply(XU.c[, 1:2] ** 2, 1, min)) +
+  #     Matrix::norm(in.prods, 'F') +
+  #     sum((XU.c[clusters == 1, 3] - XU.c[clusters == 1, 4]) ** 2) +
+  #     sum((XU.c[clusters == 2, 3] + XU.c[clusters == 2, 4]) ** 2)
+  # )
 }
@@
                     method = 'Nelder-Mead',
-                    control = list(abstol = 1e-10, maxit = 1e6))$par
+                    control = list(abstol = 1e-10, maxit = 1e3))$par
   R <- construct.Q(angles.2)
@@
 $S_3(-\tau) = \begin{bmatrix}
-\cosh \tau & 0 & 0 & \sinh(-\tau) \\
+\cosh \tau & 0 & 0 & -\sinh \tau \\
 0 & 1 & 0 & 0 \\
 0 & 0 & 1 & 0 \\
-\sinh(-\tau) & 0 & 0 & \cosh \tau
+-\sinh \tau & 0 & 0 & \cosh \tau
 \end{bmatrix}$

### 2020-07-11 pabm-scratch.R

@@ -1,0 +1,429 @@
+import::from(magrittr, `%>%`)
+library(plot.matrix)
+import::from(foreach, `%do%`)
+library(CVXR)
+
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+
+doMC::registerDoMC(8)
+
+embedding <- function(A, p = NULL, q = NULL,
+                      eps = 1e-6) {
+  eigen.A <- eigen(A, symmetric = TRUE)
+  if (is.null(p) | is.null(q)) {
+    keep <- (abs(eigen.A$values) > eps)
+  } else {
+    keep <- c(seq(p), seq(n, n - q + 1))
+  }
+  
+  U <- eigen.A$vectors[, keep]
+  S <- diag(sqrt(abs(eigen.A$values[keep])))
+  return(U %*% S)
+}
+
+draw.graph <- function(P) {
+  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
+  A[lower.tri(A)] <- 0
+  diag(A) <- 0
+  A <- A + t(A)
+  return(A)
+}
+
+n1 <- 2 ** 5
+n2 <- n1
+n <- n1 + n2
+
+z <- c(rep(1, n1), 
+       rep(2, n2))
+
+Ipq <- diag(c(1, 1, 1, -1))
+
+lambda11 <- rbeta(n1, 5, 1)
+lambda22 <- rbeta(n2, 5, 1)
+lambda12 <- rbeta(n1, 1, 5)
+lambda21 <- rbeta(n2, 1, 5)
+
+X <- cbind(c(lambda11, rep(0, n2)),
+           c(lambda12, rep(0, n2)),
+           c(rep(0, n1), lambda21),
+           c(rep(0, n1), lambda22))
+Y <- cbind(c(lambda11, rep(0, n2)),
+           c(rep(0, n1), lambda21),
+           c(lambda12, rep(0, n2)),
+           c(rep(0, n1), lambda22))
+P <- X %*% t(Y)
+A <- draw.graph(P)
+
+Z <- embedding(P)
+Zhat <- embedding(A, 3, 1)
+pairs(Z, col = z, asp = 1)
+
+P - Z %*% Ipq %*% t(Z)
+
+XU <- cbind(c(lambda11, rep(0, n2)),
+            c(rep(0, n1), lambda22),
+            c(lambda12 / sqrt(2), lambda21 / sqrt(2)),
[…]

### 2020-07-12 summary.Rmd
still working on writeup

--- 2020-06-21 pabm-notes.Rmd
@@
 ---
-title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph"
-output: pdf_document
-# output: html_document
+title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
+# output: pdf_document
+output: html_document
 # geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
@@
 - \usepackage{mathtools}
+abstract: |
+  this
+  is
+  an
+  abstract
 ---
@@
 
-```{r include = FALSE}
-embedding <- function(A, p = NULL, q = NULL,
-                      eps = 1e-6) {
-  eigen.A <- eigen(A, symmetric = TRUE)
-  if (is.null(p) | is.null(q)) {
-    keep <- (abs(eigen.A$values) > eps)
-  } else {
-    keep <- c(seq(p), seq(n, n - q + 1))
-  }
-
-  U <- eigen.A$vectors[, keep]
-  S <- diag(sqrt(abs(eigen.A$values[keep])))
-  return(U %*% S)
-}
-
-draw.graph <- function(P) {
-  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
-  A[lower.tri(A)] <- 0
-  diag(A) <- 0
-  A <- A + t(A)
-  return(A)
-}
-
-rot.mat.2 <- function(angle, axis = 1) {
-  # rotation matrix for K=2
-
-  if (axis %in% c(1, 3)) {
-    rotation <- matrix(c(cos(angle), -sin(angle), sin(angle), cos(angle)),
-                       nrow = 2, ncol = 2, byrow = TRUE)
-    if (axis == 1) {
-      return(as.matrix(Matrix::bdiag(1, rotation, -1)))
-    } else {
-      return(as.matrix(Matrix::bdiag(rotation, 1, -1)))
-    }
-  } else if (axis == 2) {
-    return(matrix(c(cos(angle), 0, sin(angle), 0,
-                    0, 1, 0, 0,
-                    -sin(angle), 0, cos(angle), 0,
-                    0, 0, 0, -1),
-                  nrow = 4, ncol = 4,
-                  byrow = TRUE))
-  } else {
-    stop(simpleError('axis must be 1, 2, or 3'))
-  }
-}
-
-hyp.rot.mat.2 <- function(angle, axis = 1) {
-  # hyperbolic rotation matrix for $K=2
-
-  if (axis == 1) {
-    return(matrix(c(1, 0, 0, 0,
-                    0, 1, 0, 0, 
-                    0, 0, cosh(angle), sinh(angle),
-                    0, 0, sinh(angle), cosh(angle)), 
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else if (axis == 2) {
-    return(matrix(c(1, 0, 0, 0,
-                    0, cosh(angle), 0, sinh(angle),
-                    0, 0, 1, 0,
-                    0, sinh(angle), 0, cosh(angle)),
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else if (axis == 3) {
-    return(matrix(c(cosh(angle), 0, 0, sinh(angle),
-                    0, 1, 0, 0, 
-                    0, 0, 1, 0, 
-                    sinh(angle), 0, 0, cosh(angle)),
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else {
-    stop(simpleError('axis must be 1, 2, or 3'))
-  }
-}
-
-construct.Q <- function(angles) {
-  rot.angles <- angles[1:3]
-  hyp.angles <- c(angles[1], angles[4], -angles[4])
-  rot.matrices <- lapply(seq(3), function(i) rot.mat.2(rot.angles[i], i))
-  hyp.matrices <- lapply(seq(3), function(i) hyp.rot.mat.2(hyp.angles[i], i))
-  matrices <- c(rot.matrices, hyp.matrices)
-  return(Reduce(`%*%`, matrices))
-}
-
-obj.fun <- function(angles, Z.hat) {
-  R <- construct.Q(angles)
-
-  Z.candidate <- Z.hat %*% R
-  B <- Z.candidate %*% t(Z.candidate)
-  return(sum(apply(Z.candidate[, 1:2] ** 2, 1, min)) + 
-           sum(apply(cbind((Z.candidate[, 3] - Z.candidate[, 4]) ** 2,
-                           (Z.candidate[, 3] + Z.candidate[, 4]) ** 2), 
-                     1, min)))
-}
-
-transform.embedding <- function(Z.hat) {
-  init.angles <- runif(4, -pi, pi)
-  angles.1 <- optim(init.angles, function(x) obj.fun(x, Z.hat), 
-                    method = 'SANN')$par
-  angles.2 <- optim(angles.1, function(x) obj.fun(x, Z.hat),
-                    method = 'Nelder-Mead',
-                    control = list(abstol = 1e-10, maxit = 1e6))$par
-  R <- construct.Q(angles.2)
-  return(Z.hat %*% R)
-}
-```
-
 \newcommand{\diag}{\text{diag}}
 
-### The popularity adjusted block model (PABM)[^pabm]
+# 1 Introduction
+
+## 1.1 Previous work
+
+# 2 Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
+
+## 2.1 The popularity adjusted block model (PABM)[^pabm]
 
@@
 
-### The generalized random dot product graph (GRDPG)[^grdpg]
+## 2.2 The generalized random dot product graph (GRDPG)[^grdpg]
 
@@
 
-### Connecting the PABM to the GRDPG for $K = 2$
+## 2.3 Connecting the PABM to the GRDPG for $K = 2$
 
@@
 
-### Generalization to $K > 2$
+## 2.4 Generalization to $K > 2$
 
@@
 
-#### Example: $K = 3$
+**Example: $K = 3$**
 
@@
 
-Then $Y = X \Pi$ where 
+Then $P = X Y$ and $Y = X \Pi$ where 
 $\Pi = \begin{bmatrix} 
@@
 
-#### Additional notes and observations
+# 3 Methodology
 
-* $XU$ is a convenient but not necessarily unique characterization of the 
-latent positions. This comes into play when we try to estimate the latent 
-positions---the estimate is only unique up to matrix multiplication by an 
-arbitrary matrix in $\mathbb{O}(p, q)$.
-    * i.e., let $Q \in \mathbb{O}(p, q)$, and let $P = (XU) I_{pq} (XU)^\top$. 
-    Then we can also decompose $P = (XUQ) I_{pq} (XUQ)^\top$
-    $= (XU) (Q I_{pq} Q^\top) (XU)^\top$
-    $= (XU) I_{pq} (XU)^\top = P$
-    * $\mathbb{O}(p, q) = 
-    \{Q \in \mathbb{R}^{(p+q) \times (p+q)} : Q I_{pq} Q^\top = I_{pq}\}$
-* If we use the latent positions described by $XU$, the dot product of two 
-points in different clusters equals 0. Points that are not within the same 
-cluster are orthogonal to each other (again, if we take $XU$ as the true 
-latent positions).
-* Even though the latent positions $XU$ follow these conditions, we can 
-multiply by any matrix $Q \in \mathbb{O}(p, q)$ and have these conditions 
-fail.
+Two inference objectives arise from the PABM:
 
-### Recovery/Estimation/Inference
+1. Cluster membership identification (up to permutation).
+2. Parameter estimation (estimating $\lambda^{(kl)}$'s).
 
-The main goals are:
+Here, we will focus more on (1) and pose possible methods for (2). In our 
+methods, we assume that $K$, the number of clusters, is known beforehand.
 
-1. Determining the underlying cluster/community memberships/structure
-2. Estimating the latent positions $XU$
-3. Estimating $\lambda^{(kl)}$ for $k, l = 1, ..., K$
+## 3.1 Cluster membership identification
 
-#### Previous results
+### 3.1.1 Using edge probability matrix $P$
 
-Theorem 5 from the GRDPG paper[^grdpg] says the following:
+We previously stated one possible set of latent positions that result in the 
+edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
+(or can estimate) $XU$ directly, then both the clustering and parameter 
+identification problem are trivial since $U$ is orthonormal and fixed for each 
+value of $K$. However, direct identification or estimation of $XU$ is not 
+possible.
 
-* Let 
-    * $X$ be an latent position matrix with latent positions as $n$ rows 
-    $x_i^\top$ in $d$ dimensions
-    * $A \sim GRDPG_{p, q}(X)$
-    * $\hat{x}_i$'s be the adjacency embedding of $A$
-* Then 
-    * $\max\limits_{i = 1, ..., n} || Q \hat{x}_i - x_i || = 
-    O_P(\frac{(\log n)^c}{n^{1/2}})$
-    * $Q \in \mathbb{O}(p, q)$ (unidentifiable)
-    * $c > 0$
+If we decompose $P = Z I_{pq} Z^\top$, then
+$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
+the exact edge probability matrix, we cannot recover the "original" latent 
+positions $XU$. Note that unlike in the case of the regular random dot product 
+graph, $Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
+$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
+$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
+us from using the properties of $XU$ directly. In particular, if 
+$Q \in \mathbb{O}(n)$, then we could use the fact that 
+$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
+and $j$ are in different clusters.
 
-Under some sparsity conditions.
+We can note from the explicit form of $XU$ that it represents points in 
+$\mathbb{R}^{K^2}$ such that points within each cluster lie on $K$-dimensional 
+subspaces. Furthermore, the subspaces are orthogonal to each other. 
+Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
+but retains the property that each cluster is represented by a $K$-dimensional 
+subspace. Using this property, previous work proposes the use of subspace 
+clustering while acknowledging some of its shortcomings.
 
-So we can only identify the embedding up to matrix multiplication.
+**Theorem 1.** 
+*Let $P = V D V^\top$ be the spectral decomposition of the edge probability
+matrix.*
 
-Both the GRDPG paper and the PABM paper[^pabm] starting with the spectral 
-embedding (adjacency or Laplacian) of $A$ and performing a clustering step 
-($K$-means, GMM, etc.). Unfortunately, there aren't any theoretical guarantees 
-for the clustering step.
+### 3.1.2 Using adjacency matrix $A$
 
-The PABM paper suggests subspace clustering on the embedding, and we can 
-clearly see the justification by looking at $XU$. The rows of $XU$ that are 
-within the same cluster lie in a $K$-dimensional subspace. We can further 
-see that the $K$ subspaces defined by the $K$ clusters are all orthogonal to 
-each other. However, in practice, the unidentifiable $Q$ matrix breaks this 
-property when we decompose $A$.
+## 3.2 Parameter estimation
 
-***TODO: Investigate CLT***
-
-#### Proposed Methods
-
-Let $z_i^\top$ denote the $i$^th^ row of $XU$.  
-Let $\gamma(x, y)$ denote the cosine similarity between $x$ and $y$.  
-Then $\gamma(z_i, z_j) = 0$ if vertices $i$ and $j$ belong to different 
-clusters, and $\gamma(z_i, z_j) > 0$ almost surely if $i$ and $j$ belong to the 
-same cluster.
-
-This then gives us the result:
-
-* Let $\hat{Z}$ denote the adjacency embedding of 
-$A \sim PABM(\{\lambda^{(kl)}\}_K)$ with rows $\hat{z}_i^\top$.
-* Then $\exists Q \in \mathbb{O}(p, q)$ s.t. 
-$\max\limits_{i, j = 1, ...,n} | \gamma(Q \hat{z}_i, Q \hat{z}_j) | =
-O_P(\frac{(\log n)^c}{n^{1/2}})$ if vertices $i$ and $j$ are in different 
-clusters.[^convergence]
-* We can state a similar result using the Laplacian embedding of $A$.
-
-[^convergence]: TODO: Check that the convergence rate is true (I'm pretty sure 
-it is since we are taking the dot product of two vectors and then dividing 
-by their norms, so we should get the same convergence rate as in the GRDPG 
-result). Or alternatively, 
-$|(Q \hat{z}_i)^\top (Q \hat{z}_j)| = O_P(\frac{(\log n)^{2c}}{n})$.
-
-***TODO: CLT results***
-
-Thus if we can recover $Q$, we can assign cluster memberships based on the behavior of the cosine similarity matrix.[^clt] The problem here is that we cannot get rid of the intractable matrix $Q$.
-
-Also note that $Q$ is not merely a rotation matrix. Multiplying by $Q$ changes 
-inter-point distances (and angles).
-
-[^clt]: TODO: Use CLT to describe deviation from 1 or 0, e.g., if $i$, $j$ in 
-different clusters then what is the distribution/density of 
-$\gamma(\hat{z}_i, \hat{z}_j)$, which should be centered at 0.
-
-However, perhaps we can use the properties of $XU$ to estimate $Q$. $XU$ 
-consists of points on $K$ $K$-dimensional orthogonal hyperplanes. The PABM 
-paper uses the hyperplanes part to justify subspace clustering but does not 
-use the fact that they are orthogonal hyperplanes---the orthogonality is 
-broken by the decomposition of $P$. If $Z$ is the adjacency embedding of $P$, 
-then $Z = X U Q^{-1}$ for some unknown $Q \in \mathbb{O}(p, q)$. However, 
-the multiplication by $Q^{-1}$ does not affect the fact that the points lie 
-along (not necessarily orthogonal) hyperplanes, which is a justification for 
-using subspace clustering to identify the community memberships.
-
-For now, we will focus on he case $K = 2$. In this scenario, we have the 
-following: We wish to find $Q \in \mathbb{O}(3, 1)$ such that $Z Q$ "looks like" 
-$XU$. $X U = \begin{bmatrix} 
-\lambda^{(11)} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(12)} / \sqrt{2} \\
-0 & \lambda^{(22)} & \lambda^{(21)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2}
-\end{bmatrix}$, and we can observe the following properties:
-
-1. If two rows of $XU$ are not in the same cluster, they are orthogonal.
-2. If vertex $i$ is in the first cluster, its second entry is $0$. Likewise,
-if vertex $j$ is in the second cluster, its first entry is $0$. 
-3. If vertex $i$ is in the first cluster, its third and fourth entries are 
-identical. Similarly, if vertex $j$ is in the second cluster, the sum of its 
-third and fourth entries is equal to $0$.
-
-It is straightforward to generalize these properties to $K > 2$.
-
-So we need to find $Q \in \mathbb{O}(3, 1)$ such that $\hat{Z} Q$ meets these 
-criteria as closely as possible. One optimization problem that could arise 
-from this is:
-
-$$\begin{split}
-  \arg\min_Q & ||\gamma^{(1)}||^2 + ||\gamma^{(2)}||^2 \\
-  \text{s.t. } & \Xi = \hat{Z} Q \text{ with rows } \xi_i^\top \\
-  & \gamma^{(1)}_i = \min (|\xi_{i1}|, |\xi_{i2}|) \\
-  & \gamma^{(2)}_i = \min(|\xi_{i3} + \xi_{i4}|, |\xi_{i3} - \xi_{i4}|) \\
-  & Q \in \mathbb{O}(3, 1)
-\end{split}$$
-
-Again, this part is easily generalizable to $K > 2$.
-
-If we embed $P$ instead of $A$, the optimal value of this objective is $0$.
-
-A possible pseudo-algorithm for this is as follows:
-
-1. Start with affinity matrix $A$ and its adjacency embedding $\hat{Z}$
-2. Pick some $\tilde{Q} \in \mathbb{O}(3, 1)$.
-3. Compute $\Xi = \hat{Z} \tilde{Q}$ and vectors $\gamma^{(1)}$, $\gamma^{(2)}$.
-4. Compute the objective function.
-5. Pick a new $\tilde{Q} \in \mathbb{O}(3, 1)$ that lowers the objective 
-function, and repeat steps 3-5.
-
-Note that this doesn't include orthogonality between points in different 
-clusters, but the other two criteria imply orthogonality.
-
-The advantage of this approach over others is that we can directly estimate 
-$X U$ and therefore $X$ (since we know what $U$ is exactly for any given $K$) 
-instead of relying on subspace clustering. Once we have the embedding 
-$\hat{Z}$ and we find $\hat{Q}$, we can set $\hat{X} = \hat{Z} \hat{Q} U^\top$, 
-and the elements of $\hat{X}$ will correspond to the PABM vectors 
-$\lambda^{(kl)}$ (exactly if the embedding is of $P$, approximately if the 
-embedding is of $A$).
-
-The difficulty is then in choosing $Q$ from $\mathbb{O}(3, 1)$. It turns out 
-that $\mathbb{O}(3, 1)$ is the Lorentz group[^lorentz] and matrices of this 
-group are equivalent to Lorentz transformations, which describe the warping of 
-spacetime in special relativity[^convention]. Any matrix 
-$Q \in \mathbb{O}(3, 1)$ can be expressed as a product of six matrices that 
-depend on four parameters[^generator], i.e., 
-$Q = R_1(\theta) R_2(\phi) R_3(\psi) S_1(\theta) S_2(\tau) S_3(-\tau)$.The six 
-matrices are analogous to three rotation matrices and six squeeze mappings.
-
-$R_1(\theta) = \begin{bmatrix}
-1 & 0 & 0 & 0 \\
-0 & \cos \theta & \sin \theta & 0 \\
-0 & -\sin \theta & \cos \theta & 0 \\
-0 & 0 & 0 & -1
-\end{bmatrix}$
-
-$R_2(\phi) = \begin{bmatrix}
-\cos \phi & 0 & \sin \phi & 0 \\
-0 & 1 & 0 & 0 \\
--\sin \phi & 0 & \cos \phi & 0 \\
-0 & 0 & 0 & -1
-\end{bmatrix}$
-
-$R_3(\psi) = \begin{bmatrix}
-\cos \psi & \sin \psi & 0 & 0 \\
--\sin \psi & \cos \psi & 0 & 0 \\
-0 & 0 & 1 & 0 \\
-0 & 0 & 0 & 1
-\end{bmatrix}$
-
-$S_1(\theta) = \begin{bmatrix}
-1 & 0 & 0 & 0 \\
-0 & 1 & 0 & 0 \\
-0 & 0 & \cosh \theta & \sinh \theta \\
-0 & 0 & \sinh \theta & \cosh \theta
-\end{bmatrix}$
-
-$S_2(\tau) = \begin{bmatrix}
-1 & 0 & 0 & 0 \\
-0 & \cosh \tau & 0 & \sinh \tau \\
-0 & 0 & 1 & 0 \\
-0 & \sinh \tau & 0 & \cosh \tau
-\end{bmatrix}$
-
-$S_3(-\tau) = \begin{bmatrix}
-\cosh \tau & 0 & 0 & \sinh(-\tau) \\
-0 & 1 & 0 & 0 \\
-0 & 0 & 1 & 0 \\
-\sinh(-\tau) & 0 & 0 & \cosh \tau
-\end{bmatrix}$
-
-We can write $Q = Q(\theta, \phi, \psi, \tau)$ and tweak the four angles until 
-convergence.
-
-**TODO: Show that for arbitrary $K$, we need $K^2$ angles and $K (K+1)$ 
-matrices.**
-
-**TODO: Compute gradients(?)**
-
-[^lorentz]: https://en.wikipedia.org/wiki/Lorentz_group
-
-[^convention]: There is actually a multiplication by $-1$ to account for here.
-
-[^generator]: https://iopscience.iop.org/book/978-1-6817-4254-0/chapter/bk978-1-6817-4254-0ch1
-
-#### Alternative approaches
-
-1. If we know the cluster memberships (or if we estimate them using subspace 
-clustering), then we can write a perhaps easier or more robust optimization 
-problem:  
-Rearrange $\hat{Z}$ such that the first $n_1$ rows are in cluster 1 and the 
-next $n_2$ rows are in cluster 2. Then solve the following.
-$$\begin{split}
-\arg\min_Q & ||\Xi_{1:n_1, 1}||^2 + ||\Xi_{n_1+1 : n, 2}||^2 + 
-||\Xi_{1:n_1, 3} - \Xi_{1:n1, 4}||^2 + 
-||\xi_{n_1+1 : n, 3} + \xi_{n_1 + 1, 4}||^2 \\
-\text{s.t. } & \Xi = \hat{Z} Q \\
-& Q \in \mathbb{O}(3, 1)
-\end{split}$$  
-Numerical expeirments thus far suggest that this method does not perform any 
-better than estimating $X U$ directly without a clustering step (although 
-more investigation is required to say anything for sure).
-
-2. $Q \in \mathbb{O}(3, 1)$ can be written as:  
-$Q = \begin{bmatrix} M & -a \\ -b^\top & \gamma \end{bmatrix}$[^decomp]. This 
-might produce a better parameterization of the $Q$ matrices, although so far I 
-haven't had much luck here. We can see that if $Q I_{3, 1} Q^\top = I_{3, 1}$, 
-then we get the following constraints:
-    * $M M^\top - a a^\top = I$
-    * $\gamma a = M b$
-    * $\gamma^2 = b^\top b + 1$
-
-[^decomp]: https://en.wikipedia.org/wiki/Lorentz_transformation
-
-### Numerical example
-
-#### PABM with $K = 2$
-
-We will draw the elements of $\lambda^{(11)}$ and $\lambda^{(22)}$ 
-independently  from a beta distribution with shape parameters $a$ and $b$ such 
-that $a > b$. 
-The elements of $\lambda^{(12)}$ and $\lambda^{(21)}$ will also be 
-independently drawn from a beta distribution, this time with shape parameters 
-$c$ and $d$ such that $c < d$. This ensures that it is more likely for a pair 
-of vertices from the same cluster to be connected than it is for a pair of 
-vertices from different clusters.
-
-```{r}
-# parameters
-n1 <- 2 ** 5
-n2 <- n1
-n <- n1 + n2
-Ipq <- diag(c(1, 1, 1, -1))
-
-# cluster assignments
-z <- c(rep(1, n1), 
-       rep(2, n2))
-
-# construct P
-lambda11 <- rbeta(n1, 3, 1)
-lambda22 <- rbeta(n2, 3, 1)
-lambda12 <- rbeta(n1, 1, 3)
-lambda21 <- rbeta(n2, 1, 3)
-X <- cbind(c(lambda11, rep(0, n2)),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(rep(0, n1), lambda22))
-Y <- cbind(c(lambda11, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda22))
-P <- X %*% t(Y)
-
-# draw affinity matrix
-A <- draw.graph(P)
-
-# U for K=2
-U <- cbind(c(1, 0, 0, 0),
-           c(0, 0, 0, 1),
-           c(0, 1 / sqrt(2), 1 / sqrt(2), 0),
-           c(0, 1 / sqrt(2), -1 / sqrt(2), 0))
-
-# latent positions
-XU <- X %*% U
-
-# print(XU)
-
-Z <- embedding(P)
-Zhat <- embedding(A, 3, 1)
-```
-
-Probability matrix:
-
-```{r, fig.width = 3}
-plot(P, border = NA)
-```
-
-One realization $A$ of $P$:
-
-```{r}
-qgraph::qgraph(A, groups = factor(z))
-```
-
-```{r fig.width = 3}
-plot(A, border = NA)
-```
-
-The "true" latent positions $X U$:
-
-```{r fig.width = 4}
-pairs(XU, col = z, asp = 1)
-```
-
-$Z$, the adjacency embedding of $P$:
-
-```{r fig.width = 4}
-pairs(Z, asp = 1, col = z)
-```
-
-$\hat{Z}$, the adjacency embedding of $A$:
-
-```{r fig.width = 4}
-pairs(Zhat, asp = 1, col = z)
-```
-
-Recovering $X U$ from $Z$:
-
-```{r cache = FALSE, fig.width = 4}
-XU.prime <- transform.embedding(Z)
-pairs(XU.prime, asp = 1, col = z)
-
-# XU.prime
-```
-
-Recovering $X U$ from $\hat{Z}$:
-
-```{r cache = TRUE, fig.width = 4}
-XU.hat <- transform.embedding(Zhat)
-pairs(XU.hat, asp = 1, col = z)
-
-# XU.hat
-```
-
-Some brief checking shows that we are able to recover the $XU$ and the 
-$\lambda^{(kl)}$'s from $P$ exactly, which suggests that this method should 
-work asymptotically, but it is not as well behaved when we use $A$.
-
-### Additional TODOs
-
-* We are currently treating the optimization step as a black box (just using 
-Nelder-Mead or simulated annealing at the moment). Since 
-$Q = Q(\theta, \phi, \psi, \tau)$ and is smooth with respect to each of those 
-four parameters, we should be able to use a more disciplined 
-gradient/Hessian-based approach.
-* Doing the above should provide us a good way to describe the convergence 
-behavior of estimating $Q$.
-* Is the connection between PABM and special relativity a happy coincidence, or 
-is there something more to explore here?
-* Generalize optimization step to $K > 2$.
-* The estimate $\hat{Z} \hat{Q}$ doesn't necessarily obey GRDPG rules. How 
-can we force it to?
-* What happens when we let $\lambda^{(kl)}$'s be random variables instead of 
-fixed vectors, e.g., $\lambda^{(kl)}_i \sim Beta(a_{kl}, b_{kl})$ (or some 
-Dirichlet equivalent)?
+# 4 Simulated Examples

### 2020-07-12 pabm-scratch.R

@@ -1,11 +1,11 @@
 import::from(magrittr, `%>%`)
 library(plot.matrix)
 import::from(foreach, `%do%`)
-library(CVXR)
+# library(CVXR)
 
 source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
 
-doMC::registerDoMC(8)
+doMC::registerDoMC(parallel::detectCores())
 
 embedding <- function(A, p = NULL, q = NULL,
                       eps = 1e-6) {
@@ -172,23 +172,6 @@ B <- lsa::cosine(t(XU))
 
 Q.inv <- MASS::ginv(XU) %*% Z
 Q <- solve(Q.inv)
-
-q <- Variable(4, 4)
-objective <- Maximize(cvxr_norm(Z %*% q %*% t(Z), 2))
-objective <- Minimize(cvxr_norm(q %*% Z[1, ] %*% t(Z[64, ]) %*% t(q), 2))
-objective <- Minimize(cvxr_norm(Z %*% q, 'fro'))
-constraints <- list(q %*% Ipq %*% t(q) - Ipq == 0)
-constraints <- list(cvxr_norm(q %*% Ipq %*% t(q) - Ipq, 'fro') == 0)
-constraints <- list(cvxr_norm(t(q) - Ipq %*% solve(q) %*% Ipq, 'fro') == 0)
-constraints <- list(matrix_frac(q[, 1], Ipq) == 1,
-                    matrix_frac(q[, 2], Ipq) == 1,
-                    matrix_frac(q[, 3], Ipq) == 1,
-                    matrix_frac(q[, 4], Ipq) == -1)
-constraints <- list(matrix_trace(q) == 2,
-                    sum_squares(q) == 4)
-problem <- Problem(objective, constraints)
-solution <- solve(problem, verbose = TRUE, gp = TRUE)
-solution$getValue(q)
 
 out <- vegan::procrustes(XU, Z)
 

### 2020-07-13 summary.Rmd
finishing up simulations

@@
 title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
-# output: pdf_document
+# output:
+#   pdf_document:
+#     citation_package: natbib
+#     number_sections: yes
 output: html_document
@@
 - \usepackage{mathtools}
+- \usepackage{natbib}
+- \usepackage[linesnumbered,vlined,ruled]{algorithm2e} 
+- \setcitestyle{numbers,square}
+bibliography: misc.bib
 abstract: |
-  this
-  is
-  an
-  abstract
+  In this paper, we connect two probabolistic models for graphs, the Popularity 
+  Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
+  (GRDPG) and use properties established in this connection to aid in 
+  clustering and parameter estimation. In particular, we note that the PABM 
+  can be represented as latent positions such that points within the same 
+  cluster lie on a subspace, and the subspaces that represent different 
+  clusters are orthogonal to one another. Using this property as well as the 
+  asymptotic properties of Adjacency Spectral Embedding of the GRDPG, we are 
+  able to establish theoretical asymptotic results of our clustering and 
+  parameter estimation methods for the PABM.
 ---
@@
 import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
 library(plot.matrix)
+library(mclust)
+library(ggplot2)
 
 set.seed(314159)
+
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+
+doMC::registerDoMC(parallel::detectCores())
+
+embedding <- function(A, p = NULL, q = NULL,
+                      eps = 1e-6) {
+  eigen.A <- eigen(A, symmetric = TRUE)
+  if (is.null(p) | is.null(q)) {
+    keep <- (abs(eigen.A$values) > eps)
+  } else {
+    keep <- c(seq(p), seq(n, n - q + 1))
+  }
+
+  U <- eigen.A$vectors[, keep]
+  S <- diag(sqrt(abs(eigen.A$values[keep])))
+  return(U %*% S)
+}
+
+draw.graph <- function(P) {
+  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
+  A[lower.tri(A)] <- 0
+  diag(A) <- 0
+  A <- A + t(A)
+  return(A)
+}
 ```
@@
 
-# 1 Introduction
+# Introduction
 
-## 1.1 Previous work
+The Popularity Adjusted Block Model (PABM) was introduced by 
+\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the Stochastic 
+Block Model (SBM) to address the heterogeneity of edge probabilities within and 
+between communities or clusters.
 
-# 2 Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
+## Previous work
 
-## 2.1 The popularity adjusted block model (PABM)[^pabm]
+\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
+to identify the cluster memberships given either an edge probability matrix $P$ 
+or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
+memberships can be identified exactly (up to permutation). A similar procedure 
+can be applied if $P$ is unknown and we have an observation $A$, but the 
+theoretical guarantees of this method applied to the PABM are unknown. In 
+particular, the method requires spherical Gaussian noise. The authors of this 
+paper then use point estimators for $\{\lambda^{(kl)}\}$ with the results of 
+SSC.
 
-[^pabm]: https://arxiv.org/abs/1910.01931
+# Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
 
+## The popularity adjusted block model (PABM) \cite{noroozi2019estimation}
+
+**Definition 1**
 Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
@@
 
-## 2.2 The generalized random dot product graph (GRDPG)[^grdpg]
+## The generalized random dot product graph (GRDPG) \cite{rubindelanchy2017statistical}
 
-[^grdpg]: https://arxiv.org/abs/1709.05506
-
+**Definition 2**
 Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
@@
 
-## 2.3 Connecting the PABM to the GRDPG for $K = 2$
+## Connecting the PABM to the GRDPG
 
+### Case where $K = 2$
+
+**Theorem 1**. 
+Let $X = \begin{bmatrix} 
+\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
+0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
+\end{bmatrix}$ 
+where the $\lambda^{(kl)}$'s are defined as in Definition 1 for $K=2$, and 
+let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
+0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
+0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
+0 & 1 & 0 & 0 \end{bmatrix}$. 
+Then $A \sim GRDPG_{3, 1}(3, 1)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
+equivalent.
+
+*Proof*.
 Let $X = \begin{bmatrix} 
@@
 
-## 2.4 Generalization to $K > 2$
+### Generalization to $K > 2$
 
+**Theorem 2**. There exists a block diagonal matrix 
+$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
+$\{\lambda^{(kl)}\}_K$ and $U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
+for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
+$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
+
+*Proof*
 Let $\Lambda^{(k)} = 
@@
 $X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in [0, 1]^{n \times K^2}$.  
-
 Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
@@
 [0, 1]^{n \times K^2}$.  
-
 Then $P = X Y^\top$.  
-
 Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
 $\Pi$, so $P = X \Pi X^\top$.  
-
 The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
@@
 a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
-$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$[^pairs].
-
-[^pairs]: TODO: describe pairs $(s, t)$ in more elegant/succinct closed form
+$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
 
@@
 
-**Example: $K = 3$**
-
-Using the same notation as before:
+**Example** for $K = 3$. Using the same notation as before:
 
@@
 
-Then $P = X Y$ and $Y = X \Pi$ where 
+Then $P = X Y^\top$ and $Y = X \Pi$ where 
 $\Pi = \begin{bmatrix} 
@@
 
-Perhaps a simpler way to look at this is:
+Another way to look at this is:
 
@@
 
-# 3 Methodology
+# Methods
 
@@
 Here, we will focus more on (1) and pose possible methods for (2). In our 
-methods, we assume that $K$, the number of clusters, is known beforehand.
+methods, we assume that $K$, the number of clusters, is known beforehand and 
+does not require estimation.
 
-## 3.1 Cluster membership identification
+## Clustering
 
-### 3.1.1 Using edge probability matrix $P$
+### Using edge probability matrix $P$
 
@@
 value of $K$. However, direct identification or estimation of $XU$ is not 
-possible.
+possible \cite{rubindelanchy2017statistical}.
 
@@
 subspace. Using this property, previous work proposes the use of subspace 
-clustering while acknowledging some of its shortcomings.
+clustering while acknowledging some of its shortcomings 
+\cite{noroozi2019estimation} \cite{soltanolkotabi2014}.
 
-**Theorem 1.** 
-*Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix.*
+**Theorem 3.** 
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability
+matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
+of different clusters.
 
-### 3.1.2 Using adjacency matrix $A$
+*Proof (sketch)*
+By projection, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
+in Theorem 2. Since $X$ is block diagonal with each block corresponding 
+to one cluster, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
+with each block corresponding to a cluster and zeros elsewhere. Therefore, 
+if vertices $i$ and $j$ belong to different clusters, then the $ij$^th^ element 
+of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
 
-## 3.2 Parameter estimation
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \KwData{Edge probability matrix $P$, number of clusters $K$}
+  \KwResult{Cluster assignments $1, ..., K$}
+    Compute the spectral decomposition $P = V D V^\top$\;
+    Compute the inner product matrix $B = V V^\top$\;
+    Identify the connected components and map each to clusters $1..K$\;
+  \caption{PABM clustering on the edge probability matrix}
+\end{algorithm}
 
-# 4 Simulated Examples
+### Using adjacency matrix $A$
+
+The adjacency embedding of $A$ approaches latent positions that form $P$ as 
+the number of vertices $n$ increases. More 
+precisely, let $\{\lambda^{(kl)}\}_K \sim \mathcal{F}_K$ for some joint 
+distribution consisting of $K$ underlying clusters $\mathcal{F}_K$. Then the 
+latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution 
+with $K$ underlying clusters $\mathcal{G}_K$. Denote $Z_n$ as a sample of size 
+$n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge 
+probability matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the 
+adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
+\citet*{rubindelanchy2017statistical},
+
+$$\max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)$$
+
+for some $c>0$ and sequence of $Q_n \in \mathbb{O}(p, q)$. In addition, 
+\citeauthor{rubindelanchy2017statistical} produce a central limit theorem 
+result.
+
+**Theorem 4**. Let $\hat{V}^{(n)} \in \mathbb{R}^{n \times K^2}$ be the matrix 
+of $K^2$ eigenvectors of $A_n$ corresponding to the $K (K+1) / 2$ most positive 
+eigenvalues and $K (K-1) / 2$ most negative eigenvalues with rows
+$(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
+different clusters. Then for some $c > 0$,
+
+$$\max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = O_P\bigg( \frac{(\log n)^c}{n}\bigg)$$
+
+In addition, $(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a generalized 
+chi-square distribution.
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \KwData{Adjacency matrix $A$, number of clusters $K$}
+  \KwResult{Cluster assignments $1, ..., K$}
+    Compute the eigenvectors of $A$ that correspond to the $K (K+1) / 2$ most 
+    positive eigenvalues and $K (K-1) / 2$ most negative eigenvalues. Construct 
+    $V$ using these eigenvectors as its columns.\;
+    Compute $B = V V^\top$.\;
+    Construct graph $G$ using $B$ as its similarity matrix.\;
+    Partition $G$ into $K$ disconnected components 
+    (e.g., using edge thresholding or spectral clustering).\;
+    Map each connected components as the clusters $1, ..., K$.\;
+  \caption{PABM clustering on the adjacency matrix}
+\end{algorithm}
+
+## Parameter estimation
+
+For this section, we will focus on the case $K = 2$. Under this condition, the 
+PABM is equivalent to the GRDPG with signature $(3, 1)$.
+
+The adjacency spectral embeddings of both $A$ and $P$ are not unique. In 
+particular, let $Z$ be the ASE of $P$. Then $P = Z I_{3, 1} Z^\top$. However, 
+for any $Q \in \mathbb{O}(3, 1)$, 
+$(ZQ) I_{3, 1} (ZQ)^\top = Z (Q I_{3, 1} Q^\top) Z^\top = Z I_{3, 1} Z^\top = P$,
+so $ZQ$ is also a valid ASE Of $P$. If we can find $Q \in \mathbb{O}(3, 1)$ 
+such that $ZQ = XU$, we can compute the parameters $\{\lambda^{(kl)}\}$ 
+directly. Furthermore, if we instead use the ASE of the adjacency matrix $A$, 
+$\exists Q \in \mathbb{O}(3, 1)$ such that $\max\limits_i ||Q \hat{z}_i - XU||$ 
+is minimized (and goes to zero under a probabolistic model for
+$\{\lambda^{(kl)}\}$'s).
+
+Thus if we can identify $Q$ such that the $ZQ = XU$ in the case of embedding 
+$P$ or $\hat{Z} Q - XU$ is minimized in the case of embedding $A$, we can 
+identify or estimate the PABM parameters directly from the embedding. $XU$ is 
+unknown, but we can still obtain an embedding $ZQ$ that follows the properties 
+of $XU$, which will yield estimates $\{\hat{\lambda}^{(kl)}\}$ that are valid 
+in that they will produce the edge probability matrix $P$.[^lambdas]
+
+[^lambdas]: Note that the set $\{\lambda^{(kl)}\}_K$ that produces a unique 
+$P$ is not unique \citep{noroozi2019estimation}.
+
+In particular, we can note the following properties of $XU$:
+
+1. If row $i$ is in cluster 1, then its second index is 0. If it is in cluster 
+2, its first index is 0.
+2. If row $i$ is in cluster 1, then its third and fourth indices are equal. If 
+it is in cluster 2, the fourth index is the negative of its third index.
+3. If rows $i$ and $j$ are in different clusters, their dot product is 0.
+
+If we have cluster memberships (either known *a priori* or estimated using a 
+clustering method), then we can estimate $XU$ by picking
+$Q \in \mathbb{O}(3, 1)$ such that $ZQ$ best fits these properties.
+
+$\mathbb{O}(3, 1)$ happens to be the Lorentz group, and each 
+$Q \in \mathbb{O}(3, 1)$ can be represented as the product of six matrices that 
+depend on a total of four parameters \cite{10.1088/978-1-6817-4254-0}. This 
+lends itself as an optimization problem:
+
+$$\begin{split}
+  \min_{\theta, \phi, \psi, \tau} & ||\xi^{(1, 1)}||^2 + ||\xi^{(2, 2)}||^2 + ||\xi^{(1, 3)} - \xi^{(1, 4)}||^2 + ||\xi^{(2, 3)} + \xi^{(2, 4)}||^2 \\
+  \text{s.t. } & \Xi = Z Q(\theta, \phi, \psi, \tau) \\
+  & \xi^{(i, j)} \text{ is the } j^{th} \text{ column of } \Xi \text{ with rows from the } i^{th} \text{ cluster}
+\end{split}$$
+
+Note that if $Q$ can be properly estimated, the asymptotic results from 
+\citet{rubindelanchy2017statistical} can be applied here.
+
+# Simulated Examples
+
+For these examples, we will set the following parameters:
+
+* $K = 2$
+* Mixture parameter $\alpha = .5$
+* $\lambda^{(kk)} \stackrel{indep}{\sim} Beta(2, 1)$
+* $\lambda^{(kl)} \stackrel{indep}{\sim} Beta(1, 2)$ for $k \neq l$
+
+```{r}
+K <- 2
+p <- K * (K + 1) / 2
+q <- K * (K - 1) / 2
+alpha <- .5
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+```
+
+## Clustering
+
+In this part, we will assess Algorithm 1's performance for sample sizes 
+$n = 64, 128, 256, 512, 1024$. For each sample size $n$, $100$ sets of 
+$\lambda^{(kl)}$'s are drawn and for each set of parameters, one adjacency 
+matrix $A$ is drawn and clustered.
+
+```{r}
+n.vec <- c(64, 128, 256, 512, 1024)
+iter <- 100
+```
+
+```{r}
+draw.pabm.beta.2 <- function(n, a1, b1, a2, b2, alpha = .5) {
+  z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
+    z <- sort(z)
+
+    n1 <- sum(z == 1)
+    n2 <- sum(z == 2)
+
+    lambda11 <- rbeta(n1, a1, b1)
+    lambda22 <- rbeta(n2, a1, b1)
+    lambda12 <- rbeta(n1, a2, b2)
+    lambda21 <- rbeta(n2, a2, b2)
+
+    X <- cbind(c(lambda11, rep(0, n2)),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(rep(0, n1), lambda22))
+    Y <- cbind(c(lambda11, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda22))
+    P <- X %*% t(Y)
+    A <- draw.graph(P)
+
+    return(list(A = A, z = z))
+}
+
+normalized.laplacian <- function(W) {
+  n <- nrow(W)
+  D.neg.sqrt <- diag(colSums(W) ** -.5)
+  I <- diag(n)
+  return(I - D.neg.sqrt %*% W %*% D.neg.sqrt)
+}
+
+cluster.pabm <- function(A, K) {
+  p <- K * (K + 1) / 2
+  q <- K * (K - 1) / 2
+  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(p), seq(n, n - q + 1))]
+  V <- V / mean(V)
+  B <- abs(V %*% t(V))
+  # L <- graph.laplacian(B)
+  L <- normalized.laplacian(B)
+  eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - K)]
+  clustering <- mclust::Mclust(eigenmap, K)$classification
+}
+```
+
+```{r}
+results.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
+  plyr::ldply(seq(iter), function(i) {
+    Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
+    A <- Az$A
+    z <- Az$z
+    clustering <- cluster.pabm(A, K)
+    accuracy <- mean(clustering == z)
+    if (accuracy < .5) accuracy <- 1 - accuracy
+    dplyr::tibble(n = n, accuracy = accuracy) %>% 
+      return()
+  }, .parallel = TRUE) %>% 
+    return()
+}
+
+results.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mean.acc = mean(accuracy),
+                   sd.acc = sd(accuracy)) %>% 
+  ggplot() + 
+  labs(y = 'accuracy') + 
+  geom_line(aes(x = n, y = mean.acc)) + 
+  geom_errorbar(aes(x = n, ymin = mean.acc - sd.acc, ymax = mean.acc + sd.acc))
+```
+
+## Parameter estimation
+
+# References

### 2020-07-13 pabm-scratch.R

@@ -38,10 +38,10 @@ z <- c(rep(1, n1),
 
 Ipq <- diag(c(1, 1, 1, -1))
 
-lambda11 <- rbeta(n1, 5, 1)
-lambda22 <- rbeta(n2, 5, 1)
-lambda12 <- rbeta(n1, 1, 5)
-lambda21 <- rbeta(n2, 1, 5)
+lambda11 <- rbeta(n1, 2, 1)
+lambda22 <- rbeta(n2, 2, 1)
+lambda12 <- rbeta(n1, 1, 2)
+lambda21 <- rbeta(n2, 1, 2)
 
 X <- cbind(c(lambda11, rep(0, n2)),
            c(lambda12, rep(0, n2)),

### 2020-07-13 summary.Rmd
updates on writeup

@@
 title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
-# output:
-#   pdf_document:
-#     citation_package: natbib
-#     number_sections: yes
-output: html_document
+output:
+  pdf_document:
+    citation_package: natbib
+    number_sections: yes
+# output: html_document
 # geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
@@
 library(ggplot2)
+
+theme_set(theme_bw())
 
@@
 
-$$\begin{split}
+\begin{equation}
+\begin{split}
   \min_{\theta, \phi, \psi, \tau} & ||\xi^{(1, 1)}||^2 + ||\xi^{(2, 2)}||^2 + ||\xi^{(1, 3)} - \xi^{(1, 4)}||^2 + ||\xi^{(2, 3)} + \xi^{(2, 4)}||^2 \\
@@
   & \xi^{(i, j)} \text{ is the } j^{th} \text{ column of } \Xi \text{ with rows from the } i^{th} \text{ cluster}
-\end{split}$$
+\end{split}
+\end{equation}
 
@@
 matrix $A$ is drawn and clustered.
+
+We will not consider clustering using the edge probability matrix $P$ since 
+this will always result exact recovery of the original clusters.
 
@@
 
-```{r}
+```{r fig.cap = 'IQR clustering error using Algorithm 1 for sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.', cache = TRUE}
 results.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
@@
     clustering <- cluster.pabm(A, K)
-    accuracy <- mean(clustering == z)
-    if (accuracy < .5) accuracy <- 1 - accuracy
-    dplyr::tibble(n = n, accuracy = accuracy) %>% 
+    error <- mean(clustering == z)
+    if (error > .5) error <- 1 - error
+    dplyr::tibble(n = n, error = error) %>% 
       return()
@@
   dplyr::group_by(n) %>% 
-  dplyr::summarise(mean.acc = mean(accuracy),
-                   sd.acc = sd(accuracy)) %>% 
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75)) %>% 
   ggplot() + 
-  labs(y = 'accuracy') + 
-  geom_line(aes(x = n, y = mean.acc)) + 
-  geom_errorbar(aes(x = n, ymin = mean.acc - sd.acc, ymax = mean.acc + sd.acc))
+  labs(y = 'error') + 
+  geom_line(aes(x = n, y = med.err)) + 
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q))
+```
+
+We can also examine how the distribution of $(\hat{v}_i)^\top \hat{v_j}$ varies 
+with $n$:
+
+```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', cache = TRUE}
+densities.df <- plyr::ldply(n.vec, function(n) {
+  Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
+  A <- Az$A
+  z <- Az$z
+  n1 <- sum(z == 1)
+  n2 <- sum(z == 2)
+  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(3), n)]
+  B <- V %*% t(V)
+  inner.prods <- as.numeric(B[(n1+1):n, (n1+1):n])
+  return(dplyr::tibble(n = n, inner.prods = inner.prods))
+}, .parallel = TRUE)
+
+ggplot(densities.df) + 
+  geom_density(aes(x = inner.prods, colour = factor(n))) + 
+  labs(x = 'between cluster inner products',
+       colour = 'n') + 
+  scale_colour_brewer(palette = 'Set1')
 ```
@@
 
+```{r}
+rot.mat.2 <- function(angle, axis = 1) {
+  # rotation matrix for K=2
+
+  if (axis %in% c(1, 3)) {
+    rotation <- matrix(c(cos(angle), -sin(angle), sin(angle), cos(angle)),
+                       nrow = 2, ncol = 2, byrow = TRUE)
+    if (axis == 1) {
+      return(as.matrix(Matrix::bdiag(1, rotation, -1)))
+    } else {
+      return(as.matrix(Matrix::bdiag(rotation, 1, -1)))
+    }
+  } else if (axis == 2) {
+    return(matrix(c(cos(angle), 0, sin(angle), 0,
+                    0, 1, 0, 0,
+                    -sin(angle), 0, cos(angle), 0,
+                    0, 0, 0, -1),
+                  nrow = 4, ncol = 4,
+                  byrow = TRUE))
+  } else {
+    stop(simpleError('axis must be 1, 2, or 3'))
+  }
+}
+
+hyp.rot.mat.2 <- function(angle, axis = 1) {
+  # hyperbolic rotation matrix for $K=2
+
+  if (axis == 1) {
+    return(matrix(c(1, 0, 0, 0,
+                    0, 1, 0, 0, 
+                    0, 0, cosh(angle), sinh(angle),
+                    0, 0, sinh(angle), cosh(angle)), 
+                  byrow = TRUE,
+                  nrow = 4, ncol = 4))
+  } else if (axis == 2) {
+    return(matrix(c(1, 0, 0, 0,
+                    0, cosh(angle), 0, sinh(angle),
+                    0, 0, 1, 0,
+                    0, sinh(angle), 0, cosh(angle)),
+                  byrow = TRUE,
+                  nrow = 4, ncol = 4))
+  } else if (axis == 3) {
+    return(matrix(c(cosh(angle), 0, 0, sinh(angle),
+                    0, 1, 0, 0, 
+                    0, 0, 1, 0, 
+                    sinh(angle), 0, 0, cosh(angle)),
+                  byrow = TRUE,
+                  nrow = 4, ncol = 4))
+  } else {
+    stop(simpleError('axis must be 1, 2, or 3'))
+  }
+}
+
+construct.Q <- function(angles) {
+  rot.angles <- angles[1:3]
+  hyp.angles <- c(angles[1], angles[4], -angles[4])
+  rot.matrices <- lapply(seq(3), function(i) rot.mat.2(rot.angles[i], i))
+  hyp.matrices <- lapply(seq(3), function(i) hyp.rot.mat.2(hyp.angles[i], i))
+  matrices <- c(rot.matrices, hyp.matrices)
+  return(Reduce(`%*%`, matrices))
+}
+
+obj.fun <- function(angles, Z.hat, clusters) {
+  R <- construct.Q(angles)
+
+  XU.c <- Z.hat %*% R
+  in.prods <- XU.c[clusters == 1, ] %*% t(XU.c[clusters == 2, , drop = FALSE])
+  return(
+    sum(XU.c[clusters == 1, 1] ** 2) + 
+      sum(XU.c[clusters == 2, 2] ** 2) + 
+      sum((XU.c[clusters == 1, 3] + XU.c[clusters == 1, 4]) ** 2) + 
+      sum((XU.c[clusters == 2, 3] - XU.c[clusters == 2, 4]) ** 2)
+  )
+}
+
+transform.embedding <- function(Z.hat, clusters) {
+  # init.angles <- runif(4, -pi, pi)
+  init.angles <- rep(0, 4)
+  angles.1 <- optim(init.angles, function(x) obj.fun(x, Z.hat, clusters), 
+                    method = 'SANN')$par
+  angles.2 <- optim(angles.1, function(x) obj.fun(x, Z.hat, clusters),
+                    method = 'Nelder-Mead',
+                    control = list(abstol = 1e-10, maxit = 1e3))$par
+  R <- construct.Q(angles.2)
+  return(Z.hat %*% R)
+}
+```
+
+```{r}
+n <- 64
+z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
+z <- sort(z)
+
+n1 <- sum(z == 1)
+n2 <- sum(z == 2)
+
+lambda11 <- rbeta(n1, a1, b1)
+lambda22 <- rbeta(n2, a1, b1)
+lambda12 <- rbeta(n1, a2, b2)
+lambda21 <- rbeta(n2, a2, b2)
+
+X <- cbind(c(lambda11, rep(0, n2)),
+           c(lambda12, rep(0, n2)),
+           c(rep(0, n1), lambda21),
+           c(rep(0, n1), lambda22))
+Y <- cbind(c(lambda11, rep(0, n2)),
+           c(rep(0, n1), lambda21),
+           c(lambda12, rep(0, n2)),
+           c(rep(0, n1), lambda22))
+P <- X %*% t(Y)
+A <- draw.graph(P)
+Z <- embedding(P)
+Zhat <- embedding(A, 3, 1)
+```
+
+Setting $n = 64$ and generating one example set of $\{\lambda^{(kl)}\}$, we can 
+see the latent positions $XU$ follow the expected properties 
+(Fig. \ref{fig:xu}).
+
+```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework. Note that the the points follow the properties outlined in Section 3.2.'}
+perm <- cbind(c(1, 0, 0, 0),
+              c(0, 0, 1, 0),
+              c(0, 1, 0, 0),
+              c(0, 0, 0, 1))
+U <- eigen(perm)$vectors[, c(3, 1, 2, 4)]
+XU <- X %*% U
+
+pairs(XU, asp = 1, col = z)
+```
+
+If we instead take the ASE of $P = X \Pi X^\top$, we fail to obtain an 
+embedding that has the same properties (Fig. \ref{fig:ase_of_p}).
+
+```{r ase_of_p, fig.cap = 'The adjacency spectral embedding of the edge probability matrix $P$.
+Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.'}
+pairs(Z, asp = 1, col = z)
+```
+
+Denoting $Z$ as the ASE of $P$, we can recover $ZQ$ that is equivalent to $XU$ 
+(Fig. \ref{fig:q_trans}).
+
+```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to a 90-degree rotation.'}
+XU.prime <- transform.embedding(Z, z)
+pairs(XU.prime, col = z, asp = 1)
+```
+
+The ASE of the adjacency matrix $A$ (as one realization of $P$) has the 
+resulting pairs plot:
+
+```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.'}
+pairs(Zhat, col = z, asp = 1)
+```
+
+Minimizing the objective function in (1) results in the following embedding:
+
+```{r}
+XU.hat <- transform.embedding(Zhat, z)
+pairs(XU.hat, col = z, asp = 1)
+```
+
 # References

### 2020-07-13 summary.Rmd
minor updates to writeup

@@
   \KwResult{Cluster assignments $1, ..., K$}
-    Compute the spectral decomposition $P = V D V^\top$\;
-    Compute the inner product matrix $B = V V^\top$\;
-    Identify the connected components and map each to clusters $1..K$\;
+    Compute the spectral decomposition $P = V D V^\top$.\;
+    Compute the inner product matrix $B = V V^\top$.\;
+    Construct graph $G$ using $B$ as edge similarities.\;
+    Identify the connected components and map each to cluster labels $1, ..., K.$\;
   \caption{PABM clustering on the edge probability matrix}
@@
     (e.g., using edge thresholding or spectral clustering).\;
-    Map each connected components as the clusters $1, ..., K$.\;
+    Map each of the connected components to the cluster labels $1, ..., K$.\;
   \caption{PABM clustering on the adjacency matrix}
@@
 
-```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to a 90-degree rotation.'}
+```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.'}
 XU.prime <- transform.embedding(Z, z)
@@
 
-```{r}
+```{r q_from_a, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
 XU.hat <- transform.embedding(Zhat, z)
 pairs(XU.hat, col = z, asp = 1)
+
+# X.hat <- XU.hat %*% t(U)
+# plot(abs(X.hat) > .25)
 ```

### 2020-07-13 summary.Rmd
minor updates to writeup

@@
 particular, the method requires spherical Gaussian noise. The authors of this 
-paper then use point estimators for $\{\lambda^{(kl)}\}$ with the results of 
-SSC.
+paper then use point estimators for $\{\lambda^{(kl)}\}$ using cluster labels 
+obtained via SSC.
 
@@
 
-**Definition 1**
+**Definition 1**.
 Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
@@
 
-**Definition 2**
+**Definition 2**.
 Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
@@
 matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.
+
+**Remark**. 
+Like the "ungeneralized" random dot product graph, the latent positions of a 
+generalized random dot product graph (RDPG) are not unique. More specifically, 
+if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
+$Q \in \{R : R^\top I_{pq} R = I_{pq} \}$, 
+$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
+x_i^\top I_{pq} x_j = P_{ij}$ (this is the indefinite orthogonal group, 
+$\mathbb{O}(p, q)$).
+Unlike in the RDPG case, transforming the latent positions by multiplication 
+with $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
+distances.
 

### 2020-07-14 summary.Rmd
minor updates to writeup

@@
   can be represented as latent positions such that points within the same 
-  cluster lie on a subspace, and the subspaces that represent different 
-  clusters are orthogonal to one another. Using this property as well as the 
-  asymptotic properties of Adjacency Spectral Embedding of the GRDPG, we are 
-  able to establish theoretical asymptotic results of our clustering and 
+  cluster lie on a subspace, and the subspaces that represent each 
+  cluster are orthogonal to one another. Using this property as well as the 
+  asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
+  are able to establish theoretical asymptotic results of our clustering and 
   parameter estimation methods for the PABM. 
@@
 
+One popular probabolistic model for graphs with underlying communities is the 
+Stochastic Block Model (SBM), which assumes that the edge probability from 
+one cluster to another follows a Beronulli distribution with a fixed 
+probability for each pair of communities. 
 The Popularity Adjusted Block Model (PABM) was introduced by 
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the Stochastic 
-Block Model (SBM) to address the heterogeneity of edge probabilities within and 
-between communities or clusters.
+\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
+address the heterogeneity of edge probabilities within and between communities 
+or clusters while still maintaining the community structure.
+
+The Random Dot Product Graph (RDPG) \cite{athreya2017statistical} is another 
+probabolistic model with Beronulli edge probabilities. Under this model, each 
+vertex of the graph can be represented by a point in latent space such that 
+the edge probability between any pair of vertices is given by their 
+corresponding dot product in the latent space. The SBM is equivalent to a 
+special case of the RDPG model in which all vertices in a given community share 
+the same position in the latent space. An analogous property exists for the 
+PABM not under the RDPG model but under the *Generalized* Random Dot Product 
+Graph model.
 
@@
     Construct graph $G$ using $B$ as edge similarities.\;
-    Identify the connected components and map each to cluster labels $1, ..., K.$\;
+    Identify the connected components of $G$ and map each to cluster labels 
+    $1, ..., K.$\;
   \caption{PABM clustering on the edge probability matrix}
@@
     (e.g., using edge thresholding or spectral clustering).\;
-    Map each of the connected components to the cluster labels $1, ..., K$.\;
+    Map each partition to the cluster labels $1, ..., K$.\;
   \caption{PABM clustering on the adjacency matrix}

### 2020-07-14 summary.Rmd
minor updates to writeup

@@
 0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
-\end{bmatrix}$ 
-where the $\lambda^{(kl)}$'s are defined as in Definition 1 for $K=2$, and 
-let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
+\end{bmatrix}$, and let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
 0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
 0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}$. 
-Then $A \sim GRDPG_{3, 1}(3, 1)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
+0 & 1 & 0 & 0 \end{bmatrix}$, as in Definition 1. 
+Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
 equivalent.
@@
 \begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
-\in [0, 1]^{n_k \times K}$.  
+\in \mathbb{R}^{n_k \times K}$.  
 Let $X$ be a block diagonal matrix 
-$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in [0, 1]^{n \times K^2}$.  
+$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
 Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
 $l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
-[0, 1]^{n \times K}$.  
+\mathbb{R}^{n \times K}$.  
 Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
-[0, 1]^{n \times K^2}$.  
+\mathbb{R}^{n \times K^2}$.  
 Then $P = X Y^\top$.  
@@
 
-In addition, $(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a generalized 
-chi-square distribution.
+In addition, by the central limit theorem, 
+$(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a normal distribution 
+centered at $0$.
 

### 2020-07-14 summary.Rmd
cleaned up some typos and such

@@
 **Remark**. 
-Like the "ungeneralized" random dot product graph, the latent positions of a 
-generalized random dot product graph (RDPG) are not unique. More specifically, 
-if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
+Like the RDPG, the latent positions of a GRDPG are not unique. More 
+specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
 $Q \in \{R : R^\top I_{pq} R = I_{pq} \}$, 
@@
 the exact edge probability matrix, we cannot recover the "original" latent 
-positions $XU$. Note that unlike in the case of the regular random dot product 
-graph, $Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
+positions $XU$. Note that unlike in the case of the RDPG, 
+$Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
 $||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
@@
 \begin{split}
-  \min_{\theta, \phi, \psi, \tau} & ||\xi^{(1, 1)}||^2 + ||\xi^{(2, 2)}||^2 + ||\xi^{(1, 3)} - \xi^{(1, 4)}||^2 + ||\xi^{(2, 3)} + \xi^{(2, 4)}||^2 \\
+  \min_{\theta, \phi, \psi, \tau} & ||\xi^{(1, 2)}||^2 + ||\xi^{(2, 1)}||^2 + 
+  ||\xi^{(1, 3)} - \xi^{(1, 4)}||^2 + ||\xi^{(2, 3)} + \xi^{(2, 4)}||^2 \\
   \text{s.t. } & \Xi = Z Q(\theta, \phi, \psi, \tau) \\
@@
   return(
-    sum(XU.c[clusters == 1, 1] ** 2) + 
-      sum(XU.c[clusters == 2, 2] ** 2) + 
-      sum((XU.c[clusters == 1, 3] + XU.c[clusters == 1, 4]) ** 2) + 
-      sum((XU.c[clusters == 2, 3] - XU.c[clusters == 2, 4]) ** 2)
+    sum(XU.c[clusters == 1, 2] ** 2) + 
+      sum(XU.c[clusters == 2, 1] ** 2) + 
+      sum((XU.c[clusters == 1, 3] - XU.c[clusters == 1, 4]) ** 2) + 
+      sum((XU.c[clusters == 2, 3] + XU.c[clusters == 2, 4]) ** 2)
   )
@@
 
-transform.embedding <- function(Z.hat, clusters) {
+transform.embedding <- function(Z.hat, clusters, init.angles) {
   # init.angles <- runif(4, -pi, pi)
-  init.angles <- rep(0, 4)
+  # init.angles <- rep(0, 4)
   angles.1 <- optim(init.angles, function(x) obj.fun(x, Z.hat, clusters), 
@@
 ```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.'}
-XU.prime <- transform.embedding(Z, z)
+XU.prime <- transform.embedding(Z, z, runif(4, 0, 0))
 pairs(XU.prime, col = z, asp = 1)
@@
 ```{r q_from_a, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
-XU.hat <- transform.embedding(Zhat, z)
+XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
 pairs(XU.hat, col = z, asp = 1)

### 2020-07-14 summary.Rmd
cleaned up some typos and such

@@
         xtable.table.placement = 'H')
+```
 
+```{r}
 import::from(magrittr, `%>%`)
@@
 theme_set(theme_bw())
-
-set.seed(314159)
 
@@
 ```{r fig.cap = 'IQR clustering error using Algorithm 1 for sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.', cache = TRUE}
+set.seed(314159)
+
 results.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
@@
 ```{r}
+set.seed(123456)
+
 n <- 64
@@
 ```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.'}
-XU.prime <- transform.embedding(Z, z, runif(4, 0, 0))
+XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
 pairs(XU.prime, col = z, asp = 1)

### 2020-07-14 summary.Rmd
cleaned up some typos and such

@@
 abstract: |
-  In this paper, we connect two probabolistic models for graphs, the Popularity 
+  In this paper, we connect two probabilistic models for graphs, the Popularity 
   Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
@@
 
-One popular probabolistic model for graphs with underlying communities is the 
+One popular probabilistic model for graphs with underlying communities is the 
 Stochastic Block Model (SBM), which assumes that the edge probability from 
@@
 The Random Dot Product Graph (RDPG) \cite{athreya2017statistical} is another 
-probabolistic model with Beronulli edge probabilities. Under this model, each 
+probabilistic model with Beronulli edge probabilities. Under this model, each 
 vertex of the graph can be represented by a point in latent space such that 
@@
 special case of the RDPG model in which all vertices in a given community share 
-the same position in the latent space. An analogous property exists for the 
+the same position in the latent space. It has also been shown that similar 
+probabilistic graph models, (such as the Mixed Membership Stochastic Block Model
+\cite{rubindelanchy2017statistical}). An analogous property exists for the 
 PABM not under the RDPG model but under the *Generalized* Random Dot Product 
@@
 $\exists Q \in \mathbb{O}(3, 1)$ such that $\max\limits_i ||Q \hat{z}_i - XU||$ 
-is minimized (and goes to zero under a probabolistic model for
+is minimized (and goes to zero under a probabilistic model for
 $\{\lambda^{(kl)}\}$'s).

### 2020-07-14 summary.Rmd
cleaned up some typos and such

@@
     Compute the spectral decomposition $P = V D V^\top$.\;
-    Compute the inner product matrix $B = V V^\top$.\;
+    Compute the inner product matrix $B = |V V^\top|$, applying $|\cdot|$ 
+    entry-wise.\;
     Construct graph $G$ using $B$ as edge similarities.\;
@@
     $V$ using these eigenvectors as its columns.\;
-    Compute $B = V V^\top$.\;
+    Compute $B = |V V^\top|$, applying $|\cdot|$ entry-wise.\;
     Construct graph $G$ using $B$ as its similarity matrix.\;

### 2020-07-19 summary.Rmd
added parameter estimation methods discussed previously, removed lorentz transformation based methods, still working on adding comparisons to other methods

@@
 - \setcitestyle{numbers,square}
+- \usepackage{verbatim}
 bibliography: misc.bib
@@
     $1, ..., K.$\;
-  \caption{PABM clustering on the edge probability matrix}
+  \caption{PABM clustering on the edge probability matrix.}
 \end{algorithm}
@@
     Map each partition to the cluster labels $1, ..., K$.\;
-  \caption{PABM clustering on the adjacency matrix}
+  \caption{PABM clustering on the adjacency matrix.}
 \end{algorithm}
@@
 ## Parameter estimation
+
+For any $P$ edge probability matrix for the PABM such that the rows and 
+columns are organized by cluster, the $kl$^th^ block is an outer product 
+of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
+Therefore, given $P^{(kl)}$, $\lambda%{(kl)}$ and $\lambda^{(lk)}$ are 
+solvable exactly (up to multiplication by $-1$) using singular value 
+decomposition. More specifically, let $P = \sigma^2 u v^\top$ be the 
+singular value decomposition of $P$. $u \in \mathbb{R}^{n_k}$ and 
+$v \in \mathbb{R}^{n_l}$ are vectors and 
+$\sigma^2 > 0$ is a scalar. Then $\lambda^{(kl)} = \pm \sigma u$ and 
+$\lambda^{(lk)} = \pm \sigma v$.
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{PABM parameter estimation using the edge probability matrix.}
+  \KwData{Edge probability matrix $P$, cluster assignments $1, ..., K$.}
+  \KwResult{PABM parameters $\{\lambda^{(kl)}\}_K$}
+  Arrange the rows and columns of $P$ by cluster such that each $P^{(kl)}$ 
+  block consists of edge probabilities between clusters $k$ and $l$.\; 
+  \For{$k, l = 1, ..., K$, $k \geq l$} {
+    Compute $P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top$, the SVD of 
+    the $kl^{th}$ block.\;
+    Assign $\lambda^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
+    $\lambda^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
+  }
+\end{algorithm}
+
+A similar method can be applied using $\hat{Z}$, the adjacency spectral 
+embedding of $A$.
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{PABM parameter estimation using the adjacency matrix.}
+  \KwData{Adjacency matrix $A$, cluster assignments $1, ..., K$}
+  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
+  Assign $p \leftarrow K (K + 1) / 2$ and $q \leftarrow K (K - 1) / 2$.\;
+  Compute $\hat{Z}$, the adjacency spectral embedding of $A$ using signature
+  $(p, q)$.\;
+  Compute $\hat{P} = \hat{Z} I_{pq} \hat{Z}^\top$, the estimate of the edge 
+  probability matrix based on the adjacency spectral embedding.\;
+  Arrange the rows and columns of $\hat{P}$ by cluster such that each 
+  $\hat{P}^{(kl)}$ block consists of estimated edge probabilities between 
+  clusters $k$ and $l$.\;
+  \For{$k, l = 1, ..., K$, $k \leq l$} {
+    Compute $\hat{P}^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+    block.\;
+    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
+    Assign $\sigma^2 \leftarrow \Sigma_{11}$.\;
+    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
+    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
+  }
+\end{algorithm}
+
+**Theorem 5**. Let $\lambda^{(kl)}_{n_k} \sim F_{n_k}$ for $k, l = 1, ..., K$, 
+and let $\{\hat{\lambda}^{(kl)}_{n_k}\}_K$ be the set of estimates computed by 
+Algorithm 4. 
+Then ... ***TODO***
+
+\begin{comment}
+
+### Using Adjacency Spectral Embedding
 
@@
 
+\end{comment}
+
 # Simulated Examples
@@
 
-In this part, we will assess Algorithm 1's performance for sample sizes 
+In this part, we will assess Algorithm 2's performance for sample sizes 
 $n = 64, 128, 256, 512, 1024$. For each sample size $n$, $100$ sets of 
@@
 ## Parameter estimation
+
+\begin{comment}
 
@@
 
-```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.'}
+```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1).
+This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE}
 XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
@@
 
-```{r q_from_a, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
+```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
 XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
@@
 
+\end{comment}
+
+```{r}
+estimate.lambda.block <- function(P.block, within = FALSE) {
+  if (within) {
+    P.svd <- svd(P.block)
+    u <- P.svd$u[, 1]
+    if (mean(u) < 0) u <- -u
+    s <- sqrt(P.svd$d[1])
+    return(s * u)
+  } else {
+    P.svd <- svd(P.block)
+    u <- P.svd$u[, 1]
+    if (mean(u) < 0) u <- -u
+    v <- P.svd$v[, 1]
+    if (mean(v) < 0) v <- -v
+    s <- sqrt(P.svd$d[1])
+    return(list(s * u, s * v))
+  }
+}
+
+lambda.rmse <- function(lambda.matrix, Phat, clustering) {
+  K <- max(clustering)
+  n <- nrow(Phat)
+  n.vector <- sapply(seq(K), function(k) sum(clustering == k))
+  sapply(seq(K), function(k) {
+    low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
+    high.ind.k <- sum(n.vector[seq(k)])
+    sapply(seq(k), function(l) {
+      P.block <- Phat[clustering == k, clustering == l]
+      if (k == l) {
+        lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
+        lambda.hat <- estimate.lambda.block(P.block, TRUE)
+        return(sum((lambda - lambda.hat) ** 2))
+      } else {
+        low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
+        high.ind.l <- sum(n.vector[seq(l)])
+        lambda.kl <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
+        lambda.lk <- as.numeric(lambda.matrix[seq(low.ind.l, high.ind.l), k])
+        lambda.hats <- estimate.lambda.block(P.block)
+        return(sum((lambda.kl - lambda.hats[[1]]) ** 2) + 
+                 sum((lambda.lk - lambda.hats[[2]]) ** 2))
+      }
+    }) %>% 
+      sum() %>% 
+      return()
+  }) %>% 
+    sum() %>% 
+    magrittr::divide_by(n * K) %>% 
+    sqrt() %>% 
+    return()
+}
+```
+
+Figure \ref{fig:lambda_est} shows the medians and interquartile ranges of root 
+mean square errors for Algorithm 4 over 100 simulations using the same 
+parameters as before.
+
+```{r lambda_est, cache = TRUE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
+set.seed(314159)
+
+Ipq <- diag(c(1, 1, 1, -1))
+
+rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  plyr::ldply(seq(iter), function(i) {
+    z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
+    z <- sort(z)
+
+    n1 <- sum(z == 1)
+    n2 <- sum(z == 2)
+
+    lambda11 <- rbeta(n1, a1, b1)
+    lambda22 <- rbeta(n2, a1, b1)
+    lambda12 <- rbeta(n1, a2, b2)
+    lambda21 <- rbeta(n2, a2, b2)
+
+    X <- cbind(c(lambda11, rep(0, n2)),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(rep(0, n1), lambda22))
+    Y <- cbind(c(lambda11, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda22))
+    P <- X %*% t(Y)
+    A <- draw.graph(P)
+    lambda.matrix <- cbind(c(lambda11, lambda21),
+                           c(lambda12, lambda22))
+    Zhat <- embedding(A, p, q)
+    Phat <- Zhat %*% Ipq %*% t(Zhat)
+    rmse <- lambda.rmse(lambda.matrix, Phat, z)
+    dplyr::tibble(n = n, rmse = rmse) %>% 
+      return()
+  }, .parallel = TRUE) %>% 
+    return()
+}
+
+rmse.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = median.rmse)) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse)) + 
+  labs(y = 'RMSE')
+```
+
 # References

### 2020-07-19 pabm-scratch.R

@@ -410,3 +410,8 @@ S <- matrix(c(cosh(angle), sinh(angle), 0, 0,
               0, 0, 0, 1), 
             byrow = TRUE, 
             nrow = 4, ncol = 4)
+
+
+p11 <- P[1:n1, 1:n1]
+p22 <- P[(n1+1):n, (n1+1):n]
+p12 <- P[1:n1, (n1+1):n]

### 2020-07-22 summary.Rmd
still working on comparisons

@@
   ggplot() + 
+  # scale_y_log10() + 
   labs(y = 'error') + 
@@
   ggplot() + 
+  scale_y_log10() + 
+  scale_x_log10() + 
   geom_line(aes(x = n, y = median.rmse)) + 

### 2020-07-22 summary.Rmd
moving stuff from rmarkdown to separate simulation scripts

@@
 import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`)
 library(plot.matrix)
-library(mclust)
 library(ggplot2)
-
-theme_set(theme_bw())
-
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-
-doMC::registerDoMC(parallel::detectCores())
-
-embedding <- function(A, p = NULL, q = NULL,
-                      eps = 1e-6) {
-  eigen.A <- eigen(A, symmetric = TRUE)
-  if (is.null(p) | is.null(q)) {
-    keep <- (abs(eigen.A$values) > eps)
-  } else {
-    keep <- c(seq(p), seq(n, n - q + 1))
-  }
-
-  U <- eigen.A$vectors[, keep]
-  S <- diag(sqrt(abs(eigen.A$values[keep])))
-  return(U %*% S)
-}
-
-draw.graph <- function(P) {
-  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
-  A[lower.tri(A)] <- 0
-  diag(A) <- 0
-  A <- A + t(A)
-  return(A)
-}
 ```
@@
 
-```{r}
-K <- 2
-p <- K * (K + 1) / 2
-q <- K * (K - 1) / 2
-alpha <- .5
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
-```
-
 ## Clustering
@@
 
-```{r}
-n.vec <- c(64, 128, 256, 512, 1024)
-iter <- 100
-```
-
-```{r}
-draw.pabm.beta.2 <- function(n, a1, b1, a2, b2, alpha = .5) {
-  z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
-  z <- sort(z)
-
-  n1 <- sum(z == 1)
-  n2 <- sum(z == 2)
-
-  lambda11 <- rbeta(n1, a1, b1)
-  lambda22 <- rbeta(n2, a1, b1)
-  lambda12 <- rbeta(n1, a2, b2)
-  lambda21 <- rbeta(n2, a2, b2)
-
-  X <- cbind(c(lambda11, rep(0, n2)),
-             c(lambda12, rep(0, n2)),
-             c(rep(0, n1), lambda21),
-             c(rep(0, n1), lambda22))
-  Y <- cbind(c(lambda11, rep(0, n2)),
-             c(rep(0, n1), lambda21),
-             c(lambda12, rep(0, n2)),
-             c(rep(0, n1), lambda22))
-  P <- X %*% t(Y)
-  A <- draw.graph(P)
-
-  return(list(A = A, z = z))
-}
-
-normalized.laplacian <- function(W) {
-  n <- nrow(W)
-  D.neg.sqrt <- diag(colSums(W) ** -.5)
-  I <- diag(n)
-  return(I - D.neg.sqrt %*% W %*% D.neg.sqrt)
-}
-
-cluster.pabm <- function(A, K) {
-  p <- K * (K + 1) / 2
-  q <- K * (K - 1) / 2
-  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(p), seq(n, n - q + 1))]
-  V <- V / mean(V)
-  B <- abs(V %*% t(V))
-  # L <- graph.laplacian(B)
-  L <- normalized.laplacian(B)
-  eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - K)]
-  clustering <- mclust::Mclust(eigenmap, K)$classification
-}
-```
-
-```{r fig.cap = 'IQR clustering error using Algorithm 1 for sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.', cache = TRUE}
-set.seed(314159)
-
-results.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
-  plyr::ldply(seq(iter), function(i) {
-    Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
-    A <- Az$A
-    z <- Az$z
-    clustering <- cluster.pabm(A, K)
-    error <- mean(clustering == z)
-    if (error > .5) error <- 1 - error
-    dplyr::tibble(n = n, error = error) %>% 
-      return()
-  }, .parallel = TRUE) %>% 
-    return()
-}
-
-results.df %>% 
+```{r fig.cap = 'IQR clustering error using Algorithm 1 for sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.', cache = FALSE}
+clustering.df <- readr::read_csv('clustering.csv')
+clustering.df %>%
   dplyr::group_by(n) %>%
@@
 
-```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', cache = TRUE}
-densities.df <- plyr::ldply(n.vec, function(n) {
-  Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
-  A <- Az$A
-  z <- Az$z
-  n1 <- sum(z == 1)
-  n2 <- sum(z == 2)
-  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(3), n)]
-  B <- V %*% t(V)
-  inner.prods <- as.numeric(B[(n1+1):n, (n1+1):n])
-  return(dplyr::tibble(n = n, inner.prods = inner.prods))
-}, .parallel = TRUE)
+```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', cache = FALSE}
+densities.df <- readr::read_csv('densities.csv')
 
@@
 \begin{comment}
-
-```{r}
-rot.mat.2 <- function(angle, axis = 1) {
-  # rotation matrix for K=2
-
-  if (axis %in% c(1, 3)) {
-    rotation <- matrix(c(cos(angle), -sin(angle), sin(angle), cos(angle)),
-                       nrow = 2, ncol = 2, byrow = TRUE)
-    if (axis == 1) {
-      return(as.matrix(Matrix::bdiag(1, rotation, -1)))
-    } else {
-      return(as.matrix(Matrix::bdiag(rotation, 1, -1)))
-    }
-  } else if (axis == 2) {
-    return(matrix(c(cos(angle), 0, sin(angle), 0,
-                    0, 1, 0, 0,
-                    -sin(angle), 0, cos(angle), 0,
-                    0, 0, 0, -1),
-                  nrow = 4, ncol = 4,
-                  byrow = TRUE))
-  } else {
-    stop(simpleError('axis must be 1, 2, or 3'))
-  }
-}
-
-hyp.rot.mat.2 <- function(angle, axis = 1) {
-  # hyperbolic rotation matrix for $K=2
-
-  if (axis == 1) {
-    return(matrix(c(1, 0, 0, 0,
-                    0, 1, 0, 0, 
-                    0, 0, cosh(angle), sinh(angle),
-                    0, 0, sinh(angle), cosh(angle)), 
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else if (axis == 2) {
-    return(matrix(c(1, 0, 0, 0,
-                    0, cosh(angle), 0, sinh(angle),
-                    0, 0, 1, 0,
-                    0, sinh(angle), 0, cosh(angle)),
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else if (axis == 3) {
-    return(matrix(c(cosh(angle), 0, 0, sinh(angle),
-                    0, 1, 0, 0, 
-                    0, 0, 1, 0, 
-                    sinh(angle), 0, 0, cosh(angle)),
-                  byrow = TRUE,
-                  nrow = 4, ncol = 4))
-  } else {
-    stop(simpleError('axis must be 1, 2, or 3'))
-  }
-}
-
-construct.Q <- function(angles) {
-  rot.angles <- angles[1:3]
-  hyp.angles <- c(angles[1], angles[4], -angles[4])
-  rot.matrices <- lapply(seq(3), function(i) rot.mat.2(rot.angles[i], i))
-  hyp.matrices <- lapply(seq(3), function(i) hyp.rot.mat.2(hyp.angles[i], i))
-  matrices <- c(rot.matrices, hyp.matrices)
-  return(Reduce(`%*%`, matrices))
-}
-
-obj.fun <- function(angles, Z.hat, clusters) {
-  R <- construct.Q(angles)
-
-  XU.c <- Z.hat %*% R
-  in.prods <- XU.c[clusters == 1, ] %*% t(XU.c[clusters == 2, , drop = FALSE])
-  return(
-    sum(XU.c[clusters == 1, 2] ** 2) + 
-      sum(XU.c[clusters == 2, 1] ** 2) + 
-      sum((XU.c[clusters == 1, 3] - XU.c[clusters == 1, 4]) ** 2) + 
-      sum((XU.c[clusters == 2, 3] + XU.c[clusters == 2, 4]) ** 2)
-  )
-}
-
-transform.embedding <- function(Z.hat, clusters, init.angles) {
-  # init.angles <- runif(4, -pi, pi)
-  # init.angles <- rep(0, 4)
-  angles.1 <- optim(init.angles, function(x) obj.fun(x, Z.hat, clusters), 
-                    method = 'SANN')$par
-  angles.2 <- optim(angles.1, function(x) obj.fun(x, Z.hat, clusters),
-                    method = 'Nelder-Mead',
-                    control = list(abstol = 1e-10, maxit = 1e3))$par
-  R <- construct.Q(angles.2)
-  return(Z.hat %*% R)
-}
-```
 
@@
 ```{r lambda_est, cache = TRUE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
-set.seed(314159)
-
-Ipq <- diag(c(1, 1, 1, -1))
-
-rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-  plyr::ldply(seq(iter), function(i) {
-    z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
-    z <- sort(z)
-
-    n1 <- sum(z == 1)
-    n2 <- sum(z == 2)
-
-    lambda11 <- rbeta(n1, a1, b1)
-    lambda22 <- rbeta(n2, a1, b1)
-    lambda12 <- rbeta(n1, a2, b2)
-    lambda21 <- rbeta(n2, a2, b2)
-
-    X <- cbind(c(lambda11, rep(0, n2)),
-               c(lambda12, rep(0, n2)),
-               c(rep(0, n1), lambda21),
-               c(rep(0, n1), lambda22))
-    Y <- cbind(c(lambda11, rep(0, n2)),
-               c(rep(0, n1), lambda21),
-               c(lambda12, rep(0, n2)),
-               c(rep(0, n1), lambda22))
-    P <- X %*% t(Y)
-    A <- draw.graph(P)
-    lambda.matrix <- cbind(c(lambda11, lambda21),
-                           c(lambda12, lambda22))
-    Zhat <- embedding(A, p, q)
-    Phat <- Zhat %*% Ipq %*% t(Zhat)
-    rmse <- lambda.rmse(lambda.matrix, Phat, z)
-    dplyr::tibble(n = n, rmse = rmse) %>% 
-      return()
-  }, .parallel = TRUE) %>% 
-    return()
-}
+rmse.df <- readr::read_csv('rmse.csv')
 

### 2020-07-22 clustering-sim.R

@@ -1,0 +1,60 @@
+# simulation using proposed clustering method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+library(mclust)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, draw.pabm.beta.2,
+             normalized.laplacian,
+             cluster.pabm,
+             .from = 'functions.R')
+# library(ggplot2)
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation parameters
+K <- 2
+p <- K * (K + 1) / 2
+q <- K * (K - 1) / 2
+alpha <- .5
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+n.vec <- c(64, 128, 256, 512, 1024)
+iter <- 100
+set.seed(314159)
+
+# clustering simulation
+clustering.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
+  plyr::ldply(seq(iter), function(i) {
+    Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
+    A <- Az$A
+    z <- Az$z
+    clustering <- cluster.pabm(A, K)
+    error <- mean(clustering == z)
+    if (error > .5) error <- 1 - error
+    dplyr::tibble(n = n, error = error) %>% 
+      return()
+  }, .parallel = TRUE) %>% 
+    return()
+}
+
+# clustering.df %>% 
+#   dplyr::group_by(n) %>% 
+#   dplyr::summarise(med.err = median(error),
+#                    first.q = quantile(error, .25),
+#                    third.q = quantile(error, .75)) %>% 
+#   ggplot() + 
+#   scale_y_log10() +
+#   scale_x_log10() +
+#   labs(y = 'error') + 
+#   geom_line(aes(x = n, y = med.err)) + 
+#   geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q))
+
+# export as csv
+readr::write_csv(clustering.df, 'clustering.csv')

### 2020-07-22 functions.R

@@ -1,0 +1,208 @@
+embedding <- function(A, p = NULL, q = NULL,
+                      eps = 1e-6) {
+  eigen.A <- eigen(A, symmetric = TRUE)
+  if (is.null(p) | is.null(q)) {
+    keep <- (abs(eigen.A$values) > eps)
+  } else {
+    keep <- c(seq(p), seq(n, n - q + 1))
+  }
+  
+  U <- eigen.A$vectors[, keep]
+  S <- diag(sqrt(abs(eigen.A$values[keep])))
+  return(U %*% S)
+}
+
+draw.graph <- function(P) {
+  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
+  A[lower.tri(A)] <- 0
+  diag(A) <- 0
+  A <- A + t(A)
+  return(A)
+}
+
+I.pq <- function(p, q) {
+  return(diag(c(rep(1, p), rep(-1, q))))
+}
+
+draw.pabm.beta.2 <- function(n, a1, b1, a2, b2, alpha = .5) {
+  z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
+  z <- sort(z)
+  
+  n1 <- sum(z == 1)
+  n2 <- sum(z == 2)
+  
+  lambda11 <- rbeta(n1, a1, b1)
+  lambda22 <- rbeta(n2, a1, b1)
+  lambda12 <- rbeta(n1, a2, b2)
+  lambda21 <- rbeta(n2, a2, b2)
+  
+  X <- cbind(c(lambda11, rep(0, n2)),
+             c(lambda12, rep(0, n2)),
+             c(rep(0, n1), lambda21),
+             c(rep(0, n1), lambda22))
+  Y <- cbind(c(lambda11, rep(0, n2)),
+             c(rep(0, n1), lambda21),
+             c(lambda12, rep(0, n2)),
+             c(rep(0, n1), lambda22))
+  P <- X %*% t(Y)
+  A <- draw.graph(P)
+  
+  return(list(A = A, z = z))
+}
+
+normalized.laplacian <- function(W) {
+  n <- nrow(W)
+  d.neg.sqrt <- colSums(W) ** -.5
+  d.neg.sqrt[is.infinite(d.neg.sqrt)] <- 0
+  D.neg.sqrt <- diag(d.neg.sqrt)
+  I <- diag(n)
[…]

### 2020-07-22 inner-prods-sim.R

@@ -1,0 +1,46 @@
+# simulation using proposed clustering method
+# looking specifically at distributions of between-cluster inner products
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+library(mclust)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, draw.pabm.beta.2,
+             normalized.laplacian,
+             cluster.pabm,
+             .from = 'functions.R')
+# library(ggplot2)
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation parameters
+K <- 2
+p <- K * (K + 1) / 2
+q <- K * (K - 1) / 2
+alpha <- .5
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+n.vec <- c(64, 128, 256, 512, 1024)
+iter <- 100
+set.seed(314159)
+
+# simulation
+densities.df <- plyr::ldply(n.vec, function(n) {
+  Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
+  A <- Az$A
+  z <- Az$z
+  n1 <- sum(z == 1)
+  n2 <- sum(z == 2)
+  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(3), n)]
+  B <- V %*% t(V)
+  inner.prods <- as.numeric(B[(n1+1):n, (n1+1):n])
+  return(dplyr::tibble(n = n, inner.prods = inner.prods))
+}, .parallel = TRUE)
+
+readr::write_csv(densities.df, 'densities.csv')

### 2020-07-22 rmse-sim.R

@@ -1,0 +1,64 @@
+# simulation using proposed parameter estimation method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, I.pq, 
+             estimate.lambda.block, lambda.rmse,
+             .from = 'functions.R')
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation params
+K <- 2
+p <- K * (K + 1) / 2
+q <- K * (K - 1) / 2
+alpha <- .5
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+n.vec <- c(64, 128, 256, 512, 1024)
+iter <- 100
+Ipq <- I.pq(p, q)
+set.seed(314159)
+
+# simulation
+rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  plyr::ldply(seq(iter), function(i) {
+    z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
+    z <- sort(z)
+    
+    n1 <- sum(z == 1)
+    n2 <- sum(z == 2)
+    
+    lambda11 <- rbeta(n1, a1, b1)
+    lambda22 <- rbeta(n2, a1, b1)
+    lambda12 <- rbeta(n1, a2, b2)
+    lambda21 <- rbeta(n2, a2, b2)
+    
+    X <- cbind(c(lambda11, rep(0, n2)),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(rep(0, n1), lambda22))
+    Y <- cbind(c(lambda11, rep(0, n2)),
+               c(rep(0, n1), lambda21),
+               c(lambda12, rep(0, n2)),
+               c(rep(0, n1), lambda22))
+    P <- X %*% t(Y)
+    A <- draw.graph(P)
+    lambda.matrix <- cbind(c(lambda11, lambda21),
+                           c(lambda12, lambda22))
+    Zhat <- embedding(A, p, q)
+    Phat <- Zhat %*% Ipq %*% t(Zhat)
+    rmse <- lambda.rmse(lambda.matrix, Phat, z)
[…]

### 2020-07-22 summary.Rmd
still in the process of moving simulations to separate scripts

@@
 library(ggplot2)
+
+theme_set(theme_bw())
 ```
@@
 
-```{r}
+```{r eval = FALSE}
 set.seed(123456)
@@
 
-```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework. Note that the the points follow the properties outlined in Section 3.2.'}
+```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework.
+Note that the the points follow the properties outlined in Section 3.2.', eval = FALSE}
 perm <- cbind(c(1, 0, 0, 0),
@@
 ```{r ase_of_p, fig.cap = 'The adjacency spectral embedding of the edge probability matrix $P$.
-Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.'}
+Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.', eval = FALSE}
 pairs(Z, asp = 1, col = z)
@@
 ```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1).
-This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE}
+This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE, eval = FALSE}
 XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
@@
 
-```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.'}
+```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.', eval = FALSE}
 pairs(Zhat, col = z, asp = 1)
@@
 
-```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
+```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).', eval = FALSE}
 XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
@@
 
-```{r}
-estimate.lambda.block <- function(P.block, within = FALSE) {
-  if (within) {
-    P.svd <- svd(P.block)
-    u <- P.svd$u[, 1]
-    if (mean(u) < 0) u <- -u
-    s <- sqrt(P.svd$d[1])
-    return(s * u)
-  } else {
-    P.svd <- svd(P.block)
-    u <- P.svd$u[, 1]
-    if (mean(u) < 0) u <- -u
-    v <- P.svd$v[, 1]
-    if (mean(v) < 0) v <- -v
-    s <- sqrt(P.svd$d[1])
-    return(list(s * u, s * v))
-  }
-}
-
-lambda.rmse <- function(lambda.matrix, Phat, clustering) {
-  K <- max(clustering)
-  n <- nrow(Phat)
-  n.vector <- sapply(seq(K), function(k) sum(clustering == k))
-  sapply(seq(K), function(k) {
-    low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
-    high.ind.k <- sum(n.vector[seq(k)])
-    sapply(seq(k), function(l) {
-      P.block <- Phat[clustering == k, clustering == l]
-      if (k == l) {
-        lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
-        lambda.hat <- estimate.lambda.block(P.block, TRUE)
-        return(sum((lambda - lambda.hat) ** 2))
-      } else {
-        low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
-        high.ind.l <- sum(n.vector[seq(l)])
-        lambda.kl <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
-        lambda.lk <- as.numeric(lambda.matrix[seq(low.ind.l, high.ind.l), k])
-        lambda.hats <- estimate.lambda.block(P.block)
-        return(sum((lambda.kl - lambda.hats[[1]]) ** 2) + 
-                 sum((lambda.lk - lambda.hats[[2]]) ** 2))
-      }
-    }) %>% 
-      sum() %>% 
-      return()
-  }) %>% 
-    sum() %>% 
-    magrittr::divide_by(n * K) %>% 
-    sqrt() %>% 
-    return()
-}
-```
-
 Figure \ref{fig:lambda_est} shows the medians and interquartile ranges of root 
@@
 
-```{r lambda_est, cache = TRUE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
+```{r lambda_est, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
 rmse.df <- readr::read_csv('rmse.csv')

### 2020-07-23 clustering-sim.R
comparison simulations

@@ -3,13 +3,13 @@
 
 # packages, functions, etc.
 import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
 library(mclust)
 source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
 setwd('~/dev/pabm-grdpg')
 import::here(embedding, draw.graph, draw.pabm.beta.2,
              normalized.laplacian,
-             cluster.pabm,
+             cluster.pabm, ssc,
              .from = 'functions.R')
 # library(ggplot2)
 
@@ -25,36 +25,48 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
+sparsity <- .01
 n.vec <- c(64, 128, 256, 512, 1024)
 iter <- 100
 set.seed(314159)
 
 # clustering simulation
-clustering.df <- foreach(n = n.vec,.combine = dplyr::bind_rows) %do% {
-  plyr::ldply(seq(iter), function(i) {
+clustering.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  foreach(i = seq(iter), 
+          .combine = dplyr::bind_rows, 
+          .errorhandling = 'remove') %dopar% {
     Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
     A <- Az$A
     z <- Az$z
     clustering <- cluster.pabm(A, K)
     error <- mean(clustering == z)
     if (error > .5) error <- 1 - error
-    dplyr::tibble(n = n, error = error) %>% 
+    clustering.ssc <- ssc(A, K, sparsity)
+    error.ssc <- mean(clustering.ssc == z)
+    if (error.ssc > .5) error.ssc <- 1 - error.ssc
+    dplyr::tibble(n = n, error = error, error.ssc = error.ssc) %>% 
       return()
-  }, .parallel = TRUE) %>% 
+  } %>% 
     return()
 }
 
-# clustering.df %>% 
-#   dplyr::group_by(n) %>% 
+# clustering.df %>%
+#   dplyr::group_by(n) %>%
 #   dplyr::summarise(med.err = median(error),
 #                    first.q = quantile(error, .25),
-#                    third.q = quantile(error, .75)) %>% 
-#   ggplot() + 
-#   scale_y_log10() +
-#   scale_x_log10() +
-#   labs(y = 'error') + 
[…]

### 2020-07-23 functions.R

@@ -206,3 +206,75 @@ lambda.rmse <- function(lambda.matrix, Phat, clustering) {
     sqrt() %>% 
     return()
 }
+
+ssc <- function(A, 
+                K = 2,
+                lambda = .01,
+                parallel = FALSE) {
+  p <- K * (K + 1) / 2
+  q <- K * (K - 1) / 2
+  
+  Y <- t(embedding(A, p, q))
+  
+  n <- nrow(Y)
+  N <- ncol(Y)
+  B <- plyr::aaply(seq(N), 1, function(i) {
+    y <- Y[, i]
+    X <- Y[, -i]
+    betahat <- glmnet::glmnet(X, y, lambda = lambda) %>% 
+      coef() %>% 
+      as.numeric()
+    return(betahat)
+  }, .parallel = parallel) %>% 
+    abs()
+  W <- B + t(B)
+  L <- normalized.laplacian(W)
+  L.eigen <- eigen(L, symmetric = TRUE)
+  X <- diag(colSums(W) ** -.5) %*% L.eigen$vectors[, seq(N - 1, N - K)]
+  clustering <- mclust::Mclust(X, K)$classification
+  return(clustering)
+}
+
+lambda.rmse.mle <- function(lambda.matrix, A, z) {
+  K <- max(z)
+  n <- nrow(A)
+  n.vector <- sapply(seq(K), function(k) sum(z == k))
+  sapply(seq(K), function(k) {
+    low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
+    high.ind.k <- sum(n.vector[seq(k)])
+    n.k <- n.vector[k]
+    e.n.k <- rep(1, n.k)
+    sapply(seq(k), function(l) {
+      if (k == l) {
+        A.kk <- A[z == k, z == l]
+        lambda.hat <- 
+          as.numeric((A.kk %*% e.n.k) / 
+                       as.numeric(sqrt(t(e.n.k) %*% A.kk %*% e.n.k)))
+        lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
+        return(sum((lambda - lambda.hat) ** 2))
+      } else {
+        n.l <- n.vector[l]
+        e.n.l <- rep(1, n.l)
+        low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
+        high.ind.l <- sum(n.vector[seq(l)])
[…]

### 2020-07-23 rmse-sim.R

@@ -6,7 +6,7 @@ import::from(magrittr, `%>%`)
 import::from(foreach, foreach, `%do%`)
 setwd('~/dev/pabm-grdpg')
 import::here(embedding, draw.graph, I.pq, 
-             estimate.lambda.block, lambda.rmse,
+             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
              .from = 'functions.R')
 
 # parallel backend
@@ -55,7 +55,8 @@ rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
     Zhat <- embedding(A, p, q)
     Phat <- Zhat %*% Ipq %*% t(Zhat)
     rmse <- lambda.rmse(lambda.matrix, Phat, z)
-    dplyr::tibble(n = n, rmse = rmse) %>% 
+    rmse.mle <- lambda.rmse.mle(lambda.matrix, A, z)
+    dplyr::tibble(n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
       return()
   }, .parallel = TRUE) %>% 
     return()

### 2020-07-23 summary.Rmd
some comparisons done

@@
 $\lambda^{(kl)}$'s are drawn and for each set of parameters, one adjacency 
-matrix $A$ is drawn and clustered.
+matrix $A$ is drawn and clustered. We will not consider clustering using the 
+edge probability matrix $P$ since this will always result exact recovery of the
+original clusters (up to permutation).
 
-We will not consider clustering using the edge probability matrix $P$ since 
-this will always result exact recovery of the original clusters.
-
-```{r fig.cap = 'IQR clustering error using Algorithm 1 for sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.', cache = FALSE}
+```{r clust_err_sim, fig.cap = 'IQR clustering error using Algorithm 1 (red) compared against subspace clustering (blue) for sample sizes from 64 to 1024.
+Simulations were repeated 100 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', cache = FALSE}
 clustering.df <- readr::read_csv('clustering.csv')
@@
                    first.q = quantile(error, .25),
-                   third.q = quantile(error, .75)) %>%
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75)) %>%
+  dplyr::ungroup() %>% 
   ggplot() +
-  # scale_y_log10() +
-  labs(y = 'error') +
-  geom_line(aes(x = n, y = med.err)) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q))
+  scale_y_log10() +
+  labs(y = 'clustering error', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'Algorithm 1')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'Algorithm 1')) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC')) + 
+  scale_colour_brewer(palette = 'Set1')
 ```
@@
 mean square errors for Algorithm 4 over 100 simulations using the same 
-parameters as before.
+parameters as before. Comparison against an MLE-based method 
+\cite{noroozi2019estimation} suggests similar performance.
 
-```{r lambda_est, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
+```{r lambda_est, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 1024.
+Simulations were repeated 100 times for each sample size.'}
 rmse.df <- readr::read_csv('rmse.csv')
@@
                    q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75)) %>% 
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
   dplyr::ungroup() %>% 
@@
   scale_x_log10() +
-  geom_line(aes(x = n, y = median.rmse)) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse)) + 
-  labs(y = 'RMSE')
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 4')) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
+                   colour = 'MLE-based')) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(y = 'RMSE', colour = NULL)
 ```
 
+# Discussion
+
 # References

### 2020-07-24 summary.Rmd
minor changes

@@
 ---
-title: "Clustering and Parameter Estimation for the Popularity Adjusted Block Model"
+title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation"
 output:
@@
 # output: html_document
+fontsize: 12pt
 # geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"

### 2020-07-29 rmse-sim.R
updated simulation

@@ -21,7 +21,7 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
-n.vec <- c(64, 128, 256, 512, 1024)
+n.vec <- c(64, 128, 256, 512, 1024, 2048)
 iter <- 100
 Ipq <- I.pq(p, q)
 set.seed(314159)
@@ -63,3 +63,4 @@ rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
 }
 
 readr::write_csv(rmse.df, 'rmse.csv')
+

### 2020-07-29 summary.Rmd
still working on proof for theorem 5

@@
 
-$$\max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = O_P\bigg( \frac{(\log n)^c}{n}\bigg)$$
+$$\max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)$$
+where $\lim\limits_{n \to 0} \rho_n = 0$ is a sparsity parameter introduced 
+by \citet{307cbeb9b1be48299388437423d94bf1}. More precisely, 
+$\rho_n o(\log n / \sqrt{n})$.
 
@@
 
-**Theorem 5**. Let $\lambda^{(kl)}_{n_k} \sim F_{n_k}$ for $k, l = 1, ..., K$, 
-and let $\{\hat{\lambda}^{(kl)}_{n_k}\}_K$ be the set of estimates computed by 
-Algorithm 4. 
-Then ... ***TODO***
+**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1},
+
+$$\sqrt{\frac{1}{n K} \sum_{k, l = 1}^K 
+||\hat{\lambda}^{(kl)} - \lambda^{(kl)}||^2} 
+\stackrel{P}{\to} 0$$
+
+*Proof*. ***TODO***
 
@@
 * Mixture parameter $\alpha = .5$
-* $\lambda^{(kk)} \stackrel{indep}{\sim} Beta(2, 1)$
-* $\lambda^{(kl)} \stackrel{indep}{\sim} Beta(1, 2)$ for $k \neq l$
+* $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* $\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
 

### 2020-07-29 clustering-sim.R
updated simulations

@@ -26,7 +26,7 @@ b1 <- 1
 a2 <- 1
 b2 <- 2
 sparsity <- .01
-n.vec <- c(64, 128, 256, 512, 1024)
+n.vec <- c(64, 128, 256, 512, 1024, 2048)
 iter <- 100
 set.seed(314159)
 
@@ -63,9 +63,9 @@ clustering.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
 #   # scale_x_log10() +
 #   labs(y = 'error') +
 #   geom_line(aes(x = n, y = med.err)) +
-#   geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q)) + 
-#   geom_line(aes(x = n, y = med.err.ssc, colour = 'ssc')) + 
-#   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc, 
+#   geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q)) +
+#   geom_line(aes(x = n, y = med.err.ssc, colour = 'ssc')) +
+#   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
 #                     colour = 'ssc'))
 
 # export as csv

### 2020-07-29 summary.Rmd
fixed typo

@@
   scale_y_log10() +
+  # scale_x_log10() + 
   labs(y = 'clustering error', 

### 2020-07-29 functions.R
function for cluster evaluation

@@ -278,3 +278,8 @@ lambda.rmse.mle <- function(lambda.matrix, A, z) {
     sqrt() %>% 
     return()
 }
+
+clust.acc <- function(clust1, clust2) {
+  n <- length(clust1)
+  (1 - fossil::rand.index(clust1, clust2)) * choose(n, 2) / n
+}

### 2020-07-30 summary.Rmd
still working on proof of theorem 5 and additional examples

--- 2020-07-29 summary.Rmd
@@
 by \citet{307cbeb9b1be48299388437423d94bf1}. More precisely, 
-$\rho_n o(\log n / \sqrt{n})$.
+$\rho_n = o(\log n / \sqrt{n})$.
 
@@
 centered at $0$.
+
+This leads to the following algorithm:
 
@@
 \end{algorithm}
+
+Theorem 4 implies that as $n \to \infty$, the number (not just proportion) of 
+misclassified vertices, up to permutation, outputted by algorithm 2 goes to 0.
 
@@
 
-**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1},
+**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1}, the estimators outputted by 
+algorithm 4 converge to the true values as follows:
 
@@
 *Proof*. ***TODO***
+
+The convergence rate of algorithm 4 is the same as that of the methods 
+described by \citet{307cbeb9b1be48299388437423d94bf1}.
 
@@
 
+## Case $K = 2$
+
 For these examples, we will set the following parameters:
@@
 
-## Clustering
+### Clustering
 
 In this part, we will assess Algorithm 2's performance for sample sizes 
-$n = 64, 128, 256, 512, 1024$. For each sample size $n$, $100$ sets of 
+$n = 64, 128, 256, 512, 1024, 2048$. For each sample size $n$, $100$ sets of 
 $\lambda^{(kl)}$'s are drawn and for each set of parameters, one adjacency 
@@
 
-```{r clust_err_sim, fig.cap = 'IQR clustering error using Algorithm 1 (red) compared against subspace clustering (blue) for sample sizes from 64 to 1024.
+```{r clust_err_sim, fig.cap = 'IQR clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048.
 Simulations were repeated 100 times for each sample size.
@@
   scale_y_log10() +
-  labs(y = 'clustering error', 
+  # scale_x_log10() + 
+  labs(y = 'misclustering proportion', 
        colour = NULL) +
   geom_line(aes(x = n, y = med.err,
-                colour = 'Algorithm 1')) +
+                colour = 'Algorithm 2')) +
   geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 1')) + 
+                    colour = 'Algorithm 2')) + 
   geom_line(aes(x = n, y = med.err.ssc,
@@
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC')) + 
+  scale_colour_brewer(palette = 'Set1')
+```
+
+Theorem 4 implies that algorithm 2 will result in not just in the error rate 
+converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
+
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048.
+Simulations were repeated 100 times for each sample size.', cache = FALSE}
+clustering.df <- readr::read_csv('clustering.csv')
+clustering.df %>%
+  dplyr::group_by(n) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75)) %>%
+  dplyr::ungroup() %>% 
+  ggplot() +
+  # scale_y_log10() +
+  # scale_x_log10() + 
+  labs(y = 'misclustering count', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err * n,
+                colour = 'Algorithm 2')) +
+  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
+                    colour = 'Algorithm 2')) + 
+  geom_line(aes(x = n, y = med.err.ssc * n,
+                colour = 'SSC')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                     colour = 'SSC')) + 
@@
 
-## Parameter estimation
-
-\begin{comment}
-
-```{r eval = FALSE}
-set.seed(123456)
-
-n <- 64
-z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
-z <- sort(z)
-
-n1 <- sum(z == 1)
-n2 <- sum(z == 2)
-
-lambda11 <- rbeta(n1, a1, b1)
-lambda22 <- rbeta(n2, a1, b1)
-lambda12 <- rbeta(n1, a2, b2)
-lambda21 <- rbeta(n2, a2, b2)
-
-X <- cbind(c(lambda11, rep(0, n2)),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(rep(0, n1), lambda22))
-Y <- cbind(c(lambda11, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda22))
-P <- X %*% t(Y)
-A <- draw.graph(P)
-Z <- embedding(P)
-Zhat <- embedding(A, 3, 1)
-```
-
-Setting $n = 64$ and generating one example set of $\{\lambda^{(kl)}\}$, we can 
-see the latent positions $XU$ follow the expected properties 
-(Fig. \ref{fig:xu}).
-
-```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework.
-Note that the the points follow the properties outlined in Section 3.2.', eval = FALSE}
-perm <- cbind(c(1, 0, 0, 0),
-              c(0, 0, 1, 0),
-              c(0, 1, 0, 0),
-              c(0, 0, 0, 1))
-U <- eigen(perm)$vectors[, c(3, 1, 2, 4)]
-XU <- X %*% U
-
-pairs(XU, asp = 1, col = z)
-```
-
-If we instead take the ASE of $P = X \Pi X^\top$, we fail to obtain an 
-embedding that has the same properties (Fig. \ref{fig:ase_of_p}).
-
-```{r ase_of_p, fig.cap = 'The adjacency spectral embedding of the edge probability matrix $P$.
-Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.', eval = FALSE}
-pairs(Z, asp = 1, col = z)
-```
-
-Denoting $Z$ as the ASE of $P$, we can recover $ZQ$ that is equivalent to $XU$ 
-(Fig. \ref{fig:q_trans}).
-
-```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1).
-This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE, eval = FALSE}
-XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
-pairs(XU.prime, col = z, asp = 1)
-```
-
-The ASE of the adjacency matrix $A$ (as one realization of $P$) has the 
-resulting pairs plot:
-
-```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.', eval = FALSE}
-pairs(Zhat, col = z, asp = 1)
-```
-
-Minimizing the objective function in (1) results in the following embedding:
-
-```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).', eval = FALSE}
-XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
-pairs(XU.hat, col = z, asp = 1)
-
-# X.hat <- XU.hat %*% t(U)
-# plot(abs(X.hat) > .25)
-```
-
-\end{comment}
+### Parameter estimation
 
@@
 parameters as before. Comparison against an MLE-based method 
-\cite{noroozi2019estimation} suggests similar performance.
+\cite{noroozi2019estimation} suggests similar performance, close to the rate 
+$n^{-1/2}$.
 
-```{r lambda_est, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 1024.
+```{r lambda_est, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 2048.
 Simulations were repeated 100 times for each sample size.'}
@@
 
+## Higher values of $K$
+
+**TODO**
+
 # Discussion

### 2020-08-05 summary.Rmd
minor update

@@
 ```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048.
-Simulations were repeated 100 times for each sample size.', cache = FALSE}
+Simulations were repeated 100 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', cache = FALSE}
 clustering.df <- readr::read_csv('clustering.csv')
@@
   ggplot() +
-  # scale_y_log10() +
+  scale_y_log10() +
   # scale_x_log10() + 
@@
 
-**TODO**
+***TODO***
+
+# Real data examples
+
+***TODO***
 

### 2020-08-06 functions.R
minor update

@@ -175,27 +175,55 @@ estimate.lambda.block <- function(P.block, within = FALSE) {
   }
 }
 
-lambda.rmse <- function(lambda.matrix, Phat, clustering) {
+# lambda.rmse <- function(lambda.matrix, Phat, clustering) {
+#   K <- max(clustering)
+#   n <- nrow(Phat)
+#   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
+#   sapply(seq(K), function(k) {
+#     low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
+#     high.ind.k <- sum(n.vector[seq(k)])
+#     sapply(seq(k), function(l) {
+#       P.block <- Phat[clustering == k, clustering == l]
+#       if (k == l) {
+#         lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
+#         lambda.hat <- estimate.lambda.block(P.block, TRUE)
+#         return(sum((lambda - lambda.hat) ** 2))
+#       } else {
+#         low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
+#         high.ind.l <- sum(n.vector[seq(l)])
+#         lambda.kl <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
+#         lambda.lk <- as.numeric(lambda.matrix[seq(low.ind.l, high.ind.l), k])
+#         lambda.hats <- estimate.lambda.block(P.block)
+#         return(sum((lambda.kl - lambda.hats[[1]]) ** 2) + 
+#                  sum((lambda.lk - lambda.hats[[2]]) ** 2))
+#       }
+#     }) %>% 
+#       sum() %>% 
+#       return()
+#   }) %>% 
+#     sum() %>% 
+#     magrittr::divide_by(n * K) %>% 
+#     sqrt() %>% 
+#     return()
+# }
+
+lambda.rmse <- function(P, Phat, clustering) {
   K <- max(clustering)
   n <- nrow(Phat)
   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
   sapply(seq(K), function(k) {
-    low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
-    high.ind.k <- sum(n.vector[seq(k)])
[…]

### 2020-08-06 rmse-sim.R

@@ -50,12 +50,12 @@ rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
                c(rep(0, n1), lambda22))
     P <- X %*% t(Y)
     A <- draw.graph(P)
-    lambda.matrix <- cbind(c(lambda11, lambda21),
-                           c(lambda12, lambda22))
+    # lambda.matrix <- cbind(c(lambda11, lambda21),
+    #                        c(lambda12, lambda22))
     Zhat <- embedding(A, p, q)
     Phat <- Zhat %*% Ipq %*% t(Zhat)
-    rmse <- lambda.rmse(lambda.matrix, Phat, z)
-    rmse.mle <- lambda.rmse.mle(lambda.matrix, A, z)
+    rmse <- lambda.rmse(P, Phat, z)
+    rmse.mle <- lambda.rmse.mle(P, A, z)
     dplyr::tibble(n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
       return()
   }, .parallel = TRUE) %>% 

### 2020-08-06 functions.R
minor updates

@@ -335,6 +335,10 @@ lambda.rmse.mle <- function(P, A, z) {
         lambda.hat.lk <- 
           as.numeric((t(A.kl) %*% e.n.k) / 
                        as.numeric(sqrt(t(e.n.k) %*% A.kl %*% e.n.l)))
+        lambda.hat.kl <- dplyr::if_else(is.nan(lambda.hat.kl),
+                                        0, lambda.hat.kl)
+        lambda.hat.lk <- dplyr::if_else(is.nan(lambda.hat.lk), 
+                                        0, lambda.hat.lk)
         return(sum((lambdas[[1]] - lambda.hat.kl) ** 2) + 
                  sum((lambdas[[2]] - lambda.hat.lk) ** 2))
       }

### 2020-08-06 rmse-sim-k.R

@@ -1,0 +1,71 @@
+# simulation using proposed parameter estimation method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packgaes, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, I.pq, generate.P.beta,
+             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
+             .from = 'functions.R')
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation params
+K.vec <- c(4, 8)
+alpha <- .5
+a1 <- b2 <- 2
+a2 <- b1 <- 1
+n.vec <- c(64, 128, 256, 512, 1024, 2048)
+iter <- 100
+set.seed(314159)
+
+# simulation
+rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  p <- K * (K + 1) / 2
+  q <- K * (K - 1) / 2
+  Ipq <- I.pq(p, q)
+  
+  foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    plyr::ldply(seq(iter), function(i) {
+      P.z <- generate.P.beta(n, K)
+      P <- P.z$P
+      z <- P.z$clustering
+      A <- draw.graph(P)
+      Zhat <- embedding(A, p, q)
+      Phat <- Zhat %*% Ipq %*% t(Zhat)
+      rmse <- lambda.rmse(P, Phat, z)
+      rmse.mle <- lambda.rmse.mle(P, A, z)
+      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
+        return()
+    }, .parallel = TRUE) %>% 
+      return()
+  } %>%
+    return()
+}
+
+# rmse.df %>% 
+#   na.omit() %>% 
+#   dplyr::group_by(K, n) %>% 
+#   dplyr::summarise(median.rmse = median(rmse),
+#                    q1.rmse = quantile(rmse, .25),
+#                    q3.rmse = quantile(rmse, .75),
+#                    median.rmse.mle = median(rmse.mle),
+#                    q1.rmse.mle = quantile(rmse.mle, .25),
+#                    q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
[…]

### 2020-08-06 clustering-sim-k.R
minor updates

@@ -1,0 +1,74 @@
+# simulation using proposed clustering method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, 
+             generate.P.beta,
+             normalized.laplacian,
+             cluster.pabm, ssc,
+             .from = 'functions.R')
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation parameters
+K.vec <- c(2, 4, 8)
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- .01
+n.vec <- c(64, 128, 256, 512, 1024, 2048)
+iter <- 100
+set.seed(314159)
+
+# clustering simulation
+clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('K =', K, ', n =', n))
+    foreach(i = seq(iter), 
+            .combine = dplyr::bind_rows, 
+            .errorhandling = 'remove') %dopar% {
+      Pz <- generate.P.beta(n, K, a1, a2, b1, b2)
+      P <- Pz$P
+      z <- Pz$clustering
+      A <- draw.graph(P)
+      clustering <- cluster.pabm(A, K)
+      error <- 1 - fossil::adj.rand.index(z, clustering)
+      clustering.ssc <- ssc(A, K, sparsity)
+      error.ssc <- 1 - fossil::adj.rand.index(z, clustering.ssc)
+      dplyr::tibble(K = K, n = n, error = error, 
+                    error.ssc = error.ssc) %>% 
+        return()
+    } %>% 
+      return()
+  } %>%
+    return()
+}
+
+clustering.df %>%
+  dplyr::group_by(K, n) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
[…]

### 2020-08-10 summary.Rmd
minor updates

@@
 
-For these examples, we will set the following parameters:
+For these examples, we set the following parameters:
 
-* $K = 2$
+* Number of underlying clusters $K = 2$
 * Mixture parameter $\alpha = .5$
-* $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
-* $\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
+* Within-cluster popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* Between-cluster popularities 
+$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
 
@@
 
-***TODO***
+For these examples, we set the following parameters:
+
+* Number of underlying clusters $K = 2, 4, 8$
+* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$
+* Within-cluster popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* Between-cluster popularities 
+$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
+
+### Clustering
+
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+Simulations were repeated 100 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', fig.width = 8}
+clustering.df <- readr::read_csv('clustering-k.csv')
+clustering.df %>%
+  dplyr::group_by(K, n) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75)) %>%
+  dplyr::ungroup() %>% 
+  ggplot() +
+  scale_y_log10() +
+  # scale_x_log10() + 
+  labs(y = '1 - adjusted rand index', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'Algorithm 2')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'Algorithm 2')) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC')) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both')
+```
+
+### Parameter estimation
+
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+Simulations were repeated 100 times for each sample size.', fig.width = 8}
+rmse.df <- readr::read_csv('rmse.csv') %>% 
+  dplyr::transmute(K = 2,
+                   n = n,
+                   rmse = rmse,
+                   rmse.mle = rmse.mle) %>% 
+  dplyr::bind_rows(readr::read_csv('rmse-k.csv'))
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() + 
+  scale_y_log10() + 
+  scale_x_log10() +
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 4')) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
+                   colour = 'MLE-based')) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(y = 'RMSE', colour = NULL) + 
+  facet_wrap(~ K, labeller = 'label_both')
+```
 

### 2020-08-19 summary.Rmd
updated with real data example

@@
 
-One popular probabilistic model for graphs with underlying communities is the 
+Statistical analysis on graphs and networks often involve the partitioning of a 
+graph into clusters. This is often motivated by the belief that there is an 
+underlying and unobserved community or cluster label for each vertex of the 
+graph, and edges between pairs of vertices are determined by drawing from a 
+probability distribution based on the communities assigned to each pair. 
+One such probabilistic model is the 
 Stochastic Block Model (SBM), which assumes that the edge probability from 
-one cluster to another follows a Beronulli distribution with a fixed 
-probability for each pair of communities. 
-The Popularity Adjusted Block Model (PABM) was introduced by 
+one vertex to another follows a Beronulli distribution with fixed probabilities 
+for each pair of community labels. 
+The Popularity Adjusted Block Model (PABM) was then introduced by 
 \citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
 address the heterogeneity of edge probabilities within and between communities 
-or clusters while still maintaining the community structure.
+while still maintaining the community structure.
 
-The Random Dot Product Graph (RDPG) \cite{athreya2017statistical} is another 
-probabilistic model with Beronulli edge probabilities. Under this model, each 
+The Random Dot Product Graph (RDPG) model \cite{athreya2017statistical} is 
+another graph model with Beronulli edge probabilities. Under this model, each 
 vertex of the graph can be represented by a point in latent space such that 
@@
 PABM not under the RDPG model but under the *Generalized* Random Dot Product 
-Graph model.
+Graph (GRDPG) model.
 
-## Previous work
-
-\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
-to identify the cluster memberships given either an edge probability matrix $P$ 
-or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
-memberships can be identified exactly (up to permutation). A similar procedure 
-can be applied if $P$ is unknown and we have an observation $A$, but the 
-theoretical guarantees of this method applied to the PABM are unknown. In 
-particular, the method requires spherical Gaussian noise. The authors of this 
-paper then use point estimators for $\{\lambda^{(kl)}\}$ using cluster labels 
-obtained via SSC.
+In this paper, we will only consider undirected graphs, that is the edge value 
+from vertices $i$ and $j$ is equal to the edge in the opposite direction, from 
+vertices $j$ and $i$. Furthermore, we will only consider unweighted graphs with 
+binary edge values. We will also assume that graphs are hollow, i.e., there 
+are no self-loops.
 
@@
 **Definition 1**. 
-Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
-affinity matrix $A \in \{0, 1\}^{n \times n}$. Then $A$ is a random matrix 
-with corresponding edge probability matrix $P$ such that 
+Let $P \in [0, 1]^{n \times n}$ be an edge probability matrix for a set of $n$ 
+vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
+columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
+$P^{(kl})$ describes the edge probabilities between vertices in communities 
+$k$ and $l$. 
+Let graph $G = (V, E)$ be an undirected, unweighted graph such 
+that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
+realization of $Bernoulli(P)$, i.e., 
 $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ 
-($A_{ji} = A_{ij}$ and $A_{ii} = 0$). Let there exist $K$ underlying 
-communities in $G$, and let $n_{k}$ be the size of the $k$^th^ community in 
-$G$ such that $\sum_{k=1}^K n_k = n$.
+($A_{ij} = A_{ji}$ and $A_{ii} = 0$).
 
-If $A$ and $P$ are organized such that $n_k \times n_l$ blocks $A^{(kl)}$ and 
-$P^{(kl)}$ describe the edges and edge probabilities between communities $k$ 
-and $l$, then $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$ for a set of 
-fixed vectors $\{\lambda^{(st)}\}_{s, t = 1, ..., K}$. Each $\lambda^{(st)}$ 
-for $s, t = 1, ..., K$ is a column vector of length $n_s$ (i.e., the community 
-corresponding to the first index provides the vector length).
+If each block $P^{(kl)}$ can be written as the outer product of two vectors 
+$\lambda^{(kl)} (\lambda^{(lk)})^\top$ for a set of fixed vectors 
+$\{\lambda^{(st)}\}_{s, t = 1}^K$ where each $\lambda^{(st)}$ is a column vector 
+of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
+is a realization of a popularity adjusted block model with parameters 
+$\{\lambda^{(st)}\}_{s, t = 1}^K$.
 
 We will use the notation $A \sim PABM(\{\lambda^{(kl)}\}_K)$ to denote a random 
-affinity matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
+adjacency matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
 consisting of $K$ underlying clusters/communities.
@@
 graph $G$. $X$ consists of row vectors $x_i^\top$. Let 
-$A \in \{0, 1\}^{n \times n}$ be the corresponding affinity matrix.
+$A \in \{0, 1\}^{n \times n}$ be the corresponding adjacency matrix.
 
@@
 Then $G = (V, E)$ is a generalized random dot product graph with signature 
-$(p, q)$ and patent positions $X$ iff its random affinity matrix can be 
+$(p, q)$ and patent positions $X$ iff its random adjacency matrix can be 
 described as $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ where 
@@
 
-We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random affinity 
+We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random adjacency 
 matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.
@@
 does not require estimation.
+
+## Previous work
+
+\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
+to identify the cluster memberships given either an edge probability matrix $P$ 
+or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
+memberships can be identified exactly (up to permutation). A similar procedure 
+can be applied if $P$ is unknown and we have an observation $A$, but the 
+theoretical guarantees of this method applied to the PABM are unknown. In 
+particular, the method requires spherical Gaussian noise. The authors of this 
+paper then use point estimators for $\{\lambda^{(kl)}\}$ using cluster labels 
+obtained via SSC.
 
@@
 
-# Real data examples
+# Real data example
 
-***TODO***
+```{r}
+library(mclust)
+import::here(cluster.pabm, .from = '~/dev/pabm-grdpg/functions.R')
+
+n.edges <- 20566
+
+butterfly <- R.matlab::readMat('~/Downloads/Raw_butterfly_network.mat')
+
+A <- butterfly$W.butterfly0
+z <- butterfly$labels
+
+keep <- c(6, 2, 9, 4)
+A <- A[z %in% keep, z %in% keep]
+z <- z[z %in% keep]
+n <- length(z)
+K <- 4
+
+delta <- quantile(A[upper.tri(A)], 1 - n.edges / (n * (n - 1) / 2))
+A <- (A > delta) * 1
+
+z.hat <- cluster.pabm(A, K)
+A.star <- A[order(z.hat), order(z.hat)]
+```
+
+For the real data example, we applied Algorithm 2 to the Leeds Butterfly 
+dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
+butterflies across 10 species. The graph was modified to match the example 
+from \citet{noroozi2019estimation}: Only the 4 most frequent species were 
+considered, and the similarities were discretized to $\{0, 1}\$ via 
+thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
+by the resultant the clustering.
+
+Comparing against the ground truth species labels, Algorithm 2 achieves an 
+accuracy of 63\% and an adjusted Rand index of 79\%. In comparison, 
+\citet{noroozi2019estimation} achieved an adjusted Rand index of 73\%.
+
+```{r butterfly, fig.height = 5, fig.width = 4, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.'}
+library(plot.matrix)
+plot(A.star, border = NA, key = NULL, 
+     xlab = '', ylab = '', main = '',
+     axis.row = NULL, axis.col = NULL)
+```
 

### 2020-08-19 functions.R

@@ -175,38 +175,6 @@ estimate.lambda.block <- function(P.block, within = FALSE) {
   }
 }
 
-# lambda.rmse <- function(lambda.matrix, Phat, clustering) {
-#   K <- max(clustering)
-#   n <- nrow(Phat)
-#   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
-#   sapply(seq(K), function(k) {
-#     low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
-#     high.ind.k <- sum(n.vector[seq(k)])
-#     sapply(seq(k), function(l) {
-#       P.block <- Phat[clustering == k, clustering == l]
-#       if (k == l) {
-#         lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
-#         lambda.hat <- estimate.lambda.block(P.block, TRUE)
-#         return(sum((lambda - lambda.hat) ** 2))
-#       } else {
-#         low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
-#         high.ind.l <- sum(n.vector[seq(l)])
-#         lambda.kl <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
-#         lambda.lk <- as.numeric(lambda.matrix[seq(low.ind.l, high.ind.l), k])
-#         lambda.hats <- estimate.lambda.block(P.block)
-#         return(sum((lambda.kl - lambda.hats[[1]]) ** 2) + 
-#                  sum((lambda.lk - lambda.hats[[2]]) ** 2))
-#       }
-#     }) %>% 
-#       sum() %>% 
-#       return()
-#   }) %>% 
-#     sum() %>% 
-#     magrittr::divide_by(n * K) %>% 
-#     sqrt() %>% 
-#     return()
-# }
-
 lambda.rmse <- function(P, Phat, clustering) {
   K <- max(clustering)
   n <- nrow(Phat)
@@ -262,50 +230,6 @@ ssc <- function(A,
   clustering <- mclust::Mclust(X, K)$classification
   return(clustering)
 }
-
-# lambda.rmse.mle <- function(lambda.matrix, A, z) {
-#   K <- max(z)
-#   n <- nrow(A)
-#   n.vector <- sapply(seq(K), function(k) sum(z == k))
[…]

### 2020-08-19 summary.Rmd
fixed typo

@@
 A.star <- A[order(z.hat), order(z.hat)]
+
+ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
 ```
@@
 from \citet{noroozi2019estimation}: Only the 4 most frequent species were 
-considered, and the similarities were discretized to $\{0, 1}\$ via 
+considered, and the similarities were discretized to $\{0, 1\}$ via 
 thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
@@
 Comparing against the ground truth species labels, Algorithm 2 achieves an 
-accuracy of 63\% and an adjusted Rand index of 79\%. In comparison, 
-\citet{noroozi2019estimation} achieved an adjusted Rand index of 73\%.
+accuracy of 63\% and an adjusted Rand index of `r ari`\%. In comparison, 
+\citet{noroozi2019estimation} achieved an adjusted Rand index of 73\% using 
+sparse subspace clustering.
 
-```{r butterfly, fig.height = 5, fig.width = 4, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.'}
+```{r butterfly, fig.height = 3, fig.width = 3, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
 library(plot.matrix)
@@
      xlab = '', ylab = '', main = '',
-     axis.row = NULL, axis.col = NULL)
+     axis.row = NULL, axis.col = NULL, asp = 1)
 ```

### 2020-08-19 summary.Rmd
fixed typo

@@
 title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation"
+author: John Koo, Minh Tang, Michael Trosset
 output:
@@
 
-**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1}, the estimators outputted by 
-algorithm 4 converge to the true values as follows:
+**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1}, the root mean square errors of the 
+estimators outputted by algorithm 4 converge to the true values as follows:
 
@@
 
-```{r butterfly, fig.height = 3, fig.width = 3, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
+```{r butterfly, fig.height = 4, fig.width = 4, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
 library(plot.matrix)

### 2020-08-26 summary.Rmd
still working on additional example

@@
 library(ggplot2)
+library(plot.matrix)
 
@@
 
-*Proof*
+*Proof*. 
 Let $\Lambda^{(k)} = 
@@
 
-**Theorem 5**. Under assumptions 1-6 as laid out by \citet{307cbeb9b1be48299388437423d94bf1}, the root mean square errors of the 
-estimators outputted by algorithm 4 converge to the true values as follows:
+**Theorem 5**. Under regularity and sparsity assumptions, and under the further 
+assumption that $K$ is fixed,
 
-$$\sqrt{\frac{1}{n K} \sum_{k, l = 1}^K 
-||\hat{\lambda}^{(kl)} - \lambda^{(kl)}||^2} 
-\stackrel{P}{\to} 0$$
+$$\max_{k, l \in \{1, ..., K\}} 
+||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
+O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)$$
 
-*Proof*. ***TODO***
+*Proof (sketch)*.  
+Let $P$ and $A$ be organized by cluster such that the elements of blocks 
+$P^{(kl)}$ and $A^{(kl)}$ correspond to edges between clusters $k$ and $l$. 
+By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. This implies 
+that $P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ has one singular value 
+and one pair of singular vectors. By convention, we can let 
+$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. Let 
+$\hat{U}^{(kl)} \hat{\Sigma}^{(kl)} (\hat{V}^{kl})^\top$ be the singular value 
+decomposition of $A^{(kl)}$, and let 
+$\hat{\sigma}_{kl} \hat{u}^{(kl)} (\hat{v}^{(kl)})^\top$ be its one-dimensional 
+approximation. Then 
+$\hat{\lambda}^{(kl)} = \hat{\sigma}_{kl} \hat{u}^{(kl)}$.[^alg4] Thus 
+$\hat{\lambda}^{(kl)}$ is a one-dimensional adjacency embedding approximation 
+of $\lambda^{(kl)}$.  
+Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
+embedding $\hat{\lambda}^{(kl)}$ approximates $\lambda^{(kl)}$ at rate 
+$\frac{(\log n_k)^c}{n_k^{1/2}}$.
 
-The convergence rate of algorithm 4 is the same as that of the methods 
-described by \citet{307cbeb9b1be48299388437423d94bf1}.
-
-\begin{comment}
-
-### Using Adjacency Spectral Embedding
-
-For this section, we will focus on the case $K = 2$. Under this condition, the 
-PABM is equivalent to the GRDPG with signature $(3, 1)$.
-
-The adjacency spectral embeddings of both $A$ and $P$ are not unique. In 
-particular, let $Z$ be the ASE of $P$. Then $P = Z I_{3, 1} Z^\top$. However, 
-for any $Q \in \mathbb{O}(3, 1)$, 
-$(ZQ) I_{3, 1} (ZQ)^\top = Z (Q I_{3, 1} Q^\top) Z^\top = Z I_{3, 1} Z^\top = P$,
-so $ZQ$ is also a valid ASE Of $P$. If we can find $Q \in \mathbb{O}(3, 1)$ 
-such that $ZQ = XU$, we can compute the parameters $\{\lambda^{(kl)}\}$ 
-directly. Furthermore, if we instead use the ASE of the adjacency matrix $A$, 
-$\exists Q \in \mathbb{O}(3, 1)$ such that $\max\limits_i ||Q \hat{z}_i - XU||$ 
-is minimized (and goes to zero under a probabilistic model for
-$\{\lambda^{(kl)}\}$'s).
-
-Thus if we can identify $Q$ such that the $ZQ = XU$ in the case of embedding 
-$P$ or $\hat{Z} Q - XU$ is minimized in the case of embedding $A$, we can 
-identify or estimate the PABM parameters directly from the embedding. $XU$ is 
-unknown, but we can still obtain an embedding $ZQ$ that follows the properties 
-of $XU$, which will yield estimates $\{\hat{\lambda}^{(kl)}\}$ that are valid 
-in that they will produce the edge probability matrix $P$.[^lambdas]
-
-[^lambdas]: Note that the set $\{\lambda^{(kl)}\}_K$ that produces a unique 
-$P$ is not unique \citep{noroozi2019estimation}.
-
-In particular, we can note the following properties of $XU$:
-
-1. If row $i$ is in cluster 1, then its second index is 0. If it is in cluster 
-2, its first index is 0.
-2. If row $i$ is in cluster 1, then its third and fourth indices are equal. If 
-it is in cluster 2, the fourth index is the negative of its third index.
-3. If rows $i$ and $j$ are in different clusters, their dot product is 0.
-
-If we have cluster memberships (either known *a priori* or estimated using a 
-clustering method), then we can estimate $XU$ by picking
-$Q \in \mathbb{O}(3, 1)$ such that $ZQ$ best fits these properties.
-
-$\mathbb{O}(3, 1)$ happens to be the Lorentz group, and each 
-$Q \in \mathbb{O}(3, 1)$ can be represented as the product of six matrices that 
-depend on a total of four parameters \cite{10.1088/978-1-6817-4254-0}. This 
-lends itself as an optimization problem:
-
-\begin{equation}
-\begin{split}
-  \min_{\theta, \phi, \psi, \tau} & ||\xi^{(1, 2)}||^2 + ||\xi^{(2, 1)}||^2 + 
-  ||\xi^{(1, 3)} - \xi^{(1, 4)}||^2 + ||\xi^{(2, 3)} + \xi^{(2, 4)}||^2 \\
-  \text{s.t. } & \Xi = Z Q(\theta, \phi, \psi, \tau) \\
-  & \xi^{(i, j)} \text{ is the } j^{th} \text{ column of } \Xi \text{ with rows from the } i^{th} \text{ cluster}
-\end{split}
-\end{equation}
-
-Note that if $Q$ can be properly estimated, the asymptotic results from 
-\citet{rubindelanchy2017statistical} can be applied here.
-
-\end{comment}
+[^alg4]: Note that this estimator is similar to but slightly worse than the 
+estimator from Algorithm 4.
 
@@
 
-# Real data example
+# Real data examples
 
@@
 \citet{noroozi2019estimation} achieved an adjusted Rand index of 73\% using 
-sparse subspace clustering.
+sparse subspace clustering on the same dataset.
 
-```{r butterfly, fig.height = 4, fig.width = 4, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
-library(plot.matrix)
-plot(A.star, border = NA, key = NULL, 
+```{r}
+plot.A <- function(A, margins = c(0, 0, 0, 0)) {
+  default.mar <- par()$mar
+  par(mar = margins)
+  plot(A, border = NA, key = NULL, 
        xlab = '', ylab = '', main = '',
        axis.row = NULL, axis.col = NULL, asp = 1)
+  par(mar = default.mar)
+}
+```
+
+```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
+plot.A(A.star)
 ```

### 2020-09-02 summary.Rmd
updated theorem 5 and added second example

@@
 - \usepackage{natbib}
-- \usepackage[linesnumbered,vlined,ruled]{algorithm2e} 
+- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
 - \setcitestyle{numbers,square}
@@
 \newcommand{\diag}{\text{diag}}
+\newcommand{\tr}{\text{Tr}}
 
@@
 
-Statistical analysis on graphs and networks often involve the partitioning of a 
-graph into clusters. This is often motivated by the belief that there is an 
-underlying and unobserved community or cluster label for each vertex of the 
-graph, and edges between pairs of vertices are determined by drawing from a 
-probability distribution based on the communities assigned to each pair. 
-One such probabilistic model is the 
-Stochastic Block Model (SBM), which assumes that the edge probability from 
+Statistical analysis on graphs or networks often involves the partitioning of 
+a graph into disconnected subgraphs or clusters. This is often motivated by the 
+belief that there exist underlying and unobserved communities to which each 
+vertex of the graph belongs, and edges between pairs of vertices are determined 
+by drawing from a probability distribution based on the community relationships 
+between each pair. The goal of this analysis then is population community 
+detection, or the recovery of the true underlying community labels for each 
+vertex, up to permutation (with some additional parameter estimation being of 
+possible interest), assuming some probabilistic graph model. One such model is 
+the Stochastic Block Model (SBM), which assumes that the edge probability from 
 one vertex to another follows a Beronulli distribution with fixed probabilities 
-for each pair of community labels. 
-The Popularity Adjusted Block Model (PABM) was then introduced by 
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
-address the heterogeneity of edge probabilities within and between communities 
-while still maintaining the community structure.
+for each pair of community labels. The Popularity Adjusted Block Model (PABM) 
+was then introduced by \citet*{307cbeb9b1be48299388437423d94bf1} as a 
+generalization of the SBM to address the heterogeneity of edge probabilities 
+within and between communities while still maintaining the community structure.
 
@@
 another graph model with Beronulli edge probabilities. Under this model, each 
-vertex of the graph can be represented by a point in latent space such that 
-the edge probability between any pair of vertices is given by their 
+vertex of the graph can be represented by a point in a latent Euclidean space 
+such that the edge probability between any pair of vertices is given by their 
 corresponding dot product in the latent space. The SBM is equivalent to a 
-special case of the RDPG model in which all vertices in a given community share 
+special case of the RDPG model in which all vertices of a given community share 
 the same position in the latent space. It has also been shown that similar 
-probabilistic graph models, (such as the Mixed Membership Stochastic Block Model
-\cite{rubindelanchy2017statistical}). An analogous property exists for the 
-PABM not under the RDPG model but under the *Generalized* Random Dot Product 
-Graph (GRDPG) model.
+probabilistic graph models, such as the Mixed Membership Stochastic Block 
+Model, can be represented in this way \cite{rubindelanchy2017statistical}. An 
+analogous property exists for the PABM, not under the RDPG model but under the *Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
+explored in this paper and exploited to construct algorithms for community 
+detection and parameter estimation for the PABM.
 
-In this paper, we will only consider undirected graphs, that is the edge value 
-from vertices $i$ and $j$ is equal to the edge in the opposite direction, from 
-vertices $j$ and $i$. Furthermore, we will only consider unweighted graphs with 
-binary edge values. We will also assume that graphs are hollow, i.e., there 
-are no self-loops.
+In this paper, we will only consider undirected graphs, that is the edge weight 
+from vertex $i$ to vertex $j$ is equal to the edge weight in the opposite 
+direction, from vertex $j$ to vertex $i$. Furthermore, we will only consider 
+unweighted graphs with binary ($0, 1$) edge weights We will also assume that 
+graphs are hollow, i.e., there are no edges from a vertex to itself.
 
@@
 
-## The popularity adjusted block model (PABM) \cite{noroozi2019estimation}
+## The popularity adjusted block model (PABM) \cite{noroozi2019estimation} and the generalized random dot product graph \cite{rubindelanchy2017statistical}
 
 **Definition 1**. 
-Let $P \in [0, 1]^{n \times n}$ be an edge probability matrix for a set of $n$ 
+Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
+set of $n$ 
 vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
@@
 
-## The generalized random dot product graph (GRDPG) \cite{rubindelanchy2017statistical}
-
 **Definition 2**.
-Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
-graph $G$. $X$ consists of row vectors $x_i^\top$. Let 
-$A \in \{0, 1\}^{n \times n}$ be the corresponding adjacency matrix.
-
-Fix $p$, $q$ such that $p + q = d$ and define 
-$I_{pq} = \begin{bmatrix} I_p & 0 \\ 0 & -I_q \end{bmatrix}$.
-
-Then $G = (V, E)$ is a generalized random dot product graph with signature 
-$(p, q)$ and patent positions $X$ iff its random adjacency matrix can be 
-described as $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ where 
-$P_{ij} = x_i^\top I_{pq} x_j$.
+Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
+set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that 
+$P = X I_{pq} X^\top$ for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
+graph $G = (V, E)$ with adjacency matrix $A$ such that $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and $A_{ii} = 0$) is a draw from the generalized random dot product graph model with latent positions $X$ and signature $(p, q)$.
+More precisely, if vertices $i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains the latent positions as rows $x_i^\top$.
 
@@
 
+**Definition 3**. The indefinite orthogonal group with signature $(p, q)$ is 
+the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
+denoted as $\mathbb{O}(p, q)$ \cite{rubindelanchy2017statistical}.
+
 **Remark**. 
-Like the RDPG, the latent positions of a GRDPG are not unique. More 
-specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
-$Q \in \{R : R^\top I_{pq} R = I_{pq} \}$, 
+Like the RDPG, the latent positions of a GRDPG are not unique 
+\cite{rubindelanchy2017statistical}. 
+More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
+$Q \in \mathbb{O}(p, q)$, 
 $(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
-x_i^\top I_{pq} x_j = P_{ij}$ (this is the indefinite orthogonal group, 
-$\mathbb{O}(p, q)$).
+x_i^\top I_{pq} x_j = P_{ij}$. 
 Unlike in the RDPG case, transforming the latent positions by multiplication 
@@
 
-* Positions 1, 5, 9 are fixed
-* The cycles of order 2 are
-    * $(2, 4)$
-    * $(3, 7)$
-    * $(6, 8)$
+* Positions 1, 5, 9 are fixed.
+* The cycles of order 2 are $(2, 4)$, $(3, 7)$, and $(6, 8)$.
 
@@
   \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
-  Assign $p \leftarrow K (K + 1) / 2$ and $q \leftarrow K (K - 1) / 2$.\;
-  Compute $\hat{Z}$, the adjacency spectral embedding of $A$ using signature
-  $(p, q)$.\;
-  Compute $\hat{P} = \hat{Z} I_{pq} \hat{Z}^\top$, the estimate of the edge 
-  probability matrix based on the adjacency spectral embedding.\;
-  Arrange the rows and columns of $\hat{P}$ by cluster such that each 
-  $\hat{P}^{(kl)}$ block consists of estimated edge probabilities between 
+  Arrange the rows and columns of $A$ by cluster such that each 
+  $A^{(kl)}$ block consists of estimated edge probabilities between 
   clusters $k$ and $l$.\;
   \For{$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $\hat{P}^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
     block.\;
@@
 **Theorem 5**. Under regularity and sparsity assumptions, and under the further 
-assumption that $K$ is fixed,
+assumption that $K$ is fixed and community labels are known,
 
@@
 *Proof (sketch)*. 
-Let $P$ and $A$ be organized by cluster such that the elements of blocks 
-$P^{(kl)}$ and $A^{(kl)}$ correspond to edges between clusters $k$ and $l$. 
-By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. This implies 
-that $P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ has one singular value 
-and one pair of singular vectors. By convention, we can let 
-$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. Let 
-$\hat{U}^{(kl)} \hat{\Sigma}^{(kl)} (\hat{V}^{kl})^\top$ be the singular value 
-decomposition of $A^{(kl)}$, and let 
-$\hat{\sigma}_{kl} \hat{u}^{(kl)} (\hat{v}^{(kl)})^\top$ be its one-dimensional 
-approximation. Then 
-$\hat{\lambda}^{(kl)} = \hat{\sigma}_{kl} \hat{u}^{(kl)}$.[^alg4] Thus 
-$\hat{\lambda}^{(kl)}$ is a one-dimensional adjacency embedding approximation 
-of $\lambda^{(kl)}$.  
+Let $P$ and $A$ be organized by community such that the elements of blocks 
+$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between clusters $k$ and $l$.
+
+* *Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
+probabilities and edges for community $k$.  
+By definition, 
+$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
+signular value decomposition 
+$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
+one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
+singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
+Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
+value decomposition of $A^{(kk)}$, and let 
+$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
+one-dimensional approximation. Define 
+$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
+$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
+$\lambda^{(kk)}$.  
 Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
-embedding $\hat{\lambda}^{(kl)}$ approximates $\lambda^{(kl)}$ at rate 
-$\frac{(\log n_k)^c}{n_k^{1/2}}$.
+spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
+$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
 
-[^alg4]: Note that this estimator is similar to but slightly worse than the 
-estimator from Algorithm 4.
+* *Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
+edges between communities $k$ and $l$.  
+Let $R^{(kl)} = P^{(kl)} P^{(lk)}$. By definition, 
+$P^{(kl)} P^{(lk)} = 
+\lambda^{(kl)} (\lambda^{(lk)})^\top \lambda^{(lk)} (\lambda^{(kl)})^\top$. 
+Noting that the two middle terms compose a scalar, we can rewrite 
+$R^{(kk)} = c \lambda^{(kl)} (\lambda^{(kl)})^\top$, where $c = \tr(P^{(lk)})$. 
+As in the $k = l$ case, we note that the singular value decomposition 
+$R^{(kl)} = \sigma_{kl}^2 u^{(kl)} (u^{(kl)})^\top$ is one-dimensional and 
+$\lambda^{(kl)} = \sigma_{kl} u^{(kl)} / c$.  
+Let $B^{(kl)} = A^{(kl)} A^{(lk)}$, and let 
+$\hat{U}^{(kl)} \hat{\Sigma}^{(kl)} (\hat{U}^{(kl)})^\top$ be the singular 
+value decomposition of $B^{(kl)}$. Let 
+$\hat{\sigma}_{kl}^2 \hat{u}^{kl} (\hat{u}^{(kl)})^\top$ be the one-dimensional 
+approximation of $B^{(kl)}$ based on the singular value decomposition. Define 
+$\hat{\lambda}^{(kl)} = \hat{\sigma}_{kl} \hat{u}^{(kl)} / \tr(A^{(lk)})$. Then 
+$\hat{\lambda}^{(kl)}$ is the adjacency spectral embedding approximation of 
+$\lambda^{(kl)}$.  
+Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
+spectral embedding $\hat{\lambda}^{(kl)}$ approximates $\lambda^{(kl)}$ at rate 
+$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
 
@@
 
-For the real data example, we applied Algorithm 2 to the Leeds Butterfly 
+In the first real data example, we applied Algorithm 2 to the Leeds Butterfly 
 dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
@@
 thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
-by the resultant the clustering.
+by the resultant clustering.
 
@@
 
+```{r}
+# this portion is from the code provided by Sengupta and Chen
+
+setwd('~/Downloads/codes_and_data')
+library("igraph")
+library("clue")
+
+##### Data input and pre-processing #####
+foo=read.table("BritishMPcomm.txt",header=T)
+mat=read.table("politicsuk-retweets.mtx",header=F)
+mat = mat[-1,]
+A = matrix(0,nrow(foo),nrow(foo))
+for (k in 1:nrow(mat)){ #adjacency matrix with 5 parties
+  edgei = mat[k,1]
+  edgej = mat[k,2]
+  i = which(foo[,1]==edgei)
+  j = which(foo[,1]==edgej)
+  A[i,j]<-1;A[j,i]<-1
+}
+b.true <- foo[,2]
+levels(b.true) <- 1:5
+b.true <- as.numeric(b.true)
+nodes1 = which(b.true==1); nodes2 = which(b.true==2)
+nodes = c(nodes1, nodes2)
+A <- A[nodes,nodes] # Select only the two big communities
+G<-graph.adjacency(A, mode="undirected")
+foo1 <- clusters(G, mode="strong") 
+nodes2 <- which(foo1$membership==1) # largest connected component (basically 31 nodes are disconnected)
+A <- A[nodes2,nodes2] # form network with 329 nodes
+b.true <- b.true[nodes2]
+```
+
+```{r}
+zhat <- cluster.pabm(A, 2, use.all = TRUE, normalize = FALSE)
+clust.acc <- mean(zhat == b.true)
+```
+
+```{r mp, cache = TRUE, fig.width = 2, fig.height = 2, fig.cap = 'Adjacency matrix of the British MPs dataset after sorting by the clustering outputted by Algorithm 2.'}
+plot.A(A[order(zhat), order(zhat)])
+```
+
+In the second example, we applied Algorithm 2 to the British MPs Twitter 
+network \cite{greene2013producing}. The original data consists of 419 vertices, 
+each corresponding to a British MP belonging to one of five political parties. 
+For this data analysis, we subsetted the data as described by 
+\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
+dataset, resulting in 329 vertices and two parties ($n = 329, K = 2$). 
+\citet{307cbeb9b1be48299388437423d94bf1} were able to obtain a classification 
+accuracy of `r round(328 / 329 * 100, 1)`\%. Algorithm 2 obtained a lower 
+classification accuracy of `r round(clust.acc * 100, 1)`\%.
+
 # Discussion

### 2020-09-02 summary.Rmd
typos

@@
 adjacency matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
-consisting of $K$ underlying clusters/communities.
+consisting of $K$ underlying communities.
 
@@
 
-1. Cluster membership identification (up to permutation).
+1. Community membership identification (up to permutation).
 2. Parameter estimation (estimating $\lambda^{(kl)}$'s).
@@
 Here, we will focus more on (1) and pose possible methods for (2). In our 
-methods, we assume that $K$, the number of clusters, is known beforehand and 
+methods, we assume that $K$, the number of communities, is known beforehand and 
 does not require estimation.
@@
 \citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
-to identify the cluster memberships given either an edge probability matrix $P$ 
-or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
+to identify the community memberships given either an edge probability matrix $P$ 
+or an adjacency matrix $A$. In the case that $P$ is known, the community 
 memberships can be identified exactly (up to permutation). A similar procedure 
@@
 
-## Clustering
+## Community detection
 
@@
 edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
-(or can estimate) $XU$ directly, then both the clustering and parameter 
+(or can estimate) $XU$ directly, then both the community detection and parameter 
 identification problem are trivial since $U$ is orthonormal and fixed for each 
@@
 $\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
-and $j$ are in different clusters.
+and $j$ are in different communities.
 
 We can note from the explicit form of $XU$ that it represents points in 
-$\mathbb{R}^{K^2}$ such that points within each cluster lie on $K$-dimensional 
+$\mathbb{R}^{K^2}$ such that points within each community lie on $K$-dimensional 
 subspaces. Furthermore, the subspaces are orthogonal to each other. 
 Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
-but retains the property that each cluster is represented by a $K$-dimensional 
+but retains the property that each community is represented by a $K$-dimensional 
 subspace. Using this property, previous work proposes the use of subspace 
@@
 matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
-of different clusters.
+of different communities.
 
@@
 in Theorem 2. Since $X$ is block diagonal with each block corresponding 
-to one cluster, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
-with each block corresponding to a cluster and zeros elsewhere. Therefore, 
-if vertices $i$ and $j$ belong to different clusters, then the $ij$^th^ element 
-of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
+to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
+with each block corresponding to a community and zeros elsewhere. Therefore, 
+if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
+element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
 
@@
   \SetAlgoLined
-  \KwData{Edge probability matrix $P$, number of clusters $K$}
-  \KwResult{Cluster assignments $1, ..., K$}
+  \KwData{Edge probability matrix $P$, number of communities $K$}
+  \KwResult{Community labels $1, ..., K$}
     Compute the spectral decomposition $P = V D V^\top$.\;
@@
     Construct graph $G$ using $B$ as edge similarities.\;
-    Identify the connected components of $G$ and map each to cluster labels 
+    Identify the connected components of $G$ and map each to community labels 
     $1, ..., K.$\;
@@
 precisely, let $\{\lambda^{(kl)}\}_K \sim \mathcal{F}_K$ for some joint 
-distribution consisting of $K$ underlying clusters $\mathcal{F}_K$. Then the 
+distribution consisting of $K$ underlying communities $\mathcal{F}_K$. Then the 
 latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution 
-with $K$ underlying clusters $\mathcal{G}_K$. Denote $Z_n$ as a sample of size 
+with $K$ underlying communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size 
 $n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge 
@@
 $(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
-different clusters. Then for some $c > 0$,
+different communities. Then for some $c > 0$,
 
@@
   \SetAlgoLined
-  \KwData{Adjacency matrix $A$, number of clusters $K$}
-  \KwResult{Cluster assignments $1, ..., K$}
+  \KwData{Adjacency matrix $A$, number of communities $K$}
+  \KwResult{Community assignments $1, ..., K$}
     Compute the eigenvectors of $A$ that correspond to the $K (K+1) / 2$ most 
@@
     (e.g., using edge thresholding or spectral clustering).\;
-    Map each partition to the cluster labels $1, ..., K$.\;
+    Map each partition to the community labels $1, ..., K$.\;
   \caption{PABM clustering on the adjacency matrix.}
@@
 For any $P$ edge probability matrix for the PABM such that the rows and 
-columns are organized by cluster, the $kl$^th^ block is an outer product 
+columns are organized by community, the $kl$^th^ block is an outer product 
 of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
@@
   \caption{PABM parameter estimation using the edge probability matrix.}
-  \KwData{Edge probability matrix $P$, cluster assignments $1, ..., K$.}
+  \KwData{Edge probability matrix $P$, community assignments $1, ..., K$.}
   \KwResult{PABM parameters $\{\lambda^{(kl)}\}_K$}
-  Arrange the rows and columns of $P$ by cluster such that each $P^{(kl)}$ 
-  block consists of edge probabilities between clusters $k$ and $l$.\; 
+  Arrange the rows and columns of $P$ by community such that each $P^{(kl)}$ 
+  block consists of edge probabilities between communities $k$ and $l$.\; 
   \For{$k, l = 1, ..., K$, $k \geq l$} {
@@
   \caption{PABM parameter estimation using the adjacency matrix.}
-  \KwData{Adjacency matrix $A$, cluster assignments $1, ..., K$}
+  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
   \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
-  Arrange the rows and columns of $A$ by cluster such that each 
+  Arrange the rows and columns of $A$ by community such that each 
   $A^{(kl)}$ block consists of estimated edge probabilities between 
-  clusters $k$ and $l$.\;
+  communities $k$ and $l$.\;
   \For{$k, l = 1, ..., K$, $k \leq l$} {
@@
 Let $P$ and $A$ be organized by community such that the elements of blocks 
-$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between clusters $k$ and $l$.
+$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
+$l$.
 
@@
 
-* Number of underlying clusters $K = 2$
+* Number of underlying communities $K = 2$
 * Mixture parameter $\alpha = .5$
-* Within-cluster popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
-* Between-cluster popularities 
+* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* Between-group popularities 
 $\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
 
-### Clustering
+### Community detection
 
@@
 $\lambda^{(kl)}$'s are drawn and for each set of parameters, one adjacency 
-matrix $A$ is drawn and clustered. We will not consider clustering using the 
+matrix $A$ is drawn and clustered. We will not consider community detection using the 
 edge probability matrix $P$ since this will always result exact recovery of the
-original clusters (up to permutation).
+original communities (up to permutation).
 
@@
 
-* Number of underlying clusters $K = 2, 4, 8$
+* Number of underlying communities $K = 2, 4, 8$
 * Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$
-* Within-cluster popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
-* Between-cluster popularities 
+* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* Between-group popularities 
 $\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
 
-### Clustering
+### Community detection
 

### 2020-09-02 functions.R

@@ -1,5 +1,6 @@
 embedding <- function(A, p = NULL, q = NULL,
                       eps = 1e-6) {
+  n <- nrow(A)
   eigen.A <- eigen(A, symmetric = TRUE)
   if (is.null(p) | is.null(q)) {
     keep <- (abs(eigen.A$values) > eps)
@@ -59,15 +60,26 @@ normalized.laplacian <- function(W) {
   return(I - D.neg.sqrt %*% W %*% D.neg.sqrt)
 }
 
-cluster.pabm <- function(A, K) {
+graph.laplacian <- function(W) {
+  D <- diag(colSums(W))
+  L <- D - W
+  return(L)
+}
+
+cluster.pabm <- function(A, K, normalize = TRUE, use.all = FALSE) {
+  n <- nrow(A)
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(p), seq(n, n - q + 1))]
-  V <- V / mean(V)
+  if (normalize) V <- V / mean(V)
   B <- abs(V %*% t(V))
   # L <- graph.laplacian(B)
   L <- normalized.laplacian(B)
-  eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - K)]
+  if (use.all) {
+    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n, n - K)]
+  } else {
+    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - K)]
+  }
   clustering <- mclust::Mclust(eigenmap, K)$classification
 }
 

### 2020-09-07 summary.Rmd
fixed proof of theorem 5

@@
 0 & 1 & 0 & 0 \end{bmatrix}$, as in Definition 1. 
-Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
+Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
 equivalent.
@@
 * *Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
-edges between communities $k$ and $l$.  
-Let $R^{(kl)} = P^{(kl)} P^{(lk)}$. By definition, 
-$P^{(kl)} P^{(lk)} = 
-\lambda^{(kl)} (\lambda^{(lk)})^\top \lambda^{(lk)} (\lambda^{(kl)})^\top$. 
-Noting that the two middle terms compose a scalar, we can rewrite 
-$R^{(kk)} = c \lambda^{(kl)} (\lambda^{(kl)})^\top$, where $c = \tr(P^{(lk)})$. 
-As in the $k = l$ case, we note that the singular value decomposition 
-$R^{(kl)} = \sigma_{kl}^2 u^{(kl)} (u^{(kl)})^\top$ is one-dimensional and 
-$\lambda^{(kl)} = \sigma_{kl} u^{(kl)} / c$.  
-Let $B^{(kl)} = A^{(kl)} A^{(lk)}$, and let 
-$\hat{U}^{(kl)} \hat{\Sigma}^{(kl)} (\hat{U}^{(kl)})^\top$ be the singular 
-value decomposition of $B^{(kl)}$. Let 
-$\hat{\sigma}_{kl}^2 \hat{u}^{kl} (\hat{u}^{(kl)})^\top$ be the one-dimensional 
-approximation of $B^{(kl)}$ based on the singular value decomposition. Define 
-$\hat{\lambda}^{(kl)} = \hat{\sigma}_{kl} \hat{u}^{(kl)} / \tr(A^{(lk)})$. Then 
-$\hat{\lambda}^{(kl)}$ is the adjacency spectral embedding approximation of 
-$\lambda^{(kl)}$.  
-Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
-spectral embedding $\hat{\lambda}^{(kl)}$ approximates $\lambda^{(kl)}$ at rate 
-$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
+edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
+By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
+$k = l$ case, we note that the singular value decomposition 
+$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
+$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
+$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
+$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
+$v^{(kl)} = u^{(lk)}$.)  
+Now consider the Hermitian dilation 
+$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$, 
+which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
+that the eigendecomposition of $M^{(kl)}$ is  
+$M^{(kl)} = 
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
+\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$.  
+Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
+latent positions in $\mathbb{R}^2$ given by 
+$\begin{bmatrix} 
+  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
+  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
+\end{bmatrix} = 
+\begin{bmatrix} 
+  \lambda^{(kl)} & \lambda^{(kl)} \\ 
+  \lambda^{(lk)} & -\lambda^{(lk)} 
+\end{bmatrix}$.  
+Now consider $M' = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$. 
+Then $M' = M + E'$ where 
+$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$ and $E$ is the 
+$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
+in $A^{(kl)}$.  
+Then $M'$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
+spectral embedding, given by 
+$\begin{bmatrix} 
+  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
+  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
+\end{bmatrix}$ 
+where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
+the latent positions of $M^{(kl)}$ by the rate in Theorem 5 of 
+\citet{rubindelanchy2017statistical}.  
+In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
+result \cite{rubindelanchy2017statistical} is of the form 
+$\hat{U}^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
+for this GRDPG is $(1, 1)$, and $\hat{U}^\top \hat{U}$ is block diagonal, 
+resulting in $W_* = I$. Therefore, the adjacency spectral embedding of $M'$ is 
+a direct estimation of the specific latent positions outlined for $M^{(kl)}$.
 

### 2020-09-09 summary.Rmd
typo
minor tweak to clustering function

@@
 resulting in $W_* = I$. Therefore, the adjacency spectral embedding of $M'$ is 
-a direct estimation of the specific latent positions outlined for $M^{(kl)}$.
+a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
+up to sign flip.
 
@@
 \citet{307cbeb9b1be48299388437423d94bf1} were able to obtain a classification 
-accuracy of `r round(328 / 329 * 100, 1)`\%. Algorithm 2 obtained a lower 
-classification accuracy of `r round(clust.acc * 100, 1)`\%.
+accuracy of `r round(328 / 329 * 100, 1)`\% via modularity maximization. 
+Algorithm 2 obtained a lower 
+classification accuracy of `r round(clust.acc * 100, 1)`\%. We can also note 
+that Algorithm 2 had a much shorter runtime than modularity maximization.
 

### 2020-09-09 functions.R

@@ -66,15 +66,23 @@ graph.laplacian <- function(W) {
   return(L)
 }
 
-cluster.pabm <- function(A, K, normalize = TRUE, use.all = FALSE) {
+cluster.pabm <- function(A, K, 
+                         normalize = TRUE, 
+                         use.all = FALSE,
+                         laplacian = 'normalized') {
   n <- nrow(A)
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(p), seq(n, n - q + 1))]
-  if (normalize) V <- V / mean(V)
+  if (normalize) {
+    V <- sweep(V, 2, abs(A.eigen$values[c(seq(p), seq(n, n - q + 1))]) ** .5, `*`)
+  }
   B <- abs(V %*% t(V))
-  # L <- graph.laplacian(B)
-  L <- normalized.laplacian(B)
+  if (laplacian == 'graph') {
+    L <- graph.laplacian(B)
+  } else {
+    L <- normalized.laplacian(B)
+  }
   if (use.all) {
     eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n, n - K)]
   } else {

### 2020-09-14 summary.Rmd
can now be compiled without R

@@
 knitr::opts_chunk$set(echo = FALSE, 
+                      # eval = FALSE,
                       comment = NA, 
@@
 Model, can be represented in this way \cite{rubindelanchy2017statistical}. An 
-analogous property exists for the PABM, not under the RDPG model but under the *Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
+analogous property exists for the PABM, not under the RDPG model but under the 
+*Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
 explored in this paper and exploited to construct algorithms for community 
@@
 Comparing against the ground truth species labels, Algorithm 2 achieves an 
-accuracy of 63\% and an adjusted Rand index of `r ari`\%. In comparison, 
+accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
 \citet{noroozi2019estimation} achieved an adjusted Rand index of 73\% using 
@@
 Algorithm 2 obtained a lower 
-classification accuracy of `r round(clust.acc * 100, 1)`\%. We can also note 
+classification accuracy of 99.1\%. We can also note 
 that Algorithm 2 had a much shorter runtime than modularity maximization.

### 2020-09-16 summary.Rmd
can now compile without files outside repository

@@
 library(ggplot2)
-library(plot.matrix)
 
@@
 \end{bmatrix}$.  
-Now consider $M' = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$. 
-Then $M' = M + E'$ where 
+Now consider $\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$. 
+Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where 
 $E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$ and $E$ is the 
@@
 in $A^{(kl)}$.  
-Then $M'$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
+Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
 spectral embedding, given by 
@@
 result \cite{rubindelanchy2017statistical} is of the form 
-$\hat{U}^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
-for this GRDPG is $(1, 1)$, and $\hat{U}^\top \hat{U}$ is block diagonal, 
-resulting in $W_* = I$. Therefore, the adjacency spectral embedding of $M'$ is 
+$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
+for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
+resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
 a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
@@
 
-butterfly <- R.matlab::readMat('~/Downloads/Raw_butterfly_network.mat')
+butterfly <- R.matlab::readMat('data/Raw_butterfly_network.mat')
 
@@
 ```{r}
-plot.A <- function(A, margins = c(0, 0, 0, 0)) {
-  default.mar <- par()$mar
-  par(mar = margins)
-  plot(A, border = NA, key = NULL, 
-       xlab = '', ylab = '', main = '',
-       axis.row = NULL, axis.col = NULL, asp = 1)
-  par(mar = default.mar)
+# plot.A <- function(A, margins = c(0, 0, 0, 0)) {
+#   default.mar <- par()$mar
+#   par(mar = margins)
+#   plot(A, border = NA, key = NULL, 
+#        xlab = '', ylab = '', main = '',
+#        axis.row = NULL, axis.col = NULL, asp = 1)
+#   par(mar = default.mar)
+# }
+
+# plot.A <- function(A, labels) {
+#   heatmap(A[order(labels), order(labels)],
+#           margins = c(0, 0),
+#           revC = TRUE)
+# }
+
+plot.A <- function(A, labels) {
+  # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
+  A[order(labels), order(labels)] %>% 
+    tibble::as_tibble() %>% 
+    tibble::rowid_to_column(var = 'X') %>% 
+    tidyr::gather(key = 'Y', value = 'Z', -1) %>% 
+    dplyr::mutate(Y = as.numeric(gsub('V', '', Y))) %>% 
+    ggplot() + 
+    geom_tile(aes(x = X, y = Y, fill = Z)) + 
+    coord_fixed() + 
+    labs(x = NULL, y = NULL) + 
+    theme_void() + 
+    theme(legend.position = 'none',
+          axis.title.x = element_blank(),
+          axis.text.x = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.y = element_blank()) + 
+    scale_x_continuous(expand = c(0, 0)) + 
+    scale_y_continuous(expand = c(0, 0)) + 
+    scale_fill_gradient(low = 'yellow', high = 'red')
 }
@@
 # this portion is from the code provided by Sengupta and Chen
+# the data were generated by scripts from S&C, then saved as RDS files
 
-setwd('~/Downloads/codes_and_data')
-library("igraph")
-library("clue")
+bmp <- readRDS('data/bmp.rds')
+A.bmp <- bmp$A
+z.bmp <- bmp$z
 
-##### Data input and pre-processing #####
-foo=read.table("BritishMPcomm.txt",header=T)
-mat=read.table("politicsuk-retweets.mtx",header=F)
-mat = mat[-1,]
-A = matrix(0,nrow(foo),nrow(foo))
-for (k in 1:nrow(mat)){ #adjacency matrix with 5 parties
-  edgei = mat[k,1]
-  edgej = mat[k,2]
-  i = which(foo[,1]==edgei)
-  j = which(foo[,1]==edgej)
-  A[i,j]<-1;A[j,i]<-1
-}
-b.true <- foo[,2]
-levels(b.true) <- 1:5
-b.true <- as.numeric(b.true)
-nodes1 = which(b.true==1); nodes2 = which(b.true==2)
-nodes = c(nodes1, nodes2)
-A <- A[nodes,nodes] # Select only the two big communities
-G<-graph.adjacency(A, mode="undirected")
-foo1 <- clusters(G, mode="strong") 
-nodes2 <- which(foo1$membership==1) # largest connected component (basically 31 nodes are disconnected)
-A <- A[nodes2,nodes2] # form network with 329 nodes
-b.true <- b.true[nodes2]
+pb <- readRDS('data/pb.rds')
+A.pb <- pb$A
+z.pb <- pb$z
+
+dblp <- readRDS('data/dblp.rds')
+A.dblp <- dblp$A
+z.dblp <- dblp$z
 ```
 
-```{r}
-zhat <- cluster.pabm(A, 2, use.all = TRUE, normalize = FALSE)
-clust.acc <- mean(zhat == b.true)
-```
+```{r, cache = TRUE}
+zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = TRUE)
+bmp.acc <- mean(zhat.bmp == z.bmp)
 
-```{r mp, cache = TRUE, fig.width = 2, fig.height = 2, fig.cap = 'Adjacency matrix of the British MPs dataset after sorting by the clustering outputted by Algorithm 2.'}
-plot.A(A[order(zhat), order(zhat)])
+zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = TRUE, 
+                        p = 2, q = 0)
+pb.acc <- mean(zhat.pb == z.pb)
+
+zhat.dblp <- cluster.pabm(A.dblp, 2, use.all = TRUE, normalize = FALSE)
+dblp.acc <- 1 - mean(zhat.dblp == z.dblp)
 ```
@@
 In the second example, we applied Algorithm 2 to the British MPs Twitter 
-network \cite{greene2013producing}. The original data consists of 419 vertices, 
-each corresponding to a British MP belonging to one of five political parties. 
-For this data analysis, we subsetted the data as described by 
+network \cite{greene2013producing}, the Political Blogs network and the DBLP network. For this data analysis, we subsetted the data as described by 
 \citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
-dataset, resulting in 329 vertices and two parties ($n = 329, K = 2$). 
-\citet{307cbeb9b1be48299388437423d94bf1} were able to obtain a classification 
-accuracy of `r round(328 / 329 * 100, 1)`\% via modularity maximization. 
-Algorithm 2 obtained a lower 
-classification accuracy of 99.1\%. We can also note 
-that Algorithm 2 had a much shorter runtime than modularity maximization.
+networks. Our methods underperformed compared to modularity maximization, although performance is comparable for the British MPs and DBLP networks.
+
+```{r}
+dplyr::tibble(network = c('British MPs', 'Political blogs', 'DBLP'),
+              vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
+              `error (modularity maximization)` = c(round(1 / 329, 3),
+                                                    round(61 / 1222, 3),
+                                                    round(62 / 2203, 3)),
+              `error (algorithm 2)` = round(1 - c(bmp.acc, 
+                                                  pb.acc,
+                                                  dblp.acc), 3)) %>% 
+  knitr::kable(caption = paste('Community detection error rates', 
+                               'for modularity maximization',
+                               '(\\citet{307cbeb9b1be48299388437423d94bf1})',
+                               'and Algorithm 2.'))
+```
+
+```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.'}
+gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
+                        plot.A(A.pb, zhat.pb),
+                        plot.A(A.dblp, zhat.dblp),
+                        ncol = 3)
+
+```
 

### 2020-09-16 functions.R

@@ -68,20 +68,32 @@ graph.laplacian <- function(W) {
 
 cluster.pabm <- function(A, K, 
                          normalize = TRUE, 
-                         use.all = FALSE,
+                         use.all = TRUE,
+                         p = NULL, q = NULL, 
                          laplacian = 'normalized') {
   n <- nrow(A)
-  p <- K * (K + 1) / 2
-  q <- K * (K - 1) / 2
-  V <- eigen(A, symmetric = TRUE)$vectors[, c(seq(p), seq(n, n - q + 1))]
+  if (is.null(p)) {
+    p <- K * (K + 1) / 2
+  }
+  if (is.null(q)) {
+    q <- K * (K - 1) / 2
+  }
+  if (p * q > 0) {
+    indices <- c(seq(p), seq(n, n - q + 1))
+  } else if (p == 0) {
+    indices <- seq(n, n - q + q)
+  } else if (q == 0) {
+    indices <- seq(p)
+  }
+  V <- eigen(A, symmetric = TRUE)$vectors[, indices]
   if (normalize) {
-    v.norm <- apply(V, 2, function(v) sum(v ** 2))
+    v.norm <- sqrt(apply(V, 1, function(v) sum(v ** 2)))
     V <- sweep(V, 1, v.norm, `/`)
   }
   B <- abs(V %*% t(V))
   if (laplacian == 'graph') {
     L <- graph.laplacian(B)
-  } else {
+  } else if (laplacian == 'normalized') {
     L <- normalized.laplacian(B)
   }
   if (use.all) {

### 2020-09-16 summary.Rmd
forgot to add minor addition

@@
 In the second example, we applied Algorithm 2 to the British MPs Twitter 
-network \cite{greene2013producing}, the Political Blogs network and the DBLP network. For this data analysis, we subsetted the data as described by 
+network \cite{greene2013producing}, the Political Blogs network and the DBLP n
+etwork. For this data analysis, we subsetted the data as described by 
 \citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
-networks. Our methods underperformed compared to modularity maximization, although performance is comparable for the British MPs and DBLP networks.
+networks. Our methods underperformed compared to modularity maximization, 
+although performance is comparable for the British MPs and DBLP networks. In 
+addition, Algorithm 2's runtime is much lower than that of modularity 
+maximization.
 

### 2020-09-16 summary.Rmd
forgot to add some citations
fixed figures

@@
 library(ggplot2)
+library(mclust)
 
@@
 where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
-the latent positions of $M^{(kl)}$ by the rate in Theorem 5 of 
-\citet{rubindelanchy2017statistical}.  
+the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
+by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
 In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
@@
 ```{r}
-library(mclust)
 import::here(cluster.pabm, .from = '~/dev/pabm-grdpg/functions.R')
@@
 
-ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
+# ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
 ```
@@
     tidyr::gather(key = 'Y', value = 'Z', -1) %>% 
-    dplyr::mutate(Y = as.numeric(gsub('V', '', Y))) %>% 
+    dplyr::mutate(Y = rev(as.numeric(gsub('V', '', Y)))) %>% 
     ggplot() + 
@@
     scale_y_continuous(expand = c(0, 0)) + 
-    scale_fill_gradient(low = 'yellow', high = 'red')
+    scale_fill_gradient(low = 'red', high = 'yellow')
 }
@@
 In the second example, we applied Algorithm 2 to the British MPs Twitter 
-network \cite{greene2013producing}, the Political Blogs network and the DBLP n
-etwork. For this data analysis, we subsetted the data as described by 
+network \cite{greene2013producing}, the Political Blogs network 
+\cite{10.1145/1134271.1134277}, and the DBLP network \cite{NIPS2009_3855} 
+\cite{10.1007/978-3-642-15880-3_42}. For this data analysis, we subsetted the 
+data as described by 
 \citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
@@
 
-```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.'}
+```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 25}
 gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),

### 2020-09-21 summary.Rmd
intermediate commit to move some files

@@
 ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods, colour = factor(n))) + 
+  geom_density(aes(x = inner.prods * sqrt(n), colour = factor(n))) + 
   labs(x = 'between cluster inner products',
@@
 z.hat <- cluster.pabm(A, K)
-A.star <- A[order(z.hat), order(z.hat)]
 
@@
 
-plot.A <- function(A, labels) {
+plot.A <- function(A, labels, max.size = 500) {
   # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
-  A[order(labels), order(labels)] %>% 
+  A <- A[order(labels), order(labels)]
+  n <- nrow(A)
+  if (n > max.size) {
+    ind <- sort(sample(seq(n), max.size))
+    A <- A[ind, ind]
+  }
+  A %>% 
     tibble::as_tibble() %>% 
@@
 
-```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = TRUE}
-plot.A(A.star)
+```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = FALSE}
+plot.A(A, z.hat)
 ```
@@
 ```{r, cache = TRUE}
-zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = TRUE)
+zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = FALSE)
 bmp.acc <- mean(zhat.bmp == z.bmp)
 
-zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = TRUE, 
-                        p = 2, q = 0)
+zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = FALSE, 
+                        p = 2, q = 0, d.eigenmap = 2)
 pb.acc <- mean(zhat.pb == z.pb)
@@
 
-```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 25}
+```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
 gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
@@
                         ncol = 3)
-
 ```

### 2020-09-21 functions.R

@@ -69,7 +69,7 @@ graph.laplacian <- function(W) {
 cluster.pabm <- function(A, K, 
                          normalize = TRUE, 
                          use.all = TRUE,
-                         p = NULL, q = NULL, 
+                         p = NULL, q = NULL, d.eigenmap = K + 1, 
                          laplacian = 'normalized') {
   n <- nrow(A)
   if (is.null(p)) {
@@ -87,7 +87,7 @@ cluster.pabm <- function(A, K,
   }
   V <- eigen(A, symmetric = TRUE)$vectors[, indices]
   if (normalize) {
-    v.norm <- sqrt(apply(V, 1, function(v) sum(v ** 2)))
+    v.norm <- sqrt(rowSums(V ** 2))
     V <- sweep(V, 1, v.norm, `/`)
   }
   B <- abs(V %*% t(V))
@@ -97,9 +97,9 @@ cluster.pabm <- function(A, K,
     L <- normalized.laplacian(B)
   }
   if (use.all) {
-    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n, n - K)]
+    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n, n - d.eigenmap + 1)]
   } else {
-    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - K)]
+    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - d.eigenmap + 1)]
   }
   clustering <- mclust::Mclust(eigenmap, K)$classification
 }

### 2020-09-23 summary.Rmd
intermediate commit to move stuff around

@@
 bmp.acc <- mean(zhat.bmp == z.bmp)
+bmp.acc <- ifelse(bmp.acc < .5, 1 - bmp.acc, bmp.acc)
 
@@
 pb.acc <- mean(zhat.pb == z.pb)
+pb.acc <- ifelse(pb.acc < .5, 1 - pb.acc, pb.acc)
 
 zhat.dblp <- cluster.pabm(A.dblp, 2, use.all = TRUE, normalize = FALSE)
-dblp.acc <- 1 - mean(zhat.dblp == z.dblp)
+dblp.acc <- mean(zhat.dblp == z.dblp)
+dblp.acc <- ifelse(dblp.acc < .5, 1 - dblp.acc, dblp.acc)
 ```

### 2020-09-23 summary.Rmd
still wip

@@
 ```{r}
-import::here(cluster.pabm, .from = '~/dev/pabm-grdpg/functions.R')
+import::here(cluster.pabm, plot.A, .from = '~/dev/pabm-grdpg/functions.R')
 
@@
 sparse subspace clustering on the same dataset.
-
-```{r}
-# plot.A <- function(A, margins = c(0, 0, 0, 0)) {
-#   default.mar <- par()$mar
-#   par(mar = margins)
-#   plot(A, border = NA, key = NULL, 
-#        xlab = '', ylab = '', main = '',
-#        axis.row = NULL, axis.col = NULL, asp = 1)
-#   par(mar = default.mar)
-# }
-
-# plot.A <- function(A, labels) {
-#   heatmap(A[order(labels), order(labels)],
-#           margins = c(0, 0),
-#           revC = TRUE)
-# }
-
-plot.A <- function(A, labels, max.size = 500) {
-  # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
-  A <- A[order(labels), order(labels)]
-  n <- nrow(A)
-  if (n > max.size) {
-    ind <- sort(sample(seq(n), max.size))
-    A <- A[ind, ind]
-  }
-  A %>% 
-    tibble::as_tibble() %>% 
-    tibble::rowid_to_column(var = 'X') %>% 
-    tidyr::gather(key = 'Y', value = 'Z', -1) %>% 
-    dplyr::mutate(Y = rev(as.numeric(gsub('V', '', Y)))) %>% 
-    ggplot() + 
-    geom_tile(aes(x = X, y = Y, fill = Z)) + 
-    coord_fixed() + 
-    labs(x = NULL, y = NULL) + 
-    theme_void() + 
-    theme(legend.position = 'none',
-          axis.title.x = element_blank(),
-          axis.text.x = element_blank(),
-          axis.ticks.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks.y = element_blank()) + 
-    scale_x_continuous(expand = c(0, 0)) + 
-    scale_y_continuous(expand = c(0, 0)) + 
-    scale_fill_gradient(low = 'red', high = 'yellow')
-}
-```
 
@@
 
+In the third example,
+
+```{r, cache = TRUE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.'}
+A.hh <- readRDS('data/village-hh-59.rds')
+zhat.hh <- cluster.pabm(A.hh, 6)
+plot.A(A.hh, zhat.hh)
+```
+
 # Discussion

### 2020-09-23 functions.R

@@ -339,3 +339,33 @@ generate.P.beta <- function(n, K = 2, a1 = 2, b1 = 1, a2 = 1, b2 = 2) {
   }
   return(list(P = P, clustering = clustering))
 }
+
+plot.A <- function(A, labels, max.size = 500) {
+  # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
+  A <- A[order(labels), order(labels)]
+  n <- nrow(A)
+  if (n > max.size) {
+    ind <- sort(sample(seq(n), max.size))
+    A <- A[ind, ind]
+  }
+  A %>% 
+    tibble::as_tibble() %>% 
+    tibble::rowid_to_column(var = 'X') %>% 
+    tidyr::gather(key = 'Y', value = 'Z', -1) %>% 
+    dplyr::mutate(Y = rev(as.numeric(gsub('V', '', Y)))) %>% 
+    ggplot() + 
+    geom_tile(aes(x = X, y = Y, fill = Z)) + 
+    coord_fixed() + 
+    labs(x = NULL, y = NULL) + 
+    theme_void() + 
+    theme(legend.position = 'none',
+          axis.title.x = element_blank(),
+          axis.text.x = element_blank(),
+          axis.ticks.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks.y = element_blank()) + 
+    scale_x_continuous(expand = c(0, 0)) + 
+    scale_y_continuous(expand = c(0, 0)) + 
+    scale_fill_gradient(low = 'red', high = 'yellow')
+}

### 2020-09-23 summary.Rmd
updated density plot

@@
 ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods * sqrt(n), colour = factor(n))) + 
-  labs(x = 'between cluster inner products',
+  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
+  labs(x = expression(n * v[i]^T * v[j]),
        colour = 'n') + 
@@
 
-```{r, cache = TRUE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.'}
+```{r, cache = TRUE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
 A.hh <- readRDS('data/village-hh-59.rds')

### 2020-09-23 summary.Rmd
fixed citation for pabm

@@
 
-## The popularity adjusted block model (PABM) \cite{noroozi2019estimation} and the generalized random dot product graph \cite{rubindelanchy2017statistical}
+## The popularity adjusted block model (PABM) \cite{307cbeb9b1be48299388437423d94bf1} and the generalized random dot product graph \cite{rubindelanchy2017statistical}
 

### 2020-09-30 mod-max-sim.R
modularity maximization results

@@ -1,0 +1,95 @@
+# simulation using proposed clustering method
+# run on IU RED (not a good idea to attempt on laptop)
+# this code is mostly from scripts provided by sengupta & chen
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+library(igraph)
+library(clue)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, 
+             generate.P.beta,
+             normalized.laplacian,
+             cluster.pabm, ssc,
+             .from = 'functions.R')
+setwd('codes_and_data')
+source('functions.R')
+source('EPalgosim.R')
+setwd('..')
+
+mod.max.clust <- function(A) {
+  # code copy-pasted from sengupta & chen
+  Eps <- (0:19)/20
+  b.can <- NULL
+  for (ieps in 1:length(Eps)){
+    b.can<-c(b.can,list(EPalgo(A,eps=Eps[ieps]))) # eps=0 for unperturbed version
+  }
+  
+  b.can.unpert <- b.can[[1]]     # matrix of candidate assignments for eps = 0
+  n.can.unpert <- dim(b.can.unpert)[2]  # total number of candidate assignments
+  Q.PA.unpert = rep(NA, n.can.unpert)  # array to store modularity values
+  
+  ##### modularity calculation #####
+  ptm<-proc.time()
+  for (ican in 1:n.can.unpert){
+    Q.PA.unpert[ican] = Q.PA(A, b=b.can.unpert[,ican])
+  }
+  index.PA<-which(Q.PA.unpert==max(Q.PA.unpert))
+  b.PA.unpert = b.can.unpert[,index.PA]
+  
+  return(b.PA.unpert)
+}
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation parameters
+K <- 2
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- .01
+n.vec <- c(64, 128, 256, 512, 1024, 2048)
+iter <- 100
+set.seed(314159)
+
+# clustering simulation
+clustering.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  print(paste('K =', K, ', n =', n))
+  foreach(i = seq(iter), 
[…]

### 2020-09-30 summary.Rmd
temporary push

@@
 \newcommand{\tr}{\text{Tr}}
+\newcommand{\blockdiag}{\text{blockdiag}}
 
@@
 Let $X$ be a block diagonal matrix 
-$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
+$X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
 Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
-$l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
+$l = 1, ..., K$. $L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
 \mathbb{R}^{n \times K}$.  
@@
 
-```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
+```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
 gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
@@
 
-```{r, cache = TRUE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
-A.hh <- readRDS('data/village-hh-59.rds')
-zhat.hh <- cluster.pabm(A.hh, 6)
-plot.A(A.hh, zhat.hh)
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
+hh <- readRDS('data/village-hh-59.rds')
+A.hh <- hh$A
+z.hh <- hh$z
+K <- 6
+zhat.hh <- cluster.pabm(A.hh, K, 
+                        use.all = TRUE, normalize = FALSE, p = K, q = 0)
+
+zhat.mres <- readRDS('data/mres-result.rds')
+
+gridExtra::grid.arrange(plot.A(A.hh, zhat.hh), 
+                        plot.A(A.hh, zhat.mres),
+                        ncol = 2)
 ```

### 2020-09-30 functions.R

@@ -340,15 +340,17 @@ generate.P.beta <- function(n, K = 2, a1 = 2, b1 = 1, a2 = 1, b2 = 2) {
   return(list(P = P, clustering = clustering))
 }
 
-plot.A <- function(A, labels, max.size = 500) {
+plot.A <- function(A, labels, lines = TRUE, max.size = 500) {
   # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
   A <- A[order(labels), order(labels)]
+  labels <- sort(labels)
   n <- nrow(A)
   if (n > max.size) {
     ind <- sort(sample(seq(n), max.size))
     A <- A[ind, ind]
+    labels <- labels[ind]
   }
-  A %>% 
+  out.plot <- A %>% 
     tibble::as_tibble() %>% 
     tibble::rowid_to_column(var = 'X') %>% 
     tidyr::gather(key = 'Y', value = 'Z', -1) %>% 
@@ -357,15 +359,25 @@ plot.A <- function(A, labels, max.size = 500) {
     geom_tile(aes(x = X, y = Y, fill = Z)) + 
     coord_fixed() + 
     labs(x = NULL, y = NULL) + 
-    theme_void() + 
+    theme_void() +
     theme(legend.position = 'none',
           axis.title.x = element_blank(),
           axis.text.x = element_blank(),
           axis.ticks.x = element_blank(),
           axis.title.y = element_blank(),
           axis.text.y = element_blank(),
-          axis.ticks.y = element_blank()) + 
+          axis.ticks.y = element_blank()) +
     scale_x_continuous(expand = c(0, 0)) + 
     scale_y_continuous(expand = c(0, 0)) + 
     scale_fill_gradient(low = 'red', high = 'yellow')
+  
+  if (lines) {
+    positions.x <- which(diff(labels) != 0)
+    positions.y <- which(diff(rev(labels)) != 0)
+    out.plot <- out.plot + 
+      geom_hline(yintercept = positions.y) + 
+      geom_vline(xintercept = positions.x)
+  }
+  
+  return(out.plot)
 }

### 2020-09-30 summary.Rmd
results from mod max

@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (green) and modularity maximization (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
 Simulations were repeated 100 times for each sample size.
@@
 clustering.df <- readr::read_csv('clustering-k.csv')
+mod.max.df <- readr::read_csv('mod-max-k.csv') %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>% 
+  dplyr::ungroup()
 clustering.df %>%
@@
                     colour = 'SSC')) + 
+  geom_line(data = mod.max.df,
+            aes(x = n, y = med.err.mm,
+                colour = 'modularity maximization')) + 
+  geom_errorbar(data = mod.max.df,
+                aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'modularity maximization')) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left) and results of \\citet{fosdick2016multiresolution} (right).', fig.height = 2}
 hh <- readRDS('data/village-hh-59.rds')

### 2020-09-30 density-sim.R
fixed density simulation

@@ -1,0 +1,52 @@
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, draw.pabm.beta.2,
+             normalized.laplacian,
+             cluster.pabm, ssc, 
+             embedding, 
+             .from = 'functions.R')
+library(ggplot2)
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores())
+
+# simulation parameters
+K <- 2
+p <- K * (K + 1) / 2
+q <- K * (K - 1) / 2
+alpha <- .5
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- .01
+# n.vec <- c(64, 128, 256, 512, 1024, 2048)
+n.vec <- c(512, 1024, 2048, 3072, 4096)
+iter <- 100
+subsample <- 1e3
+set.seed(314159)
+
+# clustering simulation
+densities.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
+  A <- Az$A
+  z <- Az$z
+  A.eigen <- eigen(A, symmetric = TRUE)
+  V <- A.eigen$vectors[, c(1:3, n)]
+  B <- V %*% t(V)
+  between.clust <- B[z == 1, z == 2]
+  between.clust <- between.clust[upper.tri(between.clust)]
+  between.clust <- sample(between.clust, subsample)
+  dplyr::tibble(n = n, inner.prods = between.clust) %>% 
+    return()
+}
+
+ggplot(densities.df) + 
+  geom_density(aes(x = inner.prods * n, colour = factor(n)))
+
+readr::write_csv(densities.df, 'densities.csv')

### 2020-10-08 summary.Rmd
updated final real data example

@@
 * Number of underlying communities $K = 2, 4, 8$
-* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$
+* Mixture parameters $\alpha_K = 1 / K$ for $k = 1, ..., K$
+* Community labels 
+$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
 * Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
@@
 ```{r}
-import::here(cluster.pabm, plot.A, .from = '~/dev/pabm-grdpg/functions.R')
+import::here(cluster.pabm, ssc, plot.A, .from = '~/dev/pabm-grdpg/functions.R')
 
@@
 
-In the third example,
+In the third example, we consider the Karantaka villages data studied by 
+\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` network from 
+village 12 at the household level. The label of interest is the religious 
+affiliation. The network was truncated to religions "1" and "2", and vertices 
+of degree 0 were removed. The resulting network has 141 vertices.
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left) and results of \\citet{fosdick2016multiresolution} (right).', fig.height = 2}
-hh <- readRDS('data/village-hh-59.rds')
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left) and sparse subspace clustering (right).', fig.height = 2, fig.width = 6, eval = TRUE}
+hh <- readRDS('data/village-hh-12.rds')
 A.hh <- hh$A
 z.hh <- hh$z
-K <- 6
+K <- 2
 zhat.hh <- cluster.pabm(A.hh, K,
-                        use.all = TRUE, normalize = FALSE, p = K, q = 0)
-
-zhat.mres <- readRDS('data/mres-result.rds')
+                        use.all = TRUE, normalize = FALSE, p = 3, q = 1)
+zhat.hh.ssc <- ssc(A.hh, K, lambda = .1)
 
 gridExtra::grid.arrange(plot.A(A.hh, zhat.hh),
-                        plot.A(A.hh, zhat.mres),
+                        plot.A(A.hh, zhat.hh.ssc),
                         ncol = 2)
+```
+
+```{r, eval = TRUE}
+hh.acc <- mean(zhat.hh == z.hh)
+hh.acc <- ifelse(hh.acc < .5, 1 - hh.acc, hh.acc)
+hh.acc.ssc <- mean(zhat.hh.ssc == z.hh)
+hh.acc.ssc <- ifelse(hh.acc.ssc < .5, 1 - hh.acc.ssc, hh.acc.ssc)
+
+knitr::kable(data.frame(method = c('Algorithm 2', 'SSC'),
+                        accuracy = round(c(hh.acc, hh.acc.ssc), 3)))
 ```

### 2020-10-08 functions.R

@@ -97,9 +97,11 @@ cluster.pabm <- function(A, K,
     L <- normalized.laplacian(B)
   }
   if (use.all) {
-    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n, n - d.eigenmap + 1)]
+    eigenmap <- 
+      eigen(L, symmetric = TRUE)$vectors[, seq(n, n - d.eigenmap + 1)]
   } else {
-    eigenmap <- eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - d.eigenmap + 1)]
+    eigenmap <- 
+      eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - d.eigenmap + 1)]
   }
   clustering <- mclust::Mclust(eigenmap, K)$classification
 }
@@ -340,15 +342,15 @@ generate.P.beta <- function(n, K = 2, a1 = 2, b1 = 1, a2 = 1, b2 = 2) {
   return(list(P = P, clustering = clustering))
 }
 
-plot.A <- function(A, labels, lines = TRUE, max.size = 500) {
+plot.A <- function(A, z, lines = TRUE, max.size = 500) {
   # https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html
-  A <- A[order(labels), order(labels)]
-  labels <- sort(labels)
+  A <- A[order(z), order(z)]
+  z <- sort(z)
   n <- nrow(A)
   if (n > max.size) {
     ind <- sort(sample(seq(n), max.size))
     A <- A[ind, ind]
-    labels <- labels[ind]
+    z <- z[ind]
   }
   out.plot <- A %>% 
     tibble::as_tibble() %>% 
@@ -372,11 +374,11 @@ plot.A <- function(A, labels, lines = TRUE, max.size = 500) {
     scale_fill_gradient(low = 'red', high = 'yellow')
   
   if (lines) {
-    positions.x <- which(diff(labels) != 0)
-    positions.y <- which(diff(rev(labels)) != 0)
+    positions.x <- which(diff(z) != 0)
+    positions.y <- which(diff(rev(z)) != 0)
     out.plot <- out.plot + 
-      geom_hline(yintercept = positions.y) + 
-      geom_vline(xintercept = positions.x)
+      geom_hline(yintercept = positions.y + .5) + 
+      geom_vline(xintercept = positions.x + .5)
   }
   
   return(out.plot)

### 2020-10-14 clustering-sim-k.R
updating simulations

@@ -14,23 +14,36 @@ import::here(embedding, draw.graph,
              cluster.pabm, ssc,
              .from = 'functions.R')
 
+cluster.acc <- function(yhat, yobs) {
+  table(yhat, yobs) %>% 
+    as.matrix() %>% 
+    apply(1, max) %>% 
+    sum() %>% 
+    magrittr::divide_by(length(yhat)) %>% 
+    return()
+}
+
 # parallel backend
-doMC::registerDoMC(parallel::detectCores())
+doMC::registerDoMC(parallel::detectCores() / 4)
 
 # simulation parameters
-K.vec <- c(2, 4, 8)
+K.vec <- c(2, 3, 4)
 a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
 sparsity <- .01
-n.vec <- c(64, 128, 256, 512, 1024, 2048)
-iter <- 100
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+iter <- 50
+# n.vec <- c(256, 512, 1024)
+# iter <- 10
+n.vec <- rev(n.vec)
+K.vec <- rev(K.vec)
 set.seed(314159)
 
 # clustering simulation
 clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
-  foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
     print(paste('K =', K, ', n =', n))
     foreach(i = seq(iter), 
             .combine = dplyr::bind_rows, 
@@ -40,16 +53,17 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       z <- Pz$clustering
       A <- draw.graph(P)
       clustering <- cluster.pabm(A, K, use.all = TRUE)
-      error <- 1 - fossil::adj.rand.index(z, clustering)
+      error <- 1 - cluster.acc(clustering, z)
       clustering.ssc <- ssc(A, K, sparsity)
-      error.ssc <- 1 - fossil::adj.rand.index(z, clustering.ssc)
+      error.ssc <- 1 - cluster.acc(clustering.ssc, z)
       dplyr::tibble(K = K, n = n, error = error, 
                     error.ssc = error.ssc) %>% 
         return()
     } %>% 
       return()
-  } %>%
-    return()
+  }
+  gc()
+  return(out.df)
 }
 
 clustering.df %>%

### 2020-10-14 functions.R
minor update to functions

@@ -104,6 +104,8 @@ cluster.pabm <- function(A, K,
       eigen(L, symmetric = TRUE)$vectors[, seq(n - 1, n - d.eigenmap + 1)]
   }
   clustering <- mclust::Mclust(eigenmap, K)$classification
+  
+  return(clustering)
 }
 
 rot.mat.2 <- function(angle, axis = 1) {
@@ -311,8 +313,14 @@ lambda.rmse.mle <- function(P, A, z) {
     return()
 }
 
-generate.P.beta <- function(n, K = 2, a1 = 2, b1 = 1, a2 = 1, b2 = 2) {
-  clustering <- sample(seq(K), n, replace = TRUE)
+generate.P.beta <- function(n, K = 2, a1 = 2, b1 = 1, a2 = 1, b2 = 2,
+                            unbalanced = FALSE) {
+  if (unbalanced) {
+    clustering <- rmultinom(n, 1, seq(K) ** -1) %>% 
+      apply(2, function(x) which(x == 1))
+  } else {
+    clustering <- sample(seq(K), n, replace = TRUE)
+  }
   clustering <- sort(clustering)
   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
   

### 2020-10-14 clustering-sim-k-imbalanaced.R
more simulations

@@ -1,0 +1,89 @@
+# simulation using proposed clustering method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, 
+             generate.P.beta,
+             normalized.laplacian,
+             cluster.pabm, ssc,
+             .from = 'functions.R')
+
+cluster.acc <- function(yhat, yobs) {
+  table(yhat, yobs) %>% 
+    as.matrix() %>% 
+    apply(1, max) %>% 
+    sum() %>% 
+    magrittr::divide_by(length(yhat)) %>% 
+    return()
+}
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores() / 4)
+
+# simulation parameters
+K.vec <- c(2, 3, 4)
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- .01
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+iter <- 50
+# n.vec <- c(256, 512, 1024)
+# iter <- 10
+n.vec <- rev(n.vec)
+K.vec <- rev(K.vec)
+set.seed(314159)
+
+# clustering simulation
+clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('K =', K, ', n =', n))
+    foreach(i = seq(iter), 
+            .combine = dplyr::bind_rows, 
+            .errorhandling = 'remove') %dopar% {
+      Pz <- generate.P.beta(n, K, a1, a2, b1, b2, unbalanced = TRUE)
+      P <- Pz$P
+      z <- Pz$clustering
+      A <- draw.graph(P)
+      clustering <- cluster.pabm(A, K, use.all = TRUE)
+      error <- 1 - cluster.acc(clustering, z)
+      clustering.ssc <- ssc(A, K, sparsity)
+      error.ssc <- 1 - cluster.acc(clustering.ssc, z)
+      dplyr::tibble(K = K, n = n, error = error, 
+                    error.ssc = error.ssc) %>% 
+        return()
[…]

### 2020-10-14 summary.Rmd
more simulations

--- 2020-09-30 summary.Rmd
@@
 
-## Case $K = 2$
+## Balanced communities
+
+In each simulation, community labels $z_1, ..., z_n$ were drawn 
+from a multinomial distribution with mixture parameters 
+$\{\alpha_1, ..., \alpha_K\}$, then $\{\lambda^{(kl)}\}_K$ according to the 
+drawn community labels, $P$ was constructed using the drawn 
+$\{\lambda^{(kl)}\}_K$, and $A$ was drawn from $P$ by 
+$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$. Each simulation has a unique
+edge probability matrix $P$.
 
@@
 
-* Number of underlying communities $K = 2$
-* Mixture parameter $\alpha = .5$
+* Number of vertices $n = 256, 512, 1024, 2048, 4096$
+* Number of underlying communities $K = 2, 3, 4$
+* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
+community has an equal probability of being drawn)
+* Community labels 
+$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
 * Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
@@
 
-### Community detection
+$50$ simulations were performed for each $(n, K)$ pair.
 
-In this part, we will assess Algorithm 2's performance for sample sizes 
-$n = 64, 128, 256, 512, 1024, 2048$. For each sample size $n$, $100$ sets of 
-$\lambda^{(kl)}$'s are drawn and for each set of parameters, one adjacency 
-matrix $A$ is drawn and clustered. We will not consider community detection using the 
-edge probability matrix $P$ since this will always result exact recovery of the
-original communities (up to permutation).
-
-```{r clust_err_sim, fig.cap = 'IQR clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048.
-Simulations were repeated 100 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', cache = FALSE}
-clustering.df <- readr::read_csv('clustering.csv')
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+Simulations were repeated 50 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k.csv')
+# mod.max.df <- readr::read_csv('mod-max-k.csv') %>% 
+#   dplyr::group_by(K, n) %>% 
+#   dplyr::summarise(med.err.mm = median(error.mm),
+#                    first.q.mm = quantile(error.mm, .25),
+#                    third.q.mm = quantile(error.mm, .75)) %>% 
+#   dplyr::ungroup()
 clustering.df %>%
-  dplyr::group_by(n) %>%
+  dplyr::group_by(K, n) %>%
   dplyr::summarise(med.err = median(error),
@@
   ggplot() +
-  scale_y_log10() +
   # scale_x_log10() + 
-  labs(y = 'misclustering proportion', 
+  labs(y = 'community detection error', 
        colour = NULL) +
@@
                     colour = 'SSC')) + 
-  scale_colour_brewer(palette = 'Set1')
+  # geom_line(data = mod.max.df,
+  #           aes(x = n, y = med.err.mm,
+  #               colour = 'modularity maximization')) + 
+  # geom_errorbar(data = mod.max.df,
+  #               aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+  #                   colour = 'modularity maximization')) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
+  scale_y_log10()
 ```
@@
 Simulations were repeated 100 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', cache = FALSE}
-clustering.df <- readr::read_csv('clustering.csv')
+The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k.csv')
 clustering.df %>%
-  dplyr::group_by(n) %>%
+  dplyr::group_by(n, K) %>%
   dplyr::summarise(med.err = median(error),
@@
   ggplot() +
-  scale_y_log10() +
   # scale_x_log10() + 
@@
                     colour = 'SSC')) + 
-  scale_colour_brewer(palette = 'Set1')
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10() +
+  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096))
 ```
 
-We can also examine how the distribution of $(\hat{v}_i)^\top \hat{v_j}$ varies 
-with $n$:
-
-```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', cache = FALSE}
-densities.df <- readr::read_csv('densities.csv')
-
-ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
-  labs(x = expression(n * v[i]^T * v[j]),
-       colour = 'n') + 
-  scale_colour_brewer(palette = 'Set1')
-```
-
-### Parameter estimation
-
-Figure \ref{fig:lambda_est} shows the medians and interquartile ranges of root 
-mean square errors for Algorithm 4 over 100 simulations using the same 
-parameters as before. Comparison against an MLE-based method 
-\cite{noroozi2019estimation} suggests similar performance, close to the rate 
-$n^{-1/2}$.
-
-```{r lambda_est, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 2048.
-Simulations were repeated 100 times for each sample size.'}
-rmse.df <- readr::read_csv('rmse.csv')
-
-rmse.df %>% 
-  dplyr::group_by(n) %>% 
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-  dplyr::ungroup() %>% 
-  ggplot() + 
-  scale_y_log10() + 
-  scale_x_log10() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 4')) + 
-  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based')) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  labs(y = 'RMSE', colour = NULL)
-```
-
-## Higher values of $K$
-
-For these examples, we set the following parameters:
-
-* Number of underlying communities $K = 2, 4, 8$
-* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$
-* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
-* Between-group popularities 
-$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
-
-### Community detection
-
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
-Simulations were repeated 100 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', fig.width = 8}
-clustering.df <- readr::read_csv('clustering-k.csv')
-clustering.df %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75)) %>%
-  dplyr::ungroup() %>% 
-  ggplot() +
-  scale_y_log10() +
-  # scale_x_log10() + 
-  labs(y = '1 - adjusted rand index', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'Algorithm 2')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 2')) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC')) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both')
-```
-
-### Parameter estimation
-
 ```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 2048 and number of clusters from 2 to 8.
-Simulations were repeated 100 times for each sample size.', fig.width = 8}
+Simulations were repeated 50 times for each sample size.', fig.width = 10}
 rmse.df <- readr::read_csv('rmse.csv') %>% 
@@
   ggplot() +
-  scale_y_log10() + 
-  scale_x_log10() +
   geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
@@
   labs(y = 'RMSE', colour = NULL) + 
-  facet_wrap(~ K, labeller = 'label_both')
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10() + 
+  scale_x_log10(breaks = c(64, 128, 256, 512, 1024, 2048))
+```
+
+## Imbalanced communities
+
+Simulations performed in this section are similar to those in the previous 
+section with the exception of the mixture parameters 
+$\{\alpha_1, ..., \alpha_K\}$ used to draw community labels from the 
+multinomial distribution. For these examples, we set the following parameters:
+
+* Number of vertices $n = 256, 512, 1024, 2048, 4096$
+* Number of underlying communities $K = 2, 3, 4$
+* Mixture parameters $\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}$ for 
+$k = 1, ..., K$
+* Community labels 
+$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
+* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+* Between-group popularities 
+$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
+
+$50$ simulations were performed for each $(n, K)$ pair.
+
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+Simulations were repeated 50 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
+clustering.df %>%
+  dplyr::group_by(K, n) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75)) %>%
+  dplyr::ungroup() %>% 
+  ggplot() +
+  # scale_x_log10() + 
+  labs(y = 'community detection error', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'Algorithm 2')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'Algorithm 2')) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC')) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
+  scale_y_log10()
+```
+
+## Replication and comparison of simulations from \citet*{307cbeb9b1be48299388437423d94bf1}
+
+```{r clust_err_sc, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) for sample size equal to 400.
+Simulations were repeated 50 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', fig.height = 2, fig.width = 6}
+clustering.mm.df <- readr::read_csv('clustering-sim-sc.csv')
+clustering.ase.df <- readr::read_csv('clustering-sim-sc-ase.csv')
+clustering.df <- dplyr::bind_rows(
+  clustering.mm.df %>% 
+    dplyr::transmute(method = 'Modularity maximization',
+                     homophily = homophily,
+                     error = pa.err),
+  clustering.ase.df %>% 
+    dplyr::transmute(method = 'Algorithm 2',
+                     homophily = homophily,
+                     error = pa.err)
+)
+clustering.df %>%
+  dplyr::group_by(method, homophily) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75)) %>%
+  dplyr::ungroup() %>% 
+  ggplot() +
+  labs(y = 'community detection \nerror', 
+       colour = NULL) +
+  geom_line(aes(x = homophily, y = med.err, colour = method)) +
+  geom_errorbar(aes(x = homophily, ymin = first.q, ymax = third.q,
+                    colour = method)) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  # coord_fixed(1) + 
+  scale_y_log10()
+```
+
+## Additional experiments
+
+Using the same set of parameters for generating $P$ and $A$ as in the 
+balanced communities examples for $K = 2$, we generated one instance of $A$ for
+each $n$ and constructed $B$ according to algorithm 4 to verify that 
+as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
+in different clusters. Furthermore, the distribution of these inner products 
+should be approximately normal.
+
+```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', fig.height = 2, fig.width = 4}
+densities.df <- readr::read_csv('densities.csv')
+
+ggplot(densities.df) + 
+  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
+  labs(x = expression(n * v[i]^T * v[j]),
+       colour = 'n') + 
+  scale_x_continuous(breaks = seq(-5, 5)) + 
+  scale_colour_brewer(palette = 'Set1')
 ```
@@
 
-In the third example,
+In the third example, we consider the Karantaka villages data studied by 
+\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` network from 
+village 12 at the household level. The label of interest is the religious 
+affiliation. The network was truncated to religions "1" and "2", and vertices 
+of degree 0 were removed. The resulting network has 141 vertices.
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
-hh <- readRDS('data/village-hh-59.rds')
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left), sparse subspace clustering (middle), and modularity maximization (right).', fig.height = 2, fig.width = 8, eval = TRUE}
+hh <- readRDS('data/village-hh-12.rds')
 A.hh <- hh$A
 z.hh <- hh$z
-K <- 6
+K <- 2
 zhat.hh <- cluster.pabm(A.hh, K,
-                        use.all = TRUE, normalize = FALSE, p = K, q = 0)
+                        use.all = TRUE, normalize = FALSE, p = 3, q = 1)
+zhat.hh.ssc <- ssc(A.hh, K, lambda = .1)
 
-zhat.mres <- readRDS('data/mres-result.rds')
+# this part of code taken from sengupta & chen
+setwd('~/dev/pabm-grdpg/codes_and_data')
+require(MASS);require("clue")
+source("EPalgosim.R") 
+source("functions.R")
+b.can = EPalgo(A.hh,eps=0) # EP algorithm (no perturbation)
+Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
+for (i in 1:ncol(b.can)){
+  #check if any cluster is empty
+  foo = rep(NA, max(z.hh))
+  for (clus in 1:max(z.hh)) {foo[clus]=sum(b.can[,i]==clus)}
+  if (min(foo)==0) {stop('Empty groups are not allowed')} 
+  Q.PA.can[i] = Q.PA(A.hh, b=b.can[,i])   # fit PABM
+} # end of i for loop
+foo1 = order(-Q.PA.can)[1] 
+zhat.hh.mm = b.can[,foo1]   # community assignment that maximises Q.PA
 
 gridExtra::grid.arrange(plot.A(A.hh, zhat.hh),
-                        plot.A(A.hh, zhat.mres),
-                        ncol = 2)
+                        plot.A(A.hh, zhat.hh.ssc),
+                        plot.A(A.hh, zhat.hh.mm),
+                        nrow = 1)
+```
+
+```{r, eval = TRUE}
+hh.acc <- mean(zhat.hh == z.hh)
+hh.acc <- ifelse(hh.acc < .5, 1 - hh.acc, hh.acc)
+hh.acc.ssc <- mean(zhat.hh.ssc == z.hh)
+hh.acc.ssc <- ifelse(hh.acc.ssc < .5, 1 - hh.acc.ssc, hh.acc.ssc)
+hh.acc.mm <- mean(zhat.hh.mm == z.hh)
+hh.acc.mm <- ifelse(hh.acc.mm < .5, 1 - hh.acc.mm, hh.acc.mm)
+
+knitr::kable(data.frame(method = c('Algorithm 2', 'SSC', 
+                                   'Modularity maximization'),
+                        error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
+             caption = paste('Community detection error rates for identifying',
+                             'household religion.'))
 ```

### 2020-10-14 summary.Rmd
typo
updated descriptions/captions

@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.
@@
 
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 256 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048 and number of clusters from 2 to 8.
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.
@@
 networks. Our methods underperformed compared to modularity maximization, 
-although performance is comparable for the British MPs and DBLP networks. In 
+although performance is comparable. In 
 addition, Algorithm 2's runtime is much lower than that of modularity 

### 2020-10-20 summary.Rmd
temp commit

@@
 
+In these simulations, Algorithm 2 is compared against sparse subspace clustering \cite{soltanolkotabi2014} \cite{noroozi2019estimation} and modularity maximization \cite{307cbeb9b1be48299388437423d94bf1}, and Algorithm 3 is compared against an
+MLE-based estimator \cite{noroozi2019estimation} and modularity maximization 
+\cite{307cbeb9b1be48299388437423d94bf1}.
+
 ## Balanced communities

### 2020-10-21 clustering-sim-k-imbalanaced.R
more simulations

@@ -14,6 +14,12 @@ import::here(embedding, draw.graph,
              cluster.pabm, ssc,
              .from = 'functions.R')
 
+mod.max <- function(A) {
+  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
+  clustering <- cluster_louvain(Ag)
+  return(membership(clustering))
+}
+
 cluster.acc <- function(yhat, yobs) {
   table(yhat, yobs) %>% 
     as.matrix() %>% 
@@ -24,7 +30,7 @@ cluster.acc <- function(yhat, yobs) {
 }
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores() / 4)
+doMC::registerDoMC(20)
 
 # simulation parameters
 K.vec <- c(2, 3, 4)
@@ -32,9 +38,9 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
-sparsity <- .01
+sparsity <- 1e-4
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-iter <- 50
+iter <- 40
 # n.vec <- c(256, 512, 1024)
 # iter <- 10
 n.vec <- rev(n.vec)

### 2020-10-21 clustering-sim-k.R

@@ -6,6 +6,7 @@ import::from(magrittr, `%>%`)
 import::from(foreach, foreach, `%do%`, `%dopar%`)
 library(mclust)
 library(ggplot2)
+library(igraph)
 source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
 setwd('~/dev/pabm-grdpg')
 import::here(embedding, draw.graph, 
@@ -13,6 +14,12 @@ import::here(embedding, draw.graph,
              normalized.laplacian,
              cluster.pabm, ssc,
              .from = 'functions.R')
+
+mod.max <- function(A) {
+  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
+  clustering <- cluster_louvain(Ag)
+  return(membership(clustering))
+}
 
 cluster.acc <- function(yhat, yobs) {
   table(yhat, yobs) %>% 
@@ -24,7 +31,7 @@ cluster.acc <- function(yhat, yobs) {
 }
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores() / 4)
+doMC::registerDoMC(parallel::detectCores() / 2)
 
 # simulation parameters
 K.vec <- c(2, 3, 4)
@@ -32,9 +39,9 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
-sparsity <- .01
+sparsity <- 1e-4
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-iter <- 50
+iter <- 40
 # n.vec <- c(256, 512, 1024)
 # iter <- 10
 n.vec <- rev(n.vec)
@@ -56,8 +63,12 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       error <- 1 - cluster.acc(clustering, z)
       clustering.ssc <- ssc(A, K, sparsity)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
-      dplyr::tibble(K = K, n = n, error = error, 
-                    error.ssc = error.ssc) %>% 
+      clustering.mm <- mod.max(A)
+      error.mm <- 1 - cluster.acc(clustering.mm, z)
+      dplyr::tibble(K = K, n = n, 
+                    error = error, 
+                    error.ssc = error.ssc,
+                    error.mm = error.mm) %>% 
         return()
     } %>% 
       return()
@@ -66,6 +77,8 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
[…]

### 2020-10-21 clustering-sim-k-imbalanaced.R
more simulations

@@ -62,8 +62,12 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       error <- 1 - cluster.acc(clustering, z)
       clustering.ssc <- ssc(A, K, sparsity)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
-      dplyr::tibble(K = K, n = n, error = error, 
-                    error.ssc = error.ssc) %>% 
+      clustering.mm <- mod.max(A)
+      error.mm <- 1 - cluster.acc(clustering.mm, z)
+      dplyr::tibble(K = K, n = n, 
+                    error = error, 
+                    error.ssc = error.ssc,
+                    error.mm = error.mm) %>% 
         return()
     } %>% 
       return()
@@ -72,6 +76,8 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   return(out.df)
 }
 
+gc()
+
 clustering.df %>%
   dplyr::group_by(K, n) %>%
   dplyr::summarise(med.err = median(error),
@@ -79,7 +85,10 @@ clustering.df %>%
                    third.q = quantile(error, .75),
                    med.err.ssc = median(error.ssc),
                    first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75)) %>%
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
   ggplot() +
   scale_y_log10() +
   # scale_x_log10() +
@@ -89,6 +98,9 @@ clustering.df %>%
   geom_line(aes(x = n, y = med.err.ssc, colour = 'ssc')) +
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                     colour = 'ssc')) + 
+  geom_line(aes(x = n, y = med.err.mm, colour = 'mm')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'mm')) + 
   facet_wrap(~ K)
 
 # export as csv

### 2020-10-21 summary.Rmd
ore simulations

--- 2020-10-14 summary.Rmd
@@
 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
@@
 
+In these simulations, Algorithm 2 is compared against sparse subspace 
+clustering \cite{soltanolkotabi2014} \cite{noroozi2019estimation} and 
+modularity maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}, 
+and Algorithm 4 is compared against an MLE-based estimator 
+\cite{noroozi2019estimation} \cite{307cbeb9b1be48299388437423d94bf1}.
+
 ## Balanced communities
@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4.
-Simulations were repeated 50 times for each sample size.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 256 to 4096 and number of underlying communities from 2 to 4.
+Simulations were repeated 40 times for each sample size.
 The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k.csv')
-# mod.max.df <- readr::read_csv('mod-max-k.csv') %>% 
-#   dplyr::group_by(K, n) %>% 
-#   dplyr::summarise(med.err.mm = median(error.mm),
-#                    first.q.mm = quantile(error.mm, .25),
-#                    third.q.mm = quantile(error.mm, .75)) %>% 
-#   dplyr::ungroup()
 clustering.df %>%
@@
                    first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75)) %>%
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
   dplyr::ungroup() %>% 
@@
                     colour = 'SSC')) + 
-  # geom_line(data = mod.max.df,
-  #           aes(x = n, y = med.err.mm,
-  #               colour = 'modularity maximization')) + 
-  # geom_errorbar(data = mod.max.df,
-  #               aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-  #                   colour = 'modularity maximization')) + 
+  geom_line(aes(x = n, y = med.err.mm,
+                colour = 'modularity maximization')) +
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'modularity maximization')) +
   scale_colour_brewer(palette = 'Set1') + 
@@
 
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 64 to 2048.
-Simulations were repeated 100 times for each sample size.
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
+Simulations were repeated 40 times for each sample size.
 The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 256 to 4096 and number of clusters from 2 to 4.
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.

### 2020-10-21 functions.R

@@ -369,7 +369,7 @@ plot.A <- function(A, z, lines = TRUE, max.size = 500) {
     geom_tile(aes(x = X, y = Y, fill = Z)) + 
     coord_fixed() + 
     labs(x = NULL, y = NULL) + 
-    theme_void() +
+    # theme_void() +
     theme(legend.position = 'none',
           axis.title.x = element_blank(),
           axis.text.x = element_blank(),
@@ -379,14 +379,18 @@ plot.A <- function(A, z, lines = TRUE, max.size = 500) {
           axis.ticks.y = element_blank()) +
     scale_x_continuous(expand = c(0, 0)) + 
     scale_y_continuous(expand = c(0, 0)) + 
-    scale_fill_gradient(low = 'red', high = 'yellow')
+    scale_fill_gradient(low = 'white', high = 'black')
   
   if (lines) {
     positions.x <- which(diff(z) != 0)
     positions.y <- which(diff(rev(z)) != 0)
     out.plot <- out.plot + 
-      geom_hline(yintercept = positions.y + .5) + 
-      geom_vline(xintercept = positions.x + .5)
+      geom_hline(yintercept = positions.y + .5, 
+                 colour = 'red',
+                 size = .1) + 
+      geom_vline(xintercept = positions.x + .5, 
+                 colour = 'red',
+                 size = .1)
   }
   
   return(out.plot)

### 2020-10-21 clustering-sim-k-imbalanaced.R
more simulations

@@ -38,7 +38,7 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
-sparsity <- 1e-4
+sparsity <- 1e-3
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
 iter <- 40
 # n.vec <- c(256, 512, 1024)

### 2020-10-21 summary.Rmd
more simulations

@@
 * Community labels 
-$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
+$z_k \stackrel{iid}{\sim} Multinomial(\alpha_1, ..., \alpha_K)$
 * Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
@@
   scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
-  scale_y_log10()
-```
-
-## Replication and comparison of simulations from \citet*{307cbeb9b1be48299388437423d94bf1}
-
-```{r clust_err_sc, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) for sample size equal to 400.
-Simulations were repeated 50 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', fig.height = 2, fig.width = 6}
-clustering.mm.df <- readr::read_csv('clustering-sim-sc.csv')
-clustering.ase.df <- readr::read_csv('clustering-sim-sc-ase.csv')
-clustering.df <- dplyr::bind_rows(
-  clustering.mm.df %>% 
-    dplyr::transmute(method = 'Modularity maximization',
-                     homophily = homophily,
-                     error = pa.err),
-  clustering.ase.df %>% 
-    dplyr::transmute(method = 'Algorithm 2',
-                     homophily = homophily,
-                     error = pa.err)
-)
-clustering.df %>%
-  dplyr::group_by(method, homophily) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75)) %>%
-  dplyr::ungroup() %>% 
-  ggplot() +
-  labs(y = 'community detection \nerror', 
-       colour = NULL) +
-  geom_line(aes(x = homophily, y = med.err, colour = method)) +
-  geom_errorbar(aes(x = homophily, ymin = first.q, ymax = third.q,
-                    colour = method)) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  # coord_fixed(1) + 
   scale_y_log10()
@@
 dblp.acc <- ifelse(dblp.acc < .5, 1 - dblp.acc, dblp.acc)
+
+zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3)
+bmp.acc.ssc <- mean(zhat.bmp.ssc == z.bmp)
+bmp.acc.ssc <- ifelse(bmp.acc.ssc < .5, 1 - bmp.acc.ssc, bmp.acc.ssc)
+
+zhat.pb.ssc <- ssc(A.pb, 2, 1e-4)
+pb.acc.ssc <- mean(zhat.pb.ssc == z.pb)
+pb.acc.ssc <- ifelse(pb.acc.ssc < .5, 1 - pb.acc.ssc, pb.acc.ssc)
+
+zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2)
+dblp.acc.ssc <- mean(zhat.dblp.ssc == z.dblp)
+dblp.acc.ssc <- ifelse(dblp.acc.ssc < .5, 1 - dblp.acc.ssc, dblp.acc.ssc)
 ```
@@
               vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
-              `error (modularity maximization)` = c(round(1 / 329, 3),
+              `error (mod. max.)` = c(round(1 / 329, 3),
                                       round(61 / 1222, 3),
                                       round(62 / 2203, 3)),
-              `error (algorithm 2)` = round(1 - c(bmp.acc, 
+              `error (SSC)` = round(1 - c(bmp.acc.ssc,
+                                          pb.acc.ssc,
+                                          dblp.acc.ssc), 3),
+              `error (alg. 2)` = round(1 - c(bmp.acc, 
                                                   pb.acc,
@@
   knitr::kable(caption = paste('Community detection error rates', 
-                               'for modularity maximization',
-                               '(\\citet{307cbeb9b1be48299388437423d94bf1})',
+                               'for modularity maximization,',
+                               'sparse subspace clustering, ',
                                'and Algorithm 2.'))

### 2020-10-21 clustering-sim-k-imbalanaced.R
more simulations

@@ -5,6 +5,7 @@
 import::from(magrittr, `%>%`)
 import::from(foreach, foreach, `%do%`, `%dopar%`)
 library(mclust)
+library(igraph)
 library(ggplot2)
 source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
 setwd('~/dev/pabm-grdpg')

### 2020-10-21 rmse-sim-k.R
more simulations

@@ -10,15 +10,16 @@ import::here(embedding, draw.graph, I.pq, generate.P.beta,
              .from = 'functions.R')
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores())
+doMC::registerDoMC(parallel::detectCores() / 2)
 
 # simulation params
-K.vec <- c(4, 8)
+K.vec <- c(4, 3, 2)
 alpha <- .5
 a1 <- b2 <- 2
 a2 <- b1 <- 1
-n.vec <- c(64, 128, 256, 512, 1024, 2048)
-iter <- 100
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+n.vec <- rev(n.vec)
+iter <- 40
 set.seed(314159)
 
 # simulation
@@ -27,7 +28,8 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   q <- K * (K - 1) / 2
   Ipq <- I.pq(p, q)
   
-  foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('n =', n, 'K =', K))
     plyr::ldply(seq(iter), function(i) {
       P.z <- generate.P.beta(n, K)
       P <- P.z$P
@@ -41,31 +43,32 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
         return()
     }, .parallel = TRUE) %>% 
       return()
-  } %>%
-    return()
+  }
+  gc()
+  return(out)
 }
 
-# rmse.df %>% 
-#   na.omit() %>% 
-#   dplyr::group_by(K, n) %>% 
-#   dplyr::summarise(median.rmse = median(rmse),
-#                    q1.rmse = quantile(rmse, .25),
-#                    q3.rmse = quantile(rmse, .75),
-#                    median.rmse.mle = median(rmse.mle),
-#                    q1.rmse.mle = quantile(rmse.mle, .25),
-#                    q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-#   dplyr::ungroup() %>% 
-#   ggplot() + 
-#   scale_y_log10() + 
-#   scale_x_log10() +
-#   geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
-#   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
[…]

### 2020-10-21 summary.Rmd

@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 256 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
 Simulations were repeated 40 times for each sample size.
@@
   facet_wrap(~ K, labeller = 'label_both') + 
-  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
+  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
   scale_y_log10()
@@
   scale_y_log10() +
-  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096))
+  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096))
 ```
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and subspace clustering (green) for sample sizes from 128 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.
@@
                     colour = 'SSC')) + 
+  geom_line(aes(x = n, y = med.err.mm, colour = 'modularity maximization')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm, 
+                    colour = 'modularity maximization')) + 
   scale_colour_brewer(palette = 'Set1') + 
   facet_wrap(~ K, labeller = 'label_both') + 
-  scale_x_continuous(breaks = c(256, 512, 1024, 2048, 4096)) +
+  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
   scale_y_log10()

### 2020-10-21 rmse-sim-k.R
more simulations

@@ -4,6 +4,7 @@
 # packgaes, functions, etc.
 import::from(magrittr, `%>%`)
 import::from(foreach, foreach, `%do%`)
+library(ggplot2)
 setwd('~/dev/pabm-grdpg')
 import::here(embedding, draw.graph, I.pq, generate.P.beta,
              estimate.lambda.block, lambda.rmse, lambda.rmse.mle,

### 2020-10-21 summary.Rmd
more simulations

@@
     Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
-    Assign $\sigma^2 \leftarrow \Sigma_{11}$.\;
+    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
     Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
@@
   geom_line(aes(x = n, y = med.err.mm,
-                colour = 'modularity maximization')) +
+                colour = 'Modularity maximization')) +
   geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'modularity maximization')) +
+                    colour = 'Modularity maximization')) +
   scale_colour_brewer(palette = 'Set1') + 
@@
                     colour = 'SSC')) + 
+  geom_line(aes(x = n, y = med.err.mm * n,
+                colour = 'Modularity maximization')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
+                    colour = 'Modularity maximization')) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
 
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 256 to 4096 and number of clusters from 2 to 4.
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
-rmse.df <- readr::read_csv('rmse.csv') %>% 
-  dplyr::transmute(K = 2,
-                   n = n,
-                   rmse = rmse,
-                   rmse.mle = rmse.mle) %>% 
-  dplyr::bind_rows(readr::read_csv('rmse-k.csv'))
+rmse.df <- readr::read_csv('rmse-k.csv')
 
@@
   scale_y_log10() + 
-  scale_x_log10(breaks = c(64, 128, 256, 512, 1024, 2048))
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
 ```

### 2020-10-21 rmse-sim-k-imbalanced.R
more simulations

@@ -1,0 +1,75 @@
+# simulation using proposed parameter estimation method
+# run on IU RED (not a good idea to attempt on laptop)
+
+# packgaes, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+import::here(embedding, draw.graph, I.pq, generate.P.beta,
+             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
+             .from = 'functions.R')
+
+# parallel backend
+doMC::registerDoMC(parallel::detectCores() / 2)
+
+# simulation params
+K.vec <- c(4, 3, 2)
+alpha <- .5
+a1 <- b2 <- 2
+a2 <- b1 <- 1
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+n.vec <- rev(n.vec)
+iter <- 40
+set.seed(314159)
+
+# simulation
+rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  p <- K * (K + 1) / 2
+  q <- K * (K - 1) / 2
+  Ipq <- I.pq(p, q)
+  
+  out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('n =', n, 'K =', K))
+    plyr::ldply(seq(iter), function(i) {
+      P.z <- generate.P.beta(n, K, a1, a2, b1, b2, unbalanced = TRUE)
+      P <- P.z$P
+      z <- P.z$clustering
+      A <- draw.graph(P)
+      Zhat <- embedding(A, p, q)
+      Phat <- Zhat %*% Ipq %*% t(Zhat)
+      rmse <- lambda.rmse(P, Phat, z)
+      rmse.mle <- lambda.rmse.mle(P, A, z)
+      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
+        return()
+    }, .parallel = TRUE) %>% 
+      return()
+  }
+  gc()
+  return(out)
+}
+
+rmse.df %>%
+  na.omit() %>%
+  dplyr::group_by(K, n) %>%
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
[…]

### 2020-10-21 clustering-sim-k.R
more simulations

@@ -31,7 +31,7 @@ cluster.acc <- function(yhat, yobs) {
 }
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores() / 2)
+doMC::registerDoMC(20)
 
 # simulation parameters
 K.vec <- c(2, 3, 4)
@@ -39,9 +39,9 @@ a1 <- 2
 b1 <- 1
 a2 <- 1
 b2 <- 2
-sparsity <- 1e-4
+sparsity <- 5e-4
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-iter <- 40
+iter <- 50
 # n.vec <- c(256, 512, 1024)
 # iter <- 10
 n.vec <- rev(n.vec)

### 2020-10-21 summary.Rmd
more simulations

@@
 * Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
-community has an equal probability of being drawn)
+community label has an equal probability of being drawn)
 * Community labels 
-$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
+$z_k \stackrel{iid}{\sim} Multinomial(\alpha_1, ..., \alpha_K)$
 * Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
@@
 ```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
-Simulations were repeated 40 times for each sample size.
+Simulations were repeated 50 times for each sample size.
 The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
@@
   ggplot() +
-  # scale_x_log10() + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
   labs(y = 'community detection error', 
@@
   geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 2')) + 
+                    colour = 'Algorithm 2'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc,
@@
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC')) + 
+                    colour = 'SSC'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm,
@@
   geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'Modularity maximization')) +
+                    colour = 'Modularity maximization'), width = .1) +
   scale_colour_brewer(palette = 'Set1') + 
   facet_wrap(~ K, labeller = 'label_both') +
-  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
   scale_y_log10()
@@
 ```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 40 times for each sample size.
+Simulations were repeated 50 times for each sample size.
 The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
@@
   ggplot() +
-  # scale_x_log10() + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
   labs(y = 'misclustering count', 
@@
   geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'Algorithm 2')) + 
+                    colour = 'Algorithm 2'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc * n,
@@
   geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC')) + 
+                    colour = 'SSC'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm * n,
@@
   geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
-                    colour = 'Modularity maximization')) + 
+                    colour = 'Modularity maximization'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
   facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10() +
-  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096))
+  scale_y_log10()
 ```
@@
   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 4')) + 
+                    colour = 'Algorithm 4'), width = .1) + 
   geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
   geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based')) + 
+                   colour = 'MLE-based'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC')) + 
-  geom_line(aes(x = n, y = med.err.mm, colour = 'modularity maximization')) + 
+                    colour = 'SSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.mm, colour = 'Modularity maximization')) + 
   geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm, 
-                    colour = 'modularity maximization')) + 
+                    colour = 'Modularity maximization'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
   facet_wrap(~ K, labeller = 'label_both') +
-  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
   scale_y_log10()
+```
+
+```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
+Simulations were repeated 50 times for each sample size.
+The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
+  dplyr::ungroup() %>% 
+  ggplot() +
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  labs(y = 'misclustering count', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err * n,
+                colour = 'Algorithm 2')) +
+  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
+                    colour = 'Algorithm 2'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc * n,
+                colour = 'SSC')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
+                    colour = 'SSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.mm * n,
+                colour = 'Modularity maximization')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
+                    colour = 'Modularity maximization'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10()
+```
+
+```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+Simulations were repeated 50 times for each sample size.', fig.width = 10}
+rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 4'), width = .1) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
+                   colour = 'MLE-based'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(y = 'RMSE', colour = NULL) + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10() + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
 ```
@@
 ```{r}
-dplyr::tibble(network = c('British MPs', 'Political blogs', 'DBLP'),
-              vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
-              `error (mod. max.)` = c(round(1 / 329, 3),
+dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
+              Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
+              `Error (Mod. Max.)` = c(round(1 / 329, 3),
                                       round(61 / 1222, 3),
                                       round(62 / 2203, 3)),
-              `error (SSC)` = round(1 - c(bmp.acc.ssc,
+              `Error (SSC)` = round(1 - c(bmp.acc.ssc,
                                           pb.acc.ssc,
                                           dblp.acc.ssc), 3),
-              `error (alg. 2)` = round(1 - c(bmp.acc, 
+              `Error (Alg. 2)` = round(1 - c(bmp.acc, 
                                                   pb.acc,
@@
                                'sparse subspace clustering, ',
-                               'and Algorithm 2.'))
+                               'and Alg. 2.'))
 ```
@@
 
-knitr::kable(data.frame(method = c('Algorithm 2', 'SSC', 
-                                   'Modularity maximization'),
-                        error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
+knitr::kable(data.frame(Method = c('Alg. 2', 'SSC', 
+                                   'Mod. Max.'),
+                        Error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
              caption = paste('Community detection error rates for identifying',

### 2020-10-22 clustering-sim-k-imbalanaced.R
more simulations

@@ -31,7 +31,7 @@ cluster.acc <- function(yhat, yobs) {
 }
 
 # parallel backend
-doMC::registerDoMC(20)
+doMC::registerDoMC(18)
 
 # simulation parameters
 K.vec <- c(2, 3, 4)
@@ -41,7 +41,7 @@ a2 <- 1
 b2 <- 2
 sparsity <- 1e-3
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-iter <- 40
+iter <- 50
 # n.vec <- c(256, 512, 1024)
 # iter <- 10
 n.vec <- rev(n.vec)

### 2020-10-22 rmse-sim-k-imbalanced.R
more simulations

@@ -20,7 +20,7 @@ a1 <- b2 <- 2
 a2 <- b1 <- 1
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
 n.vec <- rev(n.vec)
-iter <- 40
+iter <- 50
 set.seed(314159)
 
 # simulation

### 2020-10-28 summary.Rmd
updated writeup

@@
   (GRDPG) and use properties established in this connection to aid in 
-  clustering and parameter estimation. In particular, we note that the PABM 
-  can be represented as latent positions such that points within the same 
-  cluster lie on a subspace, and the subspaces that represent each 
-  cluster are orthogonal to one another. Using this property as well as the 
+  community detection and parameter estimation. In particular, we note that the
+  PABM can be represented as latent positions such that points within the same 
+  community lie on a subspace, and the subspaces that represent each 
+  community are orthogonal to one another. Using this property as well as the 
   asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
-  are able to establish theoretical asymptotic results of our clustering and 
-  parameter estimation methods for the PABM. 
+  are able to establish theoretical asymptotic results of our community 
+  detection and parameter estimation methods for the PABM. 
 ---
@@
 \newcommand{\blockdiag}{\text{blockdiag}}
+\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
+\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
+\newcommand{\Bernoulli}{\text{Bernoulli}}
+\newcommand{\Betadist}{\text{Beta}}
 
@@
 possible interest), assuming some probabilistic graph model. One such model is 
-the Stochastic Block Model (SBM), which assumes that the edge probability from 
-one vertex to another follows a Beronulli distribution with fixed probabilities 
-for each pair of community labels. The Popularity Adjusted Block Model (PABM) 
-was then introduced by \citet*{307cbeb9b1be48299388437423d94bf1} as a 
-generalization of the SBM to address the heterogeneity of edge probabilities 
-within and between communities while still maintaining the community structure.
-
-The Random Dot Product Graph (RDPG) model \cite{athreya2017statistical} is 
-another graph model with Beronulli edge probabilities. Under this model, each 
-vertex of the graph can be represented by a point in a latent Euclidean space 
-such that the edge probability between any pair of vertices is given by their 
-corresponding dot product in the latent space. The SBM is equivalent to a 
-special case of the RDPG model in which all vertices of a given community share 
-the same position in the latent space. It has also been shown that similar 
-probabilistic graph models, such as the Mixed Membership Stochastic Block 
-Model, can be represented in this way \cite{rubindelanchy2017statistical}. An 
-analogous property exists for the PABM, not under the RDPG model but under the 
+the Stochastic Block Model (SBM), first proposed by 
+\citet{doi:10.1080/0022250X.1971.9989788}, which assumes that the edge 
+probability from one vertex to another follows a Beronulli distribution with 
+fixed probabilities for each pair of community labels. Other probabilistic 
+graph models have been proposed and studied, such as the Degree-Corrected Block 
+Model (DCBM), introduced by \citet{Karrer_2011} and the Mixed Membership 
+Stochastic Block Model (MMSBM), introduced by  \citet*{NIPS2008_3578}. The 
+Popularity Adjusted Block Model (PABM) was then introduced by 
+\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
+address the heterogeneity of edge probabilities within and between communities 
+while still maintaining distinct community structure.
+
+The underlying similarity among the SBM, PABM, and other such models is that 
+they involve a symmetric edge probability matrix $P \in [0, 1]^{n \times n}$ 
+where $n$ is the number of vertices in the graph. An undirected and unweighted 
+graph is then drawn from this edge probability matrix such that the existence 
+of an edge between each pair of vertices $i$ and $j$ is given by 
+$\Bernoulli(P_{ij})$. For example, for the SBM with two communities for which 
+the within-community edge probability is $\xi$ and the between-community edge 
+probability is $\eta$, the entries of $P$ consist of $\xi$ and $\eta$.
+
+The Random Dot Product Graph (RDPG) model proposed by 
+\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli edge 
+probabilities. Under this model, each vertex of the graph can be represented by 
+a point in a latent Euclidean space such that the edge probability between any 
+pair of vertices is given by their corresponding dot product in the latent 
+space, i.e., given a latent positions $x_1, ..., x_n \in \mathbb{R}^d$, the 
+edge probability matrix is $P = X X^\top$ where 
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. The SBM is 
+equivalent to a special case of the RDPG model in which all vertices of a 
+given community share the same position in the latent space 
+\cite{lyzinski2015perfect}. It has also been shown that similar 
+probabilistic graph models, including both the MMSBM and DCBM, can be 
+represented in this way \cite{rubindelanchy2017consistency} 
+\cite{lyzinski2015perfect}. An analogous 
+property exists for the PABM but not for the RDPG model but under the 
 *Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
@@
 unweighted graphs with binary ($0, 1$) edge weights We will also assume that 
-graphs are hollow, i.e., there are no edges from a vertex to itself.
+graphs are hollow, i.e., there are no edges from a vertex to itself. All such 
+graphs can be represented by a symmetric adjacency matrix 
+$A \in \{0, 1\}^{n \times n}$ for which $A_{ij} = 1$ if there exists an edge 
+between vertices $i$ and $j$ and $0$ otherwise, and $A$ is an element-wise 
+independent Bernoulli draw from a symmetric edge probability matrix 
+$P \in [0, 1]^{n \times n}$.
 
@@
 
-## The popularity adjusted block model (PABM) \cite{307cbeb9b1be48299388437423d94bf1} and the generalized random dot product graph \cite{rubindelanchy2017statistical}
-
-**Definition 1**. 
+## The popularity adjusted block model (PABM) and the generalized random dot product graph
+
+**Definition 1** (Popularity Adjusted Block Model) 
+\cite{307cbeb9b1be48299388437423d94bf1}. 
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
@@
 columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
-$P^{(kl})$ describes the edge probabilities between vertices in communities 
-$k$ and $l$. 
+$P^{(kl)}$ describes the edge probabilities between vertices in communities 
+$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
 Let graph $G = (V, E)$ be an undirected, unweighted graph such 
 that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
-realization of $Bernoulli(P)$, i.e., 
-$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ 
+realization of $\Bernoulli(P)$, i.e., 
+$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ 
 ($A_{ij} = A_{ji}$ and $A_{ii} = 0$).
 
-If each block $P^{(kl)}$ can be written as the outer product of two vectors 
-$\lambda^{(kl)} (\lambda^{(lk)})^\top$ for a set of fixed vectors 
-$\{\lambda^{(st)}\}_{s, t = 1}^K$ where each $\lambda^{(st)}$ is a column vector 
+If each block $P^{(kl)}$ can be written as the outer product of two vectors:
+
+\begin{equation}
+  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
+\end{equation}
+
+for a set of $K^2$ fixed vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each 
+$\lambda^{(st)}$ is a column vector 
 of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
@@
 
-**Definition 2**.
+**Definition 2** (Generalized Random Dot Product Graph) 
+\cite{rubindelanchy2017statistical}.
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
 set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that
-$P = X I_{pq} X^\top$ for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
-graph $G = (V, E)$ with adjacency matrix $A$ such that $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and $A_{ii} = 0$) is a draw from the generalized random dot product graph model with latent positions $X$ and signature $(p, q)$.
-More precisely, if vertices $i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains the latent positions as rows $x_i^\top$.
+
+\begin{equation}
+  P = X I_{pq} X^\top
+\end{equation}
+
+for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
+graph $G = (V, E)$ with adjacency matrix $A$ such that 
+$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
+$A_{ii} = 0$) is a draw from the generalized random dot product graph model 
+with latent positions $X$ and signature $(p, q)$. More precisely, if vertices 
+$i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge 
+probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains 
+the latent positions as rows $x_i^\top$.
 
@@
 x_i^\top I_{pq} x_j = P_{ij}$. 
-Unlike in the RDPG case, transforming the latent positions by multiplication 
-with $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
+Unlike in the RDPG case, transforming the latent positions via multiplication 
+by $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
 distances.
@@
 
-### Case where $K = 2$
-
-**Theorem 1**. 
-Let $X = \begin{bmatrix} 
+**Theorem 1** (Connecting the PABM to the GRDPG for $K = 2$). 
+Let
+
+$$X = \begin{bmatrix} 
 \lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
 0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
-\end{bmatrix}$, and let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
+\end{bmatrix}$$
+
+$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
 0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
 0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}$, as in Definition 1. 
+0 & 1 & 0 & 0 \end{bmatrix}$$
+
+as in Definition 1. 
 Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
@@
 *Proof*. 
-Let $X = \begin{bmatrix} 
-\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
-0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
-\end{bmatrix}$ and 
-$Y = \begin{bmatrix} \lambda^{(11)} & 0 & \lambda^{(12)} & 0 \\
-0 & \lambda^{(21)} & 0 & \lambda^{(22)} \end{bmatrix}$. 
-Then $P = X Y^\top$.
-
-We can note that $Y = X \Pi$ where $\Pi$ is the permutation matrix 
-$\Pi = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0 & 0 & 1 & 0 \\ 
-0 & 1 & 0 & 0 \\ 0 & 0 & 0 & 1 \end{bmatrix}$. 
-Therefore, $P = X \Pi X^\top$.
-
-Taking the spectral decomposition of $\Pi = U D U^\top$, we can see that 
-$P = (X U) D (X U)^\top$. We can then denote $\Sigma = |D|^{1/2}$, the 
-square root of the absolute values of the (diagonal) entries of $D$ and 
-obtain $P = (X U \Sigma) I_{pq} (X U \Sigma)^\top$ where $p$ and $q$ correspond 
-to the number of positive and negative eigenvalues of $\Pi$, respectively. 
-Therefore, the PABM with $K = 2$ is a special case of the GRDPG. We can however 
-expand upon this a bit further.
-
-The permutation described by $\Pi$ has two fixed points and one cycle of order 
-2. The two fixed points are at positions $1$ and $4$, so $\Pi$ has two 
-eigenvalues equal to $1$ and corresponding eigenvectors $e_1$ and $e_4$. The 
-cycle of order 2 switching positions $2$ and $3$ corresponds to eigenvalues 
-$1$ and $-1$ with corresponding eigenvalues $(e_2 + e_3) / \sqrt{2}$ and 
-$(e_2 - e_3) / \sqrt{2}$ respectively. Therefore, 
-$D = \begin{bmatrix} 
+This is given by straightforward matrix multiplication. It suffices to show 
+that
+
+$$X U I_{3, 1} U^\top X^\top = 
+\begin{bmatrix} 
+  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
+  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
+\end{bmatrix}$$
+
+**Remark**. While we can just perform the matrix multiplication to show the 
+equivalence, it is more illustrative to look at a few intermediate steps. Note 
+that the product of the three inner matrices results in a permutation matrix 
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swappoing 
+positions $2$ and $3$:
+
+$$U I_{3, 1} U^\top = \begin{bmatrix} 
   1 & 0 & 0 & 0 \\ 
+  0 & 0 & 1 & 0 \\ 
   0 & 1 & 0 & 0 \\ 
-0 & 0 & 1 & 0 \\
-0 & 0 & 0 & -1
-\end{bmatrix} = I_{3, 1}$ and 
-$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
-0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
-0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}$.
-
-Putting it all together, we get $P = (X U) I_{3, 1} (X U)^\top$. Therefore, 
-the PABM with $K = 2$ is a GRDPG with $p = 3$, $q = 1$, $d = K^2 = 4$, and 
-latent positions 
-$X U = \begin{bmatrix} 
-\lambda^{(11)} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(12)} / \sqrt{2} \\
-0 & \lambda^{(22)} & \lambda^{(21)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2}
-\end{bmatrix}$.
-
-### Generalization to $K > 2$
-
-**Theorem 2**. There exists a block diagonal matrix 
+  0 & 0 & 0 & 1 
+\end{bmatrix} = \Pi$$
+
+Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$ 
+is a spectral decomposition of this permutation matrix. Note that the two fixed 
+points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$ 
+where $i = 1, 4$ corresponding to the locations of the fixed points, and the 
+cycle of order two results in two eigenvalues $\pm 1$ with corresponding 
+eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is 
+swapped.
+
+**Theorem 2** (Generalization to $K > 2$). There exists a block diagonal matrix 
 $X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
@@
 *Proof*. 
-Let $\Lambda^{(k)} = 
+Let 
+$$\Lambda^{(k)} = 
 \begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
-\in \mathbb{R}^{n_k \times K}$.  
-Let $X$ be a block diagonal matrix 
-$X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
-Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
-$l = 1, ..., K$. $L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
-\mathbb{R}^{n \times K}$.  
-Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
-\mathbb{R}^{n \times K^2}$.  
+\in \mathbb{R}^{n_k \times K}$$
+
+\begin{equation}
+X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
+\end{equation}
+
+$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
+\mathbb{R}^{n \times K}$$
+
+$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
+\mathbb{R}^{n \times K^2}$$
+
 Then $P = X Y^\top$.
-Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
-$\Pi$, so $P = X \Pi X^\top$.  
+
+Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
+$\Pi$, resulting in $P = X \Pi X^\top$.  
 The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
@@
 
-So $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
+Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
 eigenvalues equal to $-1$. $\Pi$ has the decomposed form
-$\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top$, and we can describe the 
-PABM with $K$ communities as a GRDPG with latent positions $X U$ with signature 
-$\Big( K (K + 1) / 2, K (K - 1) / 2 \Big)$.
-
-**Example** for $K = 3$. Using the same notation as before:
-
-$X = \begin{bmatrix} 
+
+\begin{equation}
+\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
+\end{equation}
+
+The edge probability matrix then can be written as:
+
+\begin{equation}
+P = X U I_{p, q} (X U)^\top
+\end{equation}
+
+\begin{equation}
+p = K (K + 1) / 2
+\end{equation}
+
+\begin{equation}
+q = K (K - 1) / 2
+\end{equation}
+
+and we can describe the PABM with $K$ communities as a GRDPG with latent 
+positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
+
+**Example** ($K = 3$). Using the same notation as before:
+
+$$X = \begin{bmatrix} 
 \lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
@@
 0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
-\end{bmatrix}$
-
-$Y = \begin{bmatrix} 
+\end{bmatrix}$$
+
+$$Y = \begin{bmatrix} 
 \lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
@@
 0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
-\end{bmatrix}$
-
-Then $P = X Y^\top$ and $Y = X \Pi$ where 
-$\Pi = \begin{bmatrix} 
+\end{bmatrix}$$
+
+Then $P = X Y^\top$ and $Y = X \Pi$ where $\Pi$ is a permutation matrix 
+consisting of $3$ fixed points and $3$ cycles of order 2:
+
+$$\Pi = \begin{bmatrix} 
 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
@@
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
-\end{bmatrix}$
-
-Another way to look at this is:
+\end{bmatrix}$$
 
@@
 The latent positions are the rows of  
-$XU = \begin{bmatrix}
-\lambda^{(11)} & 0 & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 & \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
-0 & \lambda^{(22)} & 0 & \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} & -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
-0 & 0 & \lambda^{(33)} & 0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} & 0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
-\end{bmatrix}$.
+$$XU = \begin{bmatrix}
+  \lambda^{(11)} & 0 & 0 & 
+  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 & 
+  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
+  0 & \lambda^{(22)} & 0 & 
+  \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} & 
+  -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
+  0 & 0 & \lambda^{(33)} & 
+  0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} & 
+  0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
+\end{bmatrix}$$
 
@@
 
-Here, we will focus more on (1) and pose possible methods for (2). In our 
-methods, we assume that $K$, the number of communities, is known beforehand and 
-does not require estimation.
-
-## Previous work
+In our methods, we assume that $K$, the number of communities, is known 
+beforehand and does not require estimation.
+
+## Related work
+
+\citet*{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, used 
+modularity maximization for clustering and parameter estimation. They were able 
+to show that as the sample size increases, the proportion of misclassified 
+community labels (up to permutation) goes to 0.
 
 \citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
-to identify the community memberships given either an edge probability matrix $P$ 
-or an adjacency matrix $A$. In the case that $P$ is known, the community 
-memberships can be identified exactly (up to permutation). A similar procedure 
-can be applied if $P$ is unknown and we have an observation $A$, but the 
-theoretical guarantees of this method applied to the PABM are unknown. In 
-particular, the method requires spherical Gaussian noise. The authors of this 
-paper then use point estimators for $\{\lambda^{(kl)}\}$ using cluster labels 
-obtained via SSC.
+to identify the community memberships given either an edge probability matrix 
+$P$ or an adjacency matrix $A$. In the case that $P$ is known, the community 
+memberships can be identified exactly (up to permutation). This is supported by 
+our Theorem 2, which shows that the ASE of edge probability matrix $P$ consists 
+of $K$ $K$-dimensional subspaces lying in $\mathbb{R}^{K^2}$. A similar 
+procedure can be applied if $P$ is unknown and we have an observation $A$, but 
+the theoretical guarantees of this method applied to the PABM are unknown.
 
@@
 subspace. Using this property, previous work proposes the use of subspace 
-clustering while acknowledging some of its shortcomings 
-\cite{noroozi2019estimation} \cite{soltanolkotabi2014}.
+clustering \cite{noroozi2019estimation} \cite{soltanolkotabi2014}.
+
+**Lemma 1.** 
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
+matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
+defined as in equation (3).
+
+*Proof*. 
+By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in equation (3) and $p$ and $q$ are defined as in equations (6) and (7).
+Alternatively, the spectral decomposition can be written as $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ and $|\cdot|^{1/2}$ is applied entry-wise.
+Thus for some $Q \in \mathbb{O}(p, q)$,
+
+$$X U Q = V |D|^{1/2}$$
+
+Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,
+
+$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
+(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$
+
+The righthand side becomes
+
+$$\begin{split}
+  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
+    = X U Q Q^{-1} U^\top (X^\top X)^{-1} U (Q^\top)^{-1} Q^\top U^\top X^\top \\
+  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
+  & = X (X^\top X)^{-1} X^\top
+\end{split}$$
+
+The lefthand side becomes:
+
+$$\begin{split}
+  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
+    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
+  & = V V^\top
+\end{split}$$
 
@@
 
-*Proof (sketch)*
-By projection, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
+*Proof*.
+By Lemma 1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
 in Theorem 2. Since $X$ is block diagonal with each block corresponding 
@@
 adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
-\citet*{rubindelanchy2017statistical},
-
-$$\max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)$$
+\citet{rubindelanchy2017statistical},
+
+\begin{equation}
+  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
+  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
+\end{equation}
 
@@
 
-$$\max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)$$
+\begin{equation}
+  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
+    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
+\end{equation}
+
 where $\lim\limits_{n \to 0} \rho_n = 0$ is a sparsity parameter introduced 
 by \citet{307cbeb9b1be48299388437423d94bf1}. More precisely, 
-$\rho_n = o(\log n / \sqrt{n})$.
-
-In addition, by the central limit theorem, 
-$(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a normal distribution 
-centered at $0$.
+$\rho_n = o(\log n / \sqrt{n})$ \cite{rubindelanchy2017statistical}.
 
@@
 
-A similar method can be applied using $\hat{Z}$, the adjacency spectral 
-embedding of $A$.
+Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
+simply use plug-in estimators:
 
@@
 
-$$\max_{k, l \in \{1, ..., K\}} 
+\begin{equation}
+\max_{k, l \in \{1, ..., K\}} 
 ||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
-O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)$$
+O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
+\end{equation}
 
@@
 
-* *Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
+*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
 probabilities and edges for community $k$.  
@@
 
-* *Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
+*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
 edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
@@
 Now consider the Hermitian dilation
-$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$, 
+
+$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$
+
 which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
 that the eigendecomposition of $M^{(kl)}$ is
-$M^{(kl)} = 
+
+$$M^{(kl)} = 
 \begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
 \begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$.  
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$
+
 Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
 latent positions in $\mathbb{R}^2$ given by
-$\begin{bmatrix} 
+
+$$\begin{bmatrix} 
   \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
@@
   \lambda^{(lk)} & -\lambda^{(lk)} 
-\end{bmatrix}$.  
-Now consider $\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$. 
+\end{bmatrix}$$
+
+Now consider
+
+$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
 Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where
-$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$ and $E$ is the 
+
+$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$
+
+and $E$ is the 
 $n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
@@
 spectral embedding, given by
-$\begin{bmatrix} 
+
+$$\begin{bmatrix} 
   \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
   \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
-\end{bmatrix}$ 
+\end{bmatrix}$$
+
 where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
@@
 
+Both Algorithm 2 and Sparse Subspace Clustering produce an affinity matrix $B$ 
+for which $B_{ij} \to 0$ if vertices $i$ and $j$ belong in different 
+communities. Spectral clustering was performed on $B$ to obtain the final 
+community estimates.
+
+For estimating $\{\lambda^{(kl)}_K$, both methods usethe ground truth community 
+labels.
+
 ## Balanced communities
@@
 
+Fig \ref{fig:clust_err_k} show that for large $n$, Algorithm 2 results in 
+a misclustering rate of 0. Sparse Subspace Clustering produces similar results 
+for $K > 2$.
+
 ```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
-Simulations were repeated 50 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
+Simulations were repeated 50 times for each sample size.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k.csv')
@@
 converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
+We explore this in Fig \ref{fig:clust_err_ct_sim}.
 
 ```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 50 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
+Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k.csv')
@@
 
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+Given ground truth community labels, Algorithm 5 and the MLE-based plug-in 
+estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
+perform similarly, with root mean square error decaying at rate approximately 
+$n^{-1/2}$.
+
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 
+Fig \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_k_imba} show 
+similar results as in the balanced communities case, with both Algorithm 2 and 
+Sparse Subspace Clustering resulting in no misclustered vertices for a 
+sufficiently large sample size. However, Fig \ref{fig:lambda_est_k_imba} 
+suggests that while Algorithm 5 retains $\sqrt{n}$ efficiency, the MLE-based 
+plug-in estimator fails to do so.
+
 ```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and subspace clustering (green) for sample sizes from 128 to 4096 and number of clusters from 2 to 4.
-Simulations were repeated 50 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
+Simulations were repeated 50 times for each sample size.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
@@
 ```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 50 times for each sample size.
-The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
+Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
@@
 
-```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 
+# z.hat.mm <- A %>% 
+#   igraph::graph_from_adjacency_matrix(mode = 'undirected') %>% 
+#   igraph::cluster_louvain() %>% 
+#   igraph::membership()
+
 # ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
@@
 
-zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3)
+zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3, normalize = TRUE)
 bmp.acc.ssc <- mean(zhat.bmp.ssc == z.bmp)
@@
 
-zhat.pb.ssc <- ssc(A.pb, 2, 1e-4)
+zhat.pb.ssc <- ssc(A.pb, 2, .008, normalize = TRUE)
 pb.acc.ssc <- mean(zhat.pb.ssc == z.pb)
@@
 
-zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2)
+zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2, normalize = FALSE)
 dblp.acc.ssc <- mean(zhat.dblp.ssc == z.dblp)
 dblp.acc.ssc <- ifelse(dblp.acc.ssc < .5, 1 - dblp.acc.ssc, dblp.acc.ssc)
+
 ```
@@
 In the third example, we consider the Karantaka villages data studied by 
-\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` network from 
-village 12 at the household level. The label of interest is the religious 
-affiliation. The network was truncated to religions "1" and "2", and vertices 
-of degree 0 were removed. The resulting network has 141 vertices.
+\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` networks from 
+villages 12, 31, and 46 at the household level. The label of interest is the 
+religious affiliation. The networks were truncated to religions "1" and "2", 
+and vertices of degree 0 were removed.
 
 ```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left), sparse subspace clustering (middle), and modularity maximization (right).', fig.height = 2, fig.width = 8, eval = TRUE}
-hh <- readRDS('data/village-hh-12.rds')
-A.hh <- hh$A
-z.hh <- hh$z
 K <- 2
-zhat.hh <- cluster.pabm(A.hh, K,
-                        use.all = TRUE, normalize = FALSE, p = 3, q = 1)
-zhat.hh.ssc <- ssc(A.hh, K, lambda = .1)
+hh.12 <- readRDS('data/village-hh-12.rds')
+hh.31 <- readRDS('data/village-hh-31.rds')
+hh.46 <- readRDS('data/village-hh-46.rds')
 
@@
 source("functions.R")
+mm <- function(A.hh) {
   b.can = EPalgo(A.hh,eps=0) # EP algorithm (no perturbation)
@@
     #check if any cluster is empty
-  foo = rep(NA, max(z.hh))
-  for (clus in 1:max(z.hh)) {foo[clus]=sum(b.can[,i]==clus)}
+    foo = rep(NA, 2)
+    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
     if (min(foo)==0) {stop('Empty groups are not allowed')} 
@@
   foo1 = order(-Q.PA.can)[1] 
-zhat.hh.mm = b.can[,foo1]   # community assignment that maximises Q.PA
-
-gridExtra::grid.arrange(plot.A(A.hh, zhat.hh),
-                        plot.A(A.hh, zhat.hh.ssc),
-                        plot.A(A.hh, zhat.hh.mm),
+  return(b.can[,foo1])   # community assignment that maximises Q.PA
+}
+
+A.hh.12 <- hh.12$A
+z.hh.12 <- hh.12$z
+zhat.hh.12 <- cluster.pabm(A.hh.12, K,
+                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
+zhat.hh.ssc.12 <- ssc(A.hh.12, K, lambda = .1)
+zhat.hh.mm.12 <- mm(A.hh.12)
+
+A.hh.31 <- hh.31$A
+z.hh.31 <- hh.31$z
+zhat.hh.31 <- cluster.pabm(A.hh.31, K,
+                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
+zhat.hh.ssc.31 <- ssc(A.hh.31, K, lambda = .05)
+zhat.hh.mm.31 <- mm(A.hh.31)
+
+A.hh.46 <- hh.46$A
+z.hh.46 <- hh.46$z
+zhat.hh.46 <- cluster.pabm(A.hh.46, K,
+                           use.all = TRUE, normalize = FALSE, p = 2, q = 0)
+zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .0005)
+zhat.hh.mm.46 <- mm(A.hh.46)
+
+gridExtra::grid.arrange(plot.A(A.hh.12, zhat.hh.12),
+                        plot.A(A.hh.31, zhat.hh.31),
+                        plot.A(A.hh.46, zhat.hh.46),
                         nrow = 1)
@@
 ```{r, eval = TRUE}
-hh.acc <- mean(zhat.hh == z.hh)
-hh.acc <- ifelse(hh.acc < .5, 1 - hh.acc, hh.acc)
-hh.acc.ssc <- mean(zhat.hh.ssc == z.hh)
-hh.acc.ssc <- ifelse(hh.acc.ssc < .5, 1 - hh.acc.ssc, hh.acc.ssc)
-hh.acc.mm <- mean(zhat.hh.mm == z.hh)
-hh.acc.mm <- ifelse(hh.acc.mm < .5, 1 - hh.acc.mm, hh.acc.mm)
+hh.acc.12 <- mean(zhat.hh.12 == z.hh.12)
+hh.acc.12 <- ifelse(hh.acc.12 < .5, 1 - hh.acc.12, hh.acc.12)
+hh.acc.ssc.12 <- mean(zhat.hh.ssc.12 == z.hh.12)
+hh.acc.ssc.12 <- ifelse(hh.acc.ssc.12 < .5, 1 - hh.acc.ssc.12, hh.acc.ssc.12)
+hh.acc.mm.12 <- mean(zhat.hh.mm.12 == z.hh.12)
+hh.acc.mm.12 <- ifelse(hh.acc.mm.12 < .5, 1 - hh.acc.mm.12, hh.acc.mm.12)
 
@@
                                    'Mod. Max.'),
-                        Error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
+                        Error = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3)),
              caption = paste('Community detection error rates for identifying',

### 2020-10-29 summary.Rmd
minor tweaks to writeup

@@
 
-For estimating $\{\lambda^{(kl)}_K$, both methods usethe ground truth community 
+For estimating $\{\lambda^{(kl)}\}_K$, both methods usethe ground truth community 
 labels.
@@
 
-* Number of vertices $n = 256, 512, 1024, 2048, 4096$
+* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
 * Number of underlying communities $K = 2, 3, 4$
@@
 * Community labels 
-$z_k \stackrel{iid}{\sim} Multinomial(\alpha_1, ..., \alpha_K)$
-* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
+$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
+* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
 * Between-group popularities 
-$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$
+$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$
 
@@
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error', 
+  labs(y = 'community detection error rate', 
        colour = NULL) +
@@
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'misclustering count', 
+  labs(y = 'community detection error count', 
        colour = NULL) +
@@
 
-Fig \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_k_imba} show 
+Fig \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
 similar results as in the balanced communities case, with both Algorithm 2 and 
@@
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left), sparse subspace clustering (middle), and modularity maximization (right).', fig.height = 2, fig.width = 8, eval = TRUE}
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left).
+The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
 K <- 2
@@
 
-knitr::kable(data.frame(Method = c('Alg. 2', 'SSC', 
-                                   'Mod. Max.'),
-                        Error = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3)),
+hh.acc.31 <- mean(zhat.hh.31 == z.hh.31)
+hh.acc.31 <- ifelse(hh.acc.31 < .5, 1 - hh.acc.31, hh.acc.31)
+hh.acc.ssc.31 <- mean(zhat.hh.ssc.31 == z.hh.31)
+hh.acc.ssc.31 <- ifelse(hh.acc.ssc.31 < .5, 1 - hh.acc.ssc.31, hh.acc.ssc.31)
+hh.acc.mm.31 <- mean(zhat.hh.mm.31 == z.hh.31)
+hh.acc.mm.31 <- ifelse(hh.acc.mm.31 < .5, 1 - hh.acc.mm.31, hh.acc.mm.31)
+
+hh.acc.46 <- mean(zhat.hh.46 == z.hh.46)
+hh.acc.46 <- ifelse(hh.acc.46 < .5, 1 - hh.acc.46, hh.acc.46)
+hh.acc.ssc.46 <- mean(zhat.hh.ssc.46 == z.hh.46)
+hh.acc.ssc.46 <- ifelse(hh.acc.ssc.46 < .5, 1 - hh.acc.ssc.46, hh.acc.ssc.46)
+hh.acc.mm.46 <- mean(zhat.hh.mm.46 == z.hh.46)
+hh.acc.mm.46 <- ifelse(hh.acc.mm.46 < .5, 1 - hh.acc.mm.46, hh.acc.mm.46)
+
+knitr::kable(dplyr::tibble(
+  Method = c('Alg. 2', 'SSC', 'Mod. Max.'),
+  `Error (Village 12)` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
+  `Error (Village 31)` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
+  `Error (Village 46)` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
   caption = paste('Community detection error rates for identifying',

### 2020-10-29 villages-data.R

@@ -1,0 +1,88 @@
+import::from(magrittr, `%>%`)
+import::from(readr, read_csv)
+library(mclust)
+library(haven)
+library(ggplot2)
+
+type <- 'visitgo'
+villageno <- 0
+
+# villageno <- villageno + 1
+
+classification <- 'hohreligion'
+keep <- c(1, 2)
+villageno <- 12
+
+setwd('~/dev/multiresolution_networks/data/indian_village_raw/datav4.0/Data')
+
+villageno <- villageno + 1
+print(villageno)
+
+A <- read_csv(paste0("1. Network Data/Adjacency Matrices/adj_", type, "_HH_vilno_", villageno, ".csv"), 
+              col_names = FALSE) %>% 
+  as.matrix() %>% 
+  magrittr::set_colnames(NULL)
+deg <- colSums(A)
+A <- A[deg > 0, deg > 0]
+household_characteristics <- haven::read_dta("2. Demographics and Outcomes/household_characteristics.dta") %>% 
+  dplyr::filter(village == villageno)
+household_characteristics <- household_characteristics[deg > 0, ]
+z <- as.numeric(household_characteristics[[classification]])
+A <- A[z %in% keep, z %in% keep]
+z <- z[z %in% keep]
+deg <- colSums(A)
+A <- A[deg > 0, deg > 0]
+z <- z[deg > 0]
+table(z)
+qgraph::qgraph(A, groups = factor(z))
+
+saveRDS(list(A = A, z = z), 
+        paste0('~/dev/pabm-grdpg/data/village-hh-', villageno, '.rds'))
+
+K <- length(unique(z))
+
+p <- 3
+q <- 1
+normalize <- FALSE
+d.eigenmap <- K + 1
+laplacian <- 'normalized'
+use.all <- TRUE
+n <- nrow(A)
+if (is.null(p)) {
+  p <- K * (K + 1) / 2
+}
+if (is.null(q)) {
+  q <- K * (K - 1) / 2
+}
+if (p * q > 0) {
+  indices <- c(seq(p), seq(n, n - q + 1))
+} else if (p == 0) {
+  indices <- seq(n, n - q + q)
+} else if (q == 0) {
+  indices <- seq(p)
+}
+V <- eigen(A, symmetric = TRUE)$vectors[, indices]
+if (normalize) {
+  v.norm <- sqrt(rowSums(V ** 2))
[…]

### 2020-11-04 clustering-sim-k.R
more simulations

@@ -59,7 +59,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       P <- Pz$P
       z <- Pz$clustering
       A <- draw.graph(P)
-      clustering <- cluster.pabm(A, K, use.all = TRUE)
+      clustering <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error <- 1 - cluster.acc(clustering, z)
       clustering.ssc <- ssc(A, K, sparsity, normalize = TRUE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)

### 2020-11-04 summary.Rmd
added algorithms for ssc and mm

@@
 communities. Spectral clustering was performed on $B$ to obtain the final 
-community estimates.
-
-For estimating $\{\lambda^{(kl)}\}_K$, both methods usethe ground truth community 
-labels.
+community estimates. More specifically, clustering was performed on a 
+normalized Laplacian eigenmap of dimension $K + 1$ by fitting Gaussian Mixture 
+Model.
+
+For estimating $\{\lambda^{(kl)}\}_K$, both methods use the ground truth 
+community labels.
+
+### Sparse subspace Clustering and Modularity Maximization
+
+**TODO** find citations for these
+
+For each simulation, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
+from a joint distribution, the edge probability matrix $P$ is constructed using 
+the popularity vectors, and finally an unweighted and undirected adjacency 
+matrix $A$ is drawn from $P$. Three community detection methods are applied to 
+$A$: Algorithm 2 from this paper, Sparse Subspace Clustering, and Modularity 
+Maximization. The latter two algorithms are detailed as follows:
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{Sparse Subspace Clustering.}
+  \KwData{Adjacency matrix $A$, number of communities $K$.}
+  \KwResult{Community assignments $1, ..., K$.}
+  Embed $A$ in $d$ dimensions using ASE to obtain 
+  $\hat{Y} \in \mathbb{R}^{n \times d}$.\;
+  Fit a sparse linear regression model to each row of $\hat{Y}$ using the 
+  remaining $n - 1$ rows of $Y$ as predictors. The models must exclude the 
+  intercept term.\;
+  Construct $n \times n$ symmetric matrix $C$ such that $c_{ij}$ is the $j^{th}$
+  coefficient for the $i^{th}$ model.\;
+  Normalize the columns of $|C|$ to obtain $\tilde{C}$.\;
+  Compute the affinity matrix $B = \tilde{C} + \tilde{C}^\top$.\;
+  Construct graph $G$ using $B$ as its similarity matrix.\;
+  Partition $G$ into $K$ disconnected components (e.g., using edge thresholding 
+  or spectral clustering).\;
+  Map each partition to the community labels $1, ..., K$\;
+\end{algorithm}
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{Modularity Maximization (Louvain Method)}
+  \KwData{Adjacency matrix $A$, number of communities $K$.}
+  \KwResult{Community assignments $1, ..., K$}
+  Initialize the clustering by assigning to each vertex a unique community 
+  label.\;
+  For each vertex $v_i$, reassign the label to match one of its neighboring 
+  vertices, and compute the change in modularity for each reassignment.\;
+  Reassign $v_i$ to the community that results in the most positive change in 
+  modularity, if possible. If modularity does not increase by changing the 
+  label, skip the vertex.\;
+  Group vertices with the same label and treat as one vertex.\;
+  Repeat steps (3) and (4) until termination or the desired number of 
+  communities is estimated.
+\end{algorithm}
 

### 2020-11-05 summary.Rmd
fixed a few things in writeup

@@
 
-\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
+\citet*{noroozi2019estimation} proposed using sparse subspace clustering 
 to identify the community memberships given either an edge probability matrix 
@@
 
-Both Algorithm 2 and Sparse Subspace Clustering produce an affinity matrix $B$ 
+Both Algorithm 2 and sparse subspace clustering produce an affinity matrix $B$ 
 for which $B_{ij} \to 0$ if vertices $i$ and $j$ belong in different 
@@
 
-### Sparse subspace Clustering and Modularity Maximization
+### Sparse subspace clustering and modularity maximization
 
@@
 matrix $A$ is drawn from $P$. Three community detection methods are applied to 
-$A$: Algorithm 2 from this paper, Sparse Subspace Clustering, and Modularity 
-Maximization. The latter two algorithms are detailed as follows:
+$A$: Algorithm 2 from this paper, sparse subspace clustering, and modularity 
+maximization.
+
+Sparse subspace clustering is described by the following:
 
@@
   \SetAlgoLined
-  \caption{Sparse Subspace Clustering.}
+  \caption{Sparse subspace clustering.}
   \KwData{Adjacency matrix $A$, number of communities $K$.}
@@
 
+For modularity maximization, \citet{307cbeb9b1be48299388437423d94bf1} used the 
+Extreme Points (EP) algorithm \cite{le2016}, which is specific to the case 
+$K = 2$. In our simulation studies, we used the EP algorithm for $K = 2$ and 
+the Louvain algorithm for $K > 2$. The Louvain algorithm is described by the 
+following:
+
 \begin{algorithm}[H]
@@
   \SetAlgoLined
-  \caption{Modularity Maximization (Louvain Method)}
+  \caption{Modularity maximization (Louvain Method)}
   \KwData{Adjacency matrix $A$, number of communities $K$.}
@@
 Fig \ref{fig:clust_err_k} show that for large $n$, Algorithm 2 results in 
-a misclustering rate of 0. Sparse Subspace Clustering produces similar results 
+a misclustering rate of 0. Sparse subspace clustering produces similar results 
 for $K > 2$.
@@
 similar results as in the balanced communities case, with both Algorithm 2 and 
-Sparse Subspace Clustering resulting in no misclustered vertices for a 
+sparse subspace clustering resulting in no misclustered vertices for a 
 sufficiently large sample size. However, Fig \ref{fig:lambda_est_k_imba} 

### 2020-11-05 functions.R

@@ -244,7 +244,7 @@ ssc <- function(A,
                 K = 2,
                 lambda = .01,
                 parallel = FALSE,
-                normalize = FALSE) {
+                normalize = TRUE) {
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   
@@ -405,4 +405,46 @@ plot.A <- function(A, z, lines = TRUE, max.size = 500) {
   }
   
   return(out.plot)
+}
+
+mod.max.exchange <- function(A, init.clust) {
+  K <- max(init.clust)
+  n <- nrow(A)
+  indices <- seq(n)
+  clustering <- init.clust
+  clusters <- seq(K)
+  current.modularity <- Q.PA(A, clustering)
+  prev.modularity <- current.modularity - 1
+  print(current.modularity)
+  
+  iter <- 0
+  
+  while (prev.modularity != current.modularity) {
+    prev.modularity <- current.modularity
+    # indices <- sample(indices)
+    for (i in indices) {
+      current.label <- clustering[i]
+      neighbors <- which(A[i, ] == 1)
+      neighboring.labels <- unique(clustering[neighbors])
+      if ((length(neighboring.labels) > 1) | 
+          (neighboring.labels[1] != current.label)) {
+        # print(i)
+        for (k in neighboring.labels) {
+          if (k != current.label) {
+            proposed.clustering <- clustering
+            proposed.clustering[i] <- k
+            proposed.modularity <- Q.PA(A, proposed.clustering)
+            if (proposed.modularity > current.modularity) {
+              clustering <- proposed.clustering
+              current.modularity <- proposed.modularity
+            }
+          }
+        }
+      }
+    }
+    iter <- iter + 1
+    print(iter)
+    print(current.modularity)
+  }
+  return(clustering)
 }

### 2020-11-11 summary.Rmd
minor update

@@
 $K = 2$. In our simulation studies, we used the EP algorithm for $K = 2$ and 
-the Louvain algorithm for $K > 2$. The Louvain algorithm is described by the 
-following:
+the Louvain algorithm \cite{Blondel_2008} for $K > 2$.
+
+The EP algorithm is described by the following:
 
@@
   \SetAlgoLined
-  \caption{Modularity maximization (Louvain Method)}
+  \caption{Modularity maximization (EP algorithm)}
+  \KwData{Adjacency matrix $A$, number of communities $K = 2$.}
+  \KwResult{Community assignments $1, ..., K$.}
+  Embed vertices of the graph described by $A$ into the subspace constructed 
+  by a low rank approximation of $A$ (e.g., by the first few eigenvectors of 
+  $A$).\;
+  Project the hypercube $[-1, 1]^n$ to $U_A$.\;
+\end{algorithm}
+
+The Louvain algorithm is described by the following:
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{Modularity maximization (Louvain method)}
   \KwData{Adjacency matrix $A$, number of communities $K$.}
-  \KwResult{Community assignments $1, ..., K$}
+  \KwResult{Community assignments $1, ..., K$.}
   Initialize the clustering by assigning to each vertex a unique community 

### 2020-11-18 summary.Rmd
minor rewording

--- 2020-11-05 summary.Rmd
@@
 
-*Proof (sketch)*. 
+*Proof*. 
 Let $P$ and $A$ be organized by community such that the elements of blocks 
@@
 
-In these simulations, Algorithm 2 is compared against sparse subspace 
-clustering \cite{soltanolkotabi2014} \cite{noroozi2019estimation} and 
-modularity maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}, 
-and Algorithm 4 is compared against an MLE-based estimator 
-\cite{noroozi2019estimation} \cite{307cbeb9b1be48299388437423d94bf1}.
-
-Both Algorithm 2 and sparse subspace clustering produce an affinity matrix $B$ 
-for which $B_{ij} \to 0$ if vertices $i$ and $j$ belong in different 
-communities. Spectral clustering was performed on $B$ to obtain the final 
-community estimates. More specifically, clustering was performed on a 
-normalized Laplacian eigenmap of dimension $K + 1$ by fitting Gaussian Mixture 
-Model.
-
-For estimating $\{\lambda^{(kl)}\}_K$, both methods use the ground truth 
-community labels.
-
-### Sparse subspace clustering and modularity maximization
-
-**TODO** find citations for these
-
 For each simulation, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
@@
 the popularity vectors, and finally an unweighted and undirected adjacency 
-matrix $A$ is drawn from $P$. Three community detection methods are applied to 
-$A$: Algorithm 2 from this paper, sparse subspace clustering, and modularity 
-maximization.
-
-Sparse subspace clustering is described by the following:
-
-\begin{algorithm}[H]
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{Sparse subspace clustering.}
-  \KwData{Adjacency matrix $A$, number of communities $K$.}
-  \KwResult{Community assignments $1, ..., K$.}
-  Embed $A$ in $d$ dimensions using ASE to obtain 
-  $\hat{Y} \in \mathbb{R}^{n \times d}$.\;
-  Fit a sparse linear regression model to each row of $\hat{Y}$ using the 
-  remaining $n - 1$ rows of $Y$ as predictors. The models must exclude the 
-  intercept term.\;
-  Construct $n \times n$ symmetric matrix $C$ such that $c_{ij}$ is the $j^{th}$
-  coefficient for the $i^{th}$ model.\;
-  Normalize the columns of $|C|$ to obtain $\tilde{C}$.\;
-  Compute the affinity matrix $B = \tilde{C} + \tilde{C}^\top$.\;
-  Construct graph $G$ using $B$ as its similarity matrix.\;
-  Partition $G$ into $K$ disconnected components (e.g., using edge thresholding 
-  or spectral clustering).\;
-  Map each partition to the community labels $1, ..., K$\;
-\end{algorithm}
-
-For modularity maximization, \citet{307cbeb9b1be48299388437423d94bf1} used the 
-Extreme Points (EP) algorithm \cite{le2016}, which is specific to the case 
-$K = 2$. In our simulation studies, we used the EP algorithm for $K = 2$ and 
-the Louvain algorithm for $K > 2$. The Louvain algorithm is described by the 
-following:
-
-\begin{algorithm}[H]
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{Modularity maximization (Louvain Method)}
-  \KwData{Adjacency matrix $A$, number of communities $K$.}
-  \KwResult{Community assignments $1, ..., K$}
-  Initialize the clustering by assigning to each vertex a unique community 
-  label.\;
-  For each vertex $v_i$, reassign the label to match one of its neighboring 
-  vertices, and compute the change in modularity for each reassignment.\;
-  Reassign $v_i$ to the community that results in the most positive change in 
-  modularity, if possible. If modularity does not increase by changing the 
-  label, skip the vertex.\;
-  Group vertices with the same label and treat as one vertex.\;
-  Repeat steps (3) and (4) until termination or the desired number of 
-  communities is estimated.
-\end{algorithm}
+matrix $A$ is drawn from $P$. Algorithm 2 is then used for community detection, 
+and this method is compared against sparse subspace clustering 
+\cite{soltanolkotabi2014} \cite{noroozi2019estimation} and modularity 
+maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}. True 
+community labels are used with Algorithm 4 to estimate the popularity 
+vectors, and this method is then compared against an MLE-based estimator 
+described in \citet{noroozi2019estimation} and
+\citet{307cbeb9b1be48299388437423d94bf1}.
+
+Sparse subspace clustering is described as follows. First, an embedding of the 
+affinity matrix $A$ is required, and for this, we use spectral decomposition 
+using the first $K (K + 1) / 2$ and last $K (K - 1) / 2$ eigenvectors. Then a 
+sparse linear regression model is fit to each row of the embedding data matrix 
+using the remaining rows as predictors, the rows of which are stored in 
+matrix $C$. Finally, an affinity matrix $B = |C| + |C^\top|$ is constructed, 
+where $|\cdot|$ is applied entry-wise. If the data in the embedding lie 
+perfectly in $K$ subspaces, then $B_{ij} = 0$ for $i, j$ not in the same 
+subspace. If the embedding of $P$ is used, then this property holds, but it 
+is true in general for the embedding of $A$, in which case a final clustering 
+step is required. For these simulations, spectral clustering with the 
+normalized Laplacian and Gaussian mixture models is used for this final step.
+
+Modularity maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
+used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$. 
+For these simulations, the EP algorithm was used for $K = 2$, and for $K > 2$, 
+the Louvain algorithm \cite{Blondel_2008} was used instead.
 

### 2020-12-16 summary.Rmd
fixed a few citations
just syncing some stuff up

@@
 abstract: |
-  In this paper, we connect two probabilistic models for graphs, the Popularity 
+  In this paper, we connect two random graph models, the Popularity 
   Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
@@
 vertex, up to permutation (with some additional parameter estimation being of 
-possible interest), assuming some probabilistic graph model. One such model is 
+possible interest), assuming some random graph model. One such model is 
 the Stochastic Block Model (SBM), first proposed by 
@@
 probability from one vertex to another follows a Beronulli distribution with 
-fixed probabilities for each pair of community labels. Other probabilistic 
+fixed probabilities for each pair of community labels. Other random 
 graph models have been proposed and studied, such as the Degree-Corrected Block 
@@
 given community share the same position in the latent space 
-\cite{lyzinski2015perfect}. It has also been shown that similar 
-probabilistic graph models, including both the MMSBM and DCBM, can be 
+\cite{lyzinski2014}. It has also been shown that similar 
+random graph models, including both the MMSBM and DCBM, can be 
 represented in this way \cite{rubindelanchy2017consistency} 
-\cite{lyzinski2015perfect}. An analogous 
+\cite{lyzinski2014}. An analogous 
 property exists for the PABM but not for the RDPG model but under the 
@@
 to show that as the sample size increases, the proportion of misclassified 
-community labels (up to permutation) goes to 0.
+community labels (up to permutation) goes to 0. However, because modularity 
+maximization is NP-hard, the authors of this paper used an approximation 
+based on the Extreme Points algorithm \cite{le2016}.
 
@@
 using the remaining rows as predictors, the rows of which are stored in 
-matrix $C$. Finally, an affinity matrix $B = |C| + |C^\top|$ is constructed, 
+matrix $C$. Finally, an affinity matrix $B = |C| + |C|^\top$ is constructed, 
 where $|\cdot|$ is applied entry-wise. If the data in the embedding lie 

### 2020-12-27 summary.Rmd
added additional simulation results

@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix, for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k.csv')
-clustering.df %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'Algorithm 2')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 2'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm,
-                colour = 'Modularity maximization')) +
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'Modularity maximization'), width = .1) +
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') +
-  scale_y_log10()
-```
-
-Theorem 4 implies that algorithm 2 will result in not just in the error rate 
-converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
-We explore this in Fig \ref{fig:clust_err_ct_sim}.
-
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k.csv')
 clustering.df %>%
@@
   dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
+  ggplot() +
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  labs(y = 'community detection error count', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'OSC')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'OSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC-A')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC-A'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.mm,
+                colour = 'MM-EP')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'MM-EP'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc2,
+                colour = 'SSC-W')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
+                    colour = 'SSC-W'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10()
+```
+
+Theorem 4 implies that algorithm 2 will result in not just in the error rate 
+converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
+We explore this in Fig \ref{fig:clust_err_ct_sim}.
+
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix, for sample sizes from 128 to 4096.
+Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k.csv')
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
+  dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
   ggplot() +
@@
   geom_line(aes(x = n, y = med.err * n,
-                colour = 'Algorithm 2')) +
+                colour = 'OSC')) +
   geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'Algorithm 2'), width = .1) + 
+                    colour = 'OSC'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC')) + 
+                colour = 'SSC-A')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC'), width = .1) + 
+                    colour = 'SSC-A'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm * n,
-                colour = 'Modularity maximization')) + 
+                colour = 'MM-EP')) + 
   geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
-                    colour = 'Modularity maximization'), width = .1) + 
+                    colour = 'MM-EP'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc2 * n,
+                colour = 'SSC-W')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
+                    colour = 'SSC-W'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 

### 2020-12-29 summary.Rmd
updated simulations

--- 2020-12-16 summary.Rmd
@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+Communities are approximately balanced.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k.csv')
-clustering.df %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'Algorithm 2')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 2'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm,
-                colour = 'Modularity maximization')) +
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'Modularity maximization'), width = .1) +
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') +
-  scale_y_log10()
-```
-
-Theorem 4 implies that algorithm 2 will result in not just in the error rate 
-converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
-We explore this in Fig \ref{fig:clust_err_ct_sim}.
-
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k.csv')
 clustering.df %>%
@@
   dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
+  ggplot() +
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  labs(y = 'community detection error count', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'OSC')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'OSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC-ASE')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC-ASE'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.mm,
+                colour = 'MM-EP')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'MM-EP'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc2,
+                colour = 'SSC-A')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
+                    colour = 'SSC-A'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10()
+```
+
+Theorem 4 implies that algorithm 2 will result in not just in the error rate 
+converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
+We explore this in Fig \ref{fig:clust_err_ct_sim}.
+
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+Communities are approximately balanced.
+Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k.csv')
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
+  dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
   ggplot() +
@@
   geom_line(aes(x = n, y = med.err * n,
-                colour = 'Algorithm 2')) +
+                colour = 'OSC')) +
   geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'Algorithm 2'), width = .1) + 
+                    colour = 'OSC'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC')) + 
+                colour = 'SSC-ASE')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC'), width = .1) + 
+                    colour = 'SSC-ASE'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm * n,
-                colour = 'Modularity maximization')) + 
+                colour = 'MM-EP')) + 
   geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
-                    colour = 'Modularity maximization'), width = .1) + 
+                    colour = 'MM-EP'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc2 * n,
+                colour = 'SSC-A')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
+                    colour = 'SSC-A'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and subspace clustering (green) for sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+Communities are imbalanced.
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
 clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
-clustering.df %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'Algorithm 2')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'Algorithm 2'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm, colour = 'Modularity maximization')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm, 
-                    colour = 'Modularity maximization'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') +
-  scale_y_log10()
-```
-
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
 clustering.df %>%
@@
   dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
+  ggplot() +
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  labs(y = 'community detection error count', 
+       colour = NULL) +
+  geom_line(aes(x = n, y = med.err,
+                colour = 'OSC')) +
+  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
+                    colour = 'OSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc,
+                colour = 'SSC-ASE')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
+                    colour = 'SSC-ASE'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.mm,
+                colour = 'MM-EP')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+                    colour = 'MM-EP'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc2,
+                colour = 'SSC-A')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
+                    colour = 'SSC-A'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10()
+```
+
+```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+Communities are imbalanced.
+Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
+clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
+ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(med.err = median(error),
+                   first.q = quantile(error, .25),
+                   third.q = quantile(error, .75),
+                   med.err.ssc = median(error.ssc),
+                   first.q.ssc = quantile(error.ssc, .25),
+                   third.q.ssc = quantile(error.ssc, .75),
+                   med.err.mm = median(error.mm),
+                   first.q.mm = quantile(error.mm, .25),
+                   third.q.mm = quantile(error.mm, .75)) %>%
+  dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
+                       first.q.ssc2 = quantile(error.ssc2, .25),
+                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
   ggplot() +

### 2020-12-29 summary.Rmd
typo

@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
 Communities are approximately balanced.
@@
 
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
 Communities are approximately balanced.
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
 Communities are imbalanced.
@@
 
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (green), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (purple), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
 Communities are imbalanced.

### 2021-01-06 summary.Rmd
typo

@@
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error count', 
+  labs(y = 'community detection error rate', 
        colour = NULL) +
@@
   geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC-ASE')) + 
+                colour = 'SSC-A')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC-ASE'), width = .1) + 
+                    colour = 'SSC-A'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm,
@@
   geom_line(aes(x = n, y = med.err.ssc2,
-                colour = 'SSC-A')) + 
+                colour = 'SSC-ASE')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
-                    colour = 'SSC-A'), width = .1) + 
+                    colour = 'SSC-ASE'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
   geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) + 
+                colour = 'SSC-A')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) + 
+                    colour = 'SSC-A'), width = .1) + 
   geom_line(aes(x = n, y = med.err.mm * n,
@@
   geom_line(aes(x = n, y = med.err.ssc2 * n,
-                colour = 'SSC-A')) + 
+                colour = 'SSC-ASE')) + 
   geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
-                    colour = 'SSC-A'), width = .1) + 
+                    colour = 'SSC-ASE'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 

### 2021-02-09 summary.Rmd
rewrote some stuff for clarity

@@
 a graph into disconnected subgraphs or clusters. This is often motivated by the 
-belief that there exist underlying and unobserved communities to which each 
+assumption that there exist underlying and unobserved communities to which each 
 vertex of the graph belongs, and edges between pairs of vertices are determined 
 by drawing from a probability distribution based on the community relationships 
-between each pair. The goal of this analysis then is population community 
+between each pair. The goal of the analysis then is population community 
 detection, or the recovery of the true underlying community labels for each 
 vertex, up to permutation (with some additional parameter estimation being of 
-possible interest), assuming some random graph model. One such model is 
-the Stochastic Block Model (SBM), first proposed by 
+possible interest), assuming some underlying probability model. One such model
+is the Stochastic Block Model (SBM), first proposed by 
 \citet{doi:10.1080/0022250X.1971.9989788}, which assumes that the edge 
@@
 graph models have been proposed and studied, such as the Degree-Corrected Block 
-Model (DCBM), introduced by \citet{Karrer_2011} and the Mixed Membership 
-Stochastic Block Model (MMSBM), introduced by  \citet*{NIPS2008_3578}. The 
-Popularity Adjusted Block Model (PABM) was then introduced by 
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
+Model (DCBM), introduced by \citet{Karrer_2011}, which is a generalization of 
+the SBM. The Popularity Adjusted Block Model (PABM) was then introduced by 
+\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the DCBM to 
 address the heterogeneity of edge probabilities within and between communities 
@@
 The Random Dot Product Graph (RDPG) model proposed by 
-\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli edge 
-probabilities. Under this model, each vertex of the graph can be represented by 
-a point in a latent Euclidean space such that the edge probability between any 
-pair of vertices is given by their corresponding dot product in the latent 
-space, i.e., given a latent positions $x_1, ..., x_n \in \mathbb{R}^d$, the 
-edge probability matrix is $P = X X^\top$ where 
+\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli 
+edge probabilities. Under this model, each vertex of the graph can be
+represented by a point in some latent space such that the edge
+probability between any pair of vertices is given by their corresponding dot
+product in the latent space, i.e., given a latent positions
+$x_1, ..., x_n \in \mathbb{R}^d$, the edge probability matrix is 
+$P = X X^\top$ where 
 $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. The SBM is 
@@
 \cite{lyzinski2014}. It has also been shown that similar 
-random graph models, including both the MMSBM and DCBM, can be 
-represented in this way \cite{rubindelanchy2017consistency} 
-\cite{lyzinski2014}. An analogous 
+random graph models, including the DCBM, can be represented in this way
+\cite{rubindelanchy2017consistency} \cite{lyzinski2014}. An analogous 
 property exists for the PABM but not for the RDPG model but under the 
@@
 
-*Proof*. 
-This is given by straightforward matrix multiplication. It suffices to show 
-that
-
-$$X U I_{3, 1} U^\top X^\top = 
-\begin{bmatrix} 
-  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}$$
-
-**Remark**. While we can just perform the matrix multiplication to show the 
-equivalence, it is more illustrative to look at a few intermediate steps. Note 
-that the product of the three inner matrices results in a permutation matrix 
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swappoing 
-positions $2$ and $3$:
-
-$$U I_{3, 1} U^\top = \begin{bmatrix} 
-  1 & 0 & 0 & 0 \\ 
-  0 & 0 & 1 & 0 \\ 
-  0 & 1 & 0 & 0 \\ 
-  0 & 0 & 0 & 1 
-\end{bmatrix} = \Pi$$
-
-Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$ 
-is a spectral decomposition of this permutation matrix. Note that the two fixed 
-points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$ 
-where $i = 1, 4$ corresponding to the locations of the fixed points, and the 
-cycle of order two results in two eigenvalues $\pm 1$ with corresponding 
-eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is 
-swapped.
-
 **Theorem 2** (Generalization to $K > 2$). There exists a block diagonal matrix 
@@
 $A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
-
-*Proof*. 
-Let 
-$$\Lambda^{(k)} = 
-\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
-\in \mathbb{R}^{n_k \times K}$$
-
-\begin{equation}
-X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
-\end{equation}
-
-$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
-\mathbb{R}^{n \times K}$$
-
-$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
-\mathbb{R}^{n \times K^2}$$
-
-Then $P = X Y^\top$.
-
-Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
-$\Pi$, resulting in $P = X \Pi X^\top$.  
-The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
-$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
-$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
-$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
-a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
-$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
-
-Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
-eigenvalues equal to $-1$. $\Pi$ has the decomposed form
-
-\begin{equation}
-\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
-\end{equation}
-
-The edge probability matrix then can be written as:
-
-\begin{equation}
-P = X U I_{p, q} (X U)^\top
-\end{equation}
-
-\begin{equation}
-p = K (K + 1) / 2
-\end{equation}
-
-\begin{equation}
-q = K (K - 1) / 2
-\end{equation}
-
-and we can describe the PABM with $K$ communities as a GRDPG with latent 
-positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
 
@@
 \citet*{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, used 
-modularity maximization for clustering and parameter estimation. They were able 
-to show that as the sample size increases, the proportion of misclassified 
-community labels (up to permutation) goes to 0. However, because modularity 
-maximization is NP-hard, the authors of this paper used an approximation 
-based on the Extreme Points algorithm \cite{le2016}.
-
-\citet*{noroozi2019estimation} proposed using sparse subspace clustering 
-to identify the community memberships given either an edge probability matrix 
-$P$ or an adjacency matrix $A$. In the case that $P$ is known, the community 
-memberships can be identified exactly (up to permutation). This is supported by 
-our Theorem 2, which shows that the ASE of edge probability matrix $P$ consists 
-of $K$ $K$-dimensional subspaces lying in $\mathbb{R}^{K^2}$. A similar 
-procedure can be applied if $P$ is unknown and we have an observation $A$, but 
-the theoretical guarantees of this method applied to the PABM are unknown.
+a relaxed version of Modularity Maximization (MM) based on the Extreme Points
+(EP) algorithm \cite{le2016} for clustering and parameter estimation. They were
+able to show that as the sample size increases, the proportion of misclassified
+community labels (up to permutation) goes to 0.
+
+\citet*{noroozi2019estimation} proposed using Sparse Subspace Clustering (SSC)
+on the adjacency matrix $A$ or edge probability matrix $P$ to identify the
+community memberships. They were able to show that their SSC procedure on $P$
+results in perfect community detection, although they were unable to determine
+error bounds for their procedure on $A$. This is somewhat consistent with
+Theorem 2, which shows that the ASE of edge probability matrix $P$ consists of
+$K$ $K$-dimensional subspaces in $\mathbb{R}^{K^2}$.
+
+Much of the following is built on top of work from
+\citet{rubindelanchy2017statistical}, in particular, the properties of the ASE 
+of $A$.
 
 ## Community detection
-
-### Using edge probability matrix $P$
 
@@
 edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
-(or can estimate) $XU$ directly, then both the community detection and parameter 
-identification problem are trivial since $U$ is orthonormal and fixed for each 
-value of $K$. However, direct identification or estimation of $XU$ is not 
-possible \cite{rubindelanchy2017statistical}.
+(or can estimate) $XU$ directly, then both the community detection and 
+parameter identification problem are trivial since $U$ is orthonormal and 
+fixed for each value of $K$. However, direct identification or estimation of
+$XU$ is not possible \cite{rubindelanchy2017statistical}.
 
@@
 
-We can note from the explicit form of $XU$ that it represents points in 
-$\mathbb{R}^{K^2}$ such that points within each community lie on $K$-dimensional 
-subspaces. Furthermore, the subspaces are orthogonal to each other. 
+The explicit form of $XU$ represents points in 
+$\mathbb{R}^{K^2}$ such that points within each community lie on 
+$K$-dimensional orthogonal subspaces. 
 Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
-but retains the property that each community is represented by a $K$-dimensional 
-subspace. Using this property, previous work proposes the use of subspace 
-clustering \cite{noroozi2019estimation} \cite{soltanolkotabi2014}.
-
-**Lemma 1.** 
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in equation (3).
-
-*Proof*. 
-By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in equation (3) and $p$ and $q$ are defined as in equations (6) and (7).
-Alternatively, the spectral decomposition can be written as $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ and $|\cdot|^{1/2}$ is applied entry-wise.
-Thus for some $Q \in \mathbb{O}(p, q)$,
-
-$$X U Q = V |D|^{1/2}$$
-
-Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,
-
-$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$
-
-The righthand side becomes
-
-$$\begin{split}
-  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-    = X U Q Q^{-1} U^\top (X^\top X)^{-1} U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-  & = X (X^\top X)^{-1} X^\top
-\end{split}$$
-
-The lefthand side becomes:
-
-$$\begin{split}
-  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-  & = V V^\top
-\end{split}$$
+but retains the property that each community is represented by a 
+$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
+correspond to each community, suggesting the use of SSC. In this paper, we will 
+use a different and leave the properties of SSC on the ASE of $A$ as future
+work.
 
@@
 matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
-of different communities.
-
-*Proof*.
-By Lemma 1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
-in Theorem 2. Since $X$ is block diagonal with each block corresponding 
-to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
-with each block corresponding to a community and zeros elsewhere. Therefore, 
-if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
-element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
-
-\begin{algorithm}[H]
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \KwData{Edge probability matrix $P$, number of communities $K$}
-  \KwResult{Community labels $1, ..., K$}
-    Compute the spectral decomposition $P = V D V^\top$.\;
-    Compute the inner product matrix $B = |V V^\top|$, applying $|\cdot|$ 
-    entry-wise.\;
-    Construct graph $G$ using $B$ as edge similarities.\;
-    Identify the connected components of $G$ and map each to community labels 
-    $1, ..., K.$\;
-  \caption{PABM clustering on the edge probability matrix.}
-\end{algorithm}
-
-### Using adjacency matrix $A$
-
-The adjacency embedding of $A$ approaches latent positions that form $P$ as 
+from different communities.
+
+Theorem 3 provides perfect community detection given $P$. Letting $|B|$ be the 
+affinity matrix for graph $G$, $G$ is partitioned into $K$ disjoint subgraphs 
+that correspond to each community. Using $A$ introduces some error in the ASE, 
+resulting in points that lie near subspaces but not exactly on them. Using 
+results from \cite{rubindelaunchy2017statistical}, we can show that as $n$ 
+grows, the ASE of $A$ approaches the ASE of $P$ and $B_{ij} \to 0$ for $i$ and 
+$j$ in different clusters.
+
+The ASE of $A$ approaches latent positions that form $P$ as 
 the number of vertices $n$ increases. More 
-precisely, let $\{\lambda^{(kl)}\}_K \sim \mathcal{F}_K$ for some joint 
-distribution consisting of $K$ underlying communities $\mathcal{F}_K$. Then the 
-latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution 
-with $K$ underlying communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size 
-$n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge 
-probability matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the 
-adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
-\citet{rubindelanchy2017statistical},
+precisely, let each $\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and $k, l = 1, ..., K$.
+Then the corresponding latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying communities $\mathcal{G}_K$.
+Denote $Z_n$ as a sample of size $n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability matrix $P_n = Z_n I_{pq} Z_n^\top$.
+Let $\hat{Z}_n$ be the adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$.
+Then by \citet{rubindelanchy2017statistical},
 
@@
 
-for some $c>0$ and sequence of $Q_n \in \mathbb{O}(p, q)$. In addition, 
-\citeauthor{rubindelanchy2017statistical} produce a central limit theorem 
-result.
+for some $c > 0$ and sequence of $Q_n \in \mathbb{O}(p, q)$.
 
@@
 $(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
-different communities. Then for some $c > 0$,
+different communities. Let $n \rho_n = \omega((\log n)^{4c}))$ as described in
+\cite{rubindelanchy2017statistical}. Then for some $c > 0$,
 
@@
 
-where $\lim\limits_{n \to 0} \rho_n = 0$ is a sparsity parameter introduced 
-by \citet{307cbeb9b1be48299388437423d94bf1}. More precisely, 
-$\rho_n = o(\log n / \sqrt{n})$ \cite{rubindelanchy2017statistical}.
-
-This leads to the following algorithm:
+Thus, as $n \to \infty$, the partitioning of $\hat{B}^{(n)}$ results in 
+perfect community detection. 
+This leads to the Orthogonal Spectral Clustering (OSE) algorithm (Alg. 1).
 
@@
     Map each partition to the community labels $1, ..., K$.\;
-  \caption{PABM clustering on the adjacency matrix.}
+  \caption{Orthogonal Spectral Clustering.}
 \end{algorithm}
 
-Theorem 4 implies that as $n \to \infty$, the number (not just proportion) of 
-misclassified vertices, up to permutation, outputted by algorithm 2 goes to 0.
-
 ## Parameter estimation
 
-For any $P$ edge probability matrix for the PABM such that the rows and 
+For any edge probability matrix $P$ for the PABM such that the rows and 
 columns are organized by community, the $kl$^th^ block is an outer product 
 of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
-Therefore, given $P^{(kl)}$, $\lambda%{(kl)}$ and $\lambda^{(lk)}$ are 
+Therefore, given $P^{(kl)}$, $\lambda^{(kl)}$ and $\lambda^{(lk)}$ are 
 solvable exactly (up to multiplication by $-1$) using singular value 
-decomposition. More specifically, let $P = \sigma^2 u v^\top$ be the 
-singular value decomposition of $P$. $u \in \mathbb{R}^{n_k}$ and 
+decomposition. More specifically, let $P^{(kl)} = \sigma^2 u v^\top$ be the 
+singular value decomposition of $P^{(kl)}$. $u \in \mathbb{R}^{n_k}$ and 
 $v \in \mathbb{R}^{n_l}$ are vectors and 
@@
 $\lambda^{(lk)} = \pm \sigma v$. 
-
-\begin{algorithm}[H]
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{PABM parameter estimation using the edge probability matrix.}
-  \KwData{Edge probability matrix $P$, community assignments $1, ..., K$.}
-  \KwResult{PABM parameters $\{\lambda^{(kl)}\}_K$}
-  Arrange the rows and columns of $P$ by community such that each $P^{(kl)}$ 
-  block consists of edge probabilities between communities $k$ and $l$.\; 
-  \For{$k, l = 1, ..., K$, $k \geq l$} {
-    Compute $P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top$, the SVD of 
-    the $kl^{th}$ block.\;
-    Assign $\lambda^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
-    $\lambda^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
-  }
-\end{algorithm}
-
 Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
-simply use plug-in estimators:
+simply use plug-in estimators (Alg. 2), which converge to the true parameters 
+(Theorem 5).
 
@@
 
-*Proof*. 
-Let $P$ and $A$ be organized by community such that the elements of blocks 
-$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
-$l$.
-
-*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
-probabilities and edges for community $k$.  
-By definition, 
-$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
-signular value decomposition 
-$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
-one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
-singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
-Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
-value decomposition of $A^{(kk)}$, and let 
-$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
-one-dimensional approximation. Define 
-$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
-$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
-$\lambda^{(kk)}$.  
-Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
-spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
-$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
-
-*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
-edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
-By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
-$k = l$ case, we note that the singular value decomposition 
-$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
-$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
-$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
-$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
-$v^{(kl)} = u^{(lk)}$.)  
-Now consider the Hermitian dilation
-
-$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$
-
-which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
-that the eigendecomposition of $M^{(kl)}$ is
-
-$$M^{(kl)} = 
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
-\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$
-
-Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
-latent positions in $\mathbb{R}^2$ given by
-
-$$\begin{bmatrix} 
-  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
-  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
-\end{bmatrix} = 
-\begin{bmatrix} 
-  \lambda^{(kl)} & \lambda^{(kl)} \\ 
-  \lambda^{(lk)} & -\lambda^{(lk)} 
-\end{bmatrix}$$
-
-Now consider
-
-$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
-Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where
-
-$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$
-
-and $E$ is the 
-$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
-in $A^{(kl)}$.  
-Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
-spectral embedding, given by
-
-$$\begin{bmatrix} 
-  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
-  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
-\end{bmatrix}$$
-
-where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
-the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
-by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
-In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
-result \cite{rubindelanchy2017statistical} is of the form 
-$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
-for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
-resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
-a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
-up to sign flip.
-
 # Simulated Examples
 
-For each simulation, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
-from a joint distribution, the edge probability matrix $P$ is constructed using 
+For each simulation, community labels are drawn from a multinomial 
+distribution, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
+from two types of joint distributions depending on whether $k = l$, the edge
+probability matrix $P$ is constructed using 
 the popularity vectors, and finally an unweighted and undirected adjacency 
-matrix $A$ is drawn from $P$. Algorithm 2 is then used for community detection, 
-and this method is compared against sparse subspace clustering 
-\cite{soltanolkotabi2014} \cite{noroozi2019estimation} and modularity 
-maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}. True 
-community labels are used with Algorithm 4 to estimate the popularity 
-vectors, and this method is then compared against an MLE-based estimator 
-described in \citet{noroozi2019estimation} and
+matrix $A$ is drawn from $P$. OSE is then used for community detection, 
+and this method is compared against SSC \cite{noroozi2019estimation}
+\cite{soltanolkotabi2014} and MM \cite{igraph}
+\cite{307cbeb9b1be48299388437423d94bf1}. True 
+community labels are used with Algorithm 2 to estimate the popularity 
+vectors $\{\lambda^{(kl)}\}_K$, and this method is then compared against an
+MLE-based estimator described in \citet{noroozi2019estimation} and
 \citet{307cbeb9b1be48299388437423d94bf1}.
 
-Sparse subspace clustering is described as follows. First, an embedding of the 
-affinity matrix $A$ is required, and for this, we use spectral decomposition 
-using the first $K (K + 1) / 2$ and last $K (K - 1) / 2$ eigenvectors. Then a 
-sparse linear regression model is fit to each row of the embedding data matrix 
-using the remaining rows as predictors, the rows of which are stored in 
-matrix $C$. Finally, an affinity matrix $B = |C| + |C|^\top$ is constructed, 
-where $|\cdot|$ is applied entry-wise. If the data in the embedding lie 
-perfectly in $K$ subspaces, then $B_{ij} = 0$ for $i, j$ not in the same 
-subspace. If the embedding of $P$ is used, then this property holds, but it 
-is true in general for the embedding of $A$, in which case a final clustering 
-step is required. For these simulations, spectral clustering with the 
-normalized Laplacian and Gaussian mixture models is used for this final step.
-
-Modularity maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
+Modularity Maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
 used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$. 
@@
 the Louvain algorithm \cite{Blondel_2008} was used instead.
+
+Two implementations of SSC are shown here. The first method, denoted as 
+SSC-A, treats the columns of the adjacency matrix $A$ as points in
+$\mathbb{R}^n$. The second method, denoted as SSC-ASE, first embeds $A$ and 
+then performs SSC on the embedding.
+
+SSC is performed by solving an optimization problem for each observed point.
+Given $X \in \mathbb{R}^{n \times d}$ with vectors 
+$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
+$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
+for each $i = 1, ..., n$. The solutions are collected collected into matrix 
+$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
+affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
+subspaces, $B$ describes an undirected graph consisting of $K$ subgraphs, i.e.,
+$B_{ij} = 0$ if $i, j$ are in different subspaces. 
+If $X$ instead represents points near $K$ subspaces with some noise, a final 
+graph partitioning step is performed (e.g., edge thresholding or spectral 
+clustering). For our simulations, we used spectral clustering by fitting a 
+Gaussian Mixture Model on the $K$-dimensional Laplacian eigenmap of $B$.
+
+In practice, this is often done by solving the LASSO problems
+
+\begin{equation}
+\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
+\end{equation}
+
+for some sparsity parameter $\lambda > 0$. For our simulations, $\lambda$ was 
+chosen via a preliminary cross validation experiment.
 
@@
 
-Fig \ref{fig:clust_err_k} show that for large $n$, Algorithm 2 results in 
+Fig. \ref{fig:clust_err_k} show that for large $n$, OSE results in 
 a misclustering rate of 0. Sparse subspace clustering produces similar results 
@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+```{r clust_err_k, fig.cap = 'IQR of clustering error using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
@@
 
-Theorem 4 implies that algorithm 2 will result in not just in the error rate 
-converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
-We explore this in Fig \ref{fig:clust_err_ct_sim}.
-
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+Theorem 4 implies that OSE will result in not just in the error rate 
+converging to 0 but the error *count* as well.
+We explore this in Fig. \ref{fig:clust_err_ct_sim}.
+
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
@@
 
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 2 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 
-Fig \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
-similar results as in the balanced communities case, with both Algorithm 2 and 
-sparse subspace clustering resulting in no misclustered vertices for a 
-sufficiently large sample size. However, Fig \ref{fig:lambda_est_k_imba} 
-suggests that while Algorithm 5 retains $\sqrt{n}$ efficiency, the MLE-based 
-plug-in estimator fails to do so.
-
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4.
+Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
+similar results as in the balanced communities case, with both OSE and 
+SSC resulting in no misclustered vertices for a 
+sufficiently large sample size. However, Fig. \ref{fig:lambda_est_k_imba} 
+suggests that while Algorithm 2 retains $\sqrt{n}$ efficiency, the MLE-based 
+plug-in estimator is less efficient for this setup.
+
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
@@
 
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4.
+```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
@@
 
-```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4.
+```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 2 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 
-In the first real data example, we applied Algorithm 2 to the Leeds Butterfly 
+In the first real data example, we applied OSE to the Leeds Butterfly 
 dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
@@
 
-Comparing against the ground truth species labels, Algorithm 2 achieves an 
+Comparing against the ground truth species labels, OSE achieves an 
 accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
@@
 
-```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = FALSE}
+```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSE.', cache = FALSE}
 plot.A(A, z.hat)
@@
 
-In the second example, we applied Algorithm 2 to the British MPs Twitter 
+In the second example, we applied OSE to the British MPs Twitter 
 network \cite{greene2013producing}, the Political Blogs network 
@@
 although performance is comparable. In 
-addition, Algorithm 2's runtime is much lower than that of modularity 
+addition, OSE's runtime is much lower than that of modularity 
 maximization.
@@
                                           dblp.acc.ssc), 3),
-              `Error (Alg. 2)` = round(1 - c(bmp.acc, 
+              `Error (OSE)` = round(1 - c(bmp.acc, 
                                                   pb.acc,
@@
                                'sparse subspace clustering, ',
-                               'and Alg. 2.'))
-```
-
-```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
+                               'and OSE.'))
+```
+
+```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSE.', fig.dpi = 10}
 gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
@@
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left).
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSE (left).
 The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
@@
 knitr::kable(dplyr::tibble(
-  Method = c('Alg. 2', 'SSC', 'Mod. Max.'),
+  Method = c('OSE', 'SSC', 'Mod. Max.'),
   `Error (Village 12)` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
@@
 
+# Proofs
+
+**Proof of Theorem 1**. 
+This is given by straightforward matrix multiplication. It suffices to show 
+that
+
+$$X U I_{3, 1} U^\top X^\top = 
+\begin{bmatrix} 
+  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
+  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
+\end{bmatrix}$$
+
+*Remark*. While we can just perform the matrix multiplication to show the 
+equivalence, it is more illustrative to look at a few intermediate steps. Note 
+that the product of the three inner matrices results in a permutation matrix 
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swappoing 
+positions $2$ and $3$:
+
+$$U I_{3, 1} U^\top = \begin{bmatrix} 
+  1 & 0 & 0 & 0 \\ 
+  0 & 0 & 1 & 0 \\ 
+  0 & 1 & 0 & 0 \\ 
+  0 & 0 & 0 & 1 
+\end{bmatrix} = \Pi$$
+
+Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$ 
+is a spectral decomposition of this permutation matrix. Note that the two fixed 
+points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$ 
+where $i = 1, 4$ corresponding to the locations of the fixed points, and the 
+cycle of order two results in two eigenvalues $\pm 1$ with corresponding 
+eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is 
+swapped.
+
+**Proof of Theorem 2**. 
+Let 
+$$\Lambda^{(k)} = 
+\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
+\in \mathbb{R}^{n_k \times K}$$
+
+\begin{equation}
+X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
+\end{equation}
+
+$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
+\mathbb{R}^{n \times K}$$
+
+$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
+\mathbb{R}^{n \times K^2}$$
+
+Then $P = X Y^\top$.
+
+Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
+$\Pi$, resulting in $P = X \Pi X^\top$.  
+The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
+$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
+$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
+$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
+a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
+$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
+
+Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
+eigenvalues equal to $-1$. $\Pi$ has the decomposed form
+
+\begin{equation}
+\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
+\end{equation}
+
+The edge probability matrix then can be written as:
+
+\begin{equation}
+P = X U I_{p, q} (X U)^\top
+\end{equation}
+
+\begin{equation}
+p = K (K + 1) / 2
+\end{equation}
+
+\begin{equation}
+q = K (K - 1) / 2
+\end{equation}
+
+and we can describe the PABM with $K$ communities as a GRDPG with latent 
+positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
+
+**Lemma 1.** 
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
+matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
+defined as in equation (6).
+
+**Proof of Lemma 1**. 
+By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in equation (6) and $p$ and $q$ are defined as in equations (9) and (10).
+Alternatively, the spectral decomposition can be written as $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ and $|\cdot|^{1/2}$ is applied entry-wise.
+Thus for some $Q \in \mathbb{O}(p, q)$,
+
+$$X U Q = V |D|^{1/2}$$
+
+Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,
+
+$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
+(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$
+
+The righthand side becomes
+
+$$\begin{split}
+  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
+    = X U Q Q^{-1} U^\top (X^\top X)^{-1} U (Q^\top)^{-1} Q^\top U^\top X^\top \\
+  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
+  & = X (X^\top X)^{-1} X^\top
+\end{split}$$
+
+The lefthand side becomes:
+
+$$\begin{split}
+  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
+    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
+  & = V V^\top
+\end{split}$$
+
+**Proof of Theorem 3**.
+By Lemma 1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
+in Theorem 2. Since $X$ is block diagonal with each block corresponding 
+to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
+with each block corresponding to a community and zeros elsewhere. Therefore, 
+if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
+element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
+
+**Proof of Theorem 5**. 
+Let $P$ and $A$ be organized by community such that the elements of blocks 
+$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
+$l$.
+
+*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
+probabilities and edges for community $k$.  
+By definition, 
+$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
+signular value decomposition 
+$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
+one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
+singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
+Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
+value decomposition of $A^{(kk)}$, and let 
+$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
+one-dimensional approximation. Define 
+$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
+$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
+$\lambda^{(kk)}$.  
+Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
+spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
+$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
+
+*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
+edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
+By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
+$k = l$ case, we note that the singular value decomposition 
+$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
+$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
+$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
+$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
+$v^{(kl)} = u^{(lk)}$.)  
+Now consider the Hermitian dilation
+
+$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$
+
+which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
+that the eigendecomposition of $M^{(kl)}$ is
+
+$$M^{(kl)} = 
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
+\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$
+
+Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
+latent positions in $\mathbb{R}^2$ given by
+
+$$\begin{bmatrix} 
+  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
+  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
+\end{bmatrix} = 
+\begin{bmatrix} 
+  \lambda^{(kl)} & \lambda^{(kl)} \\ 
+  \lambda^{(lk)} & -\lambda^{(lk)} 
+\end{bmatrix}$$
+
+Now consider
+
+$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
+Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where
+
+$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$
+
+and $E$ is the 
+$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
+in $A^{(kl)}$.  
+Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
+spectral embedding, given by
+
+$$\begin{bmatrix} 
+  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
+  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
+\end{bmatrix}$$
+
+where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
+the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
+by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
+In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
+result \cite{rubindelanchy2017statistical} is of the form 
+$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
+for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
+resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
+a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
+up to sign flip.
+
 # References

### 2021-02-09 summary.Rmd
minor tweaks

@@
 affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
-subspaces, $B$ describes an undirected graph consisting of $K$ subgraphs, i.e.,
-$B_{ij} = 0$ if $i, j$ are in different subspaces. 
+subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
+subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
 If $X$ instead represents points near $K$ subspaces with some noise, a final 

### 2021-02-09 summary.Rmd
ssc stuff

@@
 resulting in points that lie near subspaces but not exactly on them. Using 
-results from \cite{rubindelaunchy2017statistical}, we can show that as $n$ 
+results from \cite{rubindelanchy2017statistical}, we can show that as $n$ 
 grows, the ASE of $A$ approaches the ASE of $P$ and $B_{ij} \to 0$ for $i$ and 

### 2021-02-09 summary.Rmd
fixed typos

@@
 
-Given ground truth community labels, Algorithm 5 and the MLE-based plug-in 
+Given ground truth community labels, Algorithm 2 and the MLE-based plug-in 
 estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
@@
   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 4'), width = .1) + 
+                    colour = 'Algorithm 2'), width = .1) + 
   geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
@@
   ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 2')) + 
   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 4'), width = .1) + 
+                    colour = 'Algorithm 2'), width = .1) + 
   geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
@@
 balanced communities examples for $K = 2$, we generated one instance of $A$ for
-each $n$ and constructed $B$ according to algorithm 4 to verify that 
+each $n$ and constructed $B$ according to algorithm 2 to verify that 
 as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
@@
 
-**Proof of Lemma 1**. 
-By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in equation (6) and $p$ and $q$ are defined as in equations (9) and (10).
-Alternatively, the spectral decomposition can be written as $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ and $|\cdot|^{1/2}$ is applied entry-wise.
-Thus for some $Q \in \mathbb{O}(p, q)$,
+*Proof of Lemma 1*. 
+By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
+equation (6) and $p$ and $q$ are defined as in equations (9) and (10).
+Alternatively, the spectral decomposition can be written as 
+$P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
+and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some 
+$Q \in \mathbb{O}(p, q)$,
 
@@
 
-where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
+where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 2, approximates 
 the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 

### 2021-02-10 summary.Rmd
typo

@@
   ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 2')) + 
   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,

### 2021-02-16 summary.Rmd
worked on ssc stuff

@@
 different communities. Let $n \rho_n = \omega((\log n)^{4c}))$ as described in
-\cite{rubindelanchy2017statistical}. Then for some $c > 0$,
+\citet{rubindelanchy2017statistical}. Then for some $c > 0$,
 
@@
 
+Theorems 3 and 4 also provide a very natural path toward SSC. We established in 
+theorem 2 that an ASE of the edge probability matrix $P$ can be constructed 
+such that the clusters lie on mutually orthogonal subspaces, and this property 
+can be recovered from the eigenvectors of $P$. Then theorems 3 and 4 provide 
+convergence rates for the eigenvectors of the edge weight matrix $A$ drawn 
+from $P$.
+
 ## Parameter estimation
@@
 simply use plug-in estimators (Alg. 2), which converge to the true parameters 
-(Theorem 5).
+(Theorem 6).
 
@@
 
-**Theorem 5**. Under regularity and sparsity assumptions, and under the further 
+**Theorem 6**. Under regularity and sparsity assumptions, and under the further 
 assumption that $K$ is fixed and community labels are known,
@@
 **Proof of Theorem 5**.
+
+**Proof of Theorem 6**. 
 Let $P$ and $A$ be organized by community such that the elements of blocks 

### 2021-02-17 summary.Rmd
adding ssc stuff to main writeup

--- 2021-02-10 summary.Rmd
@@
 a relaxed version of Modularity Maximization (MM) based on the Extreme Points
-(EP) algorithm \cite{le2016} for clustering and parameter estimation. They were
-able to show that as the sample size increases, the proportion of misclassified
-community labels (up to permutation) goes to 0.
-
-\citet*{noroozi2019estimation} proposed using Sparse Subspace Clustering (SSC)
-on the adjacency matrix $A$ or edge probability matrix $P$ to identify the
-community memberships. They were able to show that their SSC procedure on $P$
-results in perfect community detection, although they were unable to determine
-error bounds for their procedure on $A$. This is somewhat consistent with
-Theorem 2, which shows that the ASE of edge probability matrix $P$ consists of
-$K$ $K$-dimensional subspaces in $\mathbb{R}^{K^2}$.
-
-Much of the following is built on top of work from
+(EP) algorithm \cite{le2016} for community detection and parameter estimation. 
+They were able to show that as the sample size increases, the proportion of
+misclassified community labels (up to permutation) goes to 0.
+
+\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for 
+community detection in the PABM. 
+SSC is performed by solving an optimization problem for each observed point.
+Given $X \in \mathbb{R}^{n \times d}$ with vectors 
+$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
+$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
+for each $i = 1, ..., n$. The solutions are collected collected into matrix 
+$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
+affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
+subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
+subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
+If $X$ instead represents points near $K$ subspaces with some noise, a final 
+graph partitioning step is performed (e.g., edge thresholding or spectral 
+clustering).
+
+In practice, SSC is often performed by solving the LASSO problems
+
+\begin{equation}
+\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
+\end{equation}
+
+for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
+into $C$ and $B$ as before.
+
+Theorem 2 suggests that SSC is appropriate for community detection for the 
+PABM. More precisely, theorem 2 says that each community consists of a
+$K$-dimensional subspace, and together the subspaces lie in $\mathbb{R}^{K^2}$. 
+The natural approach then is to perform SSC on the ASE of $P$ or $A$.
+\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$ themselves.
+
+Much of our work is built on top of work from 
 \citet{rubindelanchy2017statistical}, in particular, the properties of the ASE 
-of $A$.
+of $A$ in relation to the ASE of $P$ built on $\{\lambda^{(kl)}\}_K$.
 
@@
 The ASE of $A$ approaches latent positions that form $P$ as 
-the number of vertices $n$ increases. More 
-precisely, let each $\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and $k, l = 1, ..., K$.
-Then the corresponding latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying communities $\mathcal{G}_K$.
-Denote $Z_n$ as a sample of size $n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability matrix $P_n = Z_n I_{pq} Z_n^\top$.
-Let $\hat{Z}_n$ be the adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$.
-Then by \citet{rubindelanchy2017statistical},
+the number of vertices $n$ increases. More precisely, let each 
+$\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and 
+$k, l = 1, ..., K$. Then the corresponding latent positions 
+$XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying
+communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size $n$ from
+$\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability 
+matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the adjacency embedding 
+of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
+\citet{rubindelanchy2017statistical},
 
@@
     Construct graph $G$ using $B$ as its similarity matrix.\;
-    Partition $G$ into $K$ disconnected components 
+    Partition $G$ into $K$ disconnected subgraphs  
     (e.g., using edge thresholding or spectral clustering).\;
@@
 \end{algorithm}
+
+Theorems 2, 3, and 4 also provide a very natural path toward using SSC for 
+community detection for the PABM. We established in theorem 2 that an ASE of 
+the edge probability matrix $P$ can be constructed such that the communities 
+lie on mutually orthogonal subspaces, and this property can be recovered from
+the eigenvectors of $P$. Then theorems 3 and 4 show that this property holds 
+for the adjacency matrix $A$ drawn from $P$ as $n \to \infty$.
+
+\begin{algorithm}[H]
+  \DontPrintSemicolon
+  \SetAlgoLined
+  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
+  \KwData{Adjacency matrix $A$, number of communities $K$, 
+  hyperparameter $\lambda$}
+  \KwResult{Community assignments $1, ..., K$}
+    Find $Z$, the ASE of $A$ using the $K (K + 1) / 2$ most positive 
+    and the $K (K - 1) / 2$ most negative eigenvalues and eigenvectors of 
+    $A$.\;
+    \For {$i = 1, ..., n$} {
+      Assign $z_i^\top$ as the $i^{th}$ row of $Z$. 
+      Assign $Z_{-i} = \begin{bmatrix} 
+      z_1 & \cdots & z_{i-1} & z_{i+1} & \cdots & z_n \end{bmatrix}^\top$.\;
+      Solve the LASSO problem 
+      $c_i = \arg\min_{\beta_i} 
+      \frac{1}{2} ||z_i - Z_{-i} \beta_i||_2^2 + \lambda ||\beta_i||_1$.\;
+      Assign $\tilde{c}_i = \begin{bmatrix} 
+      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
+      \end{bmatrix}^\top$ such that the superscript is the index of 
+      $\tilde{c}_i$.\;
+    }
+    Assign 
+    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
+    Compute $B = |C| + |C^\top|$.\;
+    Construct graph $G$ using $B$ as its similarity matrix.\;
+    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
+    thresholding or spectral clustering).\;
+    Map each partition to the community labels $1, ..., K$. 
+\end{algorithm}
+
+**Theorem 5**. Let $P_n$ describe the edge probability matrix of the PABM with 
+$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$. Let $Z_n$ be the ASE of 
+$P_n$, and let $Z_n^{(k)}$ correspond to the embedding points that correspond 
+to the $k$^th^ block. Let $\hat{Z}_n$ be the ASE of $A_n$ and $\hat{Z}_n^{(k)}$ 
+be the corresponding $k$^th^ block. If $r(Z_n^{(k)})$, the inradius of 
+$Z_n^{(k)}$ (as defined by \citet{soltanolkotabi2012}), is positive for each 
+$k$, then the SSC detection property holds for $\hat{Z}_n$ with probability 1 
+as $n \to \infty$.
 
@@
 Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
-simply use plug-in estimators (Alg. 2), which converge to the true parameters 
-(Theorem 5).
+simply use plug-in estimators (algorithm 3), which converge to the true 
+parameters (theorem 6).
 
@@
   communities $k$ and $l$.\;
-  \For{$k, l = 1, ..., K$, $k \leq l$} {
+  \For {$k, l = 1, ..., K$, $k \leq l$} {
     Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
@@
 \cite{307cbeb9b1be48299388437423d94bf1}. True 
-community labels are used with Algorithm 2 to estimate the popularity 
+community labels are used with Algorithm 3 to estimate the popularity 
 vectors $\{\lambda^{(kl)}\}_K$, and this method is then compared against an
@@
 SSC-A, treats the columns of the adjacency matrix $A$ as points in
-$\mathbb{R}^n$. The second method, denoted as SSC-ASE, first embeds $A$ and 
-then performs SSC on the embedding.
-
-SSC is performed by solving an optimization problem for each observed point.
-Given $X \in \mathbb{R}^{n \times d}$ with vectors 
-$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
-$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
-for each $i = 1, ..., n$. The solutions are collected collected into matrix 
-$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
-affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
-subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
-subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
-If $X$ instead represents points near $K$ subspaces with some noise, a final 
-graph partitioning step is performed (e.g., edge thresholding or spectral 
-clustering). For our simulations, we used spectral clustering by fitting a 
-Gaussian Mixture Model on the $K$-dimensional Laplacian eigenmap of $B$.
-
-In practice, this is often done by solving the LASSO problems
-
-\begin{equation}
-\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
-\end{equation}
-
-for some sparsity parameter $\lambda > 0$. For our simulations, $\lambda$ was 
-chosen via a preliminary cross validation experiment.
+$\mathbb{R}^n$, as described in \citet{noroozi2019estimation}. The second 
+method, denoted as SSC-ASE, first embeds $A$ and then performs SSC on the 
+embedding, as described in algorithm 2. The sparsity parameter $\lambda$ was 
+chosen via a preliminary cross-validation experiment. For the final clustering 
+step, a Gaussian Mixture Model was fit on the normalized Lapacian eigenmap of 
+the affinity matrix $B$.
 
@@
 
-Given ground truth community labels, Algorithm 2 and the MLE-based plug-in 
+Given ground truth community labels, Algorithm 3 and the MLE-based plug-in 
 estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
@@
 
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 2 (red) compared against an MLE-based method (blue).
+```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
   ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 2')) + 
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
   geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 2'), width = .1) + 
+                    colour = 'Algorithm 3'), width = .1) + 
   geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
@@
 sufficiently large sample size. However, Fig. \ref{fig:lambda_est_k_imba} 
-suggests that while Algorithm 2 retains $\sqrt{n}$ efficiency, the MLE-based 
+suggests that while Algorithm 3 retains $\sqrt{n}$ efficiency, the MLE-based 
 plug-in estimator is less efficient for this setup.
@@
 
-```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 2 (red) compared against an MLE-based method (blue).
+```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.', fig.width = 10}
@@
 balanced communities examples for $K = 2$, we generated one instance of $A$ for
-each $n$ and constructed $B$ according to algorithm 2 to verify that 
+each $n$ and constructed $B$ according to Algorithm 3 to verify that 
 as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
@@
 matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in equation (6).
+defined as in equation (7).
 
@@
 By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
-equation (6) and $p$ and $q$ are defined as in equations (9) and (10).
+equation (7) and $p$ and $q$ are defined as in equations (10) and (11).
 Alternatively, the spectral decomposition can be written as 
@@
 
-where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 2, approximates 
+where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 3, approximates 
 the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 

### 2021-02-19 summary.Rmd
very rough proofs

@@
 Theorem 2 suggests that SSC is appropriate for community detection for the 
-PABM. More precisely, theorem 2 says that each community consists of a
+PABM. More precisely, Theorem 2 says that each community consists of a
 $K$-dimensional subspace, and together the subspaces lie in $\mathbb{R}^{K^2}$. 
@@
 Theorems 2, 3, and 4 also provide a very natural path toward using SSC for 
-community detection for the PABM. We established in theorem 2 that an ASE of 
+community detection for the PABM. We established in Theorem 2 that an ASE of 
 the edge probability matrix $P$ can be constructed such that the communities 
 lie on mutually orthogonal subspaces, and this property can be recovered from
-the eigenvectors of $P$. Then theorems 3 and 4 show that this property holds 
+the eigenvectors of $P$. Then Theorems 3 and 4 show that this property holds 
 for the adjacency matrix $A$ drawn from $P$ as $n \to \infty$.
@@
 simply use plug-in estimators (algorithm 3), which converge to the true 
-parameters (theorem 6).
+parameters (Theorem 6).
 
@@
 
-**Lemma 1.** 
+**Lemma 3.1.** 
 Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
@@
 
-*Proof of Lemma 1*. 
+*Proof of Lemma 3.1*. 
 By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
@@
 **Proof of Theorem 3**.
-By Lemma 1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
+By Lemma 3.1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
 in Theorem 2. Since $X$ is block diagonal with each block corresponding 
@@
 
+**Lemma 5.1**. 
+Let $P_n = V_n \Lambda_n V_n^\top$ be defined as in Lemma 5.1. Let 
+$A_n \sim \Bernoulli(P_n)$ be the adjacency matrix drawn from $P_n$, and let 
+$A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be its approximate spectral
+decomposition using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ most 
+negative eigenvalues and their corresponding eigenvectors.
+Without loss of generality, order the rows of $V_n$ and $\hat{V}_n$ by 
+community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the $n_k \times K^2$ 
+matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
+respectively. Let $r_n^{(k)} = r(V_n^{(k)})$ and 
+$\hat{r}_n^{(k)} = r(\hat{V}_n^{(k)})$ be the inradius of points in $V_n$ and
+$\hat{V}_n$ as defined by \citet{soltanolkotabi2012} and 
+\citet{pmlr-v28-wang13}. Then
+
+\begin{equation}
+\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
+\end{equation}
+
+*Proof of Lemma 5.1*. 
+This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
+Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is smooth, 
+$\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}$.
+
+**Lemma 5.2**. 
+Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
+as defined by \citet{soltanolkotabi2012} (also note that 
+$\mu(\hat{V}_n^{(k)}) = \mu(V_n^{(k)})$ using the extension of this definition 
+by \citet{pmlr-v28-wang13}).
+
+\begin{equation}
+\mu_n^{(k)} = 0
+\end{equation}
+
+with probability 1 $\forall n, k$.
+
+*Proof of Lemma 5.2*. 
+Since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
+this falls directly out of Theorem 2.8 of \citet{soltanolkotabi2012}.
+
+**Lemma 5.3**. 
+Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
+$\hat{V}_n$ respectively. By \cite{rubindelanchy2017statistical},
+
+\begin{equation}
+\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
+\end{equation}
+
 **Proof of Theorem 5**. 
+For ease of proving, we will consider $V_n$ and $\hat{V}_n$ instead of $Z_n$ 
+and $\hat{Z}_n$.
+
+By (12) and (13),
+
+\begin{equation}
+r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r_n^{(k)} \stackrel{p}{\to} r_n^{(k)} > 0
+\end{equation}
+
+Then by Theorem 6 of \citet{pmlr-v28-wang13}, (14) and (15) imply that  
+$\Pr(\lim\limits_{n \to \infty} \text{ SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.
 

### 2021-02-19 summary.Rmd
fixed some of the proofs

@@
 
-\begin{equation}
+\begin{equation} \label{eq:pabm}
   P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
@@
 
-\begin{equation}
+\begin{equation} \label{eq:grdpg}
   P = X I_{pq} X^\top
@@
 
-\begin{equation}
+\begin{equation} \label{eq:ssc}
 \min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
@@
 
-\begin{equation}
+\begin{equation} \label{eq:thm3}
   \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:thm4}
   \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:thm6}
 \max_{k, l \in \{1, ..., K\}} 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:xy}
 X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
@@
 
-\begin{equation}
+\begin{equation} \label{eq:permutation}
 \Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
@@
 
-\begin{equation}
+\begin{equation} \label{eq:pabm-grdpg}
 P = X U I_{p, q} (X U)^\top
@@
 
-\begin{equation}
+\begin{equation} \label{eq:p}
 p = K (K + 1) / 2
@@
 
-\begin{equation}
+\begin{equation} \label{eq:q}
 q = K (K - 1) / 2
@@
 matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in equation (7).
+defined as in (\ref{eq:xy}).
 
@@
 By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
-equation (7) and $p$ and $q$ are defined as in equations (10) and (11).
-Alternatively, the spectral decomposition can be written as 
+(\ref{eq:xy}) and $p$ and $q$ are defined as in equations (\ref{eq:p}) and
+(\ref{eq:q}). Alternatively, the spectral decomposition can be written as 
 $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:r-conv}
 \hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
@@
 This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
-Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is smooth, 
-$\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}$.
+Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is 
+continuous, $r(\hat{V}_n^{(k)}) \stackrel{p}{\to} r(V_n^{(k)})$.
 
@@
 Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
-as defined by \citet{soltanolkotabi2012} (also note that 
-$\mu(\hat{V}_n^{(k)}) = \mu(V_n^{(k)})$ using the extension of this definition 
-by \citet{pmlr-v28-wang13}).
-
-\begin{equation}
-\mu_n^{(k)} = 0
-\end{equation}
-
+as defined by \citet{soltanolkotabi2012}. Then 
+$\mu_n^{(k)} = 0$ 
 with probability 1 $\forall n, k$.
@@
 **Lemma 5.3**. 
+Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the extended definition of
+$\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then
+
+\begin{equation} \label{eq:mu-conv}
+\hat{\mu}_n^{(k)} \stackrel{p}{\to} 0
+\end{equation}
+
+*Proof of Lemma 5.3*. 
+This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
+Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
+continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} \mu(V_n^{(k)}) = 0$.
+
+**Lemma 5.4**. 
 Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-$\hat{V}_n$ respectively. By \cite{rubindelanchy2017statistical},
-
-\begin{equation}
+$\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},
+
+\begin{equation} \label{eq:delta-conv}
 \delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
@@
 
-By (12) and (13),
-
-\begin{equation}
-r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r_n^{(k)} \stackrel{p}{\to} r_n^{(k)} > 0
-\end{equation}
-
-Then by Theorem 6 of \citet{pmlr-v28-wang13}, (14) and (15) imply that  
-$\Pr(\lim\limits_{n \to \infty} \text{ SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.
+By (\ref{eq:r-conv}) and (\ref{eq:mu-conv}),
+
+\begin{equation} \label{eq:thm5-proof}
+r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} 
+r_n^{(k)} - \mu_n^{(k)} = r_n^{(k)} > 0
+\end{equation}
+
+for each $k = 1, ..., K$.
+
+Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
+(\ref{eq:thm5-proof}) imply that $\forall \epsilon > 0$, $\exists N > 0$ such 
+that $\forall n > N$, 
+$\Pr(\text{SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.
 

### 2021-02-19 summary.Rmd
fixed some of the wording for theorem 5

@@
     $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
-    Compute $B = |C| + |C^\top|$.\;
+    Compute the affinity matrix $B = |C| + |C^\top|$.\;
     Construct graph $G$ using $B$ as its similarity matrix.\;
@@
 $P_n$, and let $Z_n^{(k)}$ correspond to the embedding points that correspond 
-to the $k$^th^ block. Let $\hat{Z}_n$ be the ASE of $A_n$ and $\hat{Z}_n^{(k)}$ 
-be the corresponding $k$^th^ block. If $r(Z_n^{(k)})$, the inradius of 
-$Z_n^{(k)}$ (as defined by \citet{soltanolkotabi2012}), is positive for each 
-$k$, then the SSC detection property holds for $\hat{Z}_n$ with probability 1 
-as $n \to \infty$.
+to the $k$^th^ community. Let $\hat{Z}_n$ be the ASE of $A_n$ and $\hat{Z}_n^{(k)}$ 
+be the corresponding $k$^th^ block.
+Let $\hat{B}_n$ be the affinity matrix in the SSC algorithm with input $\hat{Z}_n$ such that $\hat{B}_n^{(ij)}$ is the $ij$^th^ element of $\hat{B}_n$.
+If $r(Z_n^{(k)})$, the inradius of $Z_n^{(k)}$ (as defined by \citet{soltanolkotabi2012}), is positive for each $k$, then $\hat{B}_n^{(ij)} \stackrel{p}{\to} 0$ $\forall i, j$ in different communities.
 

### 2021-02-23 summary.Rmd
fixed typo

@@
 perfect community detection. 
-This leads to the Orthogonal Spectral Clustering (OSE) algorithm (Alg. 1).
+This leads to the Orthogonal Spectral Clustering (OSC) algorithm (Alg. 1).
 
@@
 the popularity vectors, and finally an unweighted and undirected adjacency 
-matrix $A$ is drawn from $P$. OSE is then used for community detection, 
+matrix $A$ is drawn from $P$. OSC is then used for community detection, 
 and this method is compared against SSC \cite{noroozi2019estimation}
@@
 
-Fig. \ref{fig:clust_err_k} show that for large $n$, OSE results in 
+Fig. \ref{fig:clust_err_k} show that for large $n$, OSC results in 
 a misclustering rate of 0. Sparse subspace clustering produces similar results 
@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_k, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
@@
 
-Theorem 4 implies that OSE will result in not just in the error rate 
+Theorem 4 implies that OSC will result in not just in the error rate 
 converging to 0 but the error *count* as well.
@@
 
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
@@
 Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
-similar results as in the balanced communities case, with both OSE and 
+similar results as in the balanced communities case, with both OSC and 
 SSC resulting in no misclustered vertices for a 
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
@@
 
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using OSE (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
@@
 
-In the first real data example, we applied OSE to the Leeds Butterfly 
+In the first real data example, we applied OSC to the Leeds Butterfly 
 dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
@@
 
-Comparing against the ground truth species labels, OSE achieves an 
+Comparing against the ground truth species labels, OSC achieves an 
 accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
@@
 
-```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSE.', cache = FALSE}
+```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.', cache = FALSE}
 plot.A(A, z.hat)
@@
 
-In the second example, we applied OSE to the British MPs Twitter 
+In the second example, we applied OSC to the British MPs Twitter 
 network \cite{greene2013producing}, the Political Blogs network 
@@
 although performance is comparable. In 
-addition, OSE's runtime is much lower than that of modularity 
+addition, OSC's runtime is much lower than that of modularity 
 maximization.
@@
                                           dblp.acc.ssc), 3),
-              `Error (OSE)` = round(1 - c(bmp.acc, 
+              `Error (OSC)` = round(1 - c(bmp.acc, 
                                                   pb.acc,
@@
                                'sparse subspace clustering, ',
-                               'and OSE.'))
-```
-
-```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSE.', fig.dpi = 10}
+                               'and OSC.'))
+```
+
+```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.', fig.dpi = 10}
 gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
@@
 
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSE (left).
+```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
 The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
@@
 knitr::kable(dplyr::tibble(
-  Method = c('OSE', 'SSC', 'Mod. Max.'),
+  Method = c('OSC', 'SSC', 'Mod. Max.'),
   `Error (Village 12)` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
@@
 which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
-that the eigendecomposition of $M^{(kl)}$ is
+that the spectral decomposition of $M^{(kl)}$ is
 

### 2021-02-23 summary.Rmd
minor tweaks

@@
 dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
-              Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
-              `Error (Mod. Max.)` = c(round(1 / 329, 3),
+              # Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
+              `MM` = c(round(1 / 329, 3),
                        round(61 / 1222, 3),
                        round(62 / 2203, 3)),
-              `Error (SSC)` = round(1 - c(bmp.acc.ssc,
+              `SSC-ASE` = round(1 - c(bmp.acc.ssc,
                                       pb.acc.ssc,
                                       dblp.acc.ssc), 3),
-              `Error (OSC)` = round(1 - c(bmp.acc, 
+              `OSC` = round(1 - c(bmp.acc, 
                                   pb.acc,
@@
                            use.all = TRUE, normalize = FALSE, p = 2, q = 0)
-zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .0005)
+zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .001)
 zhat.hh.mm.46 <- mm(A.hh.46)
@@
 
+# knitr::kable(dplyr::tibble(
+#   Method = c('OSC', 'SSC-ASE', 'Mod. Max.'),
+#   `Village 12` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
+#   `Village 31` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
+#   `Village 46` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
+#   caption = paste('Community detection error rates for identifying',
+#                   'household religion.'))
+
 knitr::kable(dplyr::tibble(
-  Method = c('OSC', 'SSC', 'Mod. Max.'),
-  `Error (Village 12)` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
-  `Error (Village 31)` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
-  `Error (Village 46)` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
+  Network = c('Village 12', 'Village 31', 'Village 46'),
+  MM = round(1 - c(hh.acc.mm.12, hh.acc.mm.31, hh.acc.mm.46), 3),
+  `SSC-ASE` = round(1 - c(hh.acc.ssc.12, hh.acc.ssc.31, hh.acc.ssc.46), 3),
+  `OSC` = round(1 - c(hh.acc.12, hh.acc.31, hh.acc.46), 3)
+),
 caption = paste('Community detection error rates for identifying',

### 2021-03-02 summary.Rmd
typo

@@
 $X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
-$\{\lambda^{(kl)}\}_K$ and $U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
+$\{\lambda^{(kl)}\}_K$ and orthonormal matrix 
+$U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
 for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 

### 2021-03-02 summary.Rmd
rewording some things

@@
 ---
-title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation"
+title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation"
 author: John Koo, Minh Tang, Michael Trosset
@@
 the within-community edge probability is $\xi$ and the between-community edge 
-probability is $\eta$, the entries of $P$ consist of $\xi$ and $\eta$.
+probability is $\eta$, the entries of $P$ consist of $P_{ij} = \xi$ if $i$ and 
+$j$ are in the same community and $i \neq j$, $P_{ij} = \eta$ if $i$ and $j$ 
+belong to separate communities, and $P_{ii} = 0$.
 
@@
 random graph models, including the DCBM, can be represented in this way
-\cite{rubindelanchy2017consistency} \cite{lyzinski2014}. An analogous 
+\cite{lyzinski2014} \cite{rubindelanchy2017consistency}. An analogous 
 property exists for the PABM but not for the RDPG model but under the 
@@
 direction, from vertex $j$ to vertex $i$. Furthermore, we will only consider 
-unweighted graphs with binary ($0, 1$) edge weights We will also assume that 
+unweighted graphs with binary ($0$ or $1$) edge weights We will also assume that 
 graphs are hollow, i.e., there are no edges from a vertex to itself. All such 
@@
 
-We previously stated one possible set of latent positions that result in the 
-edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
-(or can estimate) $XU$ directly, then both the community detection and 
-parameter identification problem are trivial since $U$ is orthonormal and 
-fixed for each value of $K$. However, direct identification or estimation of
-$XU$ is not possible \cite{rubindelanchy2017statistical}.
-
-If we decompose $P = Z I_{pq} Z^\top$, then
-$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
-the exact edge probability matrix, we cannot recover the "original" latent 
-positions $XU$. Note that unlike in the case of the RDPG, 
-$Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
-$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
-$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
-us from using the properties of $XU$ directly. In particular, if 
-$Q \in \mathbb{O}(n)$, then we could use the fact that 
-$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
-and $j$ are in different communities.
-
-The explicit form of $XU$ represents points in 
-$\mathbb{R}^{K^2}$ such that points within each community lie on 
-$K$-dimensional orthogonal subspaces. 
-Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
-but retains the property that each community is represented by a 
-$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
-correspond to each community, suggesting the use of SSC. In this paper, we will 
-use a different and leave the properties of SSC on the ASE of $A$ as future
-work.
-
-**Theorem 3.** 
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
-from different communities.
-
-Theorem 3 provides perfect community detection given $P$. Letting $|B|$ be the 
-affinity matrix for graph $G$, $G$ is partitioned into $K$ disjoint subgraphs 
-that correspond to each community. Using $A$ introduces some error in the ASE, 
-resulting in points that lie near subspaces but not exactly on them. Using 
-results from \cite{rubindelanchy2017statistical}, we can show that as $n$ 
-grows, the ASE of $A$ approaches the ASE of $P$ and $B_{ij} \to 0$ for $i$ and 
-$j$ in different clusters.
-
-The ASE of $A$ approaches latent positions that form $P$ as 
-the number of vertices $n$ increases. More precisely, let each 
-$\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and 
-$k, l = 1, ..., K$. Then the corresponding latent positions 
-$XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying
-communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size $n$ from
-$\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability 
-matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the adjacency embedding 
-of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
-\citet{rubindelanchy2017statistical},
-
-\begin{equation} \label{eq:thm3}
-  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
-  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
-\end{equation}
-
-for some $c > 0$ and sequence of $Q_n \in \mathbb{O}(p, q)$.
-
-**Theorem 4**. Let $\hat{V}^{(n)} \in \mathbb{R}^{n \times K^2}$ be the matrix 
-of $K^2$ eigenvectors of $A_n$ corresponding to the $K (K+1) / 2$ most positive 
-eigenvalues and $K (K-1) / 2$ most negative eigenvalues with rows
-$(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
-different communities. Let $n \rho_n = \omega((\log n)^{4c}))$ as described in
-\citet{rubindelanchy2017statistical}. Then for some $c > 0$,
-
-\begin{equation} \label{eq:thm4}
-  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
-    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
-\end{equation}
-
-Thus, as $n \to \infty$, the partitioning of $\hat{B}^{(n)}$ results in 
-perfect community detection. 
-This leads to the Orthogonal Spectral Clustering (OSC) algorithm (Alg. 1).
-
-\begin{algorithm}[H]
+\begin{algorithm}[t]
   \DontPrintSemicolon
@@
 
+We previously stated one possible set of latent positions that result in the 
+edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
+(or can estimate) $XU$ directly, then both the community detection and 
+parameter identification problem are trivial since $U$ is orthonormal and 
+fixed for each value of $K$. However, direct identification or estimation of
+$XU$ is not possible \cite{rubindelanchy2017statistical}.
+
+If we decompose $P = Z I_{pq} Z^\top$, then
+$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
+the exact edge probability matrix, we cannot recover the "original" latent 
+positions $XU$. Note that unlike in the case of the RDPG, 
+$Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
+$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
+$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
+us from using the properties of $XU$ directly. In particular, if 
+$Q \in \mathbb{O}(n)$, then we could use the fact that 
+$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
+and $j$ are in different communities.
+
+The explicit form of $XU$ represents points in 
+$\mathbb{R}^{K^2}$ such that points within each community lie on 
+$K$-dimensional orthogonal subspaces. 
+Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
+but retains the property that each community is represented by a 
+$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
+correspond to each community, suggesting the use of SSC. In this paper, we will 
+use a different and leave the properties of SSC on the ASE of $A$ as future
+work.
+
+**Theorem 3.** 
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability
+matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
+from different communities.
+
+Theorem 3 provides perfect community detection given $P$. Letting $|B|$ be the 
+affinity matrix for graph $G$, $G$ is partitioned into $K$ disjoint subgraphs 
+that correspond to each community. Using $A$ introduces some error in the ASE, 
+resulting in points that lie near subspaces but not exactly on them. Using 
+results from \cite{rubindelanchy2017statistical}, we can show that as $n$ 
+grows, the ASE of $A$ approaches the ASE of $P$ and $B_{ij} \to 0$ for $i$ and 
+$j$ in different clusters.
+
+The ASE of $A$ approaches latent positions that form $P$ as 
+the number of vertices $n$ increases. More precisely, let each 
+$\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and 
+$k, l = 1, ..., K$. Then the corresponding latent positions 
+$XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying
+communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size $n$ from
+$\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability 
+matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the adjacency embedding 
+of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
+\citet{rubindelanchy2017statistical},
+
+\begin{equation} \label{eq:thm3}
+  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
+  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
+\end{equation}
+
+for some $c > 0$ and sequence of $Q_n \in \mathbb{O}(p, q)$.
+
+**Theorem 4**. Let $\hat{V}^{(n)} \in \mathbb{R}^{n \times K^2}$ be the matrix 
+of $K^2$ eigenvectors of $A_n$ corresponding to the $K (K+1) / 2$ most positive 
+eigenvalues and $K (K-1) / 2$ most negative eigenvalues with rows
+$(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
+different communities. Let $n \rho_n = \omega((\log n)^{4c}))$ as described in
+\citet{rubindelanchy2017statistical}. Then for some $c > 0$,
+
+\begin{equation} \label{eq:thm4}
+  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
+    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
+\end{equation}
+
+Thus, as $n \to \infty$, the partitioning of $\hat{B}_n$, the affinity matrix 
+from OSE (Alg.1), results in perfect community detection with probability 1.
+
 Theorems 2, 3, and 4 also provide a very natural path toward using SSC for 
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
   \DontPrintSemicolon
@@
 **Theorem 5**. Let $P_n$ describe the edge probability matrix of the PABM with 
-$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$. Let $Z_n$ be the ASE of 
-$P_n$, and let $Z_n^{(k)}$ correspond to the embedding points that correspond 
-to the $k$^th^ community. Let $\hat{Z}_n$ be the ASE of $A_n$ and $\hat{Z}_n^{(k)}$ 
-be the corresponding $k$^th^ block.
-Let $\hat{B}_n$ be the affinity matrix in the SSC algorithm with input $\hat{Z}_n$ such that $\hat{B}_n^{(ij)}$ is the $ij$^th^ element of $\hat{B}_n$.
-If $r(Z_n^{(k)})$, the inradius of $Z_n^{(k)}$ (as defined by \citet{soltanolkotabi2012}), is positive for each $k$, then $\hat{B}_n^{(ij)} \stackrel{p}{\to} 0$ $\forall i, j$ in different communities.
+$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{Z}_n$ be the ASE 
+of $A_n$. Then $\exists \lambda > 0$ such that $\hat{B}_n$, the affinity matrix
+outputted by SSC using $\hat{Z}_n$ and $\lambda$, has $ij$^th^ entry
+$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$ $\forall i, j$ in different 
+communities.
 
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
   \DontPrintSemicolon
   \SetAlgoLined
-  \caption{PABM parameter estimation using the adjacency matrix.}
+  \caption{PABM parameter estimation.}
   \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}

### 2021-03-02 summary.Rmd
more fixes

@@
 
+**Definition 4** (Inradius) \cite{soltanolkotabi2012} 
+\cite{pmlr-v28-wang13}. 
+The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
+defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
+
 **Lemma 5.1**. 
@@
 matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
-respectively. Let $r_n^{(k)} = r(V_n^{(k)})$ and 
-$\hat{r}_n^{(k)} = r(\hat{V}_n^{(k)})$ be the inradius of points in $V_n$ and
-$\hat{V}_n$ as defined by \citet{soltanolkotabi2012} and 
-\citet{pmlr-v28-wang13}. Then
-
-\begin{equation} \label{eq:r-conv}
-\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
-\end{equation}
-
-*Proof of Lemma 5.1*. 
-This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
-Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is 
-continuous, $r(\hat{V}_n^{(k)}) \stackrel{p}{\to} r(V_n^{(k)})$.
-
-**Lemma 5.2**. 
+respectively. 
 Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
 as defined by \citet{soltanolkotabi2012}. Then
-$\mu_n^{(k)} = 0$ 
+
+\begin{equation} \label{eq:mu-zero}
+\mu_n^{(k)} = 0
+\end{equation}
+
 with probability 1 $\forall n, k$.
 
-*Proof of Lemma 5.2*. 
-Since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
+Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
+extended definition of $\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then
+
+\begin{equation} \label{eq:mu-conv}
+\hat{\mu}_n^{(k)} \stackrel{a.s.}{\to} 0
+\end{equation}
+
+*Proof of Lemma 5.1*. 
+For the first part, 
+since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
 this falls directly out of Theorem 2.8 of \citet{soltanolkotabi2012}.
 
-**Lemma 5.3**. 
-Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the extended definition of
-$\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then
-
-\begin{equation} \label{eq:mu-conv}
-\hat{\mu}_n^{(k)} \stackrel{p}{\to} 0
-\end{equation}
-
-*Proof of Lemma 5.3*. 
-This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
-Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
-continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} \mu(V_n^{(k)}) = 0$.
-
-**Lemma 5.4**. 
+For the second part, 
+this falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
+Since $\hat{V}_n^{(k)} \stackrel{a.s.}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
+continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} \mu(V_n^{(k)}) = 0$.
+
+**Lemma 5.2**. 
 Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
@@
 \begin{equation} \label{eq:delta-conv}
-\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
+\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{a.s.}{\to} 0
 \end{equation}
@@
 
-By (\ref{eq:r-conv}) and (\ref{eq:mu-conv}),
+By (\ref{eq:mu-zero}) and (\ref{eq:mu-conv}),
 
 \begin{equation} \label{eq:thm5-proof}
-r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} 
-r_n^{(k)} - \mu_n^{(k)} = r_n^{(k)} > 0
+r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} 
+r(\hat{V}_n^{(k)}) > 0
 \end{equation}
@@
 
-Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
-(\ref{eq:thm5-proof}) imply that $\forall \epsilon > 0$, $\exists N > 0$ such 
-that $\forall n > N$, 
-$\Pr(\text{SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.
+Then Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
+(\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
+$n \to infty$ with probability 1.
 
@@
 for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
-resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
-a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
-up to sign flip.
+resulting in $W_* \stackrel{a.s.}{\to} I$. Therefore, the adjacency spectral
+embedding of $\hat{M}^{(kl)}$ is a direct estimation of the specific latent 
+positions outlined for $M^{(kl)}$, up to sign flip.
 

### 2021-03-02 summary.Rmd
typo

@@
 (\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
-$n \to infty$ with probability 1.
+$n \to \infty$ with probability 1.
 

### 2021-03-02 summary.Rmd
fixed proof of theorem 5

--- 2021-03-02 summary.Rmd
@@
 
+**Proof of Theorem 4**.
+
 **Definition 4** (Inradius) \cite{soltanolkotabi2012} 
@@
 defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
+In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents 
+the inradius of the symmetric convex hull of $X$.
 
 **Lemma 5.1**. 
-Let $P_n = V_n \Lambda_n V_n^\top$ be defined as in Lemma 5.1. Let 
-$A_n \sim \Bernoulli(P_n)$ be the adjacency matrix drawn from $P_n$, and let 
-$A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be its approximate spectral
-decomposition using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ most 
-negative eigenvalues and their corresponding eigenvectors.
-Without loss of generality, order the rows of $V_n$ and $\hat{V}_n$ by 
-community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the $n_k \times K^2$ 
-matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
-respectively. 
+Let $P_n = V_n \Lambda_n V_n^\top$ be the edge probability matrix of a PABM 
+with $n$ vertices. Let $A_n \sim \Bernoulli(P_n)$ be the adjacency matrix 
+drawn from $P_n$, and let $A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be 
+its approximate spectral decomposition using the $K (K+1) / 2$ most positive 
+and $K (K-1) / 2$ most negative eigenvalues and their corresponding 
+eigenvectors.Without loss of generality, order the rows of $V_n$ and 
+$\hat{V}_n$ by community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the 
+$n_k \times K^2$ matrix of the rows of the $k$^th^ community in $V_n$ and 
+$\hat{V}_n$ respectively. 
 Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
@@
 \begin{equation} \label{eq:mu-conv}
-\hat{\mu}_n^{(k)} \stackrel{a.s.}{\to} 0
+\hat{\mu}_n^{(k)} = 0
 \end{equation}
@@
 For the second part, 
-this falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
-Since $\hat{V}_n^{(k)} \stackrel{a.s.}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
-continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} \mu(V_n^{(k)}) = 0$.
+$\mu(\hat{V}_n^{(k)}) = \max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned by $V_n^{(k)}$.
+Since if $u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ $\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem 2, this quantity must be 0.
 
@@
 \begin{equation} \label{eq:thm5-proof}
-r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} 
-r(\hat{V}_n^{(k)}) > 0
-\end{equation}
-
-for each $k = 1, ..., K$.
-
-Then Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
+r(V_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r(V_n^{(k)})
+\end{equation}
+
+for each $k = 1, ..., K$. If enough points are drawn from each community that 
+the corresponding uncorrupted embedding points span a $K$-dimensional subspace
+(Theorem 2), then $r(V_n^{(k)}) > 0$.
+
+Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
 (\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
-$n \to infty$ with probability 1.
+$n \to \infty$ with probability 1.
 

### 2021-03-09 summary.Rmd
cleaned up proof of theorem 5 a bit more

@@
 
-By (\ref{eq:mu-zero}) and (\ref{eq:mu-conv}),
-
-\begin{equation} \label{eq:thm5-proof}
-r(V_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r(V_n^{(k)})
-\end{equation}
-
-for each $k = 1, ..., K$. If enough points are drawn from each community that 
-the corresponding uncorrupted embedding points span a $K$-dimensional subspace
-(Theorem 2), then $r(V_n^{(k)}) > 0$.
-
-Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
-(\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
-$n \to \infty$ with probability 1.
+The basis of this theorem is Theorem 6 from \citet{pmlr-v28-wang13}, which 
+states that the SSC detection property holds if the noise is small enough and 
+the subspace inradius is greater than the subspace incoherence for each 
+community $k$.
+
+Let $V_{n, -i}^{(k)}$ be $V_n^{(k)}$ with the $i$^th^ entry removed. Suppose 
+that for each community $k$, there are enough vertices such that for each $i$, 
+$V_{n, -i}^{(k)}$ spans its corresponding subspace (Theorem 2). 
+Then $r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0$. 
+Thus by (\ref{eq:mu-conv}), for each $k$, $r_n^{(k)} > \mu_n^{(k)}$ where 
+$\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})$ and $n$ is large enough such that 
+$\min\limits_{k, i} rank(V_{n, -i}^{(k)}) = K$.
+
+Let $r_n = \min_k r_n^{(k)}$. By (\ref{eq:delta-conv}), 
+$\delta_n \stackrel{a.s.}{\to} 0$. Then as $n\to \infty$, 
+$\delta_n < \min_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}$ 
+with probability 1.
+
+Thus the conditions for the SSC detection property from Theorem 6 from
+\citet{pmlr-v28-wang13} are satisfied almost surely as $n \to \infty$.
 

### 2021-03-09 summary.Rmd
more cleanup

@@
 for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
-into $C$ and $B$ as before.
+into $C$ and $B$ as before. The ability for this algorithm to identify the 
+communities exactly is called the SSC detection property.
 
@@
 The natural approach then is to perform SSC on the ASE of $P$ or $A$.
-\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$ themselves.
-
-Much of our work is built on top of work from 
-\citet{rubindelanchy2017statistical}, in particular, the properties of the ASE 
-of $A$ in relation to the ASE of $P$ built on $\{\lambda^{(kl)}\}_K$.
+\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$.
+
+Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
+the SSC detection property holds for $XU$, which is an ASE of $P$. More 
+specifically, if points lie exactly on mutually orthogonal subspaces, then the 
+SSC detection property will hold with probability 1, and this is exactly the 
+case for the PABM (Theorem 2). Much of our work is then built on \citet{ribindelanchy2017statistical}, who describe the convergence behavior of 
+the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13}, who show the 
+necessary conditions for the SSC detection property to hold in noisy cases 
+where the points lie near subspaces.
 

### 2021-03-09 summary.Rmd
more cleanup

@@
 SSC detection property will hold with probability 1, and this is exactly the 
-case for the PABM (Theorem 2). Much of our work is then built on \citet{ribindelanchy2017statistical}, who describe the convergence behavior of 
+case for the PABM (Theorem 2). Much of our work is then built on \citet{rubindelanchy2017statistical}, who describe the convergence behavior of 
 the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13}, who show the 

### 2021-03-16 summary.Rmd
proof for theorem 4

--- 2021-03-09 summary.Rmd
@@
 fontsize: 12pt
-# geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
+geometry: "left=1in,right=1in,top=1in,bottom=1in"
 urlcolor: blue
@@
   are able to establish theoretical asymptotic results of our community 
-  detection and parameter estimation methods for the PABM. 
+  detection and parameter estimation methods for the PABM, including for 
+  Sparse Subspace Clustering (SSC). 
 ---
@@
 
-## The popularity adjusted block model (PABM) and the generalized random dot product graph
+## The popularity adjusted block model and the generalized random dot product graph
 
@@
 
-as in Definition 1. 
+where each $\lambda^{(kl)}$ is a vector as in Definition 1. 
 Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
@@
 $A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
+
+*Proof of Theorem 2*. 
+Define the following matrices from $\{\lambda^{(kl)}\}_K$: 
+$$\Lambda^{(k)} = 
+\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
+\in \mathbb{R}^{n_k \times K}$$
+
+\begin{equation} \label{eq:xy}
+X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
+\end{equation}
+
+$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
+\mathbb{R}^{n \times K}$$
+
+$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
+\mathbb{R}^{n \times K^2}$$
+
+Then $P = X Y^\top$.
+
+Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
+$\Pi$, resulting in $P = X \Pi X^\top$.  
+The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
+$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
+$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
+$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
+a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
+$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
+
+Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
+eigenvalues equal to $-1$. $\Pi$ has the decomposed form
+
+\begin{equation} \label{eq:permutation}
+\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
+\end{equation}
+
+The edge probability matrix then can be written as:
+
+\begin{equation} \label{eq:pabm-grdpg}
+P = X U I_{p, q} (X U)^\top
+\end{equation}
+
+\begin{equation} \label{eq:p}
+p = K (K + 1) / 2
+\end{equation}
+
+\begin{equation} \label{eq:q}
+q = K (K - 1) / 2
+\end{equation}
+
+and we can describe the PABM with $K$ communities as a GRDPG with latent 
+positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
 
@@
 \begin{equation} \label{eq:ssc}
-\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
+c_i = \arg\min_{c} \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
 \end{equation}
@@
 into $C$ and $B$ as before.
+
+**Definition 4** (Subspace Detection Property) \cite{pmlr-v28-wang13}. 
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
+points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the 
+solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of 
+$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from 
+different subspaces, then $X$ obeys the subspace detection property.
+
+**Remark**. In practice, sample $X$ often does not obey the subspace detection
+property. In such cases, $B$ is treated as an affinity matrix for a graph which 
+is then partitioned into $K$ subgraphs to obtain the clustering. On the other 
+hand, if $X$ does obey the subspace detection property, $B$ describes a graph 
+with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
+property holds, there are exactly $K$ subgraphs which map to each subspace, 
+but it could be the case that some of the subspaces are represented by 
+multiple disconnected subgraphs.
 
@@
 The natural approach then is to perform SSC on the ASE of $P$ or $A$.
-\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$ themselves.
-
-Much of our work is built on top of work from 
-\citet{rubindelanchy2017statistical}, in particular, the properties of the ASE 
-of $A$ in relation to the ASE of $P$ built on $\{\lambda^{(kl)}\}_K$.
+\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$, foregoing 
+embedding altogether.
+
+Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
+the subspace detection property holds for $XU$, which is an ASE of $P$. More 
+specifically, if points lie exactly on mutually orthogonal subspaces, then the 
+subspace detection property will hold with probability 1, and this is exactly 
+the case for the PABM (Theorem 2). Much of our work is then built on \citet{rubindelanchy2017statistical}, who describe the convergence behavior of 
+the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13}, who show the 
+necessary conditions for the subspace detection property to hold in noisy cases 
+where the points lie near subspaces.
 
@@
 the exact edge probability matrix, we cannot recover the "original" latent 
-positions $XU$. Note that unlike in the case of the RDPG, 
-$Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
+positions $XU$. Note that unlike in the case of the RDPG, $Q$ is not 
+necessarily an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
 $||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
@@
 
-**Theorem 4**. Let $\hat{V}^{(n)} \in \mathbb{R}^{n \times K^2}$ be the matrix 
-of $K^2$ eigenvectors of $A_n$ corresponding to the $K (K+1) / 2$ most positive 
-eigenvalues and $K (K-1) / 2$ most negative eigenvalues with rows
-$(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
-different communities. Let $n \rho_n = \omega((\log n)^{4c}))$ as described in
-\citet{rubindelanchy2017statistical}. Then for some $c > 0$,
+**Theorem 4**. 
+Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij})$ be the affinity matrix from OSC (Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities,
 
 \begin{equation} \label{eq:thm4}
-  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
-    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
+\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0
 \end{equation}
-
-Thus, as $n \to \infty$, the partitioning of $\hat{B}_n$, the affinity matrix 
-from OSE (Alg.1), results in perfect community detection with probability 1.
 
@@
 
-**Proof of Theorem 2**. 
-Let 
-$$\Lambda^{(k)} = 
-\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
-\in \mathbb{R}^{n_k \times K}$$
-
-\begin{equation} \label{eq:xy}
-X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
-\end{equation}
-
-$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
-\mathbb{R}^{n \times K}$$
-
-$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
-\mathbb{R}^{n \times K^2}$$
-
-Then $P = X Y^\top$.
-
-Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
-$\Pi$, resulting in $P = X \Pi X^\top$.  
-The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
-$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
-$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
-$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
-a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
-$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
-
-Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
-eigenvalues equal to $-1$. $\Pi$ has the decomposed form
-
-\begin{equation} \label{eq:permutation}
-\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
-\end{equation}
-
-The edge probability matrix then can be written as:
-
-\begin{equation} \label{eq:pabm-grdpg}
-P = X U I_{p, q} (X U)^\top
-\end{equation}
-
-\begin{equation} \label{eq:p}
-p = K (K + 1) / 2
-\end{equation}
-
-\begin{equation} \label{eq:q}
-q = K (K - 1) / 2
-\end{equation}
-
-and we can describe the PABM with $K$ communities as a GRDPG with latent 
-positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
-
 **Lemma 3.1.** 
@@
 
-The righthand side becomes
+The right-hand side becomes
 
@@
 
-The lefthand side becomes:
+The left-hand side becomes:
 
@@
 **Proof of Theorem 4**. 
-
-**Definition 4** (Inradius) \cite{soltanolkotabi2012} 
+By Theorem 3, if $i$ and $j$ belong to different communities and $v_n^{(i)}$ 
+and $v_n^{(j)}$ are the corresponding eigenvectors of $P_n$, then 
+$(v_n^{(i)})^\top v_n^{(j)} = 0$. Then by \citet{rubindelanchy2017statistical}, 
+for eigenvectors of $A_n$, $\hat{v}_n^{(i)}$ and $\hat{v}_n^{(j)}$,
+
+$$(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} \stackrel{a.s.}{\to} 
+(v_n^{(i)})^\top v_n^{(j)} = 0$$
+
+since $W_n \hat{v}_n^{(i)} \stackrel{a.s.}{\to} v_n^{(i)}$ for each 
+$i = 1, ..., n$ for some orthonormal $W_n$.
+
+**Definition 5** (Inradius) \cite{soltanolkotabi2012} 
 \cite{pmlr-v28-wang13}. 
@@
 Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
-as defined by \citet{soltanolkotabi2012}. Then
-
-\begin{equation} \label{eq:mu-zero}
-\mu_n^{(k)} = 0
-\end{equation}
-
+as defined by \citet{soltanolkotabi2012}. Then $\mu_n^{(k)} = 0$ 
 with probability 1 $\forall n, k$.
@@
 For the second part, 
-$\mu(\hat{V}_n^{(k)}) = \max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned by $V_n^{(k)}$.
-Since if $u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ $\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem 2, this quantity must be 0.
+$\mu(\hat{V}_n^{(k)}) = 
+\max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ 
+where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned 
+by $V_n^{(k)}$. Since if 
+$u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ 
+$\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem 2, this quantity must be 0.
 
@@
 The basis of this theorem is Theorem 6 from \citet{pmlr-v28-wang13}, which 
-states that the SSC detection property holds if the noise is small enough and 
-the subspace inradius is greater than the subspace incoherence for each 
+states that the subspace detection property holds if the noise is small enough 
+and the subspace inradius is greater than the subspace incoherence for each 
 community $k$.
@@
 
-Thus the conditions for the SSC detection property from Theorem 6 from
+Thus the conditions for the subspace detection property from Theorem 6 from
 \citet{pmlr-v28-wang13} are satisfied almost surely as $n \to \infty$.

### 2021-03-22 summary.Rmd
fixed theorem 4

@@
 title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation"
-author: John Koo, Minh Tang, Michael Trosset
+authors: John Koo, Minh Tang, Michael Trosset
+# author: John Koo
+# author2: Minh Tang
+# author3: Michael Trosset
 output:
@@
     number_sections: yes
+# output:
+#   bookdown::pdf_document2:
+#     citation_package: natbib
+#     number_citations: yes
 # output: html_document
+# output: rticles::rss_article
 fontsize: 12pt
-geometry: "left=1in,right=1in,top=1in,bottom=1in"
+# geometry: "left=1in,right=1in,top=1in,bottom=1in"
 urlcolor: blue
@@
 - \usepackage{verbatim}
+- \usepackage{amsthm}
 bibliography: misc.bib
 abstract: |
-  In this paper, we connect two random graph models, the Popularity 
-  Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
-  (GRDPG) and use properties established in this connection to aid in 
-  community detection and parameter estimation. In particular, we note that the
-  PABM can be represented as latent positions such that points within the same 
-  community lie on a subspace, and the subspaces that represent each 
-  community are orthogonal to one another. Using this property as well as the 
-  asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
-  are able to establish theoretical asymptotic results of our community 
-  detection and parameter estimation methods for the PABM, including for 
-  Sparse Subspace Clustering (SSC). 
+  We connect two random graph models, the Popularity Adjusted Block Model 
+  (PABM) and the Generalized Random Dot Product Graph (GRDPG), demonstrating
+  that a PABM is a GRDPG in which communities correspond to certain mutually
+  orthogonal subspaces of latent positions. This insight leads to the
+  construction of improved algorithms for community detection and parameter
+  estimation with PABM. Using established asymptotic properties of Adjacency
+  Spectral Embedding (ASE) for GRDPG, we derive asymptotic properties of these
+  algorithms, including algorithms that rely on Sparse Subspace Clustering
+  (SSC). We illustrate these properties via simulation.
 ---
@@
 
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \newcommand{\diag}{\text{diag}}
@@
 \newcommand{\Betadist}{\text{Beta}}
+\newtheorem{definition}{Definition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\theoremstyle{example}
+\newtheorem*{example}{Example}
 
@@
 
-**Definition 1** (Popularity Adjusted Block Model) 
+\begin{definition}[Popularity Adjusted Block Model]
+\label{pabm}
 \cite{307cbeb9b1be48299388437423d94bf1}. 
@@
 $\{\lambda^{(st)}\}_{s, t = 1}^K$. 
+\end{definition}
 
@@
 
-**Definition 2** (Generalized Random Dot Product Graph) 
+\begin{definition}[Generalized Random Dot Product Graph] 
+\label{grdpg} 
 \cite{rubindelanchy2017statistical}.
@@
 the latent positions as rows $x_i^\top$. 
+\end{definition}
 
@@
 
-**Definition 3**. The indefinite orthogonal group with signature $(p, q)$ is 
+\begin{definition}[Indefinite Orthogonal Group] 
+The indefinite orthogonal group with signature $(p, q)$ is 
 the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
 denoted as $\mathbb{O}(p, q)$ \cite{rubindelanchy2017statistical}.
-
-**Remark**. 
+\end{definition}
+
+\begin{remark}
 Like the RDPG, the latent positions of a GRDPG are not unique 
@@
 distances. 
+\end{remark}
 
@@
 
-**Theorem 1** (Connecting the PABM to the GRDPG for $K = 2$). 
+\begin{theorem}[Connecting the PABM to the GRDPG for $K = 2$]
+\label{theorem1}  
 Let
@@
 equivalent.
-
-**Theorem 2** (Generalization to $K > 2$). There exists a block diagonal matrix 
+\end{theorem}
+
+\begin{theorem}[Generalization to $K > 2$] 
+\label{theorem2}
+There exists a block diagonal matrix 
 $X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
@@
 $A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
-
-*Proof of Theorem 2*. 
+\end{theorem}
+
+\begin{proof}
 Define the following matrices from $\{\lambda^{(kl)}\}_K$: 
@@
 positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
-
-**Example** ($K = 3$). Using the same notation as before:
+\end{proof}
+
+\begin{example}[$K = 3$] Using the same notation as in Theorem \ref{theorem2}:
 
@@
 * Positions 1, 5, 9 are fixed.
+
 * The cycles of order 2 are $(2, 4)$, $(3, 7)$, and $(6, 8)$.
@@
 \end{bmatrix}$$
+\end{example}
 
@@
 
-**Definition 4** (Subspace Detection Property) \cite{pmlr-v28-wang13}. 
+\begin{definition}[Subspace Detection Property \cite{pmlr-v28-wang13}]. 
 Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
@@
 different subspaces, then $X$ obeys the subspace detection property. 
-
-**Remark**. In practice, sample $X$ often does not obey the subspace detection
+\end{definition}
+
+\begin{remark} 
+In practice, sample $X$ often does not obey the subspace detection
 property. In such cases, $B$ is treated as an affinity matrix for a graph which 
@@
 multiple disconnected subgraphs. 
-
-Theorem 2 suggests that SSC is appropriate for community detection for the 
-PABM. More precisely, Theorem 2 says that each community consists of a
+\end{remark}
+
+Theorem \ref{theorem2} suggests that SSC is appropriate for community 
+detection for the PABM. 
+More precisely, Theorem \ref{theorem2} says that each community consists of a
 $K$-dimensional subspace, and together the subspaces lie in $\mathbb{R}^{K^2}$. 
 The natural approach then is to perform SSC on the ASE of $P$ or $A$.
-\citet{noroozi2019estimation} instead applied SSC to $P$ and $A$, foregoing 
-embedding altogether.
-
-Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
+\citeauthor{noroozi2019estimation} instead applied SSC to $P$ and $A$, 
+foregoing embedding altogether.
+
+Using results from \citeauthor{soltanolkotabi2012}, it can be easily shown that 
 the subspace detection property holds for $XU$, which is an ASE of $P$. More 
@@
 subspace detection property will hold with probability 1, and this is exactly 
-the case for the PABM (Theorem 2). Much of our work is then built on \citet{rubindelanchy2017statistical}, who describe the convergence behavior of 
-the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13}, who show the 
-necessary conditions for the subspace detection property to hold in noisy cases 
-where the points lie near subspaces.
+the case for the PABM (Theorem \ref{theorem2}). 
+Much of our work is then built on
+\citeauthor{rubindelanchy2017statistical}, who describe the convergence 
+behavior of the ASE of $A$ to the ASE of $P$, and \citeauthor{pmlr-v28-wang13},
+who show the necessary conditions for the subspace detection property to hold 
+in noisy cases where the points lie near subspaces.
 
@@
     $V$ using these eigenvectors as its columns.\;
-    Compute $B = |V V^\top|$, applying $|\cdot|$ entry-wise.\;
+    Compute $B = |n V V^\top|$, applying $|\cdot|$ entry-wise.\;
     Construct graph $G$ using $B$ as its similarity matrix.\;
@@
 
-**Theorem 3.** 
+\begin{theorem}
+\label{theorem3}
 Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
+matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
 from different communities.
-
-Theorem 3 provides perfect community detection given $P$. Letting $|B|$ be the 
-affinity matrix for graph $G$, $G$ is partitioned into $K$ disjoint subgraphs 
-that correspond to each community. Using $A$ introduces some error in the ASE, 
-resulting in points that lie near subspaces but not exactly on them. Using 
-results from \cite{rubindelanchy2017statistical}, we can show that as $n$ 
-grows, the ASE of $A$ approaches the ASE of $P$ and $B_{ij} \to 0$ for $i$ and 
-$j$ in different clusters.
-
-The ASE of $A$ approaches latent positions that form $P$ as 
-the number of vertices $n$ increases. More precisely, let each 
-$\lambda^{(kl)}_i \sim \mathcal{F}^{(kl)}$ for $i = 1, .., n$ and 
-$k, l = 1, ..., K$. Then the corresponding latent positions 
-$XU \sim \mathcal{G}_K$ for some related joint distribution with $K$ underlying
-communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size $n$ from
-$\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge probability 
-matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the adjacency embedding 
-of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
-\citet{rubindelanchy2017statistical},
-
-\begin{equation} \label{eq:thm3}
-  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
-  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
+\end{theorem}
+
+Theorem \ref{theorem3} provides perfect community detection given $P$. 
+Letting $|B|$ be the affinity matrix for graph $G$, $G$ is partitioned into 
+at least $K$ disjoint subgraphs since each of the $K$ communities have no 
+edges between them. Similar to the subspace detection property, it could be 
+the case that some of the communities are represented by multiple disjoint
+subgraphs in $G$, in which case additional reconstruction is required to
+identify the communities exactly.
+
+Using $A$ instead of $P$ introduces error, which converges to $0$ 
+almost surely:
+
+\begin{theorem}
+\label{theorem4} 
+Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC 
+(Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities 
+and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,
+
+\begin{equation} \label{eq:thm4}
+\max_{i, j} |n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = 
+O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
 \end{equation}
 
-for some $c > 0$ and sequence of $Q_n \in \mathbb{O}(p, q)$.
-
-**Theorem 4**. 
-Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij})$ be the affinity matrix from OSC (Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities,
-
-\begin{equation} \label{eq:thm4}
-\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0
-\end{equation}
-
-Theorems 2, 3, and 4 also provide a very natural path toward using SSC for 
-community detection for the PABM. We established in Theorem 2 that an ASE of 
-the edge probability matrix $P$ can be constructed such that the communities 
-lie on mutually orthogonal subspaces, and this property can be recovered from
-the eigenvectors of $P$. Then Theorems 3 and 4 show that this property holds 
-for the adjacency matrix $A$ drawn from $P$ as $n \to \infty$.
+This provides the result that for $i, j$ in different communities, 
+$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$.
+\end{theorem}
+
+Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} 
+also provide a very natural path toward using SSC for 
+community detection for the PABM. We established in Theorem \ref{theorem2}
+that an ASE of the edge probability matrix $P$ can be constructed 
+such that the communities lie on mutually orthogonal subspaces,
+and this property can be recovered from the eigenvectors of $P$. 
+Then Theorems \ref{theorem3} and \ref{theorem4} show that this property holds 
+for the unscaled ASE of $A$ drawn from $P$ as $n \to \infty$.
 
@@
   \KwResult{Community assignments $1, ..., K$}
-    Find $Z$, the ASE of $A$ using the $K (K + 1) / 2$ most positive 
-    and the $K (K - 1) / 2$ most negative eigenvalues and eigenvectors of 
-    $A$.\;
+    Find $V$, the matrix of eigenvectors of $A$ 
+    corresponding to the $K (K + 1) / 2$ most positive 
+    and the $K (K - 1) / 2$ most negative eigenvalues.\;
     \For {$i = 1, ..., n$} {
-      Assign $z_i^\top$ as the $i^{th}$ row of $Z$. 
-      Assign $Z_{-i} = \begin{bmatrix} 
-      z_1 & \cdots & z_{i-1} & z_{i+1} & \cdots & z_n \end{bmatrix}^\top$.\;
+      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
+      Assign $V_{-i} = \begin{bmatrix} 
+      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
       Solve the LASSO problem 
-      $c_i = \arg\min_{\beta_i} 
-      \frac{1}{2} ||z_i - Z_{-i} \beta_i||_2^2 + \lambda ||\beta_i||_1$.\;
+      $c_i = \arg\min_{\beta} 
+      \frac{1}{2} ||v_i - V_{-i} \beta||_2^2 + \lambda ||\beta||_1$.\;
       Assign $\tilde{c}_i = \begin{bmatrix} 
@@
 
-**Theorem 5**. Let $P_n$ describe the edge probability matrix of the PABM with 
-$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{Z}_n$ be the ASE 
-of $A_n$. Then $\exists \lambda > 0$ such that $\hat{B}_n$, the affinity matrix
-outputted by SSC using $\hat{Z}_n$ and $\lambda$, has $ij$^th^ entry
+\begin{theorem}
+\label{theorem5}
+Let $P_n$ describe the edge probability matrix of the PABM with 
+$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{V}_n$ be the 
+matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
+positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
+$\exists \lambda > 0$ such that $\hat{B}_n$, the affinity matrix
+outputted by SSC using $\hat{V}_n$ and $\lambda$, has $ij^{th}$ entry
 $\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$ $\forall i, j$ in different 
 communities. 
+\end{theorem}
 
@@
 simply use plug-in estimators (algorithm 3), which converge to the true 
-parameters (Theorem 6).
+parameters.
 
@@
 
-**Theorem 6**. Under regularity and sparsity assumptions, and under the further 
+\begin{theorem}
+\label{theorem6}
+Under regularity and sparsity assumptions, and under the further 
 assumption that $K$ is fixed and community labels are known,
@@
 \end{equation}
+\end{theorem}
 
@@
 vectors $\{\lambda^{(kl)}\}_K$, and this method is then compared against an
-MLE-based estimator described in \citet{noroozi2019estimation} and
-\citet{307cbeb9b1be48299388437423d94bf1}.
+MLE-based estimator described in \citeauthor{noroozi2019estimation} and
+\citeauthor{307cbeb9b1be48299388437423d94bf1}.
 
@@
 SSC-A, treats the columns of the adjacency matrix $A$ as points in
-$\mathbb{R}^n$, as described in \citet{noroozi2019estimation}. The second 
+$\mathbb{R}^n$, as described in \citeauthor{noroozi2019estimation}. The second 
 method, denoted as SSC-ASE, first embeds $A$ and then performs SSC on the 
@@
 
-Theorem 4 implies that OSC will result in not just in the error rate 
-converging to 0 but the error *count* as well.
+Theorem \ref{theorem4} implies that OSC will result in not just in the error
+rate converging to 0 but the error *count* as well.
 We explore this in Fig. \ref{fig:clust_err_ct_sim}.
@@
 butterflies across 10 species. The graph was modified to match the example 
-from \citet{noroozi2019estimation}: Only the 4 most frequent species were 
+from \citeauthor{noroozi2019estimation}: Only the 4 most frequent species were 
 considered, and the similarities were discretized to $\{0, 1\}$ via 
@@
 accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
-\citet{noroozi2019estimation} achieved an adjusted Rand index of 73\% using 
-sparse subspace clustering on the same dataset.
+\citeauthor{noroozi2019estimation} achieved an adjusted Rand index of 73\% 
+using sparse subspace clustering on the same dataset.
 
@@
                                'sparse subspace clustering, ',
-                               'and OSC.'))
+                               'and OSC.'),
+               format = 'latex')
 ```
@@
 caption = paste('Community detection error rates for identifying',
-                  'household religion.'))
+                  'household religion.'),
+format = 'latex')
 ```
@@
 
-**Proof of Theorem 1**. 
+**Proof of Theorem \ref{theorem1}**. 
 This is given by straightforward matrix multiplication. It suffices to show 
@@
 
-*Remark*. While we can just perform the matrix multiplication to show the 
+\begin{remark} 
+While we can just perform the matrix multiplication to show the 
 equivalence, it is more illustrative to look at a few intermediate steps. Note 
 that the product of the three inner matrices results in a permutation matrix 
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swappoing 
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping 
 positions $2$ and $3$:
@@
 swapped.
-
-**Lemma 3.1.** 
+\end{remark}
+
+\begin{lemma}
+\label{lemma1}
 Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
@@
 defined as in (\ref{eq:xy}).
-
-*Proof of Lemma 3.1*. 
+\end{lemma}
+
+*Proof*. 
 By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
@@
   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-    = X U Q Q^{-1} U^\top (X^\top X)^{-1} U (Q^\top)^{-1} Q^\top U^\top X^\top \\
+    = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
+    U (Q^\top)^{-1} Q^\top U^\top X^\top \\
   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
@@
 
-**Proof of Theorem 3**.
+**Proof of Theorem \ref{theorem3}**.
 By Lemma 3.1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
@@
 
-**Proof of Theorem 4**. 
-By Theorem 3, if $i$ and $j$ belong to different communities and $v_n^{(i)}$ 
-and $v_n^{(j)}$ are the corresponding eigenvectors of $P_n$, then 
-$(v_n^{(i)})^\top v_n^{(j)} = 0$. Then by \citet{rubindelanchy2017statistical}, 
-for eigenvectors of $A_n$, $\hat{v}_n^{(i)}$ and $\hat{v}_n^{(j)}$,
-
-$$(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} \stackrel{a.s.}{\to} 
-(v_n^{(i)})^\top v_n^{(j)} = 0$$
-
-since $W_n \hat{v}_n^{(i)} \stackrel{a.s.}{\to} v_n^{(i)}$ for each 
-$i = 1, ..., n$ for some orthonormal $W_n$.
-
-**Definition 5** (Inradius) \cite{soltanolkotabi2012} 
-\cite{pmlr-v28-wang13}. 
+**Proof of Theorem \ref{theorem4}**. 
+Let $V_n$ and $\hat{V}_n$ be the eigenvectors of $P$ and $A$ corresponding 
+to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative
+eigenvalues. By \citeauthor{rubindelanchy2017statistical}, for some 
+$W \in \mathbb{O}(K^2)$, and $c > 0$,  
+$||\hat{V} W - V||_{2 \to \infty} = 
+O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big)$. 
+We furthermore have $||V||_{2 \to \infty} = O_P(n^{-1/2})$. Then if
+$(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ correspond to the rows of 
+$V_n$ and $\hat{V}_n$, for $i$ and $j$ in different communities, using the 
+fact that $(v_n^{(i)})^\top v_n^{(j)} = 0$,
+
+$$\begin{split}
+\max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
+= \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} - 
+(v_n^{(i)})^\top v_n^{(j)}| \\
+& = \max_{i, j} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
+(v_n^{(i)})^\top v_n^{(j)}| \\
+& = ||\hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
+& = ||2 \hat{V}_n W V^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
+2 \hat{V}_n W V^\top + V V^\top||_{2 \to \infty} \\
+& = ||2 (\hat{V}_n W - V_n) V_n^\top + 
+(\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
+& \leq 2 ||\hat{V}_n W - V_n||_{2 \to \infty} ||V_n||_{2 \to \infty} +
+||\hat{V}_n W - V_n||_{2 \to \infty}^2 \\
+& = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
+\end{split}$$
+
+Then scaling by $n$, we get 
+$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = 
+O_P \big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \big)$.
+
+\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{pmlr-v28-wang13}]
 The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
@@
 the inradius of the symmetric convex hull of $X$. 
-
-**Lemma 5.1**. 
+\end{definition}
+
+\begin{lemma}
+\label{lemma2}
 Let $P_n = V_n \Lambda_n V_n^\top$ be the edge probability matrix of a PABM 
@@
 $\hat{V}_n$ by community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the 
-$n_k \times K^2$ matrix of the rows of the $k$^th^ community in $V_n$ and 
+$n_k \times K^2$ matrix of the rows of the $k^{th}$ community in $V_n$ and 
 $\hat{V}_n$ respectively. 
 Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
-as defined by \citet{soltanolkotabi2012}. Then $\mu_n^{(k)} = 0$ 
+as defined by \citeauthor{soltanolkotabi2012}. Then $\mu_n^{(k)} = 0$ 
 with probability 1 $\forall n, k$.
@@
 Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
-extended definition of $\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then
+extended definition of $\mu(\cdot)$ from \citeauthor{pmlr-v28-wang13}. Then
 
@@
 \end{equation}
-
-*Proof of Lemma 5.1*. 
+\end{lemma}
+
+*Proof*. 
 For the first part, 
-since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
-this falls directly out of Theorem 2.8 of \citet{soltanolkotabi2012}.
+since $v_n^{(k)}$ is orthogonal to $v_n^{(l)}$ $\forall k \neq l$ 
+(Lemma \ref{lemma1}), 
+this falls directly out of Theorem 2.8 of \citeauthor{soltanolkotabi2012}.
 
@@
 where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned 
-by $V_n^{(k)}$. Since if 
+by $V_n^{(k)}$. Since 
 $u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ 
-$\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem 2, this quantity must be 0.
-
-**Lemma 5.2**. 
+$\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem \ref{theorem2}, this quantity 
+must be 0.
+
+\begin{lemma}
+\label{lemma3}
 Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-$\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},
+$\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
 
@@
 \end{equation}
-
-**Proof of Theorem 5**. 
-For ease of proving, we will consider $V_n$ and $\hat{V}_n$ instead of $Z_n$ 
-and $\hat{Z}_n$.
-
-The basis of this theorem is Theorem 6 from \citet{pmlr-v28-wang13}, which 
+\end{lemma}
+
+**Proof of Theorem \ref{theorem5}**. 
+The basis of this proof is Theorem 6 from \citeauthor{pmlr-v28-wang13}, which 
 states that the subspace detection property holds if the noise is small enough 
@@
 Thus the conditions for the subspace detection property from Theorem 6 from
-\citet{pmlr-v28-wang13} are satisfied almost surely as $n \to \infty$.
-
-**Proof of Theorem 6**. 
+\citeauthor{pmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
+
+**Proof of Theorem \ref{theorem6}**. 
 Let $P$ and $A$ be organized by community such that the elements of blocks 
@@
 $\lambda^{(kk)}$.  
-Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
+Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the adjacency 
 spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
@@
 the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
-by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
+by the rate given in Theorem 5 of \citeauthor{rubindelanchy2017statistical}.  
 In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 

### 2021-03-22 functions.R

@@ -1,4 +1,5 @@
 embedding <- function(A, p = NULL, q = NULL,
+                      scale = TRUE,
                       eps = 1e-6) {
   n <- nrow(A)
   eigen.A <- eigen(A, symmetric = TRUE)
@@ -9,8 +10,12 @@ embedding <- function(A, p = NULL, q = NULL,
   }
   
   U <- eigen.A$vectors[, keep]
-  S <- diag(sqrt(abs(eigen.A$values[keep])))
-  return(U %*% S)
+  if (scale) {
+    S <- diag(sqrt(abs(eigen.A$values[keep])))
+    return(U %*% S)
+  } else {
+    return(U)
+  }
 }
 
 draw.graph <- function(P) {
@@ -244,11 +249,12 @@ ssc <- function(A,
                 K = 2,
                 lambda = .01,
                 parallel = FALSE,
-                normalize = TRUE) {
+                normalize = TRUE,
+                scale = TRUE) {
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   
-  Y <- t(embedding(A, p, q))
+  Y <- t(embedding(A, p, q, scale))
   
   N <- ncol(Y)
   B <- plyr::aaply(seq(N), 1, function(i) {

### 2021-03-23 summary.Rmd
started writing the discussion section

@@
 
+This paper shows the connection between the PABM and the GRDPG, namely that 
+a PABM graph can be represented as a union of orthogonal subspaces in an 
+embedding under the GRDPG framework. We then exploited this relationship to 
+develop community detection and parameter estimation methods. In fact, we can 
+represent any graph with Bernoulli edges as a GRDPG, and in the PABM case, 
+it turns out that this relationship leads to a very straightforward 
+applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
+\citeauthor{soltanolkotabi2012}, and \citeauthor{pmlr-v28-wang13}, which lead 
+to asymptotically correct solutions with high probability. Similar methods can 
+be applied for other models, such as the Nested Block Model
+\cite{noroozi2021hierarchy}.
+
 # Proofs

### 2021-03-23 summary.Rmd
cleaned up a few typos

@@
 \label{pabm}
-\cite{307cbeb9b1be48299388437423d94bf1}. 
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
@@
 \label{grdpg} 
-\cite{rubindelanchy2017statistical}.
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
@@
 the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
-denoted as $\mathbb{O}(p, q)$ \cite{rubindelanchy2017statistical}.
+denoted as $\mathbb{O}(p, q)$.
 \end{definition}
@@
 
-\citet*{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, used 
-a relaxed version of Modularity Maximization (MM) based on the Extreme Points
+\citeauthor{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, 
+used Modularity Maximization (MM) and the Extreme Points
 (EP) algorithm \cite{le2016} for community detection and parameter estimation. 
@@
 \begin{equation} \label{eq:ssc}
-c_i = \arg\min_{c} \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
+c_i = \arg\min_c \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
 \end{equation}
@@
 
-\begin{definition}[Subspace Detection Property \cite{pmlr-v28-wang13}]. 
+\begin{definition}[Subspace Detection Property]. 
 Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
@@
 
-Using results from \citeauthor{soltanolkotabi2012}, it can be easily shown that 
+Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
 the subspace detection property holds for $XU$, which is an ASE of $P$. More 
@@
 \citeauthor{rubindelanchy2017statistical}, who describe the convergence 
-behavior of the ASE of $A$ to the ASE of $P$, and \citeauthor{pmlr-v28-wang13},
+behavior of the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13},
 who show the necessary conditions for the subspace detection property to hold 
@@
 data as described by 
-\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
+\citeauthor{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
 networks. Our methods underperformed compared to modularity maximization, 
@@
 & = ||\hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 \hat{V}_n W V^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
-2 \hat{V}_n W V^\top + V V^\top||_{2 \to \infty} \\
+& = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
+2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
 & = ||2 (\hat{V}_n W - V_n) V_n^\top + 

### 2021-03-23 summary.Rmd
fixed ssc scaling

@@
     $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
-    Compute the affinity matrix $B = |C| + |C^\top|$.\;
+    Compute the affinity matrix $B = \sqrt{n} (|C| + |C^\top|)$.\;
     Construct graph $G$ using $B$ as its similarity matrix.\;
@@
 positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
-$\exists \lambda > 0$ such that $\hat{B}_n$, the affinity matrix
-outputted by SSC using $\hat{V}_n$ and $\lambda$, has $ij^{th}$ entry
-$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$ $\forall i, j$ in different 
-communities. 
+$\exists \lambda > 0$ such that as $n \to \infty$, $\hat{V}_n$ obeys the 
+subspace detection property with probability 1.
 \end{theorem}

### 2021-03-23 functions.R

@@ -95,7 +95,7 @@ cluster.pabm <- function(A, K,
     v.norm <- sqrt(rowSums(V ** 2))
     V <- sweep(V, 1, v.norm, `/`)
   }
-  B <- abs(V %*% t(V))
+  B <- n * abs(V %*% t(V))
   if (laplacian == 'graph') {
     L <- graph.laplacian(B)
   } else if (laplacian == 'normalized') {
@@ -277,6 +277,7 @@ ssc <- function(A,
     B[is.nan(B)] <- 0
   }
   W <- B + t(B)
+  W <- sqrt(N) * W
   L <- normalized.laplacian(W)
   L.eigen <- eigen(L, symmetric = TRUE)
   X <- L.eigen$vectors[, seq(N, N - K)]

### 2021-03-23 summary.Rmd
typos

@@
 Define the following matrices from $\{\lambda^{(kl)}\}_K$:
+
 $$\Lambda^{(k)} = 
@@
 **Proof of Theorem \ref{theorem3}**.
-By Lemma 3.1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
-in Theorem 2. Since $X$ is block diagonal with each block corresponding 
-to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
-with each block corresponding to a community and zeros elsewhere. Therefore, 
-if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
-element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
+By Lemma \ref{lemma1}, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
+defined as in (\ref{eq:xy}). Since $X$ is block diagonal with each 
+block corresponding to one community, $X (X^\top X)^{-1} X^\top$ is also a 
+block diagonal matrix with each block corresponding to a community and zeros
+elsewhere. Therefore, if vertices $i$ and $j$ belong to different communities,
+then the $ij$^th^ element of $n X (X^\top X)^{-1} X^\top = n V V^\top = B$ 
+is 0.
 

### 2021-03-24 clustering-sim-k.R
updated ssc sim

@@ -81,7 +81,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       A <- draw.graph(P)
       clustering <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error <- 1 - cluster.acc(clustering, z)
-      clustering.ssc <- ssc(A, K, sparsity, normalize = TRUE)
+      clustering.ssc <- ssc(A, K, n ** -1 * K, normalize = TRUE, scale = FALSE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
       if (K == 2) {
         clustering.mm <- cluster.sg(A)
@@ -89,10 +89,11 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
         clustering.mm <- mod.max(A)
       }
       error.mm <- 1 - cluster.acc(clustering.mm, z)
-      dplyr::tibble(K = K, n = n, 
-                    error = error, 
+      dplyr::tibble(K = K, n = n,
+                    error = error,
                     error.ssc = error.ssc,
-                    error.mm = error.mm) %>% 
+                    error.mm = error.mm) %>%
+        return()
         return()
     } %>% 
       return()
@@ -105,15 +106,17 @@ gc()
 
 clustering.df %>%
   dplyr::group_by(K, n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
+  dplyr::summarise(
+    med.err = median(error),
+    first.q = quantile(error, .25),
+    third.q = quantile(error, .75),
+    med.err.ssc = median(error.ssc),
+    first.q.ssc = quantile(error.ssc, .25),
+    third.q.ssc = quantile(error.ssc, .75),
+    med.err.mm = median(error.mm),
+    first.q.mm = quantile(error.mm, .25),
+    third.q.mm = quantile(error.mm, .75)
+  ) %>%
   ggplot() +
   scale_y_log10() +
   scale_x_log10() +

### 2021-03-24 clustering-sim-k-imbalanaced.R
updated ssc sim

@@ -81,7 +81,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       A <- draw.graph(P)
       clustering <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error <- 1 - cluster.acc(clustering, z)
-      clustering.ssc <- ssc(A, K, sparsity, normalize = TRUE)
+      clustering.ssc <- ssc(A, K, K / n / 4, normalize = TRUE, scale = FALSE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
       if (K == 2) {
         clustering.mm <- cluster.sg(A)
@@ -89,10 +89,10 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
         clustering.mm <- mod.max(A)
       }
       error.mm <- 1 - cluster.acc(clustering.mm, z)
-      dplyr::tibble(K = K, n = n, 
-                    error = error, 
+      dplyr::tibble(K = K, n = n,
+                    error = error,
                     error.ssc = error.ssc,
-                    error.mm = error.mm) %>% 
+                    error.mm = error.mm) %>%
         return()
     } %>% 
       return()
@@ -102,7 +102,6 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
 }
 
 gc()
-
 clustering.df %>%
   dplyr::group_by(K, n) %>%
   dplyr::summarise(med.err = median(error),

### 2021-03-24 clustering-sim-k.R

@@ -81,7 +81,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       A <- draw.graph(P)
       clustering <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error <- 1 - cluster.acc(clustering, z)
-      clustering.ssc <- ssc(A, K, n ** -1 * K, normalize = TRUE, scale = FALSE)
+      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
       if (K == 2) {
         clustering.mm <- cluster.sg(A)

### 2021-04-03 short-summary.Rmd
wrote a short summary of work thus far + future work

--- 2021-03-23 summary.Rmd
@@
 ---
-title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation"
-authors: John Koo, Minh Tang, Michael Trosset
-# author: John Koo
-# author2: Minh Tang
-# author3: Michael Trosset
+title: 'Summary of Dissertation Work Thus Far and Future Steps'
+author: John Koo
 output:
@@
     citation_package: natbib
-    number_sections: yes
-# output:
-#   bookdown::pdf_document2:
-#     citation_package: natbib
-#     number_citations: yes
 # output: html_document
-# output: rticles::rss_article
-fontsize: 12pt
-# geometry: "left=1in,right=1in,top=1in,bottom=1in"
 urlcolor: blue
@@
 - \usepackage{amsthm}
-bibliography: misc.bib
-abstract: |
-  We connect two random graph models, the Popularity Adjusted Block Model 
-  (PABM) and the Generalized Random Dot Product Graph (GRDPG), demonstrating
-  that a PABM is a GRDPG in which communities correspond to certain mutually
-  orthogonal subspaces of latent positions. This insight leads to the
-  construction of improved algorithms for community detection and parameter
-  estimation with PABM. Using established asymptotic properties of Adjacency
-  Spectral Embedding (ASE) for GRDPG, we derive asymptotic properties of these
-  algorithms, including algorithms that rely on Sparse Subspace Clustering
-  (SSC). We illustrate these properties via simulation.
+bibliography: short-summary.bib
 ---
@@
 ```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      # eval = FALSE,
+knitr::opts_chunk$set(echo = TRUE, 
                       comment = NA, 
@@
                       fig.align = 'center', 
-                      fig.height = 3, 
-                      fig.width = 5, 
+                      fig.height = 5, 
+                      fig.width = 12, 
                       fig.dpi = 300)
-
 options(xtable.comment = FALSE, 
@@
 ```
-
-```{r}
-import::from(magrittr, `%>%`)
-library(plot.matrix)
-library(ggplot2)
-library(mclust)
-
-theme_set(theme_bw())
-```
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 
-# Introduction
+# Introduction and Research Goal/Scope
 
-Statistical analysis on graphs or networks often involves the partitioning of 
-a graph into disconnected subgraphs or clusters. This is often motivated by the 
-assumption that there exist underlying and unobserved communities to which each 
-vertex of the graph belongs, and edges between pairs of vertices are determined 
-by drawing from a probability distribution based on the community relationships 
-between each pair. The goal of the analysis then is population community 
-detection, or the recovery of the true underlying community labels for each 
-vertex, up to permutation (with some additional parameter estimation being of 
-possible interest), assuming some underlying probability model. One such model
-is the Stochastic Block Model (SBM), first proposed by 
-\citet{doi:10.1080/0022250X.1971.9989788}, which assumes that the edge 
-probability from one vertex to another follows a Beronulli distribution with 
-fixed probabilities for each pair of community labels. Other random 
-graph models have been proposed and studied, such as the Degree-Corrected Block 
-Model (DCBM), introduced by \citet{Karrer_2011}, which is a generalization of 
-the SBM. The Popularity Adjusted Block Model (PABM) was then introduced by 
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the DCBM to 
-address the heterogeneity of edge probabilities within and between communities 
-while still maintaining distinct community structure.
+Statistical analysis on graphs or networks often involves the partitioning of a graph into disconnected subgraphs or clusters, i.e., finding some function $f : [n] \mapsto [K]$.
+This is often motivated by the assumption that there exist underlying and unobserved communities to which each vertex of the graph belongs, and edges between pairs of vertices are determined by drawing from a probability distribution based on the community relationships between each pair.
+The goal of the partitioning or clustering then is community detection, or the recovery of the true underlying community labels for each vertex, up to permutation (with some additional parameter estimation being of possible interest), assuming some underlying probability model.
 
-The underlying similarity among the SBM, PABM, and other such models is that 
-they involve a symmetric edge probability matrix $P \in [0, 1]^{n \times n}$ 
-where $n$ is the number of vertices in the graph. An undirected and unweighted 
-graph is then drawn from this edge probability matrix such that the existence 
-of an edge between each pair of vertices $i$ and $j$ is given by 
-$\Bernoulli(P_{ij})$. For example, for the SBM with two communities for which 
-the within-community edge probability is $\xi$ and the between-community edge 
-probability is $\eta$, the entries of $P$ consist of $P_{ij} = \xi$ if $i$ and 
-$j$ are in the same community and $i \neq j$, $P_{ij} = \eta$ if $i$ and $j$ 
-belong to separate communities, and $P_{ii} = 0$.
+One example of such a model is the Stochastic Block Model (SBM) with two communities.
+In this case, we may draw edges between communities independently such that the distribution of each edge within community 1 is $Bernoulli(p)$, the distribution of each edge within community 2 is $Bernoulli(q)$, and the distribution of each edge between communities 1 and 2 is $Bernoulli(r)$.
+Other models have been proposed, such as the Degree-Corrected Block Model (DCBM), Mixed Membership Stochastic Block Model (MMSBM), and Popularity Adjusted Block Model (PABM), which are often considered generalizations of the SBM.
+All such models involve undirected, unweighted, hollow graphs (although there are some straightforward generalizations to other types of graphs, in particular to weighted graphs).
+The information in this graph is stored in adjacency $A \in \{0, 1\}^{n \times n}$ where $A_{ij} = 1$ if there is an edge between vertices $i$ and $j$, and 0 otherwise.
 
-The Random Dot Product Graph (RDPG) model proposed by 
-\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli 
-edge probabilities. Under this model, each vertex of the graph can be
-represented by a point in some latent space such that the edge
-probability between any pair of vertices is given by their corresponding dot
-product in the latent space, i.e., given a latent positions
-$x_1, ..., x_n \in \mathbb{R}^d$, the edge probability matrix is 
-$P = X X^\top$ where 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. The SBM is 
-equivalent to a special case of the RDPG model in which all vertices of a 
-given community share the same position in the latent space 
-\cite{lyzinski2014}. It has also been shown that similar 
-random graph models, including the DCBM, can be represented in this way
-\cite{lyzinski2014} \cite{rubindelanchy2017consistency}. An analogous 
-property exists for the PABM but not for the RDPG model but under the 
-*Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
-explored in this paper and exploited to construct algorithms for community 
-detection and parameter estimation for the PABM.
+The underlying similarity among all of these Bernoulli edge graphs is the edge probability matrix, $P \in [0, 1]^{n \times n}$.
+Each element $P_{ij}$ is the probability of the existence of an edge between vertices $i$ and $j$ ($P$ is then symmetric and $P_{ii} = 0$ $\forall i$).
+The adjacency matrix $A$ is then drawn such that $A_{ij} \stackrel{indep}{\sim} P_{ij}$ for each $i < j$ and $A_{ji} = A_{ij}$.
+For instance, in the SBM example from before, $P_{ij} = p$ if vertices $i$ and $j$ both belong to community 1, $P_{ij} = q$ if they both belong to community 2, and $P_{ij} = r$ if one is in community 1 and the other is in community 2.
 
-In this paper, we will only consider undirected graphs, that is the edge weight 
-from vertex $i$ to vertex $j$ is equal to the edge weight in the opposite 
-direction, from vertex $j$ to vertex $i$. Furthermore, we will only consider 
-unweighted graphs with binary ($0$ or $1$) edge weights We will also assume that 
-graphs are hollow, i.e., there are no edges from a vertex to itself. All such 
-graphs can be represented by a symmetric adjacency matrix 
-$A \in \{0, 1\}^{n \times n}$ for which $A_{ij} = 1$ if there exists an edge 
-between vertices $i$ and $j$ and $0$ otherwise, and $A$ is an element-wise 
-independent Bernoulli draw from a symmetric edge probability matrix 
-$P \in [0, 1]^{n \times n}$.
-
-# Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
-
-## The popularity adjusted block model and the generalized random dot product graph
+## The Popularity Adjusted Block Model
 
@@
 
+## The Generalized Random Dot Product Graph
+
+The Random Dot Product Graph (RDPG) is another graph model with Bernoulli edge probabilities.
+Under this model, each vertex of the graph can be represented by a point in some latent space such that the edge probability between any pair of vertices is given by their corresponding dot product in the latent space.
+It can be shown that any Bernoulli edge graph with positive semidefinite edge probability matrix $P$ can be represented as a RDPG.
+More specifically, if $P = V \Sigma V^\top$ is the spectral decomposition of $P$ such that $\Sigma$ is a diagonal matrix of $d$ positive eigenvalues, then $V \Sigma^{1/2}$ is one representation of the latent positions under the RDPG framework.
+
+\begin{example}[Connecting the SBM to the RDPG]
+Consider the same SBM as before.
+Let community 1 have $n_1$ vertices and community 2 have $n_2$ vertices such that $n_1 + n_2 = n$.
+Let the edge probability matrix $P$ be organized by community such that the $P^{(kl)}$ block represents edge probabilities between communities $k$ and $l$.
+In this case, $P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} & P^{(22)} \end{bmatrix}$ where each block is a constant value, e.g., $P^{(11)}_{ij} = p$.
+Then one RDPG representation of this SBM is:
+
+$$X = \begin{bmatrix} 
+\sqrt{p} & 0 \\
+\vdots & \vdots \\
+\sqrt{p} & 0 \\
+\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\ 
+\vdots & \vdots \\
+\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
+\end{bmatrix}
+\in \mathbb{R}^{n \times 2}$$
+
+where the first $n_1$ rows are $\begin{bmatrix} \sqrt{p} & 0 \end{bmatrix}$ and the next $n_2$ rows are $\begin{bmatrix} \sqrt{r^2 / p} & \sqrt{q - r^2 / p} \end{bmatrix}$. Then it can be shown that
+
+$$P = X X^\top$$
+\end{example}
+
+If $P$ is not positive semidefinite, then it cannot be the edge probability 
+matrix of a RDPG, but it can be the edge probability matrix of a *Generalized*
+Random Dot Product Graph (GRDPG).[^grdpg]
+
+[^grdpg]: It is trivial to show that *all* Bernoulli graphs are GRDPGs.
+
 \begin{definition}[Generalized Random Dot Product Graph] 
@@
 \begin{remark}
-Like the RDPG, the latent positions of a GRDPG are not unique 
-\cite{rubindelanchy2017statistical}. 
+Like the RDPG, the latent positions of a GRDPG are not unique. 
 More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
@@
 
-## Connecting the PABM to the GRDPG
+# Literature Review and Previous Work
+
+The PABM was introduced by \citet{307cbeb9b1be48299388437423d94bf1}, who 
+intuited that a spectral clustering type of approach would be appropriate for 
+this problem. Their work was furthered by \citet{noroozi2019estimation}, who 
+used SSC for community detection. SSC is typically performed on points in 
+Euclidean space, which suggests a spectral embedding, but the authors of this 
+paper instead performed SSC on $P$ and $A$ directly. They were able to show 
+that SSC on $P$ results in perfect community detection.
+
+SSC is described as follows. Given $x_1, ..., x_n \in \mathbb{R}^d$ such 
+that each point lies on one of $K$ subspaces, the goal is to determine 
+which subspace each point belongs to. We might also want to add some noise to 
+the points so that they don't lie exactly on their respective subspaces.
+
+SSC is performed by solving an optimization problem for each observed point.
+Given $X \in \mathbb{R}^{n \times d}$ with vectors 
+$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
+$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
+for each $i = 1, ..., n$. The solutions are collected collected into matrix 
+$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
+affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
+subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
+subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
+If $X$ instead represents points near $K$ subspaces with some noise, a final 
+graph partitioning step is performed (e.g., edge thresholding or spectral 
+clustering).
+
+In practice, SSC is often performed by solving the LASSO problems
+
+\begin{equation} \label{eq:ssc}
+c_i = \arg\min_c \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
+\end{equation}
+
+for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
+into $C$ and $B$ as before.
+
+\begin{definition}[Subspace Detection Property] 
+Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
+points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the 
+solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of 
+$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from 
+different subspaces, then $X$ obeys the subspace detection property. 
+\end{definition}
+
+\citet{pmlr-v28-wang13} showed that in the case where the subspaces are 
+orthogonal (or at least the cosine of their angles is very small) and the 
+points are close enough to their respective subspaces, the subspace detection
+property will hold given an appropriate choice of $\lambda$. It turns out that 
+this is precisely the case for a particular embedding choice for the PABM, of 
+which the properties have been described by 
+\citet{rubindelanchy2017statistical}.
+
+# Completed Work Thus Far
+
+The PABM can be expressed as a GRDPG. In particular, there is one expression 
+of this that results in a very simple structure such that each community 
+corresponds to a subspace and the subspaces are mutually orthogonal.
 
@@
 
-# Methods
+This connection leads to the following property:
 
-Two inference objectives arise from the PABM:
+\begin{theorem}
+\label{theorem3}
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability
+matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
+from different communities.
+\end{theorem}
 
-1. Community membership identification (up to permutation).
-2. Parameter estimation (estimating $\lambda^{(kl)}$'s).
-
-In our methods, we assume that $K$, the number of communities, is known 
-beforehand and does not require estimation.
-
-## Related work
-
-\citeauthor{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, 
-used Modularity Maximization (MM) and the Extreme Points
-(EP) algorithm \cite{le2016} for community detection and parameter estimation. 
-They were able to show that as the sample size increases, the proportion of
-misclassified community labels (up to permutation) goes to 0.
-
-\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for 
-community detection in the PABM. 
-SSC is performed by solving an optimization problem for each observed point.
-Given $X \in \mathbb{R}^{n \times d}$ with vectors 
-$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
-$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
-for each $i = 1, ..., n$. The solutions are collected collected into matrix 
-$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
-affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
-subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
-subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
-If $X$ instead represents points near $K$ subspaces with some noise, a final 
-graph partitioning step is performed (e.g., edge thresholding or spectral 
-clustering).
-
-In practice, SSC is often performed by solving the LASSO problems
-
-\begin{equation} \label{eq:ssc}
-c_i = \arg\min_c \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
-\end{equation}
-
-for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
-into $C$ and $B$ as before.
-
-\begin{definition}[Subspace Detection Property]. 
-Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
-points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the 
-solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of 
-$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from 
-different subspaces, then $X$ obeys the subspace detection property. 
-\end{definition}
-
-\begin{remark} 
-In practice, sample $X$ often does not obey the subspace detection
-property. In such cases, $B$ is treated as an affinity matrix for a graph which 
-is then partitioned into $K$ subgraphs to obtain the clustering. On the other 
-hand, if $X$ does obey the subspace detection property, $B$ describes a graph 
-with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
-property holds, there are exactly $K$ subgraphs which map to each subspace, 
-but it could be the case that some of the subspaces are represented by 
-multiple disconnected subgraphs. 
-\end{remark}
-
-Theorem \ref{theorem2} suggests that SSC is appropriate for community 
-detection for the PABM. 
-More precisely, Theorem \ref{theorem2} says that each community consists of a
-$K$-dimensional subspace, and together the subspaces lie in $\mathbb{R}^{K^2}$. 
-The natural approach then is to perform SSC on the ASE of $P$ or $A$.
-\citeauthor{noroozi2019estimation} instead applied SSC to $P$ and $A$, 
-foregoing embedding altogether.
-
-Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
-the subspace detection property holds for $XU$, which is an ASE of $P$. More 
-specifically, if points lie exactly on mutually orthogonal subspaces, then the 
-subspace detection property will hold with probability 1, and this is exactly 
-the case for the PABM (Theorem \ref{theorem2}). 
-Much of our work is then built on
-\citeauthor{rubindelanchy2017statistical}, who describe the convergence 
-behavior of the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13},
-who show the necessary conditions for the subspace detection property to hold 
-in noisy cases where the points lie near subspaces.
-
-## Community detection
+In practice, however, we observe $A$ rather than $P$. Using the asymptotic
+properties of the Adjacency Spectral Embedding (ASE) of the GRDPG
+\cite{rubindelanchy2017statistical}, it can be shown that constructing 
+$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}$
+corresponding to between-community pairings going to 0 with probability 1.
 
@@
 \end{algorithm}
-
-We previously stated one possible set of latent positions that result in the 
-edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
-(or can estimate) $XU$ directly, then both the community detection and 
-parameter identification problem are trivial since $U$ is orthonormal and 
-fixed for each value of $K$. However, direct identification or estimation of
-$XU$ is not possible \cite{rubindelanchy2017statistical}.
-
-If we decompose $P = Z I_{pq} Z^\top$, then
-$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
-the exact edge probability matrix, we cannot recover the "original" latent 
-positions $XU$. Note that unlike in the case of the RDPG, $Q$ is not 
-necessarily an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
-$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
-$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
-us from using the properties of $XU$ directly. In particular, if 
-$Q \in \mathbb{O}(n)$, then we could use the fact that 
-$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
-and $j$ are in different communities.
-
-The explicit form of $XU$ represents points in 
-$\mathbb{R}^{K^2}$ such that points within each community lie on 
-$K$-dimensional orthogonal subspaces. 
-Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
-but retains the property that each community is represented by a 
-$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
-correspond to each community, suggesting the use of SSC. In this paper, we will 
-use a different and leave the properties of SSC on the ASE of $A$ as future
-work.
-
-\begin{theorem}
-\label{theorem3}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
-from different communities.
-\end{theorem}
-
-Theorem \ref{theorem3} provides perfect community detection given $P$. 
-Letting $|B|$ be the affinity matrix for graph $G$, $G$ is partitioned into 
-at least $K$ disjoint subgraphs since each of the $K$ communities have no 
-edges between them. Similar to the subspace detection property, it could be 
-the case that some of the communities are represented by multiple disjoint
-subgraphs in $G$, in which case additional reconstruction is required to
-identify the communities exactly.
-
-Using $A$ instead of $P$ introduces error, which converges to $0$ 
-almost surely:
 
@@
       $c_i = \arg\min_{\beta} 
-      \frac{1}{2} ||v_i - V_{-i} \beta||_2^2 + \lambda ||\beta||_1$.\;
+      \frac{1}{2} ||\sqrt{n} v_i - \sqrt{n} V_{-i} \beta||_2^2 + 
+      \lambda ||\beta||_1$.\;
       Assign $\tilde{c}_i = \begin{bmatrix} 
@@
 positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
-$\exists \lambda > 0$ such that $\hat{B}_n$, the affinity matrix
-outputted by SSC using $\hat{V}_n$ and $\lambda$, has $ij^{th}$ entry
-$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$ $\forall i, j$ in different 
-communities. 
+$\exists \lambda > 0$ such that as $n \to \infty$, $\hat{V}_n$ obeys the 
+subspace detection property with probability 1.
 \end{theorem}
 
-## Parameter estimation
+We are also able to show convergence properties for parameter estimation using 
+the ASE, the details of which have been omitted for brevity but can be found 
+in the main paper.
 
-For any edge probability matrix $P$ for the PABM such that the rows and 
-columns are organized by community, the $kl$^th^ block is an outer product 
-of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
-Therefore, given $P^{(kl)}$, $\lambda^{(kl)}$ and $\lambda^{(lk)}$ are 
-solvable exactly (up to multiplication by $-1$) using singular value 
-decomposition. More specifically, let $P^{(kl)} = \sigma^2 u v^\top$ be the 
-singular value decomposition of $P^{(kl)}$. $u \in \mathbb{R}^{n_k}$ and 
-$v \in \mathbb{R}^{n_l}$ are vectors and 
-$\sigma^2 > 0$ is a scalar. Then $\lambda^{(kl)} = \pm \sigma u$ and 
-$\lambda^{(lk)} = \pm \sigma v$. 
-Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
-simply use plug-in estimators (algorithm 3), which converge to the true 
-parameters.
+# Future Work
 
-\begin{algorithm}[t]
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{PABM parameter estimation.}
-  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
-  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
-  Arrange the rows and columns of $A$ by community such that each 
-  $A^{(kl)}$ block consists of estimated edge probabilities between 
-  communities $k$ and $l$.\;
-  \For {$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
-    block.\;
-    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
-    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
-    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
-    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
-  }
-\end{algorithm}
+So far, we've shown that there is a very convenient GRDPG representation of 
+the PABM that makes community detection almost trivially simple. 
+More specifically, we showed that an embedding of the PABM results in 
+orthogonal subspaces and noise that goes to 0 with probability 1, which makes 
+for a very straightforward application of SSC. It can also be shown that this 
+type of embedding has the same properties of the SBM and DCBM, which are really
+just special cases of the PABM.
 
-\begin{theorem}
-\label{theorem6}
-Under regularity and sparsity assumptions, and under the further 
-assumption that $K$ is fixed and community labels are known,
+Furthermore, all Bernoulli graphs can be
+represented as a GRDPG, and it may be the case that other, more complicated 
+graphs have nice ASE representations that lead to community detection methods. 
+In particular, we are currently exploring the nature of RDPGs and GRDPGs that
+result from latent positions for which each community lies on a manifold rather
+than on a subspace. 
+Much of this work is based on work by \citet{trosset2020learning} who showed 
+convergence properties for graphs that result from latent positions that lie 
+on a one-dimensional manifold.
 
-\begin{equation} \label{eq:thm6}
-\max_{k, l \in \{1, ..., K\}} 
-||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
-O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
-\end{equation}
-\end{theorem}
-
-# Simulated Examples
-
-For each simulation, community labels are drawn from a multinomial 
-distribution, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
-from two types of joint distributions depending on whether $k = l$, the edge
-probability matrix $P$ is constructed using 
-the popularity vectors, and finally an unweighted and undirected adjacency 
-matrix $A$ is drawn from $P$. OSC is then used for community detection, 
-and this method is compared against SSC \cite{noroozi2019estimation}
-\cite{soltanolkotabi2014} and MM \cite{igraph}
-\cite{307cbeb9b1be48299388437423d94bf1}. True 
-community labels are used with Algorithm 3 to estimate the popularity 
-vectors $\{\lambda^{(kl)}\}_K$, and this method is then compared against an
-MLE-based estimator described in \citeauthor{noroozi2019estimation} and
-\citeauthor{307cbeb9b1be48299388437423d94bf1}.
-
-Modularity Maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
-used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$. 
-For these simulations, the EP algorithm was used for $K = 2$, and for $K > 2$, 
-the Louvain algorithm \cite{Blondel_2008} was used instead.
-
-Two implementations of SSC are shown here. The first method, denoted as 
-SSC-A, treats the columns of the adjacency matrix $A$ as points in
-$\mathbb{R}^n$, as described in \citeauthor{noroozi2019estimation}. The second 
-method, denoted as SSC-ASE, first embeds $A$ and then performs SSC on the 
-embedding, as described in algorithm 2. The sparsity parameter $\lambda$ was 
-chosen via a preliminary cross-validation experiment. For the final clustering 
-step, a Gaussian Mixture Model was fit on the normalized Lapacian eigenmap of 
-the affinity matrix $B$.
-
-## Balanced communities
-
-In each simulation, community labels $z_1, ..., z_n$ were drawn 
-from a multinomial distribution with mixture parameters 
-$\{\alpha_1, ..., \alpha_K\}$, then $\{\lambda^{(kl)}\}_K$ according to the 
-drawn community labels, $P$ was constructed using the drawn 
-$\{\lambda^{(kl)}\}_K$, and $A$ was drawn from $P$ by 
-$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$. Each simulation has a unique
-edge probability matrix $P$.
-
-For these examples, we set the following parameters:
-
-* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
-* Number of underlying communities $K = 2, 3, 4$
-* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
-community label has an equal probability of being drawn)
-* Community labels 
-$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
-* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
-* Between-group popularities 
-$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$
-
-$50$ simulations were performed for each $(n, K)$ pair.
-
-Fig. \ref{fig:clust_err_k} show that for large $n$, OSC results in 
-a misclustering rate of 0. Sparse subspace clustering produces similar results 
-for $K > 2$.
-
-```{r clust_err_k, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
-                    colour = 'SSC-A'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
-```
-
-Theorem \ref{theorem4} implies that OSC will result in not just in the error
-rate converging to 0 but the error *count* as well.
-We explore this in Fig. \ref{fig:clust_err_ct_sim}.
-
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error count', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm * n,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
-                    colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2 * n,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
-                    colour = 'SSC-A'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
-```
-
-Given ground truth community labels, Algorithm 3 and the MLE-based plug-in 
-estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
-perform similarly, with root mean square error decaying at rate approximately 
-$n^{-1/2}$.
-
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-rmse.df <- readr::read_csv('rmse-k.csv')
-
-rmse.df %>% 
-  na.omit() %>% 
-  dplyr::group_by(K, n) %>% 
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-  dplyr::ungroup() %>% 
-  ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 3'), width = .1) + 
-  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  labs(y = 'RMSE', colour = NULL) + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10() + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
-```
-
-## Imbalanced communities
-
-Simulations performed in this section are similar to those in the previous 
-section with the exception of the mixture parameters 
-$\{\alpha_1, ..., \alpha_K\}$ used to draw community labels from the 
-multinomial distribution. For these examples, we set the following parameters:
-
-* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
-* Number of underlying communities $K = 2, 3, 4$
-* Mixture parameters $\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}$ for 
-$k = 1, ..., K$
-* Community labels 
-$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
-* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
-* Between-group popularities 
-$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$
-
-$50$ simulations were performed for each $(n, K)$ pair.
-
-Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
-similar results as in the balanced communities case, with both OSC and 
-SSC resulting in no misclustered vertices for a 
-sufficiently large sample size. However, Fig. \ref{fig:lambda_est_k_imba} 
-suggests that while Algorithm 3 retains $\sqrt{n}$ efficiency, the MLE-based 
-plug-in estimator is less efficient for this setup.
-
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are imbalanced.
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
-                    colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
-                    colour = 'SSC-A'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
-```
-
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are imbalanced.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error count', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm * n,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
-                    colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2 * n,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
-                    colour = 'SSC-A'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
-```
-
-```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')
-
-rmse.df %>% 
-  na.omit() %>% 
-  dplyr::group_by(K, n) %>% 
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-  dplyr::ungroup() %>% 
-  ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 3'), width = .1) + 
-  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  labs(y = 'RMSE', colour = NULL) + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10() + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
-```
-
-## Additional experiments
-
-Using the same set of parameters for generating $P$ and $A$ as in the 
-balanced communities examples for $K = 2$, we generated one instance of $A$ for
-each $n$ and constructed $B$ according to Algorithm 3 to verify that 
-as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
-in different clusters. Furthermore, the distribution of these inner products 
-should be approximately normal.
-
-```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', fig.height = 2, fig.width = 4}
-densities.df <- readr::read_csv('densities.csv')
-
-ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
-  labs(x = expression(n * v[i]^T * v[j]),
-       colour = 'n') + 
-  scale_x_continuous(breaks = seq(-5, 5)) + 
-  scale_colour_brewer(palette = 'Set1')
-```
-
-# Real data examples
-
-```{r}
-import::here(cluster.pabm, ssc, plot.A, .from = '~/dev/pabm-grdpg/functions.R')
-
-n.edges <- 20566
-
-butterfly <- R.matlab::readMat('data/Raw_butterfly_network.mat')
-
-A <- butterfly$W.butterfly0
-z <- butterfly$labels
-
-keep <- c(6, 2, 9, 4)
-A <- A[z %in% keep, z %in% keep]
-z <- z[z %in% keep]
-n <- length(z)
-K <- 4
-
-delta <- quantile(A[upper.tri(A)], 1 - n.edges / (n * (n - 1) / 2))
-A <- (A > delta) * 1
-
-z.hat <- cluster.pabm(A, K)
-
-# z.hat.mm <- A %>% 
-#   igraph::graph_from_adjacency_matrix(mode = 'undirected') %>% 
-#   igraph::cluster_louvain() %>% 
-#   igraph::membership()
-
-# ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
-```
-
-In the first real data example, we applied OSC to the Leeds Butterfly 
-dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
-butterflies across 10 species. The graph was modified to match the example 
-from \citeauthor{noroozi2019estimation}: Only the 4 most frequent species were 
-considered, and the similarities were discretized to $\{0, 1\}$ via 
-thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
-by the resultant clustering.
-
-Comparing against the ground truth species labels, OSC achieves an 
-accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
-\citeauthor{noroozi2019estimation} achieved an adjusted Rand index of 73\% 
-using sparse subspace clustering on the same dataset.
-
-```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.', cache = FALSE}
-plot.A(A, z.hat)
-```
-
-```{r}
-# this portion is from the code provided by Sengupta and Chen
-# the data were generated by scripts from S&C, then saved as RDS files
-
-bmp <- readRDS('data/bmp.rds')
-A.bmp <- bmp$A
-z.bmp <- bmp$z
-
-pb <- readRDS('data/pb.rds')
-A.pb <- pb$A
-z.pb <- pb$z
-
-dblp <- readRDS('data/dblp.rds')
-A.dblp <- dblp$A
-z.dblp <- dblp$z
-```
-
-```{r, cache = TRUE}
-zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = FALSE)
-bmp.acc <- mean(zhat.bmp == z.bmp)
-bmp.acc <- ifelse(bmp.acc < .5, 1 - bmp.acc, bmp.acc)
-
-zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = FALSE, 
-                        p = 2, q = 0, d.eigenmap = 2)
-pb.acc <- mean(zhat.pb == z.pb)
-pb.acc <- ifelse(pb.acc < .5, 1 - pb.acc, pb.acc)
-
-zhat.dblp <- cluster.pabm(A.dblp, 2, use.all = TRUE, normalize = FALSE)
-dblp.acc <- mean(zhat.dblp == z.dblp)
-dblp.acc <- ifelse(dblp.acc < .5, 1 - dblp.acc, dblp.acc)
-
-zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3, normalize = TRUE)
-bmp.acc.ssc <- mean(zhat.bmp.ssc == z.bmp)
-bmp.acc.ssc <- ifelse(bmp.acc.ssc < .5, 1 - bmp.acc.ssc, bmp.acc.ssc)
-
-zhat.pb.ssc <- ssc(A.pb, 2, .008, normalize = TRUE)
-pb.acc.ssc <- mean(zhat.pb.ssc == z.pb)
-pb.acc.ssc <- ifelse(pb.acc.ssc < .5, 1 - pb.acc.ssc, pb.acc.ssc)
-
-zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2, normalize = FALSE)
-dblp.acc.ssc <- mean(zhat.dblp.ssc == z.dblp)
-dblp.acc.ssc <- ifelse(dblp.acc.ssc < .5, 1 - dblp.acc.ssc, dblp.acc.ssc)
-
-```
-
-In the second example, we applied OSC to the British MPs Twitter 
-network \cite{greene2013producing}, the Political Blogs network 
-\cite{10.1145/1134271.1134277}, and the DBLP network \cite{NIPS2009_3855} 
-\cite{10.1007/978-3-642-15880-3_42}. For this data analysis, we subsetted the 
-data as described by 
-\citeauthor{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
-networks. Our methods underperformed compared to modularity maximization, 
-although performance is comparable. In 
-addition, OSC's runtime is much lower than that of modularity 
-maximization.
-
-```{r}
-dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
-              # Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
-              `MM` = c(round(1 / 329, 3),
-                       round(61 / 1222, 3),
-                       round(62 / 2203, 3)),
-              `SSC-ASE` = round(1 - c(bmp.acc.ssc,
-                                      pb.acc.ssc,
-                                      dblp.acc.ssc), 3),
-              `OSC` = round(1 - c(bmp.acc, 
-                                  pb.acc,
-                                  dblp.acc), 3)) %>% 
-  knitr::kable(caption = paste('Community detection error rates', 
-                               'for modularity maximization,',
-                               'sparse subspace clustering, ',
-                               'and OSC.'),
-               format = 'latex')
-```
-
-```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.', fig.dpi = 10}
-gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
-                        plot.A(A.pb, zhat.pb),
-                        plot.A(A.dblp, zhat.dblp),
-                        ncol = 3)
-```
-
-In the third example, we consider the Karantaka villages data studied by 
-\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` networks from 
-villages 12, 31, and 46 at the household level. The label of interest is the 
-religious affiliation. The networks were truncated to religions "1" and "2", 
-and vertices of degree 0 were removed.
-
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
-The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
-K <- 2
-hh.12 <- readRDS('data/village-hh-12.rds')
-hh.31 <- readRDS('data/village-hh-31.rds')
-hh.46 <- readRDS('data/village-hh-46.rds')
-
-# this part of code taken from sengupta & chen
-setwd('~/dev/pabm-grdpg/codes_and_data')
-require(MASS);require("clue")
-source("EPalgosim.R") 
-source("functions.R")
-mm <- function(A.hh) {
-  b.can = EPalgo(A.hh,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A.hh, b=b.can[,i])   # fit PABM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  return(b.can[,foo1])   # community assignment that maximises Q.PA
-}
-
-A.hh.12 <- hh.12$A
-z.hh.12 <- hh.12$z
-zhat.hh.12 <- cluster.pabm(A.hh.12, K,
-                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
-zhat.hh.ssc.12 <- ssc(A.hh.12, K, lambda = .1)
-zhat.hh.mm.12 <- mm(A.hh.12)
-
-A.hh.31 <- hh.31$A
-z.hh.31 <- hh.31$z
-zhat.hh.31 <- cluster.pabm(A.hh.31, K,
-                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
-zhat.hh.ssc.31 <- ssc(A.hh.31, K, lambda = .05)
-zhat.hh.mm.31 <- mm(A.hh.31)
-
-A.hh.46 <- hh.46$A
-z.hh.46 <- hh.46$z
-zhat.hh.46 <- cluster.pabm(A.hh.46, K,
-                           use.all = TRUE, normalize = FALSE, p = 2, q = 0)
-zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .001)
-zhat.hh.mm.46 <- mm(A.hh.46)
-
-gridExtra::grid.arrange(plot.A(A.hh.12, zhat.hh.12),
-                        plot.A(A.hh.31, zhat.hh.31),
-                        plot.A(A.hh.46, zhat.hh.46),
-                        nrow = 1)
-```
-
-```{r, eval = TRUE}
-hh.acc.12 <- mean(zhat.hh.12 == z.hh.12)
-hh.acc.12 <- ifelse(hh.acc.12 < .5, 1 - hh.acc.12, hh.acc.12)
-hh.acc.ssc.12 <- mean(zhat.hh.ssc.12 == z.hh.12)
-hh.acc.ssc.12 <- ifelse(hh.acc.ssc.12 < .5, 1 - hh.acc.ssc.12, hh.acc.ssc.12)
-hh.acc.mm.12 <- mean(zhat.hh.mm.12 == z.hh.12)
-hh.acc.mm.12 <- ifelse(hh.acc.mm.12 < .5, 1 - hh.acc.mm.12, hh.acc.mm.12)
-
-hh.acc.31 <- mean(zhat.hh.31 == z.hh.31)
-hh.acc.31 <- ifelse(hh.acc.31 < .5, 1 - hh.acc.31, hh.acc.31)
-hh.acc.ssc.31 <- mean(zhat.hh.ssc.31 == z.hh.31)
-hh.acc.ssc.31 <- ifelse(hh.acc.ssc.31 < .5, 1 - hh.acc.ssc.31, hh.acc.ssc.31)
-hh.acc.mm.31 <- mean(zhat.hh.mm.31 == z.hh.31)
-hh.acc.mm.31 <- ifelse(hh.acc.mm.31 < .5, 1 - hh.acc.mm.31, hh.acc.mm.31)
-
-hh.acc.46 <- mean(zhat.hh.46 == z.hh.46)
-hh.acc.46 <- ifelse(hh.acc.46 < .5, 1 - hh.acc.46, hh.acc.46)
-hh.acc.ssc.46 <- mean(zhat.hh.ssc.46 == z.hh.46)
-hh.acc.ssc.46 <- ifelse(hh.acc.ssc.46 < .5, 1 - hh.acc.ssc.46, hh.acc.ssc.46)
-hh.acc.mm.46 <- mean(zhat.hh.mm.46 == z.hh.46)
-hh.acc.mm.46 <- ifelse(hh.acc.mm.46 < .5, 1 - hh.acc.mm.46, hh.acc.mm.46)
-
-# knitr::kable(dplyr::tibble(
-#   Method = c('OSC', 'SSC-ASE', 'Mod. Max.'),
-#   `Village 12` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
-#   `Village 31` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
-#   `Village 46` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
-#   caption = paste('Community detection error rates for identifying',
-#                   'household religion.'))
-
-knitr::kable(dplyr::tibble(
-  Network = c('Village 12', 'Village 31', 'Village 46'),
-  MM = round(1 - c(hh.acc.mm.12, hh.acc.mm.31, hh.acc.mm.46), 3),
-  `SSC-ASE` = round(1 - c(hh.acc.ssc.12, hh.acc.ssc.31, hh.acc.ssc.46), 3),
-  `OSC` = round(1 - c(hh.acc.12, hh.acc.31, hh.acc.46), 3)
-),
-caption = paste('Community detection error rates for identifying',
-                  'household religion.'),
-format = 'latex')
-```
-
-# Discussion
-
-This paper shows the connection between the PABM and the GRDPG, namely that 
-a PABM graph can be represented as a union of orthogonal subspaces in an 
-embedding under the GRDPG framework. We then exploited this relationship to 
-develop community detection and parameter estimation methods. In fact, we can 
-represent any graph with Bernoulli edges as a GRDPG, and in the PABM case, 
-it turns out that this relationship leads to a very straightforward 
-applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
-\citeauthor{soltanolkotabi2012}, and \citeauthor{pmlr-v28-wang13}, which lead 
-to asymptotically correct solutions with high probability. Similar methods can 
-be applied for other models, such as the Nested Block Model
-\cite{noroozi2021hierarchy}.
-
-# Proofs
-
-**Proof of Theorem \ref{theorem1}**. 
-This is given by straightforward matrix multiplication. It suffices to show 
-that
-
-$$X U I_{3, 1} U^\top X^\top = 
-\begin{bmatrix} 
-  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}$$
-
-\begin{remark} 
-While we can just perform the matrix multiplication to show the 
-equivalence, it is more illustrative to look at a few intermediate steps. Note 
-that the product of the three inner matrices results in a permutation matrix 
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping 
-positions $2$ and $3$:
-
-$$U I_{3, 1} U^\top = \begin{bmatrix} 
-  1 & 0 & 0 & 0 \\ 
-  0 & 0 & 1 & 0 \\ 
-  0 & 1 & 0 & 0 \\ 
-  0 & 0 & 0 & 1 
-\end{bmatrix} = \Pi$$
-
-Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$ 
-is a spectral decomposition of this permutation matrix. Note that the two fixed 
-points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$ 
-where $i = 1, 4$ corresponding to the locations of the fixed points, and the 
-cycle of order two results in two eigenvalues $\pm 1$ with corresponding 
-eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is 
-swapped.
-\end{remark}
-
-\begin{lemma}
-\label{lemma1}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in (\ref{eq:xy}).
-\end{lemma}
-
-*Proof*. 
-By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
-(\ref{eq:xy}) and $p$ and $q$ are defined as in equations (\ref{eq:p}) and
-(\ref{eq:q}). Alternatively, the spectral decomposition can be written as 
-$P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
-and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some 
-$Q \in \mathbb{O}(p, q)$,
-
-$$X U Q = V |D|^{1/2}$$
-
-Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,
-
-$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$
-
-The right-hand side becomes
-
-$$\begin{split}
-  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-    = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
-    U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-  & = X (X^\top X)^{-1} X^\top
-\end{split}$$
-
-The left-hand side becomes:
-
-$$\begin{split}
-  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-  & = V V^\top
-\end{split}$$
-
-**Proof of Theorem \ref{theorem3}**.
-By Lemma 3.1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
-in Theorem 2. Since $X$ is block diagonal with each block corresponding 
-to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
-with each block corresponding to a community and zeros elsewhere. Therefore, 
-if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
-element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.
-
-**Proof of Theorem \ref{theorem4}**. 
-Let $V_n$ and $\hat{V}_n$ be the eigenvectors of $P$ and $A$ corresponding 
-to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative
-eigenvalues. By \citeauthor{rubindelanchy2017statistical}, for some 
-$W \in \mathbb{O}(K^2)$, and $c > 0$,  
-$||\hat{V} W - V||_{2 \to \infty} = 
-O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big)$. 
-We furthermore have $||V||_{2 \to \infty} = O_P(n^{-1/2})$. Then if
-$(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ correspond to the rows of 
-$V_n$ and $\hat{V}_n$, for $i$ and $j$ in different communities, using the 
-fact that $(v_n^{(i)})^\top v_n^{(j)} = 0$,
-
-$$\begin{split}
-\max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
-= \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} - 
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& = \max_{i, j} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& = ||\hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
-2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 (\hat{V}_n W - V_n) V_n^\top + 
-(\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
-& \leq 2 ||\hat{V}_n W - V_n||_{2 \to \infty} ||V_n||_{2 \to \infty} +
-||\hat{V}_n W - V_n||_{2 \to \infty}^2 \\
-& = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
-\end{split}$$
-
-Then scaling by $n$, we get 
-$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = 
-O_P \big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \big)$.
-
-\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{pmlr-v28-wang13}]
-The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
-defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
-In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents 
-the inradius of the symmetric convex hull of $X$. 
-\end{definition}
-
-\begin{lemma}
-\label{lemma2}
-Let $P_n = V_n \Lambda_n V_n^\top$ be the edge probability matrix of a PABM 
-with $n$ vertices. Let $A_n \sim \Bernoulli(P_n)$ be the adjacency matrix 
-drawn from $P_n$, and let $A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be 
-its approximate spectral decomposition using the $K (K+1) / 2$ most positive 
-and $K (K-1) / 2$ most negative eigenvalues and their corresponding 
-eigenvectors.Without loss of generality, order the rows of $V_n$ and 
-$\hat{V}_n$ by community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the 
-$n_k \times K^2$ matrix of the rows of the $k^{th}$ community in $V_n$ and 
-$\hat{V}_n$ respectively. 
-Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
-as defined by \citeauthor{soltanolkotabi2012}. Then $\mu_n^{(k)} = 0$ 
-with probability 1 $\forall n, k$.
-
-Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
-extended definition of $\mu(\cdot)$ from \citeauthor{pmlr-v28-wang13}. Then
-
-\begin{equation} \label{eq:mu-conv}
-\hat{\mu}_n^{(k)} = 0
-\end{equation}
-\end{lemma}
-
-*Proof*. 
-For the first part, 
-since $v_n^{(k)}$ is orthogonal to $v_n^{(l)}$ $\forall k \neq l$ 
-(Lemma \ref{lemma1}), 
-this falls directly out of Theorem 2.8 of \citeauthor{soltanolkotabi2012}.
-
-For the second part, 
-$\mu(\hat{V}_n^{(k)}) = 
-\max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ 
-where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned 
-by $V_n^{(k)}$. Since 
-$u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ 
-$\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem \ref{theorem2}, this quantity 
-must be 0.
-
-\begin{lemma}
-\label{lemma3}
-Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-$\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
-
-\begin{equation} \label{eq:delta-conv}
-\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{a.s.}{\to} 0
-\end{equation}
-\end{lemma}
-
-**Proof of Theorem \ref{theorem5}**. 
-The basis of this proof is Theorem 6 from \citeauthor{pmlr-v28-wang13}, which 
-states that the subspace detection property holds if the noise is small enough 
-and the subspace inradius is greater than the subspace incoherence for each 
-community $k$.
-
-Let $V_{n, -i}^{(k)}$ be $V_n^{(k)}$ with the $i$^th^ entry removed. Suppose 
-that for each community $k$, there are enough vertices such that for each $i$, 
-$V_{n, -i}^{(k)}$ spans its corresponding subspace (Theorem 2). 
-Then $r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0$. 
-Thus by (\ref{eq:mu-conv}), for each $k$, $r_n^{(k)} > \mu_n^{(k)}$ where 
-$\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})$ and $n$ is large enough such that 
-$\min\limits_{k, i} rank(V_{n, -i}^{(k)}) = K$.
-
-Let $r_n = \min_k r_n^{(k)}$. By (\ref{eq:delta-conv}), 
-$\delta_n \stackrel{a.s.}{\to} 0$. Then as $n\to \infty$, 
-$\delta_n < \min_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}$ 
-with probability 1.
-
-Thus the conditions for the subspace detection property from Theorem 6 from
-\citeauthor{pmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
-
-**Proof of Theorem \ref{theorem6}**. 
-Let $P$ and $A$ be organized by community such that the elements of blocks 
-$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
-$l$.
-
-*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
-probabilities and edges for community $k$.  
-By definition, 
-$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
-signular value decomposition 
-$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
-one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
-singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
-Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
-value decomposition of $A^{(kk)}$, and let 
-$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
-one-dimensional approximation. Define 
-$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
-$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
-$\lambda^{(kk)}$.  
-Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the adjacency 
-spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
-$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
-
-*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
-edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
-By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
-$k = l$ case, we note that the singular value decomposition 
-$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
-$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
-$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
-$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
-$v^{(kl)} = u^{(lk)}$.)  
-Now consider the Hermitian dilation
-
-$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$
-
-which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
-that the spectral decomposition of $M^{(kl)}$ is
-
-$$M^{(kl)} = 
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
-\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$
-
-Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
-latent positions in $\mathbb{R}^2$ given by
-
-$$\begin{bmatrix} 
-  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
-  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
-\end{bmatrix} = 
-\begin{bmatrix} 
-  \lambda^{(kl)} & \lambda^{(kl)} \\ 
-  \lambda^{(lk)} & -\lambda^{(lk)} 
-\end{bmatrix}$$
-
-Now consider
-
-$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
-Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where
-
-$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$
-
-and $E$ is the 
-$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
-in $A^{(kl)}$.  
-Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
-spectral embedding, given by
-
-$$\begin{bmatrix} 
-  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
-  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
-\end{bmatrix}$$
-
-where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 3, approximates 
-the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
-by the rate given in Theorem 5 of \citeauthor{rubindelanchy2017statistical}.  
-In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
-result \cite{rubindelanchy2017statistical} is of the form 
-$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
-for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
-resulting in $W_* \stackrel{a.s.}{\to} I$. Therefore, the adjacency spectral
-embedding of $\hat{M}^{(kl)}$ is a direct estimation of the specific latent 
-positions outlined for $M^{(kl)}$, up to sign flip.
+\newpage
 

### 2021-04-03 summary.Rmd
wrote a short summary of work thus far + future work

@@
 
-\begin{definition}[Subspace Detection Property]. 
+\begin{definition}[Subspace Detection Property] 
 Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
@@
 
-Modularity Maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
-used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$. 
+Modularity Maximization is NP-hard, so
+\citeauthor{307cbeb9b1be48299388437423d94bf1} 
+used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$, 
+as a greedy relaxation of the optimization problem. 
 For these simulations, the EP algorithm was used for $K = 2$, and for $K > 2$, 

### 2021-04-03 functions.R

@@ -14,7 +14,7 @@ embedding <- function(A, p = NULL, q = NULL,
     S <- diag(sqrt(abs(eigen.A$values[keep])))
     return(U %*% S)
   } else {
-    return(U)
+    return(U * sqrt(n))
   }
 }
 
@@ -277,7 +277,6 @@ ssc <- function(A,
     B[is.nan(B)] <- 0
   }
   W <- B + t(B)
-  W <- sqrt(N) * W
   L <- normalized.laplacian(W)
   L.eigen <- eigen(L, symmetric = TRUE)
   X <- L.eigen$vectors[, seq(N, N - K)]

### 2021-04-03 short-summary.Rmd
fixed a couple of typos

@@
 $\{\lambda^{(kl)}\}_K$ and orthonormal matrix 
-$U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
+$U \in \mathbb{R}^{K^2 \times K^2}$ that is known 
 for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
@@
 $$\Lambda^{(k)} = 
-\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
+\begin{bmatrix} \lambda^{(k1)} & \cdots & \lambda^{(kK)} \end{bmatrix}
 \in \mathbb{R}^{n_k \times K}$$
@@
 \cite{rubindelanchy2017statistical}, it can be shown that constructing 
-$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}$
+$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}_n$
 corresponding to between-community pairings going to 0 with probability 1.
@@
   \SetAlgoLined
-  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
+  \caption{Sparse Subspace Clustering for the PABM using LASSO.}
   \KwData{Adjacency matrix $A$, number of communities $K$, 
@@
 More specifically, we showed that an embedding of the PABM results in 
-orthogonal subspaces and noise that goes to 0 with probability 1, which makes 
-for a very straightforward application of SSC. It can also be shown that this 
-type of embedding has the same properties of the SBM and DCBM, which are really
-just special cases of the PABM.
+orthogonal subspaces and noise that converges to 0 with probability 1, which 
+makes for a very straightforward application of SSC. It can also be shown that 
+this type of embedding has the same properties for the SBM, DCBM, and Nested 
+Block Model (NBM), which are really just special cases of the PABM
+(see \citet{noroozi2021hierarchy}, in particular Fig 2.).
 
@@
 than on a subspace. 
-Much of this work is based on work by \citet{trosset2020learning} who showed 
-convergence properties for graphs that result from latent positions that lie 
-on a one-dimensional manifold.
+Much of this work is based on work by \citet{trosset2020learning} who 
+considered the RDPG with latent positions that lie on a manifold and developed 
+methods for recovering the manifold from an adjacency matrix sampled from it.
 

### 2021-04-06 summary.Rmd
updated clustering simulations to have both EP and louvain

@@
   dplyr::group_by(n, K) %>%
-  dplyr::summarise(med.err = median(error),
+  dplyr::summarise(
+    med.err = median(error),
     first.q = quantile(error, .25),
@@
     third.q.ssc = quantile(error.ssc, .75),
-                   med.err.mm = median(error.mm),
-                   first.q.mm = quantile(error.mm, .25),
-                   third.q.mm = quantile(error.mm, .75)) %>%
+    med.err.ep = median(error.ep, na.rm = TRUE),
+    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
+    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
+    med.err.louvain = median(error.louvain),
+    first.q.louvain = quantile(error.louvain, .25),
+    third.q.louvain = quantile(error.louvain, .75)
+  ) %>% 
   dplyr::ungroup() %>% 
@@
                     colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm,
+  geom_line(aes(x = n, y = med.err.ep,
                 colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
+  geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
                     colour = 'MM-EP'), width = .1) + 
@@
                     colour = 'SSC-A'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.louvain,
+                colour = 'MM-Louvain')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
+                    colour = 'MM-Louvain'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
                     colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.mm * n,
+  geom_line(aes(x = n, y = med.err.ep * n,
                 colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
+  geom_errorbar(aes(x = n, ymin = first.q.ep * n, ymax = third.q.ep * n,
                     colour = 'MM-EP'), width = .1) + 
@@
                     colour = 'SSC-A'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.louvain * n,
+                colour = 'MM-Louvain')) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = first.q.louvain * n, ymax = third.q.louvain * n,
+                    colour = 'MM-Louvain'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 

### 2021-04-06 clustering-sim-k.R

@@ -84,16 +84,19 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z)
       if (K == 2) {
-        clustering.mm <- cluster.sg(A)
+        clustering.ep <- cluster.sg(A)
+        error.ep <- 1 - cluster.acc(clustering.ep, z)
       } else {
-        clustering.mm <- mod.max(A)
+        error.ep <- NA
       }
+      clustering.louvain <- mod.max(A)
+      error.louvain <- 1 - cluster.acc(clustering.louvain, z)
       error.mm <- 1 - cluster.acc(clustering.mm, z)
       dplyr::tibble(K = K, n = n,
                     error = error,
                     error.ssc = error.ssc,
-                    error.mm = error.mm) %>%
-        return()
+                    error.ep = error.ep,
+                    error.louvain = error.louvain) %>%
         return()
     } %>% 
       return()
@@ -113,9 +116,12 @@ clustering.df %>%
     med.err.ssc = median(error.ssc),
     first.q.ssc = quantile(error.ssc, .25),
     third.q.ssc = quantile(error.ssc, .75),
-    med.err.mm = median(error.mm),
-    first.q.mm = quantile(error.mm, .25),
-    third.q.mm = quantile(error.mm, .75)
+    med.err.ep = median(error.ep, na.rm = TRUE),
+    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
+    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
+    med.err.louvain = median(error.louvain),
+    first.q.louvain = quantile(error.louvain, .25),
+    third.q.louvain = quantile(error.louvain, .75)
   ) %>%
   ggplot() +
   scale_y_log10() +
@@ -126,9 +132,12 @@ clustering.df %>%
   geom_line(aes(x = n, y = med.err.ssc, colour = 'ssc')) +
   geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                     colour = 'ssc')) + 
-  geom_line(aes(x = n, y = med.err.mm, colour = 'mm')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
[…]

### 2021-04-06 summary.Rmd
typo

@@
 \label{theorem6}
-Under regularity and sparsity assumptions, and under the further 
-assumption that $K$ is fixed and community labels are known,
+Under regularity and sparsity assumptions, given fixed $K$,
 
@@
 
-```{r clust_err_k, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_k, fig.cap = 'IQR of community detection error error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_k_imba, fig.cap = 'IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.

### 2021-04-06 summary.Rmd
pmlr -> jmlr

@@
 \citeauthor{rubindelanchy2017statistical}, who describe the convergence 
-behavior of the ASE of $A$ to the ASE of $P$, and \citet{pmlr-v28-wang13},
+behavior of the ASE of $A$ to the ASE of $P$, and \citet{jmlr-v28-wang13},
 who show the necessary conditions for the subspace detection property to hold 
@@
   \SetAlgoLined
-  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
+  \caption{Sparse Subspace Clustering using LASSO \cite{jmlr-v28-wang13}.}
   \KwData{Adjacency matrix $A$, number of communities $K$, 
@@
 applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
-\citeauthor{soltanolkotabi2012}, and \citeauthor{pmlr-v28-wang13}, which lead 
+\citeauthor{soltanolkotabi2012}, and \citeauthor{jmlr-v28-wang13}, which lead 
 to asymptotically correct solutions with high probability. Similar methods can 
@@
 
-\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{pmlr-v28-wang13}]
+\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{jmlr-v28-wang13}]
 The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
@@
 Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
-extended definition of $\mu(\cdot)$ from \citeauthor{pmlr-v28-wang13}. Then
+extended definition of $\mu(\cdot)$ from \citeauthor{jmlr-v28-wang13}. Then
 
@@
 **Proof of Theorem \ref{theorem5}**. 
-The basis of this proof is Theorem 6 from \citeauthor{pmlr-v28-wang13}, which 
+The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which 
 states that the subspace detection property holds if the noise is small enough 
@@
 Thus the conditions for the subspace detection property from Theorem 6 from
-\citeauthor{pmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
+\citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
 

### 2021-04-13 summary.Rmd
updated draft per suggestions

@@
 - \usepackage{amsthm}
+- \usepackage{comment}
 bibliography: misc.bib
@@
 \begin{remark} 
-In practice, sample $X$ often does not obey the subspace detection
+In practice, a noisy sample $X$ often does not obey the subspace detection
 property. In such cases, $B$ is treated as an affinity matrix for a graph which 
@@
 but it could be the case that some of the subspaces are represented by 
-multiple disconnected subgraphs. 
+multiple disconnected subgraphs. The subspace detection property is contingent 
+on choosing a sufficiently large sparsity parameter $\lambda$. 
 \end{remark}
@@
 as a greedy relaxation of the optimization problem. 
-For these simulations, the EP algorithm was used for $K = 2$, and for $K > 2$, 
-the Louvain algorithm \cite{Blondel_2008} was used instead.
+For these simulations, the Louvain algorithm was used, as the EP algorithm 
+proved to be prohibitively computationally expensive for $K > 2$. For $K = 2$, 
+it was verified that EP and Louvain produce comparable results.
 
@@
 chosen via a preliminary cross-validation experiment. For the final clustering 
-step, a Gaussian Mixture Model was fit on the normalized Lapacian eigenmap of 
+step, a Gaussian Mixture Model was fit on the normalized Laplacian eigenmap of 
 the affinity matrix $B$.
+
+For comparing methods, we define the community detection error as:
+
+$$L_c(\hat{\sigma}, \sigma; \{v_i\}) = 
+\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma{v_i})$$
+
+where $\sigma(v_i)$ is the true community label of vertex $v_i$, 
+$\hat{\sigma}(v_i)$ is a community assignment function, and $\pi$ is a 
+permutation operator. This is effectively the "misclustering count" of
+$\hat{\sigma}$.
+
+We then define parameter estimation error as the MSE up to sign flip:
+
+$$MSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
+\min_{s \in \{-1, 1\}} 
+\sqrt{\frac{1}{n} \sum_{k < l} \sum_{i = 1}^{n_k} 
+(s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}$$
 
@@
 
+\begin{comment}
+
 Fig. \ref{fig:clust_err_k} show that for large $n$, OSC results in 
-a misclustering rate of 0. Sparse subspace clustering produces similar results 
-for $K > 2$.
+a misclustering rate of 0.
 
@@
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  # scale_y_log10() + 
   labs(y = 'community detection error rate', 
@@
                     colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ep,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
-                    colour = 'MM-EP'), width = .1) + 
+  # geom_line(aes(x = n, y = med.err.ep,
+  #               colour = 'MM-EP')) + 
+  # geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
+  #                   colour = 'MM-EP'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc2,
@@
   scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
+  facet_wrap(~ K, labeller = 'label_both')
 ```
@@
 
-```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+\end{comment}
+
+Fig \ref{fig:clust_err_ct_sim} shows OSC's community detection error going to 0 
+for large $n$. SSC on both the embedding and on the adjacency 
+matrix produces similar results for $K > 2$. Weaker performance of SSC for 
+$K = 2$ can be attributed to the final spectral clustering step of the 
+affinity matrix. A GMM was fit to the Laplacian eigenmap, but visual 
+inspection suggests that the communities are not distributed as a mixture 
+of Gaussians in the eigenmap. While the subspace detection property is 
+guaranteed for large $n$, in our simulations, setting a large enough 
+sparsity parameter for SSC resulted in more than $K$ disconnected subgraphs.
+
+```{r clust_err_ct_sim, fig.cap = 'Median and IQR of community detection error.
 Communities are approximately balanced.
@@
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  scale_y_log10() +
   labs(y = 'community detection error count', 
@@
                     colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ep * n,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ep * n, ymax = third.q.ep * n,
-                    colour = 'MM-EP'), width = .1) + 
+  # geom_line(aes(x = n, y = med.err.ep * n,
+  #               colour = 'MM-EP')) + 
+  # geom_errorbar(aes(x = n, ymin = first.q.ep * n, ymax = third.q.ep * n,
+  #                   colour = 'MM-EP'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc2 * n,
@@
 ```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be approximately balanced.', fig.width = 10}
 rmse.df <- readr::read_csv('rmse-k.csv')
@@
 
+\begin{comment}
+
 Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
@@
 
-```{r clust_err_k_imba, fig.cap = 'IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+```{r clust_err_k_imba, fig.cap = 'Median and IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
@@
 
-```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+\end{comment}
+
+We again see community detection error trending to 0 for OSC, as well as for 
+SSC when $K > 2$ (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3 continues to 
+see $n^{-1/2}$ decay in parameter estimation error 
+(\ref{fig:lambda_est_k_imba}).
+
+```{r clust_err_ct_sim_imba, fig.cap = 'Median and IQR of community detection error.
 Communities are imbalanced.
@@
 ```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be imbalanced.', fig.width = 10}
 rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')
@@
 
+\begin{comment}
+
 ## Additional experiments
@@
 each $n$ and constructed $B$ according to Algorithm 3 to verify that 
-as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
+as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{a.s.}{\to} 0$ for $i, j$ 
 in different clusters. Furthermore, the distribution of these inner products 
@@
 ```
+
+\end{comment}
 

### 2021-04-13 functions.R

@@ -256,6 +256,11 @@ ssc <- function(A,
   
   Y <- t(embedding(A, p, q, scale))
   
+  if (normalize) {
+    Y <- sweep(Y, 2, apply(Y, 2, function(y) sqrt(sum(y ** 2))), `/`)
+    Y[is.nan(Y)] <- 0
+  }
+  
   N <- ncol(Y)
   B <- plyr::aaply(seq(N), 1, function(i) {
     y <- Y[, i]
@@ -272,14 +277,14 @@ ssc <- function(A,
     return(betahat)
   }, .parallel = parallel) %>% 
     abs()
-  if (normalize) {
-    B <- sweep(B, 2, apply(B, 2, max), `/`)
-    B[is.nan(B)] <- 0
-  }
+  B <- sweep(B, 2, apply(B, 2, max), `/`)
+  B[is.nan(B)] <- 0
   W <- B + t(B)
   L <- normalized.laplacian(W)
   L.eigen <- eigen(L, symmetric = TRUE)
   X <- L.eigen$vectors[, seq(N, N - K)]
+  # X <- sweep(X, 2, apply(X, 2, mean), `-`)
+  # X <- sweep(X, 2, apply(X, 2, sd), `/`)
   clustering <- mclust::Mclust(X, K, verbose = FALSE)$classification
   return(clustering)
 }

### 2021-04-20 summary.Rmd
updated theorem 5

@@
 
+\begin{remark}
+The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6 
+from  \citeauthor{jmlr-v28-wang13} and the fact that the unscaled ASE of 
+$P_n$ consists of orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume 
+that the points in the embedding are all of unit length, and while we apply 
+this normalization in the simulated examples, it is not strictly necessary 
+for Theorem \ref{theorem5} due to orthogonality. 
+\end{remark}
+
 ## Parameter estimation
@@
 $$L_c(\hat{\sigma}, \sigma; \{v_i\}) = 
-\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma{v_i})$$
+\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))$$
 
 where $\sigma(v_i)$ is the true community label of vertex $v_i$, 
-$\hat{\sigma}(v_i)$ is a community assignment function, and $\pi$ is a 
+$\hat{\sigma}(v_i)$ is the predicted label of $v_i$, and $\pi$ is a 
 permutation operator. This is effectively the "misclustering count" of
-$\hat{\sigma}$.
-
-We then define parameter estimation error as the MSE up to sign flip:
-
-$$MSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
-\min_{s \in \{-1, 1\}} 
-\sqrt{\frac{1}{n} \sum_{k < l} \sum_{i = 1}^{n_k} 
+clustering function $\hat{\sigma}$.
+
+We also define parameter estimation error as the RMSE up to sign flip:
+
+$$RMSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
+\sqrt{\frac{1}{n} \sum_{k < l} \min_{s \in \{-1, 1\}} 
+\sum_{i = 1}^{n_k} 
 (s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}$$
@@
 that for each community $k$, there are enough vertices such that for each $i$, 
-$V_{n, -i}^{(k)}$ spans its corresponding subspace (Theorem 2). 
+$V_{n, -i}^{(k)}$ spans its corresponding subspace (Theorem \ref{theorem2}). 
 Then $r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0$. 
-Thus by (\ref{eq:mu-conv}), for each $k$, $r_n^{(k)} > \mu_n^{(k)}$ where 
+Thus by (\ref{eq:mu-conv}), for each $k$, $r_n^{(k)} > \mu_n^{(k)} = 0$ where 
 $\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})$ and $n$ is large enough such that 
-$\min\limits_{k, i} rank(V_{n, -i}^{(k)}) = K$.
-
-Let $r_n = \min_k r_n^{(k)}$. By (\ref{eq:delta-conv}), 
+$\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K$.
+
+Let $r_n = \min\limits_k r_n^{(k)}$. By (\ref{eq:delta-conv}), 
 $\delta_n \stackrel{a.s.}{\to} 0$. Then as $n\to \infty$, 
-$\delta_n < \min_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}$ 
+$\delta_n < \min\limits_k 
+\frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}$ 
+$= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}$
 with probability 1.
@@
 \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
+
+\begin{remark}
+Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $||v_n^{(i)}|| = 1$, 
+which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for 
+the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long 
+as the $k^{th}$ community spans its subspace, $r_n^{(k)} > 0 = \mu_n^{(k)}$.
+\end{remark}
 

### 2021-04-27 summary.Rmd
added subspace incoherence

@@
   pdf_document:
+    keep_tex: true
     citation_package: natbib
@@
 - \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
-- \setcitestyle{numbers,square}
+- \setcitestyle{numbers,square,comma}
 - \usepackage{verbatim}
@@
     and the $K (K - 1) / 2$ most negative eigenvalues.\;
+    Normalize $V \leftarrow \sqrt{n} V$.\;
     \For {$i = 1, ..., n$} {
@@
     $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
-    Compute the affinity matrix $B = \sqrt{n} (|C| + |C^\top|)$.\;
+    Compute the affinity matrix $B = |C| + |C^\top|$.\;
     Construct graph $G$ using $B$ as its similarity matrix.\;
@@
 
-\begin{definition}[Inradius \cite{soltanolkotabi2012} \cite{jmlr-v28-wang13}]
+\begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
 The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
@@
 
+\begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
+\end{definition}
+
 \begin{lemma}
 \label{lemma2}
-Let $P_n = V_n \Lambda_n V_n^\top$ be the edge probability matrix of a PABM 
-with $n$ vertices. Let $A_n \sim \Bernoulli(P_n)$ be the adjacency matrix 
-drawn from $P_n$, and let $A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be 
-its approximate spectral decomposition using the $K (K+1) / 2$ most positive 
-and $K (K-1) / 2$ most negative eigenvalues and their corresponding 
-eigenvectors.Without loss of generality, order the rows of $V_n$ and 
-$\hat{V}_n$ by community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the 
-$n_k \times K^2$ matrix of the rows of the $k^{th}$ community in $V_n$ and 
-$\hat{V}_n$ respectively. 
-Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
-as defined by \citeauthor{soltanolkotabi2012}. Then $\mu_n^{(k)} = 0$ 
-with probability 1 $\forall n, k$.
-
-Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
-extended definition of $\mu(\cdot)$ from \citeauthor{jmlr-v28-wang13}. Then
+Let $\hat{V}$ be the eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues such that the rows of $\hat{V}$ are ordered by community, and let $\hat{V}^{(k)}$ be the rows of the $k^{th}$ community in $\hat{V}$ and $\hat{V}^{(-k)}$ be the rows of $\hat{V}$ with the $k^{th}$ community omitted.
+Denote $(\hat{v}_i^{(k)})^\top$ as the rows of $\hat{V}$, $\hat{V}_{-i}^{(k)}$ as $\hat{V}^{(k)}$ with the $i^{th}$ row omitted, and $\mathcal{S}^{(k)}$ as the subspace spanned by $V^{(k)}$.
+Let $V$, $V^{(k)}$, $V^{(-k)}$, and $v_i^{(k)}$ be the corresponding values for $P_n$.
+
+Let $\nu_{i}^{(k)} = \max\limits_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2 \lambda} \eta^\top \eta$ subject to $||V_{-i}^{(k)} \eta||_\infty \leq 1$, and define the projected dual direction $w_{i}^{(k)}$ as $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$ normalized to length 1.
+Collect the projected dual directions into $W = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$.
+
+Define the subspace incoherence:
+
+$$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
+
+Then $\forall k, n$,
 
 \begin{equation} \label{eq:mu-conv}
-\hat{\mu}_n^{(k)} = 0
+\mu_n^{(k)} = 0
 \end{equation}
@@
 *Proof*. 
-For the first part, 
-since $v_n^{(k)}$ is orthogonal to $v_n^{(l)}$ $\forall k \neq l$ 
-(Lemma \ref{lemma1}), 
-this falls directly out of Theorem 2.8 of \citeauthor{soltanolkotabi2012}.
-
-For the second part, 
-$\mu(\hat{V}_n^{(k)}) = 
-\max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ 
-where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned 
-by $V_n^{(k)}$. Since 
-$u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ 
-$\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem \ref{theorem2}, this quantity 
-must be 0.
+Since by Theorem \ref{theorem2} each $\mathcal{S}^{(k)}$ are mutually orthogonal, any vector projected to 
+$\mathcal{S}^{(k)}$ must be orthogonal to each row of $V^{(-k)}$. Therefore, $W^{(k)} v = 0$ $\forall v \in \mathcal{S}^{(-k)}$.
 
@@
 the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long 
-as the $k^{th}$ community spans its subspace, $r_n^{(k)} > 0 = \mu_n^{(k)}$.
+as the $k^{th}$ community spans its subspace, $a r_n^{(k)} > 0 = \mu_n^{(k)}$ 
+$\forall a > 0$.
 \end{remark}
@@
 $P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
-signular value decomposition 
+singular value decomposition 
 $P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 

### 2021-04-27 summary.tex
forgot to add tex file

--- 2021-04-27 summary.Rmd
@@
----
-title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation"
-authors: John Koo, Minh Tang, Michael Trosset
-# author: John Koo
-# author2: Minh Tang
-# author3: Michael Trosset
-output:
-  pdf_document:
-    keep_tex: true
-    citation_package: natbib
-    number_sections: yes
-# output:
-#   bookdown::pdf_document2:
-#     citation_package: natbib
-#     number_citations: yes
-# output: html_document
-# output: rticles::rss_article
-fontsize: 12pt
-# geometry: "left=1in,right=1in,top=1in,bottom=1in"
-urlcolor: blue
-header-includes:
-- \usepackage{float}
-- \usepackage{mathtools}
-- \usepackage{natbib}
-- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
-- \setcitestyle{numbers,square,comma}
-- \usepackage{verbatim}
-- \usepackage{amsthm}
-- \usepackage{comment}
-bibliography: misc.bib
-abstract: |
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+%
+\documentclass[
+  12pt,
+]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation},
+  colorlinks=true,
+  linkcolor=Maroon,
+  filecolor=Maroon,
+  citecolor=Blue,
+  urlcolor=blue,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{natbib}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\setcitestyle{numbers,square,comma}
+\usepackage{verbatim}
+\usepackage{amsthm}
+\usepackage{comment}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
+\title{Connecting the Popularity Adjusted Block Model to the Generalized Random
+Dot Product Graph for Community Detection and Parameter Estimation}
+\author{}
+\date{\vspace{-2.5em}}
+
+\begin{document}
+\maketitle
+\begin{abstract}
 We connect two random graph models, the Popularity Adjusted Block Model
-  (PABM) and the Generalized Random Dot Product Graph (GRDPG), demonstrating
-  that a PABM is a GRDPG in which communities correspond to certain mutually
-  orthogonal subspaces of latent positions. This insight leads to the
-  construction of improved algorithms for community detection and parameter
-  estimation with PABM. Using established asymptotic properties of Adjacency
-  Spectral Embedding (ASE) for GRDPG, we derive asymptotic properties of these
-  algorithms, including algorithms that rely on Sparse Subspace Clustering
-  (SSC). We illustrate these properties via simulation.
----
-
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      # eval = FALSE,
-                      comment = NA, 
-                      warning = FALSE, 
-                      message = FALSE, 
-                      fig.pos = 'H', 
-                      fig.align = 'center', 
-                      fig.height = 3, 
-                      fig.width = 5, 
-                      fig.dpi = 300)
-
-options(xtable.comment = FALSE, 
-        xtable.table.placement = 'H')
-```
-
-```{r}
-import::from(magrittr, `%>%`)
-library(plot.matrix)
-library(ggplot2)
-library(mclust)
-
-theme_set(theme_bw())
-```
+(PABM) and the Generalized Random Dot Product Graph (GRDPG),
+demonstrating that a PABM is a GRDPG in which communities correspond to
+certain mutually orthogonal subspaces of latent positions. This insight
+leads to the construction of improved algorithms for community detection
+and parameter estimation with PABM. Using established asymptotic
+properties of Adjacency Spectral Embedding (ASE) for GRDPG, we derive
+asymptotic properties of these algorithms, including algorithms that
+rely on Sparse Subspace Clustering (SSC). We illustrate these properties
+via simulation.
+\end{abstract}
 
@@
 
-# Introduction
-
-Statistical analysis on graphs or networks often involves the partitioning of 
-a graph into disconnected subgraphs or clusters. This is often motivated by the 
-assumption that there exist underlying and unobserved communities to which each 
-vertex of the graph belongs, and edges between pairs of vertices are determined 
-by drawing from a probability distribution based on the community relationships 
-between each pair. The goal of the analysis then is population community 
-detection, or the recovery of the true underlying community labels for each 
-vertex, up to permutation (with some additional parameter estimation being of 
-possible interest), assuming some underlying probability model. One such model
-is the Stochastic Block Model (SBM), first proposed by 
-\citet{doi:10.1080/0022250X.1971.9989788}, which assumes that the edge 
-probability from one vertex to another follows a Beronulli distribution with 
-fixed probabilities for each pair of community labels. Other random 
-graph models have been proposed and studied, such as the Degree-Corrected Block 
-Model (DCBM), introduced by \citet{Karrer_2011}, which is a generalization of 
-the SBM. The Popularity Adjusted Block Model (PABM) was then introduced by 
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the DCBM to 
-address the heterogeneity of edge probabilities within and between communities 
-while still maintaining distinct community structure.
-
-The underlying similarity among the SBM, PABM, and other such models is that 
-they involve a symmetric edge probability matrix $P \in [0, 1]^{n \times n}$ 
-where $n$ is the number of vertices in the graph. An undirected and unweighted 
-graph is then drawn from this edge probability matrix such that the existence 
-of an edge between each pair of vertices $i$ and $j$ is given by 
-$\Bernoulli(P_{ij})$. For example, for the SBM with two communities for which 
-the within-community edge probability is $\xi$ and the between-community edge 
-probability is $\eta$, the entries of $P$ consist of $P_{ij} = \xi$ if $i$ and 
-$j$ are in the same community and $i \neq j$, $P_{ij} = \eta$ if $i$ and $j$ 
-belong to separate communities, and $P_{ii} = 0$.
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Statistical analysis on graphs or networks often involves the
+partitioning of a graph into disconnected subgraphs or clusters. This is
+often motivated by the assumption that there exist underlying and
+unobserved communities to which each vertex of the graph belongs, and
+edges between pairs of vertices are determined by drawing from a
+probability distribution based on the community relationships between
+each pair. The goal of the analysis then is population community
+detection, or the recovery of the true underlying community labels for
+each vertex, up to permutation (with some additional parameter
+estimation being of possible interest), assuming some underlying
+probability model. One such model is the Stochastic Block Model (SBM),
+first proposed by \citet{doi:10.1080/0022250X.1971.9989788}, which
+assumes that the edge probability from one vertex to another follows a
+Beronulli distribution with fixed probabilities for each pair of
+community labels. Other random graph models have been proposed and
+studied, such as the Degree-Corrected Block Model (DCBM), introduced by
+\citet{Karrer_2011}, which is a generalization of the SBM. The
+Popularity Adjusted Block Model (PABM) was then introduced by
+\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the
+DCBM to address the heterogeneity of edge probabilities within and
+between communities while still maintaining distinct community
+structure.
+
+The underlying similarity among the SBM, PABM, and other such models is
+that they involve a symmetric edge probability matrix
+\(P \in [0, 1]^{n \times n}\) where \(n\) is the number of vertices in
+the graph. An undirected and unweighted graph is then drawn from this
+edge probability matrix such that the existence of an edge between each
+pair of vertices \(i\) and \(j\) is given by
+\(\text{Bernoulli}(P_{ij})\). For example, for the SBM with two
+communities for which the within-community edge probability is \(\xi\)
+and the between-community edge probability is \(\eta\), the entries of
+\(P\) consist of \(P_{ij} = \xi\) if \(i\) and \(j\) are in the same
+community and \(i \neq j\), \(P_{ij} = \eta\) if \(i\) and \(j\) belong
+to separate communities, and \(P_{ii} = 0\).
 
 The Random Dot Product Graph (RDPG) model proposed by
-\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli 
-edge probabilities. Under this model, each vertex of the graph can be
-represented by a point in some latent space such that the edge
-probability between any pair of vertices is given by their corresponding dot
-product in the latent space, i.e., given a latent positions
-$x_1, ..., x_n \in \mathbb{R}^d$, the edge probability matrix is 
-$P = X X^\top$ where 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. The SBM is 
-equivalent to a special case of the RDPG model in which all vertices of a 
-given community share the same position in the latent space 
-\cite{lyzinski2014}. It has also been shown that similar 
-random graph models, including the DCBM, can be represented in this way
+\citet*{10.1007/978-3-540-77004-6_11} is another graph model with
+Bernoulli edge probabilities. Under this model, each vertex of the graph
+can be represented by a point in some latent space such that the edge
+probability between any pair of vertices is given by their corresponding
+dot product in the latent space, i.e., given a latent positions
+\(x_1, ..., x_n \in \mathbb{R}^d\), the edge probability matrix is
+\(P = X X^\top\) where
+\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The SBM
+is equivalent to a special case of the RDPG model in which all vertices
+of a given community share the same position in the latent space
+\cite{lyzinski2014}. It has also been shown that similar random graph
+models, including the DCBM, can be represented in this way
 \cite{lyzinski2014} \cite{rubindelanchy2017consistency}. An analogous
 property exists for the PABM but not for the RDPG model but under the
-*Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
-explored in this paper and exploited to construct algorithms for community 
-detection and parameter estimation for the PABM.
-
-In this paper, we will only consider undirected graphs, that is the edge weight 
-from vertex $i$ to vertex $j$ is equal to the edge weight in the opposite 
-direction, from vertex $j$ to vertex $i$. Furthermore, we will only consider 
-unweighted graphs with binary ($0$ or $1$) edge weights We will also assume that 
-graphs are hollow, i.e., there are no edges from a vertex to itself. All such 
-graphs can be represented by a symmetric adjacency matrix 
-$A \in \{0, 1\}^{n \times n}$ for which $A_{ij} = 1$ if there exists an edge 
-between vertices $i$ and $j$ and $0$ otherwise, and $A$ is an element-wise 
-independent Bernoulli draw from a symmetric edge probability matrix 
-$P \in [0, 1]^{n \times n}$.
-
-# Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
-
-## The popularity adjusted block model and the generalized random dot product graph
+\emph{Generalized} Random Dot Product Graph (GRDPG) model. This
+relationship will be explored in this paper and exploited to construct
+algorithms for community detection and parameter estimation for the
+PABM.
+
+In this paper, we will only consider undirected graphs, that is the edge
+weight from vertex \(i\) to vertex \(j\) is equal to the edge weight in
+the opposite direction, from vertex \(j\) to vertex \(i\). Furthermore,
+we will only consider unweighted graphs with binary (\(0\) or \(1\))
+edge weights We will also assume that graphs are hollow, i.e., there are
+no edges from a vertex to itself. All such graphs can be represented by
+a symmetric adjacency matrix \(A \in \{0, 1\}^{n \times n}\) for which
+\(A_{ij} = 1\) if there exists an edge between vertices \(i\) and \(j\)
+and \(0\) otherwise, and \(A\) is an element-wise independent Bernoulli
+draw from a symmetric edge probability matrix
+\(P \in [0, 1]^{n \times n}\).
+
+\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
+\section{Connecting the Popularity Adjusted Block Model to the
+Generalized Random Dot Product
+Graph}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
+
+\hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
+\subsection{The popularity adjusted block model and the generalized
+random dot product
+graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
 
@@
 that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
-realization of $\Bernoulli(P)$, i.e., 
-$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ 
+realization of $\text{Bernoulli}(P)$, i.e., 
+$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ 
 ($A_{ij} = A_{ji}$ and $A_{ii} = 0$).
@@
 
-We will use the notation $A \sim PABM(\{\lambda^{(kl)}\}_K)$ to denote a random 
-adjacency matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
-consisting of $K$ underlying communities.
+We will use the notation \(A \sim PABM(\{\lambda^{(kl)}\}_K)\) to denote
+a random adjacency matrix \(A\) drawn from a PABM with parameters
+\(\lambda^{(kl)}\) consisting of \(K\) underlying communities.
 
@@
 graph $G = (V, E)$ with adjacency matrix $A$ such that 
-$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
+$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
 $A_{ii} = 0$) is a draw from the generalized random dot product graph model 
@@
 
-We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random adjacency 
-matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.
+We will use the notation \(A \sim GRDPG_{p,q}(X)\) to denote a random
+adjacency matrix \(A\) drawn from latent positions \(X\) and signature
+\((p, q)\).
 
@@
 
-## Connecting the PABM to the GRDPG
+\hypertarget{connecting-the-pabm-to-the-grdpg}{%
+\subsection{Connecting the PABM to the
+GRDPG}\label{connecting-the-pabm-to-the-grdpg}}
 
@@
 \begin{equation} \label{eq:xy}
-X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
+X = \text{blockdiag}(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
 \end{equation}
 
-$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
+$$L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
 \mathbb{R}^{n \times K}$$
@@
 
-# Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 
-1. Community membership identification (up to permutation).
-2. Parameter estimation (estimating $\lambda^{(kl)}$'s).
-
-In our methods, we assume that $K$, the number of communities, is known 
-beforehand and does not require estimation.
-
-## Related work
-
-\citeauthor{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, 
-used Modularity Maximization (MM) and the Extreme Points
-(EP) algorithm \cite{le2016} for community detection and parameter estimation. 
-They were able to show that as the sample size increases, the proportion of
-misclassified community labels (up to permutation) goes to 0.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Community membership identification (up to permutation).
+\item
+  Parameter estimation (estimating \(\lambda^{(kl)}\)'s).
+\end{enumerate}
+
+In our methods, we assume that \(K\), the number of communities, is
+known beforehand and does not require estimation.
+
+\hypertarget{related-work}{%
+\subsection{Related work}\label{related-work}}
+
+\citeauthor{307cbeb9b1be48299388437423d94bf1}, who first proposed the
+PABM, used Modularity Maximization (MM) and the Extreme Points (EP)
+algorithm \cite{le2016} for community detection and parameter
+estimation. They were able to show that as the sample size increases,
+the proportion of misclassified community labels (up to permutation)
+goes to 0.
 
 \citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for
-community detection in the PABM. 
-SSC is performed by solving an optimization problem for each observed point.
-Given $X \in \mathbb{R}^{n \times d}$ with vectors 
-$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
-$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
-for each $i = 1, ..., n$. The solutions are collected collected into matrix 
-$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
-affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
-subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
-subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
-If $X$ instead represents points near $K$ subspaces with some noise, a final 
-graph partitioning step is performed (e.g., edge thresholding or spectral 
-clustering).
+community detection in the PABM. SSC is performed by solving an
+optimization problem for each observed point. Given
+\(X \in \mathbb{R}^{n \times d}\) with vectors
+\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
+\(\min_{c_i} ||c_i||_1\) subject to \(x_i = X c_i\) and \(\beta_i = 0\)
+is solved for each \(i = 1, ..., n\). The solutions are collected
+collected into matrix
+\(C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top\) to
+construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
+perfectly on one of \(K\) subspaces, \(B\) describes an undirected graph
+consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
+\(x_i, x_j\) are in different subspaces. If \(X\) instead represents
+points near \(K\) subspaces with some noise, a final graph partitioning
+step is performed (e.g., edge thresholding or spectral clustering).
 
@@
 
-for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
-into $C$ and $B$ as before.
+for some sparsity parameter \(\lambda > 0\). The \(c_i\) vectors are
+then collected into \(C\) and \(B\) as before.
 
@@
 Theorem \ref{theorem2} suggests that SSC is appropriate for community
-detection for the PABM. 
-More precisely, Theorem \ref{theorem2} says that each community consists of a
-$K$-dimensional subspace, and together the subspaces lie in $\mathbb{R}^{K^2}$. 
-The natural approach then is to perform SSC on the ASE of $P$ or $A$.
-\citeauthor{noroozi2019estimation} instead applied SSC to $P$ and $A$, 
-foregoing embedding altogether.
-
-Using results from \citet{soltanolkotabi2012}, it can be easily shown that 
-the subspace detection property holds for $XU$, which is an ASE of $P$. More 
-specifically, if points lie exactly on mutually orthogonal subspaces, then the 
-subspace detection property will hold with probability 1, and this is exactly 
-the case for the PABM (Theorem \ref{theorem2}). 
-Much of our work is then built on
+detection for the PABM. More precisely, Theorem \ref{theorem2} says that
+each community consists of a \(K\)-dimensional subspace, and together
+the subspaces lie in \(\mathbb{R}^{K^2}\). The natural approach then is
+to perform SSC on the ASE of \(P\) or \(A\).
+\citeauthor{noroozi2019estimation} instead applied SSC to \(P\) and
+\(A\), foregoing embedding altogether.
+
+Using results from \citet{soltanolkotabi2012}, it can be easily shown
+that the subspace detection property holds for \(XU\), which is an ASE
+of \(P\). More specifically, if points lie exactly on mutually
+orthogonal subspaces, then the subspace detection property will hold
+with probability 1, and this is exactly the case for the PABM (Theorem
+\ref{theorem2}). Much of our work is then built on
 \citeauthor{rubindelanchy2017statistical}, who describe the convergence
-behavior of the ASE of $A$ to the ASE of $P$, and \citet{jmlr-v28-wang13},
-who show the necessary conditions for the subspace detection property to hold 
-in noisy cases where the points lie near subspaces.
-
-## Community detection
+behavior of the ASE of \(A\) to the ASE of \(P\), and
+\citet{jmlr-v28-wang13}, who show the necessary conditions for the
+subspace detection property to hold in noisy cases where the points lie
+near subspaces.
+
+\hypertarget{community-detection}{%
+\subsection{Community detection}\label{community-detection}}
 
@@
 
-We previously stated one possible set of latent positions that result in the 
-edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
-(or can estimate) $XU$ directly, then both the community detection and 
-parameter identification problem are trivial since $U$ is orthonormal and 
-fixed for each value of $K$. However, direct identification or estimation of
-$XU$ is not possible \cite{rubindelanchy2017statistical}.
-
-If we decompose $P = Z I_{pq} Z^\top$, then
-$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
-the exact edge probability matrix, we cannot recover the "original" latent 
-positions $XU$. Note that unlike in the case of the RDPG, $Q$ is not 
-necessarily an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
-$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
-$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
-us from using the properties of $XU$ directly. In particular, if 
-$Q \in \mathbb{O}(n)$, then we could use the fact that 
-$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
-and $j$ are in different communities.
-
-The explicit form of $XU$ represents points in 
-$\mathbb{R}^{K^2}$ such that points within each community lie on 
-$K$-dimensional orthogonal subspaces. 
-Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
-but retains the property that each community is represented by a 
-$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
-correspond to each community, suggesting the use of SSC. In this paper, we will 
-use a different and leave the properties of SSC on the ASE of $A$ as future
-work.
+We previously stated one possible set of latent positions that result in
+the edge probability matrix of a PABM, \(P = (XU) I_{pq} (XU)^\top\). If
+we have (or can estimate) \(XU\) directly, then both the community
+detection and parameter identification problem are trivial since \(U\)
+is orthonormal and fixed for each value of \(K\). However, direct
+identification or estimation of \(XU\) is not possible
+\cite{rubindelanchy2017statistical}.
+
+If we decompose \(P = Z I_{pq} Z^\top\), then
+\(\exists Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
+start with the exact edge probability matrix, we cannot recover the
+``original'' latent positions \(XU\). Note that unlike in the case of
+the RDPG, \(Q\) is not necessarily an orthogonal matrix. If \(z_i\)'s
+are the rows of \(XU\), then
+\(||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2\), and
+\(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
+prevents us from using the properties of \(XU\) directly. In particular,
+if \(Q \in \mathbb{O}(n)\), then we could use the fact that
+\(\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0\) if
+vertices \(i\) and \(j\) are in different communities.
+
+The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
+such that points within each community lie on \(K\)-dimensional
+orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
+removes the orthogonality property but retains the property that each
+community is represented by a \(K\)-dimensional subspace. Therefore, the
+ASE of \(P\) results in subspaces that correspond to each community,
+suggesting the use of SSC. In this paper, we will use a different and
+leave the properties of SSC on the ASE of \(A\) as future work.
 
@@
 
-Theorem \ref{theorem3} provides perfect community detection given $P$. 
-Letting $|B|$ be the affinity matrix for graph $G$, $G$ is partitioned into 
-at least $K$ disjoint subgraphs since each of the $K$ communities have no 
-edges between them. Similar to the subspace detection property, it could be 
-the case that some of the communities are represented by multiple disjoint
-subgraphs in $G$, in which case additional reconstruction is required to
-identify the communities exactly.
-
-Using $A$ instead of $P$ introduces error, which converges to $0$ 
+Theorem \ref{theorem3} provides perfect community detection given \(P\).
+Letting \(|B|\) be the affinity matrix for graph \(G\), \(G\) is
+partitioned into at least \(K\) disjoint subgraphs since each of the
+\(K\) communities have no edges between them. Similar to the subspace
+detection property, it could be the case that some of the communities
+are represented by multiple disjoint subgraphs in \(G\), in which case
+additional reconstruction is required to identify the communities
+exactly.
+
+Using \(A\) instead of \(P\) introduces error, which converges to \(0\)
 almost surely:
@@
 
-Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} 
-also provide a very natural path toward using SSC for 
-community detection for the PABM. We established in Theorem \ref{theorem2}
-that an ASE of the edge probability matrix $P$ can be constructed 
-such that the communities lie on mutually orthogonal subspaces,
-and this property can be recovered from the eigenvectors of $P$. 
-Then Theorems \ref{theorem3} and \ref{theorem4} show that this property holds 
-for the unscaled ASE of $A$ drawn from $P$ as $n \to \infty$.
+Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} also provide
+a very natural path toward using SSC for community detection for the
+PABM. We established in Theorem \ref{theorem2} that an ASE of the edge
+probability matrix \(P\) can be constructed such that the communities
+lie on mutually orthogonal subspaces, and this property can be recovered
+from the eigenvectors of \(P\). Then Theorems \ref{theorem3} and
+\ref{theorem4} show that this property holds for the unscaled ASE of
+\(A\) drawn from \(P\) as \(n \to \infty\).
 
@@
 Let $P_n$ describe the edge probability matrix of the PABM with 
-$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{V}_n$ be the 
+$n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the 
 matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
@@
 
-## Parameter estimation
-
-For any edge probability matrix $P$ for the PABM such that the rows and 
-columns are organized by community, the $kl$^th^ block is an outer product 
-of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
-Therefore, given $P^{(kl)}$, $\lambda^{(kl)}$ and $\lambda^{(lk)}$ are 
-solvable exactly (up to multiplication by $-1$) using singular value 
-decomposition. More specifically, let $P^{(kl)} = \sigma^2 u v^\top$ be the 
-singular value decomposition of $P^{(kl)}$. $u \in \mathbb{R}^{n_k}$ and 
-$v \in \mathbb{R}^{n_l}$ are vectors and 
-$\sigma^2 > 0$ is a scalar. Then $\lambda^{(kl)} = \pm \sigma u$ and 
-$\lambda^{(lk)} = \pm \sigma v$. 
-Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
-simply use plug-in estimators (algorithm 3), which converge to the true 
-parameters.
+\hypertarget{parameter-estimation}{%
+\subsection{Parameter estimation}\label{parameter-estimation}}
+
+For any edge probability matrix \(P\) for the PABM such that the rows
+and columns are organized by community, the \(kl\)\textsuperscript{th}
+block is an outer product of two vectors, i.e.,
+\(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). Therefore, given
+\(P^{(kl)}\), \(\lambda^{(kl)}\) and \(\lambda^{(lk)}\) are solvable
+exactly (up to multiplication by \(-1\)) using singular value
+decomposition. More specifically, let \(P^{(kl)} = \sigma^2 u v^\top\)
+be the singular value decomposition of \(P^{(kl)}\).
+\(u \in \mathbb{R}^{n_k}\) and \(v \in \mathbb{R}^{n_l}\) are vectors
+and \(\sigma^2 > 0\) is a scalar. Then \(\lambda^{(kl)} = \pm \sigma u\)
+and \(\lambda^{(lk)} = \pm \sigma v\). Given the adjacency matrix \(A\)
+instead of edge probability matrix \(P\), we can simply use plug-in
+estimators (algorithm 3), which converge to the true parameters.
 
@@
 
-# Simulated Examples
+\hypertarget{simulated-examples}{%
+\section{Simulated Examples}\label{simulated-examples}}
 
 For each simulation, community labels are drawn from a multinomial
-distribution, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
-from two types of joint distributions depending on whether $k = l$, the edge
-probability matrix $P$ is constructed using 
-the popularity vectors, and finally an unweighted and undirected adjacency 
-matrix $A$ is drawn from $P$. OSC is then used for community detection, 
-and this method is compared against SSC \cite{noroozi2019estimation}
+distribution, the popularity vectors \(\{\lambda^{(kl)}\}_K\) are drawn
+from two types of joint distributions depending on whether \(k = l\),
+the edge probability matrix \(P\) is constructed using the popularity
+vectors, and finally an unweighted and undirected adjacency matrix \(A\)
+is drawn from \(P\). OSC is then used for community detection, and this
+method is compared against SSC \cite{noroozi2019estimation}
 \cite{soltanolkotabi2014} and MM \cite{igraph}
-\cite{307cbeb9b1be48299388437423d94bf1}. True 
-community labels are used with Algorithm 3 to estimate the popularity 
-vectors $\{\lambda^{(kl)}\}_K$, and this method is then compared against an
+\cite{307cbeb9b1be48299388437423d94bf1}. True community labels are used
+with Algorithm 3 to estimate the popularity vectors
+\(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
 MLE-based estimator described in \citeauthor{noroozi2019estimation} and
@@
 Modularity Maximization is NP-hard, so
-\citeauthor{307cbeb9b1be48299388437423d94bf1} 
-used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$, 
-as a greedy relaxation of the optimization problem. 
-For these simulations, the Louvain algorithm was used, as the EP algorithm 
-proved to be prohibitively computationally expensive for $K > 2$. For $K = 2$, 
-it was verified that EP and Louvain produce comparable results.
+\citeauthor{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
+(EP) algorithm \cite{le2016}, which is \(O(n^{K - 1})\), as a greedy
+relaxation of the optimization problem. For these simulations, the
+Louvain algorithm was used, as the EP algorithm proved to be
+prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
+was verified that EP and Louvain produce comparable results.
 
 Two implementations of SSC are shown here. The first method, denoted as
-SSC-A, treats the columns of the adjacency matrix $A$ as points in
-$\mathbb{R}^n$, as described in \citeauthor{noroozi2019estimation}. The second 
-method, denoted as SSC-ASE, first embeds $A$ and then performs SSC on the 
-embedding, as described in algorithm 2. The sparsity parameter $\lambda$ was 
-chosen via a preliminary cross-validation experiment. For the final clustering 
-step, a Gaussian Mixture Model was fit on the normalized Laplacian eigenmap of 
-the affinity matrix $B$.
+SSC-A, treats the columns of the adjacency matrix \(A\) as points in
+\(\mathbb{R}^n\), as described in \citeauthor{noroozi2019estimation}.
+The second method, denoted as SSC-ASE, first embeds \(A\) and then
+performs SSC on the embedding, as described in algorithm 2. The sparsity
+parameter \(\lambda\) was chosen via a preliminary cross-validation
+experiment. For the final clustering step, a Gaussian Mixture Model was
+fit on the normalized Laplacian eigenmap of the affinity matrix \(B\).
 
@@
 
-$$L_c(\hat{\sigma}, \sigma; \{v_i\}) = 
-\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))$$
-
-where $\sigma(v_i)$ is the true community label of vertex $v_i$, 
-$\hat{\sigma}(v_i)$ is the predicted label of $v_i$, and $\pi$ is a 
-permutation operator. This is effectively the "misclustering count" of
-clustering function $\hat{\sigma}$.
+\[L_c(\hat{\sigma}, \sigma; \{v_i\}) = 
+\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))\]
+
+where \(\sigma(v_i)\) is the true community label of vertex \(v_i\),
+\(\hat{\sigma}(v_i)\) is the predicted label of \(v_i\), and \(\pi\) is
+a permutation operator. This is effectively the ``misclustering count''
+of clustering function \(\hat{\sigma}\).
 
@@
 
-$$RMSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
+\[RMSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
 \sqrt{\frac{1}{n} \sum_{k < l} \min_{s \in \{-1, 1\}} 
 \sum_{i = 1}^{n_k} 
-(s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}$$
-
-## Balanced communities
-
-In each simulation, community labels $z_1, ..., z_n$ were drawn 
-from a multinomial distribution with mixture parameters 
-$\{\alpha_1, ..., \alpha_K\}$, then $\{\lambda^{(kl)}\}_K$ according to the 
-drawn community labels, $P$ was constructed using the drawn 
-$\{\lambda^{(kl)}\}_K$, and $A$ was drawn from $P$ by 
-$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$. Each simulation has a unique
-edge probability matrix $P$.
+(s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}\]
+
+\hypertarget{balanced-communities}{%
+\subsection{Balanced communities}\label{balanced-communities}}
+
+In each simulation, community labels \(z_1, ..., z_n\) were drawn from a
+multinomial distribution with mixture parameters
+\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(kl)}\}_K\) according
+to the drawn community labels, \(P\) was constructed using the drawn
+\(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
+\(A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})\). Each simulation has
+a unique edge probability matrix \(P\).
 
@@
 
-* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
-* Number of underlying communities $K = 2, 3, 4$
-* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
-community label has an equal probability of being drawn)
-* Community labels 
-$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
-* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
-* Between-group popularities 
-$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$
-
-$50$ simulations were performed for each $(n, K)$ pair.
+\begin{itemize}
+\tightlist
+\item
+  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
+\item
+  Number of underlying communities \(K = 2, 3, 4\)
+\item
+  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\), (i.e.,
+  each community label has an equal probability of being drawn)
+\item
+  Community labels
+  \(z_k \stackrel{\text{iid}}{\sim}Multinomial(\alpha_1, ..., \alpha_K)\)
+\item
+  Within-group popularities
+  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim}Beta(2, 1)\)
+\item
+  Between-group popularities
+  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim}Beta(1, 2)\) for
+  \(k \neq l\)
+\end{itemize}
+
+\(50\) simulations were performed for each \((n, K)\) pair.
 
@@
 
-```{r clust_err_k, fig.cap = 'IQR of community detection error error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_k-1}
+
+}
+
+\caption{IQR of community detection error error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(
-    med.err = median(error),
-    first.q = quantile(error, .25),
-    third.q = quantile(error, .75),
-    med.err.ssc = median(error.ssc),
-    first.q.ssc = quantile(error.ssc, .25),
-    third.q.ssc = quantile(error.ssc, .75),
-    med.err.ep = median(error.ep, na.rm = TRUE),
-    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
-    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
-    med.err.louvain = median(error.louvain),
-    first.q.louvain = quantile(error.louvain, .25),
-    third.q.louvain = quantile(error.louvain, .75)
-  ) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  # scale_y_log10() + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC-ASE'), width = .1) + 
-  # geom_line(aes(x = n, y = med.err.ep,
-  #               colour = 'MM-EP')) + 
-  # geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
-  #                   colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
-                    colour = 'SSC-A'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.louvain,
-                colour = 'MM-Louvain')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
-                    colour = 'MM-Louvain'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both')
-```
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k}
+\end{figure}
 
@@
 
-Fig \ref{fig:clust_err_ct_sim} shows OSC's community detection error going to 0 
-for large $n$. SSC on both the embedding and on the adjacency 
-matrix produces similar results for $K > 2$. Weaker performance of SSC for 
-$K = 2$ can be attributed to the final spectral clustering step of the 
-affinity matrix. A GMM was fit to the Laplacian eigenmap, but visual 
-inspection suggests that the communities are not distributed as a mixture 
-of Gaussians in the eigenmap. While the subspace detection property is 
-guaranteed for large $n$, in our simulations, setting a large enough 
-sparsity parameter for SSC resulted in more than $K$ disconnected subgraphs.
-
-```{r clust_err_ct_sim, fig.cap = 'Median and IQR of community detection error.
-Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(
-    med.err = median(error),
-    first.q = quantile(error, .25),
-    third.q = quantile(error, .75),
-    med.err.ssc = median(error.ssc),
-    first.q.ssc = quantile(error.ssc, .25),
-    third.q.ssc = quantile(error.ssc, .75),
-    med.err.ep = median(error.ep, na.rm = TRUE),
-    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
-    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
-    med.err.louvain = median(error.louvain),
-    first.q.louvain = quantile(error.louvain, .25),
-    third.q.louvain = quantile(error.louvain, .75)
-  ) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  scale_y_log10() +
-  labs(y = 'community detection error count', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) + 
-  # geom_line(aes(x = n, y = med.err.ep * n,
-  #               colour = 'MM-EP')) + 
-  # geom_errorbar(aes(x = n, ymin = first.q.ep * n, ymax = third.q.ep * n,
-  #                   colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2 * n,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
-                    colour = 'SSC-A'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.louvain * n,
-                colour = 'MM-Louvain')) + 
-  geom_errorbar(aes(x = n, 
-                    ymin = first.q.louvain * n, ymax = third.q.louvain * n,
-                    colour = 'MM-Louvain'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both')
-```
-
-Given ground truth community labels, Algorithm 3 and the MLE-based plug-in 
-estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
-perform similarly, with root mean square error decaying at rate approximately 
-$n^{-1/2}$.
-
-```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+Fig \ref{fig:clust_err_ct_sim} shows OSC's community detection error
+going to 0 for large \(n\). SSC on both the embedding and on the
+adjacency matrix produces similar results for \(K > 2\). Weaker
+performance of SSC for \(K = 2\) can be attributed to the final spectral
+clustering step of the affinity matrix. A GMM was fit to the Laplacian
+eigenmap, but visual inspection suggests that the communities are not
+distributed as a mixture of Gaussians in the eigenmap. While the
+subspace detection property is guaranteed for large \(n\), in our
+simulations, setting a large enough sparsity parameter for SSC resulted
+in more than \(K\) disconnected subgraphs.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim-1}
+
+}
+
+\caption{Median and IQR of community detection error. Communities are approximately balanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim}
+\end{figure}
+
+Given ground truth community labels, Algorithm 3 and the MLE-based
+plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1}
+\cite{noroozi2019estimation} perform similarly, with root mean square
+error decaying at rate approximately \(n^{-1/2}\).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_k-1}
+
+}
+
+\caption{Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.', fig.width = 10}
-rmse.df <- readr::read_csv('rmse-k.csv')
-
-rmse.df %>% 
-  na.omit() %>% 
-  dplyr::group_by(K, n) %>% 
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-  dplyr::ungroup() %>% 
-  ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 3'), width = .1) + 
-  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  labs(y = 'RMSE', colour = NULL) + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10() + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
-```
-
-## Imbalanced communities
-
-Simulations performed in this section are similar to those in the previous 
-section with the exception of the mixture parameters 
-$\{\alpha_1, ..., \alpha_K\}$ used to draw community labels from the 
-multinomial distribution. For these examples, we set the following parameters:
-
-* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
-* Number of underlying communities $K = 2, 3, 4$
-* Mixture parameters $\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}$ for 
-$k = 1, ..., K$
-* Community labels 
-$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
-* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
-* Between-group popularities 
-$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$
-
-$50$ simulations were performed for each $(n, K)$ pair.
+Communities were drawn to be approximately balanced.}\label{fig:lambda_est_k}
+\end{figure}
+
+\hypertarget{imbalanced-communities}{%
+\subsection{Imbalanced communities}\label{imbalanced-communities}}
+
+Simulations performed in this section are similar to those in the
+previous section with the exception of the mixture parameters
+\(\{\alpha_1, ..., \alpha_K\}\) used to draw community labels from the
+multinomial distribution. For these examples, we set the following
+parameters:
+
+\begin{itemize}
+\tightlist
+\item
+  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
+\item
+  Number of underlying communities \(K = 2, 3, 4\)
+\item
+  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}\)
+  for \(k = 1, ..., K\)
+\item
+  Community labels
+  \(z_k \stackrel{\text{iid}}{\sim}Multinomial(\alpha_1, ..., \alpha_K)\)
+\item
+  Within-group popularities
+  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim}Beta(2, 1)\)
+\item
+  Between-group popularities
+  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim}Beta(1, 2)\) for
+  \(k \neq l\)
+\end{itemize}
+
+\(50\) simulations were performed for each \((n, K)\) pair.
 
@@
 
-```{r clust_err_k_imba, fig.cap = 'Median and IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_k_imba-1}
+
+}
+
+\caption{Median and IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
 Communities are imbalanced.
-Simulations were repeated 50 times for each sample size.', fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(
-    med.err = median(error),
-    first.q = quantile(error, .25),
-    third.q = quantile(error, .75),
-    med.err.ssc = median(error.ssc),
-    first.q.ssc = quantile(error.ssc, .25),
-    third.q.ssc = quantile(error.ssc, .75),
-    med.err.ep = median(error.ep, na.rm = TRUE),
-    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
-    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
-    med.err.louvain = median(error.louvain),
-    first.q.louvain = quantile(error.louvain, .25),
-    third.q.louvain = quantile(error.louvain, .75)
-  ) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  labs(y = 'community detection error rate', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
-                    colour = 'SSC-ASE'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ep,
-                colour = 'MM-EP')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
-                    colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
-                    colour = 'SSC-A'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.louvain,
-                colour = 'MM-Louvain')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
-                    colour = 'MM-Louvain'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10()
-```
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k_imba}
+\end{figure}
 
@@
 
-We again see community detection error trending to 0 for OSC, as well as for 
-SSC when $K > 2$ (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3 continues to 
-see $n^{-1/2}$ decay in parameter estimation error 
+We again see community detection error trending to 0 for OSC, as well as
+for SSC when \(K > 2\) (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3
+continues to see \(n^{-1/2}\) decay in parameter estimation error
 (\ref{fig:lambda_est_k_imba}).
 
-```{r clust_err_ct_sim_imba, fig.cap = 'Median and IQR of community detection error.
-Communities are imbalanced.
-Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
-clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
-ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(
-    med.err = median(error),
-    first.q = quantile(error, .25),
-    third.q = quantile(error, .75),
-    med.err.ssc = median(error.ssc),
-    first.q.ssc = quantile(error.ssc, .25),
-    third.q.ssc = quantile(error.ssc, .75),
-    med.err.ep = median(error.ep, na.rm = TRUE),
-    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
-    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
-    med.err.louvain = median(error.louvain),
-    first.q.louvain = quantile(error.louvain, .25),
-    third.q.louvain = quantile(error.louvain, .75)
-  ) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
-                       first.q.ssc2 = quantile(error.ssc2, .25),
-                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  scale_y_log10() +
-  labs(y = 'community detection error count', 
-       colour = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) + 
-  # geom_line(aes(x = n, y = med.err.ep * n,
-  #               colour = 'MM-EP')) + 
-  # geom_errorbar(aes(x = n, ymin = first.q.ep * n, ymax = third.q.ep * n,
-  #                   colour = 'MM-EP'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc2 * n,
-                colour = 'SSC-A')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
-                    colour = 'SSC-A'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.louvain * n,
-                colour = 'MM-Louvain')) + 
-  geom_errorbar(aes(x = n, 
-                    ymin = first.q.louvain * n, ymax = third.q.louvain * n,
-                    colour = 'MM-Louvain'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  facet_wrap(~ K, labeller = 'label_both')
-```
-
-```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim_imba-1}
+
+}
+
+\caption{Median and IQR of community detection error. Communities are imbalanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim_imba}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_k_imba-1}
+
+}
+
+\caption{Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.
-Communities were drawn to be imbalanced.', fig.width = 10}
-rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')
-
-rmse.df %>% 
-  na.omit() %>% 
-  dplyr::group_by(K, n) %>% 
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
-  dplyr::ungroup() %>% 
-  ggplot() +
-  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
-                    colour = 'Algorithm 3'), width = .1) + 
-  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
-  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
-                   colour = 'MLE-based'), width = .1) + 
-  scale_colour_brewer(palette = 'Set1') + 
-  labs(y = 'RMSE', colour = NULL) + 
-  facet_wrap(~ K, labeller = 'label_both') + 
-  scale_y_log10() + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
-```
+Communities were drawn to be imbalanced.}\label{fig:lambda_est_k_imba}
+\end{figure}
 
@@
 
-```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', fig.height = 2, fig.width = 4}
-densities.df <- readr::read_csv('densities.csv')
-
-ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
-  labs(x = expression(n * v[i]^T * v[j]),
-       colour = 'n') + 
-  scale_x_continuous(breaks = seq(-5, 5)) + 
-  scale_colour_brewer(palette = 'Set1')
-```
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-2-1}
+
+}
+
+\caption{Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 
-# Real data examples
-
-```{r}
-import::here(cluster.pabm, ssc, plot.A, .from = '~/dev/pabm-grdpg/functions.R')
-
-n.edges <- 20566
-
-butterfly <- R.matlab::readMat('data/Raw_butterfly_network.mat')
-
-A <- butterfly$W.butterfly0
-z <- butterfly$labels
-
-keep <- c(6, 2, 9, 4)
-A <- A[z %in% keep, z %in% keep]
-z <- z[z %in% keep]
-n <- length(z)
-K <- 4
-
-delta <- quantile(A[upper.tri(A)], 1 - n.edges / (n * (n - 1) / 2))
-A <- (A > delta) * 1
-
-z.hat <- cluster.pabm(A, K)
-
-# z.hat.mm <- A %>% 
-#   igraph::graph_from_adjacency_matrix(mode = 'undirected') %>% 
-#   igraph::cluster_louvain() %>% 
-#   igraph::membership()
-
-# ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
-```
+\hypertarget{real-data-examples}{%
+\section{Real data examples}\label{real-data-examples}}
 
 In the first real data example, we applied OSC to the Leeds Butterfly
-dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
-butterflies across 10 species. The graph was modified to match the example 
-from \citeauthor{noroozi2019estimation}: Only the 4 most frequent species were 
-considered, and the similarities were discretized to $\{0, 1\}$ via 
-thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
-by the resultant clustering.
+dataset \cite{Wang_2018} consisting of visual similarity measurements
+among 832 butterflies across 10 species. The graph was modified to match
+the example from \citeauthor{noroozi2019estimation}: Only the 4 most
+frequent species were considered, and the similarities were discretized
+to \(\{0, 1\}\) via thresholding. Fig. \ref{fig:butterfly} shows a
+sorted adjacency matrix sorted by the resultant clustering.
 
@@
 accuracy of 63\% and an adjusted Rand index of 73\%. In comparison,
-\citeauthor{noroozi2019estimation} achieved an adjusted Rand index of 73\% 
-using sparse subspace clustering on the same dataset.
-
-```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.', cache = FALSE}
-plot.A(A, z.hat)
-```
-
-```{r}
-# this portion is from the code provided by Sengupta and Chen
-# the data were generated by scripts from S&C, then saved as RDS files
-
-bmp <- readRDS('data/bmp.rds')
-A.bmp <- bmp$A
-z.bmp <- bmp$z
-
-pb <- readRDS('data/pb.rds')
-A.pb <- pb$A
-z.pb <- pb$z
-
-dblp <- readRDS('data/dblp.rds')
-A.dblp <- dblp$A
-z.dblp <- dblp$z
-```
-
-```{r, cache = TRUE}
-zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = FALSE)
-bmp.acc <- mean(zhat.bmp == z.bmp)
-bmp.acc <- ifelse(bmp.acc < .5, 1 - bmp.acc, bmp.acc)
-
-zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = FALSE, 
-                        p = 2, q = 0, d.eigenmap = 2)
-pb.acc <- mean(zhat.pb == z.pb)
-pb.acc <- ifelse(pb.acc < .5, 1 - pb.acc, pb.acc)
-
-zhat.dblp <- cluster.pabm(A.dblp, 2, use.all = TRUE, normalize = FALSE)
-dblp.acc <- mean(zhat.dblp == z.dblp)
-dblp.acc <- ifelse(dblp.acc < .5, 1 - dblp.acc, dblp.acc)
-
-zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3, normalize = TRUE)
-bmp.acc.ssc <- mean(zhat.bmp.ssc == z.bmp)
-bmp.acc.ssc <- ifelse(bmp.acc.ssc < .5, 1 - bmp.acc.ssc, bmp.acc.ssc)
-
-zhat.pb.ssc <- ssc(A.pb, 2, .008, normalize = TRUE)
-pb.acc.ssc <- mean(zhat.pb.ssc == z.pb)
-pb.acc.ssc <- ifelse(pb.acc.ssc < .5, 1 - pb.acc.ssc, pb.acc.ssc)
-
-zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2, normalize = FALSE)
-dblp.acc.ssc <- mean(zhat.dblp.ssc == z.dblp)
-dblp.acc.ssc <- ifelse(dblp.acc.ssc < .5, 1 - dblp.acc.ssc, dblp.acc.ssc)
-
-```
-
-In the second example, we applied OSC to the British MPs Twitter 
-network \cite{greene2013producing}, the Political Blogs network 
-\cite{10.1145/1134271.1134277}, and the DBLP network \cite{NIPS2009_3855} 
-\cite{10.1007/978-3-642-15880-3_42}. For this data analysis, we subsetted the 
-data as described by 
-\citeauthor{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
-networks. Our methods underperformed compared to modularity maximization, 
-although performance is comparable. In 
-addition, OSC's runtime is much lower than that of modularity 
-maximization.
-
-```{r}
-dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
-              # Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
-              `MM` = c(round(1 / 329, 3),
-                       round(61 / 1222, 3),
-                       round(62 / 2203, 3)),
-              `SSC-ASE` = round(1 - c(bmp.acc.ssc,
-                                      pb.acc.ssc,
-                                      dblp.acc.ssc), 3),
-              `OSC` = round(1 - c(bmp.acc, 
-                                  pb.acc,
-                                  dblp.acc), 3)) %>% 
-  knitr::kable(caption = paste('Community detection error rates', 
-                               'for modularity maximization,',
-                               'sparse subspace clustering, ',
-                               'and OSC.'),
-               format = 'latex')
-```
-
-```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.', fig.dpi = 10}
-gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
-                        plot.A(A.pb, zhat.pb),
-                        plot.A(A.dblp, zhat.dblp),
-                        ncol = 3)
-```
+\citeauthor{noroozi2019estimation} achieved an adjusted Rand index of
+73\% using sparse subspace clustering on the same dataset.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/butterfly-1}
+
+}
+
+\caption{Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
+\end{figure}
+
+In the second example, we applied OSC to the British MPs Twitter network
+\cite{greene2013producing}, the Political Blogs network
+\cite{10.1145/1134271.1134277}, and the DBLP network
+\cite{NIPS2009_3855} \cite{10.1007/978-3-642-15880-3_42}. For this data
+analysis, we subsetted the data as described by
+\citeauthor{307cbeb9b1be48299388437423d94bf1} for their analysis of the
+same networks. Our methods underperformed compared to modularity
+maximization, although performance is comparable. In addition, OSC's
+runtime is much lower than that of modularity maximization.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-6}Community detection error rates for modularity maximization, sparse subspace clustering,  and OSC.}
+\centering
+\begin{tabular}[t]{l|r|r|r}
+\hline
+Network & MM & SSC-ASE & OSC\\
+\hline
+British MPs & 0.003 & 0.018 & 0.009\\
+\hline
+Political blogs & 0.050 & 0.196 & 0.062\\
+\hline
+DBLP & 0.028 & 0.087 & 0.059\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/mp-1}
+
+}
+
+\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
+\end{figure}
 
 In the third example, we consider the Karantaka villages data studied by
-\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` networks from 
-villages 12, 31, and 46 at the household level. The label of interest is the 
-religious affiliation. The networks were truncated to religions "1" and "2", 
-and vertices of degree 0 were removed.
-
-```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
-The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
-K <- 2
-hh.12 <- readRDS('data/village-hh-12.rds')
-hh.31 <- readRDS('data/village-hh-31.rds')
-hh.46 <- readRDS('data/village-hh-46.rds')
-
-# this part of code taken from sengupta & chen
-setwd('~/dev/pabm-grdpg/codes_and_data')
-require(MASS);require("clue")
-source("EPalgosim.R") 
-source("functions.R")
-mm <- function(A.hh) {
-  b.can = EPalgo(A.hh,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A.hh, b=b.can[,i])   # fit PABM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  return(b.can[,foo1])   # community assignment that maximises Q.PA
+\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
+networks from villages 12, 31, and 46 at the household level. The label
+of interest is the religious affiliation. The networks were truncated to
+religions ``1'' and ``2'', and vertices of degree 0 were removed.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-7-1}
+
 }
 
-A.hh.12 <- hh.12$A
-z.hh.12 <- hh.12$z
-zhat.hh.12 <- cluster.pabm(A.hh.12, K,
-                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
-zhat.hh.ssc.12 <- ssc(A.hh.12, K, lambda = .1)
-zhat.hh.mm.12 <- mm(A.hh.12)
-
-A.hh.31 <- hh.31$A
-z.hh.31 <- hh.31$z
-zhat.hh.31 <- cluster.pabm(A.hh.31, K,
-                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
-zhat.hh.ssc.31 <- ssc(A.hh.31, K, lambda = .05)
-zhat.hh.mm.31 <- mm(A.hh.31)
-
-A.hh.46 <- hh.46$A
-z.hh.46 <- hh.46$z
-zhat.hh.46 <- cluster.pabm(A.hh.46, K,
-                           use.all = TRUE, normalize = FALSE, p = 2, q = 0)
-zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .001)
-zhat.hh.mm.46 <- mm(A.hh.46)
-
-gridExtra::grid.arrange(plot.A(A.hh.12, zhat.hh.12),
-                        plot.A(A.hh.31, zhat.hh.31),
-                        plot.A(A.hh.46, zhat.hh.46),
-                        nrow = 1)
-```
-
-```{r, eval = TRUE}
-hh.acc.12 <- mean(zhat.hh.12 == z.hh.12)
-hh.acc.12 <- ifelse(hh.acc.12 < .5, 1 - hh.acc.12, hh.acc.12)
-hh.acc.ssc.12 <- mean(zhat.hh.ssc.12 == z.hh.12)
-hh.acc.ssc.12 <- ifelse(hh.acc.ssc.12 < .5, 1 - hh.acc.ssc.12, hh.acc.ssc.12)
-hh.acc.mm.12 <- mean(zhat.hh.mm.12 == z.hh.12)
-hh.acc.mm.12 <- ifelse(hh.acc.mm.12 < .5, 1 - hh.acc.mm.12, hh.acc.mm.12)
-
-hh.acc.31 <- mean(zhat.hh.31 == z.hh.31)
-hh.acc.31 <- ifelse(hh.acc.31 < .5, 1 - hh.acc.31, hh.acc.31)
-hh.acc.ssc.31 <- mean(zhat.hh.ssc.31 == z.hh.31)
-hh.acc.ssc.31 <- ifelse(hh.acc.ssc.31 < .5, 1 - hh.acc.ssc.31, hh.acc.ssc.31)
-hh.acc.mm.31 <- mean(zhat.hh.mm.31 == z.hh.31)
-hh.acc.mm.31 <- ifelse(hh.acc.mm.31 < .5, 1 - hh.acc.mm.31, hh.acc.mm.31)
-
-hh.acc.46 <- mean(zhat.hh.46 == z.hh.46)
-hh.acc.46 <- ifelse(hh.acc.46 < .5, 1 - hh.acc.46, hh.acc.46)
-hh.acc.ssc.46 <- mean(zhat.hh.ssc.46 == z.hh.46)
-hh.acc.ssc.46 <- ifelse(hh.acc.ssc.46 < .5, 1 - hh.acc.ssc.46, hh.acc.ssc.46)
-hh.acc.mm.46 <- mean(zhat.hh.mm.46 == z.hh.46)
-hh.acc.mm.46 <- ifelse(hh.acc.mm.46 < .5, 1 - hh.acc.mm.46, hh.acc.mm.46)
-
-# knitr::kable(dplyr::tibble(
-#   Method = c('OSC', 'SSC-ASE', 'Mod. Max.'),
-#   `Village 12` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
-#   `Village 31` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
-#   `Village 46` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
-#   caption = paste('Community detection error rates for identifying',
-#                   'household religion.'))
-
-knitr::kable(dplyr::tibble(
-  Network = c('Village 12', 'Village 31', 'Village 46'),
-  MM = round(1 - c(hh.acc.mm.12, hh.acc.mm.31, hh.acc.mm.46), 3),
-  `SSC-ASE` = round(1 - c(hh.acc.ssc.12, hh.acc.ssc.31, hh.acc.ssc.46), 3),
-  `OSC` = round(1 - c(hh.acc.12, hh.acc.31, hh.acc.46), 3)
-),
-caption = paste('Community detection error rates for identifying',
-                  'household religion.'),
-format = 'latex')
-```
-
-# Discussion
-
-This paper shows the connection between the PABM and the GRDPG, namely that 
-a PABM graph can be represented as a union of orthogonal subspaces in an 
-embedding under the GRDPG framework. We then exploited this relationship to 
-develop community detection and parameter estimation methods. In fact, we can 
-represent any graph with Bernoulli edges as a GRDPG, and in the PABM case, 
-it turns out that this relationship leads to a very straightforward 
-applications of previous work from \citeauthor{rubindelanchy2017statistical}, 
-\citeauthor{soltanolkotabi2012}, and \citeauthor{jmlr-v28-wang13}, which lead 
-to asymptotically correct solutions with high probability. Similar methods can 
-be applied for other models, such as the Nested Block Model
+\caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
+The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-8}Community detection error rates for identifying household religion.}
+\centering
+\begin{tabular}[t]{l|r|r|r}
+\hline
+Network & MM & SSC-ASE & OSC\\
+\hline
+Village 12 & 0.270 & 0.291 & 0.227\\
+\hline
+Village 31 & 0.125 & 0.066 & 0.110\\
+\hline
+Village 46 & 0.052 & 0.463 & 0.078\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+This paper shows the connection between the PABM and the GRDPG, namely
+that a PABM graph can be represented as a union of orthogonal subspaces
+in an embedding under the GRDPG framework. We then exploited this
+relationship to develop community detection and parameter estimation
+methods. In fact, we can represent any graph with Bernoulli edges as a
+GRDPG, and in the PABM case, it turns out that this relationship leads
+to a very straightforward applications of previous work from
+\citeauthor{rubindelanchy2017statistical},
+\citeauthor{soltanolkotabi2012}, and \citeauthor{jmlr-v28-wang13}, which
+lead to asymptotically correct solutions with high probability. Similar
+methods can be applied for other models, such as the Nested Block Model
 \cite{noroozi2021hierarchy}.
 
-# Proofs
-
-**Proof of Theorem \ref{theorem1}**. 
-This is given by straightforward matrix multiplication. It suffices to show 
-that
-
-$$X U I_{3, 1} U^\top X^\top = 
+\hypertarget{proofs}{%
+\section{Proofs}\label{proofs}}
+
+\textbf{Proof of Theorem \ref{theorem1}}. This is given by
+straightforward matrix multiplication. It suffices to show that
+
+\[X U I_{3, 1} U^\top X^\top = 
 \begin{bmatrix} 
@@
   \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}$$
+\end{bmatrix}\]
 
@@
 
-*Proof*. 
-By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
-(\ref{eq:xy}) and $p$ and $q$ are defined as in equations (\ref{eq:p}) and
-(\ref{eq:q}). Alternatively, the spectral decomposition can be written as 
-$P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
-and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some 
-$Q \in \mathbb{O}(p, q)$,
-
-$$X U Q = V |D|^{1/2}$$
-
-Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,
-
-$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$
+\emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
+\(X\) is defined as in (\ref{eq:xy}) and \(p\) and \(q\) are defined as
+in equations (\ref{eq:p}) and (\ref{eq:q}). Alternatively, the spectral
+decomposition can be written as
+\(P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) for the same
+\((p, q)\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
+\(Q \in \mathbb{O}(p, q)\),
+
+\[X U Q = V |D|^{1/2}\]
+
+Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
+
+\[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
+(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top\]
 
@@
 
-$$\begin{split}
+\[\begin{split}
   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
@@
   & = X (X^\top X)^{-1} X^\top
-\end{split}$$
+\end{split}\]
 
@@
 
-$$\begin{split}
+\[\begin{split}
   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
@@
   & = V V^\top
-\end{split}$$
-
-**Proof of Theorem \ref{theorem3}**.
-By Lemma \ref{lemma1}, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in (\ref{eq:xy}). Since $X$ is block diagonal with each 
-block corresponding to one community, $X (X^\top X)^{-1} X^\top$ is also a 
-block diagonal matrix with each block corresponding to a community and zeros
-elsewhere. Therefore, if vertices $i$ and $j$ belong to different communities,
-then the $ij$^th^ element of $n X (X^\top X)^{-1} X^\top = n V V^\top = B$ 
-is 0.
-
-**Proof of Theorem \ref{theorem4}**. 
-Let $V_n$ and $\hat{V}_n$ be the eigenvectors of $P$ and $A$ corresponding 
-to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative
+\end{split}\]
+
+\textbf{Proof of Theorem \ref{theorem3}}. By Lemma \ref{lemma1},
+\(V V^\top = X (X^\top X)^{-1} X^\top\) where \(X\) is defined as in
+(\ref{eq:xy}). Since \(X\) is block diagonal with each block
+corresponding to one community, \(X (X^\top X)^{-1} X^\top\) is also a
+block diagonal matrix with each block corresponding to a community and
+zeros elsewhere. Therefore, if vertices \(i\) and \(j\) belong to
+different communities, then the \(ij\)\textsuperscript{th} element of
+\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
+
+\textbf{Proof of Theorem \ref{theorem4}}. Let \(V_n\) and \(\hat{V}_n\)
+be the eigenvectors of \(P\) and \(A\) corresponding to the
+\(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
 eigenvalues. By \citeauthor{rubindelanchy2017statistical}, for some
-$W \in \mathbb{O}(K^2)$, and $c > 0$,  
-$||\hat{V} W - V||_{2 \to \infty} = 
-O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big)$. 
-We furthermore have $||V||_{2 \to \infty} = O_P(n^{-1/2})$. Then if
-$(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ correspond to the rows of 
-$V_n$ and $\hat{V}_n$, for $i$ and $j$ in different communities, using the 
-fact that $(v_n^{(i)})^\top v_n^{(j)} = 0$,
-
-$$\begin{split}
+\(W \in \mathbb{O}(K^2)\), and \(c > 0\),\\
+\(||\hat{V} W - V||_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big)\).
+We furthermore have \(||V||_{2 \to \infty} = O_P(n^{-1/2})\). Then if
+\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) correspond to the
+rows of \(V_n\) and \(\hat{V}_n\), for \(i\) and \(j\) in different
+communities, using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
+
+\[\begin{split}
 \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
@@
 & = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
-\end{split}$$
-
-Then scaling by $n$, we get 
-$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = 
-O_P \big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \big)$.
+\end{split}\]
+
+Then scaling by \(n\), we get
+\(|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \big)\).
 
@@
 
-*Proof*. 
-Since by Theorem \ref{theorem2} each $\mathcal{S}^{(k)}$ are mutually orthogonal, any vector projected to 
-$\mathcal{S}^{(k)}$ must be orthogonal to each row of $V^{(-k)}$. Therefore, $W^{(k)} v = 0$ $\forall v \in \mathcal{S}^{(-k)}$.
+\emph{Proof}. Since by Theorem \ref{theorem2} each \(\mathcal{S}^{(k)}\)
+are mutually orthogonal, any vector projected to \(\mathcal{S}^{(k)}\)
+must be orthogonal to each row of \(V^{(-k)}\). Therefore,
+\(W^{(k)} v = 0\) \(\forall v \in \mathcal{S}^{(-k)}\).
 
@@
 
-**Proof of Theorem \ref{theorem5}**. 
-The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which 
-states that the subspace detection property holds if the noise is small enough 
-and the subspace inradius is greater than the subspace incoherence for each 
-community $k$.
-
-Let $V_{n, -i}^{(k)}$ be $V_n^{(k)}$ with the $i$^th^ entry removed. Suppose 
-that for each community $k$, there are enough vertices such that for each $i$, 
-$V_{n, -i}^{(k)}$ spans its corresponding subspace (Theorem \ref{theorem2}). 
-Then $r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0$. 
-Thus by (\ref{eq:mu-conv}), for each $k$, $r_n^{(k)} > \mu_n^{(k)} = 0$ where 
-$\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})$ and $n$ is large enough such that 
-$\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K$.
-
-Let $r_n = \min\limits_k r_n^{(k)}$. By (\ref{eq:delta-conv}), 
-$\delta_n \stackrel{a.s.}{\to} 0$. Then as $n\to \infty$, 
-$\delta_n < \min\limits_k 
-\frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}$ 
-$= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}$
-with probability 1.
-
-Thus the conditions for the subspace detection property from Theorem 6 from
-\citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as $n \to \infty$.
+\textbf{Proof of Theorem \ref{theorem5}}. The basis of this proof is
+Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
+subspace detection property holds if the noise is small enough and the
+subspace inradius is greater than the subspace incoherence for each
+community \(k\).
+
+Let \(V_{n, -i}^{(k)}\) be \(V_n^{(k)}\) with the
+\(i\)\textsuperscript{th} entry removed. Suppose that for each community
+\(k\), there are enough vertices such that for each \(i\),
+\(V_{n, -i}^{(k)}\) spans its corresponding subspace (Theorem
+\ref{theorem2}). Then
+\(r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0\). Thus by
+(\ref{eq:mu-conv}), for each \(k\), \(r_n^{(k)} > \mu_n^{(k)} = 0\)
+where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
+such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).
+
+Let \(r_n = \min\limits_k r_n^{(k)}\). By (\ref{eq:delta-conv}),
+\(\delta_n \stackrel{a.s.}{\to} 0\). Then as \(n\to \infty\),
+\(\delta_n < \min\limits_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}\)
+\(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
+probability 1.
+
+Thus the conditions for the subspace detection property from Theorem 6
+from \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as
+\(n \to \infty\).
 
@@
 
-**Proof of Theorem \ref{theorem6}**. 
-Let $P$ and $A$ be organized by community such that the elements of blocks 
-$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
-$l$.
-
-*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
-probabilities and edges for community $k$.  
-By definition, 
-$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
-singular value decomposition 
-$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
-one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
-singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
-Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
-value decomposition of $A^{(kk)}$, and let 
-$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
+\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) and \(A\) be
+organized by community such that the elements of blocks \(P^{(kl)}\) and
+\(A^{(kl)}\) correspond to the edges between communities \(k\) and
+\(l\).
+
+\emph{Case \(k = l\)}. \(P^{(kk)}\) and \(A^{(kk)}\) represent
+within-community edge probabilities and edges for community \(k\).\\
+By definition, \(P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top\). This
+implies that the singular value decomposition
+\(P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top\) has one singular
+value and one pair of singular vectors (\(P^{(kk)}\) is symmetric, so
+the left and right singular vectors are identical). Then
+\(\lambda^{(kk)} = \sigma_{kk} u^{(kk)}\).\\
+Let \(\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top\) be the
+singular value decomposition of \(A^{(kk)}\), and let
+\(\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top\) be its
 one-dimensional approximation. Define
-$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
-$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
-$\lambda^{(kk)}$.  
-Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the adjacency 
-spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
-$\frac{(\log n_k)^c}{\sqrt{n_k}}$.
-
-*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
-edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
-By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
-$k = l$ case, we note that the singular value decomposition 
-$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
-$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
-$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
-$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
-$v^{(kl)} = u^{(lk)}$.)  
+\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
+\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding
+approximation of \(\lambda^{(kk)}\).\\
+Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the
+adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
+\(\lambda^{(kk)}\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
+
+\emph{Case \(k \neq l\)}. \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
+probabilities and edges between communities \(k\) and \(l\). Note that
+\(P^{(kl)} = (P^{(lk)})^\top\).\\
+By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). As
+in the \(k = l\) case, we note that the singular value decomposition
+\(P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top\) is one-dimensional
+and \(\lambda^{(kl)} = \sigma_{kl} u^{(kl)}\). (We can also note that
+the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
+\(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
+\(v^{(kl)} = u^{(lk)}\).)\\
 Now consider the Hermitian dilation
 
-$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$
-
-which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
-that the spectral decomposition of $M^{(kl)}$ is
-
-$$M^{(kl)} = 
+\[M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}\]
+
+which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. It can
+be shown that the spectral decomposition of \(M^{(kl)}\) is
+
+\[M^{(kl)} = 
 \begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
 \begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$
-
-Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
-latent positions in $\mathbb{R}^2$ given by
-
-$$\begin{bmatrix} 
+\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top\]
+
+Thus treating \(M^{(kl)}\) as the edge probability matrix of a GRDPG, we
+have latent positions in \(\mathbb{R}^2\) given by
+
+\[\begin{bmatrix} 
   \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
@@
   \lambda^{(lk)} & -\lambda^{(lk)} 
-\end{bmatrix}$$
+\end{bmatrix}\]
 
@@
 
-$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
-Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where
-
-$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$
-
-and $E$ is the 
-$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
-in $A^{(kl)}$.  
-Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
-spectral embedding, given by
-
-$$\begin{bmatrix} 
+\[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
+Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
+
+\[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
+
+and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
+generate the Bernoulli entries in \(A^{(kl)}\).\\
+Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
+so its adjacency spectral embedding, given by
+
+\[\begin{bmatrix} 
   \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
   \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
-\end{bmatrix}$$
-
-where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 3, approximates 
-the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
-by the rate given in Theorem 5 of \citeauthor{rubindelanchy2017statistical}.  
-In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
-result \cite{rubindelanchy2017statistical} is of the form 
-$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
-for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
-resulting in $W_* \stackrel{a.s.}{\to} I$. Therefore, the adjacency spectral
-embedding of $\hat{M}^{(kl)}$ is a direct estimation of the specific latent 
-positions outlined for $M^{(kl)}$, up to sign flip.
-
-# References
+\end{bmatrix}\]
+
+where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3,
+approximates the latent positions of \(M^{(kl)}\) up to indefinite
+orthogonal transformation by the rate given in Theorem 5 of
+\citeauthor{rubindelanchy2017statistical}.\\
+In this case, the indefinite orthogonal transformation \(W_*\) in the
+GRDPG result \cite{rubindelanchy2017statistical} is of the form
+\(U^\top \hat{U}\). The eigenvalues of \(M\) are distinct since the
+signature for this GRDPG is \((1, 1)\), and \(U^\top \hat{U}\) is block
+diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). Therefore, the
+adjacency spectral embedding of \(\hat{M}^{(kl)}\) is a direct
+estimation of the specific latent positions outlined for \(M^{(kl)}\),
+up to sign flip.
+
+\renewcommand\refname{References}
+  \bibliography{misc.bib}
+
+\end{document}

### 2021-05-05 summary.tex
updated document

@@
 \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
-\(\min_{c_i} ||c_i||_1\) subject to \(x_i = X c_i\) and \(\beta_i = 0\)
-is solved for each \(i = 1, ..., n\). The solutions are collected
-collected into matrix
+\(c_i = \arg\min_{c} ||c||_1\) subject to \(x_i = X c\) and
+\(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
+collected collected into matrix
 \(C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top\) to
@@
 points near \(K\) subspaces with some noise, a final graph partitioning
-step is performed (e.g., edge thresholding or spectral clustering).
+step may be performed (e.g., edge thresholding or spectral clustering).
 
@@
 ASE of \(P\) results in subspaces that correspond to each community,
-suggesting the use of SSC. In this paper, we will use a different and
-leave the properties of SSC on the ASE of \(A\) as future work.
+suggesting the use of SSC. Before exploring SSC, we will first consider
+a different approach.
 
@@
 positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
-$\exists \lambda > 0$ such that as $n \to \infty$, $\hat{V}_n$ obeys the 
-subspace detection property with probability 1.
+$\exists \lambda > 0$ and $N \in \mathbb{N}$ such that when $n > N$, 
+$\sqrt{n} \hat{V}_n$ obeys the subspace detection property with probability 1.
 \end{theorem}

### 2021-05-20 sdss-talk/slides.Rmd
uploading slides for sdss talk

---
title: Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
subtitle: SDSS 2021 Lightning Presentation
author: John Koo
date: 'June 2021'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
set.seed(314159)
```

## Overview

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

1. Block Models and the Popularity Adjusted Block Model

2. Generalized Random Dot Product Graphs

3. Connecting the PABM to the GRDPG

4. Community Detection for the PABM

# Block Models

## Networks

```{r out.width = '50%'}
knitr::include_graphics('Sna_large.png')
```

\tiny

By Screenshot taken by User:DarwinPeacock - Screenshot of free software GUESS, CC BY 3.0, https://commons.wikimedia.org/w/index.php?curid=6057981

## Bernoulli Graphs

::: columns

:::: {.column width=62.5%}

Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.

$G$ is described by adjacency matrix $A$ such that
$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

$A \sim BernoulliGraph(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \stackrel{ind}{\sim} Bernoulli(P_{ij})$ for each $i < j$.

::::

:::: {.column width=37.5%}

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / log(n)
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

## Block Models

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim BernoulliGraph(P)$ is a *block model*.

Example: Stochastic Block Model (Lorrain and White, 1971) with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

* To make this an assortative SBM, set $p q > r^2$.
* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

::::

:::

## Popularity Adjusted Block Model

Definition based on Noroozi, Rimal, and Pensky (2020); 
model first proposed by Sengupta and Chen (2017).

$A \sim PABM(\{\lambda^{(kl)}\}_K)$ iff

1. w.l.o.g., organize $P$ such that each block 
$P^{(kl)} \in [0, 1]^{n_k \times n_l}$ contains edge 
probabilities between communities $k$ and $l$.
2. Organize parameters as vectors such that 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
are the popularity parameters of members of community $k$ to community $l$.  
$\{\lambda^{(kl)}\}_K$ is the set of $K^2$ popularity vectors.
3. Then we can write each block of $P$ as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
4. Sample $A \sim BernoulliGraph(P)$.

Example: $K = 2$

$$P = \begin{bmatrix}
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} = 
\begin{bmatrix}
\lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
\lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
\end{bmatrix}$$

# Generalized Ranodm Dot Product Graphs

## Generalized Random Dot Product Graph

Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = blockdiag(I_p, -I_q)$
* $A \sim BernoulliGraph(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \stackrel{iid}{\sim} F$, then we write 
$(A, X) \sim GRDPG_{p, q}(F, n)$.

## (Generalized) Random Dot Product Graph Model

### Recovery/Estimation

Want to estimate $X$ given $A$, or alternatively, 
interpoint distances, inner products, or angles.

### Adjacency Spectral Embedding

To embed in $\mathbb{R}^{p+q}$,

1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
where $\hat{\Lambda} \in \mathbb{R}^{(p+q) \times (p+q)}$ and 
$\hat{V} \in \mathbb{R}^{n \times (p+q)}$ by using 
$p$ most positive and $q$ most negative eigenvalues.

2. Let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.

\vspace*{.5\baselineskip}

$$\max_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0
\text{ (Rubin-Delanchy et al., 2020)}$$

$$Q_n \in \mathbb{O}(p, q)$$

## Connecting Block Models to the GRDPG

::: columns

:::: column

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
par(mar = rep(3, 4))
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

::::

:::: column

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

::::

:::

# Connecting the PABM to the GRDPG

## Connecting the PABM to the GRDPG ($K = 2$)

**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to 
$A \sim GRDPG_{3, 1}(X U)$ for block diagonal $X$ 
constructed from $\{\lambda^{(kl)}\}_2$ and 
predetermined $U \in \mathbb{O}(4)$.

Proof: Decompose $P$ as follows

$$X = \begin{bmatrix}
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
0 & 0 & \lambda^{(21)} & \lambda^{(22)}
\end{bmatrix}$$

$$U = \begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & -1 / \sqrt{2} \\
0 & 1 & 0 & 0
\end{bmatrix}$$

$$P = (X U) I_{3, 1} (X U)^\top = \begin{bmatrix}
\lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
\lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
\end{bmatrix}$$

## Connecting the PABM to the GRDPG ($K \geq 2$)

**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim GRDPG_{p, q}(X U)$ such that

* $p = K (K + 1) / 2$
* $q = K (K - 1) / 2$
* $U$ is orthogonal and predetermined for each $K$
* $X$ is block diagonal and composed of $\{\lambda^{(kl)}\}_K$  
$\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $x_i^\top x_j = 0$.

**Remark** (non-uniqueness of the latent configuration):  
$A \sim GRDPG_{p, q}(X U) \implies A \sim GRDPG_{p, q}(X U Q)$ 
$\forall Q \in \mathbb{O}(p, q)$

# Community Detection for the PABM

## Sparse Subspace Clustering

**Corollary**: $X$ is block diagonal by community and 
$U$ is orthogonal $\implies$ 
each community corresponds to a subspace in $\mathbb{R}^{K^2}$.

Subspace property holds even with linear transformation 
$Q \in \mathbb{O}(p, q)$.

Noroozi et al. observed that the rank of $P$ is $K^2$ 
and the columns of $P$ belonging to each community has rank $K$ 
to justify SSC for the PABM.

$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } A_{\cdot, i} = A c_i
\text{ and } c_i^{(i)} = 0$$

GRDPG-based approach: Apply SSC to the ASE of $A$.

$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } 
\hat{x}_i = \hat{X} c_i \text{ and } c_i^{(i)} = 0$$
$$A \approx \hat{X} I_{p, q} \hat{X}^\top$$

## Sparse Subspace Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
then $B_{ij} = 0$ $\forall i, j$ in different communities.

**Theorem** (KTT):

Let

* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
$A_n \sim BernoulliGraph(P_n)$.
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

Then

* $\exists N < \infty$ such that when $n > N$, 
$\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1.

## Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
then $B_{ij} = 0$ $\forall i, j$ in different communities.

Orthogonal Spectral Clustering algorithm:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,

$$
\max_{i, j} \hat{B}^{(ij)} = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

## Simulation Results

```{r clust_err_k, fig.width = 10, fig.height = 3, out.width = '100%'}
setwd('..')
clustering.df <- readr::read_csv('clustering-k.csv')
ssc.df <- readr::read_csv('clustering-ssc-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error),
    first.q = quantile(error, .25),
    third.q = quantile(error, .75),
    med.err.ssc = median(error.ssc),
    first.q.ssc = quantile(error.ssc, .25),
    third.q.ssc = quantile(error.ssc, .75),
    med.err.ep = median(error.ep, na.rm = TRUE),
    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  scale_y_log10() +
  labs(y = 'community detection error rate', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC-ASE'), width = .1) + 
  # geom_line(aes(x = n, y = med.err.ep,
  #               colour = 'MM-EP')) + 
  # geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
  #                   colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.louvain,
                colour = 'MM-Louvain')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
                    colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

\tiny

IQR of community detection error rates using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green). Communities are approximately balanced. Simulations were repeated 50 times for each sample size.

### 2021-05-21 sdss-talk/slides.Rmd
updated sdss presentation

@@
 subtitle: SDSS 2021 Lightning Presentation
-author: John Koo
+author: John Koo, Indiana University
 date: 'June 2021'
@@
 
+## Contributors
+
+::: columns
+
+:::: {.column width='33%'}
+
+```{r, out.width = '100px'}
+knitr::include_graphics('john_koo.jpeg')
+```
+
+John Koo,  
+PhD Student in Statistical Science,  
+Indiana University
+
+::::
+
+:::: {.column width='33%'}
+
+```{r, out.width = '100px'}
+knitr::include_graphics('minh_tang.jpg')
+```
+
+Minh Tang,  
+Assistant Professor of Statistics,  
+NC State University
+
+::::
+
+:::: {.column width='33%'}
+
+```{r, out.width = '100px'}
+knitr::include_graphics('mtrosset.jpg')
+```
+
+Michael Trosset,  
+Professor of Statistics,  
+Indiana University
+
+::::
+
+:::
+
 ## Overview
@@
 
-By Screenshot taken by User:DarwinPeacock - Screenshot of free software GUESS, CC BY 3.0, https://commons.wikimedia.org/w/index.php?curid=6057981
+Screenshot taken by User:DarwinPeacock - Screenshot of free software GUESS, CC BY 3.0, https://commons.wikimedia.org/w/index.php?curid=6057981
 
@@
 
+\vspace*{1\baselineskip}
+
 $G$ is described by adjacency matrix $A$ such that
@@
 $A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.
+
+\vspace*{1\baselineskip}
 
@@
 
-# Generalized Ranodm Dot Product Graphs
+## Community Detection in Block Models
+
+Likelihood
+
+$$L = \prod_{i<j} \prod_{k, l}^K 
+\big(p_{k, l, i, j}^{A_{ij}} 
+(1 - p_{k, l, i, j})^{1 - A_{ij}} \big)^{z_{ik} z_{jl}}$$
+
+* ML method for community detection: $\hat{\vec{z}} = \arg\max_{\vec{z}} L$
+
+* NP-complete
+  * Expectation-Maximization
+  * Bayesian methods
+  * **Spectral methods**
+
+# Generalized Random Dot Product Graphs
 
@@
 
-Want to estimate $X$ given $A$, or alternatively, 
-interpoint distances, inner products, or angles.
+Want to estimate $X$ given $A$.
 
@@
 
-$$\max_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0
-\text{ (Rubin-Delanchy et al., 2020)}$$
+Rubin-Delanchy et al., 2020:  
+$$\max_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$$
 
@@
 
-# Community Detection for the PABM
-
-## Sparse Subspace Clustering
-
 **Corollary**: $X$ is block diagonal by community and 
@@
 
+# Community Detection for the PABM
+
+## Sparse Subspace Clustering
+
+*Sparse Subspace Clustering* (Elhamifar and Vidal, 2009):  
+Let the rows of $X \in \mathbb{R}^{n \times d}$ are vectors that lie on or near a small number of low-dimensional subspaces. SSC constructs a graph from $X$ as follows:
+
+1. For each $i \in [n]$, solve the optimization problem  
+$c_i = \arg\min_{c} \|c\|_1$ subject to $x_i = X c$ and $c^{(i)} = 0$.
+2. Construct a similarity graph using $B = |C| + |C^\top|$ as its affinity matrix, where $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
+
+If $X$ obeys the *Subspace Detection Property*, then it produces a similarity graph $B$ such that $B_{ij} = 0$ $\forall z_i \neq z_j$.
+
+## Sparse Subspace Clustering for Community Detection
+
 Noroozi et al. observed that the rank of $P$ is $K^2$ 
@@
 
-$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } A_{\cdot, i} = A c_i
-\text{ and } c_i^{(i)} = 0$$
+$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } A_{\cdot, i} = A c
+\text{ and } c^{(i)} = 0$$
 
@@
 
-$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } 
-\hat{x}_i = \hat{X} c_i \text{ and } c_i^{(i)} = 0$$
+$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
+\hat{x}_i = \hat{X} c \text{ and } c^{(i)} = 0$$
 $$A \approx \hat{X} I_{p, q} \hat{X}^\top$$
@@
 ## Sparse Subspace Clustering
-
-**Theorem** (KTT): 
-If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
-then $B_{ij} = 0$ $\forall i, j$ in different communities.
 
@@
 **Theorem** (KTT): 
-If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
-then $B_{ij} = 0$ $\forall i, j$ in different communities.
+If $P = V \Lambda V^\top$ is the spectral decomposition of $P$ for the PABM and 
+$V$ has rows $v_i^\top$, then $v_i^\top v_j = 0$ $\forall z_i \neq z_j$.
 
-Orthogonal Spectral Clustering algorithm:
+**Theorem** (KTT): 
+If $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ is the spectral decomposition 
+of $A$ for the PABM using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ 
+most negative eigenvalues, and we denote $\hat{v}_i^\top$ as the rows of 
+$\hat{V}$, then
+
+$$\max_{i, j : z_i \neq z_j} \hat{v}_i^\top \hat{v}_j = 
+O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
+
+Orthogonal Spectral Clustering (KTT):
 
@@
 2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
-3. Construct graph $G$ using $B$ as its similarity matrix.
-4. Partition $G$ into $K$ disconnected subgraphs.
+3. Construct graph $\hat{G}$ using $B$ as its similarity matrix.
+4. Partition $\hat{G}$ into $K$ disconnected subgraphs.
 
-**Theorem** (KTT): 
-Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
-Then $\forall$ pairs $(i, j)$ belonging to different communities 
-and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,
+## Simulation Study
 
-$$
-\max_{i, j} \hat{B}^{(ij)} = 
-O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
-$$
+Simulation setup:
 
-## Simulation Results
+1. $Z_1, ..., Z_n \stackrel{iid}{\sim} Categorical(1 / K, ..., 1 / K)$
+2. $\lambda_{ik} \stackrel{iid}{\sim} Beta(a_{ik}, b_{ik})$  
+$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$, 
+$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
+3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
+4. $A \sim BernoulliGraph(P)$
 
@@
 
-\tiny
+## Conclusion
 
-IQR of community detection error rates using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.
+1. The PABM is a recently developed flexible block model that can be used to describe many graphs with community structure.
+
+2. Likelihood maximization is NP-complete for block models, so we need to take a different approach to community detection.
+
+3. The GRDPG, which describes all block models, motivates a spectral approach to statistical inference on graphs.
+
+4. Under the GRDPG framework, the PABM is induced by a latent configuration in $\mathbb{R}^{K^2}$ consisting of $K$ $K$-dimensional subspaces that are all orthogonal to each other.
+
+5. The latent configuration of the PABM under the GRDPG framework leads to two methods for community detection, both of which have nice theoretical asymptotic properties.
+
+## Thank you
+
+Code and draft of the paper available at  
+https://github.com/johneverettkoo/pabm-grdpg
+
+R package coming soon at https://github.com/johneverettkoo/pabm

### 2021-05-31 sdss-talk/slides-4min.Rmd
slides for 4 minute presentation

--- 2021-05-21 sdss-talk/slides.Rmd
@@
 
-## Contributors
+## The Popularity Adjusted Block Model
 
-::: columns
+\newcommand{\diag}{\text{diag}}
+\newcommand{\tr}{\text{Tr}}
+\newcommand{\blockdiag}{\text{blockdiag}}
+\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
+\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
+\newcommand{\Bernoulli}{\text{Bernoulli}}
+\newcommand{\Betadist}{\text{Beta}}
+\newcommand{\BG}{\text{BernoulliGraph}}
+\newcommand{\Cat}{\text{Categorical}}
+\newcommand{\Uniform}{\text{Uniform}}
+\newcommand{\RDPG}{\text{RDPG}}
+\newcommand{\GRDPG}{\text{GRDPG}}
+\newcommand{\PABM}{\text{PABM}}
 
-:::: {.column width='33%'}
+**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
 
-```{r, out.width = '100px'}
-knitr::include_graphics('john_koo.jpeg')
-```
+Let each vertex $i \in [n]$ have $K$ popularity parameters
+$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
+Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
+e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
 
-John Koo,  
-PhD Student in Statistical Science,  
-Indiana University
+**Lemma** (Noroozi, Rimal, and Pensky, 2020):
 
-::::
+$A$ is sampled from a PABM if $P$ can be described as:
 
-:::: {.column width='33%'}
+1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
+between communities $k$ and $l$. 
+2. Organize popularity parameters as vectors 
+$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
+such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
+of the $i$^th^ vertex of community $k$ towards community $l$. 
+3. Each block can be decomposed as 
+$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
 
-```{r, out.width = '100px'}
-knitr::include_graphics('minh_tang.jpg')
-```
+**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.
 
-Minh Tang,  
-Assistant Professor of Statistics,  
-NC State University
-
-::::
-
-:::: {.column width='33%'}
-
-```{r, out.width = '100px'}
-knitr::include_graphics('mtrosset.jpg')
-```
-
-Michael Trosset,  
-Professor of Statistics,  
-Indiana University
-
-::::
-
-:::
-
-## Overview
-
-<style type="text/css">
-.caption {
-    font-size: x-small;
-}
-</style>
-
-1. Block Models and the Popularity Adjusted Block Model
-
-2. Generalized Random Dot Product Graphs
-
-3. Connecting the PABM to the GRDPG
-
-4. Community Detection for the PABM
-
-# Block Models
-
-## Networks
-
-```{r out.width = '50%'}
-knitr::include_graphics('Sna_large.png')
-```
-
-\tiny
-
-Screenshot taken by User:DarwinPeacock - Screenshot of free software GUESS, CC BY 3.0, https://commons.wikimedia.org/w/index.php?curid=6057981
-
-## Bernoulli Graphs
-
-::: columns
-
-:::: {.column width=62.5%}
-
-Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.
-
-\vspace*{1\baselineskip}
-
-$G$ is described by adjacency matrix $A$ such that
-$A_{ij} = \begin{cases} 
-1 & \exists \text{ edge between } i \text{ and } j \\
-0 & \text{else}
-\end{cases}$
-
-$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.
-
-\vspace*{1\baselineskip}
-
-$A \sim BernoulliGraph(P)$ iff:
-
-1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
-vertices.
-2. $A_{ij} \stackrel{ind}{\sim} Bernoulli(P_{ij})$ for each $i < j$.
-
-::::
-
-:::: {.column width=37.5%}
-
-```{r, fig.height = 2, fig.width = 2}
-n <- 2 ** 5
-p <- 1 / log(n)
-P <- matrix(p, nrow = n, ncol = n)
-A <- draw.graph(P)
-qgraph::qgraph(A, vsize = 4)
-```
-
-::::
-
-:::
-
-## Block Models
-
-Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
-and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
-Then $A \sim BernoulliGraph(P)$ is a *block model*.
-
-Example: Stochastic Block Model (Lorrain and White, 1971) with two communities
-
-::: columns
-
-:::: column
-
-* $z_1, ..., z_n \in \{1, 2\}$
-* $P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
-\end{cases}$
-
-* To make this an assortative SBM, set $p q > r^2$.
-* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.
-
-::::
-
-:::: column
-
-```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
-n1 <- 2 ** 5
-n2 <- 2 ** 5
-n <- n1 + n2
-z <- c(rep(1, n1), rep(2, n2))
-p <- 1/2
-q <- 1/4
-r <- 1/8
-P <- matrix(r, nrow = n, ncol = n)
-P[seq(n1), seq(n1)] <- p
-P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
-A <- draw.graph(P)
-qgraph::qgraph(A, vsize = 4, groups = factor(z))
-```
-
-::::
-
-:::
-
-## Popularity Adjusted Block Model
-
-Definition based on Noroozi, Rimal, and Pensky (2020); 
-model first proposed by Sengupta and Chen (2017).
-
-$A \sim PABM(\{\lambda^{(kl)}\}_K)$ iff
-
-1. w.l.o.g., organize $P$ such that each block 
-$P^{(kl)} \in [0, 1]^{n_k \times n_l}$ contains edge 
-probabilities between communities $k$ and $l$.
-2. Organize parameters as vectors such that 
-$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
-are the popularity parameters of members of community $k$ to community $l$.  
-$\{\lambda^{(kl)}\}_K$ is the set of $K^2$ popularity vectors.
-3. Then we can write each block of $P$ as 
-$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
-4. Sample $A \sim BernoulliGraph(P)$.
-
-Example: $K = 2$
-
-$$P = \begin{bmatrix}
-P^{(11)} & P^{(12)} \\
-P^{(21)} & P^{(22)}
-\end{bmatrix} = 
-\begin{bmatrix}
-\lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-\lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}$$
-
-## Community Detection in Block Models
-
-Likelihood
-
-$$L = \prod_{i<j} \prod_{k, l}^K 
-\big(p_{k, l, i, j}^{A_{ij}} 
-(1 - p_{k, l, i, j})^{1 - A_{ij}} \big)^{z_{ik} z_{jl}}$$
-
-* ML method for community detection: $\hat{\vec{z}} = \arg\max_{\vec{z}} L$
-
-* NP-complete
-  * Expectation-Maximization
-  * Bayesian methods
-  * **Spectral methods**
-
-# Generalized Random Dot Product Graphs
-
-## Generalized Random Dot Product Graph
+## The Generalized Random Dot Product Graph
 
@@
 
-## Connecting Block Models to the GRDPG
-
-::: columns
-
-:::: column
-
-```{r, out.width = '100%', fig.height = 4, fig.width = 4}
-par(mar = rep(3, 4))
-P.eigen = eigen(P)
-X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
-plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
-```
-
-::::
-
-:::: column
-
-```{r, out.width = '100%', fig.height = 4, fig.width = 4}
-par(mar = rep(1.75, 4))
-Pz <- generate.P.beta(n)
-P <- Pz$P
-z <- Pz$clustering
-X <- embedding(P)
-pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
-```
-
-::::
-
-:::
-
-# Connecting the PABM to the GRDPG
-
-## Connecting the PABM to the GRDPG ($K = 2$)
-
-**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to 
-$A \sim GRDPG_{3, 1}(X U)$ for block diagonal $X$ 
-constructed from $\{\lambda^{(kl)}\}_2$ and 
-predetermined $U \in \mathbb{O}(4)$.
-
-Proof: Decompose $P$ as follows
-
-$$X = \begin{bmatrix}
-\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
-0 & 0 & \lambda^{(21)} & \lambda^{(22)}
-\end{bmatrix}$$
-
-$$U = \begin{bmatrix}
-1 & 0 & 0 & 0 \\
-0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
-0 & 0 & 1 / \sqrt{2} & -1 / \sqrt{2} \\
-0 & 1 & 0 & 0
-\end{bmatrix}$$
-
-$$P = (X U) I_{3, 1} (X U)^\top = \begin{bmatrix}
-\lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-\lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}$$
-
-## Connecting the PABM to the GRDPG ($K \geq 2$)
+## Connecting the PABM to the GRDPG
 
@@
 
-# Community Detection for the PABM
-
-## Sparse Subspace Clustering
-
-*Sparse Subspace Clustering* (Elhamifar and Vidal, 2009):  
-Let the rows of $X \in \mathbb{R}^{n \times d}$ are vectors that lie on or near a small number of low-dimensional subspaces. SSC constructs a graph from $X$ as follows:
-
-1. For each $i \in [n]$, solve the optimization problem  
-$c_i = \arg\min_{c} \|c\|_1$ subject to $x_i = X c$ and $c^{(i)} = 0$.
-2. Construct a similarity graph using $B = |C| + |C^\top|$ as its affinity matrix, where $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
-
-If $X$ obeys the *Subspace Detection Property*, then it produces a similarity graph $B$ such that $B_{ij} = 0$ $\forall z_i \neq z_j$.
-
-## Sparse Subspace Clustering for Community Detection
-
-Noroozi et al. observed that the rank of $P$ is $K^2$ 
-and the columns of $P$ belonging to each community has rank $K$ 
-to justify SSC for the PABM.
-
-$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } A_{\cdot, i} = A c
-\text{ and } c^{(i)} = 0$$
-
-GRDPG-based approach: Apply SSC to the ASE of $A$.
-
-$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
-\hat{x}_i = \hat{X} c \text{ and } c^{(i)} = 0$$
-$$A \approx \hat{X} I_{p, q} \hat{X}^\top$$
-
-## Sparse Subspace Clustering
-
-**Theorem** (KTT):
-
-Let
-
-* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
-$A_n \sim BernoulliGraph(P_n)$.
-* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
-$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.
-
-Then
-
-* $\exists N < \infty$ such that when $n > N$, 
-$\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1.
-
-## Orthogonal Spectral Clustering
+## Connecting the PABM ot the GRDPG
 
@@
 
-$$\max_{i, j : z_i \neq z_j} \hat{v}_i^\top \hat{v}_j = 
+$$\max_{i, j : z_i \neq z_j} n \hat{v}_i^\top \hat{v}_j = 
 O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
@@
                     colour = 'SSC-A'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.louvain,
-                colour = 'MM-Louvain')) + 
-  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
-                    colour = 'MM-Louvain'), width = .1) + 
+  # geom_line(aes(x = n, y = med.err.louvain,
+  #               colour = 'MM-Louvain')) + 
+  # geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
+  #                   colour = 'MM-Louvain'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
 ```
-
-## Conclusion
-
-1. The PABM is a recently developed flexible block model that can be used to describe many graphs with community structure.
-
-2. Likelihood maximization is NP-complete for block models, so we need to take a different approach to community detection.
-
-3. The GRDPG, which describes all block models, motivates a spectral approach to statistical inference on graphs.
-
-4. Under the GRDPG framework, the PABM is induced by a latent configuration in $\mathbb{R}^{K^2}$ consisting of $K$ $K$-dimensional subspaces that are all orthogonal to each other.
-
-5. The latent configuration of the PABM under the GRDPG framework leads to two methods for community detection, both of which have nice theoretical asymptotic properties.
-
-## Thank you
-
-Code and draft of the paper available at  
-https://github.com/johneverettkoo/pabm-grdpg
-
-R package coming soon at https://github.com/johneverettkoo/pabm

### 2021-06-01 sdss-talk/slides-4min.Rmd
updated 4 min presentation slides

@@
 
+```{r out.width = '50%', fig.height = 3, fig.width = 4}
+Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
+P <- Pz$P
+z <- Pz$clustering
+A <- draw.graph(P)
+qgraph::qgraph(A, groups = factor(z), legend = FALSE)
+```
+
 **Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
@@
 e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
+
+## The Popularity Adjusted Block Model
 
@@
 $(A, X) \sim GRDPG_{p, q}(F, n)$.
-
-## (Generalized) Random Dot Product Graph Model
-
-### Recovery/Estimation
-
-Want to estimate $X$ given $A$.
-
-### Adjacency Spectral Embedding
-
-To embed in $\mathbb{R}^{p+q}$,
-
-1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
-where $\hat{\Lambda} \in \mathbb{R}^{(p+q) \times (p+q)}$ and 
-$\hat{V} \in \mathbb{R}^{n \times (p+q)}$ by using 
-$p$ most positive and $q$ most negative eigenvalues.
-
-2. Let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.
-
-\vspace*{.5\baselineskip}
-
-Rubin-Delanchy et al., 2020:  
-$$\max_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$$
-
-$$Q_n \in \mathbb{O}(p, q)$$
 

### 2021-06-09 sdss-talk/slides-4min.Rmd
maybe modify parameter estimation stuff

@@
 subtitle: SDSS 2021 Lightning Presentation
-author: John Koo, Indiana University
+author: |
+  | John Koo, Indiana University
+  | Minh Tang, NC State University
+  | Michael Trosset, Indiana University
 date: 'June 2021'
@@
 $\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$.  
-Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
-e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
+Then $A \sim \BG(P)$ is a PABM if each 
+$P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
 
@@
 
-Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$  
+Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
 (Rubin-Delanchy, Cape, Tang, Priebe, 2020)
@@
 * Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
-$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = blockdiag(I_p, -I_q)$
-* $A \sim BernoulliGraph(X I_{p, q} X^\top)$ where 
+$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
+* $A \sim \BG(X I_{p, q} X^\top)$ where 
 $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
@@
 If latent vectors $X_1, ..., X_n \stackrel{iid}{\sim} F$, then we write 
-$(A, X) \sim GRDPG_{p, q}(F, n)$.
+$(A, X) \sim \GRDPG_{p, q}(F, n)$.
 
@@
 
-**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
-$A \sim GRDPG_{p, q}(X U)$ such that
+**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
+$A \sim \GRDPG_{p, q}(X U)$ such that
 
@@
 **Remark** (non-uniqueness of the latent configuration):  
-$A \sim GRDPG_{p, q}(X U) \implies A \sim GRDPG_{p, q}(X U Q)$ 
+$A \sim \GRDPG_{p, q}(X U) \implies A \sim \GRDPG_{p, q}(X U Q)$ 
 $\forall Q \in \mathbb{O}(p, q)$
@@
 
-1. $Z_1, ..., Z_n \stackrel{iid}{\sim} Categorical(1 / K, ..., 1 / K)$
-2. $\lambda_{ik} \stackrel{iid}{\sim} Beta(a_{ik}, b_{ik})$  
+1. $Z_1, ..., Z_n \stackrel{iid}{\sim} \Cat(1 / K, ..., 1 / K)$
+2. $\lambda_{ik} \stackrel{iid}{\sim} \Betadist(a_{ik}, b_{ik})$  
 $a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$, 
@@
 3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
-4. $A \sim BernoulliGraph(P)$
+4. $A \sim \BG(P)$
 

### 2021-06-09 summary.tex
Update git repo

--- 2021-04-27 summary.tex
@@
 \usepackage{natbib}
+\usepackage{parskip}
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
@@
 \newtheorem*{remark}{Remark}
-\theoremstyle{example}
+\theoremstyle{definition}
 \newtheorem*{example}{Example}
@@
 \begin{proof}
-Define the following matrices from $\{\lambda^{(kl)}\}_K$:
-
-$$\Lambda^{(k)} = 
-\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
-\in \mathbb{R}^{n_k \times K}$$
-
-\begin{equation} \label{eq:xy}
-X = \text{blockdiag}(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
-\end{equation}
-
-$$L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
-\mathbb{R}^{n \times K}$$
-
-$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
-\mathbb{R}^{n \times K^2}$$
-
-Then $P = X Y^\top$.
-
-Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
+First define the following matrices from $\{\lambda^{(kl)}\}_K$: 
+\begin{gather}
+\label{eq:xy}
+\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k,1)} \mid \cdots \mid \lambda^{(k, K)} \end{bmatrix}
+\in \mathbb{R}^{n_k \times K}, \quad 
+X = \text{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
+\mathbb{R}^{n \times K^2} \\
+L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in 
+\mathbb{R}^{n \times K}, \quad 
+Y = \begin{bmatrix} L^{(1)} \mid \cdots \mid L^{(K)} \end{bmatrix} \in
+\mathbb{R}^{n \times K^2}.
+\end{gather}
+We then have $P = X Y^\top$. Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
 $\Pi$, resulting in $P = X \Pi X^\top$.  
@@
 
-Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
-eigenvalues equal to $-1$. $\Pi$ has the decomposed form
-
+Then $\Pi$ has $p = K (K + 1) / 2$ eigenvalues equal to $1$ and $q = K (K - 1) / 2$ 
+eigenvalues equal to $-1$. We therefore have
 \begin{equation} \label{eq:permutation}
@@
 \end{equation}
-
-The edge probability matrix then can be written as:
-
+where $U$ is a $K^2 \times K^2$ orthogonal matrix. The edge probability matrix can then be written as
 \begin{equation} \label{eq:pabm-grdpg}
@@
 \end{equation}
-
-\begin{equation} \label{eq:p}
-p = K (K + 1) / 2
-\end{equation}
-
-\begin{equation} \label{eq:q}
-q = K (K - 1) / 2
-\end{equation}
-
-and we can describe the PABM with $K$ communities as a GRDPG with latent 
-positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
+% \begin{equation} \label{eq:p}
+% p = K (K + 1) / 2
+% \end{equation}
+% \begin{equation} \label{eq:q}
+% q = K (K - 1) / 2
+% \end{equation}
+We can therefore describe the PABM with $K$ communities as a GRDPG with latent 
+positions $X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
+\tfrac{1}{2} K (K - 1)\bigr)$.
 \end{proof}
 
-\begin{example}[$K = 3$] Using the same notation as in Theorem \ref{theorem2}:
-
-$$X = \begin{bmatrix} 
+\begin{example}[$K = 3$] Using the same notation as in Theorem
+  \ref{theorem2}, we can define
+\begin{gather*}
+X = \begin{bmatrix} 
 \lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
@@
 0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
-\end{bmatrix}$$
-
-$$Y = \begin{bmatrix} 
+\end{bmatrix}, \\
+Y = \begin{bmatrix} 
 \lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
@@
 0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
-\end{bmatrix}$$
-
-Then $P = X Y^\top$ and $Y = X \Pi$ where $\Pi$ is a permutation matrix 
-consisting of $3$ fixed points and $3$ cycles of order 2:
-
+\end{bmatrix}.
+\end{gather*}
+Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a permutation matrix of
+the form
 $$\Pi = \begin{bmatrix} 
@@
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
-\end{bmatrix}$$
-
-* Positions 1, 5, 9 are fixed.
-
-* The cycles of order 2 are $(2, 4)$, $(3, 7)$, and $(6, 8)$.
-
-Therefore, we can decompose $\Pi = U I_{6, 3} U^\top$ where the first three 
+\end{bmatrix}.$$
+The matrix $\Pi$ corresponds to a permutation of $\{1,2,\dots,9\}$
+with the following decomposition.
+\begin{enumerate}
+\item Positions 1, 5, 9 are fixed.
+\item There are three cycles of length 2, namely $(2, 4)$, $(3, 7)$, and $(6, 8)$.
+\end{enumerate}
+We can therefore write $\Pi$ as $\Pi = U I_{6, 3} U^\top$ where the first three 
 columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the fixed 
@@
 
-The latent positions are the rows of  
+The matrix $P$ can then be written as a generalized random dot product
+graph where the latent positions are the rows of the matrix
 $$XU = \begin{bmatrix}

### 2021-06-09 summary.Rmd

@@
 $x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
-$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
+$c_i = \arg\min_{c} ||c||_1$ subject to $x_i = X c$ and $c^{(i)} = 0$ is solved 
 for each $i = 1, ..., n$. The solutions are collected collected into matrix 
@@
 If $X$ instead represents points near $K$ subspaces with some noise, a final 
-graph partitioning step is performed (e.g., edge thresholding or spectral 
+graph partitioning step may be performed (e.g., edge thresholding or spectral 
 clustering).
@@
 $K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
-correspond to each community, suggesting the use of SSC. In this paper, we will 
-use a different and leave the properties of SSC on the ASE of $A$ as future
-work.
+correspond to each community, suggesting the use of SSC. Before exploring 
+SSC, we will first consider a different approach.
 

### 2021-06-09 summary.tex
Update git repo

@@
 %
-\documentclass[
-  12pt,
-]{article}
+\documentclass[12pt]{article}
 \usepackage{lmodern}
@@
 \urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
+%\usepackage[margin=1in]{geometry}
 \usepackage{graphicx,grffile}
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
 \makeatletter
@@
 \bibliographystyle{plainnat}
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
 
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
 \newcommand{\diag}{\text{diag}}
@@
 \newtheorem*{example}{Example}
+\newpage
+\spacingset{1.5} % DON'T change the spacing!
 
@@
 Let 
-
 $$X = \begin{bmatrix} 
@@
 0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
-\end{bmatrix}$$
-
-$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
+\end{bmatrix} \quad \text{and} \quad
+U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
 0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
 0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}$$
-
-where each $\lambda^{(kl)}$ is a vector as in Definition 1. 
-Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
-equivalent.
+0 & 1 & 0 & 0 \end{bmatrix},$$
+where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1. 
+Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{(\lambda^{(kl)}\})$ are 
+identically distributed.
 \end{theorem}
@@
 \emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
-\(X\) is defined as in (\ref{eq:xy}) and \(p\) and \(q\) are defined as
-in equations (\ref{eq:p}) and (\ref{eq:q}). Alternatively, the spectral
+\(X\) is defined as in (\ref{eq:xy}) and $p = K(K+1)/2$ and $q = K(K-1)/2$. % are given by $p =
+% K(K+1)/2$ and $q = are defined as
+% in equations (\ref{eq:p}) and (\ref{eq:q}). 
+Alternatively, the spectral
 decomposition can be written as

### 2021-06-15 functions.R
updated parameter estimation simulations

@@ -1,12 +1,21 @@
 embedding <- function(A, p = NULL, q = NULL,
                       scale = TRUE,
                       eps = 1e-6) {
+  if (p + q == 0) {
+    stop('one of p or q must be > 0')
+  }
   n <- nrow(A)
   eigen.A <- eigen(A, symmetric = TRUE)
   if (is.null(p) | is.null(q)) {
     keep <- (abs(eigen.A$values) > eps)
   } else {
-    keep <- c(seq(p), seq(n, n - q + 1))
+    if (p * q > 0) {
+      keep <- c(seq(p), seq(n, n - q + 1))
+    } else if (p == 0) {
+      keep <- seq(n, n - q + 1)
+    } else {
+      keep <- seq(p)
+    }
   }
   
   U <- eigen.A$vectors[, keep]
@@ -217,31 +226,37 @@ estimate.lambda.block <- function(P.block, within = FALSE) {
   }
 }
 
-lambda.rmse <- function(P, Phat, clustering) {
+lambda.rmse <- function(P, A, clustering) {
   K <- max(clustering)
-  n <- nrow(Phat)
+  n <- nrow(A)
   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
   sapply(seq(K), function(k) {
+    n.k <- sum(clustering == k)
     sapply(seq(k), function(l) {
+      n.l <- sum(clustering == l)
       P.block <- P[clustering == k, clustering == l]
-      Phat.block <- Phat[clustering == k, clustering == l]
+      A.block <- A[clustering == k, clustering == l]
       if (k == l) {
-        lambda <- estimate.lambda.block(P.block, TRUE)
-        lambda.hat <- estimate.lambda.block(Phat.block, TRUE)
-        return(sum((lambda - lambda.hat) ** 2))
+        # lambda <- estimate.lambda.block(P.block, TRUE)
+        lambda.hat <- estimate.lambda.block(A.block, TRUE)
+        P.hat <- lambda.hat %*% t(lambda.hat)
+        # return(sum((lambda - lambda.hat) ** 2))
+        return(norm(P.hat - P.block, 'F') / (n.k))
       } else {
-        lambdas <- estimate.lambda.block(P.block)
-        lambda.hats <- estimate.lambda.block(Phat.block)
-        return(sum((lambdas[[1]] - lambda.hats[[1]]) ** 2) + 
[…]

### 2021-06-16 summary.tex
updated writeup

@@
 \newcommand{\Betadist}{\text{Beta}}
+\newcommand{\BG}{\text{BernoulliGraph}}
+\newcommand{\PABM}{\text{PABM}}
+\newcommand{\RDPG}{\text{RDPG}}
+\newcommand{\GRDPG}{\text{GRDPG}}
+\newcommand{\Multinomial}{\text{Multinomial}}
 \newtheorem{definition}{Definition}
@@
 probability distribution based on the community relationships between
-each pair. The goal of the analysis then is population community
+each pair. The goal of this analysis then is population community
 detection, or the recovery of the true underlying community labels for
-each vertex, up to permutation (with some additional parameter
-estimation being of possible interest), assuming some underlying
+each vertex, up to permutation, with some additional parameter
+estimation being of possible interest, assuming some underlying
 probability model. One such model is the Stochastic Block Model (SBM),
@@
 assumes that the edge probability from one vertex to another follows a
-Beronulli distribution with fixed probabilities for each pair of
+Bernoulli distribution with fixed probabilities for each pair of
 community labels. Other random graph models have been proposed and
@@
 \(P = X X^\top\) where
-\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The SBM
-is equivalent to a special case of the RDPG model in which all vertices
+\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The assortative
+SBM is equivalent to a special case of the RDPG for which all vertices
 of a given community share the same position in the latent space
 \cite{lyzinski2014}. It has also been shown that similar random graph
-models, including the DCBM, can be represented in this way
-\cite{lyzinski2014} \cite{rubindelanchy2017consistency}. An analogous
-property exists for the PABM but not for the RDPG model but under the
-\emph{Generalized} Random Dot Product Graph (GRDPG) model. This
-relationship will be explored in this paper and exploited to construct
-algorithms for community detection and parameter estimation for the
+models, including the DCBM, can be represented in a similar fashion
+\cite{lyzinski2014, rubindelanchy2017consistency}. An analogous
+property exists for the PABM, not for the RDPG model but the
+\emph{Generalized} Random Dot Product Graph (GRDPG) model, which is an
+extension of the RDPG to allow edge probability matrices that are not positive
+semidefinite. This relationship will be explored in this paper and exploited to
+construct algorithms for community detection and parameter estimation for the
 PABM.
 
-In this paper, we will only consider undirected graphs, that is the edge
-weight from vertex \(i\) to vertex \(j\) is equal to the edge weight in
-the opposite direction, from vertex \(j\) to vertex \(i\). Furthermore,
-we will only consider unweighted graphs with binary (\(0\) or \(1\))
-edge weights We will also assume that graphs are hollow, i.e., there are
-no edges from a vertex to itself. All such graphs can be represented by
-a symmetric adjacency matrix \(A \in \{0, 1\}^{n \times n}\) for which
-\(A_{ij} = 1\) if there exists an edge between vertices \(i\) and \(j\)
-and \(0\) otherwise, and \(A\) is an element-wise independent Bernoulli
-draw from a symmetric edge probability matrix
-\(P \in [0, 1]^{n \times n}\).
+\hypertarget{notation}{%
+\subsection{Notation and Scope}\label{notation}}
+
+Let $G = (V, E)$ be an unweighted, undirected, and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$.
+$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ and $0$ otherwise.
+Because $G$ is symmetric and hollow,
+$A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$.
+We further restrict our analyses to Bernoulli graphs.
+Let $P \in [0, 1]^{n \times n}$ be a symmetric matrix of edge probabilities.
+Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ for each $1 \leq i < j \leq n$  (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$).
+We denote $A \sim \BG(P)$ as graph with adjacency matrix $A$ sampled from edge probability matrix $P$ in this manner.
+If each vertex has a hidden label in $[K]$, they are denoted as $z_1, ..., z_n$.
+Finally, we denote $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
 
@@
 
+In this section, we show that the PABM is a special case of the GRDPG, i.e.,
+graph $G$ drawn from the PABM can be represented by a collection of latent
+vectors in Euclidean space. We further show that the latent configuration that
+induces the PABM consists of orthogonal subspaces with each subspace
+corresponding to a community.
+
 \hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
-\subsection{The popularity adjusted block model and the generalized
-random dot product
-graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
+\subsection{The Popularity Adjusted Block Model and the Generalized
+Random Dot Product
+Graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
 
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-set of $n$ 
-vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
-columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
-$P^{(kl)}$ describes the edge probabilities between vertices in communities 
+graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \BG(P)$.
+Let each vertex have a community label between $1$ and $K$.
+Then $G$ is drawn from a Popularity Adjusted Block Model if
+each vertex has $K$ popularity parameters that describe its affinity
+toward each of the $K$ communities, i.e., vertex $i$ has popularity parameters
+$\lambda_{i1}, ..., \lambda_{iK}$, and
+each $P_ij = \lambda_{i z_j} \lambda_{j z_i}$.
+
+Another characterization of the PABM is as follows.
+Let the rows and columns of $P$ be arranged by community label
+such that $n_k \times n_l$ block $P^{(kl)}$
+describes the edge probabilities between vertices in communities
 $k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$).
-Let graph $G = (V, E)$ be an undirected, unweighted graph such 
-that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
-realization of $\text{Bernoulli}(P)$, i.e., 
-$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ 
-($A_{ij} = A_{ji}$ and $A_{ii} = 0$).
-
 If each block $P^{(kl)}$ can be written as the outer product of two vectors:
@@
 
-for a set of $K^2$ fixed vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each 
-$\lambda^{(st)}$ is a column vector 
-of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
-is a realization of a popularity adjusted block model with parameters 
-$\{\lambda^{(st)}\}_{s, t = 1}^K$. 
+for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each
+$\lambda^{(st)}$ is a column vector of dimension $n_s$, then graph $G$ is drawn from a PABM with parameters $\{\lambda^{(st)}\}_K$ if its its corresponding adjacency matrix $A \sim \BG(P)$.
 \end{definition}
 
-We will use the notation \(A \sim PABM(\{\lambda^{(kl)}\}_K)\) to denote
+We will use the notation \(A \sim \PABM(\{\lambda^{(kl)}\}_K)\) to denote
 a random adjacency matrix \(A\) drawn from a PABM with parameters
@@
 \label{grdpg}
-Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
-set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that
+Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$.
+If $\exists X \in \mathbb{R}^{n \times d}$ such that
 
 \begin{equation} \label{eq:grdpg}
-  P = X I_{pq} X^\top
+  P = X I_{p,q} X^\top
 \end{equation}
@@
 for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-graph $G = (V, E)$ with adjacency matrix $A$ such that 
-$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
-$A_{ii} = 0$) is a draw from the generalized random dot product graph model 
-with latent positions $X$ and signature $(p, q)$. More precisely, if vertices 
-$i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge 
-probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains 
-the latent positions as rows $x_i^\top$. 
+$G$ is drawn from the Generalized Random Dot Product Graph with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
 \end{definition}
 
-We will use the notation \(A \sim GRDPG_{p,q}(X)\) to denote a random
+We will use the notation \(A \sim \GRDPG_{p,q}(X)\) to denote a random
 adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\).
+\((p, q)\). If instead of fixed latent positions, they are drawn $X_1, ..., X_n \iid F$, we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
 
@@
 The indefinite orthogonal group with signature $(p, q)$ is
-the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
+the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{p, q} Q^{\top} = I_{p, q}\}$,
 denoted as $\mathbb{O}(p, q)$.
@@
 \cite{rubindelanchy2017statistical}.
-More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
+More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, then we also have for any
 $Q \in \mathbb{O}(p, q)$,
-$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
-x_i^\top I_{pq} x_j = P_{ij}$. 
+$(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
+x_i^\top I_{p, q} x_j = P_{ij}$.
 Unlike in the RDPG case, transforming the latent positions via multiplication
@@
 We previously stated one possible set of latent positions that result in
-the edge probability matrix of a PABM, \(P = (XU) I_{pq} (XU)^\top\). If
+the edge probability matrix of a PABM, \(P = (XU) I_{p, q} (XU)^\top\). If
 we have (or can estimate) \(XU\) directly, then both the community
@@
 
-If we decompose \(P = Z I_{pq} Z^\top\), then
+If we decompose \(P = Z I_{p, q} Z^\top\), then
 \(\exists Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
@@
 \(P^{(kl)}\), \(\lambda^{(kl)}\) and \(\lambda^{(lk)}\) are solvable
-exactly (up to multiplication by \(-1\)) using singular value
+up to multiplicative constant using singular value
 decomposition. More specifically, let \(P^{(kl)} = \sigma^2 u v^\top\)
@@
 \(u \in \mathbb{R}^{n_k}\) and \(v \in \mathbb{R}^{n_l}\) are vectors
-and \(\sigma^2 > 0\) is a scalar. Then \(\lambda^{(kl)} = \pm \sigma u\)
-and \(\lambda^{(lk)} = \pm \sigma v\). Given the adjacency matrix \(A\)
+and \(\sigma^2 > 0\) is a scalar. Then \(\lambda^{(kl)} = s_1 u\)
+and \(\lambda^{(lk)} = s_2 v\) where $s_1 s_2 = \sigma^2$.
+Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators (algorithm 3), which converge to the true parameters.
+estimators, which converge to the true parameters, again, up to a
+multiplicative constant. Since these popularity vectors aren't identifiable,
+due to the ambiguity of this constant,
+we instead estimate each edge probability block
+$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$
+using the observed edge block $A^{(kl)}$ (Algorithm 3).
+Alternatively, we can estimate the entire edge probability matrix as
+$\hat{P} = \hat{X} I_{p,q} \hat{X}^\top$, where $\hat{X}$ is the ASE of $A$.
 
@@
   \SetAlgoLined
-  \caption{PABM parameter estimation.}
-  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
-  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
-  Arrange the rows and columns of $A$ by community such that each 
-  $A^{(kl)}$ block consists of estimated edge probabilities between 
-  communities $k$ and $l$.\;
-  \For {$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
-    block.\;
+  \caption{PABM block probability estimation.}
+  \KwData{Adjacency block $A^{(kl)}$}
+  \KwResult{Estimated edge probability block $\hat{P}^{(kl)}$}
+  Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ block.\;
   Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$.
@@
   Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and
-    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
-  }
+  $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.\;
+  Assign $\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$.
 \end{algorithm}
@@
 \label{theorem6}
+Let $n_{kl} = n_k + n_l$.
 Under regularity and sparsity assumptions, given fixed $K$,
@@
 \begin{equation} \label{eq:thm6}
-\max_{k, l \in \{1, ..., K\}} 
-||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
-O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
+\max_{k, l} \|\hat{P}_n^{(kl)} - P_n^{(kl)}\| = O_P \bigg(\frac{(\log n_{kl})}{\sqrt{n_{kl}}} \bigg)
 \end{equation}
@@
 is drawn from \(P\). OSC is then used for community detection, and this
-method is compared against SSC \cite{noroozi2019estimation}
-\cite{soltanolkotabi2014} and MM \cite{igraph}
+method is compared against SSC \cite{noroozi2019estimation,soltanolkotabi2014}
+and MM \cite{igraph}
 \cite{307cbeb9b1be48299388437423d94bf1}. True community labels are used
@@
 \(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
-MLE-based estimator described in \citeauthor{noroozi2019estimation} and
+MLE-based estimator described by \citeauthor{noroozi2019estimation} and
 \citeauthor{307cbeb9b1be48299388437423d94bf1}.
@@
 relaxation of the optimization problem. For these simulations, the
-Louvain algorithm was used, as the EP algorithm proved to be
+Louvain algorithm was used for modularity maximization,
+as \citeauthor{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
 prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
-was verified that EP and Louvain produce comparable results.
+was verified that the Louvain algorithm produces comparable results
+to EP-MM.
 
@@
 
-We also define parameter estimation error as the RMSE up to sign flip:
-
-\[RMSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
-\sqrt{\frac{1}{n} \sum_{k < l} \min_{s \in \{-1, 1\}} 
-\sum_{i = 1}^{n_k} 
-(s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}\]
+We also define parameter estimation error as the RMSE
+of the estimated edge probability blocks:
+
+\[RMSE =
+\sum_{k,l} \sqrt{\frac{1}{n_k n_l} \|P^{(kl)} - \hat{P}^{(kl)}\|_F^2}\]
 
@@
 \(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
-\(A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})\). Each simulation has
+\(A_{ij} \stackrel{indep}{\sim} \Bernoulli(P_{ij})\). Each simulation has
 a unique edge probability matrix \(P\).
@@
   Community labels
-  \(z_k \stackrel{\text{iid}}{\sim}Multinomial(\alpha_1, ..., \alpha_K)\)
+  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
 \item
   Within-group popularities
-  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim}Beta(2, 1)\)
+  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
 \item
   Between-group popularities
-  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim}Beta(1, 2)\) for
+  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
   \(k \neq l\)
@@
 & = ||\hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
+& = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top +
+\hat{V}_n W W^\top \hat{V}_n^\top -
 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
@@
 
-\emph{Case \(k = l\)}. \(P^{(kk)}\) and \(A^{(kk)}\) represent
+\emph{Case \(k = l\)}. \(P^{(kk)}_n\) and \(A^{(kk)}_n\) represent
 within-community edge probabilities and edges for community \(k\).\\
-By definition, \(P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top\). This
+By definition, \(P^{(kk)}_n = \lambda^{(kk)}_n (\lambda^{(kk)}_n)^\top\). This
 implies that the singular value decomposition
-\(P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top\) has one singular
-value and one pair of singular vectors (\(P^{(kk)}\) is symmetric, so
+\(P^{(kk)}_n = (\sigma^{(kk)}_n)^2 u^{(kk)}_n (u^{(kk)}_n)^\top\)
+has one singular
+value and one pair of singular vectors (\(P^{(kk)}_n\) is symmetric, so
 the left and right singular vectors are identical). Then
-\(\lambda^{(kk)} = \sigma_{kk} u^{(kk)}\).\\
-Let \(\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top\) be the
-singular value decomposition of \(A^{(kk)}\), and let
-\(\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top\) be its
+\(\lambda^{(kk)}_n = \sigma^{(kk)}_n u^{(kk)}_n\).\\
+Let \(\hat{U}^{(kk)}_n \hat{\Sigma}^{(kk)}_n (\hat{U}^{(kk)}_n)^\top\) be the
+singular value decomposition of \(A^{(kk)}_n\), and let
+\((\hat{\sigma}^{(kk)})^2 \hat{u}^{(kk)}_n (\hat{u}^{(kk)}_n)^\top\) be its
 one-dimensional approximation. Define
-\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
-\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding
-approximation of \(\lambda^{(kk)}\).\\
+\(\hat{\lambda}^{(kk)}_n = \hat{\sigma}^{(kk)}_n \hat{u}^{(kk)}_n\). Then
+\(\hat{\lambda}^{(kk)}_n\) is the adjacency spectral embedding
+approximation of \(\lambda^{(kk)}_n\).\\
 Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the
-adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-\(\lambda^{(kk)}\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
-
-\emph{Case \(k \neq l\)}. \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
+adjacency spectral embedding \(\hat{\lambda}^{(kk)}_n\) approximates
+\(\lambda^{(kk)}_n\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
+
+\emph{Case \(k \neq l\)}. \(P^{(kl)}_n\) and \(A^{(kl)}_n\) represent edge
 probabilities and edges between communities \(k\) and \(l\). Note that
@@
 
-which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. It can
+which is a symmetric \(n_{kl} \times n_{kl}\) matrix. It can
 be shown that the spectral decomposition of \(M^{(kl)}\) is
@@
 \[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
-Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
-
-\[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
-
-and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
-generate the Bernoulli entries in \(A^{(kl)}\).\\
 Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
@@
 approximates the latent positions of \(M^{(kl)}\) up to indefinite
-orthogonal transformation by the rate given in Theorem 5 of
-\citeauthor{rubindelanchy2017statistical}.\\
+orthogonal transformation with rate $\frac{(\log n_{kl})^c}{\sqrt{n_{kl}}}$
+\cite{rubindelanchy2017statistical}.\\
 In this case, the indefinite orthogonal transformation \(W_*\) in the

### 2021-06-16 summary.tex
minor updates

@@
 probabilities and edges between communities \(k\) and \(l\). Note that
-\(P^{(kl)} = (P^{(lk)})^\top\).\\
-By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). As
-in the \(k = l\) case, we note that the singular value decomposition
-\(P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top\) is one-dimensional
-and \(\lambda^{(kl)} = \sigma_{kl} u^{(kl)}\). (We can also note that
-the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
-\(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
-\(v^{(kl)} = u^{(lk)}\).)\\
+\(P^{(kl)}_n = (P^{(lk)}_n)^\top\).
+By definition, \(P^{(kl)}_n = \lambda^{(kl)}_n (\lambda^{(lk)}_n)^\top\).
+Therefore,
+\(P^{(kl)} = (\sigma^{(kl)}_n)^2 u^{(kl)}_n (v^{(kl)}_n)^\top\) is rank 2
+
 Now consider the Hermitian dilation
 
-\[M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}\]
+\[M^{(kl)}_n = 2 \begin{bmatrix}
+0 & P^{(kl)}_n \\ P^{(lk)}_n & 0 \end{bmatrix}\]
 
 which is a symmetric \(n_{kl} \times n_{kl}\) matrix. It can
-be shown that the spectral decomposition of \(M^{(kl)}\) is
-
-\[M^{(kl)} =
+be shown that the spectral decomposition of \(M^{(kl)}_n\) is
+
+\[M^{(kl)}_n =
 \begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times
@@
 \[\begin{bmatrix}
-  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\
-  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)}
-\end{bmatrix} =
-\begin{bmatrix}
-  \lambda^{(kl)} & \lambda^{(kl)} \\
-  \lambda^{(lk)} & -\lambda^{(lk)}
+  \lambda^{(kl)}_n & \lambda^{(kl)}_n \\
+  \lambda^{(lk)}_n & -\lambda^{(lk)}_n
 \end{bmatrix}\]
@@
 
-\[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
-Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
+\[\hat{M}^{(kl)}_n = \begin{bmatrix} 0 & A^{(kl)}_n \\ A^{(lk)}_n & 0 \end{bmatrix}\]
+Then \(\hat{M}^{(kl)}_n\) is an adjacency matrix drawn from \(M^{(kl)}_n\),
 so its adjacency spectral embedding, given by

### 2021-06-23 summary.tex
updated document

--- 2021-06-16 summary.tex
@@
 up to multiplicative constant using singular value
-decomposition. More specifically, let \(P^{(kl)} = \sigma^2 u v^\top\)
+decomposition. More specifically, 
+let \(P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top\)
 be the singular value decomposition of \(P^{(kl)}\).
-\(u \in \mathbb{R}^{n_k}\) and \(v \in \mathbb{R}^{n_l}\) are vectors
-and \(\sigma^2 > 0\) is a scalar. Then \(\lambda^{(kl)} = s_1 u\)
-and \(\lambda^{(lk)} = s_2 v\) where $s_1 s_2 = \sigma^2$.
+\(u^{(kl)} \in \mathbb{R}^{n_k}\) and 
+\(v^{(kl)} \in \mathbb{R}^{n_l}\) are vectors
+and \(\sigma^{(kl)}\) is a scalar. Then \(\lambda^{(kl)} = s_1 u^{(kl)}\)
+and \(\lambda^{(lk)} = s_2 v^{(kl)}\) 
+for unidentifiable $s_1 s_2 = (\sigma^{(kl)})^2$.
 Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators, which converge to the true parameters, again, up to a
-multiplicative constant. Since these popularity vectors aren't identifiable,
-due to the ambiguity of this constant,
-we instead estimate each edge probability block
-$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$
-using the observed edge block $A^{(kl)}$ (Algorithm 3).
-Alternatively, we can estimate the entire edge probability matrix as
-$\hat{P} = \hat{X} I_{p,q} \hat{X}^\top$, where $\hat{X}$ is the ASE of $A$.
+estimators by taking the SVD of each $A^{(kl)}$. 
+Since each $\lambda^{(kl)}$ is not strictly identifiable,
+we instead estimate each $\tilde{\lambda}^{(kl)} = \sigma^{(kl)} u^{(kl)}$.
 
@@
   \SetAlgoLined
-  \caption{PABM block probability estimation.}
-  \KwData{Adjacency block $A^{(kl)}$}
-  \KwResult{Estimated edge probability block $\hat{P}^{(kl)}$}
-  Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ block.\;
+  \caption{PABM parameter estimation.}
+  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
+  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
+  Arrange the rows and columns of $A$ by community such that each 
+  $A^{(kl)}$ block consists of estimated edge probabilities between 
+  communities $k$ and $l$.\;
+  \For {$k, l = 1, ..., K$, $k \leq l$} {
+    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+    block.\;
     Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
@@
     Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
-  $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.\;
-  Assign $\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$.
+    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
+  }
 \end{algorithm}
@@
 \label{theorem6}
-Let $n_{kl} = n_k + n_l$.
 Under regularity and sparsity assumptions, given fixed $K$,
@@
 \begin{equation} \label{eq:thm6}
-\max_{k, l} \|\hat{P}_n^{(kl)} - P_n^{(kl)}\| = O_P \bigg(\frac{(\log n_{kl})}{\sqrt{n_{kl}}} \bigg)
+\max_{k, l \in \{1, ..., K\}} 
+||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
+O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
 \end{equation}
@@
 
-We also define parameter estimation error as the RMSE
-of the estimated edge probability blocks:
-
-\[RMSE =
+We also define two types of parameter estimation error. 
+First, we estimate the popularity vectors directly and compute the RMSE. 
+In this case, the "true" popularity vectors are derived from 
+taking the SVD of each edge probability block $P^{(kl)}$ 
+to avoid the unidentifiable multiplicative constants.
+
+\[\text{RMSE}(\{\hat{\lambda}^{(kl)}\}_K, \{\lambda^{(kl)}\}_K) = 
+\sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
+\|s \hat{\lambda}^{(kl)} - \lambda^{(kl)} \|_2^2}\]
+
+We can also avoid the unidentifiable multiplicative constant more directly 
+by reconstructing each 
+$\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$, 
+which we use to define another parameter estimation error.
+
+\[\text{RMSE}(\hat{P}, P) =
 \sum_{k,l} \sqrt{\frac{1}{n_k n_l} \|P^{(kl)} - \hat{P}^{(kl)}\|_F^2}\]
@@
 \(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
-\(A_{ij} \stackrel{indep}{\sim} \Bernoulli(P_{ij})\). Each simulation has
+\(A_{ij} \indep \Bernoulli(P_{ij})\). Each simulation has
 a unique edge probability matrix \(P\).
@@
 Given ground truth community labels, Algorithm 3 and the MLE-based
-plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1}
-\cite{noroozi2019estimation} perform similarly, with root mean square
+plug-in estimators \cite{noroozi2019estimation, 307cbeb9b1be48299388437423d94bf1} 
+perform similarly, with root mean square
 error decaying at rate approximately \(n^{-1/2}\).
@@
 
-\emph{Case \(k = l\)}. \(P^{(kk)}_n\) and \(A^{(kk)}_n\) represent
+\emph{Case \(k = l\)}. \(P^{(kk)}\) and \(A^{(kk)}\) represent
 within-community edge probabilities and edges for community \(k\).\\
-By definition, \(P^{(kk)}_n = \lambda^{(kk)}_n (\lambda^{(kk)}_n)^\top\). This
+By definition, \(P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top\). This
 implies that the singular value decomposition
-\(P^{(kk)}_n = (\sigma^{(kk)}_n)^2 u^{(kk)}_n (u^{(kk)}_n)^\top\)
-has one singular
-value and one pair of singular vectors (\(P^{(kk)}_n\) is symmetric, so
+\(P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top\) has one singular
+value and one pair of singular vectors (\(P^{(kk)}\) is symmetric, so
 the left and right singular vectors are identical). Then
-\(\lambda^{(kk)}_n = \sigma^{(kk)}_n u^{(kk)}_n\).\\
-Let \(\hat{U}^{(kk)}_n \hat{\Sigma}^{(kk)}_n (\hat{U}^{(kk)}_n)^\top\) be the
-singular value decomposition of \(A^{(kk)}_n\), and let
-\((\hat{\sigma}^{(kk)})^2 \hat{u}^{(kk)}_n (\hat{u}^{(kk)}_n)^\top\) be its
+\(\lambda^{(kk)} = \sigma_{kk} u^{(kk)}\).\\
+Let \(\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top\) be the
+singular value decomposition of \(A^{(kk)}\), and let
+\(\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top\) be its
 one-dimensional approximation. Define
-\(\hat{\lambda}^{(kk)}_n = \hat{\sigma}^{(kk)}_n \hat{u}^{(kk)}_n\). Then
-\(\hat{\lambda}^{(kk)}_n\) is the adjacency spectral embedding
-approximation of \(\lambda^{(kk)}_n\).\\
+\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
+\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding
+approximation of \(\lambda^{(kk)}\).\\
 Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the
-adjacency spectral embedding \(\hat{\lambda}^{(kk)}_n\) approximates
-\(\lambda^{(kk)}_n\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
-
-\emph{Case \(k \neq l\)}. \(P^{(kl)}_n\) and \(A^{(kl)}_n\) represent edge
+adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
+\(\lambda^{(kk)}\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
+
+\emph{Case \(k \neq l\)}. \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
 probabilities and edges between communities \(k\) and \(l\). Note that
@@
 
-which is a symmetric \(n_{kl} \times n_{kl}\) matrix. It can
+which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. It can
 be shown that the spectral decomposition of \(M^{(kl)}\) is
@@
 \[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
+Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
+
+\[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
+
+and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
+generate the Bernoulli entries in \(A^{(kl)}\).\\
 Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
@@
 approximates the latent positions of \(M^{(kl)}\) up to indefinite
-orthogonal transformation with rate $\frac{(\log n_{kl})^c}{\sqrt{n_{kl}}}$
-\cite{rubindelanchy2017statistical}.\\
+orthogonal transformation by the rate given in Theorem 5 of
+\citeauthor{rubindelanchy2017statistical}.\\
 In this case, the indefinite orthogonal transformation \(W_*\) in the

### 2021-06-23 summary.tex
updated document

@@
 Given ground truth community labels, Algorithm 3 and the MLE-based
-plug-in estimators \cite{noroozi2019estimation, 307cbeb9b1be48299388437423d94bf1} 
+plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1} 
 perform similarly, with root mean square

### 2021-06-23 summary.tex
updated document

@@
 
-\caption{Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+\caption{Median and IQR RMSE for popularity vectors from Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.
 Communities were drawn to be approximately balanced.}\label{fig:lambda_est_k}
+\end{figure}
+
+We observe similar behavior when comparing errors in the edge probability blocks.
+
+\begin{figure}[H]
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_p-1}}
+\caption{Median and IQR RMSE for edge probability blocks reconstructed from the outputs of Algorithm 3 (red) compared against outputs of an MLE-based method (blue).
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be approximately balanced.}
 \end{figure}
@@
 for SSC when \(K > 2\) (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3
-continues to see \(n^{-1/2}\) decay in parameter estimation error
-(\ref{fig:lambda_est_k_imba}).
+continues to see \(n^{-1/2}\) decay in parameter estimation error; 
+however, the MLE-based estimators for the popularity vectors decay at a slower 
+rate (Fig. \ref{fig:lambda_est_k_imba}). 
+The reduced rate of decay is removed 
+if we reconstruct the edge probability blocks 
+by taking the outer products of the popularity vectors.
 
@@
 Communities were drawn to be imbalanced.}\label{fig:lambda_est_k_imba}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_p_imba-1}
+
+}
+
+\caption{Median and IQR RMSE of edge probability blocks derived from the outputs of Algorithm 3 (red) compared against an MLE-based method (blue).
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be imbalanced.}\label{fig:lambda_est_p_imba}
 \end{figure}

### 2021-06-30 summary.tex
updated document

@@
 estimation being of possible interest, assuming some underlying
-probability model. One such model is the Stochastic Block Model (SBM),
-first proposed by \citet{doi:10.1080/0022250X.1971.9989788}, which
-assumes that the edge probability from one vertex to another follows a
-Bernoulli distribution with fixed probabilities for each pair of
-community labels. Other random graph models have been proposed and
-studied, such as the Degree-Corrected Block Model (DCBM), introduced by
-\citet{Karrer_2011}, which is a generalization of the SBM. The
-Popularity Adjusted Block Model (PABM) was then introduced by
+probability model.
+
+A general family of models that generate unweighted and undirected graphs is the Bernoulli Graph, which is described by an edge probability matrix $P$.
+Edges between each pair of vertices $i$ and $j$ is drawn as a Bernoulli trial with probability $P_{ij}$, for each $0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph.
+Community detection methods often restrict this type of model by assigning each vertex a community label $z_1, ..., z_n$ and then setting each $P_{ij}$ as a function of the labels of vertices $i$ and $j$.
+Such models are called Block Models.
+
+One popular type of Block Model is the Stochastic Block Model (SBM),
+first proposed by \citet{doi:10.1080/0022250X.1971.9989788}, 
+which restricts each edge probability $P_{ij}$ as only depending on the labels 
+$z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_i, z_j}$. 
+Various generalizations of the SBM have since been utilized, 
+including the Degree-Corrected Block Model (DCBM), 
+introduced by\citet{Karrer_2011}, 
+which assigns an additional parameter $\theta_i$ to each vertex 
+and sets $P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}$. 
+The Popularity Adjusted Block Model (PABM) was then introduced by
 \citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the
@@
 
-The underlying similarity among the SBM, PABM, and other such models is
-that they involve a symmetric edge probability matrix
-\(P \in [0, 1]^{n \times n}\) where \(n\) is the number of vertices in
-the graph. An undirected and unweighted graph is then drawn from this
-edge probability matrix such that the existence of an edge between each
-pair of vertices \(i\) and \(j\) is given by
-\(\text{Bernoulli}(P_{ij})\). For example, for the SBM with two
-communities for which the within-community edge probability is \(\xi\)
-and the between-community edge probability is \(\eta\), the entries of
-\(P\) consist of \(P_{ij} = \xi\) if \(i\) and \(j\) are in the same
-community and \(i \neq j\), \(P_{ij} = \eta\) if \(i\) and \(j\) belong
-to separate communities, and \(P_{ii} = 0\).
-
-The Random Dot Product Graph (RDPG) model proposed by
-\citet*{10.1007/978-3-540-77004-6_11} is another graph model with
-Bernoulli edge probabilities. Under this model, each vertex of the graph
+Parallel to the families of Block Models is another type of Bernoulli Graph 
+called the Random Dot Product Graph (RDPG), proposed by
+\citet*{10.1007/978-3-540-77004-6_11}. Under this model, each vertex of the graph
 can be represented by a point in some latent space such that the edge
@@
 \(P = X X^\top\) where 
-\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The assortative
+\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). 
+It can be shown that all Bernoulli Graphs with 
+positive semidefinite edge probability matrix $P$ is a type of RDPG. 
+For instance, the assortative
 SBM is equivalent to a special case of the RDPG for which all vertices
 of a given community share the same position in the latent space
-\cite{lyzinski2014}. It has also been shown that similar random graph
-models, including the DCBM, can be represented in a similar fashion
-\cite{lyzinski2014, rubindelanchy2017consistency}. An analogous
-property exists for the PABM, not for the RDPG model but the
-\emph{Generalized} Random Dot Product Graph (GRDPG) model, which is an
-extension of the RDPG to allow edge probability matrices that are not positive
-semidefinite. This relationship will be explored in this paper and exploited to
-construct algorithms for community detection and parameter estimation for the
-PABM.
+\cite{lyzinski2014}. A similar property exists for the DCBM
+\cite{lyzinski2014, rubindelanchy2017consistency}. 
+Because the edge probability matrix for the PABM is not positive semidefinite, 
+we instead show that it is a special case of the \emph{Generalized} Random Dot Product Graph (GRDPG) in order to analyze the PABM as a latent space model for community detection and parameter estimation.
 
@@
 method is compared against SSC \cite{noroozi2019estimation,soltanolkotabi2014}
-and MM \cite{igraph}
-\cite{307cbeb9b1be48299388437423d94bf1}. True community labels are used
+and MM \cite{igraph, 307cbeb9b1be48299388437423d94bf1}. 
+True community labels are used
 with Algorithm 3 to estimate the popularity vectors
@@
 \cite{10.1145/1134271.1134277}, and the DBLP network
-\cite{NIPS2009_3855} \cite{10.1007/978-3-642-15880-3_42}. For this data
+\cite{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
 analysis, we subsetted the data as described by

### 2021-06-30 summary.tex
Update paper

--- 2021-06-23 summary.tex
@@
 \newcommand{\Betadist}{\text{Beta}}
-\newcommand{\BG}{\text{BernoulliGraph}}
-\newcommand{\PABM}{\text{PABM}}
-\newcommand{\RDPG}{\text{RDPG}}
-\newcommand{\GRDPG}{\text{GRDPG}}
-\newcommand{\Multinomial}{\text{Multinomial}}
 \newtheorem{definition}{Definition}
@@
 probability distribution based on the community relationships between
-each pair. The goal of this analysis then is population community
+each pair. The goal of the analysis then is population community
 detection, or the recovery of the true underlying community labels for
-each vertex, up to permutation, with some additional parameter
-estimation being of possible interest, assuming some underlying
+each vertex, up to permutation (with some additional parameter
+estimation being of possible interest), assuming some underlying
 probability model. One such model is the Stochastic Block Model (SBM),
@@
 assumes that the edge probability from one vertex to another follows a
-Bernoulli distribution with fixed probabilities for each pair of
+Beronulli distribution with fixed probabilities for each pair of
 community labels. Other random graph models have been proposed and
@@
 \(P = X X^\top\) where
-\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The assortative
-SBM is equivalent to a special case of the RDPG for which all vertices
+\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). The SBM
+is equivalent to a special case of the RDPG model in which all vertices
 of a given community share the same position in the latent space
 \cite{lyzinski2014}. It has also been shown that similar random graph
-models, including the DCBM, can be represented in a similar fashion
-\cite{lyzinski2014, rubindelanchy2017consistency}. An analogous
-property exists for the PABM, not for the RDPG model but the
-\emph{Generalized} Random Dot Product Graph (GRDPG) model, which is an
-extension of the RDPG to allow edge probability matrices that are not positive
-semidefinite. This relationship will be explored in this paper and exploited to
-construct algorithms for community detection and parameter estimation for the
+models, including the DCBM, can be represented in this way
+\cite{lyzinski2014} \cite{rubindelanchy2017consistency}. An analogous
+property exists for the PABM but not for the RDPG model but under the
+\emph{Generalized} Random Dot Product Graph (GRDPG) model. This
+relationship will be explored in this paper and exploited to construct
+algorithms for community detection and parameter estimation for the
 PABM.
 
-\hypertarget{notation}{%
-\subsection{Notation and Scope}\label{notation}}
-
-Let $G = (V, E)$ be an unweighted, undirected, and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$.
-$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ and $0$ otherwise.
-Because $G$ is symmetric and hollow,
-$A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$.
-We further restrict our analyses to Bernoulli graphs.
-Let $P \in [0, 1]^{n \times n}$ be a symmetric matrix of edge probabilities.
-Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ for each $1 \leq i < j \leq n$  (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$).
-We denote $A \sim \BG(P)$ as graph with adjacency matrix $A$ sampled from edge probability matrix $P$ in this manner.
-If each vertex has a hidden label in $[K]$, they are denoted as $z_1, ..., z_n$.
-Finally, we denote $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
+In this paper, we will only consider undirected graphs, that is the edge
+weight from vertex \(i\) to vertex \(j\) is equal to the edge weight in
+the opposite direction, from vertex \(j\) to vertex \(i\). Furthermore,
+we will only consider unweighted graphs with binary (\(0\) or \(1\))
+edge weights We will also assume that graphs are hollow, i.e., there are
+no edges from a vertex to itself. All such graphs can be represented by
+a symmetric adjacency matrix \(A \in \{0, 1\}^{n \times n}\) for which
+\(A_{ij} = 1\) if there exists an edge between vertices \(i\) and \(j\)
+and \(0\) otherwise, and \(A\) is an element-wise independent Bernoulli
+draw from a symmetric edge probability matrix
+\(P \in [0, 1]^{n \times n}\).
 
@@
 
-In this section, we show that the PABM is a special case of the GRDPG, i.e.,
-graph $G$ drawn from the PABM can be represented by a collection of latent
-vectors in Euclidean space. We further show that the latent configuration that
-induces the PABM consists of orthogonal subspaces with each subspace
-corresponding to a community.
-
 \hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
-\subsection{The Popularity Adjusted Block Model and the Generalized
-Random Dot Product
-Graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
+\subsection{The popularity adjusted block model and the generalized
+random dot product
+graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
 
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
-graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \BG(P)$.
-Let each vertex have a community label between $1$ and $K$.
-Then $G$ is drawn from a Popularity Adjusted Block Model if
-each vertex has $K$ popularity parameters that describe its affinity
-toward each of the $K$ communities, i.e., vertex $i$ has popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK}$, and
-each $P_ij = \lambda_{i z_j} \lambda_{j z_i}$.
-
-Another characterization of the PABM is as follows.
-Let the rows and columns of $P$ be arranged by community label
-such that $n_k \times n_l$ block $P^{(kl)}$
-describes the edge probabilities between vertices in communities
+set of $n$ 
+vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
+columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
+$P^{(kl)}$ describes the edge probabilities between vertices in communities 
 $k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
+Let graph $G = (V, E)$ be an undirected, unweighted graph such 
+that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
+realization of $\text{Bernoulli}(P)$, i.e., 
+$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ 
+($A_{ij} = A_{ji}$ and $A_{ii} = 0$).
+
 If each block $P^{(kl)}$ can be written as the outer product of two vectors:
@@
 
-for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each
-$\lambda^{(st)}$ is a column vector of dimension $n_s$, then graph $G$ is drawn from a PABM with parameters $\{\lambda^{(st)}\}_K$ if its its corresponding adjacency matrix $A \sim \BG(P)$.
+for a set of $K^2$ fixed vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each 
+$\lambda^{(st)}$ is a column vector 
+of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
+is a realization of a popularity adjusted block model with parameters 
+$\{\lambda^{(st)}\}_{s, t = 1}^K$. 
 \end{definition}
 
-We will use the notation \(A \sim \PABM(\{\lambda^{(kl)}\}_K)\) to denote
+We will use the notation \(A \sim PABM(\{\lambda^{(kl)}\}_K)\) to denote
 a random adjacency matrix \(A\) drawn from a PABM with parameters
@@
 \label{grdpg} 
-Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$.
-If $\exists X \in \mathbb{R}^{n \times d}$ such that
+Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
+set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that
 
 \begin{equation} \label{eq:grdpg}
-  P = X I_{p,q} X^\top
+  P = X I_{pq} X^\top
 \end{equation}
@@
 for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
-$G$ is drawn from the Generalized Random Dot Product Graph with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
+graph $G = (V, E)$ with adjacency matrix $A$ such that 
+$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
+$A_{ii} = 0$) is a draw from the generalized random dot product graph model 
+with latent positions $X$ and signature $(p, q)$. More precisely, if vertices 
+$i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge 
+probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains 
+the latent positions as rows $x_i^\top$. 
 \end{definition}
 
-We will use the notation \(A \sim \GRDPG_{p,q}(X)\) to denote a random
+We will use the notation \(A \sim GRDPG_{p,q}(X)\) to denote a random
 adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\). If instead of fixed latent positions, they are drawn $X_1, ..., X_n \iid F$, we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
+\((p, q)\).
 
@@
 The indefinite orthogonal group with signature $(p, q)$ is 
-the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{p, q} Q^{\top} = I_{p, q}\}$,
+the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
 denoted as $\mathbb{O}(p, q)$.
@@
 \cite{rubindelanchy2017statistical}. 
-More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, then we also have for any
+More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
 $Q \in \mathbb{O}(p, q)$, 
-$(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
-x_i^\top I_{p, q} x_j = P_{ij}$.
+$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
+x_i^\top I_{pq} x_j = P_{ij}$. 
 Unlike in the RDPG case, transforming the latent positions via multiplication 
@@
 We previously stated one possible set of latent positions that result in
-the edge probability matrix of a PABM, \(P = (XU) I_{p, q} (XU)^\top\). If
+the edge probability matrix of a PABM, \(P = (XU) I_{pq} (XU)^\top\). If
 we have (or can estimate) \(XU\) directly, then both the community
@@
 
-If we decompose \(P = Z I_{p, q} Z^\top\), then
+If we decompose \(P = Z I_{pq} Z^\top\), then
 \(\exists Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
@@
 \(P^{(kl)}\), \(\lambda^{(kl)}\) and \(\lambda^{(lk)}\) are solvable
-up to multiplicative constant using singular value
-decomposition. More specifically, 
-let \(P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top\)
+exactly (up to multiplication by \(-1\)) using singular value
+decomposition. More specifically, let \(P^{(kl)} = \sigma^2 u v^\top\)
 be the singular value decomposition of \(P^{(kl)}\).
-\(u^{(kl)} \in \mathbb{R}^{n_k}\) and 
-\(v^{(kl)} \in \mathbb{R}^{n_l}\) are vectors
-and \(\sigma^{(kl)}\) is a scalar. Then \(\lambda^{(kl)} = s_1 u^{(kl)}\)
-and \(\lambda^{(lk)} = s_2 v^{(kl)}\) 
-for unidentifiable $s_1 s_2 = (\sigma^{(kl)})^2$.
-Given the adjacency matrix \(A\)
+\(u \in \mathbb{R}^{n_k}\) and \(v \in \mathbb{R}^{n_l}\) are vectors
+and \(\sigma^2 > 0\) is a scalar. Then \(\lambda^{(kl)} = \pm \sigma u\)
+and \(\lambda^{(lk)} = \pm \sigma v\). Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators by taking the SVD of each $A^{(kl)}$. 
-Since each $\lambda^{(kl)}$ is not strictly identifiable,
-we instead estimate each $\tilde{\lambda}^{(kl)} = \sigma^{(kl)} u^{(kl)}$.
+estimators (algorithm 3), which converge to the true parameters.
 
@@
 is drawn from \(P\). OSC is then used for community detection, and this
-method is compared against SSC \cite{noroozi2019estimation,soltanolkotabi2014}
-and MM \cite{igraph}
+method is compared against SSC \cite{noroozi2019estimation}
+\cite{soltanolkotabi2014} and MM \cite{igraph}
 \cite{307cbeb9b1be48299388437423d94bf1}. True community labels are used
@@
 \(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
-MLE-based estimator described by \citeauthor{noroozi2019estimation} and
+MLE-based estimator described in \citeauthor{noroozi2019estimation} and
 \citeauthor{307cbeb9b1be48299388437423d94bf1}.
@@
 relaxation of the optimization problem. For these simulations, the
-Louvain algorithm was used for modularity maximization,
-as \citeauthor{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
+Louvain algorithm was used, as the EP algorithm proved to be
 prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
-was verified that the Louvain algorithm produces comparable results
-to EP-MM.
+was verified that EP and Louvain produce comparable results.
 
@@
 
-We also define two types of parameter estimation error. 
-First, we estimate the popularity vectors directly and compute the RMSE. 
-In this case, the "true" popularity vectors are derived from 
-taking the SVD of each edge probability block $P^{(kl)}$ 
-to avoid the unidentifiable multiplicative constants.
-
-\[\text{RMSE}(\{\hat{\lambda}^{(kl)}\}_K, \{\lambda^{(kl)}\}_K) = 
-\sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(kl)} - \lambda^{(kl)} \|_2^2}\]
-
-We can also avoid the unidentifiable multiplicative constant more directly 
-by reconstructing each 
-$\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$, 
-which we use to define another parameter estimation error.
-
-\[\text{RMSE}(\hat{P}, P) =
-\sum_{k,l} \sqrt{\frac{1}{n_k n_l} \|P^{(kl)} - \hat{P}^{(kl)}\|_F^2}\]
+We also define parameter estimation error as the RMSE up to sign flip:
+
+\[RMSE(\{\hat{\lambda}^{(kl)}\}, \{\lambda^{(kl)}\}) = 
+\sqrt{\frac{1}{n} \sum_{k < l} \min_{s \in \{-1, 1\}} 
+\sum_{i = 1}^{n_k} 
+(s \hat{\lambda}^{(kl)}_i - \lambda^{(kl)}_i)^2}\]
 
@@
 \(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
-\(A_{ij} \indep \Bernoulli(P_{ij})\). Each simulation has
+\(A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})\). Each simulation has
 a unique edge probability matrix \(P\).
@@
   Community labels
-  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
+  \(z_k \stackrel{\text{iid}}{\sim}Multinomial(\alpha_1, ..., \alpha_K)\)
 \item
   Within-group popularities
-  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
+  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim}Beta(2, 1)\)
 \item
   Between-group popularities
-  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
+  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim}Beta(1, 2)\) for
   \(k \neq l\)
@@
 plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1}
-perform similarly, with root mean square
+\cite{noroozi2019estimation} perform similarly, with root mean square
 error decaying at rate approximately \(n^{-1/2}\).
@@
 matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in (\ref{eq:xy}).
+defined as in Eq.\eqref{eq:xy}.
 \end{lemma}
@@
 \emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
-\(X\) is defined as in (\ref{eq:xy}) and $p = K(K+1)/2$ and $q = K(K-1)/2$. % are given by $p =
+\(X\) is defined as in Eq.~\eqref{eq:xy}), $p = K(K+1)/2$, and $q = K(K-1)/2$. % are given by $p =
 % K(K+1)/2$ and $q = are defined as
 % in equations (\ref{eq:p}) and (\ref{eq:q}). 
-Alternatively, the spectral
-decomposition can be written as
-\(P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) for the same
-\((p, q)\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
+Alternatively we have the eigendecomposition \(P = V D V^\top = V
+|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
 \(Q \in \mathbb{O}(p, q)\),
-
-\[X U Q = V |D|^{1/2}\]
-
-Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
-
-\[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top =
-(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top\]
-
-The right-hand side becomes
-
-\[\begin{split}
-  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top &
-    = X U Q Q^{-1} U^\top (X^\top X)^{-1}
-    U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-  & = X (X^\top X)^{-1} X^\top
-\end{split}\]
-
-The left-hand side becomes:
-
-\[\begin{split}
-  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-  & = V V^\top
-\end{split}\]
+\[X U Q = V |D|^{1/2}.\]
+Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence
+$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
+R^{\top} X^{\top}.$$ 
+Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
+V = I$. We thus conclude
+$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
+as desired.
+% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
+% we have
+% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
+% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
+% The right-hand side becomes
+% \[\begin{split}
+%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
+%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
+%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
+%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
+%   & = X (X^\top X)^{-1} X^\top.
+% \end{split}\]
+% The left-hand side becomes:
+% \[\begin{split}
+%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
+%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
+%   & = V V^\top.
+% \end{split}\]
 
@@
 block diagonal matrix with each block corresponding to a community and
-zeros elsewhere. Therefore, if vertices \(i\) and \(j\) belong to
-different communities, then the \(ij\)\textsuperscript{th} element of
-\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
+zeros elsewhere. Let $B = n VV^{\top}$. Then $B_{ij} = 0$ whenever
+vertices $i$ and $j$ belong to different communities.  
+%
+%then the \(ij\)\textsuperscript{th} element of
+%\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
 
 \textbf{Proof of Theorem \ref{theorem4}}. Let \(V_n\) and \(\hat{V}_n\)
-be the eigenvectors of \(P\) and \(A\) corresponding to the
+be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
 \(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
-eigenvalues. By \citeauthor{rubindelanchy2017statistical}, for some
-\(W \in \mathbb{O}(K^2)\), and \(c > 0\),\\
-\(||\hat{V} W - V||_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big)\).
-We furthermore have \(||V||_{2 \to \infty} = O_P(n^{-1/2})\). Then if
-\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) correspond to the
-rows of \(V_n\) and \(\hat{V}_n\), for \(i\) and \(j\) in different
-communities, using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
-
+eigenvalues, respectively. From Lemma~5 in
+\citeauthor{rubindelanchy2017statistical} we have, for some
+\(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
+$$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big).$$
+We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). Let
+\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th row
+row of \(V_n\) and \(\hat{V}_n\), respectively. Recall
+Theorem~\ref{theorem3} and note that $B_{ij} = n(v_n^{(i)})^{\top}
+v_n^{(j)}$. Now suppose that vertices $i$ and $j$ belongs to different
+communities. Then $B_{ij} = 0$ and we have
+% Then for \(i\) and
+% \(j\) belonging to different
+% communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
 \[\begin{split}
@@
 (v_n^{(i)})^\top v_n^{(j)}| \\
-& = ||\hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top +
-\hat{V}_n W W^\top \hat{V}_n^\top -
-2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
-& = ||2 (\hat{V}_n W - V_n) V_n^\top +
-(\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
-& \leq 2 ||\hat{V}_n W - V_n||_{2 \to \infty} ||V_n||_{2 \to \infty} +
-||\hat{V}_n W - V_n||_{2 \to \infty}^2 \\
+& \leq \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|\bigr)
++ \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
+% & \leq 
+% \hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
+% & = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
+% 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
+% & = ||2 (\hat{V}_n W - V_n) V_n^\top + 
+% (\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
+% & \leq 2 
+&\leq \|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|\hat{V}_n\|_{2 \to \infty} +
+\|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty} \\
+& \leq \|\hat{V}_n W - V_n\|_{2 \to \infty}^2 + 2 \|\hat{V}_n W -
+V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty}
+\\
 & = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
-\end{split}\]
-
-Then scaling by \(n\), we get
-\(|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \big)\).
+\end{split}.\]
+Scaling all terms in the above inequality by \(n\), we obtain
+$$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
+as desired.
+
+We now provide a proof of Theorem~\ref{theorem5}. We first recall several
+useful definitions for analyzing sparse subspace clustering.
 
@@
 Define the subspace incoherence:
-
 $$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
-
-Then $\forall k, n$,
-
+Then $\forall k$, we have
 \begin{equation} \label{eq:mu-conv}
@@
 
-\emph{Proof}. Since by Theorem \ref{theorem2} each \(\mathcal{S}^{(k)}\)
-are mutually orthogonal, any vector projected to \(\mathcal{S}^{(k)}\)
-must be orthogonal to each row of \(V^{(-k)}\). Therefore,
-\(W^{(k)} v = 0\) \(\forall v \in \mathcal{S}^{(-k)}\).
+\begin{proof} From Theorem \ref{theorem3} we have that the subspaces
+  $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
+  are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
+  \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
+  \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
+  $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
+    $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
+    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
+    \in V^{(-k)}$ and hence $\mu_n^{(k)} = 0$ as desired.
+%   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
+% must be orthogonal to each row of \(V^{(-k)}\) and hence
+%\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
+\end{proof}
 
@@
 \begin{remark}
-Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $||v_n^{(i)}|| = 1$,
+Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $\|v_n^{(i)}\| = 1$, 
 which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for 
@@
 
-\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) and \(A\) be
-organized by community such that the elements of blocks \(P^{(kl)}\) and
-\(A^{(kl)}\) correspond to the edges between communities \(k\) and
-\(l\).
-
-\emph{Case \(k = l\)}. \(P^{(kk)}\) and \(A^{(kk)}\) represent
-within-community edge probabilities and edges for community \(k\).\\
-By definition, \(P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top\). This
-implies that the singular value decomposition
-\(P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top\) has one singular
-value and one pair of singular vectors (\(P^{(kk)}\) is symmetric, so
-the left and right singular vectors are identical). Then
-\(\lambda^{(kk)} = \sigma_{kk} u^{(kk)}\).\\
-Let \(\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top\) be the
+\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) be
+organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
+n_{\ell}$ matrix obtained by keeping only the rows of $P$
+corresponding to vertices in community $k$ and the columns of $P$
+corresponding to vertices in community $\ell$. We define $A^{(k
+  \ell)}$ analogously. Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
+for the cases when $k = \ell$ versus when $k \not = \ell$.
+
+\emph{Case \(k = l\)}. let $P^{(kk)} =
+\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
+decomposition of $P^{(kk)}$. We can then define
+$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
+Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
 singular value decomposition of \(A^{(kk)}\), and let
-\(\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top\) be its
-one-dimensional approximation. Define
+$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
+best rank-one approximation of $A^{(kk)}$. Define
 \(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
-\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding
-approximation of \(\lambda^{(kk)}\).\\
-Then by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, the
-adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-\(\lambda^{(kk)}\) at rate \(\frac{(\log n_k)^c}{\sqrt{n_k}}\).
+\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
+and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
+%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
+$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
+Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.
 

### 2021-06-30 summary.tex
typo

--- 2021-06-30 summary.tex
@@
 $\lambda_{i1}, ..., \lambda_{iK}$, and
-each $P_ij = \lambda_{i z_j} \lambda_{j z_i}$.
+each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
 

### 2021-06-30 summary.Rmd

@@
 
+```{r lambda_est_p, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be approximately balanced.', fig.width = 10}
+rmse.df <- readr::read_csv('rmse-p.csv')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 3'), width = .1) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
+                   colour = 'MLE-based'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(y = 'RMSE', colour = NULL) + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10() + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
+```
+
 ## Imbalanced communities
@@
 
+```{r lambda_est_p_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+Simulations were repeated 50 times for each sample size.
+Communities were drawn to be imbalanced.', fig.width = 10}
+rmse.df <- readr::read_csv('rmse-p-imbalanced.csv')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 3'), width = .1) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
+                   colour = 'MLE-based'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(y = 'RMSE', colour = NULL) + 
+  facet_wrap(~ K, labeller = 'label_both') + 
+  scale_y_log10() + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
+```
+
 \begin{comment}

### 2021-06-30 summary.tex

--- 2021-06-30 summary.tex
@@
 Define the subspace incoherence:
-
 $$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
-
-Then $\forall k, n$,
+Then $\forall k$, we have
 

### 2021-07-02 summary.tex
Update git repo

--- 2021-06-30 summary.tex
@@
 \newcommand{\Multinomial}{\text{Multinomial}}
-\newtheorem{definition}{Definition}
 \newtheorem{theorem}{Theorem}
 \newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
 \theoremstyle{remark}
@@
 \theoremstyle{definition}
+\newtheorem{definition}{Definition}
 \newtheorem*{example}{Example}
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \BG(P)$.
+graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \mathrm{Bernoulli}(P)$.
 Let each vertex have a community label between $1$ and $K$.
@@
 toward each of the $K$ communities, i.e., vertex $i$ has popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK}$, and
-each $P_ij = \lambda_{i z_j} \lambda_{j z_i}$.
+$\lambda_{i1}, ..., \lambda_{iK}$, and  $P_{ij} = \lambda_{i z_j}
+\lambda_{j z_i}$ for all $i \leq j$.
 
@@
 Let the rows and columns of $P$ be arranged by community label
-such that $n_k \times n_l$ block $P^{(kl)}$
+such that $n_k \times n_{\ell}$ block $P^{(k \ell)}$
 describes the edge probabilities between vertices in communities
-$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$).
-If each block $P^{(kl)}$ can be written as the outer product of two vectors:
-
+$k$ and $\ell$; note that $P^{(k \ell)} = (P^{(k \ell)})^\top$.
+Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
 \begin{equation} \label{eq:pabm}
-  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
+  P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top
 \end{equation}
-
-for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each
-$\lambda^{(st)}$ is a column vector of dimension $n_s$, then graph $G$ is drawn from a PABM with parameters $\{\lambda^{(st)}\}_K$ if its its corresponding adjacency matrix $A \sim \BG(P)$.
+for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{1 \leq
+  s \leq K, 1 \leq t \leq K}$ where each
+$\lambda^{(st)}$ is a column vector of dimension $n_s$. Then a graph $G$
+is drawn from a PABM with parameters
+$\{\lambda^{(st)}\}_K$ if its adjacency matrix $A$
+satisfies $A \sim \mathrm{Bernoulli}(P)$.
 \end{definition}
@@
 GRDPG}\label{connecting-the-pabm-to-the-grdpg}}
-
-\begin{theorem}[Connecting the PABM to the GRDPG for $K = 2$]
+\begin{proposition}[Connecting the PABM to the GRDPG for $K = 2$]
 \label{theorem1}
@@
 identically distributed.
-\end{theorem}
-
+\end{proposition}
+
+\begin{proof}
+  We can show that
+\[X U I_{3, 1} U^\top X^\top =
+\begin{bmatrix}
+  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
+  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
+\end{bmatrix} \]
+by straightforward matrix multiplication. It is nevertheless
+instructive to look at a few intermediate steps. More specifically,
+the product $U I_{3, 1} U^\top$ yields a permutation matrix $\Pi$
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
+positions $2$ and $3$, i.e., 
+$$\Pi = U I_{3, 1} U^\top = \begin{bmatrix} 1 & 0 & 0 & 0 \\
+  0 & 0 & 1 & 0 \\
+  0 & 1 & 0 & 0 \\
+  0 & 0 & 0 & 1
+\end{bmatrix}.$$
+Furthermore, since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $U I_{3, 1} U^\top$
+is also an eigendecomposition of $\Pi$ where the fixed
+points of $\Pi$ are mapped to the eigenvectors $e_1$ and $e_4$
+while the cycles of order two are mapped to the eigenvectors  
+$\tfrac{1}{\sqrt{2}}(e_{2} + e_3)$ and $\tfrac{1}{\sqrt{2}}(e_{2} -
+e_3)$; here $e_i$ denote the $i$th basis vector in $\mathbb{R}^{4}$. %  and $ eigenvalues $\pm 1$ with corresponding
+% eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
+% swapped.
+\end{proof}
+% The form for the matrix $\Pi$ used in the proof of
+% Proposition~\ref{theorem1} can 
 \begin{theorem}[Generalization to $K > 2$]
 \label{theorem2}
-There exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters
-$\{\lambda^{(kl)}\}_K$ and orthonormal matrix
-$U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed
-for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and
-$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\})$ be an instance of a
+PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
+\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. Then there exists a block diagonal matrix
+$X \in \mathbb{R}^{n \times K^2}$ defined by the $\{\lambda^{(kl)}\}$ and a $K^2 \times
+K^2$ {\em fixed} orthonormal matrix $U$ such 
+that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
 \end{theorem}
@@
 \begin{proof}
-First define the following matrices from $\{\lambda^{(kl)}\}_K$:
+We first define the following matrices
 \begin{gather}
 \label{eq:xy}
-\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k,1)} \mid \cdots \mid \lambda^{(k, K)} \end{bmatrix}
+\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k1)} \mid \cdots \mid \lambda^{(kK)} \end{bmatrix}
 \in \mathbb{R}^{n_k \times K}, \quad
 X = \text{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
-\mathbb{R}^{n \times K^2} \\
+\mathbb{R}^{n \times K^2}, \\
 L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
@@
 \end{gather}
-We then have $P = X Y^\top$. Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
-$\Pi$, resulting in $P = X \Pi X^\top$.
+It is then straighforward to verify that
+\begin{gather*}
+  XY^{\top} = \text{blockdiag}(\Lambda^{(1)}, \dots,
+\Lambda^{(K)}) \begin{bmatrix} L_1^{\top} \\ \vdots \\
+  L_{K}^{\top} \end{bmatrix} = \begin{bmatrix} \Lambda^{(1)}
+  (L^{(1)})^{\top} \\ \vdots \\
+  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix} \\
+\Lambda^{(k)} (L^{(k)})^{\top} = \begin{bmatrix} \lambda^{(k1)}
+  (\lambda^{(1k)})^{\top} \mid \dots \mid \lambda^{(kK)}
+  (\lambda^{(Kk)})^{\top} \end{bmatrix} = \begin{bmatrix} P^{(k1)}
+  \mid P^{(k2)} \mid \dots \mid P^{(kK)} \end{bmatrix}.
+\end{gather*}
+We therefore have $P = X Y^\top$. Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
+$\Pi$ and hence $P = X \Pi X^\top$.
 The permutation described by $\Pi$ has $K$ fixed points, which correspond to
@@
 $\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to
-a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors
-$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.
-
-Then $\Pi$ has $p = K (K + 1) / 2$ eigenvalues equal to $1$ and $q = K (K - 1) / 2$
-eigenvalues equal to $-1$. We therefore have
+a pair of eigenvalues $\{-1,+1\}$ and a pair of eigenvectors
+$\{(e_s + e_t)/\sqrt{2},(e_s - e_t)/ \sqrt{2}\}$.
+
+Let $p = K (K + 1) / 2$ and $q = K (K - 1) / 2$. 
+We therefore have
 \begin{equation} \label{eq:permutation}
@@
 \end{equation}
-where $U$ is a $K^2 \times K^2$ orthogonal matrix. The edge probability matrix can then be written as
+where $U$ is a $K^2 \times K^2$ orthogonal matrix and hence
 \begin{equation} \label{eq:pabm-grdpg}
-P = X U I_{p, q} (X U)^\top
+P = X U I_{p, q} (X U)^\top.
 \end{equation}
@@
 % \end{equation}
-We can therefore describe the PABM with $K$ communities as a GRDPG with latent
+In summary we can describe the PABM with $K$ communities as a GRDPG with latent
 positions $X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
@@
 
-\begin{example}[$K = 3$] Using the same notation as in Theorem
+\begin{example} Let $A$ be a $3$ blocks PABM with latent vectors
+  $\{\lambda^{(k \ell)} \colon 1 \leq k \leq 3, 1 \leq \ell \leq 3\}$. Using the same notation as in Theorem
   \ref{theorem2}, we can define
@@
 \end{gather*}
-Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a permutation matrix of
+Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ permutation matrix of
 the form
-$$\Pi = \begin{bmatrix}
-1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
-0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
-0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
-\end{bmatrix}.$$
+$$\Pi = %\begin{bmatrix}
+% 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
+% 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
+% 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1%
+\Bigl[e_1 \mid e_4 \mid e_7 \mid e_2 \mid e_5 \mid e_8 \mid e_3
+\mid e_6 \mid e_9 \Bigr].
+% \end{bmatrix}.
+$$
+where $e_i$ denote the $i$th basis vector in $\mathbb{R}^{9}$. 
 The matrix $\Pi$ corresponds to a permutation of $\{1,2,\dots,9\}$
@@
 We can therefore write $\Pi$ as $\Pi = U I_{6, 3} U^\top$ where the first three
-columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the fixed
-positions $1$, $5$, and $9$, the next three columns consist of eigenvectors
-$(e_k + e_l) / \sqrt{2}$, and the last three columns consist of eigenvectors
-$(e_k - e_l) / \sqrt{2}$, where pairs $(k, l)$ correspond to the cycles of
-order 2 described above.
-
-The matrix $P$ can then be written as a generalized random dot product
-graph where the latent positions are the rows of the matrix
+columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the
+{\em fixed} points, the next three columns consist of eigenvectors
+$(e_k + e_{\ell}) / \sqrt{2}$, and the last three columns consist of eigenvectors
+$(e_k - e_{\ell}) / \sqrt{2}$ for $(k, \ell) \in
+\{(2,4),(3,7),(6,8)\}$.
+
+The matrix $P$ is then the edge probabilities matrix for generalized random dot product
+graph whose latent positions are the rows of the matrix
 $$XU = \begin{bmatrix}
@@
 \end{theorem}
+\begin{proof}
+% \begin{lemma}
+% \label{lemma1}
+% Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
+% matrix for a PABM.
+We first show that $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
+defined as in Eq.\eqref{eq:xy}. Indeed, by Theorem 2, \(P = X U I_{p,
+  q} U^\top X^\top\) for $p = K(K+1)/2$ and $q = K(K-1)/2$. % are given by $p =
+% K(K+1)/2$ and $q = are defined as
+% in equations (\ref{eq:p}) and (\ref{eq:q}). 
+The eigendecomposition \(P = V D V^\top\) also yields $P = V
+|D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ where \(|\cdot|^{1/2}\) is
+applied entry-wise. Now let $Y = XU$ and $Z = V|D|^{1/2}$; note that
+$Y$ and $Z$ both have full column ranks. 
+Since $P = Y I_{p,q} Y^{\top} = Z I_{p,q} Z^{\top}$, we have
+$$Y = Z I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}.$$
+Let $Q = I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}$ and note that
+$Y = Z Q$. We then have
+\begin{equation*}
+  \begin{split}
+  Q^{\top} I_{p,q} Q &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Z I_{p,q}
+I_{p,q} I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
+Y^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} =  I_{p,q}
+\end{split}
+\end{equation*}
+and hence $Q$ is an indefinite orthogonal matrix.
+
+Let $R = U Q |D|^{-1/2}$ and note that $V = XR$. Since $R$ is
+{\em invertible}, we have
+$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
+R^{\top} X^{\top}.$$ 
+Furthermore, as $V$ has orthonormal columns, $R^{\top} X^{\top} X R =
+V^{\top} V = I$. We thus conclude
+$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
+as desired.
+% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
+% we have
+% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
+% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
+% The right-hand side becomes
+% \[\begin{split}
+%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
+%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
+%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
+%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
+%   & = X (X^\top X)^{-1} X^\top.
+% \end{split}\]
+% The left-hand side becomes:
+% \[\begin{split}
+%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
+%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
+%   & = V V^\top.
+% \end{split}\]
+
+% \textbf{Proof of Theorem \ref{theorem3}}. 
+Recall that \(X\) is block diagonal with each block
+corresponding to one community and hence \(X (X^\top X)^{-1} X^\top\) is also a
+block diagonal matrix with each block corresponding to a community. As $B = n VV^{\top} = n X (X^\top X)^{-1} X^\top$, we
+conclude that $B_{ij} = 0$ whenever vertices $i$ and $j$ belong to different communities.  
+\end{proof}
 
@@
 
+\appendix
+\renewcommand\refname{References}
+  \bibliography{misc.bib}
+
 \hypertarget{proofs}{%
-\section{Proofs}\label{proofs}}
-
-\textbf{Proof of Theorem \ref{theorem1}}. This is given by
-straightforward matrix multiplication. It suffices to show that
-
-\[X U I_{3, 1} U^\top X^\top =
-\begin{bmatrix}
-  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}\]
-
-\begin{remark}
-While we can just perform the matrix multiplication to show the
-equivalence, it is more illustrative to look at a few intermediate steps. Note
-that the product of the three inner matrices results in a permutation matrix
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
-positions $2$ and $3$:
-
-$$U I_{3, 1} U^\top = \begin{bmatrix}
-  1 & 0 & 0 & 0 \\
-  0 & 0 & 1 & 0 \\
-  0 & 1 & 0 & 0 \\
-  0 & 0 & 0 & 1
-\end{bmatrix} = \Pi$$
-
-Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$
-is a spectral decomposition of this permutation matrix. Note that the two fixed
-points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$
-where $i = 1, 4$ corresponding to the locations of the fixed points, and the
-cycle of order two results in two eigenvalues $\pm 1$ with corresponding
-eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
-swapped.
-\end{remark}
-
+\section{Proofs of Theorem~\ref{theorem4} and Theorem~\ref{theorem5}}}
+Let \(V_n\) and \(\hat{V}_n\)
+be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
+$K^2$ largest eigenvalues (in modulus), respectively. 
+We first state an important technical lemma for bounding the maximum
+$\ell_2$ norm difference between the rows of $\hat{V}_n$ and
+$V_n$. See \cite{cape_biometrika} and 
+\cite[Lemma~5]{rubindelanchy2017statistical} for a proof. 
 \begin{lemma}
-\label{lemma1}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in Eq.\eqref{eq:xy}.
+  \label{lem:technical}
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_{K})$ be a $K$-blocks
+PABM graph on $n$ vertices and let \(V_n\) and \(\hat{V}_n\)
+be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
+$K^2$ largest eigenvalues in modulus, respectively.
+Let \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
+row of \(V_n\) and \(\hat{V}_n\), respectively. 
+Then there exists a constant $c > 1$ and an orthogonal matrix $W$ such
+that, with high probability,
+$$\max_{i} \|W \hat{v}_n^{(i)} - v_n^{(i)}\|  = O_P \Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
+In particular we can take $c = 1 + \epsilon$ for any $\epsilon > 0$. 
 \end{lemma}
 
-\emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
-\(X\) is defined as in Eq.~\eqref{eq:xy}), $p = K(K+1)/2$, and $q = K(K-1)/2$. % are given by $p =
-% K(K+1)/2$ and $q = are defined as
-% in equations (\ref{eq:p}) and (\ref{eq:q}). 
-Alternatively we have the eigendecomposition \(P = V D V^\top = V
-|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
-\(Q \in \mathbb{O}(p, q)\),
-\[X U Q = V |D|^{1/2}.\]
-Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence
-$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
-R^{\top} X^{\top}.$$ 
-Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
-V = I$. We thus conclude
-$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
-as desired.
-% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
-% we have
-% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
-% The right-hand side becomes
-% \[\begin{split}
-%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
-%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-%   & = X (X^\top X)^{-1} X^\top.
-% \end{split}\]
-% The left-hand side becomes:
-% \[\begin{split}
-%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-%   & = V V^\top.
-% \end{split}\]
-
-\textbf{Proof of Theorem \ref{theorem3}}. By Lemma \ref{lemma1},
-\(V V^\top = X (X^\top X)^{-1} X^\top\) where \(X\) is defined as in
-(\ref{eq:xy}). Since \(X\) is block diagonal with each block
-corresponding to one community, \(X (X^\top X)^{-1} X^\top\) is also a
-block diagonal matrix with each block corresponding to a community and
-zeros elsewhere. Let $B = n VV^{\top}$. Then $B_{ij} = 0$ whenever
-vertices $i$ and $j$ belong to different communities.  
 %
@@
 
-\textbf{Proof of Theorem \ref{theorem4}}. Let \(V_n\) and \(\hat{V}_n\)
-be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
-\(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
-eigenvalues, respectively. From Lemma~5 in
-\citeauthor{rubindelanchy2017statistical} we have, for some
-\(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
-$$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big).$$
-We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). Let
-\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th row
-row of \(V_n\) and \(\hat{V}_n\), respectively. Recall
-Theorem~\ref{theorem3} and note that $B_{ij} = n(v_n^{(i)})^{\top}
-v_n^{(j)}$. Now suppose that vertices $i$ and $j$ belongs to different
-communities. Then $B_{ij} = 0$ and we have
+\begin{proof}[Proof of Theorem~\ref{theorem4}]
+  Recall the notations in Lemma~\ref{lem:technical} and note that,
+  under our assumption that the latent vectors $\lambda^{(k \ell)|}$
+  are all homogeneous, we have $\max_{i} \|v_n^{(i)}\| =
+  O(n^{-1/2})$. 
+%   Let \(V_n\) and \(\hat{V}_n\)
+% be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
+% \(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
+% eigenvalues, respectively. From Lemma~5 in
+% \citeauthor{rubindelanchy2017statistical} we have, for some
+% \(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
+% $$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \big).$$
+% We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). 
+% Let
+% \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
+% row of \(V_n\) and \(\hat{V}_n\), respectively. 
+Next recall Theorem~\ref{theorem3}; in particular $B_{ij} = n(v_n^{(i)})^{\top}
+v_n^{(j)}$. % Now suppose that vertices $i$ and $j$ belongs to different
+% communities. Then $B_{ij} = 0$ and hence
 % Then for \(i\) and
@@
 % communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
+We therefore have
 \[\begin{split}
-\max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
-= \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
+ \max_{ij} |\hat{B}_{ij} - B_{ij}| &= \max_{ij} n |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
 (v_n^{(i)})^\top v_n^{(j)}| \\
-& = \max_{i, j} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
+& \leq n \max_{ij} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
 (v_n^{(i)})^\top v_n^{(j)}| \\
-& \leq \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|\bigr)
+& \leq n \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|
 + \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
@@
 % & \leq 2 
-&\leq \|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|\hat{V}_n\|_{2 \to \infty} +
-\|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty} \\
-& \leq \|\hat{V}_n W - V_n\|_{2 \to \infty}^2 + 2 \|\hat{V}_n W -
-V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty}
+% &\leq  n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|\hat{v}_n^{(j)}\| +
+% n \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr)
+% \\
+& \leq n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 +  \|W
+\hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|v_n^{(j)}\| +  \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr) 
+\\ &
+\leq n \max_{i} \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 + 2n \max_{i}
+\|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \max_{j} \|v_n^{(j)}\|
 \\
-& = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
-\end{split}.\]
-Scaling all terms in the above inequality by \(n\), we obtain
-$$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
-as desired.
+& = O \Big( \frac{\log^{c}{n}}{n^{1/2} \rho_n^{1/2}} \Big)
+\end{split}\]
+with high probability.
+Theorem~\ref{theorem4} follows from the above bound together with the
+conclusion in Theorem~\ref{theorem3} that $B_{ij} = 0$ whenever vertices $i$ and $j$
+belongs to different communities. 
+\end{proof}
+% Scaling all terms in the above inequality by \(n\), we obtain
+% $$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
+% as desired.
 
@@
 \begin{equation} \label{eq:delta-conv}
-\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{a.s.}{\to} 0
+\delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
 \end{equation}
@@
 
-\renewcommand\refname{References}
-  \bibliography{misc.bib}
-
 \end{document}

### 2021-07-07 summary.tex
Update git repo

@@
 \begin{definition}[Subspace Detection Property]
+  \label{def:subspace_detection}
 Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy
-points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the
-solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of
-$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from
-different subspaces, then $X$ obeys the subspace detection property.
+points sampled from $K$ subspaces, i.e., $x_i = y_i + z_i$ where the
+$y_i$ belongs to the union of $K$ subspaces and the $z_i$ are noise
+vectors. Let $\lambda \geq 0$ be given and
+let $C$ and $B$
+be constructed from the
+solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
+this given choice of $\lambda$. Then $X$
+is said to satisfy the subspace detection property with weighting parameter $\lambda$ if each column of
+$C$ has nonzero norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
+from different subspaces.
 \end{definition}
@@
 $n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the
-matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
-positive and $K (K - 1) / 2$ most negative eigenvalues. Then
-$\exists \lambda > 0$ and $N \in \mathbb{N}$ such that when $n > N$,
-$\sqrt{n} \hat{V}_n$ obeys the subspace detection property with probability 1.
+matrix of eigenvectors of $A_n$ corresponding to the $K^2$ largest
+eigenvalues in modulus. Then for all $\epsilon > 0$ 
+there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
+that for all $n \geq N$, $\sqrt{n} \hat{V}_n$ obeys the subspace detection property with
+probability at least $1 - \epsilon$.  
 \end{theorem}
@@
 The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
-from  \citeauthor{jmlr-v28-wang13} and the fact that the unscaled ASE of
-$P_n$ consists of orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume
+from  \citeauthor{jmlr-v28-wang13} and the fact that the columns of
+$V_n$ correspond to orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume
 that the points in the embedding are all of unit length, and while we apply
@@
 Under regularity and sparsity assumptions, given fixed $K$, 
-
 \begin{equation} \label{eq:thm6}
 \max_{k, l \in \{1, ..., K\}} 
-||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
-O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
+\|\hat{\lambda}^{(kl)} - \lambda^{(kl)}\|_{\infty} = 
+O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
 \end{equation}
+with high probability.
 \end{theorem}
@@
 
-We now provide a proof of Theorem~\ref{theorem5}. We first recall several
-useful definitions for analyzing sparse subspace clustering.
-
+We now provide a proof of Theorem~\ref{theorem5}. Our proof is based
+on verifying the sufficient conditions given in Theorem~6 of
+\cite{jmlr-v28-wang13}
+under which sparse subspace clustering based on solving the
+optimization problem in Eq.~\eqref{eq:ssc} yields an affinity matrix
+$B = |C| + |C^{\top}|$ satisfying the subspace detection property of
+Definition~\ref{def:subspace_detection}. We first recall a few
+definitions used in \cite{jmlr-v28-wang13}; for ease of exposition,
+these definitions are stated using the notations of the current paper.
+
+ % We first recall the
+% notion of the inradius of a set of points. 
 \begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
+  \label{inradius}
 The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
 defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
-In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents
-the inradius of the symmetric convex hull of $X$.
+Let $X$ be a $n \times d$ matrix with rows $x_1, x_2, \dots,
+x_n$. We then define, with a slight abuse of notation, $r(X)$ as the
+inradius of the convex hull formed by $\{\pm x_1, \pm x_2, \dots, \pm x_n\}$. 
 \end{definition}
 
-\begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
+\begin{definition}[Subspace incoherence]
+  \label{def:subspace_incoherence}
+Let $\hat{V}$ be the eigenvectors of $A_n$
+corresponding to the $K^2$ largest eigenvalues in modulus. Let
+$\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
+$\hat{V}$ corresponding to the $k^{th}$
+community and let $\hat{V}^{(-k)}$ denote the matrix formed by
+omitting the rows of $\hat{V}$
+corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
+the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
+the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
+$v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
+$P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
+rows of $V^{(k)}$.
+
+Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
+1,2,\dots,n_{k}$ as the solution of the following optimization problem
+$$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
+  \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
+  \eta\|_\infty \leq 1$.}$$
+Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
+be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
+$\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
+as
+$$w_i^{(k)} =
+\frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
+Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
+and define the subspace incoherence for $\hat{V}^{(k)}$ by
+$$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
 \end{definition}
 
-\begin{lemma}
-\label{lemma2}
-Let $\hat{V}$ be the eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues such that the rows of $\hat{V}$ are ordered by community, and let $\hat{V}^{(k)}$ be the rows of the $k^{th}$ community in $\hat{V}$ and $\hat{V}^{(-k)}$ be the rows of $\hat{V}$ with the $k^{th}$ community omitted.
-Denote $(\hat{v}_i^{(k)})^\top$ as the rows of $\hat{V}$, $\hat{V}_{-i}^{(k)}$ as $\hat{V}^{(k)}$ with the $i^{th}$ row omitted, and $\mathcal{S}^{(k)}$ as the subspace spanned by $V^{(k)}$.
-Let $V$, $V^{(k)}$, $V^{(-k)}$, and $v_i^{(k)}$ be the corresponding values for $P_n$.
-
-Let $\nu_{i}^{(k)} = \max\limits_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2 \lambda} \eta^\top \eta$ subject to $||V_{-i}^{(k)} \eta||_\infty \leq 1$, and define the projected dual direction $w_{i}^{(k)}$ as $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$ normalized to length 1.
-Collect the projected dual directions into $W = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$.
-
-Define the subspace incoherence:
-$$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
-Then $\forall k$, we have
-
-\begin{equation} \label{eq:mu-conv}
-\mu_n^{(k)} = 0
-\end{equation}
-\end{lemma}
-
-\begin{proof} From Theorem \ref{theorem3} we have that the subspaces
+% Our proof is based on Theorem~6 in \cite{jmlr-v28-wang13} that provide
+% sufficient conditions under which sparse subspace clustering based on
+% solving the optimization problem in Eq.~\eqref{eq:ssc} yields an
+% affinity matrix $B = |C| + |C^{\top}|$ satisfying the subspace
+% detection property of Definition~\ref{def:subspace_detection}.  
+% \begin{theorem}
+%   Let $d \geq 1$ and let $\mathcal{S}_1, \mathcal{S}_2, \dots, \mathcal{S}_{L}$ be
+%   subspaces of $\mathbb{R}^{d}$. 
+% Let $Y = [y_1 \mid y_2 \mid \dots \mid y_n]^{\top}$ be a $n \times d$ matrix
+% with rows $y_i \in \mathcal{S}_1 \cup \mathcal{S}_2 \cup
+% \dots \mathcal{S}_{L}$. Let $x_i = y_i + z_i$ and suppose that there
+% exists a finite $\delta \geq 0$ such that $\|z_i\| \leq \delta$ for
+% all $i$.
+% \end{theorem}
+
+% Our proof leverages
+% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
+% guaranteeing the subspace detection property of a noisy
+
+% Note that the rows
+% of $\sqrt{n}V$ satisfy the subspace detection property in
+% Definition~\ref{def:subspace_detection}. Now, by 
+% Lemma~\ref{lem:technical}, there exists an orthogonal transformation
+% $W$ such that the rows of $\sqrt{n}\hat{V}_n W$ are
+% uniformly close to the rows of $\sqrt{n} V_n$.
+
+% Our proof leverages
+% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
+% guaranteeing the subspace detection property of a noisy
+
+% together with the uniform row-wise
+% error bound between $\hat{V}$ and $V$ as given in
+% Lemma~\ref{lem:technical}.
+
+% We first recall the
+% notion of the inradius of a set of points. 
+% \begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
+% The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
+% defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
+% In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents
+% the inradius of the symmetric convex hull of $X$.
+% \end{definition}
+
+% The next
+
+% \begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
+% \end{definition}
+
+% \begin{lemma}
+% \label{lemma2} % Let $\hat{V}$ be the eigenvectors of $A_n$
+% % corresponding to the $K^2$ largest eigenvalues in modulus. Let
+% % $\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
+% % $\hat{V}$ corresponding to the $k^{th}$
+% % community and let $\hat{V}^{(-k)}$ denote the matrix formed by
+% % omitting the rows of $\hat{V}$
+% % corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
+% % the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
+% % the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
+% % $v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
+% % $P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
+% % rows of $V^{(k)}$.
+
+% % Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
+% % 1,2,\dots,n_{k}$ as the solution of the following optimization problem
+% % $$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
+% %   \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
+% %   \eta\|_\infty \leq 1$.}$$
+% % Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
+% % be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
+% % $\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
+% % as
+% % $$w_i^{(k)} =
+% % \frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
+% % Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
+% % and define the subspace incoherence for $W^{(k)}$ by
+% % $$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
+% Recall the notations in Definition~\ref{def:subspace_incoherence}. 
+% We then have
+% \begin{equation} \label{eq:mu-conv}
+% \mu^{(k)} = 0, \quad \text{for all $k = 1,2,\dots,K$.}
+% \end{equation}
+% \end{lemma}
+
+% \begin{proof} From Theorem \ref{theorem3} we have that the subspaces
+%   $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
+%   are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
+%   \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
+%   \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
+%   $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
+%     $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
+%     0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
+%     \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ as desired.
+% %   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
+% % must be orthogonal to each row of \(V^{(-k)}\) and hence
+% %\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
+% \end{proof}
+
+% \begin{lemma}
+% \label{lemma3}
+% Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
+% $\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
+
+% \begin{equation} \label{eq:delta-conv}
+% \delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
+% \end{equation}
+% \end{lemma}
+
+With the above definitions in place, we are now ready to state our
+proof of Theorem~\ref{theorem5}.
+
+\begin{proof}[Proof of Theorem~\ref{theorem5}]
+For a given $k = 1,2\dots,K$, let $r^{(k)} = \min_{i}r(V_{-i}^{(k)})$ be inradius of the convex hull formed by
+the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
+\citeauthor{jmlr-v28-wang13} states that there exists a $\lambda > 0$
+such that $\sqrt{n} \hat{V}$ satisfies
+the subspace detection property in
+Definition~\ref{def:subspace_detection} whenever
+the following two conditions are satisfied
+\begin{gather}
+  \label{eq:cond1}
+  \mu^{(k)} < r^{(k)} \quad \text{for all $k = 1,2,\dots,K$}, \\
+  \label{eq:cond2}
+  \max_{i} \|W \hat{v}_{i} - v_{i}\| \leq \min_{k} \frac{r_*(r^{(k)} -
+    \mu^{(k)})}{2 + 7 r^{(k)}}.
+\end{gather}
+% Furthermore, we can also choose the weighting parameter $\lambda$ in
+% the range
+% $$\frac{1}{r_* - 2 \delta - \delta^2} \leq \lambda \leq \min_{k}
+% \frac{r_k - \mu_k - 2 \delta_1}{\delta(1 + 
+% In other words, $\sqrt{n}\hat{V}_n$ 
+% small enough and the
+% subspace inradius is greater than the subspace incoherence for each
+% community \(k\).
+We now verify that for sufficiently large n, Eq.~\eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
+holds with high probability.
+% satisfied with high probability. 
+% Indeed, let
+% The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
+% subspace detection property holds if the noise in $\sqrt{n}\hat{V}$
+% as an estimate for $\sqrt{n} V$ is small enough and that the inradius
+% of the $\hat{V}^{(k)}$ are greater 
+% small enough and the
+% subspace inradius is greater than the subspace incoherence for each
+% community \(k\).
+ % , let \(v_{-i}^{(k)}\) be the matrix \(V^{(k)}\) with the
+% \(i\)\textsuperscript{th} row removed. Then
+
+{\bf Verifying Eq.~\eqref{eq:cond1}}. If $n$ is sufficiently large then
+there are enough vertices in each community $k$ so that
+$\mathrm{span}(V_{-i}^{(k)}) = \mathcal{S}^{(k)}$ for all $i$ and hence
+% spans its corresponding subspace (Theorem
+% \ref{theorem2}). 
+% We therefore have, for sufficiently large $n$, that 
+\(r^{(k)} = \min_{i} r(V_{-i}^{(k)}) > 0\) for
+all $k = 1,2,\dots,K$. % Eq.~\eqref{eq:cond1} then follows from the fact that $\mu^{(k)} = 0$ for all $k = 1,2,\dots,K.$
+% Indeed, from 
+Next, by Theorem \ref{theorem3} we have that the subspaces
   $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
@@
     0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
-    \in V^{(-k)}$ and hence $\mu_n^{(k)} = 0$ as desired.
-%   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
-% must be orthogonal to each row of \(V^{(-k)}\) and hence
-%\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
+    \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ for all $k
+    =1,2,\dots,K$. Therefore $\mu^{(k)} < r^{(k)}$ for all $k =
+    1=2,\dots,K$ as desired.
+
+% Lemma~\ref{lemma2} then implies, with high
+% Combining the above two facts, we have, for sufficiently large, that $n$ \(r^{(k)} > \mu_n^{(k)} = 0\) for all $k = 1,2\dots,K$.
+% where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
+% such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).
+
+{\bf Verifying Eq.~\eqref{eq:cond2}}.
+Let $\delta = \max_{i} \sqrt{n} \|W
+\hat{v}_{i} - v_{i}\|$. Then from Lemma~\ref{lem:technical}, we have
+\(\delta \stackrel{a.s.}{\to} 0\) and hence
+$$\delta < \min_{k} \frac{r_* (r^{(k)} - \mu^{(k)})}{2 + 7 r^{(k)}}$$
+asymptotically almost surely. 
+% \(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
+% probability 1.
+
+In summary $\sqrt{n} \hat{V}$ satisfies the subspace detection property
+with probability converging to $1$ as \(n \to \infty\).
 \end{proof}
 
-\begin{lemma}
-\label{lemma3}
-Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-$\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
-
-\begin{equation} \label{eq:delta-conv}
-\delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
-\end{equation}
-\end{lemma}
-
-\textbf{Proof of Theorem \ref{theorem5}}. The basis of this proof is
-Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
-subspace detection property holds if the noise is small enough and the
-subspace inradius is greater than the subspace incoherence for each
-community \(k\).
-
-Let \(V_{n, -i}^{(k)}\) be \(V_n^{(k)}\) with the
-\(i\)\textsuperscript{th} entry removed. Suppose that for each community
-\(k\), there are enough vertices such that for each \(i\),
-\(V_{n, -i}^{(k)}\) spans its corresponding subspace (Theorem
-\ref{theorem2}). Then
-\(r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0\). Thus by
-(\ref{eq:mu-conv}), for each \(k\), \(r_n^{(k)} > \mu_n^{(k)} = 0\)
-where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
-such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).
-
-Let \(r_n = \min\limits_k r_n^{(k)}\). By (\ref{eq:delta-conv}),
-\(\delta_n \stackrel{a.s.}{\to} 0\). Then as \(n\to \infty\),
-\(\delta_n < \min\limits_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}\)
-\(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
-probability 1.
-
-Thus the conditions for the subspace detection property from Theorem 6
-from \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as
-\(n \to \infty\).
-
 \begin{remark}
-Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $\|v_n^{(i)}\| = 1$, 
-which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for 
-the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long 
-as the $k^{th}$ community spans its subspace, $a r_n^{(k)} > 0 = \mu_n^{(k)}$ 
-$\forall a > 0$.
+Theorem 6 of \citeauthor{jmlr-v28-wang13} assumes that each row $v_i$
+of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. This
+assumption has the effect of scaling the $r^{(k)}$ so that $r^{(k)}
+\leq 1$ for all $k = 1,2,\dots,K$. We emphasize that this assumption
+has no effect on the proof of Theorem \ref{theorem5}. Indeed,
+since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
+spans the subspace $S^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
+for any scalar $a > 0$. 
 \end{remark}
@@
 
-\emph{Case \(k = l\)}. let $P^{(kk)} =
+\emph{Case \(k = l\)}. Let $P^{(kk)} =
 \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
@@
 %adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
+$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O\Bigl(\frac{(\log n_k)^c}{\sqrt{n_k}}\Bigr)$$
+with high probability. 
 Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.
 
-\emph{Case \(k \neq l\)}. \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
-probabilities and edges between communities \(k\) and \(l\). Note that
-\(P^{(kl)} = (P^{(lk)})^\top\).\\
-By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). As
-in the \(k = l\) case, we note that the singular value decomposition
-\(P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top\) is one-dimensional
-and \(\lambda^{(kl)} = \sigma_{kl} u^{(kl)}\). (We can also note that
-the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
-\(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
-\(v^{(kl)} = u^{(lk)}\).)\\
-Now consider the Hermitian dilation
-
-\[M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}\]
-
-which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. It can
-be shown that the spectral decomposition of \(M^{(kl)}\) is
+\emph{Case \(k \neq l\)}. % \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
+% probabilities and edges between communities \(k\) and \(l\). Note that
+% \(P^{(kl)} = (P^{(lk)})^\top\).\\
+% By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). 
+Let \(P^{(k \ell)} = \sigma_{k \ell}^2 u^{(k \ell)} (v^{(k \ell)})^\top\) and 
+\(P^{(\ell k)} = \sigma_{kl}^2 u^{(\ell k)} (v^{(\ell k)})^\top\)  be the singular
+value decompositions and note that \(\sigma_{k \ell} = \sigma_{\ell
+  k}\), \(u^{(k \ell)} = v^{(\ell k)}\), and
+\(v^{(k \ell)} = u^{(\ell k)}\). 
+Now define \(\lambda^{(k \ell)} = \sigma_{k \ell} u^{(k \ell)}\) and
+\(\lambda^{(\ell k)} = \sigma_{k \ell}
+v^{(k \ell)}\).
+
+% (We can also note that
+% the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
+% \(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
+% \(v^{(kl)} = u^{(lk)}\).)\\
+Next consider the Hermitian dilation
+
+\[M^{(k \ell)} = 2 \begin{bmatrix} 0 & P^{(k \ell)} \\ P^{(\ell k)} & 0 \end{bmatrix}\]
+
+which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. The
+eigendecomposition of \(M^{(kl)}\) is then
 
 \[M^{(kl)} = 
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
+\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix} \times 
 \begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top\]
-
-Thus treating \(M^{(kl)}\) as the edge probability matrix of a GRDPG, we
-have latent positions in \(\mathbb{R}^2\) given by
-
-\[\begin{bmatrix} 
-  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
-  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
+\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix}^\top\]
+Thus treating \(M^{(k \ell)}\) as the edge probability matrix of a GRDPG, we
+have latent positions in \(\mathbb{R}^2\) given by the $(n_k + n_{\ell}) \times 2$ matrix
+\[\Lambda^{K \ell} = \begin{bmatrix} 
+  \sigma_{k \ell} u^{(k \ell)} & \sigma_{k \ell} u^{(k \ell)} \\ 
+  \sigma_{k \ell} v^{(k \ell)} & -\sigma_{k \ell} v^{(k \ell)} 
 \end{bmatrix} = 
 \begin{bmatrix} 
-  \lambda^{(kl)} & \lambda^{(kl)} \\ 
-  \lambda^{(lk)} & -\lambda^{(lk)} 
+  \lambda^{(k \ell)} & \lambda^{(k \ell)} \\ 
+  \lambda^{( \ell k)} & -\lambda^{( \ell k)} 
+\end{bmatrix}.\]
+Now consider
+\[\hat{M}^{(k \ell)} = \begin{bmatrix} 0 & A^{(k \ell)} \\ A^{(\ell k)} & 0 \end{bmatrix}\]
+% Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
+
+% \[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
+
+% and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
+% generate the Bernoulli entries in \(A^{(kl)}\).\\
+We can then view \(\hat{M}^{(kl)}\) as an adjacency matrix drawn from
+the edge probabilities matrix \(M^{(kl)}\). Now suppose that the adjacency spectral
+embedding of $\hat{M}^{(k \ell)}$ is represented as the $(n_k +
+n_{\ell}) \times 2$ matrix
+\[\hat{\Lambda}^{(k \ell)} = \begin{bmatrix} 
+  \hat{\lambda}^{(k \ell)} & \hat{\lambda}^{(k \ell)} \\ 
+  \hat{\lambda}^{(\ell k)} & -\hat{\lambda}^{(\ell k)} 
 \end{bmatrix}\]
-
-Now consider
-
-\[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
-Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
-
-\[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
-
-and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
-generate the Bernoulli entries in \(A^{(kl)}\).\\
-Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
-so its adjacency spectral embedding, given by
-
-\[\begin{bmatrix} 
-  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
-  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
-\end{bmatrix}\]
-
-where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3,
-approximates the latent positions of \(M^{(kl)}\) up to indefinite
-orthogonal transformation by the rate given in Theorem 5 of
-\citeauthor{rubindelanchy2017statistical}.\\
-In this case, the indefinite orthogonal transformation \(W_*\) in the
-GRDPG result \cite{rubindelanchy2017statistical} is of the form
-\(U^\top \hat{U}\). The eigenvalues of \(M\) are distinct since the
-signature for this GRDPG is \((1, 1)\), and \(U^\top \hat{U}\) is block
-diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). Therefore, the
-adjacency spectral embedding of \(\hat{M}^{(kl)}\) is a direct
-estimation of the specific latent positions outlined for \(M^{(kl)}\),
-up to sign flip.
-
+where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3. Then
+by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, there
+exists an indefinite orthogonal transformation $W^{*}$ such that,  with
+high probability,
+$$\max_{i} |W^{*} \hat{\Lambda}_{i}^{(k \ell)} - \Lambda_{i}^{(k \ell)} \| =
+O\Bigl(\frac{(\log (n_k + n_{\ell}))^c}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
+with high probability. Here $\Lambda_{i}^{(k \ell)}$ and
+$\hat{\Lambda}_i^{(k \ell)}$ denote the $i$th rows of $\Lambda^{(k
+  \ell)}$ and $\hat{\Lambda}^{(k \ell)}$, respectively. 
+% approximates the latent positions of \(M^{(kl)}\) up to indefinite
+% orthogonal transformation by the rate given in Theorem 5 of
+% \citeauthor{rubindelanchy2017statistical}.\\
+Furthermore, by looking at the proof of Theorem~5 in
+\cite{rubindelanchy2017statistical}, we see that $W^{*}$ is also
+blocks diagonal with $2$ blocks where the positive eigenvalues of $M^{(k \ell)}$
+forming a block and the negative eigenvalues of $M^{(k \ell)}$ forming
+the remaining block. % \(U^\top \hat{U}\). 
+Since $M^{(k \ell)}$ has one positive eigenvalue and one negative
+eigenvalue, we see that $W^{*}$ is necessarily of the form $W^{*}
+= \Bigl[\begin{smallmatrix} 1 & 0 \\ 0 & - 1\end{smallmatrix}\Bigr]$
+% \((1, 1)\), and \(U^\top \hat{U}\) is block
+% diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). 
+Using this form for $W^{*}$, we obtain
+$$\max\{\|\hat{\lambda}^{(k \ell)} - \lambda^{(k \ell)}\|_{\infty},
+\|\hat{\lambda}^{(\ell k)} - \lambda^{(\ell k)}\|_{\infty}\}  =
+O\Bigl(\frac{(\log (n_k + n_{\ell}))^c}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
+with high probability. Combining this bound with the bound for
+$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty}$ given above
+yields Eq.~\eqref{eq:thm6} in Theorem~\ref{theorem6}. 
 \end{document}

### 2021-07-07 summary.tex
updated document

--- 2021-06-30 summary.tex
@@
 \usepackage{natbib}
+\bibliographystyle{abbrvnat}
 \usepackage{parskip}
@@
 \usepackage{comment}
-\usepackage[]{natbib}
 \bibliographystyle{plainnat}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\newcommand{\diag}{\text{diag}}
-\newcommand{\tr}{\text{Tr}}
-\newcommand{\blockdiag}{\text{blockdiag}}
-\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
-\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
-\newcommand{\Bernoulli}{\text{Bernoulli}}
-\newcommand{\Betadist}{\text{Beta}}
-\newcommand{\BG}{\text{BernoulliGraph}}
-\newcommand{\PABM}{\text{PABM}}
-\newcommand{\RDPG}{\text{RDPG}}
-\newcommand{\GRDPG}{\text{GRDPG}}
-\newcommand{\Multinomial}{\text{Multinomial}}
+\newcommand{\diag}{\mathrm{diag}}
+\newcommand{\tr}{\mathrm{Tr}}
+\newcommand{\blockdiag}{\mathrm{blockdiag}}
+\newcommand{\indep}{\stackrel{\mathrm{indep}}{\sim}}
+\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
+\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
+\newcommand{\Betadist}{\mathrm{Beta}}
+\newcommand{\BG}{\mathrm{BernoulliGraph}}
+\newcommand{\PABM}{\mathrm{PABM}}
+\newcommand{\RDPG}{\mathrm{RDPG}}
+\newcommand{\GRDPG}{\mathrm{GRDPG}}
+\newcommand{\Multinomial}{\mathrm{Multinomial}}
 \newtheorem{definition}{Definition}
@@
 where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1.
-Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{(\lambda^{(kl)}\})$ are
+Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{\lambda^{(kl)}\}_K)$ are
 identically distributed.
 \end{theorem}
+
+\begin{proof}
+This is given by
+straightforward matrix multiplication.
+
+\[X U I_{3, 1} U^\top X^\top =
+\begin{bmatrix}
+  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
+  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
+\end{bmatrix}\]
+\end{proof}
+
+\begin{remark}
+While we can just perform the matrix multiplication to show the
+equivalence, it is more illustrative to look at a few intermediate steps. Note
+that the product of the three inner matrices results in a permutation matrix
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
+positions $2$ and $3$:
+
+$$U I_{3, 1} U^\top = \begin{bmatrix}
+  1 & 0 & 0 & 0 \\
+  0 & 0 & 1 & 0 \\
+  0 & 1 & 0 & 0 \\
+  0 & 0 & 0 & 1
+\end{bmatrix} = \Pi$$
+
+Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$
+is a spectral decomposition of this permutation matrix. Note that the two fixed
+points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$
+where $i = 1, 4$ corresponding to the locations of the fixed points, and the
+cycle of order two results in two eigenvalues $\pm 1$ with corresponding
+eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
+swapped.
+\end{remark}
 
@@
 
-\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for
+\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) \cite{5206547} for
 community detection in the PABM. SSC is performed by solving an
@@
 \end{theorem}
+
+\begin{proof}
+By Theorem \ref{theorem2}, \(P = X U I_{p, q} U^\top X^\top\), where
+\(X\) is defined as in Eq.~\eqref{eq:xy}, $p = K(K+1)/2$, and $q = K(K-1)/2$.
+Alternatively we have the eigendecomposition \(P = V D V^\top = V
+|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is 
+applied entry-wise. Thus for some \(Q \in \mathbb{O}(p, q)\),
+
+\[X U Q = V |D|^{1/2}.\]
+
+Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence
+
+$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
+R^{\top} X^{\top}.$$
+
+Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
+V = I$. Thus,
+
+$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$.
+
+Again by Theorem \ref{theorem2}, $X$ is block diagonal, with each block 
+corresponding to a community. Then $n X (X^\top X)^{-1} X^\top$ is also block 
+diagonal with each block corresponding to a community, and zeros elsewhere. 
+Thus, $[n V V^\top]_{ij} = [n X (X^\top X)^{-1} X^\top]_{ij} = 0$ if vertices 
+$i$ and $j$ are from different communities.
+\end{proof}
 
@@
 
-\textbf{Proof of Theorem \ref{theorem1}}. This is given by
-straightforward matrix multiplication. It suffices to show that
-
-\[X U I_{3, 1} U^\top X^\top =
-\begin{bmatrix}
-  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}\]
-
-\begin{remark}
-While we can just perform the matrix multiplication to show the
-equivalence, it is more illustrative to look at a few intermediate steps. Note
-that the product of the three inner matrices results in a permutation matrix
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
-positions $2$ and $3$:
-
-$$U I_{3, 1} U^\top = \begin{bmatrix}
-  1 & 0 & 0 & 0 \\
-  0 & 0 & 1 & 0 \\
-  0 & 1 & 0 & 0 \\
-  0 & 0 & 0 & 1
-\end{bmatrix} = \Pi$$
-
-Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$
-is a spectral decomposition of this permutation matrix. Note that the two fixed
-points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$
-where $i = 1, 4$ corresponding to the locations of the fixed points, and the
-cycle of order two results in two eigenvalues $\pm 1$ with corresponding
-eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
-swapped.
-\end{remark}
-
-\begin{lemma}
-\label{lemma1}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in Eq.\eqref{eq:xy}.
-\end{lemma}
-
-\emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
-\(X\) is defined as in Eq.~\eqref{eq:xy}), $p = K(K+1)/2$, and $q = K(K-1)/2$. % are given by $p =
-% K(K+1)/2$ and $q = are defined as
-% in equations (\ref{eq:p}) and (\ref{eq:q}). 
-Alternatively we have the eigendecomposition \(P = V D V^\top = V
-|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
-\(Q \in \mathbb{O}(p, q)\),
-\[X U Q = V |D|^{1/2}.\]
-Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence
-$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
-R^{\top} X^{\top}.$$ 
-Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
-V = I$. We thus conclude
-$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
-as desired.
-% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
-% we have
-% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
-% The right-hand side becomes
-% \[\begin{split}
-%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
-%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-%   & = X (X^\top X)^{-1} X^\top.
-% \end{split}\]
-% The left-hand side becomes:
-% \[\begin{split}
-%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-%   & = V V^\top.
-% \end{split}\]
-
-\textbf{Proof of Theorem \ref{theorem3}}. By Lemma \ref{lemma1},
-\(V V^\top = X (X^\top X)^{-1} X^\top\) where \(X\) is defined as in
-(\ref{eq:xy}). Since \(X\) is block diagonal with each block
-corresponding to one community, \(X (X^\top X)^{-1} X^\top\) is also a
-block diagonal matrix with each block corresponding to a community and
-zeros elsewhere. Let $B = n VV^{\top}$. Then $B_{ij} = 0$ whenever
-vertices $i$ and $j$ belong to different communities.  
-%
-%then the \(ij\)\textsuperscript{th} element of
-%\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
-
-\textbf{Proof of Theorem \ref{theorem4}}. Let \(V_n\) and \(\hat{V}_n\)
-be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
+\begin{proof}[Proof of Theorem \ref{theorem4}]
+Let \(V_n\) and \(\hat{V}_n\)be the $n \times K^2$ matrices 
+whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
 \(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
@@
 as desired.
+\end{proof}
 
@@
 \begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
+Let $\nu_i(X, \lambda) = 
+\arg\max_\eta x_i^\top \eta - \frac{1}{2 \lambda} \eta^\top \eta$ 
+subject to $\|X_{-i} \eta\|_\infty \leq 1$. 
+Define the projected dual direction of $X$ to subspace $\mathcal{S}$ as 
+$v_i(X, \lambda, \mathcal{S}) = 
+\frac{\mathbb{P}_{\mathcal{S}}(\nu_i)}{\|\mathbb{P}_{\mathcal{S}}(\nu_i)\|}$.
+
+If $X = \begin{bmatrix} X^{(1)} & \cdots & X^{(K)} \end{bmatrix}^\top$ 
+is a set of points lying near $K$ subspaces $\{S^{(1)}, ..., S^{(K)}\}$ 
+and each $x^{(k)}_i$ has corresponding point $y^{(k)}_i$ which lies exactly on 
+$\mathcal{S}^{(k)}$, let $v_i^{(k)} = v_i(X^{(k)}, \lambda, \mathcal{S}^{(k)})$ 
+and 
+$V^{(k)} = \begin{bmatrix} v_1^{(k)} & \cdots & v_{n_k}^{(k)} \end{bmatrix}^\top$. 
+Then we define $X^{(k)}$ as $\mu_k$-incoherent to $X^{(-k)}$ where
+
+$$\mu_k = \max_{y \in Y^{(-k)}} \|V^{(k)} y\|_\infty$$
 \end{definition}
@@
 
-\textbf{Proof of Theorem \ref{theorem5}}. The basis of this proof is
+\begin{proof}[Proof of Theorem \ref{theorem5}]
+The basis of this proof is
 Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
@@
 \(n \to \infty\).
+\end{proof}
 
@@
 
-\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) be
-organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
-n_{\ell}$ matrix obtained by keeping only the rows of $P$
+\begin{proof}[Proof of Theorem \ref{theorem6}]
+Let \(P\) be organized by community such that \(P^{(k \ell)}\) 
+denote the $n_k \times n_{\ell}$ matrix obtained by keeping only the rows of $P$
 corresponding to vertices in community $k$ and the columns of $P$
-corresponding to vertices in community $\ell$. We define $A^{(k
-  \ell)}$ analogously. Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
+corresponding to vertices in community $\ell$. We define 
+$A^{(k  \ell)}$ analogously. 
+Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ 
+for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
 for the cases when $k = \ell$ versus when $k \not = \ell$.
@@
 and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
-%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
+
+$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O_P \bigg( \frac{(\log n_k)^c}{\sqrt{n_k}}\bigg).$$
+
 Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.
@@
 up to sign flip.
+\end{proof}
 
 \renewcommand\refname{References}
+
 \bibliography{misc.bib}

### 2021-07-07 summary.tex
re-added bibliography
removed second set of references
updated citations

--- 2021-07-07 summary.tex
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\newcommand{\diag}{\text{diag}}
-\newcommand{\tr}{\text{Tr}}
-\newcommand{\blockdiag}{\text{blockdiag}}
-\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
-\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
-\newcommand{\Bernoulli}{\text{Bernoulli}}
-\newcommand{\Betadist}{\text{Beta}}
-\newcommand{\BG}{\text{BernoulliGraph}}
-\newcommand{\PABM}{\text{PABM}}
-\newcommand{\RDPG}{\text{RDPG}}
-\newcommand{\GRDPG}{\text{GRDPG}}
-\newcommand{\Multinomial}{\text{Multinomial}}
+\newcommand{\diag}{\mathrm{diag}}
+\newcommand{\tr}{\mathrm{Tr}}
+\newcommand{\blockdiag}{\mathrm{blockdiag}}
+\newcommand{\indep}{\stackrel{\mathrm{indep}}{\sim}}
+\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
+\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
+\newcommand{\Betadist}{\mathrm{Beta}}
+\newcommand{\BG}{\mathrm{BernoulliGraph}}
+\newcommand{\PABM}{\mathrm{PABM}}
+\newcommand{\RDPG}{\mathrm{RDPG}}
+\newcommand{\GRDPG}{\mathrm{GRDPG}}
+\newcommand{\Multinomial}{\mathrm{Multinomial}}
 \newtheorem{theorem}{Theorem}
@@
 
-\citeauthor{307cbeb9b1be48299388437423d94bf1}, who first proposed the
+\citet{307cbeb9b1be48299388437423d94bf1}, who first proposed the
 PABM, used Modularity Maximization (MM) and the Extreme Points (EP)
@@
 to perform SSC on the ASE of \(P\) or \(A\).
-\citeauthor{noroozi2019estimation} instead applied SSC to \(P\) and
+\citet{noroozi2019estimation} instead applied SSC to \(P\) and
 \(A\), foregoing embedding altogether.
@@
 \ref{theorem2}). Much of our work is then built on
-\citeauthor{rubindelanchy2017statistical}, who describe the convergence
+\citet{rubindelanchy2017statistical}, who describe the convergence
 behavior of the ASE of \(A\) to the ASE of \(P\), and
@@
 The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
-from  \citeauthor{jmlr-v28-wang13} and the fact that the columns of
-$V_n$ correspond to orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume
+from  \citet{jmlr-v28-wang13} and the fact that the columns of
+$V_n$ correspond to orthogonal subspaces. \citet{jmlr-v28-wang13} assume
 that the points in the embedding are all of unit length, and while we apply
@@
 \(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
-MLE-based estimator described by \citeauthor{noroozi2019estimation} and
-\citeauthor{307cbeb9b1be48299388437423d94bf1}.
+MLE-based estimator described by \citet{noroozi2019estimation} and
+\citet{307cbeb9b1be48299388437423d94bf1}.
 
 Modularity Maximization is NP-hard, so
-\citeauthor{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
+\citet{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
 (EP) algorithm \cite{le2016}, which is \(O(n^{K - 1})\), as a greedy
@@
 Louvain algorithm was used for modularity maximization,
-as \citeauthor{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
+as \citet{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
 prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
@@
 SSC-A, treats the columns of the adjacency matrix \(A\) as points in
-\(\mathbb{R}^n\), as described in \citeauthor{noroozi2019estimation}.
+\(\mathbb{R}^n\), as described in \citet{noroozi2019estimation}.
 The second method, denoted as SSC-ASE, first embeds \(A\) and then
@@
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \citeauthor{noroozi2019estimation}: Only the 4 most
+the example from \citet{noroozi2019estimation}: Only the 4 most
 frequent species were considered, and the similarities were discretized
@@
 accuracy of 63\% and an adjusted Rand index of 73\%. In comparison,
-\citeauthor{noroozi2019estimation} achieved an adjusted Rand index of
+\citet{noroozi2019estimation} achieved an adjusted Rand index of
 73\% using sparse subspace clustering on the same dataset.
@@
 analysis, we subsetted the data as described by
-\citeauthor{307cbeb9b1be48299388437423d94bf1} for their analysis of the
+\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
 same networks. Our methods underperformed compared to modularity
@@
 to a very straightforward applications of previous work from
-\citeauthor{rubindelanchy2017statistical},
-\citeauthor{soltanolkotabi2012}, and \citeauthor{jmlr-v28-wang13}, which
+\citet{rubindelanchy2017statistical},
+\citet{soltanolkotabi2012}, and \citet{jmlr-v28-wang13}, which
 lead to asymptotically correct solutions with high probability. Similar
@@
 % eigenvalues, respectively. From Lemma~5 in
-% \citeauthor{rubindelanchy2017statistical} we have, for some
+% \citet{rubindelanchy2017statistical} we have, for some
 % \(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
@@
 % Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-% $\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
+% $\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},
 
@@
 the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
-\citeauthor{jmlr-v28-wang13} states that there exists a $\lambda > 0$
+\citet{jmlr-v28-wang13} states that there exists a $\lambda > 0$
 such that $\sqrt{n} \hat{V}$ satisfies
@@
 % Indeed, let
-% The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
+% The basis of this proof is Theorem 6 from \citet{jmlr-v28-wang13}, which states that the
 % subspace detection property holds if the noise in $\sqrt{n}\hat{V}$
@@
 \begin{remark}
-Theorem 6 of \citeauthor{jmlr-v28-wang13} assumes that each row $v_i$
+Theorem 6 of \citet{jmlr-v28-wang13} assumes that each row $v_i$
 of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. This
@@
 \(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
-and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
+and by Theorem 5 of \citet{rubindelanchy2017statistical}, we have
 %adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
@@
 where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3. Then
-by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, there
+by Theorem 5 of \citet{rubindelanchy2017statistical}, there
 exists an indefinite orthogonal transformation $W^{*}$ such that,  with
@@
 % orthogonal transformation by the rate given in Theorem 5 of
-% \citeauthor{rubindelanchy2017statistical}.\\
+% \citet{rubindelanchy2017statistical}.\\
 Furthermore, by looking at the proof of Theorem~5 in
@@
 yields Eq.~\eqref{eq:thm6} in Theorem~\ref{theorem6}.
+
 \end{document}

### 2021-07-08 summary.tex
updated document

@@
 \newcommand{\blockdiag}{\mathrm{blockdiag}}
-\newcommand{\indep}{\stackrel{\mathrm{indep}}{\sim}}
+\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
 \newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
@@
 
-A general family of models that generate unweighted and undirected graphs is the Bernoulli Graph, which is described by an edge probability matrix $P$.
-Edges between each pair of vertices $i$ and $j$ is drawn as a Bernoulli trial with probability $P_{ij}$, for each $0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph.
+A general family of models that describe unweighted and undirected graphs is the Bernoulli Graph, which is characterized by edge probability matrix $P$.
+Edges between each pair of vertices $i$ and $j$ are drawn as a Bernoulli trial with probability $P_{ij}$, for each $0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph.
 Community detection methods often restrict this type of model by assigning each vertex a community label $z_1, ..., z_n$ and then setting each $P_{ij}$ as a function of the labels of vertices $i$ and $j$.
@@
 including the Degree-Corrected Block Model (DCBM), 
-introduced by\citet{Karrer_2011}, 
+introduced by \citet{Karrer_2011}, 
 which assigns an additional parameter $\theta_i$ to each vertex 
@@
 Using \(A\) instead of \(P\) introduces error, which converges to \(0\)
-almost surely:
+almost surely.
 
@@
 \label{theorem4}
-Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC
+Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC
 (Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities
-and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,
+and sparsity factor satisfying $n \rho_n = \omega \big( (\log n)^{4c} \big)$,
 
 \begin{equation} \label{eq:thm4}
-\max_{i, j} |n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| =
+\max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
 O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
@@
 This provides the result that for $i, j$ in different communities,
-$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$.
+$\hat{B}_{ij} \stackrel{a.s.}{\to} 0$.
 \end{theorem}
 
+The sparsity factor $\rho_n$ can be characterized as 
+the average expected degree of the vertices of the graph 
+divided by the number of vertices. 
+Thus Theorem \ref{theorem4} requires the average degree 
+to grow faster than $(\log n)^{4c}$.
+
 Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} also provide
-a very natural path toward using SSC for community detection for the
-PABM. We established in Theorem \ref{theorem2} that an ASE of the edge
+a natural path toward using SSC for community detection. 
+We established in Theorem \ref{theorem2} that an ASE of the edge
 probability matrix \(P\) can be constructed such that the communities
@@
 \label{theorem5}
-Let $P_n$ describe the edge probability matrix of the PABM with
-$n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the
-matrix of eigenvectors of $A_n$ corresponding to the $K^2$ largest
+Let $P$ describe the edge probability matrix of the PABM with
+$n$ vertices, and let $A \sim \Bernoulli(P)$.  Let $\hat{V}$ be the
+matrix of eigenvectors of $A$ corresponding to the $K^2$ largest
 eigenvalues in modulus. Then for all $\epsilon > 0$ 
 there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
-that for all $n \geq N$, $\sqrt{n} \hat{V}_n$ obeys the subspace detection property with
+that for all $n \geq N$, $\sqrt{n} \hat{V}$ obeys the subspace detection property with
 probability at least $1 - \epsilon$.  
 \end{theorem}
-
-\begin{remark}
-The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
-from  \citet{jmlr-v28-wang13} and the fact that the columns of
-$V_n$ correspond to orthogonal subspaces. \citet{jmlr-v28-wang13} assume
-that the points in the embedding are all of unit length, and while we apply
-this normalization in the simulated examples, it is not strictly necessary
-for Theorem \ref{theorem5} due to orthogonality.
-\end{remark}
 
@@
 \label{theorem6}
-Under regularity and sparsity assumptions, given fixed $K$, 
+Let each $\tilde{\lambda}^{(kl)}$ be the popularity vector derived from its corresponding $P^{(kl)}$, and let $\hat{\lambda}^{(kl)}$ be its estimate from $A^{(kl)}$ (Alg.
+3).
+Then if $n \rho_n = \omega \big( (\log n)^{4 c} \big)$,
+
 \begin{equation} \label{eq:thm6}
 \max_{k, l \in \{1, ..., K\}} 
-\|\hat{\lambda}^{(kl)} - \lambda^{(kl)}\|_{\infty} = 
+\|\hat{\lambda}^{(kl)} - \tilde{\lambda}^{(kl)}\|_{\infty} = 
 O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
 \end{equation}
+
 with high probability.
@@
 since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
-spans the subspace $S^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
+spans the subspace $\mathcal{S}^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
 for any scalar $a > 0$. 
@@
 have latent positions in \(\mathbb{R}^2\) given by the $(n_k + n_{\ell}) \times 2$ matrix
-\[\Lambda^{K \ell} = \begin{bmatrix} 
+\[\Lambda^{(k \ell)} = \begin{bmatrix} 
   \sigma_{k \ell} u^{(k \ell)} & \sigma_{k \ell} u^{(k \ell)} \\ 

### 2021-07-08 summary.tex
updated document

@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \mathrm{Bernoulli}(P)$.
+graph $G = (V, E)$ with adjacency matrix $A$ 
+such that $A \sim \mathrm{Bernoulli}(P)$.
 Let each vertex have a community label between $1$ and $K$.
@@
 
-We will use the notation \(A \sim \PABM(\{\lambda^{(kl)}\}_K)\) to denote
+We will use the notation \(A \sim \PABM(\{\lambda^{(k \ell)}\}_K)\) to denote
 a random adjacency matrix \(A\) drawn from a PABM with parameters
-\(\lambda^{(kl)}\) consisting of \(K\) underlying communities.
+\(\lambda^{(k \ell)}\) consisting of \(K\) underlying communities.
 
@@
 for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-$G$ is drawn from the Generalized Random Dot Product Graph with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
+$G$ is drawn from the Generalized Random Dot Product Graph 
+with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
 \end{definition}
@@
 adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\). If instead of fixed latent positions, they are drawn $X_1, ..., X_n \iid F$, we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
+\((p, q)\). If instead of fixed latent positions, 
+they are drawn $X_1, ..., X_n \iid F$, 
+we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
+
+\begin{remark}
+We can use Adjacency Spectral Embedding (ASE) 
+\cite{rubindelanchy2017statistical} to recover the latent vectors of a GRDPG. 
+This procedure consists of taking the spectral decomposition of $A$ 
+(or $P$ if available) and choosing the $p$ most positive and $q$ most negative
+eigenvalues and their corresponding eigenvectors 
+to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
+\end{remark}
 
@@
 \begin{remark}
-Like the RDPG, the latent positions of a GRDPG are not unique
+The latent vectors that produce $X I_{p,q} X^\top = P$ are not unique
 \cite{rubindelanchy2017statistical}.
-More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, then we also have for any
-$Q \in \mathbb{O}(p, q)$,
+More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
+then we also have for any$Q \in \mathbb{O}(p, q)$,
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
@@
 0 & 1 & 0 & 0 \end{bmatrix},$$
-where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1.
-Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{(\lambda^{(kl)}\})$ are
+where the $\{\lambda^{(k \ell)}\}$ are as defined in Definition 1.
+Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and 
+$B \sim \mathrm{PABM}(\{(\lambda^{(k \ell)}\}_K)$ are
 identically distributed.
@@
 $\tfrac{1}{\sqrt{2}}(e_{2} + e_3)$ and $\tfrac{1}{\sqrt{2}}(e_{2} -
-e_3)$; here $e_i$ denote the $i$th basis vector in $\mathbb{R}^{4}$. %  and $ eigenvalues $\pm 1$ with corresponding
-% eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
-% swapped.
+e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$.
 \end{proof}
-% The form for the matrix $\Pi$ used in the proof of
-% Proposition~\ref{theorem1} can 
+
 \begin{theorem}[Generalization to $K > 2$]
@@
 \colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. Then there exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by the $\{\lambda^{(kl)}\}$ and a $K^2 \times
+$X \in \mathbb{R}^{n \times K^2}$ defined by the $\{\lambda^{(k \ell)}\}$ and a $K^2 \times
 K^2$ {\em fixed} orthonormal matrix $U$ such 
@@
 \item
-  Parameter estimation (estimating \(\lambda^{(kl)}\)'s).
+  Parameter estimation (estimating \(\lambda^{(k \ell)}\)'s).
 \end{enumerate}
@@
 
-\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for
-community detection in the PABM. SSC is performed by solving an
-optimization problem for each observed point. Given
-\(X \in \mathbb{R}^{n \times d}\) with vectors
+\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
+\cite{5206547} for community detection in the PABM. 
+The SSC algorithm can be described as follows: 
+Given \(X \in \mathbb{R}^{n \times d}\) with vectors
 \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
@@
 consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
-\(x_i, x_j\) are in different subspaces. If \(X\) instead represents
+\(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
 points near \(K\) subspaces with some noise, a final graph partitioning
-step may be performed (e.g., edge thresholding or spectral clustering).
+step may be required (e.g., edge thresholding or spectral clustering).
 
@@
 block is an outer product of two vectors, i.e.,
-\(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). Therefore, given
-\(P^{(kl)}\), \(\lambda^{(kl)}\) and \(\lambda^{(lk)}\) are solvable
+\(P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top\). Therefore, given
+\(P^{(k \ell)}\), \(\lambda^{(k \ell)}\) and \(\lambda^{(\ell k)}\) are solvable
 up to multiplicative constant using singular value
 decomposition. More specifically, 
-let \(P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top\)
-be the singular value decomposition of \(P^{(kl)}\).
-\(u^{(kl)} \in \mathbb{R}^{n_k}\) and 
-\(v^{(kl)} \in \mathbb{R}^{n_l}\) are vectors
-and \(\sigma^{(kl)}\) is a scalar. Then \(\lambda^{(kl)} = s_1 u^{(kl)}\)
-and \(\lambda^{(lk)} = s_2 v^{(kl)}\) 
-for unidentifiable $s_1 s_2 = (\sigma^{(kl)})^2$.
+let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\)
+be the singular value decomposition of \(P^{(k \ell)}\).
+\(u^{(k \ell)} \in \mathbb{R}^{n_k}\) and 
+\(v^{(k \ell)} \in \mathbb{R}^{n_\ell}\) are vectors
+and \(\sigma^{(k \ell)}\) is a scalar. 
+Then \(\lambda^{(k \ell)} = s_1 u^{(k \ell)}\)
+and \(\lambda^{(\ell k)} = s_2 v^{(k \ell)}\) 
+for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
 Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators by taking the SVD of each $A^{(kl)}$. 
-Since each $\lambda^{(kl)}$ is not strictly identifiable,
-we instead estimate each $\tilde{\lambda}^{(kl)} = \sigma^{(kl)} u^{(kl)}$.
+estimators by taking the SVD of each $A^{(k \ell)}$. 
+Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
+we instead estimate each $\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}$.
 
@@
   \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
-  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
+  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(k \ell)}\}_K$.}
   Arrange the rows and columns of $A$ by community such that each 
-  $A^{(kl)}$ block consists of estimated edge probabilities between 
+  $A^{(k \ell)}$ block consists of estimated edge probabilities between 
   communities $k$ and $l$.\;
   \For {$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
     block.\;
-    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
-    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
-    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
-    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
+    Assign $u^{(k \ell)}$ and $v^{(k \ell)}$ as the first columns of $U$ and $V$. 
+    Assign $(\sigma^{(k \ell)})^2 \leftarrow \Sigma_{11}$.\;
+    Assign $\hat{\lambda}^{(k \ell)} \leftarrow \pm \sigma^{(k \ell)} u^{(k \ell)}$ and 
+    $\hat{\lambda}^{(\ell k)} \leftarrow \pm \sigma^{(k \ell)} v^{(k \ell)}$.
   }
@@
 \label{theorem6}
-Let each $\tilde{\lambda}^{(kl)}$ be the popularity vector derived from its corresponding $P^{(kl)}$, and let $\hat{\lambda}^{(kl)}$ be its estimate from $A^{(kl)}$ (Alg.
+Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived from its corresponding $P^{(k \ell)}$, and let $\hat{\lambda}^{(k \ell)}$ be its estimate from $A^{(k \ell)}$ (Alg.
 3).
@@
 \max_{k, l \in \{1, ..., K\}} 
-\|\hat{\lambda}^{(kl)} - \tilde{\lambda}^{(kl)}\|_{\infty} = 
+\|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
 O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
@@
 For each simulation, community labels are drawn from a multinomial
-distribution, the popularity vectors \(\{\lambda^{(kl)}\}_K\) are drawn
+distribution, the popularity vectors \(\{\lambda^{(k \ell)}\}_K\) are drawn
 from two types of joint distributions depending on whether \(k = l\),
@@
 with Algorithm 3 to estimate the popularity vectors
-\(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
+\(\{\lambda^{(k \ell)}\}_K\), and this method is then compared against an
 MLE-based estimator described by \citet{noroozi2019estimation} and
@@
 In this case, the "true" popularity vectors are derived from 
-taking the SVD of each edge probability block $P^{(kl)}$ 
+taking the SVD of each edge probability block $P^{(k \ell)}$ 
 to avoid the unidentifiable multiplicative constants.
 
-\[\text{RMSE}(\{\hat{\lambda}^{(kl)}\}_K, \{\lambda^{(kl)}\}_K) = 
+\[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}_K, \{\lambda^{(k \ell)}\}_K) = 
 \sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(kl)} - \lambda^{(kl)} \|_2^2}\]
+\|s \hat{\lambda}^{(k \ell)} - \lambda^{(k \ell)} \|_2^2}\]
 
@@
 by reconstructing each 
-$\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$, 
+$\hat{P}^{(k \ell)} = \hat{\lambda}^{(k \ell)} (\hat{\lambda}^{(\ell k)})^\top$, 
 which we use to define another parameter estimation error.
@@
 \[\text{RMSE}(\hat{P}, P) =
-\sum_{k,l} \sqrt{\frac{1}{n_k n_l} \|P^{(kl)} - \hat{P}^{(kl)}\|_F^2}\]
+\sum_{k,l} \sqrt{\frac{1}{n_k n_\ell} \|P^{(k \ell)} - \hat{P}^{(k \ell)}\|_F^2}\]
 
@@
 multinomial distribution with mixture parameters
-\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(kl)}\}_K\) according
+\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(k \ell)}\}_K\) according
 to the drawn community labels, \(P\) was constructed using the drawn
-\(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
+\(\{\lambda^{(k \ell)}\}_K\), and \(A\) was drawn from \(P\) by
 \(A_{ij} \indep \Bernoulli(P_{ij})\). Each simulation has
@@
   Between-group popularities
-  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
+  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
   \(k \neq l\)
@@
   \label{lem:technical}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_{K})$ be a $K$-blocks
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell \ell)}\}_{K})$ be a $K$-blocks
 PABM graph on $n$ vertices and let \(V_n\) and \(\hat{V}_n\)
@@
 
-which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. The
-eigendecomposition of \(M^{(kl)}\) is then
-
-\[M^{(kl)} = 
+which is a symmetric \((n_k + n_\ell) \times (n_k + n_\ell)\) matrix. The
+eigendecomposition of \(M^{(k \ell)}\) is then
+
+\[M^{(k \ell)} = 
 \begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix} \times 
@@
 
-% and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
+% and \(E\) is the \(n_k \times n_\ell\) matrix of independent noise (to
 % generate the Bernoulli entries in \(A^{(kl)}\).\\
-We can then view \(\hat{M}^{(kl)}\) as an adjacency matrix drawn from
-the edge probabilities matrix \(M^{(kl)}\). Now suppose that the adjacency spectral
+We can then view \(\hat{M}^{(k \ell)}\) as an adjacency matrix drawn from
+the edge probabilities matrix \(M^{(k \ell)}\). Now suppose that the adjacency spectral
 embedding of $\hat{M}^{(k \ell)}$ is represented as the $(n_k +
@@
 \end{bmatrix}\]
-where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3. Then
+where each \(\hat{\lambda}^{(k \ell)}\) is defined as in Algorithm 3. Then
 by Theorem 5 of \citet{rubindelanchy2017statistical}, there

### 2021-07-08 summary.tex
updated document

@@
 distribution, the popularity vectors \(\{\lambda^{(k \ell)}\}_K\) are drawn
-from two types of joint distributions depending on whether \(k = l\),
+from two types of joint distributions depending on whether \(k = \ell\),
 the edge probability matrix \(P\) is constructed using the popularity
-vectors, and finally an unweighted and undirected adjacency matrix \(A\)
+vectors, and finally the adjacency matrix \(A\)
 is drawn from \(P\). OSC is then used for community detection, and this
@@
 \(\{\lambda^{(k \ell)}\}_K\), and this method is then compared against an
-MLE-based estimator described by \citet{noroozi2019estimation} and
-\citet{307cbeb9b1be48299388437423d94bf1}.
+MLE-based estimator described in \cite{noroozi2019estimation} and
+\cite{307cbeb9b1be48299388437423d94bf1}.
 
@@
 The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in algorithm 2. The sparsity
+performs SSC on the embedding, as described in Algorithm 2. The sparsity
 parameter \(\lambda\) was chosen via a preliminary cross-validation
@@
 to the drawn community labels, \(P\) was constructed using the drawn
-\(\{\lambda^{(k \ell)}\}_K\), and \(A\) was drawn from \(P\) by
-\(A_{ij} \indep \Bernoulli(P_{ij})\). Each simulation has
-a unique edge probability matrix \(P\).
+\(\{\lambda^{(k \ell)}\}_K\), and \(A\) was drawn from \(P\).
 
@@
 performance of SSC for \(K = 2\) can be attributed to the final spectral
-clustering step of the affinity matrix. A GMM was fit to the Laplacian
-eigenmap, but visual inspection suggests that the communities are not
-distributed as a mixture of Gaussians in the eigenmap. While the
-subspace detection property is guaranteed for large \(n\), in our
-simulations, setting a large enough sparsity parameter for SSC resulted
+clustering step of the affinity matrix. 
+While the subspace detection property is guaranteed for large \(n\), in our
+simulations, setting the sparsity parameter to the required value resulted
 in more than \(K\) disconnected subgraphs. 
+We instead chose a smaller sparsity parameter, 
+necessitating this final clustering step. 
+A GMM was fit to the normalized Laplacian eigenmap of $B$,
+but visual inspection suggests that the communities are not
+distributed as a mixture of Gaussians in the eigenmap. 
+A different choice of mixture distribution may result in better performance.
 
@@
 We observe similar behavior when comparing errors in the edge probability blocks. 
+Both algorithms perform similarly with $n^{-1/2}$ convergence.
 
@@
 \item
-  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}\)
+  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
   for \(k = 1, ..., K\)
@@
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \citet{noroozi2019estimation}: Only the 4 most
+the example from \cite{noroozi2019estimation}: Only the 4 most
 frequent species were considered, and the similarities were discretized
@@
 \cite{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
-analysis, we subsetted the data as described by
-\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
+analysis, we subsetted the data as described in
+\cite{307cbeb9b1be48299388437423d94bf1} for their analysis of the
 same networks. Our methods underperformed compared to modularity
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/mp-1}
+
+}
+
+\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
+\end{figure}
+
 \begin{table}
 
-\caption{\label{tab:unnamed-chunk-6}Community detection error rates for modularity maximization, sparse subspace clustering,  and OSC.}
+\caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
 \centering
@@
 \hline
-Political blogs & 0.050 & 0.196 & 0.062\\
+Political Blogs & 0.050 & 0.196 & 0.062\\
 \hline
@@
 \end{table}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/mp-1}
-
-}
-
-\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
-\end{figure}
 

### 2021-07-09 summary.tex
updated document

@@
 of interest is the religious affiliation. The networks were truncated to
-religions ``1'' and ``2'', and vertices of degree 0 were removed.
+religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig \ref{fig:households-figure} and Table \ref{tab:households-table}.
 
@@
 \caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
-The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:unnamed-chunk-7}
+The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:households-figure}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-8}Community detection error rates for identifying household religion.}
+\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
 \centering
@@
 
-This paper shows the connection between the PABM and the GRDPG, namely
-that a PABM graph can be represented as a union of orthogonal subspaces
-in an embedding under the GRDPG framework. We then exploited this
-relationship to develop community detection and parameter estimation
-methods. In fact, we can represent any graph with Bernoulli edges as a
-GRDPG, and in the PABM case, it turns out that this relationship leads
-to a very straightforward applications of previous work from
-\citet{rubindelanchy2017statistical},
-\citet{soltanolkotabi2012}, and \citet{jmlr-v28-wang13}, which
-lead to asymptotically correct solutions with high probability. Similar
-methods can be applied for other models, such as the Nested Block Model
-\cite{noroozi2021hierarchy}.
+This paper connects the Popularity Adjusted Block Model to the 
+Generalized Random Dot Product Graph and demonstrates how the connection 
+can be exploited for community detection and parameter estimation. 
+Similar to how, under the GRDPG framework, 
+the Stochastic Block Model can be generated by point masses and 
+the Degree-Corrected Block Model can be generated by line segments, 
+we show that the latent configuration that induces the PABM 
+consists of orthogonal subspaces 
+with each subspace corresponding to a community. 
+This latent structure can be recovered via Adjacency Spectral Embedding. 
+We then apply clustering algorithms on the embedding 
+that take advantage of this structure: 
+Orthogonal Spectral Clustering, which is a novel algorithm, 
+and Sparse Subspace Clustering, which has already been applied 
+in some way to the PABM but 
+without asymptotic guarantees \cite{noroozi2019estimation}. 
+Finally, we show that both algorithms result in perfect community detection 
+for large graphs. 
+These theoretical results are further demonstrated via simulation studies.
 

### 2021-07-14 summary.tex
minor edits

@@
 Dot Product Graph for Community Detection and Parameter Estimation}
-\author{}
-\date{\vspace{-2.5em}}
+
+\author{John Koo\\Department of Statistics, Indiana University\\
+and\\
+Minh Tang\\Department of Statistics, North Carolina State University\\
+and\\
+Michael Trosset\\Department of Statistics, Indiana University}
+
+% \date{\vspace{-2.5em}}
+\date{}
 
 \begin{document}
+
 \maketitle
+
 \begin{abstract}
@@
 
-A general family of models that describe unweighted and undirected graphs is the Bernoulli Graph, which is characterized by edge probability matrix $P$.
-Edges between each pair of vertices $i$ and $j$ are drawn as a Bernoulli trial with probability $P_{ij}$, for each $0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph.
-Community detection methods often restrict this type of model by assigning each vertex a community label $z_1, ..., z_n$ and then setting each $P_{ij}$ as a function of the labels of vertices $i$ and $j$.
+A general family of models that describe unweighted and undirected graphs is 
+the Bernoulli Graph, which is characterized by edge probability matrix $P$. 
+Edges between each pair of vertices $i$ and $j$ are drawn 
+as a Bernoulli trial with probability $P_{ij}$, for each 
+$0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph. 
+Community detection methods often restrict this type of model 
+by assigning each vertex a community label $z_1, ..., z_n$ 
+and then setting each $P_{ij}$ as a function of the labels $z_i$ and $z_j$. 
 Such models are called Block Models.
@@
 called the Random Dot Product Graph (RDPG), proposed by
-\citet*{10.1007/978-3-540-77004-6_11}. Under this model, each vertex of the graph
-can be represented by a point in some latent space such that the edge
+\citet*{10.1007/978-3-540-77004-6_11}. 
+Under this model, each vertex of the graph
+can be represented by a vector in some latent space such that the edge
 probability between any pair of vertices is given by their corresponding
@@
 of a given community share the same position in the latent space
-\cite{lyzinski2014}. A similar property exists for the DCBM
+\cite{lyzinski2014}. A similar property exists for the DCBM, 
+which is equivalent to a RDPG in which vertices that belong to the same 
+community all lie in a line segment in the latent space 
 \cite{lyzinski2014, rubindelanchy2017consistency}. 
 Because the edge probability matrix for the PABM is not positive semidefinite, 
-we instead show that it is a special case of the \emph{Generalized} Random Dot Product Graph (GRDPG) in order to analyze the PABM as a latent space model for community detection and parameter estimation.
+we instead show that it is a special case of the 
+\emph{Generalized} Random Dot Product Graph (GRDPG) 
+in order to analyze the PABM as a latent space model 
+for community detection and parameter estimation.
 
@@
 
-Let $G = (V, E)$ be an unweighted, undirected, and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$.
-$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ and $0$ otherwise.
-Because $G$ is symmetric and hollow,
+Let $G = (V, E)$ be an unweighted, undirected, 
+and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. 
+$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ 
+such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ 
+and $0$ otherwise. Because $G$ is symmetric and hollow,
 $A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$. 
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric matrix of edge probabilities. 
-Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ for each $1 \leq i < j \leq n$  (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$).
-We denote $A \sim \BG(P)$ as graph with adjacency matrix $A$ sampled from edge probability matrix $P$ in this manner.
-If each vertex has a hidden label in $[K]$, they are denoted as $z_1, ..., z_n$.
-Finally, we denote $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
+Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ 
+for each $1 \leq i < j \leq n$ (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$). 
+We denote $A \sim \BG(P)$ as a graph with adjacency matrix $A$ 
+sampled from edge probability matrix $P$ in this manner. 
+If each vertex has a hidden label in $[K]$, 
+they are denoted as $z_1, ..., z_n$. 
+Finally, we denote 
+$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ 
+as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
 
@@
 Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
+
 \begin{equation} \label{eq:pabm}
@@
 \end{equation}
+
 for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{1 \leq
@@
 is drawn from a PABM with parameters
-$\{\lambda^{(st)}\}_K$ if its adjacency matrix $A$
+$\{\lambda^{(st)}\}$ if its adjacency matrix $A$
 satisfies $A \sim \mathrm{Bernoulli}(P)$.
@@
 a random adjacency matrix \(A\) drawn from a PABM with parameters
-\(\lambda^{(k \ell)}\) consisting of \(K\) underlying communities.
+\(\{\lambda^{(k \ell)}\}\) consisting of \(K\) underlying communities.
 
@@
 GRDPG}\label{connecting-the-pabm-to-the-grdpg}}
+
 \begin{proposition}[Connecting the PABM to the GRDPG for $K = 2$]
@@
 Let
+
 $$X = \begin{bmatrix}
@@
 0 & 1 & 0 & 0 \end{bmatrix},$$
+
 where the $\{\lambda^{(k \ell)}\}$ are as defined in Definition 1.
@@
 \label{theorem2}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\})$ be an instance of a
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K)$ be an instance of a
 PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
-\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. Then there exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by the $\{\lambda^{(k \ell)}\}$ and a $K^2 \times
-K^2$ {\em fixed} orthonormal matrix $U$ such 
+\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. 
+Then there exists a block diagonal matrix
+$X \in \mathbb{R}^{n \times K^2}$ defined by $\{\lambda^{(k \ell)}\}$ and a 
+$K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
 that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
@@
 \end{gather*}
-We therefore have $P = X Y^\top$. Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
+We therefore have $P = X Y^\top$. 
+Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
 $\Pi$ and hence $P = X \Pi X^\top$.
@@
 We therefore have
+
 \begin{equation} \label{eq:permutation}
@@
 \end{equation}
-% \begin{equation} \label{eq:p}
-% p = K (K + 1) / 2
-% \end{equation}
-% \begin{equation} \label{eq:q}
-% q = K (K - 1) / 2
-% \end{equation}
+
 In summary we can describe the PABM with $K$ communities as a GRDPG with latent
@@
 \end{gather*}
-Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ permutation matrix of
-the form
-$$\Pi = %\begin{bmatrix}
-% 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
-% 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
-% 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
-% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1%
+Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ 
+permutation matrix of the form
+
+$$\Pi = 
 \Bigl[e_1 \mid e_4 \mid e_7 \mid e_2 \mid e_5 \mid e_8 \mid e_3
-\mid e_6 \mid e_9 \Bigr].
-% \end{bmatrix}.
-$$
-where $e_i$ denote the $i$th basis vector in $\mathbb{R}^{9}$. 
+\mid e_6 \mid e_9 \Bigr].$$
+
+where $e_i$ denotes the $i^{th}$ basis vector in $\mathbb{R}^{9}$. 
 The matrix $\Pi$ corresponds to a permutation of $\{1,2,\dots,9\}$
@@
 vectors. Let $\lambda \geq 0$ be given and
-let $C$ and $B$
-be constructed from the
+let $C$ and $B$ be constructed from the
 solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
 this given choice of $\lambda$. Then $X$
-is said to satisfy the subspace detection property with weighting parameter $\lambda$ if each column of
+is said to satisfy the subspace detection property 
+with sparsity parameter $\lambda$ if each column of
 $C$ has nonzero norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
@@
 In practice, a noisy sample $X$ often does not obey the subspace detection
-property. In such cases, $B$ is treated as an affinity matrix for a graph which
-is then partitioned into $K$ subgraphs to obtain the clustering. On the other
+property. In such cases, $B$ is treated as an affinity matrix for a graph, 
+which requires a partitioning step to obtain the final clustering. On the other
 hand, if $X$ does obey the subspace detection property, $B$ describes a graph
 with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
-property holds, there are exactly $K$ subgraphs which map to each subspace,
+property holds, there are exactly $K$ subgraphs which map onto each subspace,
 but it could be the case that some of the subspaces are represented by
@@
 \(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
-prevents us from using the properties of \(XU\) directly. In particular,
-if \(Q \in \mathbb{O}(n)\), then we could use the fact that
-\(\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0\) if
-vertices \(i\) and \(j\) are in different communities.
+prevents us from using the orthogonality property of \(XU\) directly.
 
@@
 community is represented by a \(K\)-dimensional subspace. Therefore, the
-ASE of \(P\) results in subspaces that correspond to each community,
+ASE of \(P\) still consists of subspaces 
+that correspond to communities of the PABM, even if they are not orthogonal, 
 suggesting the use of SSC. Before exploring SSC, we will first consider
@@
 Theorem \ref{theorem3} provides perfect community detection given \(P\).
-Letting \(|B|\) be the affinity matrix for graph \(G\), \(G\) is
+Letting \(|B|\) be the affinity matrix for graph \(G'\), \(G'\) is
 partitioned into at least \(K\) disjoint subgraphs since each of the
@@
 detection property, it could be the case that some of the communities
-are represented by multiple disjoint subgraphs in \(G\), in which case
+are represented by multiple disjoint subgraphs in \(G'\), in which case
 additional reconstruction is required to identify the communities
@@
 Using \(A\) instead of \(P\) introduces error, which converges to \(0\)
-almost surely.
+almost surely. We use this fact to motivate Orthogonal Spectral Clustering 
+(Alg. 1).
 
@@
 We established in Theorem \ref{theorem2} that an ASE of the edge
-probability matrix \(P\) can be constructed such that the communities
-lie on mutually orthogonal subspaces, and this property can be recovered
-from the eigenvectors of \(P\). Then Theorems \ref{theorem3} and
-\ref{theorem4} show that this property holds for the unscaled ASE of
-\(A\) drawn from \(P\) as \(n \to \infty\).
+probability matrix \(P\) can be constructed from a latent vector configuration 
+consisting of orthogonal subspaces. Theorem \ref{theorem3} shows how 
+this property can be recovered from the eigenvectors of \(P\). 
+Then Theorem \ref{theorem4} shows that replacing $P$ with $A$ 
+approximates these properties with error that goes to zero as \(n \to \infty\). 
+This allows us to obtain the subspace detection property 
+in Theorem \ref{theorem5}.
 
@@
 for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
+Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
+we instead estimate each 
+$\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}$. 
 Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators by taking the SVD of each $A^{(k \ell)}$. 
-Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
-we instead estimate each $\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}$.
+estimators by taking the SVD of each $A^{(k \ell)}$ to obtain 
+$\hat{\lambda}^{(k \ell)} = \hat{\sigma}^{(k \ell)} \hat{u}^{(k \ell)}$ 
+using the largest signular value of $A$ and its corresponding singular vectors.
 
@@
 
-\[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}_K, \{\lambda^{(k \ell)}\}_K) = 
+\[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}, \{\tilde{\lambda}^{(k \ell)}\}) = 
 \sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(k \ell)} - \lambda^{(k \ell)} \|_2^2}\]
+\|s \hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)} \|_2^2}\]
 
@@
 
-Fig \ref{fig:clust_err_ct_sim} shows OSC's community detection error
+Fig. \ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
 going to 0 for large \(n\). SSC on both the embedding and on the
-adjacency matrix produces similar results for \(K > 2\). Weaker
+adjacency matrix produces similar trends for \(K > 2\). Weaker
 performance of SSC for \(K = 2\) can be attributed to the final spectral
@@
 We instead chose a smaller sparsity parameter, 
-necessitating this final clustering step. 
+necessitating a final clustering step. 
 A GMM was fit to the normalized Laplacian eigenmap of $B$,
@@
 Given ground truth community labels, Algorithm 3 and the MLE-based
-plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1} 
-perform similarly, with root mean square
-error decaying at rate approximately \(n^{-1/2}\).
+plug-in estimators perform similarly, with root mean square
+error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:lambda_est_k}). 
+We observe similar behavior when comparing errors 
+in the edge probability blocks (Fig. \ref{fig:p_block_est}).
 
@@
 Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.}\label{fig:lambda_est_k}
+Communities were drawn to be approximately balanced.}
+\label{fig:lambda_est_k}
 \end{figure}
-
-We observe similar behavior when comparing errors in the edge probability blocks. 
-Both algorithms perform similarly with $n^{-1/2}$ convergence.
 
@@
 Communities were drawn to be approximately balanced.}
+\label{fig:p_block_est}
 \end{figure}
@@
 
-Simulations performed in this section are similar to those in the
+Simulations performed in this section are the same as those in the
 previous section with the exception of the mixture parameters
@@
 of interest is the religious affiliation. The networks were truncated to
-religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig \ref{fig:households-figure} and Table \ref{tab:households-table}.
+religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig. \ref{fig:households-figure} and Table \ref{tab:households-table}.
 

### 2021-07-15 summary.tex
minor edits

@@
   \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
-  \(k \neq l\)
+  \(k \neq \ell\)
 \end{itemize}
@@
 if we reconstruct the edge probability blocks 
-by taking the outer products of the popularity vectors.
+by taking the outer products of the popularity vectors 
+(Fig. \ref{fig:lambda_est_p_imba}).
 
@@
 
-Comparing against the ground truth species labels, OSC achieves an
-accuracy of 63\% and an adjusted Rand index of 73\%. In comparison,
-\citet{noroozi2019estimation} achieved an adjusted Rand index of
-73\% using sparse subspace clustering on the same dataset.
+Comparing against the ground truth species labels, OSC achieves an adjusted Rand index of 92\%, while SSC on the ASE achieves an adjusted Rand index of 96\%.
+In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
 
@@
 
-In the third example, we consider the Karantaka villages data studied by
-\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
+In the third example, we analyzed the Karantaka villages data studied by
+\citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
 networks from villages 12, 31, and 46 at the household level. The label
 of interest is the religious affiliation. The networks were truncated to
-religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig. \ref{fig:households-figure} and Table \ref{tab:households-table}.
+religions ``1'' and ``2'', and vertices of degree 0 were removed. 
+The results are displayed in Fig. \ref{fig:households-figure} and 
+Table \ref{tab:households-table}.
 

### 2021-07-19 summary.Rmd
updated document
updated writeup

@@
 
-\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for 
+\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) \cite{5206547} for 
 community detection in the PABM. 
@@
 
-# ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
+ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
 ```

### 2021-07-23 summary.tex
updated document per meeting/discussion

@@
 
-\title{Connecting the Popularity Adjusted Block Model to the Generalized Random
-Dot Product Graph for Community Detection and Parameter Estimation}
+\title{Popularity Adjusted Block Models are Generalized Random Dot Product Graphs}
 
@@
 \newtheorem{definition}{Definition}
-\newtheorem*{example}{Example}
+\newtheorem{example}{Example}
 \newpage
@@
 
-Statistical analysis on graphs or networks often involves the
+Statistical inference on random graphs requires a suitable probability model.
+A general probability model for unweighted and undirected graphs is the Bernoulli Graph, which assumes that edges occur as independent Bernoulli trials.
+A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
+A trivial example of a Bernoulli Graph is the Erd\"{o}s-R\'{e}nyi model proposed by Gilbert \cite{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
+The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
+The present work relates two lines of generalization.
+
+Network analysis is often concerned with community detection.
+One form of community detection assumes that each vertex belongs to an unobserved community, with the probability of an edge between vertices $i$ and $j$ depending on the communities to which $i$ and $j$ belong.
+Formally, one assigns each vertex $v_i$ a community label $z_i$ and assumes a Bernoulli Graph in which $P_{ij}$ is a function of $z_i$ and $z_j$.
+Such models, called Block Models, define the goal of community detection as a problem in statistical inference: identify the true community (up to permutation of labels) to which each vertex belongs.
+
+The classical Stochastic Block Model (SBM) of \citet{doi:10.1080/0022250X.1971.9989788} specifies that each edge probability $P_{ij}$ depends only on the labels $z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_,z_j}$.
+Subsequent researchers have weakened this assumption.
+The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_,z_j}$.
+The Popularity Adjusted Block Model (PABM) of \citet*{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
+
+Another type of Bernoulli Graph was proposed by \citet*{10.1007/978-3-540-77004-6_11}.
+A Random Dot Product Graph (RDPG) specifies that each vertex corresponds to a latent position vector in Euclidean space and that the probability of an edge between two vertices is the dot product of their latent position vectors.
+Thus, if the latent positions $x_1,\ldots,x_n \in \mathbb{R}^d$ and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$, then the edge probability matrix is $P = XX^\top$.
+Clearly, any Bernoulli Graph with positive semidefinite $P$ is an RDPG.
+The positive definite Euclidean inner product in the RDPG model was replaced by an indefinite inner product in \cite{rubindelanchy2017consistency}, resulting in the {\em Generalized}\/ RDPG (GRDPG).
+
+In contrast to Block Models, neither RDPGs nor GRDPGs inherently specify distinct communities.
+However, one can easily impose community structure by assuming that the latent positions lie in distinct clusters.
+Hence, it is not surprising that Block Models can be studied by reformulating them as RDPGs or GRDPGs.
+For example, an assortative SBM (an SBM for which $P$ is positive semidefinite) is equivalent to an RDPG for which all vertices in the same community correspond to the same latent position vector \cite{lyzinski2014}.
+Likewise, the DCBM is equivalent to an RDPG for which all vertices in the same community correspond to latent position vectors that lie on a straight line \cite{lyzinski2014, rubindelanchy2017consistency}.
+
+Because the edge probability matrix of a PABM is not necessarily positive semidefinite, a PABM is not necessarily an RDPG.
+In Section \ref{connecting-the-pabm-to-the-grdpg} we demonstrate that every PABM is in fact a specific type of GRDPG for which the latent position vectors lie in distinct orthogonal subspaces, each subspace corresponding to a community.
+This identification is our central result.
+In Section \ref{methods}, we use the geometry of the GRDPG to derive more efficient algorithms for detecting the communities and estimating the parameters in the PABM.
+We report the results of simulation studies in Section \ref{simulated-examples} and apply our methods to three well-known data sets in Section \ref{real-data-examples}.
+Section \ref{discussion} concludes.
+
+\begin{comment}
+
+Statistical inference on random graphs involves the
 partitioning of a graph into disconnected subgraphs or clusters. This is
@@
 
+\end{comment}
+
+\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
+\section{PABMs are GRDPGs}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
+
+In this section, we show that the PABM is a special case of the GRDPG. 
+More specifically, graph $G$ drawn from the PABM can be represented by 
+a collection of latent vectors in Euclidean space. 
+We further show that the latent configuration that
+induces the PABM consists of orthogonal subspaces with each subspace
+corresponding to a community.
+
 \hypertarget{notation}{%
@@
 
-\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
-\section{Connecting the Popularity Adjusted Block Model to the
-Generalized Random Dot Product
-Graph}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
-
-In this section, we show that the PABM is a special case of the GRDPG, i.e.,
-graph $G$ drawn from the PABM can be represented by a collection of latent
-vectors in Euclidean space. We further show that the latent configuration that
-induces the PABM consists of orthogonal subspaces with each subspace
-corresponding to a community.
-
 \hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
-\subsection{The Popularity Adjusted Block Model and the Generalized
-Random Dot Product
-Graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
+\subsection{Two Probability Models for Graphs}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
 
@@
 Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$.
-If $\exists X \in \mathbb{R}^{n \times d}$ such that
+If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
 
@@
 \hypertarget{connecting-the-pabm-to-the-grdpg}{%
-\subsection{Connecting the PABM to the
-GRDPG}\label{connecting-the-pabm-to-the-grdpg}}
-
-\begin{proposition}[Connecting the PABM to the GRDPG for $K = 2$]
-\label{theorem1}
-Let
+\subsection{The Geometry of PABMs}\label{connecting-the-pabm-to-the-grdpg}}
+
+Now that we defined the PABM and GRDPG, 
+we show the special geometry of the PABM when viewed as a GRDPG.
+
+\begin{theorem}[The latent configuration of the PABM]
+\label{theorem2}
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K)$ be an instance of a
+PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
+\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. 
+Then there exists a block diagonal matrix
+$X \in \mathbb{R}^{n \times K^2}$ defined by $\{\lambda^{(k \ell)}\}$ and a 
+$K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
+that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
+\end{theorem}
+
+\begin{proof}
+We will prove this theorem in two parts. First, for demonstration purposes, we focus on the case for $K = 2$. Then we generalize this to $K \geq 2$.
+
+For the $K = 2$ case, the proof is straightforward. Let
 
@@
 0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix},$$
-
-where the $\{\lambda^{(k \ell)}\}$ are as defined in Definition 1.
-Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and 
-$B \sim \mathrm{PABM}(\{(\lambda^{(k \ell)}\}_K)$ are
-identically distributed.
-\end{proposition}
-
-\begin{proof}
-  We can show that
+0 & 1 & 0 & 0 \end{bmatrix}.$$
+
+Then by straightforward matrix multiplication, we obtain
+
 \[X U I_{3, 1} U^\top X^\top =
@@
   \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix} \]
-by straightforward matrix multiplication. It is nevertheless
-instructive to look at a few intermediate steps. More specifically,
-the product $U I_{3, 1} U^\top$ yields a permutation matrix $\Pi$
-with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
-positions $2$ and $3$, i.e., 
+\end{bmatrix}, \]
+
+which is a GRDPG with latent vectors described by $X U$.
+
+It is nevertheless instructive to look at a few intermediate steps. 
+More specifically, the product $U I_{3, 1} U^\top$ 
+yields a permutation matrix $\Pi$ with fixed points at positions $1$ and $4$ 
+and a cycle of order 2 swapping positions $2$ and $3$, i.e., 
 $$\Pi = U I_{3, 1} U^\top = \begin{bmatrix} 1 & 0 & 0 & 0 \\
@@
 e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$.
-\end{proof}
-
-\begin{theorem}[Generalization to $K > 2$]
-\label{theorem2}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K)$ be an instance of a
-PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
-\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. 
-Then there exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by $\{\lambda^{(k \ell)}\}$ and a 
-$K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
-that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
-\end{theorem}
-
-\begin{proof}
-We first define the following matrices
+
+For the general case, we first define the following matrices
 \begin{gather}
@@
 
-The matrix $P$ is then the edge probabilities matrix for generalized random dot product
-graph whose latent positions are the rows of the matrix
+The matrix $P$ is then the edge probabilities matrix for a 
+Generalized Random Dot Product Graph whose latent positions 
+are the rows of the matrix
 $$XU = \begin{bmatrix}
   \lambda^{(11)} & 0 & 0 &
-  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 &
-  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
+  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 &
+  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 \\
   0 & \lambda^{(22)} & 0 &
-  \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} &
-  -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
+  \frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} &
+  -\frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} \\
   0 & 0 & \lambda^{(33)} &
-  0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} &
-  0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
-\end{bmatrix}$$
+  0 & \frac{\lambda^{(31)}}{\sqrt{2}} & \frac{\lambda^{(32)}}{\sqrt{2}} &
+  0 & -\frac{\lambda^{(31)}}{\sqrt{2}} & -\frac{\lambda^{(32)}}{\sqrt{2}}
+\end{bmatrix}.$$
 \end{example}
@@
 \hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section{Algorithms}\label{methods}}
 
@@
 \hypertarget{related-work}{%
-\subsection{Related work}\label{related-work}}
-
-\citet{307cbeb9b1be48299388437423d94bf1}, who first proposed the
-PABM, used Modularity Maximization (MM) and the Extreme Points (EP)
+\subsection{Previous Work}\label{related-work}}
+
+\citet{307cbeb9b1be48299388437423d94bf1} 
+used Modularity Maximization (MM) and the Extreme Points (EP)
 algorithm \cite{le2016} for community detection and parameter
@@
 \hypertarget{community-detection}{%
-\subsection{Community detection}\label{community-detection}}
+\subsection{Algorithms for Community Detection}\label{community-detection}}
 
@@
 If we decompose \(P = Z I_{p, q} Z^\top\), then
-\(\exists Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
+\(\exists\) \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
 start with the exact edge probability matrix, we cannot recover the
@@
 \hypertarget{parameter-estimation}{%
-\subsection{Parameter estimation}\label{parameter-estimation}}
+\subsection{Algorithms for Parameter Estimation}\label{parameter-estimation}}
 
@@
 \hypertarget{simulated-examples}{%
-\section{Simulated Examples}\label{simulated-examples}}
+\section{Simulation Study}\label{simulated-examples}}
 
@@
 \hypertarget{balanced-communities}{%
-\subsection{Balanced communities}\label{balanced-communities}}
+\subsection{Balanced Communities}\label{balanced-communities}}
 
@@
 
-\begin{comment}
-
-Fig. \ref{fig:clust_err_k} show that for large $n$, OSC results in
-a misclustering rate of 0.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/clust_err_k-1}
-
-}
-
-\caption{IQR of community detection error error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are approximately balanced.
-Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k}
-\end{figure}
-
-Theorem \ref{theorem4} implies that OSC will result in not just in the error
-rate converging to 0 but the error *count* as well.
-We explore this in Fig. \ref{fig:clust_err_ct_sim}.
-
-\end{comment}
-
 Fig. \ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
@@
 \hypertarget{imbalanced-communities}{%
-\subsection{Imbalanced communities}\label{imbalanced-communities}}
+\subsection{Imbalanced Communities}\label{imbalanced-communities}}
 
@@
 
-\begin{comment}
-
-Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show
-similar results as in the balanced communities case, with both OSC and
-SSC resulting in no misclustered vertices for a
-sufficiently large sample size. However, Fig. \ref{fig:lambda_est_k_imba}
-suggests that while Algorithm 3 retains $\sqrt{n}$ efficiency, the MLE-based
-plug-in estimator is less efficient for this setup.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/clust_err_k_imba-1}
-
-}
-
-\caption{Median and IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
-Communities are imbalanced.
-Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k_imba}
-\end{figure}
-
-\end{comment}
-
 We again see community detection error trending to 0 for OSC, as well as
@@
 
-\begin{comment}
-
-## Additional experiments
-
-Using the same set of parameters for generating $P$ and $A$ as in the
-balanced communities examples for $K = 2$, we generated one instance of $A$ for
-each $n$ and constructed $B$ according to Algorithm 3 to verify that
-as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{a.s.}{\to} 0$ for $i, j$
-in different clusters. Furthermore, the distribution of these inner products
-should be approximately normal.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-2-1}
-
-}
-
-\caption{Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.}\label{fig:unnamed-chunk-2}
-\end{figure}
-
-\end{comment}
-
 \hypertarget{real-data-examples}{%
-\section{Real data examples}\label{real-data-examples}}
+\section{Applications}\label{real-data-examples}}
+
+In this section, our algorithms were applied
 
@@
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \cite{noroozi2019estimation}: Only the 4 most
+the example from \cite{noroozi2019estimation}: Only the $K=4$ most
 frequent species were considered, and the similarities were discretized
@@
 
-This paper connects the Popularity Adjusted Block Model to the 
-Generalized Random Dot Product Graph and demonstrates how the connection 
+The central result of this paper illustrates that the 
+Popularity Adjusted Block Model is a special case of the 
+Generalized Random Dot Product Graph and how this connection 
 can be exploited for community detection and parameter estimation. 
-Similar to how, under the GRDPG framework, 
-the Stochastic Block Model can be generated by point masses and 
-the Degree-Corrected Block Model can be generated by line segments, 
-we show that the latent configuration that induces the PABM 
-consists of orthogonal subspaces 
-with each subspace corresponding to a community. 
-This latent structure can be recovered via Adjacency Spectral Embedding. 
-We then apply clustering algorithms on the embedding 
-that take advantage of this structure: 
-Orthogonal Spectral Clustering, which is a novel algorithm, 
-and Sparse Subspace Clustering, which has already been applied 
-in some way to the PABM but 
-without asymptotic guarantees \cite{noroozi2019estimation}. 
-Finally, we show that both algorithms result in perfect community detection 
-for large graphs. 
-These theoretical results are further demonstrated via simulation studies.
+Specifically, the PABM with $K$ communities is a GRDPG 
+for which each community is represented by a $K$-dimensional subspace 
+in the latent space, and the subspaces are orthogonal to each other. 
+This result can be viewed as a continuation of previous results 
+that connect the Stochastic Block Model and the Degree Corrected Block Model 
+to the Random Dot Product Graph: 
+The SBM, DCBM, and PABM are successive generalizations of Block Models. 
+Similarly, the SBM is a RDPG for which each community corresponds to a point 
+(zero-dimensional subspace) \cite{lyzinski2014}, the DCBM is a RDPG 
+for which each community corresponds to a line (one-dimensional subspace) 
+\cite{rubindelanchy2017consistency}, 
+and the PABM is a GRDPG for which each community is a $K$-dimensional subspace.
+
+The latent structure of the PABM can be recovered 
+via Adjacency Spectral Embedding \cite{rubindelanchy2017statistical}, 
+which allows us to take advantage of this latent structure 
+for statistical inference. 
+Using the fact that the communities are represented by orthogonal subspaces, 
+we propose Orthogonal Spectral Clustering for community detection, 
+for which we have the theoretical guarantee that 
+the number of misclustered vertices approaches zero 
+with high probability as the size of the graph increases. 
+This is a stronger result than previously proposed algorithms
+\cite{307cbeb9b1be48299388437423d94bf1},
+which only guarantee that the error rate (and not count) 
+approaches zero asymptotically.
+Parameter estimation can be performed in a similar fashion using the ASE, 
+for which we also prove that 
+the per-parameter error approaches zero asymptotically. 
+Both of these results are demonstrated via simulation study.
+
+One proposed approach to community detection for the PABM 
+is an application of Sparse Subspace Clustering \cite{noroozi2019estimation}. 
+We show that the latent structure of the PABM leads to 
+both a straightforward justification for SSC for the PABM 
+as well as an improvement over the previous implementation.
+Our approach applies SSC on the ASE, 
+and we prove that the ASE of the PABM obeys the Subspace Detection Property
+with high probability if the graph is large. 
+This result is also demonstrated via simulation study.
 

### 2021-07-25 summary.tex
updated document

@@
 (PABM) and the Generalized Random Dot Product Graph (GRDPG),
-demonstrating that a PABM is a GRDPG in which communities correspond to
-certain mutually orthogonal subspaces of latent positions. This insight
+by demonstrating that a PABM is a GRDPG in which communities correspond to 
+mutually orthogonal subspaces of latent vectors. This insight
 leads to the construction of improved algorithms for community detection
 and parameter estimation with PABM. Using established asymptotic
-properties of Adjacency Spectral Embedding (ASE) for GRDPG, we derive
+properties of Adjacency Spectral Embedding for GRDPG, we derive
 asymptotic properties of these algorithms, including algorithms that
-rely on Sparse Subspace Clustering (SSC). We illustrate these properties
-via simulation.
+rely on Sparse Subspace Clustering. We illustrate these properties
+via simulation study.
 \end{abstract}

### 2021-08-02 summary.tex
finally figured out everything for asa format (hopefully)

@@
 \usepackage{mathtools}
-\usepackage{natbib}
-\bibliographystyle{abbrvnat}
+% \usepackage{natbib}
+\usepackage[backend=bibtex,style=authoryear,giveninits,natbib]{biblatex}
+\addbibresource{misc.bib}
+\renewbibmacro{in:}{%
+  \ifentrytype{article}{}{%
+  \printtext{\bibstring{in}\intitlepunct}}}
 \usepackage{parskip}
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\setcitestyle{numbers,square,comma}
+% \setcitestyle{numbers,square,comma}
 \usepackage{verbatim}
@@
 \usepackage{comment}
-\bibliographystyle{plainnat}
 \addtolength{\oddsidemargin}{-.5in}%
@@
 A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
-A trivial example of a Bernoulli Graph is the Erd\"{o}s-R\'{e}nyi model proposed by Gilbert \cite{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
+A trivial example of a Bernoulli Graph is the Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
 The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
@@
 The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_,z_j}$.
-The Popularity Adjusted Block Model (PABM) of \citet*{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
+The Popularity Adjusted Block Model (PABM) of \citet{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
 
@@
 Clearly, any Bernoulli Graph with positive semidefinite $P$ is an RDPG.
-The positive definite Euclidean inner product in the RDPG model was replaced by an indefinite inner product in \cite{rubindelanchy2017consistency}, resulting in the {\em Generalized}\/ RDPG (GRDPG).
+The positive definite Euclidean inner product in the RDPG model was replaced by an indefinite inner product in \citet{rubindelanchy2017consistency}, resulting in the {\em Generalized}\/ RDPG (GRDPG).
 
@@
 Hence, it is not surprising that Block Models can be studied by reformulating them as RDPGs or GRDPGs.
-For example, an assortative SBM (an SBM for which $P$ is positive semidefinite) is equivalent to an RDPG for which all vertices in the same community correspond to the same latent position vector \cite{lyzinski2014}.
-Likewise, the DCBM is equivalent to an RDPG for which all vertices in the same community correspond to latent position vectors that lie on a straight line \cite{lyzinski2014, rubindelanchy2017consistency}.
+For example, an assortative SBM (an SBM for which $P$ is positive semidefinite) is equivalent to an RDPG for which all vertices in the same community correspond to the same latent position vector (\cite{lyzinski2014}).
+Likewise, the DCBM is equivalent to an RDPG for which all vertices in the same community correspond to latent position vectors that lie on a straight line (\cite{lyzinski2014, rubindelanchy2017consistency}).
 
@@
 used Modularity Maximization (MM) and the Extreme Points (EP)
-algorithm \cite{le2016} for community detection and parameter
+algorithm (\cite{le2016}) for community detection and parameter
 estimation. They were able to show that as the sample size increases,
@@
 
-\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
-\cite{5206547} for community detection in the PABM. 
+\citet{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
+(\cite{5206547}) for community detection in the PABM. 
 The SSC algorithm can be described as follows: 
@@
 identification or estimation of \(XU\) is not possible
-\cite{rubindelanchy2017statistical}.
+(\cite{rubindelanchy2017statistical}).
 
@@
   \SetAlgoLined
-  \caption{Sparse Subspace Clustering using LASSO \cite{jmlr-v28-wang13}.}
+  \caption{Sparse Subspace Clustering using LASSO.}
   \KwData{Adjacency matrix $A$, number of communities $K$,
@@
 is drawn from \(P\). OSC is then used for community detection, and this
-method is compared against SSC \cite{noroozi2019estimation,soltanolkotabi2014}
-and MM \cite{igraph, 307cbeb9b1be48299388437423d94bf1}. 
+method is compared against SSC (\cite{noroozi2019estimation,soltanolkotabi2014})
+and MM (\cite{igraph, 307cbeb9b1be48299388437423d94bf1}). 
 True community labels are used
@@
 \(\{\lambda^{(k \ell)}\}_K\), and this method is then compared against an
-MLE-based estimator described in \cite{noroozi2019estimation} and
-\cite{307cbeb9b1be48299388437423d94bf1}.
+MLE-based estimator described in \citet{noroozi2019estimation} and
+\citet{307cbeb9b1be48299388437423d94bf1}.
 
@@
 \citet{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
-(EP) algorithm \cite{le2016}, which is \(O(n^{K - 1})\), as a greedy
+(EP) algorithm (\cite{le2016}), which is \(O(n^{K - 1})\), as a greedy
 relaxation of the optimization problem. For these simulations, the
@@
 
-In this section, our algorithms were applied
-
-In the first real data example, we applied OSC to the Leeds Butterfly
-dataset \cite{Wang_2018} consisting of visual similarity measurements
+In the first example, we applied OSC to the Leeds Butterfly
+dataset \citep{Wang_2018} consisting of visual similarity measurements
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \cite{noroozi2019estimation}: Only the $K=4$ most
+the example from \citet{noroozi2019estimation}: Only the $K=4$ most
 frequent species were considered, and the similarities were discretized
@@
 
-\caption{Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
+\caption{Adjacency matrix of the Leeds Butterfly dataset 
+after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
 \end{figure}
@@
 In the second example, we applied OSC to the British MPs Twitter network
-\cite{greene2013producing}, the Political Blogs network
-\cite{10.1145/1134271.1134277}, and the DBLP network
-\cite{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
+\citep{greene2013producing}, the Political Blogs network
+\citep{10.1145/1134271.1134277}, and the DBLP network
+\citep{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
 analysis, we subsetted the data as described in
-\cite{307cbeb9b1be48299388437423d94bf1} for their analysis of the
+\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
 same networks. Our methods underperformed compared to modularity
@@
 Similarly, the SBM is a RDPG for which each community corresponds to a point 
-(zero-dimensional subspace) \cite{lyzinski2014}, the DCBM is a RDPG 
+(zero-dimensional subspace) \citep{lyzinski2014}, the DCBM is a RDPG 
 for which each community corresponds to a line (one-dimensional subspace) 
-\cite{rubindelanchy2017consistency}, 
+\citep{rubindelanchy2017consistency}, 
 and the PABM is a GRDPG for which each community is a $K$-dimensional subspace.
@@
 The latent structure of the PABM can be recovered 
-via Adjacency Spectral Embedding \cite{rubindelanchy2017statistical}, 
+via Adjacency Spectral Embedding \citep{rubindelanchy2017statistical}, 
 which allows us to take advantage of this latent structure 
@@
 This is a stronger result than previously proposed algorithms
-\cite{307cbeb9b1be48299388437423d94bf1},
+\citep{307cbeb9b1be48299388437423d94bf1},
 which only guarantee that the error rate (and not count) 
@@
 One proposed approach to community detection for the PABM 
-is an application of Sparse Subspace Clustering \cite{noroozi2019estimation}. 
+is an application of Sparse Subspace Clustering \citep{noroozi2019estimation}. 
 We show that the latent structure of the PABM leads to 
@@
 \renewcommand\refname{References}
-  \bibliography{misc.bib}
+% \bibliographystyle{plain}
+% \bibliography{misc.bib}
+\printbibliography
 
@@
 $\ell_2$ norm difference between the rows of $\hat{V}_n$ and
-$V_n$. See \cite{cape_biometrika} and 
-\cite[Lemma~5]{rubindelanchy2017statistical} for a proof. 
+$V_n$. See \citet{cape_biometrika} and 
+\citet[Lemma~5]{rubindelanchy2017statistical} for a proof. 
 \begin{lemma}
@@
 % notion of the inradius of a set of points. 
-\begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
+\begin{definition}[Inradius \citep{soltanolkotabi2012, jmlr-v28-wang13}]
   \label{inradius}
@@
 Furthermore, by looking at the proof of Theorem~5 in
-\cite{rubindelanchy2017statistical}, we see that $W^{*}$ is also
+\citep{rubindelanchy2017statistical}, we see that $W^{*}$ is also
 blocks diagonal with $2$ blocks where the positive eigenvalues of $M^{(k \ell)}$

### 2021-08-19 summary.tex
fixed bibliography
updated per discussion on 8/18

@@
 % \usepackage{natbib}
-\usepackage[backend=bibtex,style=authoryear,giveninits,natbib]{biblatex}
+\usepackage[backend=bibtex,style=authoryear,giveninits=true,natbib]{biblatex}
+\DeclareNameAlias{sortname}{last-first}
 \addbibresource{misc.bib}
@@
 by demonstrating that a PABM is a GRDPG in which communities correspond to 
-mutually orthogonal subspaces of latent vectors. This insight
-leads to the construction of improved algorithms for community detection
-and parameter estimation with PABM. Using established asymptotic
-properties of Adjacency Spectral Embedding for GRDPG, we derive
-asymptotic properties of these algorithms, including algorithms that
-rely on Sparse Subspace Clustering. We illustrate these properties
-via simulation study.
+mutually orthogonal subspaces of latent vectors. 
+This insight allows us to construct new algorithms for community detection
+and parameter estimation with the PABM, 
+as well as improve an existing algorithm 
+that relies on Sparse Subspace Clustering. 
+Using established asymptotic properties of 
+Adjacency Spectral Embedding for the GRDPG, 
+we derive asymptotic properties of these algorithms. 
+In particular, we demonstrate that the absolute number of 
+community detection errors tends to zero as 
+the number of graph vertices tends to infinity. 
+Simulation experiments illustrate these properties. 
 \end{abstract}
@@
 Statistical inference on random graphs requires a suitable probability model.
-A general probability model for unweighted and undirected graphs is the Bernoulli Graph, which assumes that edges occur as independent Bernoulli trials.
+A general probability model for unweighted and undirected graphs is the Bernoulli Graph (also known as the inhomogeneous Er\"{o}s-R\'{e}nyi model), which assumes that edges occur as independent Bernoulli trials.
 A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
@@
 We can use Adjacency Spectral Embedding (ASE) 
-\cite{rubindelanchy2017statistical} to recover the latent vectors of a GRDPG. 
+\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
 This procedure consists of taking the spectral decomposition of $A$ 
@@
 The latent vectors that produce $X I_{p,q} X^\top = P$ are not unique
-\cite{rubindelanchy2017statistical}.
+\citep{rubindelanchy2017statistical}.
 More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
-then we also have for any$Q \in \mathbb{O}(p, q)$,
+then we also have for any $Q \in \mathbb{O}(p, q)$,
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
@@
 corresponding to one community and hence \(X (X^\top X)^{-1} X^\top\) is also a
-block diagonal matrix with each block corresponding to a community. As $B = n VV^{\top} = n X (X^\top X)^{-1} X^\top$, we
-conclude that $B_{ij} = 0$ whenever vertices $i$ and $j$ belong to different communities.  
+block diagonal matrix with each block corresponding to a community. 
+As $B = n VV^{\top} = n X (X^\top X)^{-1} X^\top$, 
+we conclude that $B_{ij} = 0$ 
+whenever vertices $i$ and $j$ belong to different communities.  
 \end{proof}
@@
 (Alg. 1). 
+In order to show that OSC is consistent, 
+we introduce the sparsity factor $\rho_n \in (0, 1]$, 
+which can be characterized as the average expected degree of the vertices 
+divided by the number of vertices. 
+The situation is the same as in \cite{307cbeb9b1be48299388437423d94bf1} (PABM) 
+and \cite{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
+are scaled as $\rho_n^{1/2} \lambda^{(k \ell)}$. 
+For ease of notation, we restrict our results to the case in which all 
+nonzero popularity parameters are greater than some positive minimum value.
 
@@
 \end{theorem}
-
-The sparsity factor $\rho_n$ can be characterized as 
-the average expected degree of the vertices of the graph 
-divided by the number of vertices. 
-Thus Theorem \ref{theorem4} requires the average degree 
-to grow faster than $(\log n)^{4c}$.
 
@@
 
-The central result of this paper illustrates that the 
-Popularity Adjusted Block Model is a special case of the 
-Generalized Random Dot Product Graph and how this connection 
-can be exploited for community detection and parameter estimation. 
-Specifically, the PABM with $K$ communities is a GRDPG 
-for which each community is represented by a $K$-dimensional subspace 
-in the latent space, and the subspaces are orthogonal to each other. 
-This result can be viewed as a continuation of previous results 
-that connect the Stochastic Block Model and the Degree Corrected Block Model 
-to the Random Dot Product Graph: 
-The SBM, DCBM, and PABM are successive generalizations of Block Models. 
-Similarly, the SBM is a RDPG for which each community corresponds to a point 
-(zero-dimensional subspace) \citep{lyzinski2014}, the DCBM is a RDPG 
-for which each community corresponds to a line (one-dimensional subspace) 
-\citep{rubindelanchy2017consistency}, 
-and the PABM is a GRDPG for which each community is a $K$-dimensional subspace.
-
-The latent structure of the PABM can be recovered 
-via Adjacency Spectral Embedding \citep{rubindelanchy2017statistical}, 
-which allows us to take advantage of this latent structure 
-for statistical inference. 
-Using the fact that the communities are represented by orthogonal subspaces, 
-we propose Orthogonal Spectral Clustering for community detection, 
-for which we have the theoretical guarantee that 
-the number of misclustered vertices approaches zero 
-with high probability as the size of the graph increases. 
-This is a stronger result than previously proposed algorithms
-\citep{307cbeb9b1be48299388437423d94bf1},
-which only guarantee that the error rate (and not count) 
-approaches zero asymptotically.
-Parameter estimation can be performed in a similar fashion using the ASE, 
-for which we also prove that 
-the per-parameter error approaches zero asymptotically. 
-Both of these results are demonstrated via simulation study.
-
-One proposed approach to community detection for the PABM 
-is an application of Sparse Subspace Clustering \citep{noroozi2019estimation}. 
-We show that the latent structure of the PABM leads to 
-both a straightforward justification for SSC for the PABM 
-as well as an improvement over the previous implementation.
-Our approach applies SSC on the ASE, 
-and we prove that the ASE of the PABM obeys the Subspace Detection Property
-with high probability if the graph is large. 
-This result is also demonstrated via simulation study.
+Our central result states that the Popularity Adjusted Block Model 
+is a special case of the Generalized Random Dot Product Graph. 
+In particular, the PABM with $K$ communities is a GRDPG 
+for which the communities are represented by 
+mutually orthogonal $K$-dimensional subspaces of the latent space. 
+This result extends previous results that connected the Stochastic Block Model and the Degree Corrected Block Model to Random Dot Product Graphs. 
+Replacing RDPGs with GRDPGs is a critical step in this line of research, as a PABM is not necessarily a RDPG.
+
+Because all Bernoulli Graphs are GRDPGs, it should be possible to invent and study new families of Bernoulli Graphs by characterizing them as special cases of GRDPGs and expliting the latent structures that define them.
+The present work illustrates the powr of this approach. 
+We recover the latent structure of the PABM by Adjacency Spectral Embedding, then exploit that structure to improve statistical inference. 
+Exploiting the fact that PABM communities correspond to orthogonal subspaces, we propose Orthogonal Spectral Clustering for community detection and demonstrate that the number of misclassified vertices approaches zero with high probability as the size of the graph increases.
+This is a stronger result than previously proposed algorithms \citep{307cbeb9b1be48299388437423d94bf1}, which only guarantee that the error rate (and not count) approaches zero asymptotically. 
+Parameter estimation can be performed in a similar fashion using the ASE, for which we also prove that the per-parameter error approaches zero asymptotically.
+
+A secondary benefit of the GRDPG approach is that the latent structure may be used to improve existing algorithms. 
+For example, one algorithm for PABM community detection \citep{noroozi2019estimation} relies on Sparse Subspace Clustering. 
+The latent structure of the PABM provides a natural justification for SSC for the PABM and leads to an improvement over the previous implementation. 
+The improved algorithm applies SSC to the ASE, and we prove that the ASE of the PABM obeys the Subspace Detection Property with high probability if the graph is large.
+
+Finally, one might well inquire what one gains and what one sacrifices by assuming that a Bernoulli Graph is a PABM. 
+The GRDPG model offers a plausible way to pursue this inquiry. 
+Absent a known latent structure that can be exploited by specialized methods, the GRDPG-ASE approach transforms the problem of network community detection to the much-studied problem of clustering vectors in Euclidean space.
+Communities of vertices are defined as clusters of latent vectors. 
+After ASE, a standard clustering algorithm, e.g., single linkage, is used ot infer the communities. 
+In future research, we intend to use such general algorithms as baselines and measure the efficiency of the PABM algorithms (and other specialized algorithms) by studying how much they improve on general algorithms when the specified latent structure obtains.
 
@@
 
-% Our proof is based on Theorem~6 in \cite{jmlr-v28-wang13} that provide
-% sufficient conditions under which sparse subspace clustering based on
-% solving the optimization problem in Eq.~\eqref{eq:ssc} yields an
-% affinity matrix $B = |C| + |C^{\top}|$ satisfying the subspace
-% detection property of Definition~\ref{def:subspace_detection}.  
-% \begin{theorem}
-%   Let $d \geq 1$ and let $\mathcal{S}_1, \mathcal{S}_2, \dots, \mathcal{S}_{L}$ be
-%   subspaces of $\mathbb{R}^{d}$. 
-% Let $Y = [y_1 \mid y_2 \mid \dots \mid y_n]^{\top}$ be a $n \times d$ matrix
-% with rows $y_i \in \mathcal{S}_1 \cup \mathcal{S}_2 \cup
-% \dots \mathcal{S}_{L}$. Let $x_i = y_i + z_i$ and suppose that there
-% exists a finite $\delta \geq 0$ such that $\|z_i\| \leq \delta$ for
-% all $i$.
-% \end{theorem}
-
-% Our proof leverages
-% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
-% guaranteeing the subspace detection property of a noisy
-
-% Note that the rows
-% of $\sqrt{n}V$ satisfy the subspace detection property in
-% Definition~\ref{def:subspace_detection}. Now, by 
-% Lemma~\ref{lem:technical}, there exists an orthogonal transformation
-% $W$ such that the rows of $\sqrt{n}\hat{V}_n W$ are
-% uniformly close to the rows of $\sqrt{n} V_n$.
-
-% Our proof leverages
-% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
-% guaranteeing the subspace detection property of a noisy
-
-% together with the uniform row-wise
-% error bound between $\hat{V}$ and $V$ as given in
-% Lemma~\ref{lem:technical}.
-
-% We first recall the
-% notion of the inradius of a set of points. 
-% \begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
-% The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
-% defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
-% In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents
-% the inradius of the symmetric convex hull of $X$.
-% \end{definition}
-
-% The next
-
-% \begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
-% \end{definition}
-
-% \begin{lemma}
-% \label{lemma2} % Let $\hat{V}$ be the eigenvectors of $A_n$
-% % corresponding to the $K^2$ largest eigenvalues in modulus. Let
-% % $\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
-% % $\hat{V}$ corresponding to the $k^{th}$
-% % community and let $\hat{V}^{(-k)}$ denote the matrix formed by
-% % omitting the rows of $\hat{V}$
-% % corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
-% % the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
-% % the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
-% % $v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
-% % $P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
-% % rows of $V^{(k)}$.
-
-% % Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
-% % 1,2,\dots,n_{k}$ as the solution of the following optimization problem
-% % $$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
-% %   \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
-% %   \eta\|_\infty \leq 1$.}$$
-% % Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
-% % be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
-% % $\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
-% % as
-% % $$w_i^{(k)} =
-% % \frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
-% % Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
-% % and define the subspace incoherence for $W^{(k)}$ by
-% % $$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
-% Recall the notations in Definition~\ref{def:subspace_incoherence}. 
-% We then have
-% \begin{equation} \label{eq:mu-conv}
-% \mu^{(k)} = 0, \quad \text{for all $k = 1,2,\dots,K$.}
-% \end{equation}
-% \end{lemma}
-
-% \begin{proof} From Theorem \ref{theorem3} we have that the subspaces
-%   $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
-%   are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
-%   \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
-%   \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
-%   $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
-%     $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
-%     0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
-%     \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ as desired.
-% %   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
-% % must be orthogonal to each row of \(V^{(-k)}\) and hence
-% %\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
-% \end{proof}
-
-% \begin{lemma}
-% \label{lemma3}
-% Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-% $\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},
-
-% \begin{equation} \label{eq:delta-conv}
-% \delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
-% \end{equation}
-% \end{lemma}
-
 With the above definitions in place, we are now ready to state our
@@
 the following two conditions are satisfied
+
 \begin{gather}
@@
 \end{gather}
-% Furthermore, we can also choose the weighting parameter $\lambda$ in
-% the range
-% $$\frac{1}{r_* - 2 \delta - \delta^2} \leq \lambda \leq \min_{k}
-% \frac{r_k - \mu_k - 2 \delta_1}{\delta(1 + 
-% In other words, $\sqrt{n}\hat{V}_n$ 
-% small enough and the
-% subspace inradius is greater than the subspace incoherence for each
-% community \(k\).
+
 We now verify that for sufficiently large n, Eq.~\eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
 holds with high probability.
-% satisfied with high probability. 
-% Indeed, let
-% The basis of this proof is Theorem 6 from \citet{jmlr-v28-wang13}, which states that the
-% subspace detection property holds if the noise in $\sqrt{n}\hat{V}$
-% as an estimate for $\sqrt{n} V$ is small enough and that the inradius
-% of the $\hat{V}^{(k)}$ are greater 
-% small enough and the
-% subspace inradius is greater than the subspace incoherence for each
-% community \(k\).
- % , let \(v_{-i}^{(k)}\) be the matrix \(V^{(k)}\) with the
-% \(i\)\textsuperscript{th} row removed. Then
 
@@
 $\mathrm{span}(V_{-i}^{(k)}) = \mathcal{S}^{(k)}$ for all $i$ and hence
-% spans its corresponding subspace (Theorem
-% \ref{theorem2}). 
-% We therefore have, for sufficiently large $n$, that 
 \(r^{(k)} = \min_{i} r(V_{-i}^{(k)}) > 0\) for
-all $k = 1,2,\dots,K$. % Eq.~\eqref{eq:cond1} then follows from the fact that $\mu^{(k)} = 0$ for all $k = 1,2,\dots,K.$
-% Indeed, from 
+all $k = 1,2,\dots,K$.
+
 Next, by Theorem \ref{theorem3} we have that the subspaces
@@
 
-% Lemma~\ref{lemma2} then implies, with high
-% Combining the above two facts, we have, for sufficiently large, that $n$ \(r^{(k)} > \mu_n^{(k)} = 0\) for all $k = 1,2\dots,K$.
-% where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
-% such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).
-
 {\bf Verifying Eq.~\eqref{eq:cond2}}.
@@
 asymptotically almost surely.
-% \(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
-% probability 1.
 

### 2021-08-24 summary.tex
fixed a couple typos

@@
 This insight allows us to construct new algorithms for community detection
-and parameter estimation with the PABM, 
+and parameter estimation for the PABM, 
 as well as improve an existing algorithm 
@@
 Statistical inference on random graphs requires a suitable probability model. 
-A general probability model for unweighted and undirected graphs is the Bernoulli Graph (also known as the inhomogeneous Er\"{o}s-R\'{e}nyi model), which assumes that edges occur as independent Bernoulli trials.
+A general probability model for unweighted and undirected graphs is the Bernoulli Graph (also known as the inhomogeneous Erd\"{o}s-R\'{e}nyi model), which assumes that edges occur as independent Bernoulli trials.
 A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
-A trivial example of a Bernoulli Graph is the Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
+A trivial example of a Bernoulli Graph is the (homogeneous) Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
 The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
@@
 A Random Dot Product Graph (RDPG) specifies that each vertex corresponds to a latent position vector in Euclidean space and that the probability of an edge between two vertices is the dot product of their latent position vectors.
-Thus, if the latent positions $x_1,\ldots,x_n \in \mathbb{R}^d$ and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$, then the edge probability matrix is $P = XX^\top$.
+Thus, if the latent positions are $x_1,\ldots,x_n \in \mathbb{R}^d$ and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$, then the edge probability matrix is $P = XX^\top$.
 Clearly, any Bernoulli Graph with positive semidefinite $P$ is an RDPG.
@@
 Section \ref{discussion} concludes.
-
-\begin{comment}
-
-Statistical inference on random graphs involves the
-partitioning of a graph into disconnected subgraphs or clusters. This is
-often motivated by the assumption that there exist underlying and
-unobserved communities to which each vertex of the graph belongs, and
-edges between pairs of vertices are determined by drawing from a
-probability distribution based on the community relationships between
-each pair. The goal of this analysis then is population community
-detection, or the recovery of the true underlying community labels for
-each vertex, up to permutation, with some additional parameter
-estimation being of possible interest, assuming some underlying
-probability model.
-
-A general family of models that describe unweighted and undirected graphs is 
-the Bernoulli Graph, which is characterized by edge probability matrix $P$. 
-Edges between each pair of vertices $i$ and $j$ are drawn 
-as a Bernoulli trial with probability $P_{ij}$, for each 
-$0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph. 
-Community detection methods often restrict this type of model 
-by assigning each vertex a community label $z_1, ..., z_n$ 
-and then setting each $P_{ij}$ as a function of the labels $z_i$ and $z_j$. 
-Such models are called Block Models.
-
-One popular type of Block Model is the Stochastic Block Model (SBM),
-first proposed by \citet{doi:10.1080/0022250X.1971.9989788}, 
-which restricts each edge probability $P_{ij}$ as only depending on the labels 
-$z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_i, z_j}$. 
-Various generalizations of the SBM have since been utilized, 
-including the Degree-Corrected Block Model (DCBM), 
-introduced by \citet{Karrer_2011}, 
-which assigns an additional parameter $\theta_i$ to each vertex 
-and sets $P_{ij} = \theta_i \theta_j \omega_{z_i, z_j}$. 
-The Popularity Adjusted Block Model (PABM) was then introduced by
-\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the
-DCBM to address the heterogeneity of edge probabilities within and
-between communities while still maintaining distinct community
-structure.
-
-Parallel to the families of Block Models is another type of Bernoulli Graph 
-called the Random Dot Product Graph (RDPG), proposed by
-\citet*{10.1007/978-3-540-77004-6_11}. 
-Under this model, each vertex of the graph
-can be represented by a vector in some latent space such that the edge
-probability between any pair of vertices is given by their corresponding
-dot product in the latent space, i.e., given a latent positions
-\(x_1, ..., x_n \in \mathbb{R}^d\), the edge probability matrix is
-\(P = X X^\top\) where 
-\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\). 
-It can be shown that all Bernoulli Graphs with 
-positive semidefinite edge probability matrix $P$ is a type of RDPG. 
-For instance, the assortative
-SBM is equivalent to a special case of the RDPG for which all vertices
-of a given community share the same position in the latent space
-\cite{lyzinski2014}. A similar property exists for the DCBM, 
-which is equivalent to a RDPG in which vertices that belong to the same 
-community all lie in a line segment in the latent space 
-\cite{lyzinski2014, rubindelanchy2017consistency}. 
-Because the edge probability matrix for the PABM is not positive semidefinite, 
-we instead show that it is a special case of the 
-\emph{Generalized} Random Dot Product Graph (GRDPG) 
-in order to analyze the PABM as a latent space model 
-for community detection and parameter estimation.
-
-\end{comment}
 
@@
 
-for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{1 \leq
-  s \leq K, 1 \leq t \leq K}$ where each
+for a set of $K^2$ popularity vectors 
+$\{\lambda^{(st)}\}_{1 \leq s \leq K, 1 \leq t \leq K}$ where each
 $\lambda^{(st)}$ is a column vector of dimension $n_s$. Then a graph $G$
@@
 $\{\lambda^{(st)}\}$ if its adjacency matrix $A$
-satisfies $A \sim \mathrm{Bernoulli}(P)$.
+satisfies $A \sim \mathrm{BernoulliGraph}(P)$.
 \end{definition}
@@
 for which the communities are represented by 
-mutually orthogonal $K$-dimensional subspaces of the latent space. 
-This result extends previous results that connected the Stochastic Block Model and the Degree Corrected Block Model to Random Dot Product Graphs. 
-Replacing RDPGs with GRDPGs is a critical step in this line of research, as a PABM is not necessarily a RDPG.
+mutually orthogonal $K$-dimensional subspaces 
+of the $K^2$-dimensional latent space. 
+This result extends previous results that connected 
+the Stochastic Block Model and the Degree Corrected Block Model 
+to Random Dot Product Graphs. 
+Replacing RDPGs with GRDPGs is a critical step in this line of research, 
+as a PABM is not necessarily a RDPG.
 
 Because all Bernoulli Graphs are GRDPGs, it should be possible to invent and study new families of Bernoulli Graphs by characterizing them as special cases of GRDPGs and expliting the latent structures that define them.
-The present work illustrates the powr of this approach. 
+The present work illustrates the power of this approach. 
 We recover the latent structure of the PABM by Adjacency Spectral Embedding, then exploit that structure to improve statistical inference. 

### 2021-08-24 summary.tex
Premerge

--- 2021-08-02 summary.tex
@@
 \usepackage{mathtools}
-% \usepackage{natbib}
-\usepackage[backend=bibtex,style=authoryear,giveninits,natbib]{biblatex}
-\addbibresource{misc.bib}
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
+\usepackage{natbib}
 \usepackage{parskip}
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-% \setcitestyle{numbers,square,comma}
+\setcitestyle{numbers,square,comma}
 \usepackage{verbatim}
@@
 \usepackage{comment}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 \addtolength{\oddsidemargin}{-.5in}%
@@
 
-\title{Popularity Adjusted Block Models are Generalized Random Dot Product Graphs}
-
-\author{John Koo\\Department of Statistics, Indiana University\\
-and\\
-Minh Tang\\Department of Statistics, North Carolina State University\\
-and\\
-Michael Trosset\\Department of Statistics, Indiana University}
-
-% \date{\vspace{-2.5em}}
-\date{}
+\title{Connecting the Popularity Adjusted Block Model to the Generalized Random
+Dot Product Graph for Community Detection and Parameter Estimation}
+\author{}
+\date{\vspace{-2.5em}}
 
 \begin{document}
-
 \maketitle
-
 \begin{abstract}
@@
 (PABM) and the Generalized Random Dot Product Graph (GRDPG),
-by demonstrating that a PABM is a GRDPG in which communities correspond to 
-mutually orthogonal subspaces of latent vectors. This insight
+demonstrating that a PABM is a GRDPG in which communities correspond to
+certain mutually orthogonal subspaces of latent positions. This insight
 leads to the construction of improved algorithms for community detection
 and parameter estimation with PABM. Using established asymptotic
-properties of Adjacency Spectral Embedding for GRDPG, we derive
+properties of Adjacency Spectral Embedding (ASE) for GRDPG, we derive
 asymptotic properties of these algorithms, including algorithms that
-rely on Sparse Subspace Clustering. We illustrate these properties
-via simulation study.
+rely on Sparse Subspace Clustering (SSC). We illustrate these properties
+via simulation.
 \end{abstract}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\newcommand{\diag}{\mathrm{diag}}
-\newcommand{\tr}{\mathrm{Tr}}
-\newcommand{\blockdiag}{\mathrm{blockdiag}}
-\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
-\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
-\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
-\newcommand{\Betadist}{\mathrm{Beta}}
-\newcommand{\BG}{\mathrm{BernoulliGraph}}
-\newcommand{\PABM}{\mathrm{PABM}}
-\newcommand{\RDPG}{\mathrm{RDPG}}
-\newcommand{\GRDPG}{\mathrm{GRDPG}}
-\newcommand{\Multinomial}{\mathrm{Multinomial}}
+\newcommand{\diag}{\text{diag}}
+\newcommand{\tr}{\text{Tr}}
+\newcommand{\blockdiag}{\text{blockdiag}}
+\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
+\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
+\newcommand{\Bernoulli}{\text{Bernoulli}}
+\newcommand{\Betadist}{\text{Beta}}
+\newcommand{\BG}{\text{BernoulliGraph}}
+\newcommand{\PABM}{\text{PABM}}
+\newcommand{\RDPG}{\text{RDPG}}
+\newcommand{\GRDPG}{\text{GRDPG}}
+\newcommand{\Multinomial}{\text{Multinomial}}
 \newtheorem{theorem}{Theorem}
@@
 \newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
+\newtheorem*{example}{Example}
 \newpage
@@
 
-Statistical inference on random graphs requires a suitable probability model.
-A general probability model for unweighted and undirected graphs is the Bernoulli Graph, which assumes that edges occur as independent Bernoulli trials.
-A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
-A trivial example of a Bernoulli Graph is the Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
-The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
-The present work relates two lines of generalization.
-
-Network analysis is often concerned with community detection.
-One form of community detection assumes that each vertex belongs to an unobserved community, with the probability of an edge between vertices $i$ and $j$ depending on the communities to which $i$ and $j$ belong.
-Formally, one assigns each vertex $v_i$ a community label $z_i$ and assumes a Bernoulli Graph in which $P_{ij}$ is a function of $z_i$ and $z_j$.
-Such models, called Block Models, define the goal of community detection as a problem in statistical inference: identify the true community (up to permutation of labels) to which each vertex belongs.
-
-The classical Stochastic Block Model (SBM) of \citet{doi:10.1080/0022250X.1971.9989788} specifies that each edge probability $P_{ij}$ depends only on the labels $z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_,z_j}$.
-Subsequent researchers have weakened this assumption.
-The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_,z_j}$.
-The Popularity Adjusted Block Model (PABM) of \citet{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
-
-Another type of Bernoulli Graph was proposed by \citet*{10.1007/978-3-540-77004-6_11}.
-A Random Dot Product Graph (RDPG) specifies that each vertex corresponds to a latent position vector in Euclidean space and that the probability of an edge between two vertices is the dot product of their latent position vectors.
-Thus, if the latent positions $x_1,\ldots,x_n \in \mathbb{R}^d$ and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$, then the edge probability matrix is $P = XX^\top$.
-Clearly, any Bernoulli Graph with positive semidefinite $P$ is an RDPG.
-The positive definite Euclidean inner product in the RDPG model was replaced by an indefinite inner product in \citet{rubindelanchy2017consistency}, resulting in the {\em Generalized}\/ RDPG (GRDPG).
-
-In contrast to Block Models, neither RDPGs nor GRDPGs inherently specify distinct communities.
-However, one can easily impose community structure by assuming that the latent positions lie in distinct clusters.
-Hence, it is not surprising that Block Models can be studied by reformulating them as RDPGs or GRDPGs.
-For example, an assortative SBM (an SBM for which $P$ is positive semidefinite) is equivalent to an RDPG for which all vertices in the same community correspond to the same latent position vector (\cite{lyzinski2014}).
-Likewise, the DCBM is equivalent to an RDPG for which all vertices in the same community correspond to latent position vectors that lie on a straight line (\cite{lyzinski2014, rubindelanchy2017consistency}).
-
-Because the edge probability matrix of a PABM is not necessarily positive semidefinite, a PABM is not necessarily an RDPG.
-In Section \ref{connecting-the-pabm-to-the-grdpg} we demonstrate that every PABM is in fact a specific type of GRDPG for which the latent position vectors lie in distinct orthogonal subspaces, each subspace corresponding to a community.
-This identification is our central result.
-In Section \ref{methods}, we use the geometry of the GRDPG to derive more efficient algorithms for detecting the communities and estimating the parameters in the PABM.
-We report the results of simulation studies in Section \ref{simulated-examples} and apply our methods to three well-known data sets in Section \ref{real-data-examples}.
-Section \ref{discussion} concludes.
-
-\begin{comment}
-
-Statistical inference on random graphs involves the
+Statistical analysis on graphs or networks often involves the
 partitioning of a graph into disconnected subgraphs or clusters. This is
@@
 
-A general family of models that describe unweighted and undirected graphs is 
-the Bernoulli Graph, which is characterized by edge probability matrix $P$. 
-Edges between each pair of vertices $i$ and $j$ are drawn 
-as a Bernoulli trial with probability $P_{ij}$, for each 
-$0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph. 
-Community detection methods often restrict this type of model 
-by assigning each vertex a community label $z_1, ..., z_n$ 
-and then setting each $P_{ij}$ as a function of the labels $z_i$ and $z_j$. 
+A general family of models that generate unweighted and undirected graphs is the Bernoulli Graph, which is described by an edge probability matrix $P$.
+Edges between each pair of vertices $i$ and $j$ is drawn as a Bernoulli trial with probability $P_{ij}$, for each $0 \leq i < j \leq n$ and $n$ is the number of vertices in the graph.
+Community detection methods often restrict this type of model by assigning each vertex a community label $z_1, ..., z_n$ and then setting each $P_{ij}$ as a function of the labels of vertices $i$ and $j$.
 Such models are called Block Models.
@@
 including the Degree-Corrected Block Model (DCBM), 
-introduced by \citet{Karrer_2011}, 
+introduced by\citet{Karrer_2011}, 
 which assigns an additional parameter $\theta_i$ to each vertex 
@@
 called the Random Dot Product Graph (RDPG), proposed by
-\citet*{10.1007/978-3-540-77004-6_11}. 
-Under this model, each vertex of the graph
-can be represented by a vector in some latent space such that the edge
+\citet*{10.1007/978-3-540-77004-6_11}. Under this model, each vertex of the graph
+can be represented by a point in some latent space such that the edge
 probability between any pair of vertices is given by their corresponding
@@
 of a given community share the same position in the latent space
-\cite{lyzinski2014}. A similar property exists for the DCBM, 
-which is equivalent to a RDPG in which vertices that belong to the same 
-community all lie in a line segment in the latent space 
+\cite{lyzinski2014}. A similar property exists for the DCBM
 \cite{lyzinski2014, rubindelanchy2017consistency}. 
 Because the edge probability matrix for the PABM is not positive semidefinite, 
-we instead show that it is a special case of the 
-\emph{Generalized} Random Dot Product Graph (GRDPG) 
-in order to analyze the PABM as a latent space model 
-for community detection and parameter estimation.
-
-\end{comment}
-
-\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
-\section{PABMs are GRDPGs}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
-
-In this section, we show that the PABM is a special case of the GRDPG. 
-More specifically, graph $G$ drawn from the PABM can be represented by 
-a collection of latent vectors in Euclidean space. 
-We further show that the latent configuration that
-induces the PABM consists of orthogonal subspaces with each subspace
-corresponding to a community.
+we instead show that it is a special case of the \emph{Generalized} Random Dot Product Graph (GRDPG) in order to analyze the PABM as a latent space model for community detection and parameter estimation.
 
@@
 
-Let $G = (V, E)$ be an unweighted, undirected, 
-and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. 
-$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ 
-such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ 
-and $0$ otherwise. Because $G$ is symmetric and hollow,
+Let $G = (V, E)$ be an unweighted, undirected, and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$.
+$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ and $0$ otherwise.
+Because $G$ is symmetric and hollow,
 $A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$.
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric matrix of edge probabilities.
-Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ 
-for each $1 \leq i < j \leq n$ (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$). 
-We denote $A \sim \BG(P)$ as a graph with adjacency matrix $A$ 
-sampled from edge probability matrix $P$ in this manner. 
-If each vertex has a hidden label in $[K]$, 
-they are denoted as $z_1, ..., z_n$. 
-Finally, we denote 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ 
-as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
+Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ for each $1 \leq i < j \leq n$  (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$).
+We denote $A \sim \BG(P)$ as graph with adjacency matrix $A$ sampled from edge probability matrix $P$ in this manner.
+If each vertex has a hidden label in $[K]$, they are denoted as $z_1, ..., z_n$.
+Finally, we denote $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
+
+\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
+\section{Connecting the Popularity Adjusted Block Model to the
+Generalized Random Dot Product
+Graph}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
+
+In this section, we show that the PABM is a special case of the GRDPG, i.e.,
+graph $G$ drawn from the PABM can be represented by a collection of latent
+vectors in Euclidean space. We further show that the latent configuration that
+induces the PABM consists of orthogonal subspaces with each subspace
+corresponding to a community.
 
 \hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
-\subsection{Two Probability Models for Graphs}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
+\subsection{The Popularity Adjusted Block Model and the Generalized
+Random Dot Product
+Graph}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
 
@@
 Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-graph $G = (V, E)$ with adjacency matrix $A$ 
-such that $A \sim \mathrm{Bernoulli}(P)$.
+graph $G = (V, E)$ with adjacency matrix $A$ such that $A \sim \mathrm{Bernoulli}(P)$.
 Let each vertex have a community label between $1$ and $K$.
@@
 Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
-
 \begin{equation} \label{eq:pabm}
@@
 \end{equation}
-
 for a set of $K^2$ popularity vectors $\{\lambda^{(st)}\}_{1 \leq
@@
 is drawn from a PABM with parameters
-$\{\lambda^{(st)}\}$ if its adjacency matrix $A$
+$\{\lambda^{(st)}\}_K$ if its adjacency matrix $A$
 satisfies $A \sim \mathrm{Bernoulli}(P)$.
@@
 
-We will use the notation \(A \sim \PABM(\{\lambda^{(k \ell)}\}_K)\) to denote
+We will use the notation \(A \sim \PABM(\{\lambda^{(kl)}\}_K)\) to denote
 a random adjacency matrix \(A\) drawn from a PABM with parameters
-\(\{\lambda^{(k \ell)}\}\) consisting of \(K\) underlying communities.
+\(\lambda^{(kl)}\) consisting of \(K\) underlying communities.
 
@@
 Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$.
-If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
+If $\exists X \in \mathbb{R}^{n \times d}$ such that
 
@@
 for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-$G$ is drawn from the Generalized Random Dot Product Graph 
-with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
+$G$ is drawn from the Generalized Random Dot Product Graph with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
 \end{definition}
@@
 adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\). If instead of fixed latent positions, 
-they are drawn $X_1, ..., X_n \iid F$, 
-we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
-
-\begin{remark}
-We can use Adjacency Spectral Embedding (ASE) 
-\cite{rubindelanchy2017statistical} to recover the latent vectors of a GRDPG. 
-This procedure consists of taking the spectral decomposition of $A$ 
-(or $P$ if available) and choosing the $p$ most positive and $q$ most negative
-eigenvalues and their corresponding eigenvectors 
-to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
-\end{remark}
+\((p, q)\). If instead of fixed latent positions, they are drawn $X_1, ..., X_n \iid F$, we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
 
@@
 \begin{remark}
-The latent vectors that produce $X I_{p,q} X^\top = P$ are not unique
+Like the RDPG, the latent positions of a GRDPG are not unique
 \cite{rubindelanchy2017statistical}.
-More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
-then we also have for any$Q \in \mathbb{O}(p, q)$,
+More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, then we also have for any
+$Q \in \mathbb{O}(p, q)$,
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
@@
 \hypertarget{connecting-the-pabm-to-the-grdpg}{%
-\subsection{The Geometry of PABMs}\label{connecting-the-pabm-to-the-grdpg}}
-
-Now that we defined the PABM and GRDPG, 
-we show the special geometry of the PABM when viewed as a GRDPG.
-
-\begin{theorem}[The latent configuration of the PABM]
-\label{theorem2}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K)$ be an instance of a
-PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
-\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. 
-Then there exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by $\{\lambda^{(k \ell)}\}$ and a 
-$K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
-that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
-\end{theorem}
-
-\begin{proof}
-We will prove this theorem in two parts. First, for demonstration purposes, we focus on the case for $K = 2$. Then we generalize this to $K \geq 2$.
-
-For the $K = 2$ case, the proof is straightforward. Let
-
+\subsection{Connecting the PABM to the
+GRDPG}\label{connecting-the-pabm-to-the-grdpg}}
+\begin{proposition}[Connecting the PABM to the GRDPG for $K = 2$]
+\label{theorem1}
+Let
 $$X = \begin{bmatrix}
@@
 0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}.$$
-
-Then by straightforward matrix multiplication, we obtain
-
+0 & 1 & 0 & 0 \end{bmatrix},$$
+where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1.
+Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{(\lambda^{(kl)}\})$ are
+identically distributed.
+\end{proposition}
+
+\begin{proof}
+  We can show that
 \[X U I_{3, 1} U^\top X^\top =
@@
   \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}, \]
-
-which is a GRDPG with latent vectors described by $X U$.
-
-It is nevertheless instructive to look at a few intermediate steps. 
-More specifically, the product $U I_{3, 1} U^\top$ 
-yields a permutation matrix $\Pi$ with fixed points at positions $1$ and $4$ 
-and a cycle of order 2 swapping positions $2$ and $3$, i.e., 
+\end{bmatrix} \]
+by straightforward matrix multiplication. It is nevertheless
+instructive to look at a few intermediate steps. More specifically,
+the product $U I_{3, 1} U^\top$ yields a permutation matrix $\Pi$
+with fixed points at positions $1$ and $4$ and a cycle of order 2 swapping
+positions $2$ and $3$, i.e., 
 $$\Pi = U I_{3, 1} U^\top = \begin{bmatrix} 1 & 0 & 0 & 0 \\
@@
 $\tfrac{1}{\sqrt{2}}(e_{2} + e_3)$ and $\tfrac{1}{\sqrt{2}}(e_{2} -
-e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$.
-
-For the general case, we first define the following matrices
+e_3)$; here $e_i$ denote the $i$th basis vector in $\mathbb{R}^{4}$. %  and $ eigenvalues $\pm 1$ with corresponding
+% eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is
+% swapped.
+\end{proof}
+% The form for the matrix $\Pi$ used in the proof of
+% Proposition~\ref{theorem1} can 
+\begin{theorem}[Generalization to $K > 2$]
+\label{theorem2}
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\})$ be an instance of a
+PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
+\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. Then there exists a block diagonal matrix
+$X \in \mathbb{R}^{n \times K^2}$ defined by the $\{\lambda^{(kl)}\}$ and a $K^2 \times
+K^2$ {\em fixed} orthonormal matrix $U$ such 
+that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
+\end{theorem}
+
+\begin{proof}
+We first define the following matrices
 \begin{gather}
@@
 \end{gather*}
-We therefore have $P = X Y^\top$. 
-Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
+We therefore have $P = X Y^\top$. Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
 $\Pi$ and hence $P = X \Pi X^\top$.
@@
 We therefore have
-
 \begin{equation} \label{eq:permutation}
@@
 \end{equation}
-
+% \begin{equation} \label{eq:p}
+% p = K (K + 1) / 2
+% \end{equation}
+% \begin{equation} \label{eq:q}
+% q = K (K - 1) / 2
+% \end{equation}
 In summary we can describe the PABM with $K$ communities as a GRDPG with latent
@@
 \end{gather*}
-Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ 
-permutation matrix of the form
-
-$$\Pi = 
+Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ permutation matrix of
+the form
+$$\Pi = %\begin{bmatrix}
+% 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
+% 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
+% 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
+% 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1%
 \Bigl[e_1 \mid e_4 \mid e_7 \mid e_2 \mid e_5 \mid e_8 \mid e_3
-\mid e_6 \mid e_9 \Bigr].$$
-
-where $e_i$ denotes the $i^{th}$ basis vector in $\mathbb{R}^{9}$. 
+\mid e_6 \mid e_9 \Bigr].
+% \end{bmatrix}.
+$$
+where $e_i$ denote the $i$th basis vector in $\mathbb{R}^{9}$. 
 The matrix $\Pi$ corresponds to a permutation of $\{1,2,\dots,9\}$
@@
 
-The matrix $P$ is then the edge probabilities matrix for a 
-Generalized Random Dot Product Graph whose latent positions 
-are the rows of the matrix
+The matrix $P$ is then the edge probabilities matrix for generalized random dot product
+graph whose latent positions are the rows of the matrix
 $$XU = \begin{bmatrix}
   \lambda^{(11)} & 0 & 0 &
-  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 &
-  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 \\
+  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 &
+  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
   0 & \lambda^{(22)} & 0 &
-  \frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} &
-  -\frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} \\
+  \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} &
+  -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
   0 & 0 & \lambda^{(33)} &
-  0 & \frac{\lambda^{(31)}}{\sqrt{2}} & \frac{\lambda^{(32)}}{\sqrt{2}} &
-  0 & -\frac{\lambda^{(31)}}{\sqrt{2}} & -\frac{\lambda^{(32)}}{\sqrt{2}}
-\end{bmatrix}.$$
+  0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} &
+  0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
+\end{bmatrix}$$
 \end{example}
@@
 \hypertarget{methods}{%
-\section{Algorithms}\label{methods}}
+\section{Methods}\label{methods}}
 
@@
 \item
-  Parameter estimation (estimating \(\lambda^{(k \ell)}\)'s).
+  Parameter estimation (estimating \(\lambda^{(kl)}\)'s).
 \end{enumerate}
@@
 \hypertarget{related-work}{%
-\subsection{Previous Work}\label{related-work}}
-
-\citet{307cbeb9b1be48299388437423d94bf1} 
-used Modularity Maximization (MM) and the Extreme Points (EP)
-algorithm (\cite{le2016}) for community detection and parameter
+\subsection{Related work}\label{related-work}}
+The authors of \cite{307cbeb9b1be48299388437423d94bf1}, who first proposed the
+PABM, used Modularity Maximization (MM) and the Extreme Points (EP)
+algorithm \cite{le2016} for community detection and parameter
 estimation. They were able to show that as the sample size increases,
@@
 
-\citet{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
-(\cite{5206547}) for community detection in the PABM. 
-The SSC algorithm can be described as follows: 
-Given \(X \in \mathbb{R}^{n \times d}\) with vectors
+\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for
+community detection in the PABM. SSC is performed by solving an
+optimization problem for each observed point. Given
+\(X \in \mathbb{R}^{n \times d}\) with vectors
 \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
@@
 consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
-\(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
+\(x_i, x_j\) are in different subspaces. If \(X\) instead represents
 points near \(K\) subspaces with some noise, a final graph partitioning
-step may be required (e.g., edge thresholding or spectral clustering).
+step may be performed (e.g., edge thresholding or spectral clustering).
 
@@
 vectors. Let $\lambda \geq 0$ be given and
-let $C$ and $B$ be constructed from the
+let $C$ and $B$
+be constructed from the
 solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
 this given choice of $\lambda$. Then $X$
-is said to satisfy the subspace detection property 
-with sparsity parameter $\lambda$ if each column of
+is said to satisfy the subspace detection property with weighting parameter $\lambda$ if each column of
 $C$ has nonzero norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
@@
 In practice, a noisy sample $X$ often does not obey the subspace detection
-property. In such cases, $B$ is treated as an affinity matrix for a graph, 
-which requires a partitioning step to obtain the final clustering. On the other
+property. In such cases, $B$ is treated as an affinity matrix for a graph which
+is then partitioned into $K$ subgraphs to obtain the clustering. On the other
 hand, if $X$ does obey the subspace detection property, $B$ describes a graph
 with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
-property holds, there are exactly $K$ subgraphs which map onto each subspace,
+property holds, there are exactly $K$ subgraphs which map to each subspace,
 but it could be the case that some of the subspaces are represented by
@@
 to perform SSC on the ASE of \(P\) or \(A\).
-\citet{noroozi2019estimation} instead applied SSC to \(P\) and
+\cite{noroozi2019estimation} instead applied SSC to \(P\) and
 \(A\), foregoing embedding altogether.
@@
 \ref{theorem2}). Much of our work is then built on
-\citet{rubindelanchy2017statistical}, who describe the convergence
+\cite{rubindelanchy2017statistical}, who describe the convergence
 behavior of the ASE of \(A\) to the ASE of \(P\), and
@@
 \hypertarget{community-detection}{%
-\subsection{Algorithms for Community Detection}\label{community-detection}}
+\subsection{Community detection}\label{community-detection}}
 
@@
 identification or estimation of \(XU\) is not possible
-(\cite{rubindelanchy2017statistical}).
+\cite{rubindelanchy2017statistical}.
 
 If we decompose \(P = Z I_{p, q} Z^\top\), then
-\(\exists\) \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
+\(\exists Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
 start with the exact edge probability matrix, we cannot recover the
@@
 \(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
-prevents us from using the orthogonality property of \(XU\) directly.
+prevents us from using the properties of \(XU\) directly. In particular,
+if \(Q \in \mathbb{O}(n)\), then we could use the fact that
+\(\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0\) if
+vertices \(i\) and \(j\) are in different communities.
 
@@
 community is represented by a \(K\)-dimensional subspace. Therefore, the
-ASE of \(P\) still consists of subspaces 
-that correspond to communities of the PABM, even if they are not orthogonal, 
+ASE of \(P\) results in subspaces that correspond to each community,
 suggesting the use of SSC. Before exploring SSC, we will first consider
@@
 Theorem \ref{theorem3} provides perfect community detection given \(P\).
-Letting \(|B|\) be the affinity matrix for graph \(G'\), \(G'\) is
+Letting \(|B|\) be the affinity matrix for graph \(G\), \(G\) is
 partitioned into at least \(K\) disjoint subgraphs since each of the
@@
 detection property, it could be the case that some of the communities
-are represented by multiple disjoint subgraphs in \(G'\), in which case
+are represented by multiple disjoint subgraphs in \(G\), in which case
 additional reconstruction is required to identify the communities
@@
 Using \(A\) instead of \(P\) introduces error, which converges to \(0\)
-almost surely. We use this fact to motivate Orthogonal Spectral Clustering 
-(Alg. 1).
+almost surely:
 
@@
 \label{theorem4}
-Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC
+Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC
 (Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities
-and sparsity factor satisfying $n \rho_n = \omega \big( (\log n)^{4c} \big)$,
+and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,
 
 \begin{equation} \label{eq:thm4}
-\max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
+\max_{i, j} |n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| =
 O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
@@
 This provides the result that for $i, j$ in different communities,
-$\hat{B}_{ij} \stackrel{a.s.}{\to} 0$.
+$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$.
 \end{theorem}
 
-The sparsity factor $\rho_n$ can be characterized as 
-the average expected degree of the vertices of the graph 
-divided by the number of vertices. 
-Thus Theorem \ref{theorem4} requires the average degree 
-to grow faster than $(\log n)^{4c}$.
-
 Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} also provide
-a natural path toward using SSC for community detection. 
-We established in Theorem \ref{theorem2} that an ASE of the edge
-probability matrix \(P\) can be constructed from a latent vector configuration 
-consisting of orthogonal subspaces. Theorem \ref{theorem3} shows how 
-this property can be recovered from the eigenvectors of \(P\). 
-Then Theorem \ref{theorem4} shows that replacing $P$ with $A$ 
-approximates these properties with error that goes to zero as \(n \to \infty\). 
-This allows us to obtain the subspace detection property 
-in Theorem \ref{theorem5}.
+a very natural path toward using SSC for community detection for the
+PABM. We established in Theorem \ref{theorem2} that an ASE of the edge
+probability matrix \(P\) can be constructed such that the communities
+lie on mutually orthogonal subspaces, and this property can be recovered
+from the eigenvectors of \(P\). Then Theorems \ref{theorem3} and
+\ref{theorem4} show that this property holds for the unscaled ASE of
+\(A\) drawn from \(P\) as \(n \to \infty\).
 
@@
   \SetAlgoLined
-  \caption{Sparse Subspace Clustering using LASSO.}
+  \caption{Sparse Subspace Clustering using LASSO \cite{jmlr-v28-wang13}.}
   \KwData{Adjacency matrix $A$, number of communities $K$,
@@
 \label{theorem5}
-Let $P$ describe the edge probability matrix of the PABM with
-$n$ vertices, and let $A \sim \Bernoulli(P)$.  Let $\hat{V}$ be the
-matrix of eigenvectors of $A$ corresponding to the $K^2$ largest
+Let $P_n$ describe the edge probability matrix of the PABM with
+$n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the
+matrix of eigenvectors of $A_n$ corresponding to the $K^2$ largest
 eigenvalues in modulus. Then for all $\epsilon > 0$ 
 there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
-that for all $n \geq N$, $\sqrt{n} \hat{V}$ obeys the subspace detection property with
+that for all $n \geq N$, $\sqrt{n} \hat{V}_n$ obeys the subspace detection property with
 probability at least $1 - \epsilon$.  
@@
 
+\begin{remark}
+The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
+from  \cite{jmlr-v28-wang13} and the fact that the columns of
+$V_n$ correspond to orthogonal subspaces. \cite{jmlr-v28-wang13} assume
+that the points in the embedding are all of unit length, and while we apply
+this normalization in the simulated examples, it is not strictly necessary
+for Theorem \ref{theorem5} due to orthogonality.
+\end{remark}
+
 \hypertarget{parameter-estimation}{%
-\subsection{Algorithms for Parameter Estimation}\label{parameter-estimation}}
+\subsection{Parameter estimation}\label{parameter-estimation}}
 
@@
 block is an outer product of two vectors, i.e.,
-\(P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top\). Therefore, given
-\(P^{(k \ell)}\), \(\lambda^{(k \ell)}\) and \(\lambda^{(\ell k)}\) are solvable
+\(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). Therefore, given
+\(P^{(kl)}\), \(\lambda^{(kl)}\) and \(\lambda^{(lk)}\) are solvable
 up to multiplicative constant using singular value
 decomposition. More specifically, 
-let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\)
-be the singular value decomposition of \(P^{(k \ell)}\).
-\(u^{(k \ell)} \in \mathbb{R}^{n_k}\) and 
-\(v^{(k \ell)} \in \mathbb{R}^{n_\ell}\) are vectors
-and \(\sigma^{(k \ell)}\) is a scalar. 
-Then \(\lambda^{(k \ell)} = s_1 u^{(k \ell)}\)
-and \(\lambda^{(\ell k)} = s_2 v^{(k \ell)}\) 
-for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
-Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
-we instead estimate each 
-$\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}$. 
+let \(P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top\)
+be the singular value decomposition of \(P^{(kl)}\).
+\(u^{(kl)} \in \mathbb{R}^{n_k}\) and 
+\(v^{(kl)} \in \mathbb{R}^{n_l}\) are vectors
+and \(\sigma^{(kl)}\) is a scalar. Then \(\lambda^{(kl)} = s_1 u^{(kl)}\)
+and \(\lambda^{(lk)} = s_2 v^{(kl)}\) 
+for unidentifiable $s_1 s_2 = (\sigma^{(kl)})^2$.
 Given the adjacency matrix \(A\)
 instead of edge probability matrix \(P\), we can simply use plug-in
-estimators by taking the SVD of each $A^{(k \ell)}$ to obtain 
-$\hat{\lambda}^{(k \ell)} = \hat{\sigma}^{(k \ell)} \hat{u}^{(k \ell)}$ 
-using the largest signular value of $A$ and its corresponding singular vectors.
+estimators by taking the SVD of each $A^{(kl)}$. 
+Since each $\lambda^{(kl)}$ is not strictly identifiable,
+we instead estimate each $\tilde{\lambda}^{(kl)} = \sigma^{(kl)} u^{(kl)}$.
 
@@
   \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
-  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(k \ell)}\}_K$.}
+  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
   Arrange the rows and columns of $A$ by community such that each 
-  $A^{(k \ell)}$ block consists of estimated edge probabilities between 
+  $A^{(kl)}$ block consists of estimated edge probabilities between 
   communities $k$ and $l$.\;
   \For {$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
     block.\;
-    Assign $u^{(k \ell)}$ and $v^{(k \ell)}$ as the first columns of $U$ and $V$. 
-    Assign $(\sigma^{(k \ell)})^2 \leftarrow \Sigma_{11}$.\;
-    Assign $\hat{\lambda}^{(k \ell)} \leftarrow \pm \sigma^{(k \ell)} u^{(k \ell)}$ and 
-    $\hat{\lambda}^{(\ell k)} \leftarrow \pm \sigma^{(k \ell)} v^{(k \ell)}$.
+    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
+    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
+    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
+    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
   }
@@
 \label{theorem6}
-Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived from its corresponding $P^{(k \ell)}$, and let $\hat{\lambda}^{(k \ell)}$ be its estimate from $A^{(k \ell)}$ (Alg.
-3).
-Then if $n \rho_n = \omega \big( (\log n)^{4 c} \big)$,
-
+Under regularity and sparsity assumptions, given fixed $K$, 
 \begin{equation} \label{eq:thm6}
 \max_{k, l \in \{1, ..., K\}} 
-\|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
+\|\hat{\lambda}^{(kl)} - \lambda^{(kl)}\|_{\infty} = 
 O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
 \end{equation}
-
 with high probability.
@@
 \hypertarget{simulated-examples}{%
-\section{Simulation Study}\label{simulated-examples}}
+\section{Simulated Examples}\label{simulated-examples}}
 
 For each simulation, community labels are drawn from a multinomial
-distribution, the popularity vectors \(\{\lambda^{(k \ell)}\}_K\) are drawn
-from two types of joint distributions depending on whether \(k = \ell\),
+distribution, the popularity vectors \(\{\lambda^{(kl)}\}_K\) are drawn
+from two types of joint distributions depending on whether \(k = l\),
 the edge probability matrix \(P\) is constructed using the popularity
-vectors, and finally the adjacency matrix \(A\)
+vectors, and finally an unweighted and undirected adjacency matrix \(A\)
 is drawn from \(P\). OSC is then used for community detection, and this
-method is compared against SSC (\cite{noroozi2019estimation,soltanolkotabi2014})
-and MM (\cite{igraph, 307cbeb9b1be48299388437423d94bf1}). 
+method is compared against SSC \cite{noroozi2019estimation,soltanolkotabi2014}
+and MM \cite{igraph, 307cbeb9b1be48299388437423d94bf1}. 
 True community labels are used
 with Algorithm 3 to estimate the popularity vectors
-\(\{\lambda^{(k \ell)}\}_K\), and this method is then compared against an
-MLE-based estimator described in \citet{noroozi2019estimation} and
-\citet{307cbeb9b1be48299388437423d94bf1}.
+\(\{\lambda^{(kl)}\}_K\), and this method is then compared against an
+MLE-based estimator described by \cite{noroozi2019estimation} and
+\cite{307cbeb9b1be48299388437423d94bf1}.
 
 Modularity Maximization is NP-hard, so
-\citet{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
-(EP) algorithm (\cite{le2016}), which is \(O(n^{K - 1})\), as a greedy
+\cite{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
+(EP) algorithm \cite{le2016}, which is \(O(n^{K - 1})\), as a greedy
 relaxation of the optimization problem. For these simulations, the
 Louvain algorithm was used for modularity maximization,
-as \citet{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
+as \cite{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
 prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
@@
 SSC-A, treats the columns of the adjacency matrix \(A\) as points in
-\(\mathbb{R}^n\), as described in \citet{noroozi2019estimation}.
+\(\mathbb{R}^n\), as described in \cite{noroozi2019estimation}.
 The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in Algorithm 2. The sparsity
+performs SSC on the embedding, as described in algorithm 2. The sparsity
 parameter \(\lambda\) was chosen via a preliminary cross-validation
@@
 In this case, the "true" popularity vectors are derived from 
-taking the SVD of each edge probability block $P^{(k \ell)}$ 
+taking the SVD of each edge probability block $P^{(kl)}$ 
 to avoid the unidentifiable multiplicative constants.
 
-\[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}, \{\tilde{\lambda}^{(k \ell)}\}) = 
+\[\text{RMSE}(\{\hat{\lambda}^{(kl)}\}_K, \{\lambda^{(kl)}\}_K) = 
 \sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)} \|_2^2}\]
+\|s \hat{\lambda}^{(kl)} - \lambda^{(kl)} \|_2^2}\]
 
@@
 by reconstructing each 
-$\hat{P}^{(k \ell)} = \hat{\lambda}^{(k \ell)} (\hat{\lambda}^{(\ell k)})^\top$, 
+$\hat{P}^{(kl)} = \hat{\lambda}^{(kl)} (\hat{\lambda}^{(lk)})^\top$, 
 which we use to define another parameter estimation error.
@@
 \[\text{RMSE}(\hat{P}, P) =
-\sum_{k,l} \sqrt{\frac{1}{n_k n_\ell} \|P^{(k \ell)} - \hat{P}^{(k \ell)}\|_F^2}\]
+\sum_{k,l} \sqrt{\frac{1}{n_k n_l} \|P^{(kl)} - \hat{P}^{(kl)}\|_F^2}\]
 
 \hypertarget{balanced-communities}{%
-\subsection{Balanced Communities}\label{balanced-communities}}
+\subsection{Balanced communities}\label{balanced-communities}}
 
@@
 multinomial distribution with mixture parameters
-\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(k \ell)}\}_K\) according
+\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(kl)}\}_K\) according
 to the drawn community labels, \(P\) was constructed using the drawn
-\(\{\lambda^{(k \ell)}\}_K\), and \(A\) was drawn from \(P\).
+\(\{\lambda^{(kl)}\}_K\), and \(A\) was drawn from \(P\) by
+\(A_{ij} \indep \Bernoulli(P_{ij})\). Each simulation has
+a unique edge probability matrix \(P\).
 
@@
   Between-group popularities
-  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
-  \(k \neq \ell\)
+  \(\lambda^{(kl)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
+  \(k \neq l\)
 \end{itemize}
@@
 
-Fig. \ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
+\begin{comment}
+
+Fig. \ref{fig:clust_err_k} show that for large $n$, OSC results in
+a misclustering rate of 0.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_k-1}
+
+}
+
+\caption{IQR of community detection error error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+Communities are approximately balanced.
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k}
+\end{figure}
+
+Theorem \ref{theorem4} implies that OSC will result in not just in the error
+rate converging to 0 but the error *count* as well.
+We explore this in Fig. \ref{fig:clust_err_ct_sim}.
+
+\end{comment}
+
+Fig \ref{fig:clust_err_ct_sim} shows OSC's community detection error
 going to 0 for large \(n\). SSC on both the embedding and on the
-adjacency matrix produces similar trends for \(K > 2\). Weaker
+adjacency matrix produces similar results for \(K > 2\). Weaker
 performance of SSC for \(K = 2\) can be attributed to the final spectral
-clustering step of the affinity matrix. 
-While the subspace detection property is guaranteed for large \(n\), in our
-simulations, setting the sparsity parameter to the required value resulted
+clustering step of the affinity matrix. A GMM was fit to the Laplacian
+eigenmap, but visual inspection suggests that the communities are not
+distributed as a mixture of Gaussians in the eigenmap. While the
+subspace detection property is guaranteed for large \(n\), in our
+simulations, setting a large enough sparsity parameter for SSC resulted
 in more than \(K\) disconnected subgraphs.
-We instead chose a smaller sparsity parameter, 
-necessitating a final clustering step. 
-A GMM was fit to the normalized Laplacian eigenmap of $B$,
-but visual inspection suggests that the communities are not
-distributed as a mixture of Gaussians in the eigenmap. 
-A different choice of mixture distribution may result in better performance.
 
@@
 Given ground truth community labels, Algorithm 3 and the MLE-based
-plug-in estimators perform similarly, with root mean square
-error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:lambda_est_k}). 
-We observe similar behavior when comparing errors 
-in the edge probability blocks (Fig. \ref{fig:p_block_est}).
+plug-in estimators \cite{307cbeb9b1be48299388437423d94bf1} 
+perform similarly, with root mean square
+error decaying at rate approximately \(n^{-1/2}\).
 
@@
 Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.}
-\label{fig:lambda_est_k}
+Communities were drawn to be approximately balanced.}\label{fig:lambda_est_k}
 \end{figure}
+
+We observe similar behavior when comparing errors in the edge probability blocks.
 
@@
 Communities were drawn to be approximately balanced.}
-\label{fig:p_block_est}
 \end{figure}
@@
 \hypertarget{imbalanced-communities}{%
-\subsection{Imbalanced Communities}\label{imbalanced-communities}}
-
-Simulations performed in this section are the same as those in the
+\subsection{Imbalanced communities}\label{imbalanced-communities}}
+
+Simulations performed in this section are similar to those in the
 previous section with the exception of the mixture parameters
@@
 \item
-  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
+  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}\)
   for \(k = 1, ..., K\)
@@
 \(50\) simulations were performed for each \((n, K)\) pair.
+
+\begin{comment}
+
+Fig. \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show
+similar results as in the balanced communities case, with both OSC and
+SSC resulting in no misclustered vertices for a
+sufficiently large sample size. However, Fig. \ref{fig:lambda_est_k_imba}
+suggests that while Algorithm 3 retains $\sqrt{n}$ efficiency, the MLE-based
+plug-in estimator is less efficient for this setup.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/clust_err_k_imba-1}
+
+}
+
+\caption{Median and IQR of community detection error using OSC (blue) compared against SSC on the ASE of A (purple), MM (red), and SSC on the adjacency matrix (green).
+Communities are imbalanced.
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_k_imba}
+\end{figure}
+
+\end{comment}
 
@@
 if we reconstruct the edge probability blocks 
-by taking the outer products of the popularity vectors 
-(Fig. \ref{fig:lambda_est_p_imba}).
+by taking the outer products of the popularity vectors.
 
@@
 
+\begin{comment}
+
+## Additional experiments
+
+Using the same set of parameters for generating $P$ and $A$ as in the
+balanced communities examples for $K = 2$, we generated one instance of $A$ for
+each $n$ and constructed $B$ according to Algorithm 3 to verify that
+as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{a.s.}{\to} 0$ for $i, j$
+in different clusters. Furthermore, the distribution of these inner products
+should be approximately normal.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-2-1}
+
+}
+
+\caption{Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+\end{comment}
+
 \hypertarget{real-data-examples}{%
-\section{Applications}\label{real-data-examples}}
-
-In the first example, we applied OSC to the Leeds Butterfly
-dataset \citep{Wang_2018} consisting of visual similarity measurements
+\section{Real data examples}\label{real-data-examples}}
+
+In the first real data example, we applied OSC to the Leeds Butterfly
+dataset \cite{Wang_2018} consisting of visual similarity measurements
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \citet{noroozi2019estimation}: Only the $K=4$ most
+the example from \cite{noroozi2019estimation}: Only the 4 most
 frequent species were considered, and the similarities were discretized
@@
 
-Comparing against the ground truth species labels, OSC achieves an adjusted Rand index of 92\%, while SSC on the ASE achieves an adjusted Rand index of 96\%.
-In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
+Comparing against the ground truth species labels, OSC achieves an
+accuracy of 63\% and an adjusted Rand index of 73\%. In comparison,
+\cite{noroozi2019estimation} achieved an adjusted Rand index of
+73\% using sparse subspace clustering on the same dataset.
 
@@
 
-\caption{Adjacency matrix of the Leeds Butterfly dataset 
-after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
+\caption{Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
 \end{figure}
@@
 In the second example, we applied OSC to the British MPs Twitter network
-\citep{greene2013producing}, the Political Blogs network
-\citep{10.1145/1134271.1134277}, and the DBLP network
-\citep{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
-analysis, we subsetted the data as described in
-\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
+\cite{greene2013producing}, the Political Blogs network
+\cite{10.1145/1134271.1134277}, and the DBLP network
+\cite{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
+analysis, we subsetted the data as described by
+\cite{307cbeb9b1be48299388437423d94bf1} for their analysis of the
 same networks. Our methods underperformed compared to modularity
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/mp-1}
-
-}
-
-\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
-\end{figure}
-
 \begin{table}
 
-\caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
+\caption{\label{tab:unnamed-chunk-6}Community detection error rates for modularity maximization, sparse subspace clustering,  and OSC.}
 \centering
@@
 \hline
-Political Blogs & 0.050 & 0.196 & 0.062\\
+Political blogs & 0.050 & 0.196 & 0.062\\
 \hline
@@
 
-In the third example, we analyzed the Karantaka villages data studied by
-\citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
+\begin{figure}[H]
+
+{\centering \includegraphics{summary_files/figure-latex/mp-1}
+
+}
+
+\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
+\end{figure}
+
+In the third example, we consider the Karantaka villages data studied by
+\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
 networks from villages 12, 31, and 46 at the household level. The label
@@
 religions ``1'' and ``2'', and vertices of degree 0 were removed.
-The results are displayed in Fig. \ref{fig:households-figure} and 
-Table \ref{tab:households-table}.
 
@@
 \caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
-The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:households-figure}
+The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
+\caption{\label{tab:unnamed-chunk-8}Community detection error rates for identifying household religion.}
 \centering
@@
 
-The central result of this paper illustrates that the 
-Popularity Adjusted Block Model is a special case of the 
-Generalized Random Dot Product Graph and how this connection 
-can be exploited for community detection and parameter estimation. 
-Specifically, the PABM with $K$ communities is a GRDPG 
-for which each community is represented by a $K$-dimensional subspace 
-in the latent space, and the subspaces are orthogonal to each other. 
-This result can be viewed as a continuation of previous results 
-that connect the Stochastic Block Model and the Degree Corrected Block Model 
-to the Random Dot Product Graph: 
-The SBM, DCBM, and PABM are successive generalizations of Block Models. 
-Similarly, the SBM is a RDPG for which each community corresponds to a point 
-(zero-dimensional subspace) \citep{lyzinski2014}, the DCBM is a RDPG 
-for which each community corresponds to a line (one-dimensional subspace) 
-\citep{rubindelanchy2017consistency}, 
-and the PABM is a GRDPG for which each community is a $K$-dimensional subspace.
-
-The latent structure of the PABM can be recovered 
-via Adjacency Spectral Embedding \citep{rubindelanchy2017statistical}, 
-which allows us to take advantage of this latent structure 
-for statistical inference. 
-Using the fact that the communities are represented by orthogonal subspaces, 
-we propose Orthogonal Spectral Clustering for community detection, 
-for which we have the theoretical guarantee that 
-the number of misclustered vertices approaches zero 
-with high probability as the size of the graph increases. 
-This is a stronger result than previously proposed algorithms
-\citep{307cbeb9b1be48299388437423d94bf1},
-which only guarantee that the error rate (and not count) 
-approaches zero asymptotically.
-Parameter estimation can be performed in a similar fashion using the ASE, 
-for which we also prove that 
-the per-parameter error approaches zero asymptotically. 
-Both of these results are demonstrated via simulation study.
-
-One proposed approach to community detection for the PABM 
-is an application of Sparse Subspace Clustering \citep{noroozi2019estimation}. 
-We show that the latent structure of the PABM leads to 
-both a straightforward justification for SSC for the PABM 
-as well as an improvement over the previous implementation.
-Our approach applies SSC on the ASE, 
-and we prove that the ASE of the PABM obeys the Subspace Detection Property
-with high probability if the graph is large. 
-This result is also demonstrated via simulation study.
+This paper shows the connection between the PABM and the GRDPG, namely
+that a PABM graph can be represented as a union of orthogonal subspaces
+in an embedding under the GRDPG framework. We then exploited this
+relationship to develop community detection and parameter estimation
+methods. In fact, we can represent any graph with Bernoulli edges as a
+GRDPG, and in the PABM case, it turns out that this relationship leads
+to a very straightforward applications of previous work from
+\cite{rubindelanchy2017statistical},
+\cite{soltanolkotabi2012}, and \cite{jmlr-v28-wang13}, which
+lead to asymptotically correct solutions with high probability. Similar
+methods can be applied for other models, such as the Nested Block Model
+\cite{noroozi2021hierarchy}.
 
@@
 \renewcommand\refname{References}
-% \bibliographystyle{plain}
-% \bibliography{misc.bib}
-\printbibliography
+  \bibliography{misc.bib}
 
@@
 $\ell_2$ norm difference between the rows of $\hat{V}_n$ and
-$V_n$. See \citet{cape_biometrika} and 
-\citet[Lemma~5]{rubindelanchy2017statistical} for a proof. 
+$V_n$. See \cite{cape_biometrika} and 
+\cite[Lemma~5]{rubindelanchy2017statistical} for a proof. 
 \begin{lemma}
   \label{lem:technical}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell \ell)}\}_{K})$ be a $K$-blocks
+Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_{K})$ be a $K$-blocks
 PABM graph on $n$ vertices and let \(V_n\) and \(\hat{V}_n\)
@@
 % eigenvalues, respectively. From Lemma~5 in
-% \citet{rubindelanchy2017statistical} we have, for some
+% \citeauthor{rubindelanchy2017statistical} we have, for some
 % \(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
@@
 % notion of the inradius of a set of points. 
-\begin{definition}[Inradius \citep{soltanolkotabi2012, jmlr-v28-wang13}]
+\begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
   \label{inradius}
@@
 % Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
-% $\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},
+% $\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},
 
@@
 the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
-\citet{jmlr-v28-wang13} states that there exists a $\lambda > 0$
+\cite{jmlr-v28-wang13} states that there exists a $\lambda > 0$
 such that $\sqrt{n} \hat{V}$ satisfies
@@
 % Indeed, let
-% The basis of this proof is Theorem 6 from \citet{jmlr-v28-wang13}, which states that the
+% The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
 % subspace detection property holds if the noise in $\sqrt{n}\hat{V}$
@@
 \begin{remark}
-Theorem 6 of \citet{jmlr-v28-wang13} assumes that each row $v_i$
+Theorem 6 of \cite{jmlr-v28-wang13} assumes that each row $v_i$
 of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. This
@@
 since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
-spans the subspace $\mathcal{S}^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
+spans the subspace $S^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
 for any scalar $a > 0$. 
@@
 \(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
-and by Theorem 5 of \citet{rubindelanchy2017statistical}, we have
+and by Theorem 5 of \cite{rubindelanchy2017statistical}, we have
 %adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
@@
 
-which is a symmetric \((n_k + n_\ell) \times (n_k + n_\ell)\) matrix. The
-eigendecomposition of \(M^{(k \ell)}\) is then
-
-\[M^{(k \ell)} = 
+which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. The
+eigendecomposition of \(M^{(kl)}\) is then
+
+\[M^{(kl)} = 
 \begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix} \times 
@@
 have latent positions in \(\mathbb{R}^2\) given by the $(n_k + n_{\ell}) \times 2$ matrix
-\[\Lambda^{(k \ell)} = \begin{bmatrix} 
+\[\Lambda^{K \ell} = \begin{bmatrix} 
   \sigma_{k \ell} u^{(k \ell)} & \sigma_{k \ell} u^{(k \ell)} \\ 
@@
 
-% and \(E\) is the \(n_k \times n_\ell\) matrix of independent noise (to
+% and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
 % generate the Bernoulli entries in \(A^{(kl)}\).\\
-We can then view \(\hat{M}^{(k \ell)}\) as an adjacency matrix drawn from
-the edge probabilities matrix \(M^{(k \ell)}\). Now suppose that the adjacency spectral
+We can then view \(\hat{M}^{(kl)}\) as an adjacency matrix drawn from
+the edge probabilities matrix \(M^{(kl)}\). Now suppose that the adjacency spectral
 embedding of $\hat{M}^{(k \ell)}$ is represented as the $(n_k +
@@
 \end{bmatrix}\]
-where each \(\hat{\lambda}^{(k \ell)}\) is defined as in Algorithm 3. Then
-by Theorem 5 of \citet{rubindelanchy2017statistical}, there
+where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3. Then
+by Theorem 5 of \cite{rubindelanchy2017statistical}, there
 exists an indefinite orthogonal transformation $W^{*}$ such that,  with
@@
 % orthogonal transformation by the rate given in Theorem 5 of
-% \citet{rubindelanchy2017statistical}.\\
+% \citeauthor{rubindelanchy2017statistical}.\\
 Furthermore, by looking at the proof of Theorem~5 in
-\citep{rubindelanchy2017statistical}, we see that $W^{*}$ is also
+\cite{rubindelanchy2017statistical}, we see that $W^{*}$ is also
 blocks diagonal with $2$ blocks where the positive eigenvalues of $M^{(k \ell)}$
@@
 yields Eq.~\eqref{eq:thm6} in Theorem~\ref{theorem6}. 
-
 \end{document}

### 2021-08-25 summary.tex

--- 2021-08-24 summary.tex
@@
 \usepackage{amssymb,amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \usepackage{ifxetex,ifluatex}
@@
 \author{John Koo\\Department of Statistics, Indiana University\\
-and\\
 Minh Tang\\Department of Statistics, North Carolina State University\\
-and\\
 Michael Trosset\\Department of Statistics, Indiana University}
@@
 
-Another type of Bernoulli Graph was proposed by \citet*{10.1007/978-3-540-77004-6_11}.
-A Random Dot Product Graph (RDPG) specifies that each vertex corresponds to a latent position vector in Euclidean space and that the probability of an edge between two vertices is the dot product of their latent position vectors.
-Thus, if the latent positions are $x_1,\ldots,x_n \in \mathbb{R}^d$ and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$, then the edge probability matrix is $P = XX^\top$.
-Clearly, any Bernoulli Graph with positive semidefinite $P$ is an RDPG.
-The positive definite Euclidean inner product in the RDPG model was replaced by an indefinite inner product in \citet{rubindelanchy2017consistency}, resulting in the {\em Generalized}\/ RDPG (GRDPG).
-
-In contrast to Block Models, neither RDPGs nor GRDPGs inherently specify distinct communities.
-However, one can easily impose community structure by assuming that the latent positions lie in distinct clusters.
-Hence, it is not surprising that Block Models can be studied by reformulating them as RDPGs or GRDPGs.
-For example, an assortative SBM (an SBM for which $P$ is positive semidefinite) is equivalent to an RDPG for which all vertices in the same community correspond to the same latent position vector (\cite{lyzinski2014}).
-Likewise, the DCBM is equivalent to an RDPG for which all vertices in the same community correspond to latent position vectors that lie on a straight line (\cite{lyzinski2014, rubindelanchy2017consistency}).
-
-Because the edge probability matrix of a PABM is not necessarily positive semidefinite, a PABM is not necessarily an RDPG.
-In Section \ref{connecting-the-pabm-to-the-grdpg} we demonstrate that every PABM is in fact a specific type of GRDPG for which the latent position vectors lie in distinct orthogonal subspaces, each subspace corresponding to a community.
-This identification is our central result.
-In Section \ref{methods}, we use the geometry of the GRDPG to derive more efficient algorithms for detecting the communities and estimating the parameters in the PABM.
-We report the results of simulation studies in Section \ref{simulated-examples} and apply our methods to three well-known data sets in Section \ref{real-data-examples}.
-Section \ref{discussion} concludes.
+Another type of Bernoulli Graph was proposed by
+\citet*{10.1007/978-3-540-77004-6_11}.  A Random Dot Product Graph
+(RDPG) specifies that each vertex corresponds to a latent position
+vector in Euclidean space and that the probability of an edge between
+two vertices is the dot product of their latent position vectors.
+Thus, if the latent positions are $x_1,\ldots,x_n \in \mathbb{R}^d$
+and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$,
+then the edge probability matrix is $P = XX^\top$.  Clearly, any
+Bernoulli Graph with positive semidefinite $P$ is an RDPG.  The
+positive definite Euclidean inner product in the RDPG model was
+replaced by an indefinite inner product in
+\citet{rubindelanchy2017statistical}, resulting in the {\em
+Generalized}\/ RDPG (GRDPG).
+
+In contrast to Block Models, neither RDPGs nor GRDPGs inherently
+specify distinct communities.  However, one can easily impose
+community structure by assuming that the latent positions lie in
+distinct clusters.  Hence, it is not surprising that Block Models can
+be studied by reformulating them as RDPGs or GRDPGs.  For example, an
+assortative SBM (an SBM for which $P$ is positive semidefinite) is
+equivalent to an RDPG for which all vertices in the same community
+correspond to the same latent position vector.
+% (\cite{lyzinski2014}).
+Likewise, the DCBM is equivalent to an RDPG for which all vertices in
+the same community correspond to latent position vectors that lie on a
+straight line. %(\cite{lyzinski2014, rubindelanchy2017statistical}).
+
+Because the edge probability matrix of a PABM is not necessarily
+positive semidefinite, a PABM is not necessarily an RDPG.  In Section
+\ref{connecting-the-pabm-to-the-grdpg} we demonstrate that every PABM
+is in fact a specific type of GRDPG for which the latent position
+vectors lie in distinct orthogonal subspaces, each subspace
+corresponding to a community.  This identification is our central
+result.  In Section \ref{methods}, we use the geometry of the GRDPG to
+derive more efficient algorithms for detecting the communities and
+estimating the parameters in the PABM.  We report the results of
+simulation studies in Section \ref{simulated-examples} and apply our
+methods to three well-known data sets in Section
+\ref{real-data-examples}. Section \ref{discussion} concludes.
 
@@
 In this section, we show that the PABM is a special case of the GRDPG. 
-More specifically, graph $G$ drawn from the PABM can be represented by 
+More specifically, a graph $G$ drawn from the PABM can be represented by 
 a collection of latent vectors in Euclidean space. 
@@
 Let $G = (V, E)$ be an unweighted, undirected, 
-and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. 
+and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. The matrix
 $A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ 
@@
 Finally, we denote 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$ 
-as the collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
+$X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^\top \in \mathbb{R}^{n \times d}$ 
+as the matrix corresponding to a collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
 
@@
 Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
-
 \begin{equation} \label{eq:pabm}
-  P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top
+  P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}
 \end{equation}
-
 for a set of $K^2$ popularity vectors 
@@
 \label{grdpg}
-Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$.
+Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$ for some symmetric $P \in
+[0,1]^{n \times n}$. 
 If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
-
 \begin{equation} \label{eq:grdpg}
@@
 \end{equation}
-
 for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
@@
 The indefinite orthogonal group with signature $(p, q)$ is
-the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{p, q} Q^{\top} = I_{p, q}\}$,
-denoted as $\mathbb{O}(p, q)$.
+the set $\{Q \in \mathbb{R}^{d \times d} \colon Q I_{p, q} Q^{\top} = I_{p, q}\}$,
+denoted as $\mathbb{O}(p, q)$. Here $d = p + q$. 
 \end{definition}
@@
 More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
-then we also have for any $Q \in \mathbb{O}(p, q)$,
+then we also have for any $Q \in \mathbb{O}(p, q)$ that
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
@@
 For the $K = 2$ case, the proof is straightforward. Let 
-
 $$X = \begin{bmatrix}
@@
 0 & 1 & 0 & 0 \end{bmatrix}.$$
-
 Then by straightforward matrix multiplication, we obtain 
-
 \[X U I_{3, 1} U^\top X^\top =
@@
 \end{bmatrix}, \]
-
 which is a GRDPG with latent vectors described by $X U$.
@@
   (L^{(1)})^{\top} \\ \vdots \\
-  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix} \\
+  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix}, \\
 \Lambda^{(k)} (L^{(k)})^{\top} = \begin{bmatrix} \lambda^{(k1)}
@@
 We therefore have
-
 \begin{equation} \label{eq:permutation}
-\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
+\Pi = U I_{p,q} U^\top
 \end{equation}
@@
 \end{equation}
-
 In summary we can describe the PABM with $K$ communities as a GRDPG with latent
@@
 permutation matrix of the form
-
 $$\Pi = 
@@
 \mid e_6 \mid e_9 \Bigr].$$
-
 where $e_i$ denotes the $i^{th}$ basis vector in $\mathbb{R}^{9}$. 
@@
 estimation. They were able to show that as the sample size increases,
-the proportion of misclassified community labels (up to permutation)
+the {\em proportion} of misclassified community labels (up to permutation)
 goes to 0.
@@
 \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
-\(c_i = \arg\min_{c} ||c||_1\) subject to \(x_i = X c\) and
+\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X c\) and
 \(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
 collected collected into matrix
-\(C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top\) to
+\(C = \bigl[ c_1 \mid c_2 \mid \cdots \mid c_n \bigr]^\top\) to
 construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
@@
 
-In practice, SSC is often performed by solving the LASSO problems
-
+In practice, due to presence of noise, SSC is often done by solving the LASSO problems
 \begin{equation} \label{eq:ssc}
-c_i = \arg\min_c \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
+c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i} c\|^2_2 + \lambda \|c\|_1
 \end{equation}
-
 for some sparsity parameter \(\lambda > 0\). The \(c_i\) vectors are
@@
   \label{def:subspace_detection}
-Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy
+Let $X = \bigl[ x_1 \mid x_2 \mid \cdots \mid x_n \bigr]^\top$ be noisy
 points sampled from $K$ subspaces, i.e., $x_i = y_i + z_i$ where the
@@
 with sparsity parameter $\lambda$ if each column of
-$C$ has nonzero norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
+$C$ has nonzero $\ell_2$ norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
 from different subspaces.
@@
 
-If we decompose \(P = Z I_{p, q} Z^\top\), then
-\(\exists\) \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Even if we
-start with the exact edge probability matrix, we cannot recover the
-``original'' latent positions \(XU\). Note that unlike in the case of
-the RDPG, \(Q\) is not necessarily an orthogonal matrix. If \(z_i\)'s
-are the rows of \(XU\), then
-\(||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2\), and
+More specifically, if we decompose \(P = Z I_{p, q} Z^\top\), then
+there exists a \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Therefore, even if we
+start with the true but unknown edge probability matrix $P$ we
+will not be able to recover the ``original'' latent positions \(XU\).
+Note that unlike in the case of the RDPG, \(Q\) is not necessarily an
+orthogonal matrix and hence if \(z_i\)'s
+are the rows of \(XU\) then
+\(\|z_i - z_j\|^2 \neq \|Q z_i - Q z_j\|^2\), and
 \(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
@@
 % \textbf{Proof of Theorem \ref{theorem3}}. 
-Recall that \(X\) is block diagonal with each block
+To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) is block diagonal with each block
 corresponding to one community and hence \(X (X^\top X)^{-1} X^\top\) is also a
@@
 Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC
-(Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities
-and sparsity factor satisfying $n \rho_n = \omega \big( (\log n)^{4c} \big)$,
-
+(Alg. 1). Then for all pairs $(i, j)$ belonging to different communities
+and sparsity factor satisfying $n \rho_n = \omega(\log^{4}{n})$, we have
 \begin{equation} \label{eq:thm4}
 \max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
-O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
+O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
 \end{equation}
-
-This provides the result that for $i, j$ in different communities,
-$\hat{B}_{ij} \stackrel{a.s.}{\to} 0$.
+with high probability. Eq.~\eqref{eq:thm4} implies that for all pairs $(i, j)$ in different communities,
+we have $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ where the convergence is
+uniform over all pairs $(i,j)$. 
 \end{theorem}
@@
       Assign $v_i^\top$ as the $i^{th}$ row of $V$.
-      Assign $V_{-i} = \begin{bmatrix}
-      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
+      Assign $V_{-i} = \bigl[
+      v_1 \mid \cdots \mid v_{i-1} \mid v_{i+1} \mid \cdots \mid v_n \bigr]^\top$.\;
       Solve the LASSO problem
       $c_i = \arg\min_{\beta}
-      \frac{1}{2} ||v_i - V_{-i} \beta||_2^2 + \lambda ||\beta||_1$.\;
-      Assign $\tilde{c}_i = \begin{bmatrix}
-      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
-      \end{bmatrix}^\top$ such that the superscript is the index of
+      \frac{1}{2} \|v_i - V_{-i} \beta\|_2^2 + \lambda \|\beta\|_1$.\;
+      Assign $\tilde{c}_i = (c_i^{(1)}, \dots, c_i^{(i-1)}, 0, c_i^{(i)}, \dots, c_i^{(n-1)})^\top$ such that the superscript is the index of
       $\tilde{c}_i$.\;
@@
     Assign
-    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
+    $C = \bigr[ \tilde{c}_1 \mid \cdots \mid \tilde{c}_n \bigr]$.\;
     Compute the affinity matrix $B = |C| + |C^\top|$.\;
@@
 matrix of eigenvectors of $A$ corresponding to the $K^2$ largest
-eigenvalues in modulus. Then for all $\epsilon > 0$ 
+eigenvalues in modulus. Then for any $\epsilon > 0$ 
 there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
@@
 
+%<<<<<<< HEAD
+%\begin{remark}
+%The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
+%from  \cite{jmlr-v28-wang13} and the fact that the columns of
+%$V_n$ correspond to orthogonal subspaces. \cite{jmlr-v28-wang13} assume
+%that the points in the embedding are all of unit length, and while we apply
+%this normalization in the simulated examples, it is not strictly necessary
+%for Theorem \ref{theorem5} due to orthogonality.
+%\end{remark}
+%
+%=======
+%>>>>>>> 08e715a2242ef7f3f92394d598119b189bdb2790
 \hypertarget{parameter-estimation}{%
@@
 let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\)
-be the singular value decomposition of \(P^{(k \ell)}\).
+be the singular value decomposition of \(P^{(k \ell)}\) where
 \(u^{(k \ell)} \in \mathbb{R}^{n_k}\) and 
@@
   communities $k$ and $l$.\;
-  \For {$k, l = 1, ..., K$, $k \leq l$} {
-    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
+  \For {$k, \ell = 1, ..., K$, $k \leq \ell$} {
+    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $k\ell$-th 
     block.\;
@@
 3).
-Then if $n \rho_n = \omega \big( (\log n)^{4 c} \big)$,
-
+Then if $n \rho_n = \omega( \log^{4}{n})$,
 \begin{equation} \label{eq:thm6}
@@
 \|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
-O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
+O\bigg(\frac{\log n_k}{\sqrt{n_k}} \bigg)
 \end{equation}
-
 with high probability.
@@
 The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in Algorithm 2. The sparsity
-parameter \(\lambda\) was chosen via a preliminary cross-validation
+performs SSC on the embedding, as described in Algorithm 2. The 
+parameter \(\lambda\) that controls the sparsity of the SSC was chosen via a preliminary cross-validation
 experiment. For the final clustering step, a Gaussian Mixture Model was
@@
 For comparing methods, we define the community detection error as:
-
 \[L_c(\hat{\sigma}, \sigma; \{v_i\}) =
 \min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))\]
-
 where \(\sigma(v_i)\) is the true community label of vertex \(v_i\),
@@
 First, we estimate the popularity vectors directly and compute the RMSE. 
-In this case, the "true" popularity vectors are derived from 
+In this case, the ``true'' popularity vectors are derived from 
 taking the SVD of each edge probability block $P^{(k \ell)}$ 
-to avoid the unidentifiable multiplicative constants.
-
+to avoid the unidentifiable multiplicative constants, i.e., 
 \[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}, \{\tilde{\lambda}^{(k \ell)}\}) = 
-\sqrt{\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)} \|_2^2}\]
-
+\Bigl(\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
+\|s \hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)} \|_2^2\Bigr)^{1/2}.\]
 We can also avoid the unidentifiable multiplicative constant more directly 
@@
 $\hat{P}^{(k \ell)} = \hat{\lambda}^{(k \ell)} (\hat{\lambda}^{(\ell k)})^\top$, 
-which we use to define another parameter estimation error.
-
+which we use to define another parameter estimation error, i.e., 
 \[\text{RMSE}(\hat{P}, P) =
-\sum_{k,l} \sqrt{\frac{1}{n_k n_\ell} \|P^{(k \ell)} - \hat{P}^{(k \ell)}\|_F^2}\]
+\sum_{k,l} \sqrt{\frac{1}{n_k n_\ell} \|P^{(k \ell)} - \hat{P}^{(k \ell)}\|_F^2}.\]
 
@@
 In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
+>>>>>>> 08e715a2242ef7f3f92394d598119b189bdb2790
 

### 2021-08-25 summary.tex
Update git

@@
 % \usepackage{natbib}
-\usepackage[backend=bibtex,style=authoryear,giveninits=true,natbib]{biblatex}
+\usepackage[backend=bibtex,maxbibnames = 10,style=authoryear,giveninits=true,natbib]{biblatex}
 \DeclareNameAlias{sortname}{last-first}
@@
 \begin{algorithm}[t]
+  \label{alg:osc}
   \DontPrintSemicolon
@@
 \begin{algorithm}[t]
+  \label{alg:ssc}
   \DontPrintSemicolon
@@
 \begin{algorithm}[t]
+  \label{alg:param_est}
   \DontPrintSemicolon
@@
 vectors, and finally the adjacency matrix \(A\)
-is drawn from \(P\). OSC is then used for community detection, and this
-method is compared against SSC (\cite{noroozi2019estimation,soltanolkotabi2014})
-and MM (\cite{igraph, 307cbeb9b1be48299388437423d94bf1}). 
-True community labels are used
-with Algorithm 3 to estimate the popularity vectors
-\(\{\lambda^{(k \ell)}\}_K\), and this method is then compared against an
-MLE-based estimator described in \citet{noroozi2019estimation} and
+is drawn from \(P\). OSC (Algorithm~\ref{alg:osc}) is then used for community detection, and this
+method is compared against (1) SSC using the spectral embedding $\hat{V}$,
+c.f. Algorithm~\ref{alg:ssc}, (2) SSC using the rows of the
+observed adjacency matrix $A$ as is done in \citet{noroozi2019estimation}
+and (3) modularity maximization (MM) as is done in
+\citet{307cbeb9b1be48299388437423d94bf1}. 
+We also estimate the latent popularity vectors $\{\lambda^{(k \ell)}\}$
+by assuming that the true community labels are known and then apply
+Algorithm~\ref{alg:param_est}, and we compare this estimation method against an
+MLE-based estimator as described in \citet{noroozi2019estimation} and
 \citet{307cbeb9b1be48299388437423d94bf1}.
@@
 \citet{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
-(EP) algorithm (\cite{le2016}), which is \(O(n^{K - 1})\), as a greedy
-relaxation of the optimization problem. For these simulations, the
-Louvain algorithm was used for modularity maximization,
-as \citet{307cbeb9b1be48299388437423d94bf1}'s implementation proved to be
-prohibitively computationally expensive for \(K > 2\). For \(K = 2\), it
+(EP) algorithm (\cite{le2016}) as a greedy
+relaxation of the optimization problem; the EP algorithm has a running
+time of $O(n^{K-1})$ where $n$ is the number of vertices in the graph
+and $K$ is the number of communities.
+For these simulations we instead replace the EP algorithm with the
+Louvain algorithm for modularity maximization,
+as the implementation of the EP algorithm in
+\citet{307cbeb9b1be48299388437423d94bf1} is too computationally expensive for \(K > 2\). For \(K = 2\), it
 was verified that the Louvain algorithm produces comparable results
@@
 The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in Algorithm 2. The 
+performs SSC on the embedding, as described in Algorithm~\ref{alg:ssc}. The 
 parameter \(\lambda\) that controls the sparsity of the SSC was chosen via a preliminary cross-validation
@@
 \end{itemize}
-
-\(50\) simulations were performed for each \((n, K)\) pair.
-
-Fig. \ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
+Fifty simulations were performed for each combination of $n$ and $K$.
+
+Fig.~\ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
 going to 0 for large \(n\). SSC on both the embedding and on the
@@
 
-Given ground truth community labels, Algorithm 3 and the MLE-based
+Given ground truth community labels, Algorithm~\ref{alg:param_est} and the MLE-based
 plug-in estimators perform similarly, with root mean square
@@
 We again see community detection error trending to 0 for OSC, as well as
-for SSC when \(K > 2\) (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3
-continues to see \(n^{-1/2}\) decay in parameter estimation error; 
-however, the MLE-based estimators for the popularity vectors decay at a slower 
-rate (Fig. \ref{fig:lambda_est_k_imba}). 
+for SSC when \(K > 2\) (Fig.~\ref{fig:clust_err_ct_sim_imba}). The
+estimation error for the parameters $\{\lambda^{k \ell}\}$ as given by
+Algorithm~\ref{alg:param_est} still decays to $0$ at rate $n^{-1/2}$;  
+however, the MLE-based estimators for the popularity vectors appear to
+decay at a slower 
+rate (Fig.~\ref{fig:lambda_est_k_imba}). 
 The reduced rate of decay is removed 
@@
 {\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim_imba-1}
-
 }
@@
 
-In the first example, we applied OSC to the Leeds Butterfly
+In the first example, we applied OSC (Algorithm~\ref{alg:osc}) to the Leeds Butterfly
 dataset \citep{Wang_2018} consisting of visual similarity measurements
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \citet{noroozi2019estimation}: Only the $K=4$ most
+the example from \citet{noroozi2019estimation}, i.e., only the $K=4$ most
 frequent species were considered, and the similarities were discretized
-to \(\{0, 1\}\) via thresholding. Fig. \ref{fig:butterfly} shows a
+to \(\{0, 1\}\) via thresholding. Fig.~\ref{fig:butterfly} shows a
 sorted adjacency matrix sorted by the resultant clustering.
@@
 In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
->>>>>>> 08e715a2242ef7f3f92394d598119b189bdb2790
 
@@
 \citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
-same networks. Our methods underperformed compared to modularity
-maximization, although performance is comparable. In addition, OSC's
-runtime is much lower than that of modularity maximization.
-
-\begin{figure}[H]
-
+same networks. Our methods slightly underperformed compared to modularity
+maximization, although performance is comparable. The run time of OSC
+is however much smaller than that of modularity maximization.
+
+\begin{figure}[tp]
 {\centering \includegraphics{summary_files/figure-latex/mp-1}
-
 }
-
 \caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
@@
 \begin{table}
-
 \caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
@@
 religions ``1'' and ``2'', and vertices of degree 0 were removed. 
-The results are displayed in Fig. \ref{fig:households-figure} and 
-Table \ref{tab:households-table}.
-
-\begin{figure}[H]
-
+The results are displayed in Fig.~\ref{fig:households-figure} and 
+Table~\ref{tab:households-table}.
+
+\begin{figure}[tp]
 {\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-7-1}
-
 }
-
 \caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
@@
 \begin{table}
-
 \caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
@@
 
-Because all Bernoulli Graphs are GRDPGs, it should be possible to invent and study new families of Bernoulli Graphs by characterizing them as special cases of GRDPGs and expliting the latent structures that define them.
-The present work illustrates the power of this approach. 
-We recover the latent structure of the PABM by Adjacency Spectral Embedding, then exploit that structure to improve statistical inference. 
-Exploiting the fact that PABM communities correspond to orthogonal subspaces, we propose Orthogonal Spectral Clustering for community detection and demonstrate that the number of misclassified vertices approaches zero with high probability as the size of the graph increases.
-This is a stronger result than previously proposed algorithms \citep{307cbeb9b1be48299388437423d94bf1}, which only guarantee that the error rate (and not count) approaches zero asymptotically. 
-Parameter estimation can be performed in a similar fashion using the ASE, for which we also prove that the per-parameter error approaches zero asymptotically.
-
-A secondary benefit of the GRDPG approach is that the latent structure may be used to improve existing algorithms. 
-For example, one algorithm for PABM community detection \citep{noroozi2019estimation} relies on Sparse Subspace Clustering. 
-The latent structure of the PABM provides a natural justification for SSC for the PABM and leads to an improvement over the previous implementation. 
-The improved algorithm applies SSC to the ASE, and we prove that the ASE of the PABM obeys the Subspace Detection Property with high probability if the graph is large.
-
-Finally, one might well inquire what one gains and what one sacrifices by assuming that a Bernoulli Graph is a PABM. 
-The GRDPG model offers a plausible way to pursue this inquiry. 
-Absent a known latent structure that can be exploited by specialized methods, the GRDPG-ASE approach transforms the problem of network community detection to the much-studied problem of clustering vectors in Euclidean space.
+Because all Bernoulli Graphs are GRDPGs, it should be possible to
+invent and study new families of Bernoulli Graphs by characterizing
+them as special cases of GRDPGs and exploiting the latent structures
+that define them.  The present work illustrates the power of this
+approach.  We recover the latent structure of the PABM by Adjacency
+Spectral Embedding, then exploit that structure to improve statistical
+inference.  Exploiting the fact that PABM communities correspond to
+orthogonal subspaces, we propose Orthogonal Spectral Clustering for
+community detection and demonstrate that the number of misclassified
+vertices approaches zero with high probability as the size of the
+graph increases.  This is a stronger result than previously proposed
+algorithms \citep{307cbeb9b1be48299388437423d94bf1}, which only
+guarantee that the error rate (and not count) approaches zero
+asymptotically.  Parameter estimation can be performed in a similar
+fashion using the ASE, for which we also prove that the per-parameter
+error approaches zero asymptotically.
+
+A secondary benefit of the GRDPG approach is that the latent structure
+may be used to improve existing algorithms.  For example, one
+algorithm for PABM community detection \citep{noroozi2019estimation}
+relies on Sparse Subspace Clustering.  The latent structure of the
+PABM provides a natural justification for SSC for the PABM and leads
+to an improvement over the previous implementation.  The improved
+algorithm applies SSC to the ASE, and we prove that the ASE of the
+PABM obeys the Subspace Detection Property with high probability if
+the graph is large.
+
+Finally, one might well inquire what one gains and what one sacrifices
+by assuming that a Bernoulli Graph is a PABM.  The GRDPG model offers
+a plausible way to pursue this inquiry.  Absent a known latent
+structure that can be exploited by specialized methods, the GRDPG-ASE
+approach transforms the problem of network community detection to the
+much-studied problem of clustering vectors in Euclidean space.
 Communities of vertices are defined as clusters of latent vectors.
-After ASE, a standard clustering algorithm, e.g., single linkage, is used ot infer the communities. 
-In future research, we intend to use such general algorithms as baselines and measure the efficiency of the PABM algorithms (and other specialized algorithms) by studying how much they improve on general algorithms when the specified latent structure obtains.
+After ASE, a standard clustering algorithm, e.g., single linkage, is
+used ot infer the communities.  In future research, we intend to use
+such general algorithms as baselines and measure the efficiency of the
+PABM algorithms (and other specialized algorithms) by studying how
+much they improve on general algorithms when the specified latent
+structure obtains.
 
@@
 \hypertarget{proofs}{%
-\section{Proofs of Theorem~\ref{theorem4} and Theorem~\ref{theorem5}}}
+\section{Proofs of Theorem~\ref{theorem4}, Theorem~\ref{theorem5}, and Theorem~\ref{theorem6}}}
 Let \(V_n\) and \(\hat{V}_n\)
@@
 Then there exists a constant $c > 1$ and an orthogonal matrix $W$ such
-that, with high probability,
-$$\max_{i} \|W \hat{v}_n^{(i)} - v_n^{(i)}\|  = O_P \Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
+that with high probability,
+$$\max_{i} \|W \hat{v}_n^{(i)} - v_n^{(i)}\|  = O\Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
 In particular we can take $c = 1 + \epsilon$ for any $\epsilon > 0$. 
@@
 Definition~\ref{def:subspace_detection}. We first recall a few
-definitions used in \cite{jmlr-v28-wang13}; for ease of exposition,
-these definitions are stated using the notations of the current paper.
-
+definitions used in \citet{soltanolkotabi2012} and \citet{jmlr-v28-wang13}; for ease of exposition,
+these definitions are stated using the notations of the current
+paper and we will drop the explicit dependency on $n$ from our
+eigenvectors $\hat{V}_n$ of $A$ and $V_n$ of $P$.
  % We first recall the
 % notion of the inradius of a set of points. 
-\begin{definition}[Inradius \citep{soltanolkotabi2012, jmlr-v28-wang13}]
+\begin{definition}[Inradius]
   \label{inradius}
@@
   \label{def:subspace_incoherence}
-Let $\hat{V}$ be the eigenvectors of $A_n$
+Let $\hat{V}$ be the eigenvectors of $A$
 corresponding to the $K^2$ largest eigenvalues in modulus. Let
@@
 $v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
-$P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
+$P$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
 rows of $V^{(k)}$.
@@
 \frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
-Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
+Now let $W^{(k)} = \bigl[ w_1^{(k)} \mid \cdots \mid w_{n_k}^{(k)} \bigr]^\top$
 and define the subspace incoherence for $\hat{V}^{(k)}$ by
@@
 the following two conditions are satisfied
-
 \begin{gather}
@@
 \end{gather}
-
 We now verify that for sufficiently large n, Eq.~\eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
@@
 
-\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) be
+\begin{proof}[Proof of Theorem \ref{theorem6}] Let \(P\) be
 organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
@@
 %adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O\Bigl(\frac{(\log n_k)^c}{\sqrt{n_k}}\Bigr)$$
+$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O\Bigl(\frac{\log n_k}{\sqrt{n_k}}\Bigr)$$
 with high probability. 
@@
 $$\max_{i} |W^{*} \hat{\Lambda}_{i}^{(k \ell)} - \Lambda_{i}^{(k \ell)} \| =
-O\Bigl(\frac{(\log (n_k + n_{\ell}))^c}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
+O\Bigl(\frac{\log (n_k + n_{\ell})}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
 with high probability. Here $\Lambda_{i}^{(k \ell)}$ and
@@
 \|\hat{\lambda}^{(\ell k)} - \lambda^{(\ell k)}\|_{\infty}\}  =
-O\Bigl(\frac{(\log (n_k + n_{\ell}))^c}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
+O\Bigl(\frac{\log(n_k + n_{\ell})}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
 with high probability. Combining this bound with the bound for
@@
 yields Eq.~\eqref{eq:thm6} in Theorem~\ref{theorem6}. 
-
+\end{proof}
 \end{document}

### 2021-08-27 summary.tex
syncing stuff

@@
   Recall the notations in Lemma~\ref{lem:technical} and note that,
-  under our assumption that the latent vectors $\lambda^{(k \ell)|}$
+  under our assumption that the latent vectors $\lambda^{(k \ell)}$
   are all homogeneous, we have $\max_{i} \|v_n^{(i)}\| =

### 2021-08-29 summary.tex
syncing stuff again

@@
 divided by the number of vertices. 
-The situation is the same as in \cite{307cbeb9b1be48299388437423d94bf1} (PABM) 
+The treatment is the same as in \cite{307cbeb9b1be48299388437423d94bf1} (PABM) 
 and \cite{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
 are scaled as $\rho_n^{1/2} \lambda^{(k \ell)}$. 
-For ease of notation, we restrict our results to the case in which all 
-nonzero popularity parameters are greater than some positive minimum value.
+This results in the edge probability matrix being scaled as $\rho_n P$. 
+Since $\rho_n$ is unidentifiable, we use the convention that the popularity parameters are drawn such that $\|X\|_F = O(\sqrt{n})$ and require the $L^2$ norm of each row is bounded away from 0 and 1.
 
@@
 \end{equation}
-with high probability. Eq.~\eqref{eq:thm4} implies that for all pairs $(i, j)$ in different communities,
-we have $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ where the convergence is
-uniform over all pairs $(i,j)$. 
+with high probability. 
 \end{theorem}
+
+Eq.~\eqref{eq:thm4} implies that for all pairs $(i, j)$ in different communities,
+we have uniform convergence $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$. 
+This allows us to identify the communities perfectly from $\hat{B}$ for large $n$. 
+For example, if we use edge thresholding to partition $\hat{B}$, 
+Theorem \ref{theorem4} implies that for any $\epsilon > 0$, 
+the number of edges of $\hat{B}$ between vertices of different communities 
+that are larger than $\epsilon$ converges to zero with probability 1.
 
@@
 
-We also define two types of parameter estimation error. 
-First, we estimate the popularity vectors directly and compute the RMSE. 
-In this case, the ``true'' popularity vectors are derived from 
-taking the SVD of each edge probability block $P^{(k \ell)}$ 
-to avoid the unidentifiable multiplicative constants, i.e., 
-\[\text{RMSE}(\{\hat{\lambda}^{(k \ell)}\}, \{\tilde{\lambda}^{(k \ell)}\}) = 
-\Bigl(\frac{1}{n} \sum_{k < l} \min_{s = \pm 1} 
-\|s \hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)} \|_2^2\Bigr)^{1/2}.\]
-We can also avoid the unidentifiable multiplicative constant more directly 
-by reconstructing each 
-$\hat{P}^{(k \ell)} = \hat{\lambda}^{(k \ell)} (\hat{\lambda}^{(\ell k)})^\top$, 
-which we use to define another parameter estimation error, i.e., 
-\[\text{RMSE}(\hat{P}, P) =
-\sum_{k,l} \sqrt{\frac{1}{n_k n_\ell} \|P^{(k \ell)} - \hat{P}^{(k \ell)}\|_F^2}.\]
+For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $p_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
+The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$:
+
+$$\mathrm{RMSE}(\hat{P}, P) = \|\hat{P} - P\|_F.$$
 

### 2021-08-30 summary.tex
updated writeup per discussion

@@
 For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $p_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
-The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$:
-
-$$\mathrm{RMSE}(\hat{P}, P) = \|\hat{P} - P\|_F.$$
+The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$ divided by the number of vertices:
+
+$$\mathrm{RMSE}(\hat{P}, P) = \frac{1}{n} \|\hat{P} - P\|_F.$$
+
+We also note that unlike the MLE-based method \citep{307cbeb9b1be48299388437423d94bf1}, the ASE method of Algorithm 3 does not require community labels if all we are interested in is the edge probability matrix, $P$.
+The estimate $\hat{P}$ we obtain by combining the popularity vectors outputted by Algorithm 3 is equivalent to using the full ASE of $A$ to reconstruct the edge probability matrix.
 
@@
 plug-in estimators perform similarly, with root mean square
-error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:lambda_est_k}). 
-We observe similar behavior when comparing errors 
-in the edge probability blocks (Fig. \ref{fig:p_block_est}).
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/lambda_est_k-1}
-
-}
-
-\caption{Median and IQR RMSE for popularity vectors from Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.}
-\label{fig:lambda_est_k}
-\end{figure}
+error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:p_block_est}).
 
@@
 {\centering \includegraphics{summary_files/figure-latex/lambda_est_p-1}}
-\caption{Median and IQR RMSE for edge probability blocks reconstructed from the outputs of Algorithm 3 (red) compared against outputs of an MLE-based method (blue).
+\caption{Median and IQR RMSE for edge probability matrices reconstructed from the outputs of Algorithm 3 (red) compared against outputs of an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.
@@
 We again see community detection error trending to 0 for OSC, as well as
-for SSC when \(K > 2\) (Fig.~\ref{fig:clust_err_ct_sim_imba}). The
-estimation error for the parameters $\{\lambda^{k \ell}\}$ as given by
-Algorithm~\ref{alg:param_est} still decays to $0$ at rate $n^{-1/2}$;  
-however, the MLE-based estimators for the popularity vectors appear to
-decay at a slower 
-rate (Fig.~\ref{fig:lambda_est_k_imba}). 
-The reduced rate of decay is removed 
-if we reconstruct the edge probability blocks 
-by taking the outer products of the popularity vectors 
+for SSC when \(K > 2\) (Fig.~\ref{fig:clust_err_ct_sim_imba}). 
+The parameter estimation error decays at rate $n^{-1/2}$ as in the balanced communities case
 (Fig. \ref{fig:lambda_est_p_imba}).
@@
 
-{\centering \includegraphics{summary_files/figure-latex/lambda_est_k_imba-1}
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_p_imba-1}
 
@@
 
-\caption{Median and IQR RMSE from Algorithm 3 (red) compared against an MLE-based method (blue).
+\caption{Median and IQR RMSE of edge probabilities derived from the outputs of Algorithm 3 (red) compared against an MLE-based method (blue).
 Simulations were repeated 50 times for each sample size.
-Communities were drawn to be imbalanced.}\label{fig:lambda_est_k_imba}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/lambda_est_p_imba-1}
-
-}
-
-\caption{Median and IQR RMSE of edge probability blocks derived from the outputs of Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be imbalanced.}\label{fig:lambda_est_p_imba}
+Communities were drawn to be imbalanced.}
+\label{fig:lambda_est_p_imba}
 \end{figure}
@@
 \citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
-networks from villages 12, 31, and 46 at the household level. The label
-of interest is the religious affiliation. The networks were truncated to
-religions ``1'' and ``2'', and vertices of degree 0 were removed. 
+networks from villages 12, 31, and 46 at the household level. 
+In these networks, each node is a household and each edge is 
+an interaction between members of pairs of households. 
+The label of interest is the religious affiliation. 
+The networks were truncated to religions ``1'' and ``2'', 
+and vertices of degree 0 were removed. 
+The villages were chosen based on there being an adequate number of nodes 
+between households within each religion.
+
 The results are displayed in Fig.~\ref{fig:households-figure} and 

### 2021-08-30 functions.R

@@ -241,14 +241,14 @@ lambda.rmse <- function(P, A, clustering) {
         lambda.hat <- estimate.lambda.block(A.block, TRUE)
         P.hat <- lambda.hat %*% t(lambda.hat)
         # return(sum((lambda - lambda.hat) ** 2))
-        return(norm(P.hat - P.block, 'F') / (n.k))
+        return(norm(P.hat - P.block, 'F') ** 2)
       } else {
         # lambdas <- estimate.lambda.block(P.block)
         lambda.hats <- estimate.lambda.block(A.block)
         # return(sum((lambdas[[1]] - lambda.hats[[1]]) ** 2) + 
         #          sum((lambdas[[2]] - lambda.hats[[2]]) ** 2))
         P.hat <- lambda.hats[[1]] %*% t(lambda.hats[[2]])
-        return(norm(P.hat - P.block, 'F') / sqrt(n.k * n.l))
+        return(2 * norm(P.hat - P.block, 'F') ** 2)
       }
     }) %>% 
       sum() %>% 
@@ -256,7 +256,8 @@ lambda.rmse <- function(P, A, clustering) {
   }) %>% 
     sum() %>% 
     # magrittr::divide_by(n * K) %>% 
-    # sqrt() %>% 
+    magrittr::divide_by(n ** 2) %>% 
+    sqrt() %>%
     return()
 }
 
@@ -321,7 +322,7 @@ lambda.rmse.mle <- function(P, A, z) {
         # lambda <- estimate.lambda.block(P.kk, TRUE)
         # return(sum((lambda - lambda.hat) ** 2))
         P.hat <- lambda.hat %*% t(lambda.hat)
-        return(norm(P.hat - P.kk, 'F') / (n.k))
+        return(norm(P.hat - P.kk, 'F') ** 2)
       } else {
         n.l <- n.vector[l]
         e.n.l <- rep(1, n.l)
@@ -341,7 +342,7 @@ lambda.rmse.mle <- function(P, A, z) {
         # return(sum((lambdas[[1]] - lambda.hat.kl) ** 2) + 
         #          sum((lambdas[[2]] - lambda.hat.lk) ** 2))
         P.hat <- lambda.hat.kl %*% t(lambda.hat.lk)
-        return(norm(P.kl - P.hat, 'F') / sqrt(n.k * n.l))
+        return(2 * norm(P.kl - P.hat, 'F') ** 2)
       }
[…]

### 2021-08-30 rmse-sim-k.R

@@ -11,7 +11,7 @@ import::here(embedding, draw.graph, I.pq, generate.P.beta,
              .from = 'functions.R')
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores() / 2)
+doMC::registerDoMC(parallel::detectCores() - 2)
 
 # simulation params
 K.vec <- c(4, 3, 2)
@@ -27,7 +27,7 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   Ipq <- I.pq(p, q)
-  
+
   out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
     print(paste('n =', n, 'K =', K))
     plyr::ldply(seq(iter), function(i) {
@@ -37,9 +37,9 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       A <- draw.graph(P)
       rmse <- lambda.rmse(P, A, z)
       rmse.mle <- lambda.rmse.mle(P, A, z)
-      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
+      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>%
         return()
-    }, .parallel = TRUE) %>% 
+    }, .parallel = TRUE) %>%
       return()
   }
   gc()

### 2021-08-30 rmse-sim.R

@@ -10,7 +10,7 @@ import::here(embedding, draw.graph, I.pq,
              .from = 'functions.R')
 
 # parallel backend
-doMC::registerDoMC(parallel::detectCores())
+doMC::registerDoMC(parallel::detectCores() - 2)
 
 # simulation params
 K <- 2
@@ -62,5 +62,5 @@ rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
     return()
 }
 
-readr::write_csv(rmse.df, 'rmse.csv')
+# readr::write_csv(rmse.df, 'rmse.csv')
 

### 2021-09-01 summary.tex
moved table captions to below tables

@@
 
-%<<<<<<< HEAD
-%\begin{remark}
-%The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
-%from  \cite{jmlr-v28-wang13} and the fact that the columns of
-%$V_n$ correspond to orthogonal subspaces. \cite{jmlr-v28-wang13} assume
-%that the points in the embedding are all of unit length, and while we apply
-%this normalization in the simulated examples, it is not strictly necessary
-%for Theorem \ref{theorem5} due to orthogonality.
-%\end{remark}
-%
-%=======
-%>>>>>>> 08e715a2242ef7f3f92394d598119b189bdb2790
 \hypertarget{parameter-estimation}{%
@@
 \begin{table}
-\caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
 \centering
@@
 \end{tabular}
+\caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
 \end{table}
 
-In the third example, we analyzed the Karantaka villages data studied by
+In the third example (Fig.~\ref{fig:households-figure} and 
+Table~\ref{tab:households-table}), 
+we analyzed the Karantaka villages data studied by
 \citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
@@
 
-The results are displayed in Fig.~\ref{fig:households-figure} and 
-Table~\ref{tab:households-table}.
-
 \begin{figure}[tp]
@@
 \begin{table}
-\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
 \centering
@@
 \end{tabular}
+\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
 \end{table}

### 2021-09-01 summary.tex
figure 5 now comes after figure 6

@@
 correspond to the same latent position vector.
-% (\cite{lyzinski2014}).
 Likewise, the DCBM is equivalent to an RDPG for which all vertices in
@@
 construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
-perfectly on one of \(K\) subspaces, \(B\) describes an undirected graph
+exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
 consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
@@
 We first show that $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in Eq.\eqref{eq:xy}. Indeed, by Theorem 2, \(P = X U I_{p,
-  q} U^\top X^\top\) for $p = K(K+1)/2$ and $q = K(K-1)/2$. % are given by $p =
-% K(K+1)/2$ and $q = are defined as
-% in equations (\ref{eq:p}) and (\ref{eq:q}). 
+defined as in Eq.~\eqref{eq:xy}. Indeed, by Theorem 2, 
+\(P = X U I_{p, q} U^\top X^\top\) for $p = K(K+1)/2$ and $q = K(K-1)/2$.
+
 The eigendecomposition \(P = V D V^\top\) also yields $P = V
@@
 $Y$ and $Z$ both have full column ranks. 
-Since $P = Y I_{p,q} Y^{\top} = Z I_{p,q} Z^{\top}$, we have
+Because $P = Y I_{p,q} Y^{\top} = Z I_{p,q} Z^{\top}$, we have
 $$Y = Z I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}.$$
@@
 
-Let $R = U Q |D|^{-1/2}$ and note that $V = XR$. Since $R$ is
-{\em invertible}, we have
+Let $R = U Q |D|^{-1/2}$ and note that $V = XR$. Because $R$ is
+invertible, we can write
+
 $$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
 R^{\top} X^{\top}.$$
+
 Furthermore, as $V$ has orthonormal columns, $R^{\top} X^{\top} X R =
@@
 as desired.
-% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
-% we have
-% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
-% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
-% The right-hand side becomes
-% \[\begin{split}
-%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
-%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
-%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
-%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
-%   & = X (X^\top X)^{-1} X^\top.
-% \end{split}\]
-% The left-hand side becomes:
-% \[\begin{split}
-%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
-%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
-%   & = V V^\top.
-% \end{split}\]
-
-% \textbf{Proof of Theorem \ref{theorem3}}. 
+
 To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) is block diagonal with each block
@@
 Letting \(|B|\) be the affinity matrix for graph \(G'\), \(G'\) is
-partitioned into at least \(K\) disjoint subgraphs since each of the
+partitioned into at least \(K\) disjoint subgraphs because each of the
 \(K\) communities have no edges between them. Similar to the subspace
@@
 
-Using \(A\) instead of \(P\) introduces error, which converges to \(0\)
-almost surely. We use this fact to motivate Orthogonal Spectral Clustering 
-(Alg. 1). 
+Using \(A\) instead of \(P\) to estimate $B$ introduces error, 
+which converges to \(0\) almost surely. 
+This is shown via properties of the GRDPG and ASE. 
+We use this fact to motivate Orthogonal Spectral Clustering (Alg. 1). 
 In order to show that OSC is consistent, 
@@
 divided by the number of vertices. 
-The treatment is the same as in \cite{307cbeb9b1be48299388437423d94bf1} (PABM) 
-and \cite{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
+The treatment is the same as in \citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
+and \citet{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
 are scaled as $\rho_n^{1/2} \lambda^{(k \ell)}$. 
-This results in the edge probability matrix being scaled as $\rho_n P$. 
-Since $\rho_n$ is unidentifiable, we use the convention that the popularity parameters are drawn such that $\|X\|_F = O(\sqrt{n})$ and require the $L^2$ norm of each row is bounded away from 0 and 1.
+This results in the edge probability matrix being scaled as 
+$P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
+Since $\rho_n$ is unidentifiable, we use the convention that the popularity parameters are drawn such that $\|\Lambda\|_F = O(\sqrt{n})$ and require that the $L^2$ norm of each row is bounded away from 0 and 1.
 
@@
 with high probability.
+
+Thus for all pairs $(i, j)$ in different communities,
+we have uniform convergence $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$. 
 \end{theorem}
 
-Eq.~\eqref{eq:thm4} implies that for all pairs $(i, j)$ in different communities,
-we have uniform convergence $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$. 
-This allows us to identify the communities perfectly from $\hat{B}$ for large $n$. 
+Theorem \ref{theorem4} provides an algorithm for community detection 
+that results in zero community detection error as $n \to \infty$. 
 For example, if we use edge thresholding to partition $\hat{B}$, 
@@
 and \(\sigma^{(k \ell)}\) is a scalar. 
-Then \(\lambda^{(k \ell)} = s_1 u^{(k \ell)}\)
-and \(\lambda^{(\ell k)} = s_2 v^{(k \ell)}\) 
+Then \(\rho_n^{1/2} \lambda^{(k \ell)} = s_1 u^{(k \ell)}\)
+and \(\rho_n^{1/2} \lambda^{(\ell k)} = s_2 v^{(k \ell)}\) 
 for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
@@
 $\hat{\lambda}^{(k \ell)} = \hat{\sigma}^{(k \ell)} \hat{u}^{(k \ell)}$ 
-using the largest signular value of $A$ and its corresponding singular vectors.
+using the largest singular value of $A$ and its corresponding singular vectors.
 
@@
 
-\begin{figure}[tp]
+\begin{figure}[H]
 {\centering \includegraphics{summary_files/figure-latex/mp-1}
@@
 on verifying the sufficient conditions given in Theorem~6 of
-\cite{jmlr-v28-wang13}
+\citet{jmlr-v28-wang13}
 under which sparse subspace clustering based on solving the

### 2021-09-01 summary.tex
syncing what i have so far

@@
 
-To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) is block diagonal with each block
-corresponding to one community and hence \(X (X^\top X)^{-1} X^\top\) is also a
+To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) 
+is block diagonal with each block corresponding to one community, 
+and hence \(X (X^\top X)^{-1} X^\top\) is also a
 block diagonal matrix with each block corresponding to a community. 
@@
 
-Theorem \ref{theorem3} provides perfect community detection given \(P\).
-Letting \(|B|\) be the affinity matrix for graph \(G'\), \(G'\) is
-partitioned into at least \(K\) disjoint subgraphs because each of the
-\(K\) communities have no edges between them. Similar to the subspace
-detection property, it could be the case that some of the communities
-are represented by multiple disjoint subgraphs in \(G'\), in which case
-additional reconstruction is required to identify the communities
-exactly.
-
-Using \(A\) instead of \(P\) to estimate $B$ introduces error, 
+Under the condition that $\|\Lambda\|_F > 0$, where 
+$\Lambda$ is the $n \times K$ matrix of popularity parameters, 
+Theorem \ref{theorem3} provides perfect community detection from \(P\).
+Letting \(|B|\) be the affinity matrix for graph \(G'\), 
+\(G'\) is partitioned into \(K\) disjoint subgraphs, as each of the
+\(K\) communities have no edges between them. 
+Then mapping each subgraph to a community provides a clustering with zero error 
+(up to permutation).
+
+In practice, we observe $A$ instead of $P$. 
+Using \(A\) instead of \(P\) to estimate $B$ as in Theorem \ref{theorem3} introduces error, 
 which converges to \(0\) almost surely. 
@@
 For comparing methods, we define the community detection error as:
-\[L_c(\hat{\sigma}, \sigma; \{v_i\}) =
+\[L_c(\hat{\sigma}, \sigma; V) =
 \min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))\]
@@
 
-For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $p_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
+For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $P_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
 The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$ divided by the number of vertices:

### 2021-09-01 summary.tex
updated stuff based on discussion

@@
 \end{bmatrix}.$$
-Furthermore, since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $U I_{3, 1} U^\top$
-is also an eigendecomposition of $\Pi$ where the fixed
+Furthermore, as $U$ is orthonormal and $I_{3, 1}$ is diagonal, 
+$U I_{3, 1} U^\top$ is also an eigendecomposition of $\Pi$ where the fixed
 points of $\Pi$ are mapped to the eigenvectors $e_1$ and $e_4$
@@
 we have (or can estimate) \(XU\) directly, then both the community
-detection and parameter identification problem are trivial since \(U\)
+detection and parameter identification problem are trivial because \(U\)
 is orthonormal and fixed for each value of \(K\). However, direct
@@
 Letting \(|B|\) be the affinity matrix for graph \(G'\), 
-\(G'\) is partitioned into \(K\) disjoint subgraphs, as each of the
-\(K\) communities have no edges between them. 
-Then mapping each subgraph to a community provides a clustering with zero error 
-(up to permutation).
-
-In practice, we observe $A$ instead of $P$. 
-Using \(A\) instead of \(P\) to estimate $B$ as in Theorem \ref{theorem3} introduces error, 
-which converges to \(0\) almost surely. 
-This is shown via properties of the GRDPG and ASE. 
+where $|\cdot\|$ is applied entry-wise, 
+\(G'\) consists of \(K\) disjoint subgraphs, 
+as $G'$ has no edges between communities. 
+Then all that is left to identify the communities is 
+to assign each subgraph a community label.
+
+In practice, we do not observe $P$ and instead observe $A$. 
+But because $A$ can be considered as an estimator of $P$, 
+one possible approach to community detection given $A$ is applying 
+Theorem \ref{theorem3} to $A$, and indeed, 
+the properties of the GRDPG and ASE suggest $\hat{B}$ constructed from $A$ 
+converges almost surely to $B$ constructed from $P$, 
+including the property that $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ 
+for pair $(i, j)$ in different communities. 
 We use this fact to motivate Orthogonal Spectral Clustering (Alg. 1). 
-In order to show that OSC is consistent, 
+In order to proceed with the proof that that OSC is consistent 
+(Theorem \ref{theorem4}), 
 we introduce the sparsity factor $\rho_n \in (0, 1]$, 
@@
 divided by the number of vertices. 
-The treatment is the same as in \citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
-and \citet{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
-are scaled as $\rho_n^{1/2} \lambda^{(k \ell)}$. 
-This results in the edge probability matrix being scaled as 
+The treatment of the sparsity factor is the same as in 
+\citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
+and \citet{rubindelanchy2017statistical} (GRDPG) 
+in which the popularity vectors are scaled as 
+$\rho_n^{1/2} \lambda^{(k \ell)}$. 
+This results in the edge probabilities being scaled as 
 $P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
-Since $\rho_n$ is unidentifiable, we use the convention that the popularity parameters are drawn such that $\|\Lambda\|_F = O(\sqrt{n})$ and require that the $L^2$ norm of each row is bounded away from 0 and 1.
+Because $\rho_n$ is unidentifiable, 
+we use the convention that the popularity parameters are drawn such that
+$\|\Lambda\|_F = O(\sqrt{n})$ and 
+require that the $L^2$ norm of each row is bounded away from 0 and 1.
 
@@
 
-Theorem \ref{theorem4} provides an algorithm for community detection 
+Theorem \ref{theorem4} provides an approach for community detection 
 that results in zero community detection error as $n \to \infty$. 
@@
 the number of edges of $\hat{B}$ between vertices of different communities 
-that are larger than $\epsilon$ converges to zero with probability 1.
+that are larger than $\epsilon$ converges to zero with probability 1 
+as $n$ increases.
 
@@
 for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
-Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
+Because each $\lambda^{(k \ell)}$ is not strictly identifiable,
 we instead estimate each 
@@
 among 832 butterflies across 10 species. The graph was modified to match
-the example from \citet{noroozi2019estimation}, i.e., only the $K=4$ most
+the example from \citet{noroozi2019estimation}: Only the $K=4$ most
 frequent species were considered, and the similarities were discretized
@@
 $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
-    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
+0$ for all $v \in V^{(-k)}$. Because $z$ is arbitrary, this implies 
+$\|W^{(k)} v\|_{\infty} = 0$ for all $v
 \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ for all $k
@@
 has no effect on the proof of Theorem \ref{theorem5}. Indeed,
-since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
+because $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
 spans the subspace $\mathcal{S}^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
@@
 the remaining block. % \(U^\top \hat{U}\). 
-Since $M^{(k \ell)}$ has one positive eigenvalue and one negative
+Because $M^{(k \ell)}$ has one positive eigenvalue and one negative
 eigenvalue, we see that $W^{*}$ is necessarily of the form $W^{*}

### 2021-09-06 summary_minh_edited.tex
Update git repo

--- 2021-09-01 summary.tex
@@
-% Options for packages loaded elsewhere
+% Options for packages loaded elsewher
 \PassOptionsToPackage{unicode}{hyperref}
@@
 (PABM) and the Generalized Random Dot Product Graph (GRDPG),
-by demonstrating that a PABM is a GRDPG in which communities correspond to 
+by demonstrating that a PABM is a special case of a GRDPG in which communities correspond to 
 mutually orthogonal subspaces of latent vectors. 
@@
 \theoremstyle{remark}
-\newtheorem*{remark}{Remark}
+\newtheorem{remark}{Remark}
 \theoremstyle{definition}
@@
 
-Statistical inference on random graphs requires a suitable probability model. 
-A general probability model for unweighted and undirected graphs is the Bernoulli Graph (also known as the inhomogeneous Erd\"{o}s-R\'{e}nyi model), which assumes that edges occur as independent Bernoulli trials.
-A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
-A trivial example of a Bernoulli Graph is the (homogeneous) Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
-The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
-The present work relates two lines of generalization.
+Statistical inference on random graphs requires a suitable probability
+model.  A general probability model for unweighted and undirected
+graphs is the Bernoulli Graph (also known as the inhomogeneous
+Erd\"{o}s-R\'{e}nyi model), which assumes that edges occur as
+independent Bernoulli trials.  A Bernoulli Graph is characterized by
+an edge probability matrix $P=[P_{ij}]$, where an edge between
+vertices $i$ and $j$ occurs with success probability $P_{ij}$.  A
+trivial example of a Bernoulli Graph is the (homogeneous)
+Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which
+the vertices of the random graph are fixed and possible edges occur
+independently with fixed probability $P_{ij} = p$ for all $i,j$.  The requirement that
+$P_{ij} \equiv p$ for all $i,j$ is too strong for most applications, and various
+researchers have weakened that requirement in various ways.  The
+present work relates two lines of generalization.
 
@@
 Thus, if the latent positions are $x_1,\ldots,x_n \in \mathbb{R}^d$
-and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$,
+and $X = \bigl[ x_1 \mid \dots \mid x_n \bigr]^\top$,
 then the edge probability matrix is $P = XX^\top$.  Clearly, any
@@
 \label{pabm}
-Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-graph $G = (V, E)$ with adjacency matrix $A$ 
-such that $A \sim \mathrm{Bernoulli}(P)$.
-Let each vertex have a community label between $1$ and $K$.
-Then $G$ is drawn from a Popularity Adjusted Block Model if
-each vertex has $K$ popularity parameters that describe its affinity
-toward each of the $K$ communities, i.e., vertex $i$ has popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK}$, and  $P_{ij} = \lambda_{i z_j}
-\lambda_{j z_i}$ for all $i \leq j$.
-
-Another characterization of the PABM is as follows.
-Let the rows and columns of $P$ be arranged by community label
-such that $n_k \times n_{\ell}$ block $P^{(k \ell)}$
-describes the edge probabilities between vertices in communities
-$k$ and $\ell$; note that $P^{(k \ell)} = (P^{(k \ell)})^\top$.
-Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
+Let $K \geq 1$ be an integer and let $\Lambda \in \mathbb{R}^{n \times
+  K}$ be a matrix with entries in $[0,1]$. Let $z_1, z_2, \dots, z_n
+\in \{1,2,\dots,K\}$. A graph $G$ with adjacency matrix $A$ is said to
+be a popularity adjusted block model graph with $K$ communities, popularity vectors
+$\Lambda$, and sparsity parameter $\rho_n \in (0,1]$ if $A \sim \BG(P)$
+where the edge probability matrix $P$ has entries of the form
+$$P_{ij} = \rho_n \lambda_{iz_j} \lambda_{jz_i}.$$
+\end{definition}
+
+\begin{remark}
+  \label{rem:pabm_view2}
+% Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
+% graph $G = (V, E)$ with adjacency matrix $A$ 
+% such that $A \sim \mathrm{Bernoulli}(P)$.
+% Let each vertex have a community label between $1$ and $K$.
+% Then $G$ is drawn from a Popularity Adjusted Block Model if
+  In a PABM each vertex $i$ has $K$ popularity parameters $\lambda_{i1}, \dots, \lambda_{iK}$, that describe its affinity
+toward each of the $K$ communities. % , i.e., vertex $i$ has popularity parameters
+% $\lambda_{i1}, ..., \lambda_{iK}$, and  $P_{ij} = \lambda_{i z_j}
+% \lambda_{j z_i}$ for all $i \leq j$. 
+Another view of a PABM is as follows.
+Let $\tilde{P}$ be the matrix obtained by permuting the rows and
+columns of $P$ so that the vertices are arranged according to
+increasing values of the community assignments $z_i \in
+\{1,2,\dots,K\}$. Denote by $\tilde{P}^{(k \ell)}$ the $n_k \times n_{\ell}$
+submatrix of $\tilde{P}$ corresponding to the
+edge probabilities between vertices in communities
+$k$ and $\ell$. Here $n_k = |\{ i \colon z_i = k\}|$ is the number of
+vertices assigned to community $k$, for $k = 1,2\dots,K$.
+Note that $\tilde{P}^{(k \ell)} = (\tilde{P}^{(\ell k)})^\top$. Next
+let $\lambda^{(k \ell)} = \{\lambda_{i \ell} \colon z_i = k\} \in
+\mathbb{R}^{n_k}$; the elements of $\lambda^{(k \ell)}$ are the affinity parameters toward the $\ell$th community of
+all vertices in the $k$th community. Define $\lambda^{(\ell k)}$
+analogously. Then each block $\tilde{P}^{(k \ell)}$ can be written as the outer product of two vectors:
 \begin{equation} \label{eq:pabm}
-  P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}
+  \tilde{P}^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}.
 \end{equation}
-for a set of $K^2$ popularity vectors 
-$\{\lambda^{(st)}\}_{1 \leq s \leq K, 1 \leq t \leq K}$ where each
-$\lambda^{(st)}$ is a column vector of dimension $n_s$. Then a graph $G$
-is drawn from a PABM with parameters
-$\{\lambda^{(st)}\}$ if its adjacency matrix $A$
-satisfies $A \sim \mathrm{BernoulliGraph}(P)$.
-\end{definition}
-
-We will use the notation \(A \sim \PABM(\{\lambda^{(k \ell)}\}_K)\) to denote
-a random adjacency matrix \(A\) drawn from a PABM with parameters
-\(\{\lambda^{(k \ell)}\}\) consisting of \(K\) underlying communities.
+% for a set of $K^2$ popularity vectors 
+% $\{\lambda^{(st)}\}_{1 \leq s \leq K, 1 \leq t \leq K}$ where each
+% $\lambda^{(st)}$ is a column vector of dimension $n_s$.
+% . Then a graph $G$
+% is drawn from a PABM with parameters
+% $\{\lambda^{(st)}\}$ if its adjacency matrix $A$
+% satisfies $A \sim \mathrm{BernoulliGraph}(P)$.
+We will henceforth use the notation \(A \sim \PABM(\{\lambda^{(k
+  \ell)}\}_K, \rho_n)\) to denote
+a random adjacency matrix \(A\) drawn from a PABM with $K$ communities, popularity parameters
+\(\{\lambda^{(k \ell)}\}\) and sparsity parameter $\rho_n$.
+\end{remark}
+
+The sparsity parameter $\rho_n$ in the definition of the PABM
+influences the degrees of the vertices in the sampled graphs $A \sim
+\BG(P)$. In particular, for a fixed $\Lambda$, the graphs become
+sparser as $\rho_n$ decreases. Note that $\rho_n$ and $\Lambda$ are not
+uniquely identifiable, i.e., we can scale $\rho_n$ by a constant $c > 0$ and
+scale $\Lambda$ by $c^{-1/2}$ without changing the edge probabilities
+in $P$. Thus, for ease of exposition, we shall assume henceforth that
+(1) $\Lambda$ is normalized to have Frobenius norm $\|\Lambda\|_{F} =
+\sqrt{n}$ and (2) the $\ell_2$ norms of the rows of
+$\|\Lambda\|$ are all bounded away from $0$. Under these two
+assumptions, the sparsity parameter $\rho_n$ can now be viewed as
+controlling the density of $A$, i.e., the average degree of
+$A$ grows at rate $n \rho_n$ and the number of edges grows at rate
+$n^2 \rho_n$. The choice $\rho_n \equiv 1$ and $\rho_n \rightarrow 0$
+then corresponds to the dense graphs regime and semi-sparse graphs
+regime, respectively.
 
@@
 \label{grdpg}
-Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$ for some symmetric $P \in
-[0,1]^{n \times n}$. 
-If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
-\begin{equation} \label{eq:grdpg}
-  P = X I_{p,q} X^\top
-\end{equation}
-for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-$G$ is drawn from the Generalized Random Dot Product Graph 
-with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
+Let $p \geq 1$ and $q \geq 0$ be integers. Define $I_{p,q}$ as the
+block diagonal matrix $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0
+  \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$ where $I_{p}$ and $I_{q}$
+are the identity matrices of dimensions $p \times p$ and $q \times q$
+respectively. Next let $\mathcal{X}$ be a subset of $\mathbb{R}^{d}$
+such that $x^{\top} I_{p,q} y \in [0,1]$ for all $x,y \in
+\mathcal{X}$. Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
+a $n \times d$ matrix with rows $x_i \in \mathcal{X}$. A graph $G$
+with adjacency matrix $A$ is said to be a generalized random dot
+product graph with latent positions $X$, sparsity parameter $\rho_n
+\in (0,1]$
+and signature $(p,q)$ if $A \sim \BG(P)$ where the edge probability
+matrix $P$ is given
+by $P = \rho_n X I_{p,q} X^{\top}$, i.e., the entries of $P$ are of
+the form
+$$P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$$
+% Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$ for some symmetric $P \in
+% [0,1]^{n \times n}$. 
+% If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
+% \begin{equation} \label{eq:grdpg}
+%   P = X I_{p,q} X^\top
+% \end{equation}
+% for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
+% $G$ is drawn from the Generalized Random Dot Product Graph 
+% with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
 \end{definition}
-
-We will use the notation \(A \sim \GRDPG_{p,q}(X)\) to denote a random
-adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\). If instead of fixed latent positions, 
-they are drawn $X_1, ..., X_n \iid F$, 
-we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
-
+We will use the notation \(A \sim \GRDPG_{p,q}(X;\rho_n)\) to denote a random
+adjacency matrix \(A\) drawn from latent positions \(X\), sparsity
+parameter $\rho_n$ and signature
+\((p, q)\). % If instead of fixed latent positions, 
+% they are drawn $X_1, ..., X_n \iid F$, 
+% we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
 \begin{remark}
+  \label{rem:ase}
 We can use Adjacency Spectral Embedding (ASE) 
@@
 \begin{remark}
+  \label{rem:non_identifiable}
 The latent vectors that produce $X I_{p,q} X^\top = P$ are not unique
@@
 \subsection{The Geometry of PABMs}\label{connecting-the-pabm-to-the-grdpg}}
-
 Now that we defined the PABM and GRDPG, 
-we show the special geometry of the PABM when viewed as a GRDPG.
-
+we show the special geometry of the PABM when viewed as a GRDPG. For
+ease of exposition, and without loss of generality, we drop the
+dependency on the sparsity parameter $\rho_n$ 
+and assume $\rho_n \equiv 1$ throughout this subsection. 
 \begin{theorem}[The latent configuration of the PABM]
@@
 $K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
-that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
+that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) /
+  2}(\tilde{\Pi}XU)$. Here $\tilde{\Pi}$ is the permutation matrix
+such that $P = \tilde{\Pi} \tilde{P} \tilde{\Pi}^{\top}$ where the
+rows and columns of $\tilde{P}$ are arranged according to increasing values of the
+community labels (see Remark~\ref{rem:pabm_view2}). 
 \end{theorem}
@@
 
-For the $K = 2$ case, the proof is straightforward. Let 
+For the $K = 2$ case, the proof is straightforward. We will first work with
+the matrix $\tilde{P}$. Note that $\tilde{P}$ has the form
+$$P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} &
+  P^{(22)} \end{bmatrix} = \begin{bmatrix} \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
+  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)}
+  (\lambda^{(22)})^\top \end{bmatrix}.$$
+  Now let
 $$X = \begin{bmatrix}
@@
 Then by straightforward matrix multiplication, we obtain 
-\[X U I_{3, 1} U^\top X^\top =
+$$X U I_{3, 1} U^\top X^\top =
 \begin{bmatrix}
@@
   \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}, \]
-which is a GRDPG with latent vectors described by $X U$.
+\end{bmatrix} = \tilde{P}$$
+and hence $\tilde{P}$ also corresponds to the edge probability matrix of GRDPG
+with latent vectors described by $X U$. As $P = \tilde{\Pi} \tilde{P}
+\tilde{\Pi}^{\top}$ we conclude that $P$ has latent vectors described
+by $\tilde{\Pi} X U$.
 
@@
 
-For the general case, we first define the following matrices
+For the general case, we once again consider $\tilde{P}$ as defined in
+Remark~\ref{rem:pabm_view2}.  
+We first define the following matrices
 \begin{gather}
@@
 \in \mathbb{R}^{n_k \times K}, \quad
-X = \text{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
+X = \mathrm{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
 \mathbb{R}^{n \times K^2}, \\
-L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
+L^{(k)} = \mathrm{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
 \mathbb{R}^{n \times K}, \quad
@@
 \end{gather}
-It is then straighforward to verify that
+It is then straightforward to verify that
 \begin{gather*}
-  XY^{\top} = \text{blockdiag}(\Lambda^{(1)}, \dots,
+  XY^{\top} = \mathrm{blockdiag}(\Lambda^{(1)}, \dots,
 \Lambda^{(K)}) \begin{bmatrix} L_1^{\top} \\ \vdots \\
@@
 \end{gather*}
-We therefore have $P = X Y^\top$. 
+We therefore have $\tilde{P} = X Y^\top$. 
 Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
-$\Pi$ and hence $P = X \Pi X^\top$.
+$\Pi$ and hence $\tilde{P} = X \Pi X^\top$.
 The permutation described by $\Pi$ has $K$ fixed points, which correspond to
@@
 \begin{equation} \label{eq:pabm-grdpg}
-P = X U I_{p, q} (X U)^\top.
+\tilde{P} = X U I_{p, q} (X U)^\top.
 \end{equation}
 In summary we can describe the PABM with $K$ communities as a GRDPG with latent
-positions $X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
+positions $\tilde{\Pi} X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
 \tfrac{1}{2} K (K - 1)\bigr)$.
@@
 \end{gather*}
-Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ 
+Then $Y = X \Pi$ and $\tilde{P} = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ 
 permutation matrix of the form
@@
 \end{enumerate}
-We can therefore write $\Pi$ as $\Pi = U I_{6, 3} U^\top$ where the first three
+We can thus write $\Pi$ as $\Pi = U I_{6, 3} U^\top$ where the first three
 columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the
-{\em fixed} points, the next three columns consist of eigenvectors
-$(e_k + e_{\ell}) / \sqrt{2}$, and the last three columns consist of eigenvectors
+fixed points, the next three columns are the eigenvectors
+$(e_k + e_{\ell}) / \sqrt{2}$, and the last three columns are the eigenvectors
 $(e_k - e_{\ell}) / \sqrt{2}$ for $(k, \ell) \in
@@
 
-The matrix $P$ is then the edge probabilities matrix for a 
+The matrix $\tilde{P}$ is then the edge probabilities matrix for a 
 Generalized Random Dot Product Graph whose latent positions 
@@
   0 & -\frac{\lambda^{(31)}}{\sqrt{2}} & -\frac{\lambda^{(32)}}{\sqrt{2}}
-\end{bmatrix}.$$
+\end{bmatrix}$$
+and the latent positions for $P$ is a permutation of the rows of
+$XU$. 
 \end{example}
@@
 collected collected into matrix
-\(C = \bigl[ c_1 \mid c_2 \mid \cdots \mid c_n \bigr]^\top\) to
+\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
 construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
 exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
-consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
+consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
 \(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
@@
 \begin{equation} \label{eq:ssc}
-c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i} c\|^2_2 + \lambda \|c\|_1
+c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i} c\|^2_2 + \vartheta \|c\|_1
 \end{equation}
-for some sparsity parameter \(\lambda > 0\). The \(c_i\) vectors are
+for some sparsity parameter \(\vartheta > 0\). The \(c_i\) vectors are
 then collected into \(C\) and \(B\) as before.
@@
   \label{def:subspace_detection}
-Let $X = \bigl[ x_1 \mid x_2 \mid \cdots \mid x_n \bigr]^\top$ be noisy
+Let $X = \bigl[ x_1 \mid \cdots \mid x_n \bigr]^\top$ be noisy
 points sampled from $K$ subspaces, i.e., $x_i = y_i + z_i$ where the
@@
 is said to satisfy the subspace detection property 
-with sparsity parameter $\lambda$ if each column of
+with sparsity parameter $\vartheta$ if each column of
 $C$ has nonzero $\ell_2$ norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
@@
 \begin{remark}
-In practice, a noisy sample $X$ often does not obey the subspace detection
-property. In such cases, $B$ is treated as an affinity matrix for a graph, 
-which requires a partitioning step to obtain the final clustering. On the other
-hand, if $X$ does obey the subspace detection property, $B$ describes a graph
-with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
-property holds, there are exactly $K$ subgraphs which map onto each subspace,
-but it could be the case that some of the subspaces are represented by
-multiple disconnected subgraphs. The subspace detection property is contingent
-on choosing a sufficiently large sparsity parameter $\lambda$.
+  One of the most common approach to show that SSC works for a noisy
+  sample $X$ is to show that 
+  $X$ satisfies the
+  subspace detection property for some choice of $\vartheta$; recall
+  that $\vartheta$ is the sparsity parameter $\vartheta$ for
+the LASSO problems in Eq.~\eqref{eq:ssc}. However, this is not sufficient to
+guarantee that SSC perfectly recovers the underlying subspaces.
+More specifically
+if $X$ satisfies the
+subspace detection property then $B$ describes a graph with {\em at least}
+$K$ disconnected subgraphs, with the ideal case being that there are exactly $K$ subgraphs which map onto each
+subspace. Nevertheless it is also possible that the $K$ subspaces are represented by
+$K' > K$
+multiple disconnected subgraphs and we cannot, at least without a subsequent
+post-processing step, recover the $K$ subspaces directly from $B$;
+see \citet{sdp_sufficiency} and \citet{liu_ssc}
+for further discussions. Therefore in practice $B$ is usually treated as an
+affinity matrix and, as we allude to earlier, the rows of $B$ are partitioned using some
+clustering algorithm to obtain the final clustering. % Finally we note that the subspace detection property is contingent
+% on choosing a sufficiently large sparsity parameter $\vartheta$ for
+% the LASSO problems in Eq.~\eqref{eq:ssc}.
 \end{remark}
 
-Theorem \ref{theorem2} suggests that SSC is appropriate for community
-detection for the PABM. More precisely, Theorem \ref{theorem2} says that
-each community consists of a \(K\)-dimensional subspace, and together
-the subspaces lie in \(\mathbb{R}^{K^2}\). The natural approach then is
-to perform SSC on the ASE of \(P\) or \(A\).
-\citet{noroozi2019estimation} instead applied SSC to \(P\) and
-\(A\), foregoing embedding altogether.
-
-Using results from \citet{soltanolkotabi2012}, it can be easily shown
-that the subspace detection property holds for \(XU\), which is an ASE
-of \(P\). More specifically, if points lie exactly on mutually
-orthogonal subspaces, then the subspace detection property will hold
-with probability 1, and this is exactly the case for the PABM (Theorem
-\ref{theorem2}). Much of our work is then built on
-\citet{rubindelanchy2017statistical}, who describe the convergence
-behavior of the ASE of \(A\) to the ASE of \(P\), and
-\citet{jmlr-v28-wang13}, who show the necessary conditions for the
+Theorem~\ref{theorem2} suggests that SSC is appropriate for community
+detection for the PABM, provided that we observe the edge probabilities
+matrix $P$. More precisely, given the matrix $\tilde{P}$ obtained by
+permuting the rows and columns of $P$ as described in
+Remark~\ref{rem:pabm_view2} we can recover $XU$ up
+to some non-identifiability indefinite orthogonal transformation $Q$. 
+% from the singular value decomposition of $P$. 
+Then using results from \citet{soltanolkotabi2012}, it can be easily shown
+that the subspace detection property holds for \(XU\). Indeed, the columns
+of \(XU\) from different communities correspond to mutually
+orthogonal subspaces. This then implies that the
+subspace detection property also holds for $XUQ$ for all invertible
+transformation $Q$ and hence the subspace detection property also holds for
+$\tilde{\Pi} X U Q$ for any $n \times n$ permutation matrix $\tilde{\Pi}$.
+
+However, because we do not observe $P$ but rather only the noisy
+adjacency matrix $A \sim \BG(P)$, the natural approach then is
+to perform SSC on the rows of the spectral embedding of \(A\). We will
+show in Theorem~\ref{theorem5} that, with probability converging to
+one as $n \rightarrow \infty$, the rows of the ASE of $A$ also
+satisfy the subspace detection property. 
+Theorem~\ref{theorem5} builds upon existing work by
+\citet{rubindelanchy2017statistical} who describe the convergence
+behavior of the ASE of \(A\) to that of \(\tilde{\Pi} XU\), and
+\citet{jmlr-v28-wang13} who show the necessary conditions for the
 subspace detection property to hold in noisy cases where the points lie
-near subspaces.
-
+near subspaces. Finally we emphasize that while
+\citet{noroozi2019estimation} also considered the use of SSC for
+community recovery in PABM, they instead applied SSC to the rows of
+\(A\) itself, foregoing
+the embedding step altogether. It is however much harder to show that
+the rows of $A$ satisfy the subspace detection property and thus, to
+the best of our knowledge, there is currently no consistency result
+regarding the application of SSC to the rows of $A$.
+
+% Using results from \citet{soltanolkotabi2012}, it can be easily shown
+% that the subspace detection property holds for \(XU\), which is an ASE
+% of \(P\). More specifically, if points lie exactly on mutually
+% orthogonal subspaces, then the subspace detection property will hold
+% with probability 1, and this is exactly the case for the PABM (Theorem
+% \ref{theorem2}). Much of our work is then built on
+% \citet{rubindelanchy2017statistical}, who describe the convergence
+% behavior of the ASE of \(A\) to the ASE of \(P\), and
+% \citet{jmlr-v28-wang13}, who show the necessary conditions for the
+% subspace detection property to hold in noisy cases where the points lie
+% near subspaces.
 \hypertarget{community-detection}{%
@@
 
-\begin{algorithm}[t]
+We previously stated in Theorem~\ref{theorem2} one possible set of latent positions that result in
+the edge probability matrix of a PABM, namely
+\(P = \tilde{\Pi} (XU) I_{p, q}
+(XU)^\top \tilde{\Pi}^{\top}\) where $X$ is block diagonal and
+$\tilde{\Pi}$ is a permutation matrix.  
+Furthermore, the explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
+such that points within each community lie on \(K\)-dimensional
+orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $z_i \not = z_j$. 
+Thus if
+we have (or can estimate) \(XU\) directly, then both the community
+detection and parameter identification problem are trivial because \(U\)
+is orthonormal and fixed for each value of \(K\). 
+However, direct
+identification or estimation of \(XU\) is possibly difficult 
+due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable}).
+when we are given only $P$. % of the latent
+% positions of a GRDPG.  
+% due to the
+% non-identifiability of \(XU\) when we are only given $P$, 
+% direct
+% identification or estimation of \(XU\) is possibly difficult 
+% if we are only given $P$. 
+% (\cite{rubindelanchy2017statistical}).
+More specifically, suppose we find a matrix $Y \in
+\mathbb{R}^{n \times K^2}$
+such that \(P = Y I_{p, q} Y^\top\). Then it is generally the case that
+we  \( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
+$Q \in \mathbb{O}(p,q)$. %  as there is no
+% general matrix decomposition that yield 
+% The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
+% such that points within each community lie on \(K\)-dimensional
+% orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
+% removes the orthogonality property but retains the property that each
+% community is represented by a \(K\)-dimensional subspace. 
+% However, because $\tilde{\Pi}$ is unknown (indeed, if we know $\tilde{\Pi}$
+% then we also know the community labels), 
+% Therefore, even if we
+% start with the true but unknown edge probability matrix $P$ we
+% will not be able to recover the latent positions \(XU\).
+However since \(Q\) is not necessarily an
+orthogonal matrix and hence, if $y_i$ denote the $i$th row of $Y$ then $\langle U^{\top} x_i, U^{\top} x_j \rangle
+\neq \langle y_i, y_j \rangle$.
+This prevents us from transferring the orthogonality property of
+\(XU\) directly to $Y$.
+
+Nevertheless by using the special geometric structure of $X$ we can circumvent the
+non-identifiability of $Y$ and $XU$ by using instead the rows of the
+matrix $V$ of eigenvectors (corresponding to the non-zero eigenvalues) of $P$. In particular $V$ is identifiable
+up to orthogonal transformations and furthermore, due to the block
+diagonal structure of $X$, the rows of $V$ also lie on $K$ distinct orthogonal
+subspaces and hence $v_i^{\top} v_j = 0$ whenever $z_i \not = z_j$. 
+% The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
+% such that points within each community lie on \(K\)-dimensional
+% orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
+% removes the orthogonality property but retains the property that each
+% community is represented by a \(K\)-dimensional subspace. Therefore, the
+% ASE of \(P\) still consists of subspaces 
+% that correspond to communities of the PABM, even if they are not orthogonal, 
+% suggesting the use of SSC. Before exploring SSC, we will first consider
+% a different approach.
+\begin{theorem}
+\label{theorem3}
+Let $P = V D V^\top$ be the spectral decomposition of the edge probability
+matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are
+from different communities.
+\end{theorem}
+\begin{proof}
+% \begin{lemma}
+% \label{lemma1}
+% Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
+% matrix for a PABM.
+We first show that $V V^\top =
+\tilde{\Pi} X (X^\top X)^{-1} X^\top \tilde{\Pi}^{\top} $ where $X$ is 
+defined as in Eq.~\eqref{eq:xy}. Indeed, by Theorem 2, 
+\(P = \tilde{\Pi} X U I_{p, q} U^\top X^\top \tilde{\Pi}\) for $p = K(K+1)/2$ and $q = K(K-1)/2$. 
+The eigendecomposition \(P = V D V^\top\) also yields $P = V
+|D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ where \(|\cdot|^{1/2}\) is
+applied entry-wise. Now let $Y = \tilde{\Pi} XU$ and $\tilde{Y} = V|D|^{1/2}$; note that
+$Y$ and $\tilde{Y}$ both have full column ranks. 
+Because $P = Y I_{p,q} Y^{\top} = \tilde{Y} I_{p,q} \tilde{Y}^{\top}$, we have
+$$Y = \tilde{Y} I_{p,q} \tilde{Y}^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}.$$
+Let $Q = I_{p,q} \tilde{Y}^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}$ and note that
+$Y = \tilde{Y} Q$. We then have
+\begin{equation*}
+  \begin{split}
+  Q^{\top} I_{p,q} Q &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} \tilde{Y} I_{p,q}
+I_{p,q} I_{p,q} \tilde{Y}^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
+Y^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} =  I_{p,q}
+\end{split}
+\end{equation*}
+and hence $Q$ is an indefinite orthogonal matrix.
+
+Let $R = U Q |D|^{-1/2}$ and note that $V = \tilde{\Pi} XR$. Because $R$ is
+invertible, we can write
+$$\tilde{\Pi} X (X^{\top} X)^{-1} X^{\top} \tilde{\Pi}^{\top} =
+\tilde{\Pi} X R (R^{\top} X^{\top} X R)^{-1}
+R^{\top} X^{\top} \tilde{\Pi}^{\top}.$$ 
+Furthermore, as $V$ has orthonormal columns, $R^{\top} X^{\top} X R =
+V^{\top} \tilde{\Pi} \tilde{\Pi}^{\top} V = V^{\top} V = I$. We thus conclude
+$$\tilde{\Pi} X (X^{\top} X)^{-1} X^{\top} \tilde{\Pi}^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
+as desired.
+
+To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) 
+is block diagonal with each block corresponding to one community, 
+and hence \(X (X^\top X)^{-1} X^\top\) is also a
+block diagonal matrix with each block corresponding to a community. 
+As $B = n VV^{\top} = n \tilde{\Pi} X (X^\top X)^{-1} X^\top \tilde{\Pi}^{\top}$, 
+we conclude that $B_{ij} = 0$ 
+whenever vertices $i$ and $j$ belong to different communities.  
+\end{proof}
+Under the condition that $\|\Lambda\|_F > 0$, where 
+$\Lambda$ is the $n \times K$ matrix of popularity parameters, 
+Theorem \ref{theorem3} provides perfect community detection from
+\(P\). More specifically let \(|B|\) be the affinity matrix for graph \(G'\), 
+where $|\cdot|$ is applied entry-wise. Then
+\(G'\) consists of exactly \(K\) disjoint subgraphs, 
+as $G'$ has no edges between communities. 
+All that is left to identify the communities is 
+to assign each subgraph a distinct community label. 
+In practice, we do not observe $P$ and instead only observe the noisy
+$A \sim \BG(P)$. A natural approach is then to use
+the affinity matrix $\hat{B} = n \hat{V} \hat{V}^{\top}$ where
+$\hat{V}$ is the matrix of eigenvectors (corresponding to the largest
+eigenvalues in modulus) of $A$. The resulting procedure, named
+Orthogonal Spectral Clustering, is presented in
+Algorithm~\ref{alg:osc}.
+The following result leverages existing theoretical properties
+of ASE for estimating of latent positions in a GRDPG \citep{rubindelanchy2017statistical} to show that
+$\hat{B}$ converges almost surely to $B$; in particular 
+ $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ 
+for pair $(i, j)$ in different communities. 
+\begin{algorithm}[tp]
   \label{alg:osc}
@@
 
-We previously stated one possible set of latent positions that result in
-the edge probability matrix of a PABM, \(P = (XU) I_{p, q} (XU)^\top\). If
-we have (or can estimate) \(XU\) directly, then both the community
-detection and parameter identification problem are trivial because \(U\)
-is orthonormal and fixed for each value of \(K\). However, direct
-identification or estimation of \(XU\) is not possible
-(\cite{rubindelanchy2017statistical}).
-
-More specifically, if we decompose \(P = Z I_{p, q} Z^\top\), then
-there exists a \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Therefore, even if we
-start with the true but unknown edge probability matrix $P$ we
-will not be able to recover the ``original'' latent positions \(XU\).
-Note that unlike in the case of the RDPG, \(Q\) is not necessarily an
-orthogonal matrix and hence if \(z_i\)'s
-are the rows of \(XU\) then
-\(\|z_i - z_j\|^2 \neq \|Q z_i - Q z_j\|^2\), and
-\(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
-prevents us from using the orthogonality property of \(XU\) directly.
-
-The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
-such that points within each community lie on \(K\)-dimensional
-orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
-removes the orthogonality property but retains the property that each
-community is represented by a \(K\)-dimensional subspace. Therefore, the
-ASE of \(P\) still consists of subspaces 
-that correspond to communities of the PABM, even if they are not orthogonal, 
-suggesting the use of SSC. Before exploring SSC, we will first consider
-a different approach.
-
-\begin{theorem}
-\label{theorem3}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are
-from different communities.
-\end{theorem}
-\begin{proof}
-% \begin{lemma}
-% \label{lemma1}
-% Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-% matrix for a PABM.
-We first show that $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in Eq.~\eqref{eq:xy}. Indeed, by Theorem 2, 
-\(P = X U I_{p, q} U^\top X^\top\) for $p = K(K+1)/2$ and $q = K(K-1)/2$.
-
-The eigendecomposition \(P = V D V^\top\) also yields $P = V
-|D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ where \(|\cdot|^{1/2}\) is
-applied entry-wise. Now let $Y = XU$ and $Z = V|D|^{1/2}$; note that
-$Y$ and $Z$ both have full column ranks. 
-Because $P = Y I_{p,q} Y^{\top} = Z I_{p,q} Z^{\top}$, we have
-$$Y = Z I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}.$$
-Let $Q = I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}$ and note that
-$Y = Z Q$. We then have
-\begin{equation*}
-  \begin{split}
-  Q^{\top} I_{p,q} Q &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Z I_{p,q}
-I_{p,q} I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
-Y^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} =  I_{p,q}
-\end{split}
-\end{equation*}
-and hence $Q$ is an indefinite orthogonal matrix.
-
-Let $R = U Q |D|^{-1/2}$ and note that $V = XR$. Because $R$ is
-invertible, we can write
-
-$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
-R^{\top} X^{\top}.$$
-
-Furthermore, as $V$ has orthonormal columns, $R^{\top} X^{\top} X R =
-V^{\top} V = I$. We thus conclude
-$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
-as desired.
-
-To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) 
-is block diagonal with each block corresponding to one community, 
-and hence \(X (X^\top X)^{-1} X^\top\) is also a
-block diagonal matrix with each block corresponding to a community. 
-As $B = n VV^{\top} = n X (X^\top X)^{-1} X^\top$, 
-we conclude that $B_{ij} = 0$ 
-whenever vertices $i$ and $j$ belong to different communities.  
-\end{proof}
-
-Under the condition that $\|\Lambda\|_F > 0$, where 
-$\Lambda$ is the $n \times K$ matrix of popularity parameters, 
-Theorem \ref{theorem3} provides perfect community detection from \(P\).
-Letting \(|B|\) be the affinity matrix for graph \(G'\), 
-where $|\cdot\|$ is applied entry-wise, 
-\(G'\) consists of \(K\) disjoint subgraphs, 
-as $G'$ has no edges between communities. 
-Then all that is left to identify the communities is 
-to assign each subgraph a community label.
-
-In practice, we do not observe $P$ and instead observe $A$. 
-But because $A$ can be considered as an estimator of $P$, 
-one possible approach to community detection given $A$ is applying 
-Theorem \ref{theorem3} to $A$, and indeed, 
-the properties of the GRDPG and ASE suggest $\hat{B}$ constructed from $A$ 
-converges almost surely to $B$ constructed from $P$, 
-including the property that $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ 
-for pair $(i, j)$ in different communities. 
-We use this fact to motivate Orthogonal Spectral Clustering (Alg. 1). 
-In order to proceed with the proof that that OSC is consistent 
-(Theorem \ref{theorem4}), 
-we introduce the sparsity factor $\rho_n \in (0, 1]$, 
-which can be characterized as the average expected degree of the vertices 
-divided by the number of vertices. 
-The treatment of the sparsity factor is the same as in 
-\citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
-and \citet{rubindelanchy2017statistical} (GRDPG) 
-in which the popularity vectors are scaled as 
-$\rho_n^{1/2} \lambda^{(k \ell)}$. 
-This results in the edge probabilities being scaled as 
-$P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
-Because $\rho_n$ is unidentifiable, 
-we use the convention that the popularity parameters are drawn such that
-$\|\Lambda\|_F = O(\sqrt{n})$ and 
-require that the $L^2$ norm of each row is bounded away from 0 and 1.
+% In order to proceed with the proof that that OSC is consistent 
+% (Theorem \ref{theorem4}), 
+% we introduce the sparsity factor $\rho_n \in (0, 1]$, 
+% which can be characterized as the average expected degree of the vertices 
+% divided by the number of vertices. 
+% The treatment of the sparsity factor is the same as in 
+% \citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
+% and \citet{rubindelanchy2017statistical} (GRDPG) 
+% in which the popularity vectors are scaled as 
+% $\rho_n^{1/2} \lambda^{(k \ell)}$. 
+% This results in the edge probabilities being scaled as 
+% $P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
+% Because $\rho_n$ is unidentifiable, 
+% we use the convention that the popularity parameters are drawn such that
+% $\|\Lambda\|_F = O(\sqrt{n})$ and 
+% require that the $L^2$ norm of each row is bounded away from 0 and 1.
 
@@
 \label{theorem4}
-Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC
-(Alg. 1). Then for all pairs $(i, j)$ belonging to different communities
-and sparsity factor satisfying $n \rho_n = \omega(\log^{4}{n})$, we have
-\begin{equation} \label{eq:thm4}
-\max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
-O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
+Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix
+obtained from OSC as described in Algorithm~\ref{alg:osc}. Then for $n \rho_n = \omega(\log^{4}{n})$, we have
+\begin{equation} \label{eq:thm4a}
+\max_{i, j} |\hat{B}_{ij} - B_{ij}| = O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
 \end{equation}
-with high probability.
-
-Thus for all pairs $(i, j)$ in different communities,
-we have uniform convergence $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$. 
+with high probability. In particular $\hat{B}_{ij} -
+B_{ij} \overset{\mathrm{a.s}}{\rightarrow} 0$ where the convergence is
+uniform over all $i,j$. Hence for all pairs $(i,j)$ in different
+communities we have
+% \begin{equation} \label{eq:thm4}
+% \max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
+% O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
+% \end{equation}
+% with high probability. 
+% Thus for all pairs $(i, j)$ in different communities,
+% we have uniform convergence 
+$\hat{B}_{ij} \overset{\mathrm{a.s}}{\rightarrow} 0$. 
 \end{theorem}
-
-Theorem \ref{theorem4} provides an approach for community detection 
-that results in zero community detection error as $n \to \infty$. 
-For example, if we use edge thresholding to partition $\hat{B}$, 
-Theorem \ref{theorem4} implies that for any $\epsilon > 0$, 
+Theorem~\ref{theorem4} guarantees that for any $\epsilon > 0$, 
 the number of edges of $\hat{B}$ between vertices of different communities 
-that are larger than $\epsilon$ converges to zero with probability 1 
-as $n$ increases.
+that are larger than $\epsilon$ converges to zero with probability
+converging to one as $n$ increases. Thus, by using $\hat{B}$, we can 
+perfectly recover all the latent community assignments $z_1, z_2,
+\dots, z_n$, i.e., the number of mis-clustered vertices is zero
+asymptotically almost surely. We note that Theorem~\ref{theorem4} is
+stronger than existing results in the literature; in particular
+Theorem~1 of \citet{307cbeb9b1be48299388437423d94bf1} (the paper that
+originally introduces the PABM model) only guarantees that the {\em
+  proportion} of mis-clustered vertices converges to $0$ as $n
+\rightarrow \infty$. Furthermore Theorem~1
+of \citet{307cbeb9b1be48299388437423d94bf1} also requires the sparsity
+parameter $\rho_n$ to satisfies $n \rho_n^2 = \omega(\log^2{n})$ which
+is a considerably stronger assumption than the assumption $n \rho_n =
+\omega(\log^{4}{n})$ used in Theorem~\ref{theorem4}. Indeed, $n
+\rho_n^2 = \omega(\log^{2}{n})$ implies $n \rho_n = \omega(n^{1/2})$. 
+We emphasize that the assumption $n \rho_n = \omega(\log^{c}{n})$ for
+some constant $c > 1$ is commonly used in the context of graph
+estimation using spectral methods.
 
@@
 a natural path toward using SSC for community detection. 
-We established in Theorem \ref{theorem2} that an ASE of the edge
+In particular we established in Theorem \ref{theorem2} that an ASE of the edge
 probability matrix \(P\) can be constructed from a latent vector configuration 
 consisting of orthogonal subspaces. Theorem \ref{theorem3} shows how 
-this property can be recovered from the eigenvectors of \(P\). 
-Then Theorem \ref{theorem4} shows that replacing $P$ with $A$ 
-approximates these properties with error that goes to zero as \(n \to \infty\). 
-This allows us to obtain the subspace detection property 
-in Theorem \ref{theorem5}.
-
+this property can also be recovered from the eigenvectors of \(P\). 
+Then Theorem \ref{theorem4} shows that, by replacing $P$ with $A$, the
+rows of $\hat{V}$ also lie on asymptotically orthogonal subspaces.
+Motivated by Theorem~\ref{theorem4}, Theorem \ref{theorem5} below
+shows that the subspace detection property also holds for the rows of
+$\sqrt{n} \hat{V}$. 
 \begin{algorithm}[t]
@@
 \hypertarget{parameter-estimation}{%
-\subsection{Algorithms for Parameter Estimation}\label{parameter-estimation}}
-
-For any edge probability matrix \(P\) for the PABM such that the rows
-and columns are organized by community, the \(kl\)\textsuperscript{th}
+\subsection{Algorithm for Parameter Estimation}\label{parameter-estimation}}
+For ease of exposition we now assume in this subsection that the
+edge probability matrix \(P\) for the PABM had been arranged so that the rows
+and columns are organized by community so that $\tilde{P} = P$ (see
+Remark~\ref{rem:pabm_view2}). Then the \(k\ell\)\textsuperscript{th}
 block is an outer product of two vectors, i.e.,
@@
 up to multiplicative constant using singular value
-decomposition. More specifically, 
+decomposition. More specifically 
 let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\)
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[tp]
   \label{alg:param_est}
@@
 \label{theorem6}
-Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived from its corresponding $P^{(k \ell)}$, and let $\hat{\lambda}^{(k \ell)}$ be its estimate from $A^{(k \ell)}$ (Alg.
-3).
+Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived
+from its corresponding $P^{(k \ell)}$ and let $\hat{\lambda}^{(k
+  \ell)}$ be its estimate obtained from $A^{(k \ell)}$ using Algorithm~\ref{alg:param_est}.
 Then if $n \rho_n = \omega( \log^{4}{n})$,
 \begin{equation} \label{eq:thm6}
-\max_{k, l \in \{1, ..., K\}} 
+\max_{k, \ell \in \{1, ..., K\}} 
 \|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
@@
 \end{equation}
-with high probability.
-\end{theorem}
-
+with high probability. Here $\|\cdot\|_{\infty}$ denotes the
+$\ell_\infty$ norm of a vector. Let $\hat{\Lambda}$ be the matrix
+$$\hat{\Lambda} = \begin{bmatrix} \hat{\lambda}^{(11)} &
+  \hat{\lambda}^{(12)} & \cdots & \hat{\lambda}^{(1 K)} \\
+  \hat{\lambda}^{(21)} &
+  \hat{\lambda}^{(22)} & \cdots & \hat{\lambda}^{(2 K)} \\
+  \vdots & \vdots & \cdots & \vdots \\
+   \hat{\lambda}^{(K1)} &
+  \hat{\lambda}^{(K2)} & \cdots & \hat{\lambda}^{(K K)} 
+\end{bmatrix}$$
+and let $\hat{P} = \hat{\Lambda} I_{p,q}
+\hat{\Lambda}^{\top}$. Eq.~\eqref{eq:thm6} then implies
+\begin{equation}
+\label{eq:conv_prob}
+\frac{1}{n} \|\rho_n^{-1} \hat{P} - \rho_n^{-1} P\|_{F} = O((n
+\rho_n)^{-1/2}), \quad \max_{ij} |\rho_n^{-1} \hat{P}_{ij} -
+\rho_n^{-1} P_{ij}| = O((n \rho_n)^{-1/2})
+\end{equation}
+with high probability. \end{theorem}
+Eq.~\eqref{eq:thm6} guarantees that $n^{-1/2} \|\rho_n^{-1/2} \hat{\Lambda} -
+\Lambda\|_{F} = O((n \rho_n)^{-1/2})$. Eq.~\eqref{eq:conv_prob} then guarantees that
+the mean square error for $\rho_n^{-1} (\hat{P} - P)$ converges to $0$ almost surely and
+furthermore the entries of $\rho_n^{-1}
+\hat{P}$ converge uniformly to the entries of $\rho_n^{-1} P$; recall that
+$\rho_n^{-1} P_{ij} = \lambda_{iz_j} \lambda_{j}z_i$. We note that
+these results are stronger than existing results in
+\citet{307cbeb9b1be48299388437423d94bf1}; for example Theorem~2 in
+\citet{307cbeb9b1be48299388437423d94bf1} only guarantees $n^{-1/2} \|\rho_n^{-1/2} \hat{\Lambda} -
+\Lambda\|_{F} = o(1)$ as $n \rightarrow \infty$.
 \hypertarget{simulated-examples}{%
@@
 distribution, the popularity vectors \(\{\lambda^{(k \ell)}\}_K\) are drawn
-from two types of joint distributions depending on whether \(k = \ell\),
-the edge probability matrix \(P\) is constructed using the popularity
-vectors, and finally the adjacency matrix \(A\)
-is drawn from \(P\). OSC (Algorithm~\ref{alg:osc}) is then used for community detection, and this
-method is compared against (1) SSC using the spectral embedding $\hat{V}$,
-c.f. Algorithm~\ref{alg:ssc}, (2) SSC using the rows of the
+from two types of joint distributions depending on whether \(k =
+\ell\) or $k \not = \ell$. The edge probability matrix \(P\) is constructed using the popularity
+vectors and finally the adjacency matrix \(A\)
+is drawn \(A \sim \mathrm{Bernoulli}(P)\). OSC (Algorithm~\ref{alg:osc}) is then used for community detection, and this
+method is compared against (1) SSC using the spectral embedding $\hat{V}$
+(Algorithm~\ref{alg:ssc}), (2) SSC using the rows of the
 observed adjacency matrix $A$ as is done in \citet{noroozi2019estimation}
 and (3) modularity maximization (MM) as is done in
-\citet{307cbeb9b1be48299388437423d94bf1}. 
+\citet{307cbeb9b1be48299388437423d94bf1}. We denote the two SSC
+implementations using the rows of $A$ and using the spectral embedding
+of $A$ as SSC-A and SSC-ASE, respectively. The parameters \(\vartheta\)
+that controls the sparsity for SSC-A and SSC-ASE were chosen via a preliminary cross-validation
+experiment. The final clustering step of SSC-A and SSC-ASE was done
+by fitting a Gaussian Mixture Model to the normalized Laplacian
+eigenmap embeddings \citep{belkin03:_laplac} of the affinity matrix \(B\).
 We also estimate the latent popularity vectors $\{\lambda^{(k \ell)}\}$
@@
 
-Two implementations of SSC are shown here. The first method, denoted as
-SSC-A, treats the columns of the adjacency matrix \(A\) as points in
-\(\mathbb{R}^n\), as described in \citet{noroozi2019estimation}.
-The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in Algorithm~\ref{alg:ssc}. The 
-parameter \(\lambda\) that controls the sparsity of the SSC was chosen via a preliminary cross-validation
-experiment. For the final clustering step, a Gaussian Mixture Model was
-fit on the normalized Laplacian eigenmap of the affinity matrix \(B\).
-
 For comparing methods, we define the community detection error as:
@@
 
-For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $P_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
-The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$ divided by the number of vertices:
-
+For parameter estimation, because the popularity parameters
+$\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge
+probabilities $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$ via the
+quantities $\hat{P}_{ij} = \hat{\lambda}_{iz_j} \hat{\lambda}_{jz_i}$. The
+parameter estimation error is then given by the normalized 
+Frobenius norm of $P$ divided by the number of vertices, i.e.,
 $$\mathrm{RMSE}(\hat{P}, P) = \frac{1}{n} \|\hat{P} - P\|_F.$$
 
-We also note that unlike the MLE-based method \citep{307cbeb9b1be48299388437423d94bf1}, the ASE method of Algorithm 3 does not require community labels if all we are interested in is the edge probability matrix, $P$.
-The estimate $\hat{P}$ we obtain by combining the popularity vectors outputted by Algorithm 3 is equivalent to using the full ASE of $A$ to reconstruct the edge probability matrix.
+We also note that unlike the MLE-based method
+\citep{307cbeb9b1be48299388437423d94bf1}, the ASE method in
+Algorithm~\ref{alg:param_est} can be trivially modified so as to not
+require the community labels if we are only interested in 
+estimating $P$. More specifically we first compute the ASE
+$\hat{Z}$ of $A$ (see Remark~\ref{rem:ase}) and then compute $\hat{P}
+= \hat{Z} I_{p,q} \hat{Z}^{\top}$. The resulting estimate $\hat{P}$
+will have the same convergence rate as that given in Eq.(xxx).
+% we obtain by
+% combining the popularity vectors outputted by
+% Algorithm~\ref{alg:param_est} is asymptotically
+% equivalent to using the full ASE of $A$ to reconstruct the edge
+% probability matrix.
 
@@
 \end{itemize}
-Fifty simulations were performed for each combination of $n$ and $K$.
-
-Fig.~\ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
-going to 0 for large \(n\). SSC on both the embedding and on the
-adjacency matrix produces similar trends for \(K > 2\). Weaker
-performance of SSC for \(K = 2\) can be attributed to the final spectral
-clustering step of the affinity matrix. 
-While the subspace detection property is guaranteed for large \(n\), in our
-simulations, setting the sparsity parameter to the required value resulted
-in more than \(K\) disconnected subgraphs. 
+Fifty simulations were performed for each combination of $n$ and
+$K$. The results for community recovery and parameter estimations are
+presented in Fig.~\ref{fig:clust_err_ct_sim} and
+Fig.~\ref{fig:p_block_est}, respectively.
+
+Fig.~\ref{fig:clust_err_ct_sim} shows that OSC recovers the community
+perfectly as $n$ increases, i.e., the number of mislabeled vertices
+goes to $0$. The performance of SSC-ASE is comparable to OSC for $K
+\geq 3$ but is noticeably worse when $K = 2$. Similarly, SSC on both the embedding and on the
+adjacency matrix produces similar trends for \(K > 2\). The difference
+in performance between SSC-A and SSC-ASE for \(K = 2\) can be attributed to the final spectral
+clustering step of the affinity matrix. While the subspace detection
+property is guaranteed for large $n$, in our simulations, setting the
+sparsity parameter $\vartheta$ to the required value usually resulted
+in more than $K$ disconnected subgraphs in the affinity matrix
+$\hat{B}$.  
+% More specifically the sparsity parameter to the required value resulted
+% in more than \(K\) disconnected subgraphs. 
 We instead chose a smaller sparsity parameter, 
 necessitating a final clustering step. 
-A GMM was fit to the normalized Laplacian eigenmap of $B$,
+A GMM was fit to the normalized Laplacian eigenmap of $\hat{B}$,
 but visual inspection suggests that the communities are not
@@
 
-\begin{figure}[H]
-
+\begin{figure}[tp]
 {\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim-1}
-
 }
@@
 
-Given ground truth community labels, Algorithm~\ref{alg:param_est} and the MLE-based
-plug-in estimators perform similarly, with root mean square
-error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:p_block_est}).
-
-\begin{figure}[H]
+Given ground truth community labels, Fig.~\ref{fig:p_block_est} shows
+that Algorithm~\ref{alg:param_est} and the MLE-based
+plug-in estimators perform comparably, with root mean square
+error decaying at rate approximately \(n^{-1/2}\) as $n$ increases.
+\begin{figure}[tp]
 {\centering \includegraphics{summary_files/figure-latex/lambda_est_p-1}}
-\caption{Median and IQR RMSE for edge probability matrices reconstructed from the outputs of Algorithm 3 (red) compared against outputs of an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.}
+\caption{Median and IQR RMSE for edge probability matrices
+  reconstructed from the outputs of Algorithm~\ref{alg:param_est}
+  (red) compared against outputs of an MLE-based method (blue) proposed in \citet{307cbeb9b1be48299388437423d94bf1}.
+  Simulations were repeated 50 times for each sample size. Communities were drawn to be approximately balanced.}
 \label{fig:p_block_est}
@@
 \end{itemize}
-
-\(50\) simulations were performed for each \((n, K)\) pair.
-
-We again see community detection error trending to 0 for OSC, as well as
-for SSC when \(K > 2\) (Fig.~\ref{fig:clust_err_ct_sim_imba}). 
-The parameter estimation error decays at rate $n^{-1/2}$ as in the balanced communities case
-(Fig. \ref{fig:lambda_est_p_imba}).
+Fifty simulations were performed for each combination of $n$ and
+$K$. The results for community recovery and parameter estimations are
+presented in Fig.~\ref{fig:clust_err_ct_sim_imba} and Fig.~\ref{fig:lambda_est_p_imba}, respectively.
+
+From Fig.~\ref{fig:clust_err_ct_sim_imba} we once again see that the
+number of mis-labeled vertices trending to $0$ for OSC. The
+performance of SSC-ASE is comparable to that of OSC for $K > 2$ but
+is worse when $K = 2$. Fig.~\ref{fig:lambda_est_p_imba} indicates that
+the parameter estimation error also decays at rate $n^{-1/2}$ similar
+to that in the balanced communities setting.
 
 \begin{figure}[H]
-
 {\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim_imba-1}
 }
-
 \caption{Median and IQR of community detection error. Communities are imbalanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim_imba}
@@
 \begin{figure}[H]
-
 {\centering \includegraphics{summary_files/figure-latex/lambda_est_p_imba-1}
-
 }
-
-\caption{Median and IQR RMSE of edge probabilities derived from the outputs of Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be imbalanced.}
+\caption{Median and IQR RMSE of edge probabilities derived from the
+  outputs of Algorithm \ref{alg:param_est} (red) compared against an
+  MLE-based method (blue) described in
+  \cite{307cbeb9b1be48299388437423d94bf1}.
+  Simulations were repeated 50 times for each sample size. Communities were drawn to be imbalanced.}
 \label{fig:lambda_est_p_imba}
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2021-09-08 summary_minh_edited.tex

@@
 (PABM) and the Generalized Random Dot Product Graph (GRDPG),
-by demonstrating that a PABM is a special case of a GRDPG in which communities correspond to 
+by demonstrating that the PABM is a special case of 
+the GRDPG in which communities correspond to 
 mutually orthogonal subspaces of latent vectors. 
@@
 % Then $G$ is drawn from a Popularity Adjusted Block Model if
-  In a PABM each vertex $i$ has $K$ popularity parameters $\lambda_{i1}, \dots, \lambda_{iK}$, that describe its affinity
+In a PABM, each vertex $i$ has $K$ popularity parameters 
+$\lambda_{i1}, \dots, \lambda_{iK}$, that describe its affinity
 toward each of the $K$ communities. % , i.e., vertex $i$ has popularity parameters
@@
 \mathbb{R}^{n_k}$; the elements of $\lambda^{(k \ell)}$ are the affinity parameters toward the $\ell$th community of
-all vertices in the $k$th community. Define $\lambda^{(\ell k)}$
+all vertices in the $k^{th}$ community. Define $\lambda^{(\ell k)}$
 analogously. Then each block $\tilde{P}^{(k \ell)}$ can be written as the outer product of two vectors:
 \begin{equation} \label{eq:pabm}
-  \tilde{P}^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}.
+  \tilde{P}^{(k \ell)} = \rho_n \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}.
 \end{equation}
@@
 $y_i$ belongs to the union of $K$ subspaces and the $z_i$ are noise
-vectors. Let $\lambda \geq 0$ be given and
+vectors. Let $\vartheta \geq 0$ be given and
 let $C$ and $B$ be constructed from the
 solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
-this given choice of $\lambda$. Then $X$
+this given choice of $\vartheta$. Then $X$
 is said to satisfy the subspace detection property 
@@
 guarantee that SSC perfectly recovers the underlying subspaces.
-More specifically
-if $X$ satisfies the
-subspace detection property then $B$ describes a graph with {\em at least}
-$K$ disconnected subgraphs, with the ideal case being that there are exactly $K$ subgraphs which map onto each
-subspace. Nevertheless it is also possible that the $K$ subspaces are represented by
+More specifically, if $X$ satisfies the 
+subspace detection property, 
+then $B$ describes a graph with {\em at least} $K$ disconnected subgraphs, 
+with the ideal case being that there are exactly $K$ subgraphs 
+which map onto each subspace. 
+Nevertheless it is also possible that the $K$ subspaces are represented by
 $K' > K$
@@
 for further discussions. Therefore in practice $B$ is usually treated as an
-affinity matrix and, as we allude to earlier, the rows of $B$ are partitioned using some
-clustering algorithm to obtain the final clustering. % Finally we note that the subspace detection property is contingent
-% on choosing a sufficiently large sparsity parameter $\vartheta$ for
-% the LASSO problems in Eq.~\eqref{eq:ssc}.
+affinity matrix and, as we allude to earlier, 
+the rows of $B$ are partitioned using some
+clustering algorithm to obtain the final clustering. 
 \end{remark}
@@
 However since \(Q\) is not necessarily an
-orthogonal matrix and hence, if $y_i$ denote the $i$th row of $Y$ then $\langle U^{\top} x_i, U^{\top} x_j \rangle
+orthogonal matrix and hence, if $y_i$ denote the $i^{th}$ row of $Y$, 
+then $\langle U^{\top} x_i, U^{\top} x_j \rangle
 \neq \langle y_i, y_j \rangle$.
@@
 \end{proof}
-Under the condition that $\|\Lambda\|_F > 0$, where 
-$\Lambda$ is the $n \times K$ matrix of popularity parameters, 
-Theorem \ref{theorem3} provides perfect community detection from
-\(P\). More specifically let \(|B|\) be the affinity matrix for graph \(G'\), 
+
+Under the condition that $\|\Lambda\|_F > 0$, 
+Theorem \ref{theorem3} provides perfect community detection from \(P\).
+\footnote{This is excluding the special case in which vectors in each 
+subspace lie on mutually orthogonal sub-subspaces.}
+More specifically let \(|B|\) be the affinity matrix for graph \(G'\), 
 where $|\cdot|$ is applied entry-wise. Then
@@
  $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$ 
-for pair $(i, j)$ in different communities. 
+for each pair $(i, j)$ in different communities.
+
 \begin{algorithm}[tp]
@@
 with high probability. In particular $\hat{B}_{ij} -
-B_{ij} \overset{\mathrm{a.s}}{\rightarrow} 0$ where the convergence is
+B_{ij} \overset{\mathrm{a.s.}}{\rightarrow} 0$ where the convergence is
 uniform over all $i,j$. Hence for all pairs $(i,j)$ in different
@@
 eigenvalues in modulus. Then for any $\epsilon > 0$ 
-there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
+there exists a choice of $\vartheta > 0$ and $N \in \mathbb{N}$ such
 that for all $n \geq N$, $\sqrt{n} \hat{V}$ obeys the subspace detection property with
@@
 = \hat{Z} I_{p,q} \hat{Z}^{\top}$. The resulting estimate $\hat{P}$
-will have the same convergence rate as that given in Eq.(xxx).
+will have the same convergence rate as that given in Eq.~(\ref{eq:conv_prob}).
 % we obtain by

### 2021-09-08 summary_minh_edited.tex
changes based on discussion from earlier

@@
 
-\author{John Koo\\Department of Statistics, Indiana University\\
-Minh Tang\\Department of Statistics, North Carolina State University\\
-Michael Trosset\\Department of Statistics, Indiana University}
+\author{John Koo\\Department of Statistics, Indiana University\\ \\
+Minh Tang\\Department of Statistics, North Carolina State University\\ \\
+Michael W. Trosset\\Department of Statistics, Indiana University\\ \\
+September 8, 2021}
 
@@
 they are denoted as $z_1, ..., z_n$. 
+$\lambda_{ik}$ denotes the popularity parameter of vertex $i$ to community $k$. 
+$\Lambda$ is the $n \times K$ matrix of popularity parameters. 
 Finally, we denote 
@@
 \label{theorem3}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are
-from different communities.
+Let $P = V D V^\top$ be the spectral decomposition 
+of the edge probability matrix. 
+Let $B = n V V^\top$. 
+Assume $\lambda_{i z_i} > 0$ for each $i \in [n]$, 
+i.e., each vertex's popularity parameter to its own community is nonzero. 
+Then if vertices $i$ and $j$ are from different communities, $B_{ij} = 0$.
 \end{theorem}
@@
 
-Under the condition that $\|\Lambda\|_F > 0$, 
-Theorem \ref{theorem3} provides perfect community detection from \(P\).
-\footnote{This is excluding the special case in which vectors in each 
-subspace lie on mutually orthogonal sub-subspaces.}
-More specifically let \(|B|\) be the affinity matrix for graph \(G'\), 
+Theorem \ref{theorem3} provides perfect community detection from \(P\)\footnote{This excludes the pathological case in which $P_{ij} = 0$ for some $z_i = z_j$.}.
+More specifically, let \(|B|\) be the affinity matrix for graph \(G'\), 
 where $|\cdot|$ is applied entry-wise. Then
@@
 that are larger than $\epsilon$ converges to zero with probability
-converging to one as $n$ increases. Thus, by using $\hat{B}$, we can 
+converging to one as $n$ increases. 
+Furthermore, if $P_{ij} > 0$ for each pair $(i, j)$ in the same community, 
+we can always find an $\epsilon > 0$ such that $\hat{B}_{ij} > \epsilon$ 
+with probability converging to one as $n$ increases. 
+Thus, by using $\hat{B}$, we can 
 perfectly recover all the latent community assignments $z_1, z_2,
@@
 $\ell_\infty$ norm of a vector. Let $\hat{\Lambda}$ be the matrix
+
 $$\hat{\Lambda} = \begin{bmatrix} \hat{\lambda}^{(11)} &
@@
 \end{bmatrix}$$
-and let $\hat{P} = \hat{\Lambda} I_{p,q}
-\hat{\Lambda}^{\top}$. Eq.~\eqref{eq:thm6} then implies
+
+and let $\hat{P} = \hat{X} U I_{p,q} U^\top \hat{X}^{\top}$ 
+where $\hat{X}$ is defined from $\hat{\Lambda}$ and $U$ is defined from $K$ as in Theorem~\ref{theorem2}. 
+Eq.~\eqref{eq:thm6} then implies
 \begin{equation}
@@
 \hat{P}$ converge uniformly to the entries of $\rho_n^{-1} P$; recall that
-$\rho_n^{-1} P_{ij} = \lambda_{iz_j} \lambda_{j}z_i$. We note that
+$\rho_n^{-1} P_{ij} = \lambda_{iz_j} \lambda_{j z_i}$. We note that
 these results are stronger than existing results in
@@
 \Lambda\|_{F} = o(1)$ as $n \rightarrow \infty$.
+
 \hypertarget{simulated-examples}{%
@@
 \citet[Lemma~5]{rubindelanchy2017statistical} for a proof.
+
 \begin{lemma}
@@
 Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell \ell)}\}_{K})$ be a $K$-blocks
-PABM graph on $n$ vertices and let \(V_n\) and \(\hat{V}_n\)
-be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
+PABM graph on $n$ vertices and let \(V\) and \(\hat{V}\)
+be the $n \times K^2$ matrices whose columns are the eigenvectors of 
+\(P\) and \(A\) corresponding to the
 $K^2$ largest eigenvalues in modulus, respectively.
-Let \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
-row of \(V_n\) and \(\hat{V}_n\), respectively. 
+Let \(v_i^\top\) and \(\hat{v}_i^\top\) denote the $i$th 
+row of \(V\) and \(\hat{V}\), respectively. 
 Then there exists a constant $c > 1$ and an orthogonal matrix $W$ such
 that with high probability,
-$$\max_{i} \|W \hat{v}_n^{(i)} - v_n^{(i)}\|  = O\Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
+$$\max_{i} \|W \hat{v}_i - v_i\|  = O\Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
 In particular we can take $c = 1 + \epsilon$ for any $\epsilon > 0$. 
 \end{lemma}
-
-%
-%then the \(ij\)\textsuperscript{th} element of
-%\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
 
@@
 under our assumption that the latent vectors $\lambda^{(k \ell)}$
-  are all homogeneous, we have $\max_{i} \|v_n^{(i)}\| =
+are all homogeneous, we have $\max_{i} \|v_i\| =
 O(n^{-1/2})$.
-%   Let \(V_n\) and \(\hat{V}_n\)
-% be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
-% \(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
-% eigenvalues, respectively. From Lemma~5 in
-% \citet{rubindelanchy2017statistical} we have, for some
-% \(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
-% $$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \big).$$
-% We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). 
-% Let
-% \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
-% row of \(V_n\) and \(\hat{V}_n\), respectively. 
-Next recall Theorem~\ref{theorem3}; in particular $B_{ij} = n(v_n^{(i)})^{\top}
-v_n^{(j)}$. % Now suppose that vertices $i$ and $j$ belongs to different
-% communities. Then $B_{ij} = 0$ and hence
-% Then for \(i\) and
-% \(j\) belonging to different
-% communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
+
+Next recall Theorem~\ref{theorem3}; in particular $B_{ij} = nv_i^{\top}
+v_j$. 
 We therefore have
 \[\begin{split}
- \max_{ij} |\hat{B}_{ij} - B_{ij}| &= \max_{ij} n |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& \leq n \max_{ij} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& \leq n \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|
-+ \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
-% & \leq 
-% \hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-% & = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
-% 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
-% & = ||2 (\hat{V}_n W - V_n) V_n^\top + 
-% (\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
-% & \leq 2 
-% &\leq  n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|\hat{v}_n^{(j)}\| +
-% n \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr)
-% \\
-& \leq n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 +  \|W
-\hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|v_n^{(j)}\| +  \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr) 
+ \max_{ij} |\hat{B}_{ij} - B_{ij}| &= \max_{ij} n |\hat{v}_i^\top \hat{v}_j -
+v_i^\top v_j| \\
+& \leq n \max_{ij} |\hat{v}_i^\top W W^\top \hat{v}_j -
+v_i^\top v_j| \\
+& \leq n \max_{i,j} \Bigl(\|W^{\top} \hat{v}_i - v_i\| \times \|\hat{v}_j\|
++ \|W^{\top} \hat{v}_j - v_j\| \times \|v_i\|\Bigr) \\
+& \leq n \Bigl(\max_{ij}  \|W \hat{v}_i  - v_i \|^2 +  \|W
+\hat{v}_i  - v_i \| \times \|v_j\| +  \|W \hat{v}_j  - v_j \| \times \|v_i\|\Bigr) 
 \\ &
-\leq n \max_{i} \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 + 2n \max_{i}
-\|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \max_{j} \|v_n^{(j)}\|
+\leq n \max_{i} \|W \hat{v}_i  - v_i \|^2 + 2n \max_{i}
+\|W \hat{v}_i  - v_i \| \times \max_{j} \|v_j\|
 \\
@@
 \end{proof}
-% Scaling all terms in the above inequality by \(n\), we obtain
-% $$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
-% as desired.
 
@@
 paper and we will drop the explicit dependency on $n$ from our
-eigenvectors $\hat{V}_n$ of $A$ and $V_n$ of $P$.
+eigenvectors $\hat{V}$ of $A$ and $V$ of $P$.
  % We first recall the

### 2021-09-08 summary_minh_edited.tex
updated draft

@@
 subspaces and hence $v_i^{\top} v_j = 0$ whenever $z_i \not = z_j$.
-% The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
-% such that points within each community lie on \(K\)-dimensional
-% orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
-% removes the orthogonality property but retains the property that each
-% community is represented by a \(K\)-dimensional subspace. Therefore, the
-% ASE of \(P\) still consists of subspaces 
-% that correspond to communities of the PABM, even if they are not orthogonal, 
-% suggesting the use of SSC. Before exploring SSC, we will first consider
-% a different approach.
+
 \begin{theorem}
@@
 i.e., each vertex's popularity parameter to its own community is nonzero. 
-Then if vertices $i$ and $j$ are from different communities, $B_{ij} = 0$.
+Then $B_{ij} = 0$ if and only if 
+vertices $i$ and $j$ are in different communities.
 \end{theorem}
+
 \begin{proof}
-% \begin{lemma}
-% \label{lemma1}
-% Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-% matrix for a PABM.
 We first show that $V V^\top =
@@
   Q^{\top} I_{p,q} Q &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} \tilde{Y} I_{p,q}
-I_{p,q} I_{p,q} \tilde{Y}^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
+I_{p,q} I_{p,q} \tilde{Y}^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ 
+&= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
 Y^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} =  I_{p,q}
@@
 
-% In order to proceed with the proof that that OSC is consistent 
-% (Theorem \ref{theorem4}), 
-% we introduce the sparsity factor $\rho_n \in (0, 1]$, 
-% which can be characterized as the average expected degree of the vertices 
-% divided by the number of vertices. 
-% The treatment of the sparsity factor is the same as in 
-% \citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
-% and \citet{rubindelanchy2017statistical} (GRDPG) 
-% in which the popularity vectors are scaled as 
-% $\rho_n^{1/2} \lambda^{(k \ell)}$. 
-% This results in the edge probabilities being scaled as 
-% $P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
-% Because $\rho_n$ is unidentifiable, 
-% we use the convention that the popularity parameters are drawn such that
-% $\|\Lambda\|_F = O(\sqrt{n})$ and 
-% require that the $L^2$ norm of each row is bounded away from 0 and 1.
-
 \begin{theorem}
 \label{theorem4}
+Assume the setting of Theorem~\ref{theorem3}. 
 Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix
-obtained from OSC as described in Algorithm~\ref{alg:osc}. Then for $n \rho_n = \omega(\log^{4}{n})$, we have
+obtained from OSC as described in Algorithm~\ref{alg:osc}. 
+Then for $n \rho_n = \omega(\log^{4}{n})$, we have
+
 \begin{equation} \label{eq:thm4a}
@@
 \end{equation}
-with high probability. In particular $\hat{B}_{ij} -
-B_{ij} \overset{\mathrm{a.s.}}{\rightarrow} 0$ where the convergence is
+
+with high probability. In particular 
+$\hat{B}_{ij} -B_{ij} \overset{\mathrm{a.s.}}{\rightarrow} 0$ 
+where the convergence is
 uniform over all $i,j$. Hence for all pairs $(i,j)$ in different
 communities we have 
-% \begin{equation} \label{eq:thm4}
-% \max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
-% O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
-% \end{equation}
-% with high probability. 
-% Thus for all pairs $(i, j)$ in different communities,
-% we have uniform convergence 
-$\hat{B}_{ij} \overset{\mathrm{a.s}}{\rightarrow} 0$. 
+$\hat{B}_{ij} \overset{\mathrm{a.s.}}{\rightarrow} 0$, 
+while for all pairs $(i, j)$ in the same community, 
+$\liminf_{n \to \infty} |\hat{B}_{ij}| > 0$ almost surely. 
 \end{theorem}
+
 Theorem~\ref{theorem4} guarantees that for any $\epsilon > 0$, 

### 2021-09-09 summary_minh_edited.tex
Remove a few sentences

@@
 identification or estimation of \(XU\) is possibly difficult 
-due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable}).
+due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable})
 when we are given only $P$. % of the latent
@@
 
-Theorem \ref{theorem3} provides perfect community detection from \(P\)\footnote{This excludes the pathological case in which $P_{ij} = 0$ for some $z_i = z_j$.}.
+Theorem \ref{theorem3} provides perfect community detection from \(P\).% \footnote{This excludes the pathological case in which $P_{ij} = 0$ for some $z_i = z_j$.}.
 More specifically, let \(|B|\) be the affinity matrix for graph \(G'\), 
@@
 Then for $n \rho_n = \omega(\log^{4}{n})$, we have
-
 \begin{equation} \label{eq:thm4a}
@@
 \end{equation}
-
 with high probability. In particular 
@@
 \end{theorem}
-
 Theorem~\ref{theorem4} guarantees that for any $\epsilon > 0$, 
@@
 converging to one as $n$ increases. 
-Furthermore, if $P_{ij} > 0$ for each pair $(i, j)$ in the same community, 
-we can always find an $\epsilon > 0$ such that $\hat{B}_{ij} > \epsilon$ 
+Furthermore we can always find an $\epsilon > 0$ such that if $i$ and
+$j$ are in the same community then $\hat{B}_{ij} > \epsilon$ 
 with probability converging to one as $n$ increases. 

### 2021-09-09 summary_minh_edited.tex
updated based on additional cfeedback

--- 2021-09-08 summary_minh_edited.tex
@@
 
-Theorem \ref{theorem3} provides perfect community detection from \(P\)\footnote{This excludes the pathological case in which $P_{ij} = 0$ for some $z_i = z_j$.}.
+Theorem \ref{theorem3} provides perfect community detection from \(P\).
 More specifically, let \(|B|\) be the affinity matrix for graph \(G'\), 
@@
 converging to one as $n$ increases. 
-Furthermore, if $P_{ij} > 0$ for each pair $(i, j)$ in the same community, 
-we can always find an $\epsilon > 0$ such that $\hat{B}_{ij} > \epsilon$ 
+We can always find an $\epsilon > 0$ such that $\hat{B}_{ij} > \epsilon$ 
 with probability converging to one as $n$ increases. 

### 2021-09-09 summary_minh_edited.tex
cleaned things up a bit more

@@
 the same community correspond to latent position vectors that lie on a
-straight line. %(\cite{lyzinski2014, rubindelanchy2017statistical}).
+straight line.
 
@@
   \label{rem:pabm_view2}
-% Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-% graph $G = (V, E)$ with adjacency matrix $A$ 
-% such that $A \sim \mathrm{Bernoulli}(P)$.
-% Let each vertex have a community label between $1$ and $K$.
-% Then $G$ is drawn from a Popularity Adjusted Block Model if
 In a PABM, each vertex $i$ has $K$ popularity parameters 
 $\lambda_{i1}, \dots, \lambda_{iK}$, that describe its affinity
-toward each of the $K$ communities. % , i.e., vertex $i$ has popularity parameters
-% $\lambda_{i1}, ..., \lambda_{iK}$, and  $P_{ij} = \lambda_{i z_j}
-% \lambda_{j z_i}$ for all $i \leq j$. 
+toward each of the $K$ communities. 
 Another view of a PABM is as follows.
@@
 \end{equation}
-% for a set of $K^2$ popularity vectors 
-% $\{\lambda^{(st)}\}_{1 \leq s \leq K, 1 \leq t \leq K}$ where each
-% $\lambda^{(st)}$ is a column vector of dimension $n_s$.
-% . Then a graph $G$
-% is drawn from a PABM with parameters
-% $\{\lambda^{(st)}\}$ if its adjacency matrix $A$
-% satisfies $A \sim \mathrm{BernoulliGraph}(P)$.
+
 We will henceforth use the notation \(A \sim \PABM(\{\lambda^{(k
@@
 the form
+
 $$P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$$
-% Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$ for some symmetric $P \in
-% [0,1]^{n \times n}$. 
-% If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
-% \begin{equation} \label{eq:grdpg}
-%   P = X I_{p,q} X^\top
-% \end{equation}
-% for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-% $G$ is drawn from the Generalized Random Dot Product Graph 
-% with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
+
 \end{definition}
@@
 parameter $\rho_n$ and signature
-\((p, q)\). % If instead of fixed latent positions, 
-% they are drawn $X_1, ..., X_n \iid F$, 
-% we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
+\((p, q)\).
+
 \begin{remark}
@@
 identification or estimation of \(XU\) is possibly difficult 
-due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable}).
-when we are given only $P$. % of the latent
-% positions of a GRDPG.  
-% due to the
-% non-identifiability of \(XU\) when we are only given $P$, 
-% direct
-% identification or estimation of \(XU\) is possibly difficult 
-% if we are only given $P$. 
-% (\cite{rubindelanchy2017statistical}).
+due to the non-identifiability of $XU$ (see Remark~\ref{rem:non_identifiable})
+when we are given only $P$. 
 More specifically, suppose we find a matrix $Y \in
@@
 we  \( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
-$Q \in \mathbb{O}(p,q)$. %  as there is no
-% general matrix decomposition that yield 
-% The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
-% such that points within each community lie on \(K\)-dimensional
-% orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
-% removes the orthogonality property but retains the property that each
-% community is represented by a \(K\)-dimensional subspace. 
-% However, because $\tilde{\Pi}$ is unknown (indeed, if we know $\tilde{\Pi}$
-% then we also know the community labels), 
-% Therefore, even if we
-% start with the true but unknown edge probability matrix $P$ we
-% will not be able to recover the latent positions \(XU\).
+$Q \in \mathbb{O}(p,q)$. 
 However since \(Q\) is not necessarily an
@@
 will have the same convergence rate as that given in Eq.~(\ref{eq:conv_prob}).
-% we obtain by
-% combining the popularity vectors outputted by
-% Algorithm~\ref{alg:param_est} is asymptotically
-% equivalent to using the full ASE of $A$ to reconstruct the edge
-% probability matrix.
 
@@
 
-% (We can also note that
-% the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
-% \(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
-% \(v^{(kl)} = u^{(lk)}\).)\\
 Next consider the Hermitian dilation
@@
 \[\hat{M}^{(k \ell)} = \begin{bmatrix} 0 & A^{(k \ell)} \\ A^{(\ell k)} & 0 \end{bmatrix}\]
-% Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
-
-% \[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
-
-% and \(E\) is the \(n_k \times n_\ell\) matrix of independent noise (to
-% generate the Bernoulli entries in \(A^{(kl)}\).\\
+
 We can then view \(\hat{M}^{(k \ell)}\) as an adjacency matrix drawn from
@@
   \ell)}$ and $\hat{\Lambda}^{(k \ell)}$, respectively.
-% approximates the latent positions of \(M^{(kl)}\) up to indefinite
-% orthogonal transformation by the rate given in Theorem 5 of
-% \citet{rubindelanchy2017statistical}.\\
+
 Furthermore, by looking at the proof of Theorem~5 in
@@
 forming a block and the negative eigenvalues of $M^{(k \ell)}$ forming
-the remaining block. % \(U^\top \hat{U}\). 
+the remaining block. 
 Because $M^{(k \ell)}$ has one positive eigenvalue and one negative
@@
 = \Bigl[\begin{smallmatrix} 1 & 0 \\ 0 & - 1\end{smallmatrix}\Bigr]$
-% \((1, 1)\), and \(U^\top \hat{U}\) is block
-% diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). 
 Using this form for $W^{*}$, we obtain

### 2021-11-26 s675/sbm-em-demo.R
notes for s675

@@ -1,0 +1,114 @@
+library(mclust)
+import::from('~/dev/pabm-grdpg/functions.R', draw.graph)
+
+em.sbm <- function(A, eps = 1e-9, maxit = 100) {
+  # EM for SBM with 2 communities
+  # using mean field approximation
+  
+  # Pi is an n by 2 matrix of estimated community probabilities
+  Pi <- initial.guess(A)
+  # n <- nrow(A)
+  # Pi <- matrix(runif(n * 2), nrow = n, ncol = 2)
+  # Pi <- sweep(Pi, 1, rowSums(Pi), `/`)
+  # Theta is a K by K matrix of community edge probabilities
+  Theta <- m.step(A, Pi)
+  
+  d.Pi <- eps + 1
+  niter <- 0
+  
+  while (d.Pi > eps) {
+    Pi.old <- Pi
+    
+    Pi <- e.step(A, Pi, Theta)
+    Theta <- m.step(A, Pi)
+    
+    niter <- niter + 1
+    if (niter >= maxit) {
+      warning('failed to converge')
+      break
+    }
+    
+    d.Pi <- norm(Pi - Pi.old, 'F') ** 2 / (n * 2)
+    # print(d.Pi)
+    zhat <- apply(Pi, 1, which.max)
+    print(compute.error(z, zhat))
+  }
+  
+  zhat <- apply(Pi, 1, which.max)
+  
+  return(list(Pi = Pi, Theta = Theta, z = zhat))
+}
+
+initial.guess <- function(A) {
+  # GMM on the laplacian eigenmap
+  
+  n <- nrow(A)
+  
+  L <- diag(colSums(A)) - A
+  L.eigen <- eigen(L, symmetric = TRUE)
+  X <- L.eigen$vectors[, n - 1]
+  Pi <- mclust::Mclust(X, 2)$z
+  
+  return(Pi)
+}
+
+e.step <- function(A, Pi, Theta) {
+  n <- nrow(A)
+  Pi.new <- matrix(0, nrow = n, ncol = 2)
+  
+  for (i in seq(n)) {
+    for (k in seq(2)) {
+      for (j in seq(n)[-i]) {
+        for (l in seq(2)) {
+          Pi.new[i, k] <- Pi.new[i, k] + 
+            Pi[j, l] * 
+            (A[i, j] * log(Theta[k, l]) + (1 - A[i, j]) * log(1 - Theta[k, l]))
+        }
+      }
[…]

### 2021-11-26 s675/sbm-em-demo.R
notes for s675

@@ -6,12 +6,14 @@ em.sbm <- function(A, eps = 1e-9, maxit = 100) {
   # using mean field approximation
   
   # Pi is an n by 2 matrix of estimated community probabilities
-  Pi <- initial.guess(A)
-  # n <- nrow(A)
-  # Pi <- matrix(runif(n * 2), nrow = n, ncol = 2)
-  # Pi <- sweep(Pi, 1, rowSums(Pi), `/`)
+  # Pi <- initial.guess(A)
+  n <- nrow(A)
+  Pi <- matrix(runif(n * 2), nrow = n, ncol = 2)
+  Pi <- sweep(Pi, 1, rowSums(Pi), `/`)
+
   # Theta is a K by K matrix of community edge probabilities
-  Theta <- m.step(A, Pi)
+  # start with some initial guess
+  Theta <- matrix(c(.9, .1, .1, .9), 2, 2)
   
   d.Pi <- eps + 1
   niter <- 0
@@ -28,7 +30,7 @@ em.sbm <- function(A, eps = 1e-9, maxit = 100) {
       break
     }
     
-    d.Pi <- norm(Pi - Pi.old, 'F') ** 2 / (n * 2)
+    d.Pi <- norm(Pi - Pi.old, 'F') ** 2 / prod(dim(Pi))
     # print(d.Pi)
     zhat <- apply(Pi, 1, which.max)
     print(compute.error(z, zhat))
@@ -99,11 +101,11 @@ compute.error <- function(z, zhat) {
   return(error)
 }
 
-n1 <- 50
-n2 <- 50
+n1 <- 20
+n2 <- 20
 n <- n1 + n2
 z <- c(rep(1, n1), rep(2, n2))
-p.within <- .3
+p.within <- .4
 p.between <- .1
 P <- matrix(p.between, nrow = n, ncol = n)
 P[seq(n1), seq(n1)] <- p.within

### 2021-11-26 s675/rdpg.Rmd
notes on rdpg and grdpg

---
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: 
        - page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

set.seed(314159)
source('~/dev/pabm-grdpg/functions.R')
```

## {.plain}

\center

\LARGE

\textcolor{darkred}{Random Dot Product Graphs}

\normalsize

STAT-S 675

Fall 2021

# Recall From Last Time ...

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newcommand{\argmin}{\mathrm{argmin}}

* Let $G = (V, E)$ be an undirected and hollow graph with $|V| = n$ 
and adjacency matrix $A$
  * $A \in \mathbb{R}^{n \times n}$ is symmetric with zero diagonals
* Suppose $G \sim F(\theta)$
  * What kind of $F(\theta)$ make sense here?
  * Given $F$ and observed $G$, how can we estimate $\theta$?

# Recall From Last Time ...

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=55%}

$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

For estimation, we need to impose some structure on $P$.

::::

:::: {.column width=45%}

**Example 1**: If every entry $P_{ij} = \theta \in (0, 1)$, 
then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / 5
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

# Random Dot Product Graph

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $P = X X^\top \in \mathbb{R}^{n \times n}$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$
* $A \sim \BG(P)$

* Estimation Objectives: 
  1. Estimate $X$ from $A$ (assume $d$ is fixed)
  2. If $x_1, ..., x_n \iid F(\theta)$, estimate $\theta$ from $A$
  (assuming certain $F$)

* Non-identifiability: For any $W \in \mathbb{O}(d)$, $X W$ is a latent configuration that produces the same $P = X X^\top = X W W^\top X^\top$

# Random Dot Product Graph

```{r}
n1 <- 50
n2 <- 50
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)
data.matrix <- cbind(c(x1, x2), c(y1, y2))
```

::: columns

:::: {.column width=50%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent space')
```

::::

:::: {.column width=50%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring', legend = FALSE)
title('RDPG')
```

::::

:::

# Maximum Likelihood Estimation

$$L(X | A) = \prod_{i < j} (x_i^\top x_j)^{A_{ij}} (1 - x_i^\top x_j)^{1 - A_{ij}}$$

* Intractable
* Not unique

# Adjacency Spectral Embedding

$$\hat{X} = \argmin_{X \in \mathbb{R}^{n \times d}} \|A - X X^\top \|_F^2$$

* Same solution as PCA
* $A = V \Lambda V^\top$
* $\hat{X} = V_d \Lambda_d^{1/2}$
  * $V_d \in \mathbb{R}^{n \times d}$
  * $\Lambda_d \in \mathbb{R}^{d \times d}$
* Not unique
  * If $\hat{X}$ is a solution, then so is $\hat{X} W$
  * Multiplying by $W$ preserves interpoint inner products and distances
* $\max_i \|\hat{x}_i - W x_i\| \stackrel{a.s.}{\to} 0$ (Athreya et al., 2018)

# Adjacency Spectral Embedding

::: columns

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent space')
```

::::

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring', legend = FALSE)
title('RDPG')
```

::::

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, c(1, 2)] %*% diag(sqrt(abs(A.eigen$values[c(1, 2)])))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

::::

:::

# Random Dot Product Graph

* For what types of graphs can we justify a latent structure?
* Any $A \sim \BG(P)$ is a RDPG if $P$ is positive semidefinite
* Whether we think of a graph as a RDPG depends on whether the latent structure is useful
* What about $P$ that are not positive semidefinite?

# Generalized Random Dot Product Graph

Generlized Random Dot Product Graph $A \sim \GRDPG(X)$  
(Rubin-Delanchy et al., 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p,q} x_j \in [0, 1]$
and $I_{p,q} = \blockdiag(I_p, -I_q)$
* $P = X I_{p,q} X^\top \in \mathbb{R}^{n \times n}$, 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times (p+q)}$,  
$A \sim \BG(P)$

* Non-identifiability: For any $Q \in \mathbb{O}(p,q)$, $X Q$ is a latent configuration that produces the same 
$P = X I_{p,q} X^\top = X Q I_{p,q} Q^\top X^\top$
  * Multiplication by $Q$ **does not** preserve interpoint inner products or distances

* Any $A \sim \BG(P)$ is a GRDPG

# Adjacency Spectral Embedding for GRDPG

$$\hat{X} = \argmin_{X \in \mathbb{R}^{n \times (p+q)}} \|A - X I_{p,q} X^\top \|_F^2$$

* $A = V \Lambda V^\top$
* $\hat{X} = V_{p,q} |\Lambda_{p,q}|^{1/2}$
  * $V_{p,q} = \begin{bmatrix} V_{1:p} & | & V_{n-q+1:n} \end{bmatrix}$
  * $\Lambda_{p,q} = \diag(\lambda_1, ..., \lambda_p, \lambda_{n-q+1}, ..., \lambda_n)$
* Not unique
  * If $\hat{X}$ is a solution, then so is $\hat{X} Q$
* $\max_i \|\hat{x}_i - Q x_i \| \stackrel{a.s.}{\to} 0$ (Rubin-Delanchy et al., 2020)

# Next Time ...

* All $A \sim \BG(P)$ are GRDPGs
  * If $P$ is positive semidefinite, then it is a RDPG
* Includes SBM, DCBM, PABM
* What kind of latent structures correspond to block models?
* Can we use the ASE to estimate block model parameters?

### 2021-11-27 s675/blockmodels-and-rdpg.Rmd
notes for friday

---
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: 
        - page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

set.seed(314159)
source('~/dev/pabm-grdpg/functions.R')
```

## {.plain}

\center

\LARGE

\textcolor{darkred}{Block Models\\and\\(Generalized) Random Dot Product Graphs}

\normalsize

STAT-S 675

Fall 2021

# Recall From Last Time ...

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newcommand{\argmin}{\mathrm{argmin}}

* Let $G = (V, E)$ be an undirected and hollow graph with $|V| = n$ 
and adjacency matrix $A$
  * $A \in \mathbb{R}^{n \times n}$ is symmetric with zero diagonals
* Suppose $G \sim F(\theta)$
  * What kind of $F(\theta)$ make sense here?
  * Given $F$ and observed $G$, how can we estimate $\theta$?

# Recall From Last Time ...

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=55%}

$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

For estimation, we need to impose some structure on $P$.

::::

:::: {.column width=45%}

**Example 1**: If every entry $P_{ij} = \theta \in (0, 1)$, 
then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / 5
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

# Recall From Last Time ...

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 2**: Stochastic Block Model with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

* To make this an assortative SBM, set $p q > r^2$.
* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

# Recall From Last Time ...

Erdos-Renyi Model (1959)

* $P_{ij} = \theta$ (not a block model)
* 1 parameter $\theta$

Stochastic Block Model (Lorrain and White, 1971)

* $P_{ij} = \theta_{z_i z_j}$
* $K (K + 1) / 2$ parameters $\theta_{kl}$

Degree Corrected Block Model (Karrer and Newman, 2011)

* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
* $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$

Popularity Adjusted Block Model (Sengupta and Chen, 2017)

* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
* $K n$ parameters $\lambda_{ik}$

# Recall From Last Time ...

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim \BG(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
* $A \sim \BG(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

# Recall From Last Time ...

### Adjacency Spectral Embedding

To embed in $\mathbb{R}^d$,

1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
where $\hat{\Lambda} \in \mathbb{R}^{d \times d}$ and 
$\hat{V} \in \mathbb{R}^{n \times d}$.  
For RDPG, use $d$ greatest eigenvalues; 
for GRDPG, use $p$ most positive and $q$ most negative eigenvalues.

2. For RDPG, let $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$; 
for GRDPG, let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.

\vspace*{.5\baselineskip}

RDPG: $\max\limits_i \|\hat{x}_i - W x_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al., 2018)  
GRDPG: 
$\max\limits_i \|\hat{x}_i - Q x_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al., 2020)

# Connecting Block Models to (G)RDPGs

All Bernoulli Graphs are RDPG (if $P$ is 
positive semidefinite)  
or GRDPG (in general).

**Example 2** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

::::

:::

# Connecting the SBM to the RDPG

**Example 2** (cont'd): If we want to perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches point masses.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from SBM')
points(X, pch = 16, col = z * 2)
```

# Connecting the DCBM to the RDPG

**Example 3**: DCBM with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p \omega_i \omega_j & z_i, z_j = 1 \\
q \omega_i \omega_j & z_i, z_j = 2 \\
r \omega_i \omega_j & z_i \neq z_j
\end{cases}$

* To make this an assortative DCBM, set $p q > r^2$.
* In this example, $p = q = 1/4$, $r = 1/2$, and $\omega_i \in (0, 1)$.

::::

:::: column

```{r, fig.height = 4, fig.width = 4, out.width = '100%'}
n1 <- 100
n2 <- 100
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/4
q <- 1/4
r <- 1/8
omega <- rbeta(n, 2, 1)
Omega <- tcrossprod(omega)
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
P <- P * Omega
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

# Connecting the DCBM to the RDPG

**Example 3** (cont'd): Assortative DCBM ($p q > r^2$) with $K = 2$

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 5, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} \omega_1 & 0 \\
\vdots & \vdots \\
\sqrt{p} \omega_{n_1} & 0 \\
\sqrt{\frac{r^2}{p}} \omega_{n_1 + 1} & \sqrt{q - \frac{r^2}{p}} \omega_{n_1 + 1} \\
\vdots & \vdots \\
\sqrt{\frac{r^2}{p}} \omega_n & \sqrt{q - \frac{r^2}{p}} \omega_n
\end{bmatrix}$$

::::

:::

# Connecting the SBM to the RDPG

**Example 3** (cont'd): If we want to perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm on $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches line segments.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, 
     # asp = 1, 
     col = z * 2, 
     xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from DCBM')
lines(rbind(c(0, 0), X[z == 1, ]), pch = 16, col = 2, lwd = 3)
lines(rbind(c(0, 0), X[z == 2, ]), pch = 16, col = 4, lwd = 3)
```

# Assortative SBM and DCBM are RDPGs

* $P$ is positive semidefinite $\implies$ RDPG
* Number of embedding dimensions $=$ number of communities
* Latent structure for the SBM are point masses,  
latent structure for the DCBM are line segments
* If the SBM or DCBM is not assortative, then $P$ is not positive semidefinite
  * GRDPG instead of RDPG
  * Number of embedding dimensions $=$ number of communities, 
  except split by positive and negative eigenvalues
  * ASE still converges to true latent positions up to multiplication by unidentifiable $Q \in \mathbb{O}(p, q)$
  * Multiplication by $Q$ doesn't change overall latent structure

# Demo

# Connecting the PABM to the GRDPG

**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.

**Lemma** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.

# Connecting the PABM to the GRDPG

**Theorem**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
* $U \in \mathbb{O}(K^2)$
* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
composed of $\{\lambda^{(kl)}\}_K$ 
with each block corresponding to a community.

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}$$

# Connecting the PABM to the GRDPG

::: columns

:::: column

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

::::

:::: column

\vspace*{2\baselineskip}

$$U \in \mathbb{O}(K^2)$$

::::

:::

$$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$

**Remark 1** (orthogonality of subspaces):
If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $y_i^\top y_j = 0$.

**Remark 2** (non-uniqueness of the latent configuration):  
If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.

# Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,

$$
\max_{i, j} \hat{B}^{(ij)}_n = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

### 2021-11-27 s675/disassortative-dcbm-demo.R

@@ -1,0 +1,33 @@
+set.seed(475675)
+
+n1 <- 200
+n2 <- 200
+n <- n1 + n2
+z <- c(rep(1, n1), rep(2, n2))
+p <- 1/50
+q <- 1/50
+r <- 1/5
+omega <- rbeta(n, 2, 1)
+Omega <- tcrossprod(omega)
+P <- matrix(r, nrow = n, ncol = n)
+P[seq(n1), seq(n1)] <- p
+P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
+P <- P * Omega
+A <- draw.graph(P)
+qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
+
+A.eigen <- eigen(A)
+X.hat <- A.eigen$vectors[, c(1, n)] %*% diag(abs(A.eigen$values[c(1, n)]) ** .5)
+plot(X.hat, 
+     asp = 1,
+     col = z * 2, 
+     xlab = NA, ylab = NA)
+
+norms <- apply(X.hat, 1, function(x) sqrt(sum(x ^ 2)))
+cosine.sim <- X.hat %*% t(X.hat) / tcrossprod(norms)
+eigen.cosine.sim <- eigen(cosine.sim, symmetric = TRUE)
+Y <- sweep(eigen.cosine.sim$vectors[, 1:2], 1, 
+           sqrt(eigen.cosine.sim$values[1:2]), `*`)
+plot(Y, col = z * 2, asp = 1)
+zhat <- kmeans(Y, 2)$cluster
+table(z, zhat)

### 2021-11-30 s675/rdpg.Rmd
doing an r demo for lecture 2 as well

@@
 * $\hat{X} = V_{p,q} |\Lambda_{p,q}|^{1/2}$
-  * $V_{p,q} = \begin{bmatrix} V_{1:p} & | & V_{n-q+1:n} \end{bmatrix}$
+  * $V_{p,q} = \begin{bmatrix} V_{1:p} \mid V_{n-q+1:n} \end{bmatrix}$
   * $\Lambda_{p,q} = \diag(\lambda_1, ..., \lambda_p, \lambda_{n-q+1}, ..., \lambda_n)$
@@
 
+# R Demo
+
 # Next Time ...
+
+We can connect block models to the (G)RDPG
 

### 2021-11-30 s675/grdpg-r-demo.R
code scraps for r demo

@@ -1,0 +1,25 @@
+A <- as.matrix(igraph::as_adjacency_matrix(karate, type = 'both'))
+z <- igraph::vertex_attr(karate)$Faction
+n <- length(z)
+qgraph::qgraph(A)
+eigen.A <- eigen(A, symmetric = TRUE)
+V <- eigen.A$vectors
+D <- eigen.A$values
+X <- sweep(V[, c(1:3, n)], 2, sqrt(abs(D[c(1:3, n)])), `/`)
+plot(X, asp = 1, col = z)
+pairs(X, asp = 1, col = z)
+
+V <- V[, c(1:3, n)]
+B <- abs(V %*% t(V))
+
+n <- 100
+theta <- rbeta(n, 1/4, 1/4)
+hist(theta)
+X <- cbind(cos(pi * theta / 2 - pi / 4), sin(pi * theta / 2 - pi / 4))
+plot(X, asp = 1)
+P <- X %*% diag(c(1, -1)) %*% t(X)
+summary(as.numeric(P))
+A <- draw.graph(P)
+qgraph::qgraph(A)
+Xhat <- embedding(A, 1, 1)
+plot(Xhat, asp = 1)

### 2021-12-28 summary_minh_edited.tex
adjustments based on reviewer comments

@@
 
+\noindent%
+{\it Keywords:} network analysis, block models, generalized random dot product graphs, community detection, sparse subspace clustering
+
 \providecommand{\tightlist}{%
@@
 \ref{real-data-examples}. Section \ref{discussion} concludes.
+Proofs of Theorems 1 and 2 are provided in the body of the text while proofs of 
+Theorems 3, 4, and 5 are provided in the Appendix.
 
@@
 
-Let $G = (V, E)$ be an unweighted, undirected, 
-and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. The matrix
+Let $G = (V, E)$ be an unweighted, undirected graph without self-loops, with vertex set $V$ ($|V| = n$) and edge set $E$. The matrix
 $A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ 
 such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ 
-and $0$ otherwise. Because $G$ is symmetric and hollow,
-$A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$. 
+and $0$ otherwise. 
+$A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$ (where $[n] = \{1, 2, ..., n\}$). 
 We further restrict our analyses to Bernoulli graphs. 
@@
 (or $P$ if available) and choosing the $p$ most positive and $q$ most negative
-eigenvalues and their corresponding eigenvectors 
+eigenvalues, $D = \diag(\lambda_1, ..., \lambda_p, \lambda_{n - q + 1}, ..., \lambda_n)$, and their corresponding eigenvectors, $V \in \mathbb{R}^{n \times (p + q)}$,
 to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
@@
 the matrix $\tilde{P}$. Note that $\tilde{P}$ has the form
-$$P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} &
+
+$$\tilde{P} = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} &
   P^{(22)} \end{bmatrix} = \begin{bmatrix} \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
@@
 In summary we can describe the PABM with $K$ communities as a GRDPG with latent
-positions $\tilde{\Pi} X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
+positions $\tilde{\Pi} X U$ and known signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
 \tfrac{1}{2} K (K - 1)\bigr)$.
@@
 \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
-\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X c\) and
-\(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
+\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
+\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = 1, ..., n\). The solutions are
 collected collected into matrix
 \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
-construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
+construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
 exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
@@
 \begin{equation} \label{eq:ssc}
-c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i} c\|^2_2 + \vartheta \|c\|_1
+c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i}^\top c\|^2_2 + \vartheta \|c\|_1
 \end{equation}
@@
 \begin{remark}
-One of the most common approach to show that SSC works for a noisy
-sample $X$ is to show that 
-$X$ satisfies the
-subspace detection property for some choice of $\vartheta$; recall
-that $\vartheta$ is the sparsity parameter $\vartheta$ for
+One common approach to show that SSC works for a noisy sample $X$ is to show that $X$ satisfies the subspace detection property for some choice of $\vartheta$; 
+recall that $\vartheta$ is the sparsity parameter for
 the LASSO problems in Eq.~\eqref{eq:ssc}. However, this is not sufficient to
@@
 adjacency matrix $A \sim \BG(P)$, the natural approach then is
-to perform SSC on the rows of the spectral embedding of \(A\). We will
-show in Theorem~\ref{theorem5} that, with probability converging to
+to perform SSC on the rows of the spectral embedding of \(A\), 
+since they will still lie approximately on the $K$ subspaces 
+\citep{rubindelanchy2017statistical}. 
+We will show in Theorem~\ref{theorem5} that, with probability converging to
 one as $n \rightarrow \infty$, the rows of the ASE of $A$ also
@@
 such that points within each community lie on \(K\)-dimensional
-orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $z_i \not = z_j$. 
-Thus if
-we have (or can estimate) \(XU\) directly, then both the community
+orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $i$ and $j$ are in different communities. 
+Thus ifwe have (or can estimate) \(XU\) directly, then both the community
 detection and parameter identification problem are trivial because \(U\)
@@
 implementations using the rows of $A$ and using the spectral embedding
-of $A$ as SSC-A and SSC-ASE, respectively. The parameters \(\vartheta\)
-that controls the sparsity for SSC-A and SSC-ASE were chosen via a preliminary cross-validation
-experiment. The final clustering step of SSC-A and SSC-ASE was done
+of $A$ as SSC-A and SSC-ASE, respectively. 
+The parameters \(\vartheta\)
+that controls the sparsity for SSC-A and SSC-ASE were chosen via a preliminary cross-validation experiment. 
+In practice, $\vartheta$ that guarantee SDP (if it is possible for the particular simulated data) often result in more than $K$ disconnected subgraphs, so a smaller $\vartheta$ that does not guarantee SDP was chosen,
+and the final clustering step of SSC-A and SSC-ASE was done
 by fitting a Gaussian Mixture Model to the normalized Laplacian
@@
 
-Comparing against the ground truth species labels, OSC achieves an adjusted Rand index of 92\%, while SSC on the ASE achieves an adjusted Rand index of 96\%.
-In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
+Comparing against the ground truth species labels, \citet{noroozi2019estimation} report that SSC on the adjacency matrix achieves an adjusted Rand index of 73\% in their implementation, whereas OSC achieves 92\% and SSC on the ASE achieves 96\%.
 
@@
 \appendix
-\renewcommand\refname{References}
-% \bibliographystyle{plain}
-% \bibliography{misc.bib}
-\printbibliography
 
@@
 \end{proof}
+
+\bigskip
+\begin{center}
+{\large\bf SUPPLEMENTAL MATERIALS}
+\end{center}
+
+\begin{description}
+
+\item[Source files:] The source files used to compile this document can be found in summary.zip.
+
+\item[Data:] The data used in section 5 can be found in data.zip.
+
+\item[Code:] The code is not uploaded here but can be found online at \\
+\url{https://github.com/johneverettkoo/pabm-grdpg}.
+
+\end{description}
+
+\renewcommand\refname{References}
+% \bibliographystyle{plain}
+% \bibliography{misc.bib}
+\printbibliography
+
 \end{document}

### 2021-12-28 s675/grdpg-r-demo.R

@@ -1,3 +1,4 @@
+data(karate, package = 'igraphdata')
 A <- as.matrix(igraph::as_adjacency_matrix(karate, type = 'both'))
 z <- igraph::vertex_attr(karate)$Faction
 n <- length(z)
@@ -5,7 +6,7 @@ qgraph::qgraph(A)
 eigen.A <- eigen(A, symmetric = TRUE)
 V <- eigen.A$vectors
 D <- eigen.A$values
-X <- sweep(V[, c(1:3, n)], 2, sqrt(abs(D[c(1:3, n)])), `/`)
+X <- sweep(V[, c(1:3, n)], 2, sqrt(abs(D[c(1:3, n)])), `*`)
 plot(X, asp = 1, col = z)
 pairs(X, asp = 1, col = z)
 

### 2021-12-28 plotting-temp.R
cleaned up simulation code

@@ -1,0 +1,59 @@
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(
+    med.err = median(error),
+    first.q = quantile(error, .25),
+    third.q = quantile(error, .75),
+    med.err.ssc = median(error.ssc),
+    first.q.ssc = quantile(error.ssc, .25),
+    third.q.ssc = quantile(error.ssc, .75),
+    med.err.ep = median(error.ep, na.rm = TRUE),
+    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
+    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
+    med.err.louvain = median(error.louvain),
+    first.q.louvain = quantile(error.louvain, .25),
+    third.q.louvain = quantile(error.louvain, .75)
+  ) %>% 
+  dplyr::ungroup() %>% 
+  dplyr::inner_join(
+    ssc.df %>% 
+      dplyr::group_by(n, K) %>% 
+      dplyr::summarise(med.err.ssc.A = median(error.ssc2),
+                       first.q.ssc.A = quantile(error.ssc2, .25),
+                       third.q.ssc.A = quantile(error.ssc2, .75)) %>% 
+      dplyr::ungroup()
+  ) %>% 
+  ggplot() +
+  theme_bw() + 
+  theme(text = element_text(size = 20)) + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
+  scale_y_log10() +
+  labs(y = 'community detection error count', 
+       colour = NULL, shape = NULL, size = NULL) +
+  geom_line(aes(x = n, y = med.err * n,
+                colour = 'OSC')) +
+  geom_point(aes(x = n, y = med.err * n,
+                 colour = 'OSC', shape = 'OSC'), size = 3) +
+  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
+                    colour = 'OSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc * n,
+                colour = 'SSC-ASE')) +
+  geom_point(aes(x = n, y = med.err.ssc * n,
+                 colour = 'SSC-ASE', shape = 'SSC-ASE'), size = 3) +
[…]

### 2021-12-28 simulations-and-data-analysis/cluster-sim-balanced.R

@@ -1,0 +1,117 @@
+# simulation study for clustering algorithms
+# compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
+# simulations are for balanced community sizes
+# configuration as-is parallelizes simulations across 10 threads
+# which uses ~20 GB memory at its peak
+# modify line 28 to increase/decrease parallelization (and memory use)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+source('code-for-review/functions.R')
+
+# simulation parameters
+K.vec <- c(2, 3, 4)
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- 1e-2
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+iter <- 50
+n.vec <- rev(n.vec)
+K.vec <- rev(K.vec)
+
+doMC::registerDoMC(10)
+
+# clustering simulation
+clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  reorder.mat <- gtools::permutations(K, K)
+  out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('K =', K, ', n =', n))
+    foreach(i = seq(iter), 
+            .combine = dplyr::bind_rows, 
+            .errorhandling = 'remove') %dopar% {
+      Pz <- generate.P.beta(n, K, a1, a2, b1, b2)
+      P <- Pz$P
+      z <- Pz$clustering
+      A <- draw.graph(P)
+      clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
+      error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
+      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
+      error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
+      clustering.louvain <- mod.max(A)
+      error.louvain <- 1 - cluster.acc(clustering.louvain, z)
+      clustering.ssc.A <- ssc2(A, K, sparsity, 
+                               normalize = TRUE, parallel = FALSE)
+      error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
[…]

### 2021-12-28 simulations-and-data-analysis/cluster-sim-imbalanced.R

@@ -1,0 +1,116 @@
+# simulation study for clustering algorithms
+# compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
+# simulations are for imbalanced community sizes
+# configuration as-is parallelizes simulations across 10 threads
+# which uses ~20 GB memory at its peak
+# modify line 29 to increase/decrease parallelization (and memory use)
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+source('simulations-and-data-analysis/functions.R')
+
+# simulation parameters
+K.vec <- c(2, 3, 4)
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- 1e-2
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+iter <- 50
+n.vec <- rev(n.vec)
+K.vec <- rev(K.vec)
+set.seed(314159)
+
+doMC::registerDoMC(10)
+
+# clustering simulation
+clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  reorder.mat <- gtools::permutations(K, K)
+  out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('K =', K, ', n =', n))
+    foreach(i = seq(iter), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+      Pz <- generate.P.beta(n, K, a1, a2, b1, b2, unbalanced = TRUE)
+      P <- Pz$P
+      z <- Pz$clustering
+      A <- draw.graph(P)
+      clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
+      error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
+      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
+      error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
+      clustering.louvain <- mod.max(A)
+      error.louvain <- 1 - cluster.acc(clustering.louvain, z)
+      clustering.ssc.A <- ssc2(A, K, sparsity, 
+                               normalize = TRUE, parallel = FALSE)
+      error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
[…]

### 2021-12-28 simulations-and-data-analysis/functions.R

@@ -1,0 +1,342 @@
+embedding <- function(A, p = NULL, q = NULL,
+                      scale = TRUE,
+                      eps = 1e-6) {
+  if (p + q == 0) {
+    stop('one of p or q must be > 0')
+  }
+  n <- nrow(A)
+  eigen.A <- eigen(A, symmetric = TRUE)
+  if (is.null(p) | is.null(q)) {
+    keep <- (abs(eigen.A$values) > eps)
+  } else {
+    if (p * q > 0) {
+      keep <- c(seq(p), seq(n, n - q + 1))
+    } else if (p == 0) {
+      keep <- seq(n, n - q + 1)
+    } else {
+      keep <- seq(p)
+    }
+  }
+  
+  U <- eigen.A$vectors[, keep]
+  if (scale) {
+    S <- diag(sqrt(abs(eigen.A$values[keep])))
+    return(U %*% S)
+  } else {
+    return(U * sqrt(n))
+  }
+}
+
+draw.graph <- function(P) {
+  n <- nrow(P)
+  A <- matrix(0, nrow = n, ncol = n)
+  A[upper.tri(A)] <- rbinom(n * (n - 1) / 2, 1, P[upper.tri(P)])
+  A <- A + t(A)
+  return(A)
+}
+
+I.pq <- function(p, q) {
+  return(diag(c(rep(1, p), rep(-1, q))))
+}
+
+normalized.laplacian <- function(W) {
+  n <- nrow(W)
+  d.neg.sqrt <- colSums(W) ** -.5
+  d.neg.sqrt[is.infinite(d.neg.sqrt)] <- 0
+  D.neg.sqrt <- diag(d.neg.sqrt)
+  I <- diag(n)
+  return(I - D.neg.sqrt %*% W %*% D.neg.sqrt)
+}
+
+graph.laplacian <- function(W) {
+  D <- diag(colSums(W))
+  L <- D - W
+  return(L)
+}
+
+cluster.pabm <- function(A, K = 2, 
+                         normalize = TRUE, 
+                         use.all = TRUE,
+                         p = NULL, q = NULL, d.eigenmap = K + 1, 
+                         laplacian = 'normalized') {
+  n <- nrow(A)
+  if (is.null(p)) {
+    p <- K * (K + 1) / 2
+  }
+  if (is.null(q)) {
+    q <- K * (K - 1) / 2
+  }
+  if (p * q > 0) {
+    indices <- c(seq(p), seq(n, n - q + 1))
+  } else if (p == 0) {
[…]

### 2021-12-28 plotting-temp.R
updated plots

@@ -25,11 +25,18 @@ clustering.df %>%
   ) %>% 
   ggplot() +
   theme_bw() + 
-  theme(text = element_text(size = 20)) + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  theme(text = element_text(size = 20),
+        legend.position = 'bottom') + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
+                labels = c(expression(2^7), 
+                           expression(2^8), 
+                           expression(2^9), 
+                           expression(2^10), 
+                           expression(2^11), 
+                           expression(2^12))) +
   # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
   scale_y_log10() +
-  labs(y = 'community detection error count', 
+  labs(y = 'error count', 
        colour = NULL, shape = NULL, size = NULL) +
   geom_line(aes(x = n, y = med.err * n,
                 colour = 'OSC')) +
@@ -57,3 +64,39 @@ clustering.df %>%
                     colour = 'MM-Louvain'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
   facet_wrap(~ K, labeller = 'label_both')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  theme_bw() + 
+  theme(text = element_text(size = 20),
+        legend.position = 'bottom') + 
+  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 3')) + 
+  geom_point(aes(x = n, y = median.rmse, colour = 'Algorithm 3', shape = 'Algorithm 3'),
+             size = 3, alpha = .5) + 
+  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
+                    colour = 'Algorithm 3'), width = .1) + 
+  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
[…]

### 2021-12-29 simulations-and-data-analysis/cluster-sim-balanced.R
finished putting together code for simulations

@@ -11,7 +11,7 @@ import::from(foreach, foreach, `%do%`, `%dopar%`)
 library(mclust)
 library(ggplot2)
 setwd('~/dev/pabm-grdpg')
-source('code-for-review/functions.R')
+source('simulations-and-data-analysis/functions.R')
 
 # simulation parameters
 K.vec <- c(2, 3, 4)

### 2021-12-29 simulations-and-data-analysis/functions.R

@@ -101,8 +101,11 @@ ssc <- function(A,
                 lambda = .01,
                 parallel = FALSE,
                 normalize = TRUE,
-                scale = TRUE) {
+                scale = TRUE, 
+                d = K + 1,
+                offset = 0) {
   # subspace clustering on the ASE of a PABM
+  
   p <- K * (K + 1) / 2
   q <- K * (K - 1) / 2
   
@@ -129,14 +132,12 @@ ssc <- function(A,
     return(betahat)
   }, .parallel = parallel) %>% 
     abs()
-  B <- sweep(B, 2, apply(B, 2, max), `/`)
+    B <- sweep(B, 2, apply(B, 2, max), `/`)
   B[is.nan(B)] <- 0
   W <- B + t(B)
   L <- normalized.laplacian(W)
   L.eigen <- eigen(L, symmetric = TRUE)
-  X <- L.eigen$vectors[, seq(N, N - K)]
-  # X <- sweep(X, 2, apply(X, 2, mean), `-`)
-  # X <- sweep(X, 2, apply(X, 2, sd), `/`)
+  X <- L.eigen$vectors[, seq(N - offset, N - d + 1 - offset)]
   clustering <- mclust::Mclust(X, K, verbose = FALSE)$classification
   return(clustering)
 }
@@ -340,3 +341,142 @@ lambda.rmse.mle <- function(P, A, z) {
     sqrt() %>%
     return()
 }
+
+cluster.sg <- function(A) {
+  # adapted from code provided by sengupta and chen:
+  # a blockmodel for node popularity in networks with community structure
+  
+  f.PA<-function(A,b){	
+    K<-max(b)       # no. of communities
+    N<-nrow(A)      # no. of nodes
+    M<-matrix(NA,nrow=N,ncol=K)  # popularity matrix
+    O<-matrix(NA,nrow=K,ncol=K)  # community interaction matrix
+    for (i in 1:N){		# calculate M
+      for (r in 1:K){
+        nodes = which(b == r)
+        M[i,r] = sum(A[i,nodes])
+      }}
+    for (r in 1:K){		# calculate O
+      for (s in r:K){
+        nodes1 = which(b == r)
+        nodes2 = which(b == s)
[…]

### 2021-12-29 simulations-and-data-analysis/karnataka.R

@@ -1,0 +1,69 @@
+# karanataka villages network data
+# https://github.com/tedwestling/multiresolution_networks.git
+
+library(mclust)
+import::from(magrittr, `%>%`)
+setwd('~/dev/pabm-grdpg')
+source('simulations-and-data-analysis/functions.R')
+
+remove.deg.0 <- function(A, z, keep = unique(z)) {
+  T <- colSums(A)
+  A <- A[T > 0, T > 0]
+  z <- z[T > 0]
+  A <- A[z %in% keep, z %in% keep]
+  z <- z[z %in% keep]
+  return(list(A = A, z = z))
+}
+
+compute.errors <- function(A, z, K = 2, lambda = 1e-1) {
+  zhat.osc <- cluster.pabm(A, K)
+  zhat.ssc <- ssc(A, K, lambda)
+  zhat.mm <- cluster.sg(A)
+  list(error.osc = 1 - cluster.acc(z, zhat.osc),
+       error.ssc = 1 - cluster.acc(z, zhat.ssc),
+       error.mm = 1 - cluster.acc(z, zhat.mm))
+}
+
+adj.path <- 'datav4.0/Data/1. Network Data/Adjacency Matrices/'
+labels.df <- haven::read_dta(file.path('datav4.0/Data', 
+                                       '2. Demographics and Outcomes',
+                                       'household_characteristics.dta'))
+
+# village 12
+A12 <- readr::read_csv(file.path(adj.path, 'adj_visitgo_HH_vilno_12.csv'),
+                       col_names = FALSE) %>% 
+  as.matrix()
+z12 <- labels.df %>% 
+  dplyr::filter(village == 12) %>%
+  .$hohreligion %>% 
+  as.numeric()
+Az12 <- remove.deg.0(A12, z12, c(1, 2))
+A12 <- Az12$A
+z12 <- Az12$z
+compute.errors(A12, z12)
+
+# village 31
+A31 <- readr::read_csv(file.path(adj.path, 'adj_visitgo_HH_vilno_31.csv'),
+                       col_names = FALSE) %>% 
+  as.matrix()
+z31 <- labels.df %>% 
+  dplyr::filter(village == 31) %>%
+  .$hohreligion %>% 
+  as.numeric()
+Az31 <- remove.deg.0(A31, z31, c(1, 2))
+A31 <- Az31$A
+z31 <- Az31$z
+compute.errors(A31, z31)
+
[…]

### 2022-01-08 simulations-and-data-analysis/cluster-sim-disassortative.R
added results from disassortative simulations

@@ -2,7 +2,7 @@
 # compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
 # simulations are for balanced community sizes and diassortative P
 # configuration as-is parallelizes simulations across 10 threads
-# which uses ~20 GB memory at its peak
+# which uses ~32 GB memory at its peak
 # modify line 28 to increase/decrease parallelization (and memory use)
 
 # packages, functions, etc.
@@ -19,13 +19,13 @@ a1 <- 1
 b1 <- 2
 a2 <- 2
 b2 <- 1
-sparsity <- 1e-2
+sparsity <- 1e-1
 n.vec <- c(128, 256, 512, 1024, 2048, 4096)
 iter <- 50
 n.vec <- rev(n.vec)
 K.vec <- rev(K.vec)
 
-doMC::registerDoMC(12)
+doMC::registerDoMC(16)
 
 # clustering simulation
 clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
@@ -37,20 +37,20 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       P <- Pz$P
       z <- Pz$clustering
       A <- draw.graph(P)
+      clustering.ssc.A <- ssc2(A, K, sparsity,
+                               normalize = TRUE, parallel = FALSE)
+      error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
       clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
-      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
+      clustering.ssc <- ssc(A, K, sparsity, normalize = TRUE, scale = FALSE)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
       clustering.louvain <- mod.max(A)
       error.louvain <- 1 - cluster.acc(clustering.louvain, z)
-      clustering.ssc.A <- ssc2(A, K, sparsity, 
-                               normalize = TRUE, parallel = FALSE)
-      error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
       out <- dplyr::tibble(K = K, n = n,
                            error.osc = error.osc,
-                           error.ssc.ase = error.ssc,
+                           error.ssc.A = error.ssc.A,
                            error.louvain = error.louvain,
[…]

### 2022-01-08 simulations-and-data-analysis/functions.R

@@ -132,7 +132,7 @@ ssc <- function(A,
     return(betahat)
   }, .parallel = parallel) %>% 
     abs()
-    B <- sweep(B, 2, apply(B, 2, max), `/`)
+  B <- sweep(B, 2, apply(B, 2, max), `/`)
   B[is.nan(B)] <- 0
   W <- B + t(B)
   L <- normalized.laplacian(W)

### 2022-01-08 simulations-and-data-analysis/param-est-disassortative.R

@@ -1,0 +1,76 @@
+# simulation using proposed parameter estimation method
+# compared against mle-based method from sengupta and chen
+# P is disassortative
+# code as is uses ~16 GB memory--adjust parallelization based on memory available
+
+# packgaes, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+source('simulations-and-data-analysis/functions.R')
+
+# parallel backend
+doMC::registerDoMC(16)
+
+# simulation params
+K.vec <- c(4, 3, 2)
+a1 <- b2 <- 1
+a2 <- b1 <- 2
+n.vec <- c(128, 256, 512, 1024, 2048, 4096)
+n.vec <- rev(n.vec)
+iter <- 50
+
+# simulation
+rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
+  p <- K * (K + 1) / 2
+  q <- K * (K - 1) / 2
+  Ipq <- I.pq(p, q)
+  
+  out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+    print(paste('n =', n, 'K =', K))
+    plyr::ldply(seq(iter), function(i) {
+      P.z <- generate.P.beta(n, K,
+                             a1 = a1, a2 = a2, b1 = b1, b2 = b2)
+      P <- P.z$P
+      z <- P.z$clustering
+      A <- draw.graph(P)
+      rmse <- lambda.rmse(P, A, z)
+      rmse.mle <- lambda.rmse.mle(P, A, z)
+      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>%
+        return()
+    }, .parallel = TRUE) %>%
+      return()
+  }
+  gc()
+  return(out)
+}
+
+readr::write_csv(rmse.df, 'rmse-k-disassortative.csv')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(K, n) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
[…]

### 2022-01-08 referee-response.Rmd
responses to referee comments

---
title: "Referee Response for JCGS Submission JCGS-21-341"
output: 
  pdf_document:
    number_sections: false
    keep_tex: true
    citation_package: natbib
# output: html_document
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: misc.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 5, 
                      fig.width = 5)
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
```

We thank the referees for their time in reviewing this paper and providing valuable comments and suggestions, which have all been taken into account for this revision. Most of the changes are in the simulation and application sections. In particular, we add a simulation study of "disassortative" PABMs and clarify the choices made in the data preprocessing.

The following sections are individual referee comments and our responses (highlighted in \textcolor{blue}{blue}).

# Referee 1 Comments

## Main comments

The interest of this paper is to provide new perspectives on PABM (theoretical and algorithmic). The theoretical results are sound. However, the algorithmic results shown in the simulations are not really impressive since they are done in settings that do not appear very challenging. Indeed, they are done with fixed numbers of communities which are quite low (no more than 4) and only with assortative structure (larger probabilities of connections within a community than between two communities). Therefore, I think the paper needs a more challenging simulation study and the question of the choice of the number of communities needs to be addressed at least with some heuristic that works in simulation studies if theoretical results are not reachable. Moreover, some points on the exposition could be made clearer. I have specific comments below where I ask for some clarifications and detail how the simulation study could be improved in my opinion.

\textcolor{blue}{
We thank the referee for the feedback. The concerns addressed here are answered in the following section. 
In particular, we make sure to address the concerns about the simulation studies. 
}

## Specific comments and questions

1. p4. What is a hollow graph?  
\textcolor{blue}{
We thank the referee for the question. 
We changed the terminology a bit here and throughout the paper 
and specify that the analysis assumes $G$ is an unweighted and undirected graph with no self-loops (no edge directly from itself to itself), 
so $A$ is binary and symmetric with zeros on the diagonal. 
}

2. p4. $G$ as a graph is non-oriented or undirected and $A$ is symmetric.  
\textcolor{blue}{
We have made minor edits clarifying the structure of $G$ and $A$. 
}

3. p4. Please define the set $[n]$.  
\textcolor{blue}{
This has been added. $[n] = \{1, 2, ..., n\}$. 
}

4. p4. Definition 1. Do you consider a probability distribution on $z_1, ..., z_n$? Please give the range of variation of $(i, j)$ when you define $P_{ij}$.  
\textcolor{blue}{
We thank the referee for the question and suggestion. 
We have included the range of $(i, j)$ in the definition. 
In regard to the probability distribution on the labels $z_1, ..., z_n$, none of the theorems require this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses. 
One thing that was discussed was a possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper. 
}

5. p4. line 50. Instead of "increasing values of the community assignments", you could write "reorganized by community memberships".  
\textcolor{blue}{
We thank the referee for the suggestion. This has been changed.
}

6. p5. Definition 2. I do not understand the definition of the subset $\mathcal{X}$. 
Is it $x \in \mathcal{X}$ if for all $y \in \mathbb{R}^d$ we have $x^\top I_{p,q} y$ or is $\mathcal{X}$ a subset on $(\mathbb{R}^d)^2$? 
Please clearly state that $d = p + q$ (not stated before Definition 3).  
\textcolor{blue}{
We thank the referee for the suggestion. 
We clarify the definition of the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ and add that $d = p + q$. 
}

7. p6. Remark 2. What are $\hat{Z}$, $V$, and $D$?  
\textcolor{blue}{
We thank the referee for the question. 
We now include in the remark the $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
}

8. top of p7. $\tilde{P}$ instead of ${P}$.  
\textcolor{blue}{
We have corrected the typo. 
}

9. p7 line 40-46. I don't understand the interest of this paragraph.

10. p8, line 14-21. I don't see why the permutation given by $\Pi$ has $K$ fixed points.  
\textcolor{blue}{
We thank the referee for the comment. 
The purpose of Example 1 is to illustrate this explicitly. 
Comparing matrices $X$ and $Y$, we can see that they share three columns: the first, fifth, and ninth. 
The other columns are permuted. 
From Example 1 (as well as the proof of Theorem 1 for $K = 2$), it should be clear how this property holds for larger $K$. 
The proof of Theorem 1 for arbitrary $K$ should then confirm the intuition based on $K = 2$ and $3$. 
}

11. Section 3.1. I found it confusing that community detection algorithms are detailed when $P$ is assumed to be known and then methods for estimating $P$ are recalled. A sentence in the beginning of this section to expose what is presented could help the reader.

12. p10, line 12. $c^{(i)}$ is not defined.  
\textcolor{blue}{
We thank the referee for pointing this out. 
$c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
This has been added to the paper.
}

13. p10. I didn't understand the exposition of the SSC algorithm.

14. p11, line 31. "... SSC on the spectral embedding of $A$". Could you support this assertion by references?  
\textcolor{blue}{
We thank the referee for the question. 
We have added, "... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}."
}

15. p12, line 13. It could help recall what the $z_i, z_j$ are.  
\textcolor{blue}{
We have changed the wording for this bit to explicitly talk about community labels. 
}

16. p14. Do you think that the OSC algorithm can retrieve other kinds of structure than assortative ones?  
\textcolor{blue}{
We thank the referee for the question.
In response to this, we added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$. 
Results of this simulation study are very similar to that of the first two and consistent with the theoretical results.
}  
\textcolor{blue}{
On a related note, it is our view that the distinction between assortative and disassortative is not very important for the PABM, at least when it is viewed as a type of GRDPG. 
In the case of the SBM and DCBM, it can be shown that the rank of $P$ (and therefore the number of embedding dimensions) is equal to $K$, but assortativity or disassortativity determines how many positive and negative eigenvalues $P$ has \citep{rubindelanchy2017statistical}. 
In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming $n_k > K$ for each $k \in [K]$). 
}

17. Sections 4 and 5. Why do you use either the ARI criterion or the misclassification rate to assess the recovery of the nodes clustering?  
\textcolor{blue}{
We thank the referee for the question.
The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
For the Leeds Butterfly dataset, while the paper with which we are comparing \citep{noroozi2019estimation} provides enough details to transform and preprocess the data to match their analysis, they do not provide code to run their clustering algorithm. 
Therefore, we use their reported benchmark value, which is in the form of ARI. 
In order to make an apples-to-apples comparison, we report the ARI for OSC and SSC-ASE here as well. 
For all other analyses (simulation and real data), we use misclassification error (whether rate or count). 
}

18. Do you consider sparsity in the simulated adjacency matrices?

19. Why do you limit your simulation to assortative structure? 
Is it possible to see the results with a larger value of $K$? 
To what extent does the value of $K$ impact the computational burden for the algorithms?

20. Again, in your application on the Leeds butterfly dataset, why did you limit your analysis to $K = 4$?  
\textcolor{blue}{
We thank the referee for the question. 
For the Leeds butterfly dataset, we wanted to compare results of OSC and SSC-ASE to an analysis performed by another group studying the PABM \citep{noroozi2019estimation}. 
In their results, they describe how they removed some of the nodes and discretized the edges, which we replicate in our analysis. 
}

21. Please specify in the text that the proofs for Theorems 3, 4, and 5 are given in the Appendix.  
\textcolor{blue}{
We have added this to the end of the introduction section.
}

### Typos

1. p 10, line 16 "if each $x_i$ lieS ..."  
\textcolor{blue}{
We have corrected this typo. 
}

# Referee 2 Comments

The first part of the paper shows that any PABM is a GRDPG and gives a constructive proof of the result. 
In particular, the proof gives the construction of the latent positions of the corresponding GRDPG.

The second part of the paper investigates the implications of this relation to the GRDPG (and the identified latent positions) on estimation algorithms for the PABM. 
New algorithms for community detection and parameter estimation are proposed. 
In particular, a new community detection method, named orthogonal spectral clustering (OSC), is proposed that directly exploits the result of Theorem 1. 
Furthermore, an improvement of the sparse subspace clustering (SSC) algorithm is proposed consisting of applying SSC to the adjacency spectral embedding of the network (ASE), referred to as SSC-ASE. For both algorithms it is shown that the subspace detection property holds and that the convergence rate of the detection error is faster than for classical SSC. 
Likewise, it is shown that parameter estimation can be improved. 
A numerical study illustrates the performance of the algorithms in comparison with existing methods. 
It shows that OSC performs the best, and also that SSC-ASE improves the results with respect to classical SSC in some scenarios. 
In the application on real datasets, OSC and SSC-ASE provide good results, but do not achieve the performance of modularity maximization (MM).

The topic of the paper is interesting as it concerns a recent graph model and the results are valuable. 
Contributions are two-fold: theoretic and computational. 
The paper is well written, and the presentation of the results and the proofs is very clear.

I only have some minor comments and a couple of questions:

* I have a general issue with latent position models like the GRDPG due to the nonidentifiability of the latent space. 
As stated in the paper, latent positions are not unique, and not even the dimension of the latent space is identifiable. 
Theorem 1 (or its proof) provide one of all possible latent positions. 
A general question is what is the consequence of this specific choice. 
Does the latent space (or its dimension) have an impact on the clustering results? 
Indeed, I lack intuition about what type of latent space would be optimal for community detection or estimation in the PABM. 
The space with the smallest possible dimension? 
Do the authors have an idea of the properties of the latent space that would give the best results for community detection? 
Related to this issue, is the latent space of the proof of Theorem 1 the smallest possible latent space to relate PABM to the GRDPG?  
\textcolor{blue}{
We thank the referee for the questions.
}

* This is clearly beyond the scope of the paper, but I wonder if there is a straightforward way to adapt the proposed methods to directed models or weighted graphs or any other types of graphs than binary undirected ones?

* In the simulation study, SSC-ASE works well for $K \geq 3$, but not for $K = 2$. 
The authors explain the bad performance by the use of gaussian mixtures in the last clustering step. 
However, I don't understand why this has an impact on the performance for $K = 2$, but not for $K \geq 3$.

* In the application section, MM performs better than OSC and SSC-ASE, while in the simulation study the opposite is observed. 
What is the reason for the different behavior?

* Typo on p. 10: $X c$ should be $X^\top c$. 
Likewise, $X_{-i} c$ is to be replaced with $X_{-i}^\top c$.

### 2022-01-08 referee-response.tex
responses to referee comments

--- 2022-01-08 referee-response.Rmd
@@
----
-title: "Referee Response for JCGS Submission JCGS-21-341"
-output: 
-  pdf_document:
-    number_sections: false
-    keep_tex: true
-    citation_package: natbib
-# output: html_document
-urlcolor: blue
-header-includes:
-- \usepackage{float}
-- \usepackage{mathtools}
-- \usepackage{natbib}
-- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
-- \setcitestyle{numbers,square,comma}
-- \usepackage{verbatim}
-- \usepackage{amsthm}
-- \usepackage{comment}
-bibliography: misc.bib
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Referee Response for JCGS Submission JCGS-21-341},
+  colorlinks=true,
+  linkcolor=Maroon,
+  filecolor=Maroon,
+  citecolor=Blue,
+  urlcolor=blue,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{natbib}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\setcitestyle{numbers,square,comma}
+\usepackage{verbatim}
+\usepackage{amsthm}
+\usepackage{comment}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = TRUE, 
-                      comment = NA, 
-                      warning = FALSE, 
-                      message = FALSE, 
-                      fig.pos = 'H', 
-                      fig.align = 'center', 
-                      fig.height = 5, 
-                      fig.width = 5)
-options(xtable.comment = FALSE, 
-        xtable.table.placement = 'H')
-```
+\title{Referee Response for JCGS Submission JCGS-21-341}
+\author{}
+\date{\vspace{-2.5em}}
 
-We thank the referees for their time in reviewing this paper and providing valuable comments and suggestions, which have all been taken into account for this revision.
-Most of the changes are in the simulation and application sections.
-In particular, we add a simulation study of "disassortative" PABMs and clarify the choices made in the data preprocessing.
+\begin{document}
+\maketitle
 
-The following sections are individual referee comments and our responses (highlighted in \textcolor{blue}{blue}).
+We thank the referees for their time in reviewing this paper and
+providing valuable comments and suggestions, which have all been taken
+into account for this revision. Most of the changes are in the
+simulation and application sections. In particular, we add a simulation
+study of ``disassortative'' PABMs and clarify the choices made in the
+data preprocessing.
 
-# Referee 1 Comments
+The following sections are individual referee comments and our responses
+(highlighted in \textcolor{blue}{blue}).
 
-## Main comments
+\hypertarget{referee-1-comments}{%
+\section{Referee 1 Comments}\label{referee-1-comments}}
 
-The interest of this paper is to provide new perspectives on PABM (theoretical and algorithmic).
-The theoretical results are sound.
-However, the algorithmic results shown in the simulations are not really impressive since they are done in settings that do not appear very challenging.
-Indeed, they are done with fixed numbers of communities which are quite low (no more than 4) and only with assortative structure (larger probabilities of connections within a community than between two communities).
-Therefore, I think the paper needs a more challenging simulation study and the question of the choice of the number of communities needs to be addressed at least with some heuristic that works in simulation studies if theoretical results are not reachable.
-Moreover, some points on the exposition could be made clearer.
-I have specific comments below where I ask for some clarifications and detail how the simulation study could be improved in my opinion.
+\hypertarget{main-comments}{%
+\subsection{Main comments}\label{main-comments}}
+
+The interest of this paper is to provide new perspectives on PABM
+(theoretical and algorithmic). The theoretical results are sound.
+However, the algorithmic results shown in the simulations are not really
+impressive since they are done in settings that do not appear very
+challenging. Indeed, they are done with fixed numbers of communities
+which are quite low (no more than 4) and only with assortative structure
+(larger probabilities of connections within a community than between two
+communities). Therefore, I think the paper needs a more challenging
+simulation study and the question of the choice of the number of
+communities needs to be addressed at least with some heuristic that
+works in simulation studies if theoretical results are not reachable.
+Moreover, some points on the exposition could be made clearer. I have
+specific comments below where I ask for some clarifications and detail
+how the simulation study could be improved in my opinion.
 
@@
 
-## Specific comments and questions
+\hypertarget{specific-comments-and-questions}{%
+\subsection{Specific comments and
+questions}\label{specific-comments-and-questions}}
 
-1. p4. What is a hollow graph?  
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  p4. What is a hollow graph?\\
   \textcolor{blue}{
@@
   }
-
-2. p4. $G$ as a graph is non-oriented or undirected and $A$ is symmetric.  
+\item
+  p4. \(G\) as a graph is non-oriented or undirected and \(A\) is
+  symmetric.\\
   \textcolor{blue}{
@@
   }
-
-3. p4. Please define the set $[n]$.  
+\item
+  p4. Please define the set \([n]\).\\
   \textcolor{blue}{
@@
   }
-
-4. p4. Definition 1. Do you consider a probability distribution on $z_1, ..., z_n$? Please give the range of variation of $(i, j)$ when you define $P_{ij}$.  
+\item
+  p4. Definition 1. Do you consider a probability distribution on
+  \(z_1, ..., z_n\)? Please give the range of variation of \((i, j)\)
+  when you define \(P_{ij}\).\\
   \textcolor{blue}{
@@
   }
-
-5. p4. line 50. Instead of "increasing values of the community assignments", you could write "reorganized by community memberships".  
+\item
+  p4. line 50. Instead of ``increasing values of the community
+  assignments'', you could write ``reorganized by community
+  memberships''.\\
   \textcolor{blue}{
@@
   }
-
-6. p5. Definition 2. I do not understand the definition of the subset $\mathcal{X}$. 
-Is it $x \in \mathcal{X}$ if for all $y \in \mathbb{R}^d$ we have $x^\top I_{p,q} y$ or is $\mathcal{X}$ a subset on $(\mathbb{R}^d)^2$? 
-Please clearly state that $d = p + q$ (not stated before Definition 3).  
+\item
+  p5. Definition 2. I do not understand the definition of the subset
+  \(\mathcal{X}\). Is it \(x \in \mathcal{X}\) if for all
+  \(y \in \mathbb{R}^d\) we have \(x^\top I_{p,q} y\) or is
+  \(\mathcal{X}\) a subset on \((\mathbb{R}^d)^2\)? Please clearly state
+  that \(d = p + q\) (not stated before Definition 3).\\
   \textcolor{blue}{
@@
   }
-
-7. p6. Remark 2. What are $\hat{Z}$, $V$, and $D$?  
+\item
+  p6. Remark 2. What are \(\hat{Z}\), \(V\), and \(D\)?\\
   \textcolor{blue}{
@@
   }
-
-8. top of p7. $\tilde{P}$ instead of ${P}$.  
+\item
+  top of p7. \(\tilde{P}\) instead of \({P}\).\\
   \textcolor{blue}{
@@
   }
-
-9. p7 line 40-46. I don't understand the interest of this paragraph.
-
-10. p8, line 14-21. I don't see why the permutation given by $\Pi$ has $K$ fixed points.  
+\item
+  p7 line 40-46. I don't understand the interest of this paragraph.
+\item
+  p8, line 14-21. I don't see why the permutation given by \(\Pi\) has
+  \(K\) fixed points.\\
   \textcolor{blue}{
@@
   }
-
-11.
-Section 3.1.
-I found it confusing that community detection algorithms are detailed when $P$ is assumed to be known and then methods for estimating $P$ are recalled.
-A sentence in the beginning of this section to expose what is presented could help the reader.
-
-12. p10, line 12. $c^{(i)}$ is not defined.  
+\item
+  Section 3.1. I found it confusing that community detection algorithms
+  are detailed when \(P\) is assumed to be known and then methods for
+  estimating \(P\) are recalled. A sentence in the beginning of this
+  section to expose what is presented could help the reader.
+\item
+  p10, line 12. \(c^{(i)}\) is not defined.\\
   \textcolor{blue}{
@@
   }
-
-13. p10. I didn't understand the exposition of the SSC algorithm.
-
-14. p11, line 31. "... SSC on the spectral embedding of $A$". Could you support this assertion by references?  
+\item
+  p10. I didn't understand the exposition of the SSC algorithm.
+\item
+  p11, line 31. ``\ldots{} SSC on the spectral embedding of \(A\)''.
+  Could you support this assertion by references?\\
   \textcolor{blue}{
@@
   }
-
-15. p12, line 13. It could help recall what the $z_i, z_j$ are.  
+\item
+  p12, line 13. It could help recall what the \(z_i, z_j\) are.\\
   \textcolor{blue}{
@@
   }
-
-16. p14. Do you think that the OSC algorithm can retrieve other kinds of structure than assortative ones?  
+\item
+  p14. Do you think that the OSC algorithm can retrieve other kinds of
+  structure than assortative ones?\\
   \textcolor{blue}{
@@
   Results of this simulation study are very similar to that of the first two and consistent with the theoretical results.
-}  
+  }\\
   \textcolor{blue}{
@@
   }
-
-17. Sections 4 and 5. Why do you use either the ARI criterion or the misclassification rate to assess the recovery of the nodes clustering?  
+\item
+  Sections 4 and 5. Why do you use either the ARI criterion or the
+  misclassification rate to assess the recovery of the nodes
+  clustering?\\
   \textcolor{blue}{
@@
   }
-
-18. Do you consider sparsity in the simulated adjacency matrices?
-
-19. Why do you limit your simulation to assortative structure? 
-Is it possible to see the results with a larger value of $K$? 
-To what extent does the value of $K$ impact the computational burden for the algorithms?
-
-20. Again, in your application on the Leeds butterfly dataset, why did you limit your analysis to $K = 4$?  
+\item
+  Do you consider sparsity in the simulated adjacency matrices?
+\item
+  Why do you limit your simulation to assortative structure? Is it
+  possible to see the results with a larger value of \(K\)? To what
+  extent does the value of \(K\) impact the computational burden for the
+  algorithms?
+\item
+  Again, in your application on the Leeds butterfly dataset, why did you
+  limit your analysis to \(K = 4\)?\\
   \textcolor{blue}{
@@
   }
-
-21. Please specify in the text that the proofs for Theorems 3, 4, and 5 are given in the Appendix.  
+\item
+  Please specify in the text that the proofs for Theorems 3, 4, and 5
+  are given in the Appendix.\\
   \textcolor{blue}{
@@
   }
+\end{enumerate}
 
-### Typos
+\hypertarget{typos}{%
+\subsubsection{Typos}\label{typos}}
 
-1. p 10, line 16 "if each $x_i$ lieS ..."  
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  p 10, line 16 ``if each \(x_i\) lieS \ldots{}''\\
   \textcolor{blue}{
@@
   }
+\end{enumerate}
 
-# Referee 2 Comments
+\hypertarget{referee-2-comments}{%
+\section{Referee 2 Comments}\label{referee-2-comments}}
 
-The first part of the paper shows that any PABM is a GRDPG and gives a constructive proof of the result. 
-In particular, the proof gives the construction of the latent positions of the corresponding GRDPG.
+The first part of the paper shows that any PABM is a GRDPG and gives a
+constructive proof of the result. In particular, the proof gives the
+construction of the latent positions of the corresponding GRDPG.
 
-The second part of the paper investigates the implications of this relation to the GRDPG (and the identified latent positions) on estimation algorithms for the PABM. 
-New algorithms for community detection and parameter estimation are proposed. 
-In particular, a new community detection method, named orthogonal spectral clustering (OSC), is proposed that directly exploits the result of Theorem 1. 
-Furthermore, an improvement of the sparse subspace clustering (SSC) algorithm is proposed consisting of applying SSC to the adjacency spectral embedding of the network (ASE), referred to as SSC-ASE.
-For both algorithms it is shown that the subspace detection property holds and that the convergence rate of the detection error is faster than for classical SSC.
-Likewise, it is shown that parameter estimation can be improved. 
-A numerical study illustrates the performance of the algorithms in comparison with existing methods. 
-It shows that OSC performs the best, and also that SSC-ASE improves the results with respect to classical SSC in some scenarios. 
-In the application on real datasets, OSC and SSC-ASE provide good results, but do not achieve the performance of modularity maximization (MM).
+The second part of the paper investigates the implications of this
+relation to the GRDPG (and the identified latent positions) on
+estimation algorithms for the PABM. New algorithms for community
+detection and parameter estimation are proposed. In particular, a new
+community detection method, named orthogonal spectral clustering (OSC),
+is proposed that directly exploits the result of Theorem 1. Furthermore,
+an improvement of the sparse subspace clustering (SSC) algorithm is
+proposed consisting of applying SSC to the adjacency spectral embedding
+of the network (ASE), referred to as SSC-ASE. For both algorithms it is
+shown that the subspace detection property holds and that the
+convergence rate of the detection error is faster than for classical
+SSC. Likewise, it is shown that parameter estimation can be improved. A
+numerical study illustrates the performance of the algorithms in
+comparison with existing methods. It shows that OSC performs the best,
+and also that SSC-ASE improves the results with respect to classical SSC
+in some scenarios. In the application on real datasets, OSC and SSC-ASE
+provide good results, but do not achieve the performance of modularity
+maximization (MM).
 
-The topic of the paper is interesting as it concerns a recent graph model and the results are valuable. 
-Contributions are two-fold: theoretic and computational. 
-The paper is well written, and the presentation of the results and the proofs is very clear.
+The topic of the paper is interesting as it concerns a recent graph
+model and the results are valuable. Contributions are two-fold:
+theoretic and computational. The paper is well written, and the
+presentation of the results and the proofs is very clear.
 
@@
 
-* I have a general issue with latent position models like the GRDPG due to the nonidentifiability of the latent space. 
-As stated in the paper, latent positions are not unique, and not even the dimension of the latent space is identifiable. 
-Theorem 1 (or its proof) provide one of all possible latent positions. 
-A general question is what is the consequence of this specific choice. 
-Does the latent space (or its dimension) have an impact on the clustering results? 
-Indeed, I lack intuition about what type of latent space would be optimal for community detection or estimation in the PABM. 
-The space with the smallest possible dimension? 
-Do the authors have an idea of the properties of the latent space that would give the best results for community detection? 
-Related to this issue, is the latent space of the proof of Theorem 1 the smallest possible latent space to relate PABM to the GRDPG?  
+\begin{itemize}
+\item
+  I have a general issue with latent position models like the GRDPG due
+  to the nonidentifiability of the latent space. As stated in the paper,
+  latent positions are not unique, and not even the dimension of the
+  latent space is identifiable. Theorem 1 (or its proof) provide one of
+  all possible latent positions. A general question is what is the
+  consequence of this specific choice. Does the latent space (or its
+  dimension) have an impact on the clustering results? Indeed, I lack
+  intuition about what type of latent space would be optimal for
+  community detection or estimation in the PABM. The space with the
+  smallest possible dimension? Do the authors have an idea of the
+  properties of the latent space that would give the best results for
+  community detection? Related to this issue, is the latent space of the
+  proof of Theorem 1 the smallest possible latent space to relate PABM
+  to the GRDPG?\\
   \textcolor{blue}{
@@
   }
+\item
+  This is clearly beyond the scope of the paper, but I wonder if there
+  is a straightforward way to adapt the proposed methods to directed
+  models or weighted graphs or any other types of graphs than binary
+  undirected ones?
+\item
+  In the simulation study, SSC-ASE works well for \(K \geq 3\), but not
+  for \(K = 2\). The authors explain the bad performance by the use of
+  gaussian mixtures in the last clustering step. However, I don't
+  understand why this has an impact on the performance for \(K = 2\),
+  but not for \(K \geq 3\).
+\item
+  In the application section, MM performs better than OSC and SSC-ASE,
+  while in the simulation study the opposite is observed. What is the
+  reason for the different behavior?
+\item
+  Typo on p.~10: \(X c\) should be \(X^\top c\). Likewise, \(X_{-i} c\)
+  is to be replaced with \(X_{-i}^\top c\).
+\end{itemize}
 
-* This is clearly beyond the scope of the paper, but I wonder if there is a straightforward way to adapt the proposed methods to directed models or weighted graphs or any other types of graphs than binary undirected ones?
+  \bibliography{misc.bib}
 
-* In the simulation study, SSC-ASE works well for $K \geq 3$, but not for $K = 2$. 
-The authors explain the bad performance by the use of gaussian mixtures in the last clustering step. 
-However, I don't understand why this has an impact on the performance for $K = 2$, but not for $K \geq 3$.
-
-* In the application section, MM performs better than OSC and SSC-ASE, while in the simulation study the opposite is observed. 
-What is the reason for the different behavior?
-
-* Typo on p. 10: $X c$ should be $X^\top c$. 
-Likewise, $X_{-i} c$ is to be replaced with $X_{-i}^\top c$.
+\end{document}

### 2022-01-08 summary_minh_edited.tex
updated plots
finished putting together code for simulations
addressed more reviewer comments
responses to referee comments

@@
   K}$ be a matrix with entries in $[0,1]$. Let $z_1, z_2, \dots, z_n
-\in \{1,2,\dots,K\}$. A graph $G$ with adjacency matrix $A$ is said to
+\in [K]$. A graph $G$ with adjacency matrix $A$ is said to
 be a popularity adjusted block model graph with $K$ communities, popularity vectors
 $\Lambda$, and sparsity parameter $\rho_n \in (0,1]$ if $A \sim \BG(P)$
-where the edge probability matrix $P$ has entries of the form
+where the edge probability matrix $P$ has entries $P_{ij}$, $i, j \in [n]$, of the form
 $$P_{ij} = \rho_n \lambda_{iz_j} \lambda_{jz_i}.$$
@@
 Let $\tilde{P}$ be the matrix obtained by permuting the rows and
-columns of $P$ so that the vertices are arranged according to
-increasing values of the community assignments $z_i \in
-\{1,2,\dots,K\}$. Denote by $\tilde{P}^{(k \ell)}$ the $n_k \times n_{\ell}$
+columns of $P$ so that the vertices are reorganized by community memberships $z_i \in \{1,2,\dots,K\}$ in increasing order. 
+Denote by $\tilde{P}^{(k \ell)}$ the $n_k \times n_{\ell}$
 submatrix of $\tilde{P}$ corresponding to the
@@
 are the identity matrices of dimensions $p \times p$ and $q \times q$
-respectively. Next let $\mathcal{X}$ be a subset of $\mathbb{R}^{d}$
-such that $x^{\top} I_{p,q} y \in [0,1]$ for all $x,y \in
-\mathcal{X}$. Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
+respectively. Next define $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ where $d = p + q$. 
+Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
 a $n \times d$ matrix with rows $x_i \in \mathcal{X}$. A graph $G$
@@
 error decaying at rate approximately \(n^{-1/2}\) as $n$ increases.
+
 \begin{figure}[tp]
@@
 }
-\caption{Median and IQR of community detection error. Communities are imbalanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim_imba}
+\caption{Median and IQR of community detection error. Communities are imbalanced. 
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim_imba}
 \end{figure}
@@
 
+\hypertarget{disassortative-models}{%
+\subsection{Disassortative Models}\label{disassortative-models}}
+
+\citet{rubindelanchy2017statistical} demonstrated the power of applying GRDPG-based approaches to disassortative block models. 
+Likewise, demonstrate that GRDPG-based algorithms OSC and SSC-ASE perform well on disassortative PABMs. 
+Simulations performed in this section are the same as those in Section \ref{balanced-communities} 
+with the exception of the distributions from which within and between-group popularity parameters are drawn. 
+Here, we draw these parameters such that the expected value is $1/3$ for the within-group popularity parameters 
+and $2/3$ for the between-group popularity parameters:
+
+\begin{itemize}
+\tightlist
+\item
+  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
+\item
+  Number of underlying communities \(K = 2, 3, 4\)
+\item
+  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
+  for \(k = 1, ..., K\)
+\item
+  Community labels
+  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
+\item
+  Within-group popularities
+  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\)
+\item
+  Between-group popularities
+  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\) for
+  \(k \neq \ell\)
+\end{itemize}
+
+Algorithms tested perform similarly on disassortative PABMs compared to the simulations in Sections \ref{balanced-communities} and \ref{imbalanced-communities} 
+(Fig. \ref{fig:clust_k_disassortative} and \ref{fig:lambda_est_disassortative}).
+
+\begin{remark}
+While we call the models in this section ``disassortative'', it is our view that the assortative/disassortative distinction is not pertinent to PABMs. 
+When viewed as GRDPGs, SBMs and DCBMs are assortative if the edge probability matrix $P$ is positive semidefinite 
+and disassortative if $P$ has negative eigenvalues \citep{rubindelanchy2017statistical}. 
+For the full rank PABM with a fixed $K$, $P$ will always have $K (K + 1) / 2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues.
+\end{remark}
+
+\begin{figure}[H]
+{\centering \includegraphics{summary_files/figure-latex/clust_err_k_disassortative}
+}
+\caption{Median and IQR of community detection error. Communities are approximately balanced. 
+Edge probabilities were drawn to be disassortative. 
+Simulations were repeated 50 times for each sample size.}\label{fig:clust_k_disassortative}
+\end{figure}
+
+\begin{figure}[H]
+{\centering \includegraphics{summary_files/figure-latex/lambda_est_disassortative}
+}
+\caption{Median and IQR RMSE of edge probabilities derived from the
+  outputs of Algorithm \ref{alg:param_est} (red) compared against an
+  MLE-based method (blue) described in
+  \cite{307cbeb9b1be48299388437423d94bf1}.
+  Simulations were repeated 50 times for each sample size. 
+  Communities were drawn to be approximately balanced.
+  Edge probabilities were drawn to be disassortative.}
+\label{fig:lambda_est_disassortative}
+\end{figure}
+
 \hypertarget{real-data-examples}{%
@@
 
-Comparing against the ground truth species labels, \citet{noroozi2019estimation} report that SSC on the adjacency matrix achieves an adjusted Rand index of 73\% in their implementation, whereas OSC achieves 92\% and SSC on the ASE achieves 96\%.
+Comparing against the ground truth species labels, \citet{noroozi2019estimation} 
+report that SSC on the adjacency matrix achieves an adjusted Rand index of 73\% in their implementation, 
+whereas OSC achieves 92\% and SSC on the ASE achieves 96\%.
 
@@
 \hline
-British MPs & 0.003 & 0.018 & 0.009\\
+British MPs & 0.003 & 0.012 & 0.006\\
 \hline
-Political Blogs & 0.050 & 0.196 & 0.062\\
+Political Blogs & 0.050 & 0.187 & 0.062\\
 \hline
-DBLP & 0.028 & 0.087 & 0.059\\
+DBLP & 0.028 & 0.072 & 0.059\\
 \hline
@@
 \hline
-Village 31 & 0.125 & 0.066 & 0.110\\
+Village 31 & 0.125 & 0.059 & 0.051\\
 \hline
-Village 46 & 0.052 & 0.463 & 0.078\\
+Village 46 & 0.052 & 0.069 & 0.056\\
 \hline

### 2022-01-11 referee-response.Rmd
more referee comment responses

@@
 9. p7 line 40-46. I don't understand the interest of this paragraph.  
+\textcolor{blue}{
+We thank the referee for the question. 
+The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
+In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of points that lie on 2 2-dimensional orthogonal subspaces.
+The idea is that this generalizes to larger $K$ such that $X$ consists of points that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2.
+This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
+We have added additional details to the paper to address this. 
+}
 
@@
 In the case of the SBM and DCBM, it can be shown that the rank of $P$ (and therefore the number of embedding dimensions) is equal to $K$, but assortativity or disassortativity determines how many positive and negative eigenvalues $P$ has \citep{rubindelanchy2017statistical}.
-In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming $n_k > K$ for each $k \in [K]$). 
+In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming each community spans its respective subspace).
 }
@@
 To what extent does the value of $K$ impact the computational burden for the algorithms?  
+\textcolor{blue}{
+We thank the referee for the questions. 
+Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
+Performance of the algorithms tested is similar to the other two simulation studies and consistent with the theory. 
+Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$. 
+Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
+}  
+\textcolor{blue}{
+Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
+For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
+For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$. 
+SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
+When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^3)$. 
+In practice, we have found that $K$ does not affect runtimes too much. 
+}
 
@@
 We thank the referee for the questions. 
+While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
+\\1. The embedding dimension is $K^2$ with $K (K + 1) / 2$ positive and $K (K - 1) / 2$ negative eigenvalues, 
+\\2.
+The latent structure will always consist of $K$ $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
+\\3. $B = n V V^\top$, where $V$ is the $n \times K^2$ matrix of eigenvectors of $P$, will always be such that $B_{ij} = 0$ if and only if vertices $i$ and $j$ are in different communities. 
+\\4. The rows of $V$ are such that the inner product of the $i^{th}$ and $j^{th}$ rows is zero if and only if vertices $i$ and $j$ are in different communities. 
+}  
+\textcolor{blue}{
+Therefore, the dimensionality of the latent configuration is not in question, only what kind of effect multiplication by unidentifiable $Q$ has on the configuration. 
+The OSC algorithm circumvents the nonidentifiability of the latent configuration by only considering the entries of $B = n V V^\top$. 
+SSC-ASE does rely on the latent configuration, but we show that the orthogonality property holds for the particular embedding we perform for this algorithm. 
+}  
+\textcolor{blue}{
+That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
+If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
+In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used).
 }
@@
 * This is clearly beyond the scope of the paper, but I wonder if there is a straightforward way to adapt the proposed methods to directed models or weighted graphs or any other types of graphs than binary undirected ones?
+\textcolor{blue}{
+We thank the referee for the question. 
+For directed graphs, the ASE provides a consistent estimator and thus should also work here \citep{doi:10.1080/01621459.2012.699795}. 
+For weighted graphs, if $A_{ij}$ can be an estimator for $P_{ij}$ (e.g., if $A_{ij} \sim \mathrm{Poisson}(P_{ij})$), then all of the theory should transfer over to those models as well, 
+and we believe that the theory can be adapted to other models where $g(A_{ij})$ is an estimator for $P_{ij}$. 
+We hope to address these topics in future research. 
+}
 
@@
 However, I don't understand why this has an impact on the performance for $K = 2$, but not for $K \geq 3$.  
+\textcolor{blue}{
+We thank the referee for raising the concern. 
+We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
+One possible reason for this may be because we failed to find the right hyperparameter for SSC-ASE. 
+}  
+\textcolor{blue}{
+Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather *at least* $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an "incorrect" $\vartheta$ for SSC and then performing the final clustering step.
+}
 
@@
 What is the reason for the different behavior?  
+\textcolor{blue}{
+We thank the referee for the question. 
+In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 
+While all three algorithms are specific to the PABM, they might all behave differently to the different ways in which the model is misspecified. 
+There is also the question of hyperparameter tuning, especially in the case of SSC-ASE, and we spent a bit more time tweaking the hyperparameters, resulting in performance that is a bit closer to that of MM.
+}
 
@@
 Likewise, $X_{-i} c$ is to be replaced with $X_{-i}^\top c$.  
+\textcolor{blue}{
+We have corrected the typo.
+}

### 2022-01-11 referee-response.tex
more referee comment responses

@@
 \item
-  p7 line 40-46. I don't understand the interest of this paragraph.
+  p7 line 40-46. I don't understand the interest of this paragraph.\\
+  \textcolor{blue}{
+  We thank the referee for the question. 
+  The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
+  In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of points that lie on 2 2-dimensional orthogonal subspaces.
+  The idea is that this generalizes to larger $K$ such that $X$ consists of points that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+  Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2.
+  This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
+  We have added additional details to the paper to address this. 
+  }
 \item
@@
   In the case of the SBM and DCBM, it can be shown that the rank of $P$ (and therefore the number of embedding dimensions) is equal to $K$, but assortativity or disassortativity determines how many positive and negative eigenvalues $P$ has \citep{rubindelanchy2017statistical}.
-  In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming $n_k > K$ for each $k \in [K]$). 
+  In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming each community spans its respective subspace).
   }
@@
   extent does the value of \(K\) impact the computational burden for the
-  algorithms?
+  algorithms?\\
+  \textcolor{blue}{
+  We thank the referee for the questions. 
+  Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
+  Performance of the algorithms tested is similar to the other two simulation studies and consistent with the theory. 
+  Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$. 
+  Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
+  }\\
+  \textcolor{blue}{
+  Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
+  For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
+  For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$. 
+  SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
+  When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^3)$. 
+  In practice, we have found that $K$ does not affect runtimes too much. 
+  }
 \item
@@
   We thank the referee for the questions. 
+  While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
+  \\1. The embedding dimension is $K^2$ with $K (K + 1) / 2$ positive and $K (K - 1) / 2$ negative eigenvalues, 
+  \\2.
+The latent structure will always consist of $K$ $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
+  \\3. $B = n V V^\top$, where $V$ is the $n \times K^2$ matrix of eigenvectors of $P$, will always be such that $B_{ij} = 0$ if and only if vertices $i$ and $j$ are in different communities. 
+  \\4. The rows of $V$ are such that the inner product of the $i^{th}$ and $j^{th}$ rows is zero if and only if vertices $i$ and $j$ are in different communities. 
+  }\\
+  \textcolor{blue}{
+  Therefore, the dimensionality of the latent configuration is not in question, only what kind of effect multiplication by unidentifiable $Q$ has on the configuration. 
+  The OSC algorithm circumvents the nonidentifiability of the latent configuration by only considering the entries of $B = n V V^\top$. 
+  SSC-ASE does rely on the latent configuration, but we show that the orthogonality property holds for the particular embedding we perform for this algorithm. 
+  }\\
+  \textcolor{blue}{
+  That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
+  If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
+  In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used).
   }
@@
   models or weighted graphs or any other types of graphs than binary
-  undirected ones?
+  undirected ones?\\
+  \textcolor{blue}{
+  We thank the referee for the question. 
+  For directed graphs, the ASE provides a consistent estimator and thus should also work here \citep{doi:10.1080/01621459.2012.699795}. 
+  For weighted graphs, if $A_{ij}$ can be an estimator for $P_{ij}$ (e.g., if $A_{ij} \sim \mathrm{Poisson}(P_{ij})$), then all of the theory should transfer over to those models as well, 
+  and we believe that the theory can be adapted to other models where $g(A_{ij})$ is an estimator for $P_{ij}$. 
+  We hope to address these topics in future research. 
+  }
 \item
@@
   understand why this has an impact on the performance for \(K = 2\),
-  but not for \(K \geq 3\).
+  but not for \(K \geq 3\).\\
+  \textcolor{blue}{
+  We thank the referee for raising the concern. 
+  We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
+  One possible reason for this may be because we failed to find the right hyperparameter for SSC-ASE. 
+  }\\
+  \textcolor{blue}{
+  Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather *at least* $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an "incorrect" $\vartheta$ for SSC and then performing the final clustering step.
+  }
 \item
@@
   while in the simulation study the opposite is observed. What is the
-  reason for the different behavior?
+  reason for the different behavior?\\
+  \textcolor{blue}{
+  We thank the referee for the question. 
+  In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 
+  While all three algorithms are specific to the PABM, they might all behave differently to the different ways in which the model is misspecified. 
+  There is also the question of hyperparameter tuning, especially in the case of SSC-ASE, and we spent a bit more time tweaking the hyperparameters, resulting in performance that is a bit closer to that of MM.
+  }
 \item
   Typo on p.~10: \(X c\) should be \(X^\top c\). Likewise, \(X_{-i} c\)
-  is to be replaced with \(X_{-i}^\top c\).
+  is to be replaced with \(X_{-i}^\top c\).\\
+  \textcolor{blue}{
+  We have corrected the typo.
+  }
 \end{itemize}

### 2022-01-11 summary_minh_edited.tex
more referee comment responses

@@
 e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$. 
-
-For the general case, we once again consider $\tilde{P}$ as defined in
+Thus, another way of decomposing the edge probability matrix is $\tilde{P} = X \Pi X^\top$ where the rows of $X$ describe 2 2-dimensional orthogonal subpsaces and $\Pi$ is a permutation matrix.
+
+For the general case, we can extend $\tilde{P} = X \Pi X^\top$ to larger $K$. 
+We once again consider $\tilde{P}$ as defined in
 Remark~\ref{rem:pabm_view2}.  
@@
 to some non-identifiability indefinite orthogonal transformation $Q$. 
-% from the singular value decomposition of $P$. 
 Then using results from \citet{soltanolkotabi2012}, it can be easily shown
@@
 
-Algorithms tested perform similarly on disassortative PABMs compared to the simulations in Sections \ref{balanced-communities} and \ref{imbalanced-communities} 
+OSC, SSC-ASE, and SSC-A perform similarly on disassortative PABMs compared to the simulations in Sections \ref{balanced-communities} and \ref{imbalanced-communities}. 
 (Fig. \ref{fig:clust_k_disassortative} and \ref{fig:lambda_est_disassortative}).
@@
 and disassortative if $P$ has negative eigenvalues \citep{rubindelanchy2017statistical}. 
-For the full rank PABM with a fixed $K$, $P$ will always have $K (K + 1) / 2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues.
+For the full rank PABM with a fixed $K$, $P$ will always have $K (K + 1) / 2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues regardless of whether edge probabilities within groups is higher than edge probabilities between groups.
 \end{remark}

### 2022-01-11 simulations-and-data-analysis/functions.R

@@ -342,7 +342,7 @@ lambda.rmse.mle <- function(P, A, z) {
     return()
 }
 
-cluster.sg <- function(A) {
+cluster.sg <- function(A, K = 2) {
   # adapted from code provided by sengupta and chen:
   # a blockmodel for node popularity in networks with community structure
   
@@ -400,13 +400,13 @@ cluster.sg <- function(A) {
     return(sum(q))
   }	# formula for Q
   
-  b.can = EPalgo(A,eps=0) # EP algorithm (no perturbation)
+  b.can = EPalgo(A,eps=0, K = K) # EP algorithm (no perturbation)
   Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
   Q.DC.can = rep(NA, ncol(b.can))	# array to store Q values
   for (i in 1:ncol(b.can)){
     #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
+    foo = rep(NA, K)
+    for (clus in 1:K) {foo[clus]=sum(b.can[,i]==clus)}
     if (min(foo)==0) {stop('Empty groups are not allowed')} 
     Q.PA.can[i] = Q.PA(A, b=b.can[,i])   # fit PABM
     Q.DC.can[i] = Q.DC(A, b=b.can[,i])   # fit DCBM
@@ -416,7 +416,7 @@ cluster.sg <- function(A) {
   return(b.PA)
 }
 
-EPalgo<-function(A,eps=0){
+EPalgo<-function(A,eps=0, K = 2){
   # provided by sengupta and chen:
   # a blockmodel for node popularity in networks with community structure
   
@@ -427,15 +427,15 @@ EPalgo<-function(A,eps=0){
   foo<-eigen(A, symmetric = TRUE)
   val = abs(foo$values)			# pick the top 2
   id = order(-val)				# eigenvalues of A and put
-  id_vec = id[1:2]				# their eigenvectors into a 2*N matrix
+  id_vec = id[1:K]				# their eigenvectors into a 2*N matrix
   # columns of foo$vectors = eigenvectors of A
   # we want a 2xn matrix whose rows are the leading eigenvectors
   X = t(foo$vectors[,id_vec])		
-  y = X[,1:2]
+  y = X[,1:K]
   
-  comms = 1:2
+  comms = 1:K
[…]

### 2022-01-11 referee-response.Rmd
more responses to referee comments

@@
 13. p10. I didn't understand the exposition of the SSC algorithm.  
+\textcolor{blue}{
+We thank the referee for the question. 
+We have added more detail to our description of the SSC algorithm. 
+In particular, we talk a bit about the intuition/motivation behind the algorithm. 
+This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
+}  
+\textcolor{blue}{The description of SSC now reads:
+}  
+\textcolor{blue}{
+The SSC algorithm can be described as follows: 
+Given \(X \in \mathbb{R}^{n \times d}\) with vectors
+\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
+\(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
+\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
+The solutions are collected collected into matrix
+\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
+construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
+exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
+consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
+The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
+Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+If \(X\) instead represents points near \(K\) subspaces with some noise, 
+then this property may only hold approximately and thus 
+a final graph partitioning step may be required 
+(e.g., edge thresholding or spectral clustering).
+}
 
@@
 For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
-For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$. 
+For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}.
 SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 

### 2022-01-11 referee-response.tex
more responses to referee comments

@@
 \item
-  p10. I didn't understand the exposition of the SSC algorithm.
+  p10. I didn't understand the exposition of the SSC algorithm.\\
+  \textcolor{blue}{
+  We thank the referee for the question. 
+  We have added more detail to our description of the SSC algorithm. 
+  In particular, we talk a bit about the intuition/motivation behind the algorithm. 
+  This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
+  }\\
+  \textcolor{blue}{The description of SSC now reads:
+  }\\
+  \textcolor{blue}{
+  The SSC algorithm can be described as follows: 
+  Given \(X \in \mathbb{R}^{n \times d}\) with vectors
+  \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
+  \(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
+  \(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
+  The solutions are collected collected into matrix
+  \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
+  construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
+  exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
+  consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
+  The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
+  Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+  If \(X\) instead represents points near \(K\) subspaces with some noise, 
+  then this property may only hold approximately and thus 
+  a final graph partitioning step may be required 
+  (e.g., edge thresholding or spectral clustering).
+  }
 \item
@@
   For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
-  For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$. 
+  For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}.
   SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 

### 2022-01-11 summary_minh_edited.tex
more responses to referee comments

@@
 \(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
-\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = 1, ..., n\). The solutions are
-collected collected into matrix
+\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
+The solutions are collected collected into matrix
 \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
@@
 exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
-consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
-\(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
-points near \(K\) subspaces with some noise, a final graph partitioning
-step may be required (e.g., edge thresholding or spectral clustering).
+consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
+The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
+Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+If \(X\) instead represents points near \(K\) subspaces with some noise, 
+then this property may only hold approximately and thus 
+a final graph partitioning step may be required 
+(e.g., edge thresholding or spectral clustering).
 
 In practice, due to presence of noise, SSC is often done by solving the LASSO problems
+
 \begin{equation} \label{eq:ssc}
@@
 \end{equation}
-for some sparsity parameter \(\vartheta > 0\). The \(c_i\) vectors are
-then collected into \(C\) and \(B\) as before.
+
+for some sparsity parameter \(\vartheta > 0\). 
+The \(c_i\) vectors are then collected into \(C\) and \(B\) as before.
 

### 2022-01-11 referee-response.Rmd
finished responding to referees

@@
 The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
-In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of points that lie on 2 2-dimensional orthogonal subspaces.
-The idea is that this generalizes to larger $K$ such that $X$ consists of points that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
-Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2.
+In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
+The idea is that this generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
+resulting in $K (K + 1) / 2$ $+1$'s and $K (K - 1) / 2$ $-1$'s, i.e., the diagonal eigenvalue matrix of $\Pi$ is exactly $I_{K (K + 1) / 2, K (K - 1) / 2}$. 
 This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
@@
 A sentence in the beginning of this section to expose what is presented could help the reader.
+\textcolor{blue}{
+We thank the referee for the suggestion. 
+The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
+This is not immediately obvious even after observing $P$, observing $P$ does make estimation much easier. 
+In the introduction to section 3 (before 3.1), we have added the following:
+}  
+\textcolor{blue}{
+``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
+To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.''
+}
 
@@
 \textcolor{blue}{
-The SSC algorithm can be described as follows: 
+``The SSC algorithm can be described as follows: 
 Given \(X \in \mathbb{R}^{n \times d}\) with vectors
@@
 a final graph partitioning step may be required 
-(e.g., edge thresholding or spectral clustering).
+(e.g., edge thresholding or spectral clustering).''
 }
@@
 \textcolor{blue}{
-Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather *at least* $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an "incorrect" $\vartheta$ for SSC and then performing the final clustering step.
+Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather {\em at least} $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an ``incorrect'' $\vartheta$ for SSC and then performing the final clustering step.
+}  
+\textcolor{blue}{
+We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
 }
@@
 }
+
+\newpage

### 2022-01-11 referee-response.tex
finished responding to referees

@@
   The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
-  In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of points that lie on 2 2-dimensional orthogonal subspaces.
-  The idea is that this generalizes to larger $K$ such that $X$ consists of points that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
-  Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2.
+  In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
+  The idea is that this generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+  Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
+  resulting in $K (K + 1) / 2$ $+1$'s and $K (K - 1) / 2$ $-1$'s, i.e., the diagonal eigenvalue matrix of $\Pi$ is exactly $I_{K (K + 1) / 2, K (K - 1) / 2}$. 
   This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
@@
   estimating \(P\) are recalled. A sentence in the beginning of this
-  section to expose what is presented could help the reader.
+  section to expose what is presented could help the reader.\\
+  \textcolor{blue}{
+  We thank the referee for the suggestion. 
+  The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
+  This is not immediately obvious even after observing $P$, observing $P$ does make estimation much easier. 
+  In the introduction to section 3 (before 3.1), we have added the following:
+  }\\
+  \textcolor{blue}{
+  ``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
+  To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.''
+  }
 \item
@@
   \textcolor{blue}{
-  The SSC algorithm can be described as follows: 
+  ``The SSC algorithm can be described as follows: 
   Given \(X \in \mathbb{R}^{n \times d}\) with vectors
@@
   a final graph partitioning step may be required 
-  (e.g., edge thresholding or spectral clustering).
+  (e.g., edge thresholding or spectral clustering).''
   }
@@
   \textcolor{blue}{
-  Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather *at least* $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an "incorrect" $\vartheta$ for SSC and then performing the final clustering step.
+  Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather {\em at least} $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an ``incorrect'' $\vartheta$ for SSC and then performing the final clustering step.
+  }\\
+  \textcolor{blue}{
+  We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
   }
@@
 
+\newpage
+
   \bibliography{misc.bib}

### 2022-01-11 summary_minh_edited.tex
finished responding to referees

@@
 e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$. 
-Thus, another way of decomposing the edge probability matrix is $\tilde{P} = X \Pi X^\top$ where the rows of $X$ describe 2 2-dimensional orthogonal subpsaces and $\Pi$ is a permutation matrix.
+Thus, another way of decomposing the edge probability matrix is $\tilde{P} = X \Pi X^\top$ where the rows of $X$ describe 2 2-dimensional orthogonal subspaces and $\Pi$ is a permutation matrix.
 
@@
 
-In our methods, we assume that \(K\), the number of communities, is
-known beforehand and does not require estimation.
+In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
+To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.
 
@@
 orthogonal subspaces, i.e. $\langle U^{\top} x_i, U^{\top} x_j \rangle = 0$ whenever $i$ and $j$ are in different communities. 
-Thus ifwe have (or can estimate) \(XU\) directly, then both the community
+Thus if we have (or can estimate) \(XU\) directly, then both the community
 detection and parameter identification problem are trivial because \(U\)
@@
 perfectly recover all the latent community assignments $z_1, z_2,
-\dots, z_n$, i.e., the number of mis-clustered vertices is zero
+\dots, z_n$, i.e., the number of misclustered vertices is zero
 asymptotically almost surely. We note that Theorem~\ref{theorem4} is
@@
 Theorem~1 of \citet{307cbeb9b1be48299388437423d94bf1} (the paper that
-originally introduces the PABM model) only guarantees that the {\em
-  proportion} of mis-clustered vertices converges to $0$ as $n
+originally introduces the PABM model) only guarantees that the {\em proportion} of misclustered vertices converges to $0$ as $n
 \rightarrow \infty$. Furthermore Theorem~1
@@
 From Fig.~\ref{fig:clust_err_ct_sim_imba} we once again see that the
-number of mis-labeled vertices trending to $0$ for OSC. The
+number of mislabeled vertices trending to $0$ for OSC. The
 performance of SSC-ASE is comparable to that of OSC for $K > 2$ but
@@
 After ASE, a standard clustering algorithm, e.g., single linkage, is
-used ot infer the communities.  In future research, we intend to use
+used to infer the communities.  In future research, we intend to use
 such general algorithms as baselines and measure the efficiency of the

### 2022-01-12 referee-response.Rmd
updated referee responses

@@
 \textcolor{blue}{
-We thank the referee for the comment. 
 The purpose of Example 1 is to illustrate this explicitly. 
@@
 \textcolor{blue}{
-We thank the referee for the suggestion. 
 The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-This is not immediately obvious even after observing $P$, observing $P$ does make estimation much easier. 
+This is not immediately obvious even after observing $P$, although observing $P$ does make estimation much easier. 
 In the introduction to section 3 (before 3.1), we have added the following:
@@
 We have added more detail to our description of the SSC algorithm. 
-In particular, we talk a bit about the intuition/motivation behind the algorithm. 
+In particular, we describe the intuition/motivation behind the algorithm. 
 This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 We have added, "... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}."
@@
 \textcolor{blue}{
-We have changed the wording for this bit to explicitly talk about community labels. 
+We have changed $z_i$ to ``community labels''.
 }
@@
 \textcolor{blue}{
-We thank the referee for the question.
 In response to this, we added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
@@
 \textcolor{blue}{
-We thank the referee for the question.
 The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
@@
 \textcolor{blue}{
-We thank the referee for the questions. 
 Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 For the Leeds butterfly dataset, we wanted to compare results of OSC and SSC-ASE to an analysis performed by another group studying the PABM \citep{noroozi2019estimation}. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 For directed graphs, the ASE provides a consistent estimator and thus should also work here \citep{doi:10.1080/01621459.2012.699795}. 
@@
 \textcolor{blue}{
-We thank the referee for raising the concern. 
 We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 

### 2022-01-12 referee-response.tex
updated referee responses

@@
   \textcolor{blue}{
-  We thank the referee for the comment. 
   The purpose of Example 1 is to illustrate this explicitly. 
@@
   \textcolor{blue}{
-  We thank the referee for the suggestion. 
   The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-  This is not immediately obvious even after observing $P$, observing $P$ does make estimation much easier. 
+  This is not immediately obvious even after observing $P$, although observing $P$ does make estimation much easier. 
   In the introduction to section 3 (before 3.1), we have added the following:
@@
   We have added more detail to our description of the SSC algorithm. 
-  In particular, we talk a bit about the intuition/motivation behind the algorithm. 
+  In particular, we describe the intuition/motivation behind the algorithm. 
   This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   We have added, "... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}."
@@
   \textcolor{blue}{
-  We have changed the wording for this bit to explicitly talk about community labels. 
+  We have changed $z_i$ to ``community labels''.
   }
@@
   \textcolor{blue}{
-  We thank the referee for the question.
   In response to this, we added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
@@
   \textcolor{blue}{
-  We thank the referee for the question.
   The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
@@
   \textcolor{blue}{
-  We thank the referee for the questions. 
   Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   For the Leeds butterfly dataset, we wanted to compare results of OSC and SSC-ASE to an analysis performed by another group studying the PABM \citep{noroozi2019estimation}. 
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   For directed graphs, the ASE provides a consistent estimator and thus should also work here \citep{doi:10.1080/01621459.2012.699795}. 
@@
   \textcolor{blue}{
-  We thank the referee for raising the concern. 
   We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 

### 2022-01-12 referee-response.Rmd
changes to referee comment responses based on discussion

@@
 \textcolor{blue}{
-We thank the referee for the feedback. The concerns addressed here are answered in the following section. 
+We thank the referee for the feedback. 
+The concerns addressed here are answered in the following section. 
 In particular, we make sure to address the concerns about the simulation studies. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 We changed the terminology a bit here and throughout the paper 
@@
 \textcolor{blue}{
-We thank the referee for the question and suggestion. 
 We have included the range of $(i, j)$ in the definition. 
@@
 \textcolor{blue}{
-We thank the referee for the suggestion. This has been changed.
+This has been changed.
 }
@@
 \textcolor{blue}{
-We thank the referee for the suggestion. 
 We clarify the definition of the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ and add that $d = p + q$. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 We now include in the remark the $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
@@
 \textcolor{blue}{
-We thank the referee for pointing this out. 
 $c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
@@
 \textcolor{blue}{
-We thank the referee for the question. 
 We have added more detail to our description of the SSC algorithm. 
@@
 If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
-In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used).
+In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).
 }

### 2022-01-12 referee-response.tex
changes to referee comment responses based on discussion

@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   We changed the terminology a bit here and throughout the paper 
@@
   \textcolor{blue}{
-  We thank the referee for the question and suggestion. 
   We have included the range of $(i, j)$ in the definition. 
@@
   \textcolor{blue}{
-  We thank the referee for the suggestion. This has been changed.
+  This has been changed.
   }
@@
   \textcolor{blue}{
-  We thank the referee for the suggestion. 
   We clarify the definition of the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ and add that $d = p + q$. 
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   We now include in the remark the $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
@@
   \textcolor{blue}{
-  We thank the referee for pointing this out. 
   $c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
@@
   \textcolor{blue}{
-  We thank the referee for the question. 
   We have added more detail to our description of the SSC algorithm. 
@@
   If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
-  In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used).
+  In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).
   }

### 2022-01-13 referee-response.Rmd
more responses to referee comments

@@
 \textcolor{blue}{
-This has been added. $[n] = \{1, 2, ..., n\}$. 
+We added $[n] = \{1, 2, ..., n\}$. 
 }
@@
 \textcolor{blue}{
-We have included the range of $(i, j)$ in the definition. 
-In regard to the probability distribution on the labels $z_1, ..., z_n$, none of the theorems require this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-One thing that was discussed was a possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
+The range of $(i, j)$ is now included in the definition. 
+In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
+One thing that was discussed was the possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
 }
@@
 \textcolor{blue}{
-We clarify the definition of the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ and add that $d = p + q$. 
+We now define the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ with $d = p + q$. 
 }
@@
 \textcolor{blue}{
-We now include in the remark the $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
+Remark 2 now includes that $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
 }
@@
 \textcolor{blue}{
-The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
+The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case by starting with $K = 2$. 
 In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
-The idea is that this generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+This then generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
 Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
@@
 The other columns are permuted. 
-From Example 1 (as well as the proof of Theorem 1 for $K = 2$), it should be clear how this property holds for larger $K$. 
-The proof of Theorem 1 for arbitrary $K$ should then confirm the intuition based on $K = 2$ and $3$. 
+The proof of Theorem 1 then generalizes this to arbitrary $K$. 
 }
@@
 The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-This is not immediately obvious even after observing $P$, although observing $P$ does make estimation much easier. 
+This is not immediately obvious even if we know $P$ beforehand, although having $P$ does make estimation much easier. 
 In the introduction to section 3 (before 3.1), we have added the following:
@@
 ``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.''
+To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
+After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.''
 }
@@
 In particular, we describe the intuition/motivation behind the algorithm. 
-This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
 }  
@@
 consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
-The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
+The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace.
+Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
 Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
@@
 \textcolor{blue}{
-We have added, "... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}."
+We have added, ``... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
 }
@@
 \textcolor{blue}{
-In response to this, we added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
-Results of this simulation study are very similar to that of the first two and consistent with the theoretical results.
+We added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
+Results of this simulation study are very similar to that of the first two and consistent with the theory.
 }  
@@
 18. Do you consider sparsity in the simulated adjacency matrices?  
+\textcolor{blue}{
+Sparsity was not considered in our analyses and algorithms.
+}
 
@@
 \textcolor{blue}{
-We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
+We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output estimated labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
+For instance, in section 4.1 for, the error count for SSC-ASE for $K = 2$ is around 20 when $n = 128$ and around 100 when $n = 4096$, resulting in an error rate of around $16\%$ for $n = 128$ and $2\%$ for $n = 4096$.
 }

### 2022-01-13 referee-response.tex
more responses to referee comments

@@
   \textcolor{blue}{
-  This has been added. $[n] = \{1, 2, ..., n\}$. 
+  We added $[n] = \{1, 2, ..., n\}$. 
   }
@@
   \textcolor{blue}{
-  We have included the range of $(i, j)$ in the definition. 
-  In regard to the probability distribution on the labels $z_1, ..., z_n$, none of the theorems require this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-  One thing that was discussed was a possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
+  The range of $(i, j)$ is now included in the definition. 
+  In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
+  One thing that was discussed was the possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
   }
@@
   \textcolor{blue}{
-  We clarify the definition of the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ and add that $d = p + q$. 
+  We now define the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ with $d = p + q$. 
   }
@@
   \textcolor{blue}{
-  We now include in the remark the $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
+  Remark 2 now includes that $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
   }
@@
   \textcolor{blue}{
-  The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case. 
+  The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case by starting with $K = 2$. 
   In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
-  The idea is that this generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
+  This then generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
   Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
@@
   The other columns are permuted. 
-  From Example 1 (as well as the proof of Theorem 1 for $K = 2$), it should be clear how this property holds for larger $K$. 
-  The proof of Theorem 1 for arbitrary $K$ should then confirm the intuition based on $K = 2$ and $3$. 
+  The proof of Theorem 1 then generalizes this to arbitrary $K$. 
   }
@@
   The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-  This is not immediately obvious even after observing $P$, although observing $P$ does make estimation much easier. 
+  This is not immediately obvious even if we know $P$ beforehand, although having $P$ does make estimation much easier. 
   In the introduction to section 3 (before 3.1), we have added the following:
@@
   ``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-  To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.''
+  To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
+  After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.''
   }
@@
   In particular, we describe the intuition/motivation behind the algorithm. 
-  This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
   }\\
@@
   consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
-  The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
+  The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace.
+  Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
   Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
@@
   \textcolor{blue}{
-  We have added, "... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}."
+  We have added, ``... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
   }
@@
   \textcolor{blue}{
-  In response to this, we added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
-  Results of this simulation study are very similar to that of the first two and consistent with the theoretical results.
+  We added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
+  Results of this simulation study are very similar to that of the first two and consistent with the theory.
   }\\
@@
 \item
-  Do you consider sparsity in the simulated adjacency matrices?
+  Do you consider sparsity in the simulated adjacency matrices?\\
+  \textcolor{blue}{
+  Sparsity was not considered in our analyses and algorithms.
+  }
 \item
@@
   \textcolor{blue}{
-  We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
+  We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output estimated labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
+  For instance, in section 4.1 for, the error count for SSC-ASE for $K = 2$ is around 20 when $n = 128$ and around 100 when $n = 4096$, resulting in an error rate of around $16\%$ for $n = 128$ and $2\%$ for $n = 4096$.
   }

### 2022-01-13 summary_minh_edited.tex
more responses to referee comments

@@
 In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-To motivate some of the methods, we also consider community detection and parameter estimation in the case where we observe the edge probability matrix, $P$, noting that community memberships and popularity parameters are not immediately discernible from observing $P$.
+To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
+After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.
 
@@
 consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
-The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
-Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace.
+Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
 If \(X\) instead represents points near \(K\) subspaces with some noise, 
-then this property may only hold approximately and thus 
-a final graph partitioning step may be required 
+then this property may only hold approximately and thus a final graph partitioning step may be required 
 (e.g., edge thresholding or spectral clustering).
@@
 \begin{remark}
-While we call the models in this section ``disassortative'', it is our view that the assortative/disassortative distinction is not pertinent to PABMs. 
-When viewed as GRDPGs, SBMs and DCBMs are assortative if the edge probability matrix $P$ is positive semidefinite 
-and disassortative if $P$ has negative eigenvalues \citep{rubindelanchy2017statistical}. 
-For the full rank PABM with a fixed $K$, $P$ will always have $K (K + 1) / 2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues regardless of whether edge probabilities within groups is higher than edge probabilities between groups.
+While we call the models in this section ``disassortative'', it is our view that the assortative/disassortative distinction is not applicable to PABMs due to its flexibility. 
+Unlike the SBM and DCBM, in the PABM, each vertex is free to have a higher affinity to its own community or to other communities, independent of each other. 
+In other words, within the same community, $v_i$ may have a larger popularity parameter to its own community whereas $v_j$ may have a larger popularity parameter to a different community. 
+Furthermore, when viewed as GRDPGs, the assortativity or disassortativity of SBMs and DCBMs affects whether $P$ is positive semidefinite, which affects ASE-based approaches to analyzing the graph \citep{rubindelanchy2017statistical},
+but in the full rank PABM, $P$ is never positive semidefinite and will always have $K (K + 1) / 2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues. 
 \end{remark}

### 2022-01-13 referee-response.Rmd
typo

@@
 That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
-If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
+If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the clustering 
 In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).

### 2022-01-13 referee-response.tex
typo

@@
   \(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
-  \(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
+  \(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
   The solutions are collected collected into matrix
@@
   That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
-  If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the embedding. 
+  If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the clustering 
   In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).

### 2022-01-19 referee-response.Rmd
typo
minor tweaks

@@
 \(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
-\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
-The solutions are collected collected into matrix
+\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
+The solutions are collected into matrix
 \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
@@
 If \(X\) instead represents points near \(K\) subspaces with some noise, 
-then this property may only hold approximately and thus 
+then this property may only hold approximately and 
 a final graph partitioning step may be required 

### 2022-01-19 summary_minh_edited.tex
typo
minor tweaks

@@
 \(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
-\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
-The solutions are collected collected into matrix
+\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
+The solutions are collected into matrix
 \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
@@
 If \(X\) instead represents points near \(K\) subspaces with some noise, 
-then this property may only hold approximately and thus a final graph partitioning step may be required 
+then this property may only hold approximately and a final graph partitioning step may be required 
 (e.g., edge thresholding or spectral clustering).

### 2022-01-19 jsm/abstract/abstract.tex
abstract for jsm

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc citation processing

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Popularity Adjusted Block Models are Generalized Random Dot
Product Graphs}

  \author{
        John Koo \\
    Department of Statistics, Indiana University\\
     and \\     Minh Tang \\
    Department of Statistics, North Carolina State University\\
     and \\     Michael W. Trosset \\
    Department of Statistics, Indiana University\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Popularity Adjusted Block Models are Generalized Random
Dot Product Graphs}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We connect two random graph models, the Popularity Adjusted Block Model
(PABM) and the Generalized Random Dot Product Graph (GRDPG),
demonstrating that a PABM is a GRDPG in which communities correspond to
certain mutually orthogonal subspaces of latent vectors. This insight
leads to the construction of new algorithms for community detection and
parameter estimation for the PABM, as well as improve an existing
algorithm that relies on Sparse Subspace Clustering. Using established
asymptotic properties of Adjacency Spectral Embedding for the GRDPG, we
derive asymptotic properties of these algorithms. In particular, we
demonstrate that the absolute number of community detection errors tends
to zero as the number of graph vertices tends to infinity. Simulation
experiments illustrate these properties.
\end{abstract}

\noindent%
{\it Keywords:} network analysis, community detection, sparse subspace
clustering, spectral clustering
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\bibliographystyle{agsm}
\bibliography{bibliography.bib}

\end{document}

### 2022-01-19 referee-response.Rmd
updated responses to referee comments

@@
 We thank the referee for the feedback. 
-The concerns addressed here are answered in the following section. 
+The concerns addressed here are answered below. 
 In particular, we make sure to address the concerns about the simulation studies. 
@@
 In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-One thing that was discussed was the possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
+More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
+While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 
 }

### 2022-01-19 referee-response.tex
typo
minor tweaks
updated responses to referee comments

@@
   In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-  One thing that was discussed was the possibility of a Bayesian analysis of the PABM, and we have some results using Variational Inference and Mean Field Approximation, but we believe that this is a topic for another paper.
+  More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
+  While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 
   }
@@
   \(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
-  The solutions are collected collected into matrix
+  The solutions are collected into matrix
   \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
@@
   If \(X\) instead represents points near \(K\) subspaces with some noise, 
-  then this property may only hold approximately and thus 
+  then this property may only hold approximately and 
   a final graph partitioning step may be required 

### 2022-01-26 referee-response.Rmd
updated response to 14

@@
 \textcolor{blue}{
-We have added, ``... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
+There are no references for this, as it is a novel idea. 
+SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
+The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
+}  
+\textcolor{blue}{
+We have added, ``...
+SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
 }

### 2022-01-26 referee-response.tex
updated response to 14

@@
   \textcolor{blue}{
-  We have added, ``... SSC on the rows of the spectral embedding of $A$, since they will still lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
+  There are no references for this, as it is a novel idea. 
+  SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
+  The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
+  }\\
+  \textcolor{blue}{
+  We have added, ``...
+SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
   }

### 2022-01-26 summary_minh_edited.tex
updated response to 14

@@
 adjacency matrix $A \sim \BG(P)$, the natural approach then is
-to perform SSC on the rows of the spectral embedding of \(A\), 
-since they will still lie approximately on the $K$ subspaces 
+to perform SSC on the rows of the spectral embedding of $A$, 
+since the embedding of $P$ consists of $K$ subspaces (Theorem~\ref{theorem2}), 
+and so the embedding of $A$ will lie approximately on the $K$ subspaces 
 \citep{rubindelanchy2017statistical}.

### 2022-01-26 summary_minh_edited.tex
updated response to 14

--- 2022-01-19 summary_minh_edited.tex
@@
 adjacency matrix $A \sim \BG(P)$, the natural approach then is
-to perform SSC on the rows of the spectral embedding of \(A\), 
-since they will still lie approximately on the $K$ subspaces 
-\citep{rubindelanchy2017statistical}. 
+to perform SSC on the rows of the spectral embedding of $A$, 
+since the embedding of $P$ consists of $K$ subspaces (Theorem~\ref{theorem2}), 
+and so the embedding of $A$ will lie approximately on the $K$ subspaces. 
 We will show in Theorem~\ref{theorem5} that, with probability converging to

### 2022-02-02 referee-response.Rmd

@@
 We have added, ``...
-SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
+SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''
 }

### 2022-02-02 referee-response.tex

@@
   We have added, ``...
-SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces \citep{rubindelanchy2017statistical}.''
+SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''
   }

### 2022-02-02 jsm/abstract/abstract.Rmd

---
title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

# to produce blinded version set to 1
blinded: 0

authors: 
- name: John Koo
  affiliation: Department of Statistics, Indiana University

- name: Minh Tang
  affiliation: Department of Statistics, North Carolina State University

- name: Michael W. Trosset
  affiliation: Department of Statistics, Indiana University

keywords:
- network analysis, community detection, sparse subspace clustering, spectral clustering

abstract: |
  We connect two random graph models, the Popularity Adjusted Block Model 
  (PABM) and the Generalized Random Dot Product Graph (GRDPG), demonstrating
  that a PABM is a GRDPG in which communities correspond to certain mutually
  orthogonal subspaces of latent vectors. 
  This insight leads to the construction of new algorithms for community 
  detection and parameter estimation for the PABM, as well as improve an 
  existing algorithm that relies on Sparse Subspace Clustering. 
  Using established asymptotic properties of Adjacency Spectral Embedding for 
  the GRDPG, we derive asymptotic properties of these algorithms. 
  In particular, we demonstrate that the absolute number of community detection 
  errors tends to zero as the number of graph vertices tends to infinity. 
  Simulation experiments illustrate these properties.

bibliography: bibliography.bib
output: rticles::asa_article
---

### 2022-02-02 s675/blockmodels-and-rdpg.Rmd

@@
 
+# Two Types of Bernoulli Graphs
+
+::: columns
+
+:::: column
+
+## Block Models
+
+* $P_{ij} = f(z_i, z_j)$
+
+* Community structure
+
+* Estimation via EM
+  * May require *ad hoc* approximations
+  * Can be sensitive to initial guess
+
+::::
+
+:::: column
+
+## (G)RDPGs
+
+* $P_{ij} = x_i^\top I_{p,q} x_j$
+
+* Latent space structure
+
+* Estimation via ASE
+  * Consistency theorem
+  * Straightforward algorithm
+  * Non-identifiable linear transformation
+
+::::
+
+:::
+
+## All Bernoulli Graph models are (G)RDPGs
+
+* Can we use this for inference on Block Models?
+
 # Connecting Block Models to (G)RDPGs
@@
 * To make this an assortative DCBM, set $p q > r^2$.
-* In this example, $p = q = 1/4$, $r = 1/2$, and $\omega_i \in (0, 1)$.
+* In this example, $p = q = 1/2$, $r = 1/8$, and $\omega_i \in (0, 1)$.
 
@@
 z <- c(rep(1, n1), rep(2, n2))
-p <- 1/4
-q <- 1/4
+p <- 1/2
+q <- 1/2
 r <- 1/8
@@
 $$P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
+p \omega_i \omega_j & z_i, z_j = 1 \\
+q \omega_i \omega_j & z_i, z_j = 2 \\
+r \omega_i \omega_j & z_i \neq z_j
 \end{cases}$$

### 2022-02-02 s675/rdpg.Rmd

@@
 
-\textcolor{darkred}{Random Dot Product Graphs}
+\textcolor{darkred}{(Generalized) Random Dot Product Graphs}
 
@@
 * Any $A \sim \BG(P)$ is a RDPG if $P$ is positive semidefinite  
+$P = V \Lambda V^\top \implies X = V \sqrt{\Lambda}$  
+$X X^\top = V \sqrt{\Lambda} \sqrt{\Lambda} V^\top = V \Lambda V^\top = P$
 * Whether we think of a graph as a RDPG depends on whether the latent structure is useful

### 2022-02-03 simulations-and-data-analysis/cluster-sim-sparsity.R
adding code for sparsity simulation

@@ -1,0 +1,107 @@
+# simulation study for clustering algorithms
+# compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
+# simulations are for balanced community sizes
+# configuration as-is parallelizes simulations across 16 threads
+
+# packages, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+library(mclust)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+source('simulations-and-data-analysis/functions.R')
+
+# simulation parameters
+n <- 2048
+K <- 3
+a1 <- 2
+b1 <- 1
+a2 <- 1
+b2 <- 2
+sparsity <- 1e-2
+rho.vec <- seq(.1, 1, .1)
+
+doMC::registerDoMC(16)
+
+# clustering simulation
+clustering.df <- foreach(rho = rho.vec, .combine = dplyr::bind_rows) %do% {
+  reorder.mat <- gtools::permutations(K, K)
+  print(paste('rho =', rho))
+  foreach(i = seq(iter), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+    Pz <- generate.P.beta(n, K, a1, a2, b1, b2)
+    P <- Pz$P * rho
+    z <- Pz$clustering
+    A <- draw.graph(P)
+    clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
+    error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
+    clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
+    error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
+    clustering.louvain <- mod.max(A)
+    error.louvain <- 1 - cluster.acc(clustering.louvain, z)
+    clustering.ssc.A <- ssc2(A, K, sparsity, 
+                             normalize = TRUE, parallel = FALSE)
+    error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
+    out.df <- dplyr::tibble(K = K, n = n,
+                            error.osc = error.osc,
+                            error.ssc.ase = error.ssc,
+                            error.louvain = error.louvain,
+                            error.ssc.A = error.ssc.A)
+    gc()
+    return(out.df)
+  } %>% 
[…]

### 2022-02-03 simulations-and-data-analysis/cluster-sim-sparsity.R
adding code for sparsity simulation

@@ -41,7 +41,7 @@ clustering.df <- foreach(rho = rho.vec, .combine = dplyr::bind_rows) %do% {
     clustering.ssc.A <- ssc2(A, K, sparsity, 
                              normalize = TRUE, parallel = FALSE)
     error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
-    out.df <- dplyr::tibble(K = K, n = n,
+    out.df <- dplyr::tibble(K = K, n = n, rho = rho, 
                             error.osc = error.osc,
                             error.ssc.ase = error.ssc,
                             error.louvain = error.louvain,
@@ -57,7 +57,7 @@ gc()
 readr::write_csv(clustering.df, 'clustering-sparsity.csv')
 
 clustering.df %>%
-  dplyr::group_by(n, K) %>%
+  dplyr::group_by(rho) %>%
   dplyr::summarise(
     med.err = median(error.osc),
     first.q = quantile(error.osc, .25),
@@ -79,29 +79,29 @@ clustering.df %>%
   scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
   labs(y = 'community detection error count', 
        colour = NULL, shape = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
+  geom_line(aes(x = rho, y = med.err * n,
                 colour = 'OSC')) +
-  geom_point(aes(x = n, y = med.err * n,
+  geom_point(aes(x = rho, y = med.err * n,
                  colour = 'OSC', shape = 'OSC')) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
+  geom_errorbar(aes(x = rho, ymin = first.q * n, ymax = third.q * n,
                     colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
+  geom_line(aes(x = rho, y = med.err.ssc * n,
                 colour = 'SSC-ASE')) +
-  geom_point(aes(x = n, y = med.err.ssc * n,
+  geom_point(aes(x = rho, y = med.err.ssc * n,
                  colour = 'SSC-ASE', shape = 'SSC-ASE')) +
-  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
+  geom_errorbar(aes(x = rho, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
[…]

### 2022-02-03 summary_minh_edited.tex
Premerge

@@
 are the identity matrices of dimensions $p \times p$ and $q \times q$
-respectively. Next define $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ where $d = p + q$. 
+respectively. Denote $d = p + q$ and let $\mathcal{X}$ be a subset of $\mathbb{R}^d$ such
+that, for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
+$x^{\top} I_{p,q} y \in [0,1]$. 
+% Next define $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ where $d = p + q$. 
 Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
@@
 by $P = \rho_n X I_{p,q} X^{\top}$, i.e., the entries of $P$ are of
-the form
-
-$$P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$$
-
+the form $P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$
 \end{definition}
@@
 
-\begin{remark}
-  \label{rem:ase}
-We can use Adjacency Spectral Embedding (ASE) 
-\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
-This procedure consists of taking the spectral decomposition of $A$ 
-(or $P$ if available) and choosing the $p$ most positive and $q$ most negative
-eigenvalues, $D = \diag(\lambda_1, ..., \lambda_p, \lambda_{n - q + 1}, ..., \lambda_n)$, and their corresponding eigenvectors, $V \in \mathbb{R}^{n \times (p + q)}$,
-to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
-\end{remark}
-
 \begin{definition}[Indefinite Orthogonal Group]
+  \label{def:indefinite}
 The indefinite orthogonal group with signature $(p, q)$ is
@@
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
-x_i^\top I_{p, q} x_j = P_{ij}$.
-Unlike in the RDPG case, transforming the latent positions via multiplication
+x_i^\top I_{p, q} x_j = P_{ij}$. Unlike in the RDPG case, transforming the latent positions via multiplication
 by $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or
 distances.
+\end{remark}
+
+\begin{remark}
+  \label{rem:ase}
+We can use Adjacency Spectral Embedding (ASE) 
+\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
+This procedure consists of taking the spectral decomposition of $A$
+and keeping the $p + q$ largest eigenvalues (in modulus) and
+corresponding eigenvectors of $A$. More specifically, let $A$ be an $n
+\times n$ adjacency matrix and denote the
+eigendecomposition of $A$ as
+$$A = \sum_{i} \hat{\lambda}_i \hat{v}_i \hat{v}_i^{\top}, \quad
+\text{where} \,\,
+|\hat{\lambda}|_1 \geq |\hat{\lambda}_2| \geq \dots \geq
+|\hat{\lambda}_n|.$$
+Now let $\hat{D}$ be a diagonal matrix whose diagonal entries are
+the eigenvalues
+$\{\hat{\lambda}_i\}_{i=1}^{d}$ arranged in decreasing values (not in
+decreasing modulus). Let $\hat{V}$ be the $n \times d$ matrix whose
+columns are the corresponding eigenvectors $\{\hat{v}_i\}_{i=1}^{d}$
+arranged in the same order as the diagonal entries of $\hat{D}$. Now
+define $\hat{Z} = \hat{V} |\hat{D}|^{1/2}$ where the $|\cdot|$
+operation is applied elementwise to the entries of $\hat{D}$. Then
+$\hat{Z}$ serves as an estimate of $X$, up to some non-identifiable
+orthogonal transformation $Q$ as described in
+Definition~\ref{def:indefinite} and
+Remark~\ref{rem:non_identifiable}. 
+% choosing the $p$ most positive and $q$ most negative
+% eigenvalues, $D = \diag(\lambda_1, ..., \lambda_p, \lambda_{n - q +
+%   1}, ..., \lambda_n)$, and their corresponding eigenvectors, $\hat{V}
+% \in \mathbb{R}^{n \times (p + q)}$, to construct the embedding
+%$\hat{Z} = V |D|^{1/2}$. More specifically, 
 \end{remark}

### 2022-02-04 simulations-and-data-analysis/cluster-sim-imbalanced.R
adding some simulation results

@@ -2,7 +2,7 @@
 # compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
 # simulations are for imbalanced community sizes
 # configuration as-is parallelizes simulations across 10 threads
-# which uses ~20 GB memory at its peak
+# which uses ~32 GB memory at its peak
 # modify line 29 to increase/decrease parallelization (and memory use)
 
 # packages, functions, etc.
@@ -26,7 +26,7 @@ n.vec <- rev(n.vec)
 K.vec <- rev(K.vec)
 set.seed(314159)
 
-doMC::registerDoMC(10)
+doMC::registerDoMC(16)
 
 # clustering simulation
 clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
@@ -40,7 +40,8 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
       A <- draw.graph(P)
       clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
       error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
-      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
+      clustering.ssc <- ssc(A, K, sparsity, normalize = TRUE, scale = FALSE,
+                            offset = 1, d = K - 1)
       error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
       clustering.louvain <- mod.max(A)
       error.louvain <- 1 - cluster.acc(clustering.louvain, z)
@@ -52,7 +53,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
                            error.ssc.ase = error.ssc,
                            error.louvain = error.louvain,
                            error.ssc.A = error.ssc.A)
-      gc()
+      # gc()
       return(out)
     } %>% 
       return()
@@ -63,7 +64,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
 
 gc()
 
-readr::write_csv(clustering.df, 'clustering-k-imbalanced.csv')
+# readr::write_csv(clustering.df, 'clustering-k-imbalanced.csv')
 
 clustering.df %>%
   dplyr::group_by(n, K) %>%
@@ -85,32 +86,41 @@ clustering.df %>%
   ggplot() +
   theme_bw() + 
   theme(text = element_text(size = 10)) + 
[…]

### 2022-02-04 referee-response.Rmd

@@
 
-We thank the referees for their time in reviewing this paper and providing valuable comments and suggestions, which have all been taken into account for this revision.
-Most of the changes are in the simulation and application sections.
-In particular, we add a simulation study of "disassortative" PABMs and clarify the choices made in the data preprocessing.
+We thank the referees for their time in reviewing this paper and
+providing valuable comments and suggestions, which have all been taken
+into account for this revision. Most of the changes are in the
+simulation and application sections. In particular, we add a
+simulation study of "disassortative" PABMs and clarify the choices
+made in the data preprocessing.
 
@@
 
-The interest of this paper is to provide new perspectives on PABM (theoretical and algorithmic).
-The theoretical results are sound.
-However, the algorithmic results shown in the simulations are not really impressive since they are done in settings that do not appear very challenging.
-Indeed, they are done with fixed numbers of communities which are quite low (no more than 4) and only with assortative structure (larger probabilities of connections within a community than between two communities).
-Therefore, I think the paper needs a more challenging simulation study and the question of the choice of the number of communities needs to be addressed at least with some heuristic that works in simulation studies if theoretical results are not reachable.
-Moreover, some points on the exposition could be made clearer.
-I have specific comments below where I ask for some clarifications and detail how the simulation study could be improved in my opinion.
+The interest of this paper is to provide new perspectives on PABM
+(theoretical and algorithmic). The theoretical results are
+sound. However, the algorithmic results shown in the simulations are
+not really impressive since they are done in settings that do not
+appear very challenging. Indeed, they are done with fixed numbers of
+communities which are quite low (no more than 4) and only with
+assortative structure (larger probabilities of connections within a
+community than between two communities). Therefore, I think the paper
+needs a more challenging simulation study and the question of the
+choice of the number of communities needs to be addressed at least
+with some heuristic that works in simulation studies if theoretical
+results are not reachable. Moreover, some points on the exposition
+could be made clearer. I have specific comments below where I ask for
+some clarifications and detail how the simulation study could be
+improved in my opinion.
 
@@
 \textcolor{blue}{
-We changed the terminology a bit here and throughout the paper 
-and specify that the analysis assumes $G$ is an unweighted and undirected graph with no self-loops (no edge directly from itself to itself), 
-so $A$ is binary and symmetric with zeros on the diagonal. 
-}
+We changed the terminology a bit here and throughout the paper and
+specify that the analysis assumes $G$ is an unweighted and undirected
+graph with no self-loops 
+(no edge directly from itself to itself), 
+so $A$ is binary and symmetric with zeros on the diagonal. The phrase
+``hollow graph'' is no longer used.  }
 
@@
 \textcolor{blue}{
-We have made minor edits clarifying the structure of $G$ and $A$. 
+We have made minor edits clarifying the structure of $G$ and $A$. In
+particular we assume throughout the paper that $G$ is undirected and $A$ is
+symmetric. 
 }
@@
 \textcolor{blue}{
-We added $[n] = \{1, 2, ..., n\}$. 
+We added $[n] = \{1, 2, ..., n\}$ at the beginning of Section~2.1. 
 }
 
-4. p4. Definition 1. Do you consider a probability distribution on $z_1, ..., z_n$? Please give the range of variation of $(i, j)$ when you define $P_{ij}$.  
+4. p4. Definition 1. Do you consider a probability distribution on
+   $z_1, ..., z_n$? 
+   Please give the range of variation of $(i, j)$ when you define $P_{ij}$.  
 \textcolor{blue}{
-The range of $(i, j)$ is now included in the definition. 
-In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
-While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 
-}
+The range of $(i, j)$ is now included in the definition; in particular
+we wrote in Section 2.1 that ``$1 \leq i < j \leq n$.''
+In regard to the probability distribution on the labels $z_1, ...,
+z_n$, none of our theoretical assume anything about the distribution
+of the $\{z_i\}$; we do, however, occassionally assume that the $n_k$
+are sufficiently large for each $k = 1,2,\dots, K$. More specifically 
+the convergence rate of Theorem 5 depends on the $n_k$, i.e., for
+convergence we require that $n_k \to \infty$ as $n \rightarrow
+\infty$. In practice, we also require that each $n_k$ is large enough
+so that the latent vectors in the $k$th community span its
+subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lies in a
+general position (are linearly independent), then we only require $n_k
+\geq K$.}
 
@@
 \textcolor{blue}{
-We now define the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ with $d = p + q$. 
-}
+We now denote $d = p + q$ in Definition 2 and define $\mathcal{X}$ as a subset of $\mathbb{R}^{d}$ such
+that for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
+$x^{\top} I_{p,q} y \in [0,1]$. }
 

### 2022-02-04 simulations-and-data-analysis/functions.R
more simulation results

@@ -239,7 +239,7 @@ ssc2 <- function(A,
   return(clustering)
 }
 
-lambda.rmse <- function(P, A, clustering) {
+lambda.rmse <- function(P, A, clustering, rho = 1) {
   K <- max(clustering)
   n <- nrow(A)
   n.vector <- sapply(seq(K), function(k) sum(clustering == k))
@@ -271,6 +271,7 @@ lambda.rmse <- function(P, A, clustering) {
     # magrittr::divide_by(n * K) %>% 
     magrittr::divide_by(n ** 2) %>% 
     sqrt() %>%
+    magrittr::divide_by(rho) %>% 
     return()
 }
 
@@ -292,7 +293,7 @@ estimate.lambda.block <- function(P.block, within = FALSE) {
   }
 }
 
-lambda.rmse.mle <- function(P, A, z) {
+lambda.rmse.mle <- function(P, A, z, rho = 1) {
   K <- max(z)
   n <- nrow(A)
   n.vector <- sapply(seq(K), function(k) sum(z == k))
@@ -339,6 +340,7 @@ lambda.rmse.mle <- function(P, A, z) {
     # magrittr::divide_by(n * K) %>% 
     magrittr::divide_by(n ** 2) %>% 
     sqrt() %>%
+    magrittr::divide_by(rho) %>% 
     return()
 }
 

### 2022-02-04 simulations-and-data-analysis/param-est-sparsity.R

@@ -1,0 +1,67 @@
+# simulation using proposed parameter estimation method
+# for varying rho, the sparsity parameter
+# compared against mle-based method from sengupta and chen
+
+# packgaes, functions, etc.
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`)
+library(ggplot2)
+setwd('~/dev/pabm-grdpg')
+source('simulations-and-data-analysis/functions.R')
+
+doMC::registerDoMC(16)
+
+# simulation params
+K <- 3
+a1 <- b2 <- 2
+a2 <- b1 <- 1
+n <- 2048
+rho.vec <- seq(.1, 1, .1)
+iter <- 50
+
+# simulation
+rmse.df <- foreach(rho = rho.vec, .combine = dplyr::bind_rows) %do% {
+  print(paste('rho =', rho))
+  plyr::ldply(seq(iter), function(i) {
+    P.z <- generate.P.beta(n, K,
+                           a1 = a1, a2 = a2, 
+                           b1 = b1, b2 = b2)
+    P <- P.z$P * rho
+    z <- P.z$clustering
+    A <- draw.graph(P)
+    rmse <- lambda.rmse(P, A, z, rho)
+    rmse.mle <- lambda.rmse.mle(P, A, z, rho)
+    dplyr::tibble(K = K, n = n, rho = rho, 
+                  rmse = rmse, rmse.mle = rmse.mle) %>%
+      return()
+  }, .parallel = TRUE) %>%
+    return()
+}
+
+readr::write_csv(rmse.df, 'rmse-sparsity.csv')
+
+rmse.df %>% 
+  na.omit() %>% 
+  dplyr::group_by(rho) %>% 
+  dplyr::summarise(median.rmse = median(rmse),
+                   q1.rmse = quantile(rmse, .25),
+                   q3.rmse = quantile(rmse, .75),
+                   median.rmse.mle = median(rmse.mle),
+                   q1.rmse.mle = quantile(rmse.mle, .25),
+                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  geom_line(aes(x = rho, y = median.rmse, colour = 'Algorithm 3')) + 
+  geom_point(aes(x = rho, y = median.rmse, colour = 'Algorithm 3', 
+                 shape = 'Algorithm 3'), size = 3) + 
[…]

### 2022-02-05 sparsity-sim/sparsity-sim.tex
supplemental materials

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc citation processing

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Supplemental Materials for ``Popuarity Adjusted Block
Models are Generalized Random Dot Product Graphs''}

  \author{
        John Koo \\
    Department of Statistics, Indiana University\\
     and \\     Minh Tang \\
    Department of Statistics, North Carolina State University\\
     and \\     Michael W. Trosset \\
    Department of Statistics, Indiana University\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Supplemental Materials for ``Popuarity Adjusted Block
Models are Generalized Random Dot Product Graphs''}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}

\end{abstract}

\noindent%
{\it Keywords:} 
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\hypertarget{sparsity-simulation-study}{%
\section{Sparsity Simulation Study}\label{sparsity-simulation-study}}

In this simulation study, we use the same setup as in the simulations
for balanced communities but with fixed \(K = 3\) and \(n = 2^{11}\) and
vary the sparsity parameter \(\rho \in (0, 1]\) (this was fixed at
\(\rho_n = 1\) for the previous simulations. More specifically,

\begin{itemize}
\tightlist
\item
  Number of vertices \(n = 2048\)
\item
  Number of underlying communities \(K = 3\)
\item
  Sparsity parameter \(\rho_n = 0.1, .03, 0.5, 0.7, 0.9\)
\item
  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\) (i.e.,
  each community label has an equal probability of being drawn)
\item
  Community labels
  \(z_k \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\)
\item
  Within-group popularities
  \(\lambda^{(kk)} \stackrel{\mathrm{iid}}{\sim}\mathrm{Beta}(2, 1)\)
\item
  Between-group popularities
  \(\lambda^{(k \ell)} \stackrel{\mathrm{iid}}{\sim}\mathrm{Beta}(1, 2)\)
  for \(k \neq \ell\)
\end{itemize}

Fifty simulations were performed for each combination of \(n\) and
\(K\). The results for community detection and parameter estimation
error are shown in Fig. \ref{fig:sparsity_sim}. Community detection
error is defined as the number of misclustered vertices. Parameter
estimation error is defined as \(\frac{1}{n} \|P - \hat{P}\|_F\) where
\(\hat{P}\) is the estimated edge probability matrix.

\begin{figure}[h]

{\centering \includegraphics[width=1\linewidth]{sparsity-sim_files/figure-latex/sparsity_sim-1}

}

\caption{Community detection error counts (left) and popularity parameter RMSEs (right) for $n = 2048$ and $K = 3$ and varying $\rho_n$. Simulations were repeated 50 times for each $\rho_n$.}\label{fig:sparsity_sim}
\end{figure}

\bibliographystyle{agsm}
\bibliography{bibliography.bib}

\end{document}

### 2022-02-05 sparsity-sim/supplemental-materials.Rmd
supplemental materials

---
title: Supplemental Materials for "Popularity Adjusted Block Models are Generalized Random Dot Product Graphs"
authors: John Koo, Minh Tang, Michael Trosset
# author: John Koo
# author2: Minh Tang
# author3: Michael Trosset
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
    number_sections: yes
# output:
#   bookdown::pdf_document2:
#     citation_package: natbib
#     number_citations: yes
# output: html_document
# output: rticles::rss_article
fontsize: 12pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: misc.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      # eval = FALSE,
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
```

```{r}
import::from(magrittr, `%>%`)
library(ggplot2)

theme_set(theme_bw())
```

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}

# Sparsity Simulation Study

In this simulation study, we use the same setup as in the simulations for balanced communities but with fixed $K = 3$ and $n = 2^{11}$ and vary the sparsity parameter $\rho \in (0, 1]$ (this was fixed at $\rho_n = 1$ for the previous simulations. 
More specifically,

* Number of vertices $n = 2048$
* Number of underlying communities $K = 3$
* Sparsity parameter $\rho_n = 0.1, 0.3, 0.5, 0.7, 0.9$
* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$ (i.e., each community label has an equal probability of being drawn)
* Community labels $z_k \iid \Multinomial(\alpha_1, ..., \alpha_K)$
* Within-group popularities $\lambda^{(kk)} \iid \Betadist(2, 1)$
* Between-group popularities $\lambda^{(k \ell)} \iid \Betadist(1, 2)$ for $k \neq \ell$

Fifty simulations were performed for each combination of $n$ and $K$. 
The results for community detection and parameter estimation error are shown in Fig. \ref{fig:sparsity_sim}. 
Community detection error is defined as the number of misclustered vertices. 
Parameter estimation error is defined as $\frac{1}{n} \|P - \hat{P}\|_F$ where 
$\hat{P}$ is the estimated edge probability matrix.

```{r sparsity_sim, fig.cap = 'Community detection error counts (left) and popularity parameter RMSEs (right) for $n = 2048$ and $K = 3$ and varying $\\rho_n$. Simulations were repeated 50 times for each $\\rho_n$.', fig.width = 13, fig.height = 5, out.width = '100%'}
clustering.df <- readr::read_csv('../clustering-sparsity.csv')

clustering.plot <- clustering.df %>%
  dplyr::group_by(n, K, rho) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(text = element_text(size = 20),
        legend.position = 'bottom') + 
  scale_x_continuous(breaks = c(0.1, 0.3, 0.5, 0.7, 0.9)) +
  scale_y_log10() +
  labs(y = 'error count', 
       x = expression(rho[n]), 
       colour = NULL, shape = NULL, size = NULL) +
  geom_line(aes(x = rho, y = med.err * n,
                colour = 'OSC')) +
  geom_point(aes(x = rho, y = med.err * n,
                 colour = 'OSC', shape = 'OSC'), size = 3) +
  geom_errorbar(aes(x = rho, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .05) + 
  geom_line(aes(x = rho, y = med.err.ssc * n,
                colour = 'SSC-ASE')) +
  geom_point(aes(x = rho, y = med.err.ssc * n,
                 colour = 'SSC-ASE', shape = 'SSC-ASE'), size = 3) +
  geom_errorbar(aes(x = rho, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-ASE'), width = .05) +
  geom_line(aes(x = rho, y = med.err.ssc.A * n,
                colour = 'SSC-A')) +
  geom_point(aes(x = rho, y = med.err.ssc.A * n,
                 colour = 'SSC-A', shape = 'SSC-A'), size = 3) +
  geom_errorbar(aes(x = rho, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'SSC-A'), width = .05) +
  geom_line(aes(x = rho, y = med.err.louvain * n,
                colour = 'MM-Louvain')) + 
  geom_point(aes(x = rho, y = med.err.louvain * n,
                 colour = 'MM-Louvain', shape = 'MM-Louvain'), size = 3) + 
  geom_errorbar(aes(x = rho, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'MM-Louvain'), width = .05) + 
  scale_colour_brewer(palette = 'Set1')

rmse.df <- readr::read_csv('../rmse-sparsity.csv')

rmse.plot <- rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n, rho) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(text = element_text(size = 20),
        legend.position = 'bottom') + 
  geom_line(aes(x = rho, y = median.rmse, colour = 'Algorithm 3')) + 
  geom_point(aes(x = rho, y = median.rmse, colour = 'Algorithm 3', shape = 'Algorithm 3'),
             size = 3, alpha = .5) + 
  geom_errorbar(aes(x = rho, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 3'), width = .05) + 
  geom_line(aes(x = rho, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_point(aes(x = rho, y = median.rmse.mle, colour = 'MLE-based', shape = 'MLE-based'),
             size = 3, alpha = .5) + 
  geom_errorbar(aes(x = rho, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                    colour = 'MLE-based'), width = .05) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', x = expression(rho[n]), colour = NULL, shape = NULL) + 
  scale_y_log10() + 
  scale_x_continuous(breaks = c(0.1, 0.3, 0.5, 0.7, 0.9))

gridExtra::grid.arrange(grobs = list(clustering.plot, rmse.plot), nrow = 1)
```

# Choice of $\vartheta$ for the Subspace Detection Property

### 2022-02-05 referee-response.Rmd
corrected typo

@@
 z_n$, none of our theoretical assume anything about the distribution
-of the $\{z_i\}$; we do, however, occassionally assume that the $n_k$
+of the $\{z_i\}$; we do, however, occasionally assume that the $n_k$
 are sufficiently large for each $k = 1,2,\dots, K$. More specifically 

### 2022-02-06 referee-response.Rmd
Premerge

--- 2022-02-04 referee-response.Rmd
@@
 that for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
-$x^{\top} I_{p,q} y \in [0,1]$. }
+$x^{\top} I_{p,q} y \in [0,1]$. One simple way to construct such a set
+$\mathcal{X}$ is to first find a collection of $K$ vectors
+$\mathcal{S} = \{\nu_1, \nu_2, \dots, \nu_K\} \subset \mathbb{R}^{d}$
+where $\nu_i^{\top} I_{p,q} \nu_j \in [0,1]$ for all $i,j$ and then define $\mathcal{X}$
+as 
+$$\mathcal{X} = \mathrm{conv}(\mathcal{S}) = \{x = \sum_{i} \lambda_i \nu_i \colon \lambda_i \geq 0 \,\, \text{for all $i$}, \sum_{i} \lambda_i =
+1\}.$$
+}
 
@@
 \textcolor{blue}{
-Remark 2 now includes that $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
-}
+We had rewrote Remark~2 to more clearly described the matrices
+$\hat{Z}$, $V$ and $D$. To avoid confusion, we now use $\hat{D}$ to
+denote the diagonal matrix whose diagonal entries are the $d = p + q$
+largest eigenvalues (in modulus) of $A$ and we use $\hat{V}$ to denote
+the $n \times d$ matrix whose columns are the corresponding
+eigenvectors. We then use $\hat{Z} = \hat{V} |\hat{D}|^{1/2}$ (where
+the $|\cdot|$ operation is applied elementwise) to denote the
+adjacency spectral embedding (ASE) of $A$ into $\mathbb{R}^{d}$; that is
+to say, the $i$th row of $\hat{Z}$ represents an estimate of the
+latent position $X_i$ (up to some non-identifiability transformation
+$Q \in \mathbb{O}(p,q)$). We also reorganized the paper slightly so
+that the definition and remarks about the non-identifiability of the
+latent positions $X$ in a GRDPG are presented before discussing ASE. }
 
@@
 \textcolor{blue}{
-The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case by starting with $K = 2$. 
-In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
-This then generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
-Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
-resulting in $K (K + 1) / 2$ $+1$'s and $K (K - 1) / 2$ $-1$'s, i.e., the diagonal eigenvalue matrix of $\Pi$ is exactly $I_{K (K + 1) / 2, K (K - 1) / 2}$. 
-This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
-We have added additional details to the paper to address this. 
+The purpose of the additional material in the proof of Theorem 1 is to
+first present the proof in the special case when $K = 2$ so as to
+build intuition for the general case of $K \geq 2$. 
+More specificially for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
+where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of
+order 2, and the latent vectors lie in the union of two $2$-dimensional
+orthogonal subspaces, i.e., and the rows of $X$ consist of
+vectors that lie in $\mathcal{S}_1 \cup \mathcal{S}_2$ where
+$\mathcal{S}_1$ and $\mathcal{S}_2$ are both $2$-dimensional
+subspaces of $\mathbb{R}^{4}$, $\mathcal{S}_1 \cap \mathcal{S}_2 = \{0\}$, and $x^{\top} y
+= 0$ for all $x \in \mathcal{S}_1, y \in \mathcal{S}_2$. 
+This then generalizes to larger $K$ such that the rows of $X$ are now
+vectors that lie in the union of $K$ orthogonal subspaces
+$\mathcal{S}_1 \cup \mathcal{S}_2 \cup \dots \mathcal{S}_K$ where each
+$\mathcal{S}_k$ is a $K$-dimensional subspace of $\mathbb{R}^{K^2}$,
+and $\Pi$ is a permutation matrix with 
+$K$ fixed points and $K (K - 1)$ cycles of order 2. The permutation
+matrix $\Pi$ then has $K(K-1)/2$ eigenvalues equal to $-1$ (due
+to the $K(K-1)$ cycles of order $2$) and $K(K+1)/2$ eigenvalues equal
+to $1$ (due to the $K(K-1)$ cycles of order $2$ together with the $K$
+fixed points). We can thus write $\Pi = U I_{K(K+1)/2,K(K-1)/2}
+U^{\top}$. We have revised the paper to more clearly present the above
+observations. See also Example~1 on page 9 of the revised manuscript
+and the response to comment $10$ below.
 }
@@
 \textcolor{blue}{
-The purpose of Example 1 is to illustrate this explicitly. 
-Comparing matrices $X$ and $Y$, we can see that they share three columns: the first, fifth, and ninth. 
-The other columns are permuted. 
-The proof of Theorem 1 then generalizes this to arbitrary $K$. 
+Example 1 illustrates these fixed points explicitly for $K = 3$. 
+More specifically, comparing the matrices $X$ and $Y$, we see that
+they have three columns in common: the first, fifth, and ninth. 
+The other columns are permuted. The proof of Theorem 1 then
+generalizes this to arbitrary $K$. In particular by the construction
+of $X$, each column of $X$ has exactly one non-zero block
+corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
+points are then the columns with indices $r(K+1) +1 $ for $0 \leq r
+\leq K-1$; these columns contain the vectors $\lambda_{kk}$ for
+$1 \leq k \leq K$. The cycles of order $2$ swap the columns of $X$
+containing the vectors $\lambda_{k \ell}$ and $\lambda_{\ell k}$ for
+the $\tbinom{K}{2}$ pairs $(k,\ell)$ with $k \not = \ell$. 
 }

### 2022-02-06 summary_minh_edited.tex
Premerge

@@
 More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
-then we also have for any $Q \in \mathbb{O}(p, q)$ that
+then for any $Q \in \mathbb{O}(p, q)$ we also have 
 $(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
@@
 |\hat{\lambda}_n|.$$
-Now let $\hat{D}$ be a diagonal matrix whose diagonal entries are
-the eigenvalues
-$\{\hat{\lambda}_i\}_{i=1}^{d}$ arranged in decreasing values (not in
-decreasing modulus). Let $\hat{V}$ be the $n \times d$ matrix whose
+Let $\hat{D}$ be a diagonal matrix whose diagonal entries are
+the eigenvalues $\{\hat{\lambda}_i\}_{i=1}^{d}$ arranged in decreasing values (not in
+decreasing modulus) and let $\hat{V}$ be the $n \times d$ matrix whose
 columns are the corresponding eigenvectors $\{\hat{v}_i\}_{i=1}^{d}$
@@
 orthogonal transformation $Q$ as described in
-Definition~\ref{def:indefinite} and
-Remark~\ref{rem:non_identifiable}. 
+Remark~\ref{rem:non_identifiable} and Definition~\ref{def:indefinite};
+see \citet{rubindelanchy2017statistical} for further details.
 % choosing the $p$ most positive and $q$ most negative
@@
 \begin{proof}
-We will prove this theorem in two parts. First, for demonstration purposes, we focus on the case for $K = 2$. Then we generalize this to $K \geq 2$.
+We will prove this theorem in two parts. First, for demonstration
+purposes, we focus on the case when $K = 2$ to build intuition. 
+The general case of $K \geq 2$ is presented later.
 
@@
 e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$. 
-Thus, another way of decomposing the edge probability matrix is $\tilde{P} = X \Pi X^\top$ where the rows of $X$ describe 2 2-dimensional orthogonal subspaces and $\Pi$ is a permutation matrix.
+Thus, another way of decomposing the edge probability matrix is
+$\tilde{P} = X \Pi X^\top$ where the rows of $X$ lie in the union of
+two 2-dimensional orthogonal subspaces and $\Pi$ is a permutation matrix.
 
@@
 $\Pi$ and hence $\tilde{P} = X \Pi X^\top$.
-The permutation described by $\Pi$ has $K$ fixed points, which correspond to
+The permutation described by $\Pi$ now has $K$ fixed points, which correspond to
 $K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where
-$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has
-$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to
+$k = r (K + 1) + 1$ for $0 \leq r \leq K - 1$. It also has
+$\tbinom{K}{2}$ cycles of order $2$. Each cycle corresponds to
 a pair of eigenvalues $\{-1,+1\}$ and a pair of eigenvectors

### 2022-02-06 referee-response.tex

@@
   \textcolor{blue}{
-  We changed the terminology a bit here and throughout the paper 
-  and specify that the analysis assumes $G$ is an unweighted and undirected graph with no self-loops (no edge directly from itself to itself), 
-  so $A$ is binary and symmetric with zeros on the diagonal. 
-  }
+  We changed the terminology a bit here and throughout the paper and
+  specify that the analysis assumes $G$ is an unweighted and undirected
+  graph with no self-loops 
+  (no edge directly from itself to itself), 
+  so $A$ is binary and symmetric with zeros on the diagonal. The phrase
+  ``hollow graph'' is no longer used.  }
 \item
@@
   \textcolor{blue}{
-  We have made minor edits clarifying the structure of $G$ and $A$. 
+  We have made minor edits clarifying the structure of $G$ and $A$. In
+  particular we assume throughout the paper that $G$ is undirected and $A$ is
+  symmetric. 
   }
@@
   \textcolor{blue}{
-  We added $[n] = \{1, 2, ..., n\}$. 
+  We added $[n] = \{1, 2, ..., n\}$ at the beginning of Section~2.1. 
   }
@@
   \textcolor{blue}{
-  The range of $(i, j)$ is now included in the definition. 
-  In regard to the probability distribution on the labels $z_1, ..., z_n$, none of our theorems use this, although some do require that each $n_k$ be sufficiently large, so it did not factor into our definitions and analyses.
-  More specifically, the convergence rate of Theorem 5 depends on each $n_k$, and for convergence, we require that $n_k \to \infty$. 
-  While none of the theorems explicitly require this, in practice, we also require that each $n_k$ is large enough for the latent vectors in the $k^th$ community to span its subspace. 
-  }
+  The range of $(i, j)$ is now included in the definition; in particular
+  we wrote in Section 2.1 that ``$1 \leq i < j \leq n$.''
+  In regard to the probability distribution on the labels $z_1, ...,
+  z_n$, none of our theoretical assume anything about the distribution
+  of the $\{z_i\}$; we do, however, occasionally assume that the $n_k$
+  are sufficiently large for each $k = 1,2,\dots, K$. More specifically 
+  the convergence rate of Theorem 5 depends on the $n_k$, i.e., for
+  convergence we require that $n_k \to \infty$ as $n \rightarrow
+  \infty$. In practice, we also require that each $n_k$ is large enough
+  so that the latent vectors in the $k$th community span its
+  subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lies in a
+  general position (are linearly independent), then we only require $n_k
+  \geq K$.}
 \item
@@
   \textcolor{blue}{
-  We now define the subset as $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p, q} y \in [0, 1]\}$ with $d = p + q$. 
-  }
+  We now denote $d = p + q$ in Definition 2 and define $\mathcal{X}$ as a subset of $\mathbb{R}^{d}$ such
+  that for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
+  $x^{\top} I_{p,q} y \in [0,1]$. }
 \item

### 2022-02-06 sparsity-sim/supplemental-materials.tex

--- 2021-09-01 summary.tex
@@
 %
-\documentclass[12pt]{article}
+\documentclass[
+  12pt,
+]{article}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
 \usepackage{ifxetex,ifluatex}
@@
 \hypersetup{
-  pdftitle={Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation},
+  pdftitle={Supplemental Materials for ``Popularity Adjusted Block Models are Generalized Random Dot Product Graphs''},
   colorlinks=true,
@@
 \urlstyle{same} % disable monospaced font for URLs
-%\usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
-
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
 \makeatletter
@@
 \usepackage{mathtools}
-% \usepackage{natbib}
-\usepackage[backend=bibtex,maxbibnames = 10,style=authoryear,giveninits=true,natbib]{biblatex}
-\DeclareNameAlias{sortname}{last-first}
-\addbibresource{misc.bib}
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-\usepackage{parskip}
+\usepackage{natbib}
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-% \setcitestyle{numbers,square,comma}
+\setcitestyle{numbers,square,comma}
 \usepackage{verbatim}
@@
 \usepackage{comment}
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 
-\title{Popularity Adjusted Block Models are Generalized Random Dot Product Graphs}
-
-\author{John Koo\\Department of Statistics, Indiana University\\
-Minh Tang\\Department of Statistics, North Carolina State University\\
-Michael Trosset\\Department of Statistics, Indiana University}
-
-% \date{\vspace{-2.5em}}
-\date{}
+\title{Supplemental Materials for ``Popularity Adjusted Block Models are
+Generalized Random Dot Product Graphs''}
+\author{}
+\date{\vspace{-2.5em}}
 
 \begin{document}
-
 \maketitle
 
-\begin{abstract}
-We connect two random graph models, the Popularity Adjusted Block Model
-(PABM) and the Generalized Random Dot Product Graph (GRDPG),
-by demonstrating that a PABM is a GRDPG in which communities correspond to 
-mutually orthogonal subspaces of latent vectors. 
-This insight allows us to construct new algorithms for community detection
-and parameter estimation for the PABM, 
-as well as improve an existing algorithm 
-that relies on Sparse Subspace Clustering. 
-Using established asymptotic properties of 
-Adjacency Spectral Embedding for the GRDPG, 
-we derive asymptotic properties of these algorithms. 
-In particular, we demonstrate that the absolute number of 
-community detection errors tends to zero as 
-the number of graph vertices tends to infinity. 
-Simulation experiments illustrate these properties. 
-\end{abstract}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \newcommand{\diag}{\mathrm{diag}}
@@
 \newcommand{\Multinomial}{\mathrm{Multinomial}}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{proposition}{Proposition}
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-\newpage
-\spacingset{1.5} % DON'T change the spacing!
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\hypertarget{sparsity-simulation-study}{%
+\section{Sparsity Simulation Study}\label{sparsity-simulation-study}}
 
-Statistical inference on random graphs requires a suitable probability model. 
-A general probability model for unweighted and undirected graphs is the Bernoulli Graph (also known as the inhomogeneous Erd\"{o}s-R\'{e}nyi model), which assumes that edges occur as independent Bernoulli trials.
-A Bernoulli Graph is characterized by an edge probability matrix $P=[P_{ij}]$, where an edge between vertices $i$ and $j$ occurs with success probability $P_{ij}$.
-A trivial example of a Bernoulli Graph is the (homogeneous) Erd\"{o}s-R\'{e}nyi model proposed by \citet{Gilbert:1959}, in which the vertices of the random graph are fixed and possible edges occur independently with fixed probability $P_{ij}=p$.
-The requirement that $p$ is fixed is too strong for most applications, and various researchers have weakened that requirement in various ways.
-The present work relates two lines of generalization.
-
-Network analysis is often concerned with community detection.
-One form of community detection assumes that each vertex belongs to an unobserved community, with the probability of an edge between vertices $i$ and $j$ depending on the communities to which $i$ and $j$ belong.
-Formally, one assigns each vertex $v_i$ a community label $z_i$ and assumes a Bernoulli Graph in which $P_{ij}$ is a function of $z_i$ and $z_j$.
-Such models, called Block Models, define the goal of community detection as a problem in statistical inference: identify the true community (up to permutation of labels) to which each vertex belongs.
-
-The classical Stochastic Block Model (SBM) of \citet{doi:10.1080/0022250X.1971.9989788} specifies that each edge probability $P_{ij}$ depends only on the labels $z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_,z_j}$.
-Subsequent researchers have weakened this assumption.
-The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_,z_j}$.
-The Popularity Adjusted Block Model (PABM) of \citet{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
-
-Another type of Bernoulli Graph was proposed by
-\citet*{10.1007/978-3-540-77004-6_11}.  A Random Dot Product Graph
-(RDPG) specifies that each vertex corresponds to a latent position
-vector in Euclidean space and that the probability of an edge between
-two vertices is the dot product of their latent position vectors.
-Thus, if the latent positions are $x_1,\ldots,x_n \in \mathbb{R}^d$
-and $X = [\begin{array}{ccc} x_1 & \cdots & x_n \end{array} ]^\top$,
-then the edge probability matrix is $P = XX^\top$.  Clearly, any
-Bernoulli Graph with positive semidefinite $P$ is an RDPG.  The
-positive definite Euclidean inner product in the RDPG model was
-replaced by an indefinite inner product in
-\citet{rubindelanchy2017statistical}, resulting in the {\em
-Generalized}\/ RDPG (GRDPG).
-
-In contrast to Block Models, neither RDPGs nor GRDPGs inherently
-specify distinct communities.  However, one can easily impose
-community structure by assuming that the latent positions lie in
-distinct clusters.  Hence, it is not surprising that Block Models can
-be studied by reformulating them as RDPGs or GRDPGs.  For example, an
-assortative SBM (an SBM for which $P$ is positive semidefinite) is
-equivalent to an RDPG for which all vertices in the same community
-correspond to the same latent position vector.
-Likewise, the DCBM is equivalent to an RDPG for which all vertices in
-the same community correspond to latent position vectors that lie on a
-straight line. %(\cite{lyzinski2014, rubindelanchy2017statistical}).
-
-Because the edge probability matrix of a PABM is not necessarily
-positive semidefinite, a PABM is not necessarily an RDPG.  In Section
-\ref{connecting-the-pabm-to-the-grdpg} we demonstrate that every PABM
-is in fact a specific type of GRDPG for which the latent position
-vectors lie in distinct orthogonal subspaces, each subspace
-corresponding to a community.  This identification is our central
-result.  In Section \ref{methods}, we use the geometry of the GRDPG to
-derive more efficient algorithms for detecting the communities and
-estimating the parameters in the PABM.  We report the results of
-simulation studies in Section \ref{simulated-examples} and apply our
-methods to three well-known data sets in Section
-\ref{real-data-examples}. Section \ref{discussion} concludes.
-
-\hypertarget{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}{%
-\section{PABMs are GRDPGs}\label{connecting-the-popularity-adjusted-block-model-to-the-generalized-random-dot-product-graph}}
-
-In this section, we show that the PABM is a special case of the GRDPG. 
-More specifically, a graph $G$ drawn from the PABM can be represented by 
-a collection of latent vectors in Euclidean space. 
-We further show that the latent configuration that
-induces the PABM consists of orthogonal subspaces with each subspace
-corresponding to a community.
-
-\hypertarget{notation}{%
-\subsection{Notation and Scope}\label{notation}}
-
-Let $G = (V, E)$ be an unweighted, undirected, 
-and hollow graph with vertex set $V$ ($|V| = n$) and edge set $E$. The matrix
-$A \in \{0, 1\}^{n \times n}$ represents the adjacency matrix of $G$ 
-such that $A_{ij} = 1$ if there exists an edge between vertices $i$ and $j$ 
-and $0$ otherwise. Because $G$ is symmetric and hollow,
-$A_{ij} = A_{ji}$ and $A_{ii} = 0$ for each $i, j \in [n]$. 
-We further restrict our analyses to Bernoulli graphs. 
-Let $P \in [0, 1]^{n \times n}$ be a symmetric matrix of edge probabilities. 
-Graph $G$ is sampled from $P$ by drawing $A_{ij} \indep \Bernoulli(P_{ij})$ 
-for each $1 \leq i < j \leq n$ (setting $A_{ji} = A_{ij}$ and $A_{ii} = 0$). 
-We denote $A \sim \BG(P)$ as a graph with adjacency matrix $A$ 
-sampled from edge probability matrix $P$ in this manner. 
-If each vertex has a hidden label in $[K]$, 
-they are denoted as $z_1, ..., z_n$. 
-Finally, we denote 
-$X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^\top \in \mathbb{R}^{n \times d}$ 
-as the matrix corresponding to a collection of $n$ latent vectors $x_1, ..., x_n \in \mathbb{R}^d$.
-
-\hypertarget{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}{%
-\subsection{Two Probability Models for Graphs}\label{the-popularity-adjusted-block-model-and-the-generalized-random-dot-product-graph}}
-
-\begin{definition}[Popularity Adjusted Block Model]
-\label{pabm}
-Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a
-graph $G = (V, E)$ with adjacency matrix $A$ 
-such that $A \sim \mathrm{Bernoulli}(P)$.
-Let each vertex have a community label between $1$ and $K$.
-Then $G$ is drawn from a Popularity Adjusted Block Model if
-each vertex has $K$ popularity parameters that describe its affinity
-toward each of the $K$ communities, i.e., vertex $i$ has popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK}$, and  $P_{ij} = \lambda_{i z_j}
-\lambda_{j z_i}$ for all $i \leq j$.
-
-Another characterization of the PABM is as follows.
-Let the rows and columns of $P$ be arranged by community label
-such that $n_k \times n_{\ell}$ block $P^{(k \ell)}$
-describes the edge probabilities between vertices in communities
-$k$ and $\ell$; note that $P^{(k \ell)} = (P^{(k \ell)})^\top$.
-Now suppose that each block $P^{(k \ell)}$ can be written as the outer product of two vectors:
-\begin{equation} \label{eq:pabm}
-  P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}
-\end{equation}
-for a set of $K^2$ popularity vectors 
-$\{\lambda^{(st)}\}_{1 \leq s \leq K, 1 \leq t \leq K}$ where each
-$\lambda^{(st)}$ is a column vector of dimension $n_s$. Then a graph $G$
-is drawn from a PABM with parameters
-$\{\lambda^{(st)}\}$ if its adjacency matrix $A$
-satisfies $A \sim \mathrm{BernoulliGraph}(P)$.
-\end{definition}
-
-We will use the notation \(A \sim \PABM(\{\lambda^{(k \ell)}\}_K)\) to denote
-a random adjacency matrix \(A\) drawn from a PABM with parameters
-\(\{\lambda^{(k \ell)}\}\) consisting of \(K\) underlying communities.
-
-\begin{definition}[Generalized Random Dot Product Graph]
-\label{grdpg}
-Let graph $G = (V, E)$ be drawn as $A \sim \BG(P)$ for some symmetric $P \in
-[0,1]^{n \times n}$. 
-If there exists a data matrix $X \in \mathbb{R}^{n \times d}$ such that
-\begin{equation} \label{eq:grdpg}
-  P = X I_{p,q} X^\top
-\end{equation}
-for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then
-$G$ is drawn from the Generalized Random Dot Product Graph 
-with latent positions $x_1, ..., x_n \in \mathbb{R}^d$ and signature $(p, q)$.
-\end{definition}
-
-We will use the notation \(A \sim \GRDPG_{p,q}(X)\) to denote a random
-adjacency matrix \(A\) drawn from latent positions \(X\) and signature
-\((p, q)\). If instead of fixed latent positions, 
-they are drawn $X_1, ..., X_n \iid F$, 
-we denote the GRDPG as $(A, X) \sim \GRDPG_{p,q}(F, n)$.
-
-\begin{remark}
-We can use Adjacency Spectral Embedding (ASE) 
-\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
-This procedure consists of taking the spectral decomposition of $A$ 
-(or $P$ if available) and choosing the $p$ most positive and $q$ most negative
-eigenvalues and their corresponding eigenvectors 
-to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
-\end{remark}
-
-\begin{definition}[Indefinite Orthogonal Group]
-The indefinite orthogonal group with signature $(p, q)$ is
-the set $\{Q \in \mathbb{R}^{d \times d} \colon Q I_{p, q} Q^{\top} = I_{p, q}\}$,
-denoted as $\mathbb{O}(p, q)$. Here $d = p + q$. 
-\end{definition}
-
-\begin{remark}
-The latent vectors that produce $X I_{p,q} X^\top = P$ are not unique
-\citep{rubindelanchy2017statistical}.
-More specifically, if $P_{ij} = x_i^\top I_{p, q} x_j$, 
-then we also have for any $Q \in \mathbb{O}(p, q)$ that
-$(Q x_i)^\top I_{p, q} (Q x_j) = x_i^\top (Q^\top I_{p, q} Q) x_j =
-x_i^\top I_{p, q} x_j = P_{ij}$.
-Unlike in the RDPG case, transforming the latent positions via multiplication
-by $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or
-distances.
-\end{remark}
-
-\hypertarget{connecting-the-pabm-to-the-grdpg}{%
-\subsection{The Geometry of PABMs}\label{connecting-the-pabm-to-the-grdpg}}
-
-Now that we defined the PABM and GRDPG, 
-we show the special geometry of the PABM when viewed as a GRDPG.
-
-\begin{theorem}[The latent configuration of the PABM]
-\label{theorem2}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K)$ be an instance of a
-PABM with $K \geq 1$ blocks and latent vectors $\{\lambda^{(k \ell)}
-\colon 1 \leq k \leq K, 1 \leq \ell \leq K\}$. 
-Then there exists a block diagonal matrix
-$X \in \mathbb{R}^{n \times K^2}$ defined by $\{\lambda^{(k \ell)}\}$ and a 
-$K^2 \times K^2$ {\em fixed} orthonormal matrix $U$ such 
-that $A \sim \mathrm{GRDPG}_{K (K+1) / 2, K (K-1) / 2}(XU)$.
-\end{theorem}
-
-\begin{proof}
-We will prove this theorem in two parts. First, for demonstration purposes, we focus on the case for $K = 2$. Then we generalize this to $K \geq 2$.
-
-For the $K = 2$ case, the proof is straightforward. Let 
-$$X = \begin{bmatrix}
-\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
-0 & 0 & \lambda^{(21)} & \lambda^{(22)}
-\end{bmatrix} \quad \text{and} \quad
-U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
-0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
-0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
-0 & 1 & 0 & 0 \end{bmatrix}.$$
-Then by straightforward matrix multiplication, we obtain 
-\[X U I_{3, 1} U^\top X^\top =
-\begin{bmatrix}
-  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
-  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
-\end{bmatrix}, \]
-which is a GRDPG with latent vectors described by $X U$.
-
-It is nevertheless instructive to look at a few intermediate steps. 
-More specifically, the product $U I_{3, 1} U^\top$ 
-yields a permutation matrix $\Pi$ with fixed points at positions $1$ and $4$ 
-and a cycle of order 2 swapping positions $2$ and $3$, i.e., 
-$$\Pi = U I_{3, 1} U^\top = \begin{bmatrix} 1 & 0 & 0 & 0 \\
-  0 & 0 & 1 & 0 \\
-  0 & 1 & 0 & 0 \\
-  0 & 0 & 0 & 1
-\end{bmatrix}.$$
-Furthermore, since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $U I_{3, 1} U^\top$
-is also an eigendecomposition of $\Pi$ where the fixed
-points of $\Pi$ are mapped to the eigenvectors $e_1$ and $e_4$
-while the cycles of order two are mapped to the eigenvectors  
-$\tfrac{1}{\sqrt{2}}(e_{2} + e_3)$ and $\tfrac{1}{\sqrt{2}}(e_{2} -
-e_3)$; here $e_i$ denote the $i^\mathrm{th}$ basis vector in $\mathbb{R}^{4}$.
-
-For the general case, we first define the following matrices
-\begin{gather}
-\label{eq:xy}
-\Lambda^{(k)} = \begin{bmatrix} \lambda^{(k1)} \mid \cdots \mid \lambda^{(kK)} \end{bmatrix}
-\in \mathbb{R}^{n_k \times K}, \quad
-X = \text{blockdiag}(\Lambda^{(1)}, \dots, \Lambda^{(K)}) \in
-\mathbb{R}^{n \times K^2}, \\
-L^{(k)} = \text{blockdiag}(\lambda^{(1k)}, \dots, \lambda^{(Kk)}) \in
-\mathbb{R}^{n \times K}, \quad
-Y = \begin{bmatrix} L^{(1)} \mid \cdots \mid L^{(K)} \end{bmatrix} \in
-\mathbb{R}^{n \times K^2}.
-\end{gather}
-It is then straighforward to verify that
-\begin{gather*}
-  XY^{\top} = \text{blockdiag}(\Lambda^{(1)}, \dots,
-\Lambda^{(K)}) \begin{bmatrix} L_1^{\top} \\ \vdots \\
-  L_{K}^{\top} \end{bmatrix} = \begin{bmatrix} \Lambda^{(1)}
-  (L^{(1)})^{\top} \\ \vdots \\
-  \Lambda^{(K)} (L^{(K)})^{\top} \end{bmatrix}, \\
-\Lambda^{(k)} (L^{(k)})^{\top} = \begin{bmatrix} \lambda^{(k1)}
-  (\lambda^{(1k)})^{\top} \mid \dots \mid \lambda^{(kK)}
-  (\lambda^{(Kk)})^{\top} \end{bmatrix} = \begin{bmatrix} P^{(k1)}
-  \mid P^{(k2)} \mid \dots \mid P^{(kK)} \end{bmatrix}.
-\end{gather*}
-We therefore have $P = X Y^\top$. 
-Similar to the $K = 2$ case, we also have $Y = X \Pi$ for some permutation matrix
-$\Pi$ and hence $P = X \Pi X^\top$.
-The permutation described by $\Pi$ has $K$ fixed points, which correspond to
-$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where
-$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has
-$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to
-a pair of eigenvalues $\{-1,+1\}$ and a pair of eigenvectors
-$\{(e_s + e_t)/\sqrt{2},(e_s - e_t)/ \sqrt{2}\}$.
-
-Let $p = K (K + 1) / 2$ and $q = K (K - 1) / 2$. 
-We therefore have
-\begin{equation} \label{eq:permutation}
-\Pi = U I_{p,q} U^\top
-\end{equation}
-where $U$ is a $K^2 \times K^2$ orthogonal matrix and hence
-\begin{equation} \label{eq:pabm-grdpg}
-P = X U I_{p, q} (X U)^\top.
-\end{equation}
-In summary we can describe the PABM with $K$ communities as a GRDPG with latent
-positions $X U$ and signature $(p,q) = \bigl( \tfrac{1}{2} K (K + 1) ,
-\tfrac{1}{2} K (K - 1)\bigr)$.
-\end{proof}
-
-\begin{example} Let $A$ be a $3$ blocks PABM with latent vectors
-  $\{\lambda^{(k \ell)} \colon 1 \leq k \leq 3, 1 \leq \ell \leq 3\}$. Using the same notation as in Theorem
-  \ref{theorem2}, we can define
-\begin{gather*}
-X = \begin{bmatrix}
-\lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
-0 & 0 & 0 & \lambda^{(21)} & \lambda^{(22)} & \lambda^{(23)} & 0 & 0 & 0 \\
-0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
-\end{bmatrix}, \\
-Y = \begin{bmatrix}
-\lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
-0 & \lambda^{(21)} & 0 & 0 & \lambda^{(22)} & 0 & 0 & \lambda^{(23)} & 0 \\
-0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
-\end{bmatrix}.
-\end{gather*}
-Then $Y = X \Pi$ and $P = X Y^{\top}$ where $\Pi$ is a $9 \times 9$ 
-permutation matrix of the form
-$$\Pi = 
-\Bigl[e_1 \mid e_4 \mid e_7 \mid e_2 \mid e_5 \mid e_8 \mid e_3
-\mid e_6 \mid e_9 \Bigr].$$
-where $e_i$ denotes the $i^{th}$ basis vector in $\mathbb{R}^{9}$. 
-The matrix $\Pi$ corresponds to a permutation of $\{1,2,\dots,9\}$
-with the following decomposition.
-\begin{enumerate}
-\item Positions 1, 5, 9 are fixed.
-\item There are three cycles of length 2, namely $(2, 4)$, $(3, 7)$, and $(6, 8)$.
-\end{enumerate}
-We can therefore write $\Pi$ as $\Pi = U I_{6, 3} U^\top$ where the first three
-columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the
-{\em fixed} points, the next three columns consist of eigenvectors
-$(e_k + e_{\ell}) / \sqrt{2}$, and the last three columns consist of eigenvectors
-$(e_k - e_{\ell}) / \sqrt{2}$ for $(k, \ell) \in
-\{(2,4),(3,7),(6,8)\}$.
-
-The matrix $P$ is then the edge probabilities matrix for a 
-Generalized Random Dot Product Graph whose latent positions 
-are the rows of the matrix
-$$XU = \begin{bmatrix}
-  \lambda^{(11)} & 0 & 0 &
-  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 &
-  \frac{\lambda^{(12)}}{\sqrt{2}} & \frac{\lambda^{(13)}}{\sqrt{2}} & 0 \\
-  0 & \lambda^{(22)} & 0 &
-  \frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} &
-  -\frac{\lambda^{(21)}}{\sqrt{2}} & 0 & \frac{\lambda^{(23)}}{\sqrt{2}} \\
-  0 & 0 & \lambda^{(33)} &
-  0 & \frac{\lambda^{(31)}}{\sqrt{2}} & \frac{\lambda^{(32)}}{\sqrt{2}} &
-  0 & -\frac{\lambda^{(31)}}{\sqrt{2}} & -\frac{\lambda^{(32)}}{\sqrt{2}}
-\end{bmatrix}.$$
-\end{example}
-
-\hypertarget{methods}{%
-\section{Algorithms}\label{methods}}
-
-Two inference objectives arise from the PABM:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Community membership identification (up to permutation).
-\item
-  Parameter estimation (estimating \(\lambda^{(k \ell)}\)'s).
-\end{enumerate}
-
-In our methods, we assume that \(K\), the number of communities, is
-known beforehand and does not require estimation.
-
-\hypertarget{related-work}{%
-\subsection{Previous Work}\label{related-work}}
-
-\citet{307cbeb9b1be48299388437423d94bf1} 
-used Modularity Maximization (MM) and the Extreme Points (EP)
-algorithm (\cite{le2016}) for community detection and parameter
-estimation. They were able to show that as the sample size increases,
-the {\em proportion} of misclassified community labels (up to permutation)
-goes to 0.
-
-\citet{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
-(\cite{5206547}) for community detection in the PABM. 
-The SSC algorithm can be described as follows: 
-Given \(X \in \mathbb{R}^{n \times d}\) with vectors
-\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
-\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X c\) and
-\(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
-collected collected into matrix
-\(C = \bigl[ c_1 \mid c_2 \mid \cdots \mid c_n \bigr]^\top\) to
-construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
-exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
-consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
-\(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
-points near \(K\) subspaces with some noise, a final graph partitioning
-step may be required (e.g., edge thresholding or spectral clustering).
-
-In practice, due to presence of noise, SSC is often done by solving the LASSO problems
-\begin{equation} \label{eq:ssc}
-c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i} c\|^2_2 + \lambda \|c\|_1
-\end{equation}
-for some sparsity parameter \(\lambda > 0\). The \(c_i\) vectors are
-then collected into \(C\) and \(B\) as before.
-
-\begin{definition}[Subspace Detection Property]
-  \label{def:subspace_detection}
-Let $X = \bigl[ x_1 \mid x_2 \mid \cdots \mid x_n \bigr]^\top$ be noisy
-points sampled from $K$ subspaces, i.e., $x_i = y_i + z_i$ where the
-$y_i$ belongs to the union of $K$ subspaces and the $z_i$ are noise
-vectors. Let $\lambda \geq 0$ be given and
-let $C$ and $B$ be constructed from the
-solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
-this given choice of $\lambda$. Then $X$
-is said to satisfy the subspace detection property 
-with sparsity parameter $\lambda$ if each column of
-$C$ has nonzero $\ell_2$ norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
-from different subspaces.
-\end{definition}
-
-\begin{remark}
-In practice, a noisy sample $X$ often does not obey the subspace detection
-property. In such cases, $B$ is treated as an affinity matrix for a graph, 
-which requires a partitioning step to obtain the final clustering. On the other
-hand, if $X$ does obey the subspace detection property, $B$ describes a graph
-with at least $K$ disconnected subgraphs. Ideally, when the subspace detection
-property holds, there are exactly $K$ subgraphs which map onto each subspace,
-but it could be the case that some of the subspaces are represented by
-multiple disconnected subgraphs. The subspace detection property is contingent
-on choosing a sufficiently large sparsity parameter $\lambda$.
-\end{remark}
-
-Theorem \ref{theorem2} suggests that SSC is appropriate for community
-detection for the PABM. More precisely, Theorem \ref{theorem2} says that
-each community consists of a \(K\)-dimensional subspace, and together
-the subspaces lie in \(\mathbb{R}^{K^2}\). The natural approach then is
-to perform SSC on the ASE of \(P\) or \(A\).
-\citet{noroozi2019estimation} instead applied SSC to \(P\) and
-\(A\), foregoing embedding altogether.
-
-Using results from \citet{soltanolkotabi2012}, it can be easily shown
-that the subspace detection property holds for \(XU\), which is an ASE
-of \(P\). More specifically, if points lie exactly on mutually
-orthogonal subspaces, then the subspace detection property will hold
-with probability 1, and this is exactly the case for the PABM (Theorem
-\ref{theorem2}). Much of our work is then built on
-\citet{rubindelanchy2017statistical}, who describe the convergence
-behavior of the ASE of \(A\) to the ASE of \(P\), and
-\citet{jmlr-v28-wang13}, who show the necessary conditions for the
-subspace detection property to hold in noisy cases where the points lie
-near subspaces.
-
-\hypertarget{community-detection}{%
-\subsection{Algorithms for Community Detection}\label{community-detection}}
-
-\begin{algorithm}[t]
-  \label{alg:osc}
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \KwData{Adjacency matrix $A$, number of communities $K$}
-  \KwResult{Community assignments $1, ..., K$}
-    Compute the eigenvectors of $A$ that correspond to the $K (K+1) / 2$ most
-    positive eigenvalues and $K (K-1) / 2$ most negative eigenvalues. Construct
-    $V$ using these eigenvectors as its columns.\;
-    Compute $B = |n V V^\top|$, applying $|\cdot|$ entry-wise.\;
-    Construct graph $G$ using $B$ as its similarity matrix.\;
-    Partition $G$ into $K$ disconnected subgraphs
-    (e.g., using edge thresholding or spectral clustering).\;
-    Map each partition to the community labels $1, ..., K$.\;
-  \caption{Orthogonal Spectral Clustering.}
-\end{algorithm}
-
-We previously stated one possible set of latent positions that result in
-the edge probability matrix of a PABM, \(P = (XU) I_{p, q} (XU)^\top\). If
-we have (or can estimate) \(XU\) directly, then both the community
-detection and parameter identification problem are trivial since \(U\)
-is orthonormal and fixed for each value of \(K\). However, direct
-identification or estimation of \(XU\) is not possible
-(\cite{rubindelanchy2017statistical}).
-
-More specifically, if we decompose \(P = Z I_{p, q} Z^\top\), then
-there exists a \(Q \in \mathbb{O}(p, q)\) such that \(XU = Z Q\). Therefore, even if we
-start with the true but unknown edge probability matrix $P$ we
-will not be able to recover the ``original'' latent positions \(XU\).
-Note that unlike in the case of the RDPG, \(Q\) is not necessarily an
-orthogonal matrix and hence if \(z_i\)'s
-are the rows of \(XU\) then
-\(\|z_i - z_j\|^2 \neq \|Q z_i - Q z_j\|^2\), and
-\(\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle\). This
-prevents us from using the orthogonality property of \(XU\) directly.
-
-The explicit form of \(XU\) represents points in \(\mathbb{R}^{K^2}\)
-such that points within each community lie on \(K\)-dimensional
-orthogonal subspaces. Multiplication by \(Q \in \mathbb{O}(p, q)\)
-removes the orthogonality property but retains the property that each
-community is represented by a \(K\)-dimensional subspace. Therefore, the
-ASE of \(P\) still consists of subspaces 
-that correspond to communities of the PABM, even if they are not orthogonal, 
-suggesting the use of SSC. Before exploring SSC, we will first consider
-a different approach.
-
-\begin{theorem}
-\label{theorem3}
-Let $P = V D V^\top$ be the spectral decomposition of the edge probability
-matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are
-from different communities.
-\end{theorem}
-\begin{proof}
-% \begin{lemma}
-% \label{lemma1}
-% Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
-% matrix for a PABM.
-We first show that $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
-defined as in Eq.~\eqref{eq:xy}. Indeed, by Theorem 2, 
-\(P = X U I_{p, q} U^\top X^\top\) for $p = K(K+1)/2$ and $q = K(K-1)/2$.
-
-The eigendecomposition \(P = V D V^\top\) also yields $P = V
-|D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ where \(|\cdot|^{1/2}\) is
-applied entry-wise. Now let $Y = XU$ and $Z = V|D|^{1/2}$; note that
-$Y$ and $Z$ both have full column ranks. 
-Because $P = Y I_{p,q} Y^{\top} = Z I_{p,q} Z^{\top}$, we have
-$$Y = Z I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}.$$
-Let $Q = I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q}$ and note that
-$Y = Z Q$. We then have
-\begin{equation*}
-  \begin{split}
-  Q^{\top} I_{p,q} Q &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Z I_{p,q}
-I_{p,q} I_{p,q} Z^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} \\ &= I_{p,q} (Y^{\top} Y)^{-1} Y^{\top} Y I_{p,q}
-Y^{\top} Y (Y^{\top} Y)^{-1} I_{p,q} =  I_{p,q}
-\end{split}
-\end{equation*}
-and hence $Q$ is an indefinite orthogonal matrix.
-
-Let $R = U Q |D|^{-1/2}$ and note that $V = XR$. Because $R$ is
-invertible, we can write
-
-$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
-R^{\top} X^{\top}.$$
-
-Furthermore, as $V$ has orthonormal columns, $R^{\top} X^{\top} X R =
-V^{\top} V = I$. We thus conclude
-$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
-as desired.
-
-To complete the proof of Theorem~\ref{theorem3}, recall that \(X\) is block diagonal with each block
-corresponding to one community and hence \(X (X^\top X)^{-1} X^\top\) is also a
-block diagonal matrix with each block corresponding to a community. 
-As $B = n VV^{\top} = n X (X^\top X)^{-1} X^\top$, 
-we conclude that $B_{ij} = 0$ 
-whenever vertices $i$ and $j$ belong to different communities.  
-\end{proof}
-
-Theorem \ref{theorem3} provides perfect community detection given \(P\).
-Letting \(|B|\) be the affinity matrix for graph \(G'\), \(G'\) is
-partitioned into at least \(K\) disjoint subgraphs because each of the
-\(K\) communities have no edges between them. Similar to the subspace
-detection property, it could be the case that some of the communities
-are represented by multiple disjoint subgraphs in \(G'\), in which case
-additional reconstruction is required to identify the communities
-exactly.
-
-Using \(A\) instead of \(P\) to estimate $B$ introduces error, 
-which converges to \(0\) almost surely. 
-This is shown via properties of the GRDPG and ASE. 
-We use this fact to motivate Orthogonal Spectral Clustering (Alg. 1). 
-In order to show that OSC is consistent, 
-we introduce the sparsity factor $\rho_n \in (0, 1]$, 
-which can be characterized as the average expected degree of the vertices 
-divided by the number of vertices. 
-The treatment is the same as in \citet{307cbeb9b1be48299388437423d94bf1} (PABM) 
-and \citet{rubindelanchy2017statistical} (GRDPG) in which the popularity vectors 
-are scaled as $\rho_n^{1/2} \lambda^{(k \ell)}$. 
-This results in the edge probability matrix being scaled as 
-$P_{ij} = \rho_n \lambda_{i z_j} \lambda_{j z_i}$. 
-Since $\rho_n$ is unidentifiable, we use the convention that the popularity parameters are drawn such that $\|\Lambda\|_F = O(\sqrt{n})$ and require that the $L^2$ norm of each row is bounded away from 0 and 1.
-
-\begin{theorem}
-\label{theorem4}
-Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC
-(Alg. 1). Then for all pairs $(i, j)$ belonging to different communities
-and sparsity factor satisfying $n \rho_n = \omega(\log^{4}{n})$, we have
-\begin{equation} \label{eq:thm4}
-\max_{i, j} |n \hat{v}_i^\top \hat{v}_j| =
-O\Big( \frac{\log n}{\sqrt{n \rho_n}} \Big)
-\end{equation}
-with high probability.
-
-Thus for all pairs $(i, j)$ in different communities,
-we have uniform convergence $\hat{B}_{ij} \stackrel{a.s.}{\to} 0$. 
-\end{theorem}
-
-Theorem \ref{theorem4} provides an algorithm for community detection 
-that results in zero community detection error as $n \to \infty$. 
-For example, if we use edge thresholding to partition $\hat{B}$, 
-Theorem \ref{theorem4} implies that for any $\epsilon > 0$, 
-the number of edges of $\hat{B}$ between vertices of different communities 
-that are larger than $\epsilon$ converges to zero with probability 1.
-
-Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} also provide
-a natural path toward using SSC for community detection. 
-We established in Theorem \ref{theorem2} that an ASE of the edge
-probability matrix \(P\) can be constructed from a latent vector configuration 
-consisting of orthogonal subspaces. Theorem \ref{theorem3} shows how 
-this property can be recovered from the eigenvectors of \(P\). 
-Then Theorem \ref{theorem4} shows that replacing $P$ with $A$ 
-approximates these properties with error that goes to zero as \(n \to \infty\). 
-This allows us to obtain the subspace detection property 
-in Theorem \ref{theorem5}.
-
-\begin{algorithm}[t]
-  \label{alg:ssc}
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{Sparse Subspace Clustering using LASSO.}
-  \KwData{Adjacency matrix $A$, number of communities $K$,
-  hyperparameter $\lambda$}
-  \KwResult{Community assignments $1, ..., K$}
-    Find $V$, the matrix of eigenvectors of $A$
-    corresponding to the $K (K + 1) / 2$ most positive
-    and the $K (K - 1) / 2$ most negative eigenvalues.\;
-    Normalize $V \leftarrow \sqrt{n} V$.\;
-    \For {$i = 1, ..., n$} {
-      Assign $v_i^\top$ as the $i^{th}$ row of $V$.
-      Assign $V_{-i} = \bigl[
-      v_1 \mid \cdots \mid v_{i-1} \mid v_{i+1} \mid \cdots \mid v_n \bigr]^\top$.\;
-      Solve the LASSO problem
-      $c_i = \arg\min_{\beta}
-      \frac{1}{2} \|v_i - V_{-i} \beta\|_2^2 + \lambda \|\beta\|_1$.\;
-      Assign $\tilde{c}_i = (c_i^{(1)}, \dots, c_i^{(i-1)}, 0, c_i^{(i)}, \dots, c_i^{(n-1)})^\top$ such that the superscript is the index of
-      $\tilde{c}_i$.\;
-    }
-    Assign
-    $C = \bigr[ \tilde{c}_1 \mid \cdots \mid \tilde{c}_n \bigr]$.\;
-    Compute the affinity matrix $B = |C| + |C^\top|$.\;
-    Construct graph $G$ using $B$ as its similarity matrix.\;
-    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge
-    thresholding or spectral clustering).\;
-    Map each partition to the community labels $1, ..., K$.
-\end{algorithm}
-
-\begin{theorem}
-\label{theorem5}
-Let $P$ describe the edge probability matrix of the PABM with
-$n$ vertices, and let $A \sim \Bernoulli(P)$.  Let $\hat{V}$ be the
-matrix of eigenvectors of $A$ corresponding to the $K^2$ largest
-eigenvalues in modulus. Then for any $\epsilon > 0$ 
-there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
-that for all $n \geq N$, $\sqrt{n} \hat{V}$ obeys the subspace detection property with
-probability at least $1 - \epsilon$.  
-\end{theorem}
-
-\hypertarget{parameter-estimation}{%
-\subsection{Algorithms for Parameter Estimation}\label{parameter-estimation}}
-
-For any edge probability matrix \(P\) for the PABM such that the rows
-and columns are organized by community, the \(kl\)\textsuperscript{th}
-block is an outer product of two vectors, i.e.,
-\(P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top\). Therefore, given
-\(P^{(k \ell)}\), \(\lambda^{(k \ell)}\) and \(\lambda^{(\ell k)}\) are solvable
-up to multiplicative constant using singular value
-decomposition. More specifically, 
-let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\)
-be the singular value decomposition of \(P^{(k \ell)}\) where
-\(u^{(k \ell)} \in \mathbb{R}^{n_k}\) and 
-\(v^{(k \ell)} \in \mathbb{R}^{n_\ell}\) are vectors
-and \(\sigma^{(k \ell)}\) is a scalar. 
-Then \(\rho_n^{1/2} \lambda^{(k \ell)} = s_1 u^{(k \ell)}\)
-and \(\rho_n^{1/2} \lambda^{(\ell k)} = s_2 v^{(k \ell)}\) 
-for unidentifiable $s_1 s_2 = (\sigma^{(k \ell)})^2$.
-Since each $\lambda^{(k \ell)}$ is not strictly identifiable,
-we instead estimate each 
-$\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}$. 
-Given the adjacency matrix \(A\)
-instead of edge probability matrix \(P\), we can simply use plug-in
-estimators by taking the SVD of each $A^{(k \ell)}$ to obtain 
-$\hat{\lambda}^{(k \ell)} = \hat{\sigma}^{(k \ell)} \hat{u}^{(k \ell)}$ 
-using the largest singular value of $A$ and its corresponding singular vectors.
-
-\begin{algorithm}[t]
-  \label{alg:param_est}
-  \DontPrintSemicolon
-  \SetAlgoLined
-  \caption{PABM parameter estimation.}
-  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
-  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(k \ell)}\}_K$.}
-  Arrange the rows and columns of $A$ by community such that each 
-  $A^{(k \ell)}$ block consists of estimated edge probabilities between 
-  communities $k$ and $l$.\;
-  \For {$k, \ell = 1, ..., K$, $k \leq \ell$} {
-    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $k\ell$-th 
-    block.\;
-    Assign $u^{(k \ell)}$ and $v^{(k \ell)}$ as the first columns of $U$ and $V$. 
-    Assign $(\sigma^{(k \ell)})^2 \leftarrow \Sigma_{11}$.\;
-    Assign $\hat{\lambda}^{(k \ell)} \leftarrow \pm \sigma^{(k \ell)} u^{(k \ell)}$ and 
-    $\hat{\lambda}^{(\ell k)} \leftarrow \pm \sigma^{(k \ell)} v^{(k \ell)}$.
-  }
-\end{algorithm}
-
-\begin{theorem}
-\label{theorem6}
-Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived from its corresponding $P^{(k \ell)}$, and let $\hat{\lambda}^{(k \ell)}$ be its estimate from $A^{(k \ell)}$ (Alg.
-3).
-Then if $n \rho_n = \omega( \log^{4}{n})$,
-\begin{equation} \label{eq:thm6}
-\max_{k, l \in \{1, ..., K\}} 
-\|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
-O\bigg(\frac{\log n_k}{\sqrt{n_k}} \bigg)
-\end{equation}
-with high probability.
-\end{theorem}
-
-\hypertarget{simulated-examples}{%
-\section{Simulation Study}\label{simulated-examples}}
-
-For each simulation, community labels are drawn from a multinomial
-distribution, the popularity vectors \(\{\lambda^{(k \ell)}\}_K\) are drawn
-from two types of joint distributions depending on whether \(k = \ell\),
-the edge probability matrix \(P\) is constructed using the popularity
-vectors, and finally the adjacency matrix \(A\)
-is drawn from \(P\). OSC (Algorithm~\ref{alg:osc}) is then used for community detection, and this
-method is compared against (1) SSC using the spectral embedding $\hat{V}$,
-c.f. Algorithm~\ref{alg:ssc}, (2) SSC using the rows of the
-observed adjacency matrix $A$ as is done in \citet{noroozi2019estimation}
-and (3) modularity maximization (MM) as is done in
-\citet{307cbeb9b1be48299388437423d94bf1}. 
-We also estimate the latent popularity vectors $\{\lambda^{(k \ell)}\}$
-by assuming that the true community labels are known and then apply
-Algorithm~\ref{alg:param_est}, and we compare this estimation method against an
-MLE-based estimator as described in \citet{noroozi2019estimation} and
-\citet{307cbeb9b1be48299388437423d94bf1}.
-
-Modularity Maximization is NP-hard, so
-\citet{307cbeb9b1be48299388437423d94bf1} used the Extreme Points
-(EP) algorithm (\cite{le2016}) as a greedy
-relaxation of the optimization problem; the EP algorithm has a running
-time of $O(n^{K-1})$ where $n$ is the number of vertices in the graph
-and $K$ is the number of communities.
-For these simulations we instead replace the EP algorithm with the
-Louvain algorithm for modularity maximization,
-as the implementation of the EP algorithm in
-\citet{307cbeb9b1be48299388437423d94bf1} is too computationally expensive for \(K > 2\). For \(K = 2\), it
-was verified that the Louvain algorithm produces comparable results
-to EP-MM.
-
-Two implementations of SSC are shown here. The first method, denoted as
-SSC-A, treats the columns of the adjacency matrix \(A\) as points in
-\(\mathbb{R}^n\), as described in \citet{noroozi2019estimation}.
-The second method, denoted as SSC-ASE, first embeds \(A\) and then
-performs SSC on the embedding, as described in Algorithm~\ref{alg:ssc}. The 
-parameter \(\lambda\) that controls the sparsity of the SSC was chosen via a preliminary cross-validation
-experiment. For the final clustering step, a Gaussian Mixture Model was
-fit on the normalized Laplacian eigenmap of the affinity matrix \(B\).
-
-For comparing methods, we define the community detection error as:
-\[L_c(\hat{\sigma}, \sigma; \{v_i\}) =
-\min_\pi \sum_i I(\pi \circ \hat{\sigma}(v_i) = \sigma(v_i))\]
-where \(\sigma(v_i)\) is the true community label of vertex \(v_i\),
-\(\hat{\sigma}(v_i)\) is the predicted label of \(v_i\), and \(\pi\) is
-a permutation operator. This is effectively the ``misclustering count''
-of clustering function \(\hat{\sigma}\).
-
-For parameter estimation, because the popularity parameters $\{\lambda_{ik}\}$ are unidentifiable, we instead estimate the edge probabilities $p_{ij} = \lambda_{i z_j} \lambda_{j, z_i}$.
-The parameter estimation error is the root mean square error between the true and estimated edge probabilities, which is equivalent to the Frobenius norm of $P$ divided by the number of vertices:
-
-$$\mathrm{RMSE}(\hat{P}, P) = \frac{1}{n} \|\hat{P} - P\|_F.$$
-
-We also note that unlike the MLE-based method \citep{307cbeb9b1be48299388437423d94bf1}, the ASE method of Algorithm 3 does not require community labels if all we are interested in is the edge probability matrix, $P$.
-The estimate $\hat{P}$ we obtain by combining the popularity vectors outputted by Algorithm 3 is equivalent to using the full ASE of $A$ to reconstruct the edge probability matrix.
-
-\hypertarget{balanced-communities}{%
-\subsection{Balanced Communities}\label{balanced-communities}}
-
-In each simulation, community labels \(z_1, ..., z_n\) were drawn from a
-multinomial distribution with mixture parameters
-\(\{\alpha_1, ..., \alpha_K\}\), then \(\{\lambda^{(k \ell)}\}_K\) according
-to the drawn community labels, \(P\) was constructed using the drawn
-\(\{\lambda^{(k \ell)}\}_K\), and \(A\) was drawn from \(P\).
-
-For these examples, we set the following parameters:
+In this simulation study, we use the same setup as in the simulations
+for balanced communities but with fixed \(K = 3\) and \(n = 2^{11}\) and
+vary the sparsity parameter \(\rho \in (0, 1]\) (this was fixed at
+\(\rho_n = 1\) for the previous simulations. More specifically,
 
@@
 \item
-  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
+  Number of vertices \(n = 2048\)
 \item
-  Number of underlying communities \(K = 2, 3, 4\)
+  Number of underlying communities \(K = 3\)
 \item
-  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\), (i.e.,
+  Sparsity parameter \(\rho_n = 0.1, 0.3, 0.5, 0.7, 0.9\)
+\item
+  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\) (i.e.,
   each community label has an equal probability of being drawn)
@@
   Community labels
-  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
+  \(z_k \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, ..., \alpha_K)\)
 \item
   Within-group popularities
-  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
+  \(\lambda^{(kk)} \stackrel{\mathrm{iid}}{\sim}\mathrm{Beta}(2, 1)\)
 \item
   Between-group popularities
-  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
-  \(k \neq \ell\)
+  \(\lambda^{(k \ell)} \stackrel{\mathrm{iid}}{\sim}\mathrm{Beta}(1, 2)\)
+  for \(k \neq \ell\)
 \end{itemize}
-Fifty simulations were performed for each combination of $n$ and $K$.
 
-Fig.~\ref{fig:clust_err_ct_sim} shows the number of mislabeled vertices
-going to 0 for large \(n\). SSC on both the embedding and on the
-adjacency matrix produces similar trends for \(K > 2\). Weaker
-performance of SSC for \(K = 2\) can be attributed to the final spectral
-clustering step of the affinity matrix. 
-While the subspace detection property is guaranteed for large \(n\), in our
-simulations, setting the sparsity parameter to the required value resulted
-in more than \(K\) disconnected subgraphs. 
-We instead chose a smaller sparsity parameter, 
-necessitating a final clustering step. 
-A GMM was fit to the normalized Laplacian eigenmap of $B$,
-but visual inspection suggests that the communities are not
-distributed as a mixture of Gaussians in the eigenmap. 
-A different choice of mixture distribution may result in better performance.
+Fifty simulations were performed for each combination of \(n\) and
+\(K\). The results for community detection and parameter estimation
+error are shown in Fig. \ref{fig:sparsity_sim}. Community detection
+error is defined as the number of misclustered vertices. Parameter
+estimation error is defined as \(\frac{1}{n} \|P - \hat{P}\|_F\) where
+\(\hat{P}\) is the estimated edge probability matrix.
 
@@
 
-{\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim-1}
+{\centering \includegraphics[width=1\linewidth]{supplemental-materials_files/figure-latex/sparsity_sim-1}
 
@@
 
-\caption{Median and IQR of community detection error. Communities are approximately balanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim}
+\caption{Community detection error counts (left) and popularity parameter RMSEs (right) for $n = 2048$ and $K = 3$ and varying $\rho_n$.
+Simulations were repeated 50 times for each $\rho_n$.}\label{fig:sparsity_sim}
 \end{figure}
 
-Given ground truth community labels, Algorithm~\ref{alg:param_est} and the MLE-based
-plug-in estimators perform similarly, with root mean square
-error decaying at rate approximately \(n^{-1/2}\) (Fig. \ref{fig:p_block_est}).
+\renewcommand\refname{Choice of \(\vartheta\) for the Subspace Detection
+Property}
+  \bibliography{misc.bib}
 
-\begin{figure}[H]
-{\centering \includegraphics{summary_files/figure-latex/lambda_est_p-1}}
-\caption{Median and IQR RMSE for edge probability matrices reconstructed from the outputs of Algorithm 3 (red) compared against outputs of an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be approximately balanced.}
-\label{fig:p_block_est}
-\end{figure}
-
-\hypertarget{imbalanced-communities}{%
-\subsection{Imbalanced Communities}\label{imbalanced-communities}}
-
-Simulations performed in this section are the same as those in the
-previous section with the exception of the mixture parameters
-\(\{\alpha_1, ..., \alpha_K\}\) used to draw community labels from the
-multinomial distribution. For these examples, we set the following
-parameters:
-
-\begin{itemize}
-\tightlist
-\item
-  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
-\item
-  Number of underlying communities \(K = 2, 3, 4\)
-\item
-  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
-  for \(k = 1, ..., K\)
-\item
-  Community labels
-  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
-\item
-  Within-group popularities
-  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
-\item
-  Between-group popularities
-  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
-  \(k \neq \ell\)
-\end{itemize}
-
-\(50\) simulations were performed for each \((n, K)\) pair.
-
-We again see community detection error trending to 0 for OSC, as well as
-for SSC when \(K > 2\) (Fig.~\ref{fig:clust_err_ct_sim_imba}). 
-The parameter estimation error decays at rate $n^{-1/2}$ as in the balanced communities case
-(Fig. \ref{fig:lambda_est_p_imba}).
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/clust_err_ct_sim_imba-1}
-}
-
-\caption{Median and IQR of community detection error. Communities are imbalanced. Simulations were repeated 50 times for each sample size.}\label{fig:clust_err_ct_sim_imba}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/lambda_est_p_imba-1}
-
-}
-
-\caption{Median and IQR RMSE of edge probabilities derived from the outputs of Algorithm 3 (red) compared against an MLE-based method (blue).
-Simulations were repeated 50 times for each sample size.
-Communities were drawn to be imbalanced.}
-\label{fig:lambda_est_p_imba}
-\end{figure}
-
-\hypertarget{real-data-examples}{%
-\section{Applications}\label{real-data-examples}}
-
-In the first example, we applied OSC (Algorithm~\ref{alg:osc}) to the Leeds Butterfly
-dataset \citep{Wang_2018} consisting of visual similarity measurements
-among 832 butterflies across 10 species. The graph was modified to match
-the example from \citet{noroozi2019estimation}, i.e., only the $K=4$ most
-frequent species were considered, and the similarities were discretized
-to \(\{0, 1\}\) via thresholding. Fig.~\ref{fig:butterfly} shows a
-sorted adjacency matrix sorted by the resultant clustering.
-
-Comparing against the ground truth species labels, OSC achieves an adjusted Rand index of 92\%, while SSC on the ASE achieves an adjusted Rand index of 96\%.
-In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{summary_files/figure-latex/butterfly-1}
-
-}
-
-\caption{Adjacency matrix of the Leeds Butterfly dataset 
-after sorting by the clustering outputted by OSC.}\label{fig:butterfly}
-\end{figure}
-
-In the second example, we applied OSC to the British MPs Twitter network
-\citep{greene2013producing}, the Political Blogs network
-\citep{10.1145/1134271.1134277}, and the DBLP network
-\citep{NIPS2009_3855, 10.1007/978-3-642-15880-3_42}. For this data
-analysis, we subsetted the data as described in
-\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the
-same networks. Our methods slightly underperformed compared to modularity
-maximization, although performance is comparable. The run time of OSC
-is however much smaller than that of modularity maximization.
-
-\begin{figure}[H]
-{\centering \includegraphics{summary_files/figure-latex/mp-1}
-}
-\caption{Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by OSC.}\label{fig:mp}
-\end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}[t]{l|r|r|r}
-\hline
-Network & MM & SSC-ASE & OSC\\
-\hline
-British MPs & 0.003 & 0.018 & 0.009\\
-\hline
-Political Blogs & 0.050 & 0.196 & 0.062\\
-\hline
-DBLP & 0.028 & 0.087 & 0.059\\
-\hline
-\end{tabular}
-\caption{\label{tab:unnamed-chunk-6}Community detection error rates on the British MPs Twitter, Political Blogs, and DBLP networks using modularity maximization, sparse subspace clustering, and OSC.}
-\end{table}
-
-In the third example (Fig.~\ref{fig:households-figure} and 
-Table~\ref{tab:households-table}), 
-we analyzed the Karantaka villages data studied by
-\citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
-networks from villages 12, 31, and 46 at the household level. 
-In these networks, each node is a household and each edge is 
-an interaction between members of pairs of households. 
-The label of interest is the religious affiliation. 
-The networks were truncated to religions ``1'' and ``2'', 
-and vertices of degree 0 were removed. 
-The villages were chosen based on there being an adequate number of nodes 
-between households within each religion.
-
-\begin{figure}[tp]
-{\centering \includegraphics{summary_files/figure-latex/unnamed-chunk-7-1}
-}
-\caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left).
-The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:households-figure}
-\end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}[t]{l|r|r|r}
-\hline
-Network & MM & SSC-ASE & OSC\\
-\hline
-Village 12 & 0.270 & 0.291 & 0.227\\
-\hline
-Village 31 & 0.125 & 0.066 & 0.110\\
-\hline
-Village 46 & 0.052 & 0.463 & 0.078\\
-\hline
-\end{tabular}
-\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
-\end{table}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Our central result states that the Popularity Adjusted Block Model 
-is a special case of the Generalized Random Dot Product Graph. 
-In particular, the PABM with $K$ communities is a GRDPG 
-for which the communities are represented by 
-mutually orthogonal $K$-dimensional subspaces 
-of the $K^2$-dimensional latent space. 
-This result extends previous results that connected 
-the Stochastic Block Model and the Degree Corrected Block Model 
-to Random Dot Product Graphs. 
-Replacing RDPGs with GRDPGs is a critical step in this line of research, 
-as a PABM is not necessarily a RDPG.
-
-Because all Bernoulli Graphs are GRDPGs, it should be possible to
-invent and study new families of Bernoulli Graphs by characterizing
-them as special cases of GRDPGs and exploiting the latent structures
-that define them.  The present work illustrates the power of this
-approach.  We recover the latent structure of the PABM by Adjacency
-Spectral Embedding, then exploit that structure to improve statistical
-inference.  Exploiting the fact that PABM communities correspond to
-orthogonal subspaces, we propose Orthogonal Spectral Clustering for
-community detection and demonstrate that the number of misclassified
-vertices approaches zero with high probability as the size of the
-graph increases.  This is a stronger result than previously proposed
-algorithms \citep{307cbeb9b1be48299388437423d94bf1}, which only
-guarantee that the error rate (and not count) approaches zero
-asymptotically.  Parameter estimation can be performed in a similar
-fashion using the ASE, for which we also prove that the per-parameter
-error approaches zero asymptotically.
-
-A secondary benefit of the GRDPG approach is that the latent structure
-may be used to improve existing algorithms.  For example, one
-algorithm for PABM community detection \citep{noroozi2019estimation}
-relies on Sparse Subspace Clustering.  The latent structure of the
-PABM provides a natural justification for SSC for the PABM and leads
-to an improvement over the previous implementation.  The improved
-algorithm applies SSC to the ASE, and we prove that the ASE of the
-PABM obeys the Subspace Detection Property with high probability if
-the graph is large.
-
-Finally, one might well inquire what one gains and what one sacrifices
-by assuming that a Bernoulli Graph is a PABM.  The GRDPG model offers
-a plausible way to pursue this inquiry.  Absent a known latent
-structure that can be exploited by specialized methods, the GRDPG-ASE
-approach transforms the problem of network community detection to the
-much-studied problem of clustering vectors in Euclidean space.
-Communities of vertices are defined as clusters of latent vectors.
-After ASE, a standard clustering algorithm, e.g., single linkage, is
-used ot infer the communities.  In future research, we intend to use
-such general algorithms as baselines and measure the efficiency of the
-PABM algorithms (and other specialized algorithms) by studying how
-much they improve on general algorithms when the specified latent
-structure obtains.
-
-\appendix
-\renewcommand\refname{References}
-% \bibliographystyle{plain}
-% \bibliography{misc.bib}
-\printbibliography
-
-\hypertarget{proofs}{%
-\section{Proofs of Theorem~\ref{theorem4}, Theorem~\ref{theorem5}, and Theorem~\ref{theorem6}}}
-Let \(V_n\) and \(\hat{V}_n\)
-be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
-$K^2$ largest eigenvalues (in modulus), respectively. 
-We first state an important technical lemma for bounding the maximum
-$\ell_2$ norm difference between the rows of $\hat{V}_n$ and
-$V_n$. See \citet{cape_biometrika} and 
-\citet[Lemma~5]{rubindelanchy2017statistical} for a proof. 
-\begin{lemma}
-  \label{lem:technical}
-Let $A \sim \mathrm{PABM}(\{\lambda^{(k \ell \ell)}\}_{K})$ be a $K$-blocks
-PABM graph on $n$ vertices and let \(V_n\) and \(\hat{V}_n\)
-be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
-$K^2$ largest eigenvalues in modulus, respectively.
-Let \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
-row of \(V_n\) and \(\hat{V}_n\), respectively. 
-Then there exists a constant $c > 1$ and an orthogonal matrix $W$ such
-that with high probability,
-$$\max_{i} \|W \hat{v}_n^{(i)} - v_n^{(i)}\|  = O\Big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \Big).$$
-In particular we can take $c = 1 + \epsilon$ for any $\epsilon > 0$. 
-\end{lemma}
-
-%
-%then the \(ij\)\textsuperscript{th} element of
-%\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.
-
-\begin{proof}[Proof of Theorem~\ref{theorem4}]
-  Recall the notations in Lemma~\ref{lem:technical} and note that,
-  under our assumption that the latent vectors $\lambda^{(k \ell)}$
-  are all homogeneous, we have $\max_{i} \|v_n^{(i)}\| =
-  O(n^{-1/2})$. 
-%   Let \(V_n\) and \(\hat{V}_n\)
-% be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
-% \(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
-% eigenvalues, respectively. From Lemma~5 in
-% \citet{rubindelanchy2017statistical} we have, for some
-% \(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
-% $$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{\log^{c}n}{n \sqrt{\rho_n}} \big).$$
-% We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). 
-% Let
-% \((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th 
-% row of \(V_n\) and \(\hat{V}_n\), respectively. 
-Next recall Theorem~\ref{theorem3}; in particular $B_{ij} = n(v_n^{(i)})^{\top}
-v_n^{(j)}$. % Now suppose that vertices $i$ and $j$ belongs to different
-% communities. Then $B_{ij} = 0$ and hence
-% Then for \(i\) and
-% \(j\) belonging to different
-% communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
-We therefore have
-\[\begin{split}
- \max_{ij} |\hat{B}_{ij} - B_{ij}| &= \max_{ij} n |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& \leq n \max_{ij} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
-(v_n^{(i)})^\top v_n^{(j)}| \\
-& \leq n \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|
-+ \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
-% & \leq 
-% \hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
-% & = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
-% 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
-% & = ||2 (\hat{V}_n W - V_n) V_n^\top + 
-% (\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
-% & \leq 2 
-% &\leq  n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|\hat{v}_n^{(j)}\| +
-% n \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr)
-% \\
-& \leq n \Bigl(\max_{ij}  \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 +  \|W
-\hat{v}_n^{(i)}  - v_n^{(i)} \| \times \|v_n^{(j)}\| +  \|W \hat{v}_n^{(j)}  - v_n^{(j)} \| \times \|v_n^{(i)}\|\Bigr) 
-\\ &
-\leq n \max_{i} \|W \hat{v}_n^{(i)}  - v_n^{(i)} \|^2 + 2n \max_{i}
-\|W \hat{v}_n^{(i)}  - v_n^{(i)} \| \times \max_{j} \|v_n^{(j)}\|
-\\
-& = O \Big( \frac{\log^{c}{n}}{n^{1/2} \rho_n^{1/2}} \Big)
-\end{split}\]
-with high probability.
-Theorem~\ref{theorem4} follows from the above bound together with the
-conclusion in Theorem~\ref{theorem3} that $B_{ij} = 0$ whenever vertices $i$ and $j$
-belongs to different communities. 
-\end{proof}
-% Scaling all terms in the above inequality by \(n\), we obtain
-% $$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
-% as desired.
-
-We now provide a proof of Theorem~\ref{theorem5}. Our proof is based
-on verifying the sufficient conditions given in Theorem~6 of
-\citet{jmlr-v28-wang13}
-under which sparse subspace clustering based on solving the
-optimization problem in Eq.~\eqref{eq:ssc} yields an affinity matrix
-$B = |C| + |C^{\top}|$ satisfying the subspace detection property of
-Definition~\ref{def:subspace_detection}. We first recall a few
-definitions used in \citet{soltanolkotabi2012} and \citet{jmlr-v28-wang13}; for ease of exposition,
-these definitions are stated using the notations of the current
-paper and we will drop the explicit dependency on $n$ from our
-eigenvectors $\hat{V}_n$ of $A$ and $V_n$ of $P$.
- % We first recall the
-% notion of the inradius of a set of points. 
-\begin{definition}[Inradius]
-  \label{inradius}
-The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
-defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
-Let $X$ be a $n \times d$ matrix with rows $x_1, x_2, \dots,
-x_n$. We then define, with a slight abuse of notation, $r(X)$ as the
-inradius of the convex hull formed by $\{\pm x_1, \pm x_2, \dots, \pm x_n\}$. 
-\end{definition}
-
-\begin{definition}[Subspace incoherence]
-  \label{def:subspace_incoherence}
-Let $\hat{V}$ be the eigenvectors of $A$
-corresponding to the $K^2$ largest eigenvalues in modulus. Let
-$\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
-$\hat{V}$ corresponding to the $k^{th}$
-community and let $\hat{V}^{(-k)}$ denote the matrix formed by
-omitting the rows of $\hat{V}$
-corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
-the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
-the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
-$v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
-$P$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
-rows of $V^{(k)}$.
-
-Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
-1,2,\dots,n_{k}$ as the solution of the following optimization problem
-$$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
-  \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
-  \eta\|_\infty \leq 1$.}$$
-Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
-be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
-$\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
-as
-$$w_i^{(k)} =
-\frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
-Now let $W^{(k)} = \bigl[ w_1^{(k)} \mid \cdots \mid w_{n_k}^{(k)} \bigr]^\top$
-and define the subspace incoherence for $\hat{V}^{(k)}$ by
-$$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
-\end{definition}
-
-With the above definitions in place, we are now ready to state our
-proof of Theorem~\ref{theorem5}.
-
-\begin{proof}[Proof of Theorem~\ref{theorem5}]
-For a given $k = 1,2\dots,K$, let $r^{(k)} = \min_{i}r(V_{-i}^{(k)})$ be inradius of the convex hull formed by
-the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
-\citet{jmlr-v28-wang13} states that there exists a $\lambda > 0$
-such that $\sqrt{n} \hat{V}$ satisfies
-the subspace detection property in
-Definition~\ref{def:subspace_detection} whenever
-the following two conditions are satisfied
-\begin{gather}
-  \label{eq:cond1}
-  \mu^{(k)} < r^{(k)} \quad \text{for all $k = 1,2,\dots,K$}, \\
-  \label{eq:cond2}
-  \max_{i} \|W \hat{v}_{i} - v_{i}\| \leq \min_{k} \frac{r_*(r^{(k)} -
-    \mu^{(k)})}{2 + 7 r^{(k)}}.
-\end{gather}
-We now verify that for sufficiently large n, Eq.~\eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
-holds with high probability.
-
-{\bf Verifying Eq.~\eqref{eq:cond1}}. If $n$ is sufficiently large then
-there are enough vertices in each community $k$ so that
-$\mathrm{span}(V_{-i}^{(k)}) = \mathcal{S}^{(k)}$ for all $i$ and hence
-\(r^{(k)} = \min_{i} r(V_{-i}^{(k)}) > 0\) for
-all $k = 1,2,\dots,K$.
-
-Next, by Theorem \ref{theorem3} we have that the subspaces
-  $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
-  are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
-  \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
-  \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
-  $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
-    $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
-    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
-    \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ for all $k
-    =1,2,\dots,K$. Therefore $\mu^{(k)} < r^{(k)}$ for all $k =
-    1=2,\dots,K$ as desired.
-
-{\bf Verifying Eq.~\eqref{eq:cond2}}.
-Let $\delta = \max_{i} \sqrt{n} \|W
-\hat{v}_{i} - v_{i}\|$. Then from Lemma~\ref{lem:technical}, we have
-\(\delta \stackrel{a.s.}{\to} 0\) and hence
-$$\delta < \min_{k} \frac{r_* (r^{(k)} - \mu^{(k)})}{2 + 7 r^{(k)}}$$
-asymptotically almost surely.
-
-In summary $\sqrt{n} \hat{V}$ satisfies the subspace detection property
-with probability converging to $1$ as \(n \to \infty\).
-\end{proof}
-
-\begin{remark}
-Theorem 6 of \citet{jmlr-v28-wang13} assumes that each row $v_i$
-of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. This
-assumption has the effect of scaling the $r^{(k)}$ so that $r^{(k)}
-\leq 1$ for all $k = 1,2,\dots,K$. We emphasize that this assumption
-has no effect on the proof of Theorem \ref{theorem5}. Indeed,
-since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
-spans the subspace $\mathcal{S}^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
-for any scalar $a > 0$. 
-\end{remark}
-
-\begin{proof}[Proof of Theorem \ref{theorem6}] Let \(P\) be
-organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
-n_{\ell}$ matrix obtained by keeping only the rows of $P$
-corresponding to vertices in community $k$ and the columns of $P$
-corresponding to vertices in community $\ell$. We define $A^{(k
-  \ell)}$ analogously. Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
-for the cases when $k = \ell$ versus when $k \not = \ell$.
-
-\emph{Case \(k = l\)}. Let $P^{(kk)} =
-\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
-decomposition of $P^{(kk)}$. We can then define
-$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
-Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
-singular value decomposition of \(A^{(kk)}\), and let
-$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
-best rank-one approximation of $A^{(kk)}$. Define
-\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
-\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
-and by Theorem 5 of \citet{rubindelanchy2017statistical}, we have
-%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
-$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O\Bigl(\frac{\log n_k}{\sqrt{n_k}}\Bigr)$$
-with high probability. 
-Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.
-
-\emph{Case \(k \neq l\)}. % \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
-% probabilities and edges between communities \(k\) and \(l\). Note that
-% \(P^{(kl)} = (P^{(lk)})^\top\).\\
-% By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). 
-Let \(P^{(k \ell)} = \sigma_{k \ell}^2 u^{(k \ell)} (v^{(k \ell)})^\top\) and 
-\(P^{(\ell k)} = \sigma_{kl}^2 u^{(\ell k)} (v^{(\ell k)})^\top\)  be the singular
-value decompositions and note that \(\sigma_{k \ell} = \sigma_{\ell
-  k}\), \(u^{(k \ell)} = v^{(\ell k)}\), and
-\(v^{(k \ell)} = u^{(\ell k)}\). 
-Now define \(\lambda^{(k \ell)} = \sigma_{k \ell} u^{(k \ell)}\) and
-\(\lambda^{(\ell k)} = \sigma_{k \ell}
-v^{(k \ell)}\).
-
-% (We can also note that
-% the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
-% \(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
-% \(v^{(kl)} = u^{(lk)}\).)\\
-Next consider the Hermitian dilation
-
-\[M^{(k \ell)} = 2 \begin{bmatrix} 0 & P^{(k \ell)} \\ P^{(\ell k)} & 0 \end{bmatrix}\]
-
-which is a symmetric \((n_k + n_\ell) \times (n_k + n_\ell)\) matrix. The
-eigendecomposition of \(M^{(k \ell)}\) is then
-
-\[M^{(k \ell)} = 
-\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix} \times 
-\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
-\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix}^\top\]
-Thus treating \(M^{(k \ell)}\) as the edge probability matrix of a GRDPG, we
-have latent positions in \(\mathbb{R}^2\) given by the $(n_k + n_{\ell}) \times 2$ matrix
-\[\Lambda^{(k \ell)} = \begin{bmatrix} 
-  \sigma_{k \ell} u^{(k \ell)} & \sigma_{k \ell} u^{(k \ell)} \\ 
-  \sigma_{k \ell} v^{(k \ell)} & -\sigma_{k \ell} v^{(k \ell)} 
-\end{bmatrix} = 
-\begin{bmatrix} 
-  \lambda^{(k \ell)} & \lambda^{(k \ell)} \\ 
-  \lambda^{( \ell k)} & -\lambda^{( \ell k)} 
-\end{bmatrix}.\]
-Now consider
-\[\hat{M}^{(k \ell)} = \begin{bmatrix} 0 & A^{(k \ell)} \\ A^{(\ell k)} & 0 \end{bmatrix}\]
-% Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
-
-% \[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]
-
-% and \(E\) is the \(n_k \times n_\ell\) matrix of independent noise (to
-% generate the Bernoulli entries in \(A^{(kl)}\).\\
-We can then view \(\hat{M}^{(k \ell)}\) as an adjacency matrix drawn from
-the edge probabilities matrix \(M^{(k \ell)}\). Now suppose that the adjacency spectral
-embedding of $\hat{M}^{(k \ell)}$ is represented as the $(n_k +
-n_{\ell}) \times 2$ matrix
-\[\hat{\Lambda}^{(k \ell)} = \begin{bmatrix} 
-  \hat{\lambda}^{(k \ell)} & \hat{\lambda}^{(k \ell)} \\ 
-  \hat{\lambda}^{(\ell k)} & -\hat{\lambda}^{(\ell k)} 
-\end{bmatrix}\]
-where each \(\hat{\lambda}^{(k \ell)}\) is defined as in Algorithm 3. Then
-by Theorem 5 of \citet{rubindelanchy2017statistical}, there
-exists an indefinite orthogonal transformation $W^{*}$ such that,  with
-high probability,
-$$\max_{i} |W^{*} \hat{\Lambda}_{i}^{(k \ell)} - \Lambda_{i}^{(k \ell)} \| =
-O\Bigl(\frac{\log (n_k + n_{\ell})}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
-with high probability. Here $\Lambda_{i}^{(k \ell)}$ and
-$\hat{\Lambda}_i^{(k \ell)}$ denote the $i$th rows of $\Lambda^{(k
-  \ell)}$ and $\hat{\Lambda}^{(k \ell)}$, respectively. 
-% approximates the latent positions of \(M^{(kl)}\) up to indefinite
-% orthogonal transformation by the rate given in Theorem 5 of
-% \citet{rubindelanchy2017statistical}.\\
-Furthermore, by looking at the proof of Theorem~5 in
-\citep{rubindelanchy2017statistical}, we see that $W^{*}$ is also
-blocks diagonal with $2$ blocks where the positive eigenvalues of $M^{(k \ell)}$
-forming a block and the negative eigenvalues of $M^{(k \ell)}$ forming
-the remaining block. % \(U^\top \hat{U}\). 
-Since $M^{(k \ell)}$ has one positive eigenvalue and one negative
-eigenvalue, we see that $W^{*}$ is necessarily of the form $W^{*}
-= \Bigl[\begin{smallmatrix} 1 & 0 \\ 0 & - 1\end{smallmatrix}\Bigr]$
-% \((1, 1)\), and \(U^\top \hat{U}\) is block
-% diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). 
-Using this form for $W^{*}$, we obtain
-$$\max\{\|\hat{\lambda}^{(k \ell)} - \lambda^{(k \ell)}\|_{\infty},
-\|\hat{\lambda}^{(\ell k)} - \lambda^{(\ell k)}\|_{\infty}\}  =
-O\Bigl(\frac{\log(n_k + n_{\ell})}{\sqrt{n_k + n_{\ell}}}\Bigr)$$
-with high probability. Combining this bound with the bound for
-$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty}$ given above
-yields Eq.~\eqref{eq:thm6} in Theorem~\ref{theorem6}. 
-\end{proof}
 \end{document}

### 2022-02-07 referee-response.Rmd
Update git repo

@@
 The concerns addressed here are answered below. 
-In particular, we make sure to address the concerns about the simulation studies. 
+In particular, we made considerable effort to address the referees' concerns about the simulation studies. 
 }
@@
 order 2, and the latent vectors lie in the union of two $2$-dimensional
-orthogonal subspaces, i.e., and the rows of $X$ consist of
+orthogonal subspaces, i.e., the rows of $X$ consist of
 vectors that lie in $\mathcal{S}_1 \cup \mathcal{S}_2$ where
 $\mathcal{S}_1$ and $\mathcal{S}_2$ are both $2$-dimensional
-subspaces of $\mathbb{R}^{4}$, $\mathcal{S}_1 \cap \mathcal{S}_2 = \{0\}$, and $x^{\top} y
+subspaces of $\mathbb{R}^{4}$ with $\mathcal{S}_1 \cap \mathcal{S}_2 = \{0\}$ and $x^{\top} y
 = 0$ for all $x \in \mathcal{S}_1, y \in \mathcal{S}_2$. 
@@
 corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
-points are then the columns with indices $r(K+1) +1 $ for $0 \leq r
+points of $\Pi$ are then the columns with indices $r(K+1) +1 $ for $0 \leq r
 \leq K-1$; these columns contain the vectors $\lambda_{kk}$ for
@@
 \textcolor{blue}{
-The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-This is not immediately obvious even if we know $P$ beforehand, although having $P$ does make estimation much easier. 
-In the introduction to section 3 (before 3.1), we have added the following:
-}  
-\textcolor{blue}{
-``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
-After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.''
+We thank the referee for pointing out this issue. We had revised the
+manuscript to include the line "In our methods, the data that are
+observed is only the adjacency matrix $A \sim
+\mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$ along with an assumed
+number of communities, $K$. To motivate our methods, we first
+consider community detection and parameter estimation in the case
+where we know the edge probability matrix, $P$, beforehand, noting
+that community memberships and popularity parameters are not
+immediately discernible from $P$ itself. After establishing methods
+for community detection and parameter estimation from $P$, we use the
+consistency property of ASE to demonstrate that the same methods work
+for $A$ almost surely as $n \to \infty$." Indeed, the assumption of
+known $P$ helps motivate Algorithm $1$ through Algorithm~3 and the
+theoretical results in Theorem~2. Theorem 3 and Theorem 4 then follows
+by leveraging the consistency properties for adjacency spectral embedding
+to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
+community recover and paraemters estimation using the true but unknown $P$. 
 }
@@
 \textcolor{blue}{
-We have added more detail to our description of the SSC algorithm. 
-In particular, we describe the intuition/motivation behind the algorithm. 
-}  
-\textcolor{blue}{The description of SSC now reads:
-}  
-\textcolor{blue}{
+We have added more details to our description of SSC to
+provide further intuition/motivation behind the
+algorithm. The description of SSC now reads:
 ``The SSC algorithm can be described as follows: 
@@
 consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
-The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace.
-Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
-Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+The intuition here is that vectors that lie on the same subspace can
+be described as linear combinations of each other, assuming the number
+of vectors in the subspace is greater than the dimensionality of the
+subspace.
+Thus, for each $c_i$, $c_i^{(j)}$ is zero whenever $x_i$ and $x_j$ belong to different subspaces and may be nonzero otherwise.
 If \(X\) instead represents points near \(K\) subspaces with some noise,
-then this property may only hold approximately and 
+then this property will only hold approximately and 
 a final graph partitioning step may be required 
@@
 \textcolor{blue}{
-There are no references for this, as it is a novel idea. 
-SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
-The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
-}  
-\textcolor{blue}{
-We have added, ``...
-SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''
+There are no references for this, as it is, to the best of our knowledge, a novel idea. 
+We emphasize that while SSC has been proposed for community detection
+for PABM (see e.g., \citep{noroozi2019estimation}), these work apply SSC
+directly on the rows of $A$ (which are binary vectors in
+$\mathbb{R}^{n}$) rather than on the ASE of $A$ (which are real-valued
+vectors in $\mathbb{R}^{d}$ where $d \ll n$). The motivation behind
+our use of SSC comes  from Theorem 1 wherein we show that the latent
+vectors for PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
+by the consistency properties of ASE, the rows of the estimated
+eigenvectors $\hat{V}$ also lies close to the union of orthogonal
+subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
+union of $K$ orthogonal subspaces in $\mathbb{R}^{n}$, the rows of $A$
+(which are now noisy binary vectors) could be quite far from the
+corresponding rows of $P$ and hence need not lie close to the union
+of any $K$ orthogonal subspaces in $\mathbb{R}^{n}$ 
 }
@@
 \textcolor{blue}{
-We added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
-Results of this simulation study are very similar to that of the first two and consistent with the theory.
+The OSC algorithm works for any PABM parameters setting and is thus
+agnostic to the particular structure (such as assortativity vs
+disassortativity). We had revised the manuscript to include an
+additional simulation demonstrating this claim (see Section 4.3). In
+particular Section~4.3 considers a PABM setting wherein
+$\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
+are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$ is a
+vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
+random variables. This result in an average within-community edge
+probability of $1/3$ and a between-community edge
+probability of $2/3$. The results for this simultation setting are
+very similar to those presented in the original manuscript and is also
+consistent with the theoretical results in the paper.
 }  
 \textcolor{blue}{
-On a related note, it is our view that the distinction between assortative and disassortative is not very important for the PABM, at least when it is viewed as a type of GRDPG. 
-In the case of the SBM and DCBM, it can be shown that the rank of $P$ (and therefore the number of embedding dimensions) is equal to $K$, but assortativity or disassortativity determines how many positive and negative eigenvalues $P$ has \citep{rubindelanchy2017statistical}.
-In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming each community spans its respective subspace).
+On a related note, it is our view that the distinction between
+assortative and disassortative is not always meaningful for the
+PABM. Indeed, unlike the SBM, in the PABM each vertex is free to have a higher affinity to its
+own community or to other communities. For example, suppose vertex $i$
+belongs to community $1$. Then the average probability of an edge
+between vertex $i$ and another vertex in community $1$ is $n^{-1}
+\rho_n 
+\lambda_{i1} \sum_{j} \lambda_{j1}$ while the average probability of an edge between vertex $i$ and another vertex in community $k \not =
+1$ is $n^{-1} \rho_n \lambda_{ik} \sum_{j} \lambda_{j1}$ and thus, on
+average, vertex $i$ have more affinity with vertices in community $k \not =
+1$ whenever $\lambda_{i1} < \lambda_{ik}$
 }
@@
 \textcolor{blue}{
-The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
-For the Leeds Butterfly dataset, while the paper with which we are comparing \citep{noroozi2019estimation} provides enough details to transform and preprocess the data to match their analysis, they do not provide code to run their clustering algorithm.
-Therefore, we use their reported benchmark value, which is in the form of ARI. 
-In order to make an apples-to-apples comparison, we report the ARI for OSC and SSC-ASE here as well. 
-For all other analyses (simulation and real data), we use misclassification error (whether rate or count). 
+The first two datasets that we analyzed in section 5 had also been analyzed
+in previous papers on the PABM. In particular the
+Leeds Butterfly dataset was analyzed in \citep{noroozi2019estimation}
+and while \citep{noroozi2019estimation} provides enough details
+for us to preprocess the data to match their analysis, they do not
+provide code to run their clustering algorithm (which is based on SSC
+applied to the rows of $A$). Therefore we used their reported
+benchmark values, which are given in terms of ARI. 
+In order to make an apples-to-apples comparison, we also report the ARI for OSC and SSC-ASE here as well. 
+For all other analyses (simulation and real data), we use the
+misclassification error rate/count. 
 }
@@
 \textcolor{blue}{
-Sparsity was not considered in our analyses and algorithms.
+Sparsity was not considered in our analyses and algorithms. (Fix this)
 }
@@
 \textcolor{blue}{
-Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
-Performance of the algorithms tested is similar to the other two simulation studies and consistent with the theory. 
+As we discussed in the response to comment
+$14$, we have added a third simulation to address the issue of disassortativity (see Section
+4.3 of the revised manuscript). The performance of all the algorithm
+for the setting in Section~4.3 is qualitatively similar to those
+presented earlier in Section~4.1 and Section~4.2 and is consistent with the theory. 
 Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$.
@@
 \textcolor{blue}{
-Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
-For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
-For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}.
-SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
-When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^3)$. 
-In practice, we have found that $K$ does not affect runtimes too much. 
+As for the computational complexity of the algorithms, the main bottleneck for OSC and
+SSC-ASE is the construction of the affinity matrix $B$. 
+For OSC, this involves a (truncated) spectral decomposition to extract the
+$K^2$ largest eigenvalues in modulus (and their corresponding
+eigenvectors) followed by a matrix multiplication of $n \times K^2$
+and $K^2 \times n$ matrices, so the time complexity is of the order $O(n^2 K^2)$. 
+For SSC-ASE (applying SSC on the rows of the adjacency spectral
+embedding $\hat{V}$) we again start with spectral decomposition, followed by
+$n$ LASSO problems with a design matrix of size 
+$K^2 \times (n - 1)$, so the complexity is $O(n^2 K^2 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}. 
+SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$.
+When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^2 K^2)$. 
+In practice the choice of parameter $K$ does not affect
+runtimes too much. For the simulation studies we only considered $K
+\leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
+comparable estimation accuracy as $K \leq 4$. 
 }
@@
 \textcolor{blue}{
-For the Leeds butterfly dataset, we wanted to compare results of OSC and SSC-ASE to an analysis performed by another group studying the PABM \citep{noroozi2019estimation}. 
-In their results, they describe how they removed some of the nodes and discretized the edges, which we replicate in our analysis. 
-}
+For the Leeds butterfly dataset, we wanted to compare our results
+using OSC and SSC-ASE to an earlier analysis using SSC on the rows of
+$A$ (see \citep{noroozi2019estimation}). In their analysis they
+describe how they removed some of the nodes (corresponding to the
+other butterfly species) so that the resulting dataset only include the
+$K = 4$ butterfly species as considered in our paper.}
 
@@
 Therefore, the dimensionality of the latent configuration is not in question, only what kind of effect multiplication by unidentifiable $Q$ has on the configuration. 
-The OSC algorithm circumvents the nonidentifiability of the latent configuration by only considering the entries of $B = n V V^\top$. 
-SSC-ASE does rely on the latent configuration, but we show that the orthogonality property holds for the particular embedding we perform for this algorithm. 
+The OSC algorithm circumvents the nonidentifiability of the latent
+configuration by only considering the entries of $\hat{B} = n \hat{V}
+\hat{V}^{\top}$ as $n\hat{V}\hat{V}^{\top}$ is, by the asymptotic properties of
+adjacency spectral embedding, close to $n VV^{\top}$ and $n V
+V^{\top}$ is always unique (the orthogonal projection onto a column
+space is always unique). Similarly, SSC-ASE also circumvents this non-identifiability
+by only using the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
+$\hat{v}_j$ of $\hat{V}$ belonging to different subspaces, i.e.,
+different communities, will also
+have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
 }  
 \textcolor{blue}{
-That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
-If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the clustering 
-In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).
+That said, looking back at the latent configuration in Theorem 1, we
+can see that a $K$-dimensional (not $K^2$-dimensional) embedding is
+sufficient for clustering, i.e., the embedding corresponding to the
+diagonal blocks $\lambda_{11}, \lambda_{22}, \dots, \lambda_{KK}$. 
+If we are able to find this embedding then this would greatly reduce
+the number of dimensions required for the community
+recovery/clustering. In practice, this isn't always possible due to
+the nonidentifiability issue; in particular, determining which of the
+$K$ out of $K^2$ eigenvectors should be used is the main difficulty.
+Nevertheless we have found via numerical experiments that community
+detection is often possible using only the eigenvectors corresponding
+to the $2^\mathrm{nd}$ to the $(K+1)^\mathrm{st}$ largest eigenvalues
+of $A$. 
 }
@@
 \textcolor{blue}{
-We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
-One possible reason for this may be because we failed to find the right hyperparameter for SSC-ASE. 
+We are not entirely clear on why this happens for $K = 2$, but when
+comparing the normalized Laplacian eigenmaps of the affinity matrix
+$B$ for various $K$, we see that when $K > 2$, the points appear much
+closer to Gaussian mixtures, whereas with $K = 2$, while there is some
+separation by community, they do not appear as Gaussian mixtures. 
+One possible reason for this may be because we failed to find the
+right value for the hyperparameter $\vartheta$ for SSC-ASE. 
 }  
 \textcolor{blue}{
-Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather {\em at least} $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an ``incorrect'' $\vartheta$ for SSC and then performing the final clustering step.
+More specifically, under ideal circumstances, we would be able to
+circumvent the issue of selecting the hyperparameter $\vartheta$ by
+using the theoretical value that ensures, with high probability, the rows of $\hat{V}$ satisfy the subspace
+detection property (SDP). However, as other researchers had pointed out (see
+e.g., \citep{sdp_sufficiency,liu_ssc}, the fact that SDP is satisfied
+does not guarantee that the affinity matrix $B$ represents a graph
+with $K$ connected components but rather a graph with {\em at least}
+$K$ connected component. Indeed, we found in our simulations that
+setting $\vartheta$ to the theoretical values that guarantees 
+SDP often results in $B$ having more than $K$ connected components.
+This is why we choose to do spectral clustering on $B$, i.e., we first
+compute the normalized Laplacian embeddings using $B$ before
+clustering the rows of this embedding using Gaussian mixtures
+modeling. 
 }  
 \textcolor{blue}{
-We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output estimated labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
-For instance, in section 4.1 for, the error count for SSC-ASE for $K = 2$ is around 20 when $n = 128$ and around 100 when $n = 4096$, resulting in an error rate of around $16\%$ for $n = 128$ and $2\%$ for $n = 4096$.
+Finally we note that in our simulations the proportion of mislabeled
+vertices does decreases as $n$ increases. 
+For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
+vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
+$n = 4096$; this correspond to an error rate of around $16\%$ and
+$2\%$, respectively.
 }
@@
 In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 
-While all three algorithms are specific to the PABM, they might all behave differently to the different ways in which the model is misspecified. 
-There is also the question of hyperparameter tuning, especially in the case of SSC-ASE, and we spent a bit more time tweaking the hyperparameters, resulting in performance that is a bit closer to that of MM.
-}
+While all three algorithms are specific to the PABM, they might behave
+differently to the different ways in which the model is
+misspecified. Finally, we note that Section~5 analyzed three real
+datasets. For the first dataset (Leeds butterfly) we didn't run the MM
+algorithm as the original implementation of MM used in 
+\cite{307cbeb9b1be48299388437423d94bf1} is extremely slow when $K = 4$. For
+the second dataset (British MP's), the accuracy of MM is slightly
+better than that of OSC. Finally, for the last dataset (Karantaka
+villages), the accuracy of MM is slightly worse than that of OSC. }
 

### 2022-02-07 summary_minh_edited.tex
Update git repo

@@
 In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
+To motivate our methods, we first consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
 After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.

### 2022-02-08 referee-response.tex
updated referee response

@@
   that for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
-  $x^{\top} I_{p,q} y \in [0,1]$. }
+  $x^{\top} I_{p,q} y \in [0,1]$. One simple way to construct such a set
+  $\mathcal{X}$ is to first find a collection of $K$ vectors
+  $\mathcal{S} = \{\nu_1, \nu_2, \dots, \nu_K\} \subset \mathbb{R}^{d}$
+  where $\nu_i^{\top} I_{p,q} \nu_j \in [0,1]$ for all $i,j$ and then define $\mathcal{X}$
+  as 
+  $$\mathcal{X} = \mathrm{conv}(\mathcal{S}) = \{x = \sum_{i} \lambda_i \nu_i \colon \lambda_i \geq 0 \,\, \text{for all $i$}, \sum_{i} \lambda_i =
+  1\}.$$
+  }
 \item
@@
   \textcolor{blue}{
-  Remark 2 now includes that $D$ is the diagonal matrix of the $p$ most positive and $q$ most negative eigenvalues of $A$ (or $P$), $V$ is the matrix of corresponding eigenvectors, and $\hat{Z}$ is the embedding $V |D|^{1/2}$.
-  }
+  We had rewrote Remark~2 to more clearly described the matrices
+  $\hat{Z}$, $V$ and $D$. To avoid confusion, we now use $\hat{D}$ to
+  denote the diagonal matrix whose diagonal entries are the $d = p + q$
+  largest eigenvalues (in modulus) of $A$ and we use $\hat{V}$ to denote
+  the $n \times d$ matrix whose columns are the corresponding
+  eigenvectors. We then use $\hat{Z} = \hat{V} |\hat{D}|^{1/2}$ (where
+  the $|\cdot|$ operation is applied elementwise) to denote the
+  adjacency spectral embedding (ASE) of $A$ into $\mathbb{R}^{d}$; that is
+  to say, the $i$th row of $\hat{Z}$ represents an estimate of the
+  latent position $X_i$ (up to some non-identifiability transformation
+  $Q \in \mathbb{O}(p,q)$). We also reorganized the paper slightly so
+  that the definition and remarks about the non-identifiability of the
+  latent positions $X$ in a GRDPG are presented before discussing ASE. }
 \item
@@
   \textcolor{blue}{
-  The purpose of the additional material in the proof of Theorem 1 is to provide intuition for the arbitrary $K$ case by starting with $K = 2$. 
-  In particular, for the $K = 2$ case, we show that $\tilde{P} = X \Pi X^\top$ where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of order 2, and the rows of $X$ consist of vectors that lie on two 2-dimensional orthogonal subspaces.
-  This then generalizes to larger $K$ such that the rows of $X$ are vectors that lie on $K$ $K$-dimensional orthogonal subspaces, and $\Pi$ is a permutation matrix with $K$ fixed points and $K^2 - K = K (K - 1)$ cycles of order 2.
-  Then the eigenvalues of $\Pi$ are broken up by the following: $K$ $1$'s corresponding to the $K$ fixed points and $K (K - 1)$ pairs of $+1$ and $-1$ corresponding to the $K (K - 1)$ cycles of order 2,
-  resulting in $K (K + 1) / 2$ $+1$'s and $K (K - 1) / 2$ $-1$'s, i.e., the diagonal eigenvalue matrix of $\Pi$ is exactly $I_{K (K + 1) / 2, K (K - 1) / 2}$. 
-  This gives us the embedding dimension ($\mathbb{R}^{K^2}$ broken up into $K (K + 1) / 2$ ``positive'' and $K (K - 1) / 2$ ``negative'' dimensions according to the eigenvalues of $\Pi$) as well as the structure ($K$ $K$-dimensional subspaces as described by the rows of $X$).
-  We have added additional details to the paper to address this. 
+  The purpose of the additional material in the proof of Theorem 1 is to
+  first present the proof in the special case when $K = 2$ so as to
+  build intuition for the general case of $K \geq 2$. 
+  More specificially for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
+  where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of
+  order 2, and the latent vectors lie in the union of two $2$-dimensional
+  orthogonal subspaces, i.e., the rows of $X$ consist of
+  vectors that lie in $\mathcal{S}_1 \cup \mathcal{S}_2$ where
+  $\mathcal{S}_1$ and $\mathcal{S}_2$ are both $2$-dimensional
+  subspaces of $\mathbb{R}^{4}$ with $\mathcal{S}_1 \cap \mathcal{S}_2 = \{0\}$ and $x^{\top} y
+  = 0$ for all $x \in \mathcal{S}_1, y \in \mathcal{S}_2$. 
+  This then generalizes to larger $K$ such that the rows of $X$ are now
+  vectors that lie in the union of $K$ orthogonal subspaces
+  $\mathcal{S}_1 \cup \mathcal{S}_2 \cup \dots \mathcal{S}_K$ where each
+  $\mathcal{S}_k$ is a $K$-dimensional subspace of $\mathbb{R}^{K^2}$,
+  and $\Pi$ is a permutation matrix with 
+  $K$ fixed points and $K (K - 1)$ cycles of order 2. The permutation
+  matrix $\Pi$ then has $K(K-1)/2$ eigenvalues equal to $-1$ (due
+  to the $K(K-1)$ cycles of order $2$) and $K(K+1)/2$ eigenvalues equal
+  to $1$ (due to the $K(K-1)$ cycles of order $2$ together with the $K$
+  fixed points). We can thus write $\Pi = U I_{K(K+1)/2,K(K-1)/2}
+  U^{\top}$. We have revised the paper to more clearly present the above
+  observations. See also Example~1 on page 9 of the revised manuscript
+  and the response to comment $10$ below.
   }
@@
   \textcolor{blue}{
-  The purpose of Example 1 is to illustrate this explicitly. 
-  Comparing matrices $X$ and $Y$, we can see that they share three columns: the first, fifth, and ninth. 
-  The other columns are permuted. 
-  The proof of Theorem 1 then generalizes this to arbitrary $K$. 
+  Example 1 illustrates these fixed points explicitly for $K = 3$. 
+  More specifically, comparing the matrices $X$ and $Y$, we see that
+  they have three columns in common: the first, fifth, and ninth. 
+  The other columns are permuted. The proof of Theorem 1 then
+  generalizes this to arbitrary $K$. In particular by the construction
+  of $X$, each column of $X$ has exactly one non-zero block
+  corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
+  points of $\Pi$ are then the columns with indices $r(K+1) +1 $ for $0 \leq r
+  \leq K-1$; these columns contain the vectors $\lambda_{kk}$ for
+  $1 \leq k \leq K$. The cycles of order $2$ swap the columns of $X$
+  containing the vectors $\lambda_{k \ell}$ and $\lambda_{\ell k}$ for
+  the $\tbinom{K}{2}$ pairs $(k,\ell)$ with $k \not = \ell$. 
   }
@@
   \textcolor{blue}{
-  The goal of estimation is not to estimate $P$ but to identify the community labels and estimate the popularity parameters. 
-  This is not immediately obvious even if we know $P$ beforehand, although having $P$ does make estimation much easier. 
-  In the introduction to section 3 (before 3.1), we have added the following:
-  }\\
-  \textcolor{blue}{
-  ``In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-  To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
-  After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.''
+  We thank the referee for pointing out this issue. We had revised the
+  manuscript to include the line "In our methods, the data that are
+  observed is only the adjacency matrix $A \sim
+  \mathrm{PABM}(\{\lambda^{(k \ell)}\}_K, \rho_n)$ along with an assumed
+  number of communities, $K$. To motivate our methods, we first
+  consider community detection and parameter estimation in the case
+  where we know the edge probability matrix, $P$, beforehand, noting
+  that community memberships and popularity parameters are not
+  immediately discernible from $P$ itself. After establishing methods
+  for community detection and parameter estimation from $P$, we use the
+  consistency property of ASE to demonstrate that the same methods work
+  for $A$ almost surely as $n \to \infty$." Indeed, the assumption of
+  known $P$ helps motivate Algorithm $1$ through Algorithm~3 and the
+  theoretical results in Theorem~2. Theorem 3 and Theorem 4 then follows
+  by leveraging the consistency properties for adjacency spectral embedding
+  to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
+  community recover and paraemters estimation using the true but unknown $P$. 
   }
@@
   \textcolor{blue}{
-  We have added more detail to our description of the SSC algorithm. 
-  In particular, we describe the intuition/motivation behind the algorithm. 
-  }\\
-  \textcolor{blue}{The description of SSC now reads:
-  }\\
-  \textcolor{blue}{
+  We have added more details to our description of SSC to
+  provide further intuition/motivation behind the
+  algorithm. The description of SSC now reads:
   ``The SSC algorithm can be described as follows: 
@@
   consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
-  The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace.
-  Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
-  Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
+  The intuition here is that vectors that lie on the same subspace can
+  be described as linear combinations of each other, assuming the number
+  of vectors in the subspace is greater than the dimensionality of the
+  subspace.
+Thus, for each $c_i$, $c_i^{(j)}$ is zero whenever $x_i$ and $x_j$ belong to different subspaces and may be nonzero otherwise.
 If \(X\) instead represents points near \(K\) subspaces with some noise,
-  then this property may only hold approximately and 
+  then this property will only hold approximately and 
   a final graph partitioning step may be required 
@@
   \textcolor{blue}{
-  There are no references for this, as it is a novel idea. 
-  SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
-  The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
-  }\\
-  \textcolor{blue}{
-  We have added, ``...
-SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''
+  There are no references for this, as it is, to the best of our knowledge, a novel idea. 
+  We emphasize that while SSC has been proposed for community detection
+  for PABM (see e.g., \citep{noroozi2019estimation}), these work apply SSC
+  directly on the rows of $A$ (which are binary vectors in
+  $\mathbb{R}^{n}$) rather than on the ASE of $A$ (which are real-valued
+  vectors in $\mathbb{R}^{d}$ where $d \ll n$). The motivation behind
+  our use of SSC comes  from Theorem 1 wherein we show that the latent
+  vectors for PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
+  by the consistency properties of ASE, the rows of the estimated
+  eigenvectors $\hat{V}$ also lies close to the union of orthogonal
+  subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
+  union of $K$ orthogonal subspaces in $\mathbb{R}^{n}$, the rows of $A$
+  (which are now noisy binary vectors) could be quite far from the
+  corresponding rows of $P$ and hence need not lie close to the union
+  of any $K$ orthogonal subspaces in $\mathbb{R}^{n}$ 
   }
@@
   \textcolor{blue}{
-  We added a third simulation study with $\lambda_{kk} \sim Beta(1,2)$ and $\lambda_{kl} \sim Beta(2,1)$, resulting in an average within-community edge probability of $1/3$ and between-community edge probability of $2/3$.
-  Results of this simulation study are very similar to that of the first two and consistent with the theory.
+  The OSC algorithm works for any PABM parameters setting and is thus
+  agnostic to the particular structure (such as assortativity vs
+  disassortativity). We had revised the manuscript to include an
+  additional simulation demonstrating this claim (see Section 4.3). In
+  particular Section~4.3 considers a PABM setting wherein
+  $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
+  are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$ is a
+  vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
+  random variables. This result in an average within-community edge
+  probability of $1/3$ and a between-community edge
+  probability of $2/3$. The results for this simultation setting are
+  very similar to those presented in the original manuscript and is also
+  consistent with the theoretical results in the paper.
   }\\
   \textcolor{blue}{
-  On a related note, it is our view that the distinction between assortative and disassortative is not very important for the PABM, at least when it is viewed as a type of GRDPG. 
-  In the case of the SBM and DCBM, it can be shown that the rank of $P$ (and therefore the number of embedding dimensions) is equal to $K$, but assortativity or disassortativity determines how many positive and negative eigenvalues $P$ has \citep{rubindelanchy2017statistical}.
-  In the case of the PABM, the number of positive eigenvalues is always $K (K + 1) / 2$ and the number of negative eigenvalues is always $K (K - 1) / 2$, (assuming each community spans its respective subspace).
+  On a related note, it is our view that the distinction between
+  assortative and disassortative is not always meaningful for the
+  PABM. Indeed, unlike the SBM, in the PABM each vertex is free to have a higher affinity to its
+  own community or to other communities. For example, suppose vertex $i$
+  belongs to community $1$. Then the average probability of an edge
+  between vertex $i$ and another vertex in community $1$ is $n^{-1}
+  \rho_n 
+  \lambda_{i1} \sum_{j} \lambda_{j1}$ while the average probability of an edge between vertex $i$ and another vertex in community $k \not =
+  1$ is $n^{-1} \rho_n \lambda_{ik} \sum_{j} \lambda_{j1}$ and thus, on
+  average, vertex $i$ have more affinity with vertices in community $k \not =
+  1$ whenever $\lambda_{i1} < \lambda_{ik}$
   }
@@
   \textcolor{blue}{
-  The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
-  For the Leeds Butterfly dataset, while the paper with which we are comparing \citep{noroozi2019estimation} provides enough details to transform and preprocess the data to match their analysis, they do not provide code to run their clustering algorithm.
-  Therefore, we use their reported benchmark value, which is in the form of ARI. 
-  In order to make an apples-to-apples comparison, we report the ARI for OSC and SSC-ASE here as well. 
-  For all other analyses (simulation and real data), we use misclassification error (whether rate or count). 
+  The first two datasets that we analyzed in section 5 had also been analyzed
+  in previous papers on the PABM. In particular the
+  Leeds Butterfly dataset was analyzed in \citep{noroozi2019estimation}
+  and while \citep{noroozi2019estimation} provides enough details
+  for us to preprocess the data to match their analysis, they do not
+  provide code to run their clustering algorithm (which is based on SSC
+  applied to the rows of $A$). Therefore we used their reported
+  benchmark values, which are given in terms of ARI. 
+  In order to make an apples-to-apples comparison, we also report the ARI for OSC and SSC-ASE here as well. 
+  For all other analyses (simulation and real data), we use the
+  misclassification error rate/count. 
   }
@@
   \textcolor{blue}{
-  Sparsity was not considered in our analyses and algorithms.
+  We have added another simulation in the supplemental materials. 
+  In this simulation, we fix $n = 2048$ and $K = 3$ and vary $\rho \in (0, 1)$. 
   }
@@
   \textcolor{blue}{
-  Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
-  Performance of the algorithms tested is similar to the other two simulation studies and consistent with the theory. 
+  As we discussed in the response to comment
+  $14$, we have added a third simulation to address the issue of disassortativity (see Section
+  4.3 of the revised manuscript). The performance of all the algorithm
+  for the setting in Section~4.3 is qualitatively similar to those
+  presented earlier in Section~4.1 and Section~4.2 and is consistent with the theory. 
   Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$.
@@
   \textcolor{blue}{
-  Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
-  For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
-  For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}.
-  SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
-  When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^3)$. 
-  In practice, we have found that $K$ does not affect runtimes too much. 
+  As for the computational complexity of the algorithms, the main bottleneck for OSC and
+  SSC-ASE is the construction of the affinity matrix $B$. 
+  For OSC, this involves a (truncated) spectral decomposition to extract the
+  $K^2$ largest eigenvalues in modulus (and their corresponding
+  eigenvectors) followed by a matrix multiplication of $n \times K^2$
+  and $K^2 \times n$ matrices, so the time complexity is of the order $O(n^2 K^2)$. 
+  For SSC-ASE (applying SSC on the rows of the adjacency spectral
+  embedding $\hat{V}$) we again start with spectral decomposition, followed by
+  $n$ LASSO problems with a design matrix of size 
+  $K^2 \times (n - 1)$, so the complexity is $O(n^2 K^2 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}. 
+  SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$.
+  When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^2 K^2)$. 
+  In practice the choice of parameter $K$ does not affect
+  runtimes too much. For the simulation studies we only considered $K
+  \leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
+  comparable estimation accuracy as $K \leq 4$. 
   }
@@
   \textcolor{blue}{
-  For the Leeds butterfly dataset, we wanted to compare results of OSC and SSC-ASE to an analysis performed by another group studying the PABM \citep{noroozi2019estimation}. 
-  In their results, they describe how they removed some of the nodes and discretized the edges, which we replicate in our analysis. 
-  }
+  For the Leeds butterfly dataset, we wanted to compare our results
+  using OSC and SSC-ASE to an earlier analysis using SSC on the rows of
+  $A$ (see \citep{noroozi2019estimation}). In their analysis they
+  describe how they removed some of the nodes (corresponding to the
+  other butterfly species) so that the resulting dataset only include the
+  $K = 4$ butterfly species as considered in our paper.}
 \item
@@
   Therefore, the dimensionality of the latent configuration is not in question, only what kind of effect multiplication by unidentifiable $Q$ has on the configuration. 
-  The OSC algorithm circumvents the nonidentifiability of the latent configuration by only considering the entries of $B = n V V^\top$. 
-  SSC-ASE does rely on the latent configuration, but we show that the orthogonality property holds for the particular embedding we perform for this algorithm. 
+  The OSC algorithm circumvents the nonidentifiability of the latent
+  configuration by only considering the entries of $\hat{B} = n \hat{V}
+  \hat{V}^{\top}$ as $n\hat{V}\hat{V}^{\top}$ is, by the asymptotic properties of
+  adjacency spectral embedding, close to $n VV^{\top}$ and $n V
+  V^{\top}$ is always unique (the orthogonal projection onto a column
+  space is always unique). Similarly, SSC-ASE also circumvents this non-identifiability
+  by only using the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
+  $\hat{v}_j$ of $\hat{V}$ belonging to different subspaces, i.e.,
+  different communities, will also
+  have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
   }\\
   \textcolor{blue}{
-  That said, looking back at the latent configuration in Theorem 1, we can see that a $K$-dimensional (not $K^2$-dimensional) latent configuration is sufficient for clustering. 
-  If we are able to obtain this exact embedding, this would greatly reduce the number of dimensions required for the clustering 
-  In practice, this isn't always possible due to the nonidentifiability issue (in particular, determining which of the $K$ out of $K^2$ eigenvectors should be used, although we have found via numerical experiments that community detection is often possible with the $2^{nd}$ to the $(K+1)^{st}$ embedding dimensions).
+  That said, looking back at the latent configuration in Theorem 1, we
+  can see that a $K$-dimensional (not $K^2$-dimensional) embedding is
+  sufficient for clustering, i.e., the embedding corresponding to the
+  diagonal blocks $\lambda_{11}, \lambda_{22}, \dots, \lambda_{KK}$. 
+  If we are able to find this embedding then this would greatly reduce
+  the number of dimensions required for the community
+  recovery/clustering. In practice, this isn't always possible due to
+  the nonidentifiability issue; in particular, determining which of the
+  $K$ out of $K^2$ eigenvectors should be used is the main difficulty.
+  Nevertheless we have found via numerical experiments that community
+  detection is often possible using only the eigenvectors corresponding
+  to the $2^\mathrm{nd}$ to the $(K+1)^\mathrm{st}$ largest eigenvalues
+  of $A$. 
   }
@@
   \textcolor{blue}{
-  We are not entirely clear on why this happens for $K = 2$, but when comparing the normalized Laplacian eigenmaps of the affinity matrix $B$ for various $K$, we see that when $K > 2$, the points appear much closer to Gaussian mixtures, whereas with $K = 2$, there is some separation by community, but they do not appear as Gaussian mixtures, at least for the simulated data.
-  One possible reason for this may be because we failed to find the right hyperparameter for SSC-ASE. 
+  We are not entirely clear on why this happens for $K = 2$, but when
+  comparing the normalized Laplacian eigenmaps of the affinity matrix
+  $B$ for various $K$, we see that when $K > 2$, the points appear much
+  closer to Gaussian mixtures, whereas with $K = 2$, while there is some
+  separation by community, they do not appear as Gaussian mixtures. 
+  One possible reason for this may be because we failed to find the
+  right value for the hyperparameter $\vartheta$ for SSC-ASE. 
   }\\
   \textcolor{blue}{
-  Under ideal circumstances, we would be able to circumvent this issue altogether via the subspace detection property, but SDP does not necessarily guarantee a partitioning of the graph into exactly $K$ disjoint subgraphs but rather {\em at least} $K$ disjoint subgraphs, and in our simulations, we found that setting $\vartheta$ to the appropriate value to obtain SDP often results in greater than $K$ subgraphs, which is why we ended up choosing an ``incorrect'' $\vartheta$ for SSC and then performing the final clustering step.
+  More specifically, under ideal circumstances, we would be able to
+  circumvent the issue of selecting the hyperparameter $\vartheta$ by
+  using the theoretical value that ensures, with high probability, the rows of $\hat{V}$ satisfy the subspace
+  detection property (SDP). However, as other researchers had pointed out (see
+  e.g., \citep{sdp_sufficiency,liu_ssc}, the fact that SDP is satisfied
+  does not guarantee that the affinity matrix $B$ represents a graph
+  with $K$ connected components but rather a graph with {\em at least}
+  $K$ connected component. Indeed, we found in our simulations that
+  setting $\vartheta$ to the theoretical values that guarantees 
+  SDP often results in $B$ having more than $K$ connected components.
+  This is why we choose to do spectral clustering on $B$, i.e., we first
+  compute the normalized Laplacian embeddings using $B$ before
+  clustering the rows of this embedding using Gaussian mixtures
+  modeling. 
   }\\
   \textcolor{blue}{
-  We also note that in our simulations, while SSC-ASE and SSC-A in the $K = 2$ case and MM-Louvain for all $K$ output estimated labels such that the number of mislabeled vertices increases with $n$, the proportion of mislabeled vertices still decreases.
-  For instance, in section 4.1 for, the error count for SSC-ASE for $K = 2$ is around 20 when $n = 128$ and around 100 when $n = 4096$, resulting in an error rate of around $16\%$ for $n = 128$ and $2\%$ for $n = 4096$.
+  Finally we note that in our simulations the proportion of mislabeled
+  vertices does decreases as $n$ increases. 
+  For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
+  vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
+  $n = 4096$; this correspond to an error rate of around $16\%$ and
+  $2\%$, respectively.
   }
@@
   In the application section, we analyze data that are not necessarily generated by the PABM (or any known model). 
-  While all three algorithms are specific to the PABM, they might all behave differently to the different ways in which the model is misspecified. 
-  There is also the question of hyperparameter tuning, especially in the case of SSC-ASE, and we spent a bit more time tweaking the hyperparameters, resulting in performance that is a bit closer to that of MM.
-  }
+  While all three algorithms are specific to the PABM, they might behave
+  differently to the different ways in which the model is
+  misspecified. Finally, we note that Section~5 analyzed three real
+  datasets. For the first dataset (Leeds butterfly) we didn't run the MM
+  algorithm as the original implementation of MM used in 
+  \cite{307cbeb9b1be48299388437423d94bf1} is extremely slow when $K = 4$. For
+  the second dataset (British MP's), the accuracy of MM is slightly
+  better than that of OSC. Finally, for the last dataset (Karantaka
+  villages), the accuracy of MM is slightly worse than that of OSC. }
 \item

### 2022-02-08 referee-response.Rmd
updated referee response
Premerge

@@
 In regard to the probability distribution on the labels $z_1, ...,
-z_n$, none of our theoretical assume anything about the distribution
+z_n$, none of our theoretical results assume anything about the distribution
 of the $\{z_i\}$; we do, however, occasionally assume that the $n_k$
@@
 so that the latent vectors in the $k$th community span its
-subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lies in a
-general position (are linearly independent), then we only require $n_k
+subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lie in 
+general positions then we only require $n_k
 \geq K$.}
@@
 build intuition for the general case of $K \geq 2$. 
-More specificially for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
+More specifically for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
 where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of
@@
 of $X$, each column of $X$ has exactly one non-zero block
-corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
+corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{n_k}$. The fixed
 points of $\Pi$ are then the columns with indices $r(K+1) +1 $ for $0 \leq r
@@
 containing the vectors $\lambda_{k \ell}$ and $\lambda_{\ell k}$ for
-the $\tbinom{K}{2}$ pairs $(k,\ell)$ with $k \not = \ell$. 
+the $\tbinom{K}{2}$ pairs $\{k,\ell\}$ with $k \not = \ell$. 
 }
@@
 consider community detection and parameter estimation in the case
-where we know the edge probability matrix, $P$, beforehand, noting
+where we know the edge probability matrix $P$ beforehand, noting
 that community memberships and popularity parameters are not
@@
 construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
-exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
+exactly on one of \(K\) subspaces then \(B\) describes an undirected graph
 consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
@@
 subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
-union of $K$ orthogonal subspaces in $\mathbb{R}^{n}$, the rows of $A$
+union of $K$ subspaces in $\mathbb{R}^{n}$, the rows of $A$
 (which are now noisy binary vectors) could be quite far from the
 corresponding rows of $P$ and hence need not lie close to the union
-of any $K$ orthogonal subspaces in $\mathbb{R}^{n}$ 
+of any $K$ subspaces in $\mathbb{R}^{n}$ 
 }
@@
 $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
-are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$ is a
+are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
+(for $k \not = \ell$) is a
 vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
@@
 \textcolor{blue}{
-Sparsity was not considered in our analyses and algorithms. (Fix this)
+We have added another simulation in the supplemental materials. 
+In this simulation, we fix $n = 2048$ and $K = 3$ and vary $\rho \in
+(0, 1)$. In particular we see that the performance of all algorithms
+deteriorate significantly as the graphs become sparser. Nevertheless
+it appears that the accuracy of OSC is always comparable and/or better
+than the remaining algorithms for all levels of the sparsity
+considered.
 }
@@
 $14$, we have added a third simulation to address the issue of disassortativity (see Section
-4.3 of the revised manuscript). The performance of all the algorithm
+4.3 of the revised manuscript). The performance of all algorithms
 for the setting in Section~4.3 is qualitatively similar to those
 presented earlier in Section~4.1 and Section~4.2 and is consistent with the theory. 
-Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$.
-Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough.
 }  
@@
 SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$.
-When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^2 K^2)$. 
 In practice the choice of parameter $K$ does not affect
@@
 \leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
-comparable estimation accuracy as $K \leq 4$. 
+comparable estimation accuracy as $K \leq 4$.  Indeed, consider for
+example estimating the entries of the edge probabilities matrix
+$P$. Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of
+$P$ when $K = 4$ is considerably simpler than estimation of $P$ when
+$K = 6$. Cursory numerical experiments suggest that OSC and
+SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
 }
@@
 While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
-\\1. The embedding dimension is $K^2$ with $K (K + 1) / 2$ positive and $K (K - 1) / 2$ negative eigenvalues, 
-\\2.
-The latent structure will always consist of $K$ $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
-\\3. $B = n V V^\top$, where $V$ is the $n \times K^2$ matrix of eigenvectors of $P$, will always be such that $B_{ij} = 0$ if and only if vertices $i$ and $j$ are in different communities. 
-\\4. The rows of $V$ are such that the inner product of the $i^{th}$ and $j^{th}$ rows is zero if and only if vertices $i$ and $j$ are in different communities. 
+\\1. If $P$ is the edge probabilities matrix for a PABM with $K$
+blocks then $P$ is of rank $K^2$ and furthermore $P$ has $K (K + 1) /
+2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues. 
+\\2. Therefore, viewed as a GRDPG, the latent vectors $X$ can be
+represented as vectors in $\mathbb{R}^{K^2}$ and 
+$K^2$ is the smallest possible dimension. Furthermore the latent
+vectors will always lie on the union of $K$ disjoint $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
+\\3. The matrix $B = n V V^\top$, where $V$ is the $n \times K^2$
+matrix of eigenvectors of $P$ (corresponding to the non-zero
+eigenvalues of $P$), is such that $B_{ij} = 0$ if and only
+if vertices $i$ and $j$ are in different communities, i.e.,
+$v_i^{\top} v_j = 0$ if and only if vertices $i$ and $j$ are in
+different communities; here $v_i$ denote the $i$th row of $V$. 
 }  
@@
 V^{\top}$ is always unique (the orthogonal projection onto a column
-space is always unique). Similarly, SSC-ASE also circumvents this non-identifiability
-by only using the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
-$\hat{v}_j$ of $\hat{V}$ belonging to different subspaces, i.e.,
-different communities, will also
+space is always unique). Similarly, SSC-ASE also circumvents the
+non-identifiability of the latent positions
+by using only the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
+$\hat{v}_j$ belonging to different communities will 
 have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
@@
 \textcolor{blue}{
-That said, looking back at the latent configuration in Theorem 1, we
-can see that a $K$-dimensional (not $K^2$-dimensional) embedding is
+That said, if we look at the latent configuration in Theorem 1 then we
+see that a $K$-dimensional (not $K^2$-dimensional) embedding is
 sufficient for clustering, i.e., the embedding corresponding to the
@@
 $K$ out of $K^2$ eigenvectors should be used is the main difficulty.
-Nevertheless we have found via numerical experiments that community
-detection is often possible using only the eigenvectors corresponding
+Nevertheless we have found via numerical experiments that accuracte community
+recovery is generally possible when we use only the eigenvectors corresponding
 to the $2^\mathrm{nd}$ to the $(K+1)^\mathrm{st}$ largest eigenvalues

### 2022-02-08 referee-response.Rmd
fixed minor typos

@@
 to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
-community recover and paraemters estimation using the true but unknown $P$. 
+community recover and parameters estimation using the true but unknown $P$. 
 }
@@
 probability of $1/3$ and a between-community edge
-probability of $2/3$. The results for this simultation setting are
+probability of $2/3$. The results for this simulation setting are
 very similar to those presented in the original manuscript and is also
@@
 $K$ out of $K^2$ eigenvectors should be used is the main difficulty.
-Nevertheless we have found via numerical experiments that accuracte community
+Nevertheless we have found via numerical experiments that accurate community
 recovery is generally possible when we use only the eigenvectors corresponding

### 2022-02-08 referee-response.tex
fixed minor typos

@@
   In regard to the probability distribution on the labels $z_1, ...,
-  z_n$, none of our theoretical assume anything about the distribution
+  z_n$, none of our theoretical results assume anything about the distribution
   of the $\{z_i\}$; we do, however, occasionally assume that the $n_k$
@@
   so that the latent vectors in the $k$th community span its
-  subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lies in a
-  general position (are linearly independent), then we only require $n_k
+  subspace; this is a rather mild requirement, e.g., if the latent vectors in the $k$th community lie in 
+  general positions then we only require $n_k
   \geq K$.}
@@
   build intuition for the general case of $K \geq 2$. 
-  More specificially for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
+  More specifically for $K = 2$ we show that $\tilde{P} = X \Pi X^\top$
   where $\Pi$ is a permutation matrix with 2 fixed points and 1 cycle of
@@
   of $X$, each column of $X$ has exactly one non-zero block
-  corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{k}$. The fixed
+  corresponding to some vector $\lambda_{k \ell} \in \mathbb{R}^{n_k}$. The fixed
   points of $\Pi$ are then the columns with indices $r(K+1) +1 $ for $0 \leq r
@@
   containing the vectors $\lambda_{k \ell}$ and $\lambda_{\ell k}$ for
-  the $\tbinom{K}{2}$ pairs $(k,\ell)$ with $k \not = \ell$. 
+  the $\tbinom{K}{2}$ pairs $\{k,\ell\}$ with $k \not = \ell$. 
   }
@@
   consider community detection and parameter estimation in the case
-  where we know the edge probability matrix, $P$, beforehand, noting
+  where we know the edge probability matrix $P$ beforehand, noting
   that community memberships and popularity parameters are not
@@
   to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
-  community recover and paraemters estimation using the true but unknown $P$. 
+  community recover and parameters estimation using the true but unknown $P$. 
   }
@@
   construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
-  exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
+  exactly on one of \(K\) subspaces then \(B\) describes an undirected graph
   consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
@@
   subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
-  union of $K$ orthogonal subspaces in $\mathbb{R}^{n}$, the rows of $A$
+  union of $K$ subspaces in $\mathbb{R}^{n}$, the rows of $A$
   (which are now noisy binary vectors) could be quite far from the
   corresponding rows of $P$ and hence need not lie close to the union
-  of any $K$ orthogonal subspaces in $\mathbb{R}^{n}$ 
+  of any $K$ subspaces in $\mathbb{R}^{n}$ 
   }
@@
   $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
-  are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$ is a
+  are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
+  (for $k \not = \ell$) is a
   vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
@@
   probability of $1/3$ and a between-community edge
-  probability of $2/3$. The results for this simultation setting are
+  probability of $2/3$. The results for this simulation setting are
   very similar to those presented in the original manuscript and is also
@@
   We have added another simulation in the supplemental materials. 
-  In this simulation, we fix $n = 2048$ and $K = 3$ and vary $\rho \in (0, 1)$. 
+  In this simulation, we fix $n = 2048$ and $K = 3$ and vary $\rho \in
+  (0, 1)$. In particular we see that the performance of all algorithms
+  deteriorate significantly as the graphs become sparser. Nevertheless
+  it appears that the accuracy of OSC is always comparable and/or better
+  than the remaining algorithms for all levels of the sparsity
+  considered.
   }
@@
   $14$, we have added a third simulation to address the issue of disassortativity (see Section
-  4.3 of the revised manuscript). The performance of all the algorithm
+  4.3 of the revised manuscript). The performance of all algorithms
   for the setting in Section~4.3 is qualitatively similar to those
   presented earlier in Section~4.1 and Section~4.2 and is consistent with the theory. 
-  Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$.
-Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough.
   }\\
@@
   SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$.
-  When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^2 K^2)$. 
   In practice the choice of parameter $K$ does not affect
@@
   \leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
-  comparable estimation accuracy as $K \leq 4$. 
+  comparable estimation accuracy as $K \leq 4$.  Indeed, consider for
+  example estimating the entries of the edge probabilities matrix
+  $P$. Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of
+  $P$ when $K = 4$ is considerably simpler than estimation of $P$ when
+  $K = 6$. Cursory numerical experiments suggest that OSC and
+  SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
   }
@@
   While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
-  \\1. The embedding dimension is $K^2$ with $K (K + 1) / 2$ positive and $K (K - 1) / 2$ negative eigenvalues, 
-  \\2.
-The latent structure will always consist of $K$ $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
-  \\3. $B = n V V^\top$, where $V$ is the $n \times K^2$ matrix of eigenvectors of $P$, will always be such that $B_{ij} = 0$ if and only if vertices $i$ and $j$ are in different communities. 
-  \\4. The rows of $V$ are such that the inner product of the $i^{th}$ and $j^{th}$ rows is zero if and only if vertices $i$ and $j$ are in different communities. 
+  \\1. If $P$ is the edge probabilities matrix for a PABM with $K$
+  blocks then $P$ is of rank $K^2$ and furthermore $P$ has $K (K + 1) /
+  2$ positive eigenvalues and $K (K - 1) / 2$ negative eigenvalues. 
+  \\2. Therefore, viewed as a GRDPG, the latent vectors $X$ can be
+  represented as vectors in $\mathbb{R}^{K^2}$ and 
+  $K^2$ is the smallest possible dimension. Furthermore the latent
+  vectors will always lie on the union of $K$ disjoint $K$-dimensional subspaces---in the configuration outlined by Theorem 1, these subspaces are orthogonal, but the multiplication by arbitrary $Q \in \mathbb{O}(p, q)$ that results in the same $P$ may skew the subspaces making them no longer orthogonal (although they will still be $K$ $K$-dimensional subspaces),
+  \\3. The matrix $B = n V V^\top$, where $V$ is the $n \times K^2$
+  matrix of eigenvectors of $P$ (corresponding to the non-zero
+  eigenvalues of $P$), is such that $B_{ij} = 0$ if and only
+  if vertices $i$ and $j$ are in different communities, i.e.,
+  $v_i^{\top} v_j = 0$ if and only if vertices $i$ and $j$ are in
+  different communities; here $v_i$ denote the $i$th row of $V$. 
   }\\
@@
   V^{\top}$ is always unique (the orthogonal projection onto a column
-  space is always unique). Similarly, SSC-ASE also circumvents this non-identifiability
-  by only using the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
-  $\hat{v}_j$ of $\hat{V}$ belonging to different subspaces, i.e.,
-  different communities, will also
+  space is always unique). Similarly, SSC-ASE also circumvents the
+  non-identifiability of the latent positions
+  by using only the rows of $n^{1/2} \hat{V}$ as two rows $\hat{v}_i$ and
+  $\hat{v}_j$ belonging to different communities will 
   have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
@@
   \textcolor{blue}{
-  That said, looking back at the latent configuration in Theorem 1, we
-  can see that a $K$-dimensional (not $K^2$-dimensional) embedding is
+  That said, if we look at the latent configuration in Theorem 1 then we
+  see that a $K$-dimensional (not $K^2$-dimensional) embedding is
   sufficient for clustering, i.e., the embedding corresponding to the
@@
   $K$ out of $K^2$ eigenvectors should be used is the main difficulty.
-  Nevertheless we have found via numerical experiments that community
-  detection is often possible using only the eigenvectors corresponding
+  Nevertheless we have found via numerical experiments that accurate community
+  recovery is generally possible when we use only the eigenvectors corresponding
   to the $2^\mathrm{nd}$ to the $(K+1)^\mathrm{st}$ largest eigenvalues

### 2022-02-08 referee-response.Rmd
fixed minor typos

@@
 \textcolor{blue}{
-We had rewrote Remark~2 to more clearly described the matrices
+We have rewritten Remark~2 to more clearly describe the matrices
 $\hat{Z}$, $V$ and $D$. To avoid confusion, we now use $\hat{D}$ to
@@
 adjacency spectral embedding (ASE) of $A$ into $\mathbb{R}^{d}$; that is
-to say, the $i$th row of $\hat{Z}$ represents an estimate of the
+to say, the $i^{th}$ row of $\hat{Z}$ represents an estimate of the
 latent position $X_i$ (up to some non-identifiability transformation
@@
 We thank the referee for pointing out this issue. We had revised the
-manuscript to include the line "In our methods, the data that are
+manuscript to include the line ``In our methods, the data that are
 observed is only the adjacency matrix $A \sim
@@
 consistency property of ASE to demonstrate that the same methods work
-for $A$ almost surely as $n \to \infty$." Indeed, the assumption of
+for $A$ almost surely as $n \to \infty$.'' Indeed, the assumption of
 known $P$ helps motivate Algorithm $1$ through Algorithm~3 and the
@@
 by leveraging the consistency properties for adjacency spectral embedding
-to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
-community recover and parameters estimation using the true but unknown $P$. 
+to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to community recovery and parameter estimation 
+using the true but unknown $P$. 
 }
@@
 our use of SSC comes  from Theorem 1 wherein we show that the latent
-vectors for PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
+vectors for the PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
 by the consistency properties of ASE, the rows of the estimated
-eigenvectors $\hat{V}$ also lies close to the union of orthogonal
+eigenvectors $\hat{V}$ also lie close to the union of orthogonal
 subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
@@
 agnostic to the particular structure (such as assortativity vs
-disassortativity). We had revised the manuscript to include an
+disassortativity). We have revised the manuscript to include an
 additional simulation demonstrating this claim (see Section 4.3). In
@@
 1$ is $n^{-1} \rho_n \lambda_{ik} \sum_{j} \lambda_{j1}$ and thus, on
-average, vertex $i$ have more affinity with vertices in community $k \not =
+average, vertex $i$ has more affinity with vertices in community $k \not =
 1$ whenever $\lambda_{i1} < \lambda_{ik}$
@@
 While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
-\\1. If $P$ is the edge probabilities matrix for a PABM with $K$
+\\1. If $P$ is the edge probability matrix for a PABM with $K$
 blocks then $P$ is of rank $K^2$ and furthermore $P$ has $K (K + 1) /
@@
 matrix of eigenvectors of $P$ (corresponding to the non-zero
-eigenvalues of $P$), is such that $B_{ij} = 0$ if and only
+eigenvalues of $P$) is such that $B_{ij} = 0$ if and only
 if vertices $i$ and $j$ are in different communities, i.e.,
 $v_i^{\top} v_j = 0$ if and only if vertices $i$ and $j$ are in
-different communities; here $v_i$ denote the $i$th row of $V$. 
+different communities; here $v_i$ denote the $i^{th}$ row of $V$. 
 }  
@@
 $\hat{v}_j$ belonging to different communities will 
-have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
+have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$ 
+(as Theorem 3 states that $n \hat{v}_i^\top \hat{v}_j \stackrel{a.s.}{\to} 0$). 
 }  
@@
 compute the normalized Laplacian embeddings using $B$ before
-clustering the rows of this embedding using Gaussian mixtures
-modeling. 
+clustering the rows of this embedding using Gaussian mixture models. 
 }  
@@
 vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
-$n = 4096$; this correspond to an error rate of around $16\%$ and
+$n = 4096$; this corresponds to an error rate of around $16\%$ and
 $2\%$, respectively.

### 2022-02-09 referee-response.Rmd
Premerge

--- 2022-02-08 referee-response.Rmd
@@
 \textcolor{blue}{
-More specifically, under ideal circumstances, we would be able to
-circumvent the issue of selecting the hyperparameter $\vartheta$ by
-using the theoretical value that ensures, with high probability, the rows of $\hat{V}$ satisfy the subspace
+More specifically, under ideal circumstances, we can set the hyperparameter $\vartheta$ to
+the theoretical values that ensure, with high probability, the rows of $\hat{V}$ satisfy the subspace
 detection property (SDP). However, as other researchers had pointed out (see
@@
 \textcolor{blue}{
-Finally we note that in our simulations the proportion of mislabeled
-vertices does decreases as $n$ increases. 
-For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
-vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
+Finally we emphasize that in our simulations the proportion of mislabeled
+vertices does decreases as $n$ increases. For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
 $n = 4096$; this correspond to an error rate of around $16\%$ and
-$2\%$, respectively.
+$2\%$, respectively. In other words SSC-ASE also works quite well for $K =
+2$ as the proportion of mis-clustered vertices decreases to $0$ as $n$
+increases; it is only when we compared the number of mis-clustered
+vertices that $K = 2$ performs worse compared to $K \geq 3$. 
 }

### 2022-02-09 referee-response.Rmd

--- 2022-02-08 referee-response.Rmd
@@
 \textcolor{blue}{
-Finally we note that in our simulations the proportion of mislabeled
-vertices does decreases as $n$ increases. 
-For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
-vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
-$n = 4096$; this corresponds to an error rate of around $16\%$ and
-$2\%$, respectively.
+Finally we emphasize that in our simulations the proportion of mislabeled
+vertices does decreases as $n$ increases. For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
+$n = 4096$; this correspond to an error rate of around $16\%$ and
+$2\%$, respectively. In other words SSC-ASE also works quite well for $K =
+2$ as the proportion of mis-clustered vertices decreases to $0$ as $n$
+increases; it is only when we compared the number of mis-clustered
+vertices that $K = 2$ performs worse compared to $K \geq 3$. 
 }

### 2022-02-09 referee-response.tex

@@
   \textcolor{blue}{
-  We had rewrote Remark~2 to more clearly described the matrices
+  We have rewritten Remark~2 to more clearly describe the matrices
   $\hat{Z}$, $V$ and $D$. To avoid confusion, we now use $\hat{D}$ to
@@
   adjacency spectral embedding (ASE) of $A$ into $\mathbb{R}^{d}$; that is
-  to say, the $i$th row of $\hat{Z}$ represents an estimate of the
+  to say, the $i^{th}$ row of $\hat{Z}$ represents an estimate of the
   latent position $X_i$ (up to some non-identifiability transformation
@@
   We thank the referee for pointing out this issue. We had revised the
-  manuscript to include the line "In our methods, the data that are
+  manuscript to include the line ``In our methods, the data that are
   observed is only the adjacency matrix $A \sim
@@
   consistency property of ASE to demonstrate that the same methods work
-  for $A$ almost surely as $n \to \infty$." Indeed, the assumption of
+  for $A$ almost surely as $n \to \infty$.'' Indeed, the assumption of
   known $P$ helps motivate Algorithm $1$ through Algorithm~3 and the
@@
   by leveraging the consistency properties for adjacency spectral embedding
-  to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to
-  community recover and parameters estimation using the true but unknown $P$. 
+  to show that community recovery and parameters estimation using $A$ is asymptotically equivalent to community recovery and parameter estimation 
+  using the true but unknown $P$. 
   }
@@
   our use of SSC comes  from Theorem 1 wherein we show that the latent
-  vectors for PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
+  vectors for the PABM lie on the union of orthogonal subspaces in $\mathbb{R}^{d}$ and hence,
   by the consistency properties of ASE, the rows of the estimated
-  eigenvectors $\hat{V}$ also lies close to the union of orthogonal
+  eigenvectors $\hat{V}$ also lie close to the union of orthogonal
   subspaces in $\mathbb{R}^{d}$. In contrast, while the rows of $P$ do lie exactly on the
@@
   agnostic to the particular structure (such as assortativity vs
-  disassortativity). We had revised the manuscript to include an
+  disassortativity). We have revised the manuscript to include an
   additional simulation demonstrating this claim (see Section 4.3). In
@@
   1$ is $n^{-1} \rho_n \lambda_{ik} \sum_{j} \lambda_{j1}$ and thus, on
-  average, vertex $i$ have more affinity with vertices in community $k \not =
+  average, vertex $i$ has more affinity with vertices in community $k \not =
   1$ whenever $\lambda_{i1} < \lambda_{ik}$
@@
   While it is true that there are an infinite number of latent configurations that generate the PABM, based on Theorem 1 and GRDPG results by \citet{rubindelanchy2017statistical}, we can conclusively say the following:
-  \\1. If $P$ is the edge probabilities matrix for a PABM with $K$
+  \\1. If $P$ is the edge probability matrix for a PABM with $K$
   blocks then $P$ is of rank $K^2$ and furthermore $P$ has $K (K + 1) /
@@
   matrix of eigenvectors of $P$ (corresponding to the non-zero
-  eigenvalues of $P$), is such that $B_{ij} = 0$ if and only
+  eigenvalues of $P$) is such that $B_{ij} = 0$ if and only
   if vertices $i$ and $j$ are in different communities, i.e.,
   $v_i^{\top} v_j = 0$ if and only if vertices $i$ and $j$ are in
-  different communities; here $v_i$ denote the $i$th row of $V$. 
+  different communities; here $v_i$ denote the $i^{th}$ row of $V$. 
   }\\
@@
   $\hat{v}_j$ belonging to different communities will 
-  have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$. 
+  have $n \hat{v}_i^{\top} \hat{v}_j \approx 0$ 
+  (as Theorem 3 states that $n \hat{v}_i^\top \hat{v}_j \stackrel{a.s.}{\to} 0$). 
   }\\
@@
   compute the normalized Laplacian embeddings using $B$ before
-  clustering the rows of this embedding using Gaussian mixtures
-  modeling. 
+  clustering the rows of this embedding using Gaussian mixture models. 
   }\\
@@
   vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
-  $n = 4096$; this correspond to an error rate of around $16\%$ and
+  $n = 4096$; this corresponds to an error rate of around $16\%$ and
   $2\%$, respectively.

### 2022-02-09 referee-response.Rmd
changes based on meeting today

@@
 $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
-are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
+are i.i.d. $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
 (for $k \not = \ell$) is a
-vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
-random variables. This result in an average within-community edge
-probability of $1/3$ and a between-community edge
-probability of $2/3$. The results for this simulation setting are
+vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d. $\mathrm{Beta}(2,1)$
+random variables. This result in an average within-community popularity 
+parameter of $1/3$ and a between-community popularity
+parameter of $2/3$. The results for this simulation setting are
 very similar to those presented in the original manuscript and is also
@@
 As we discussed in the response to comment
-$14$, we have added a third simulation to address the issue of disassortativity (see Section
+$16$, we have added a third simulation to address the issue of disassortativity (see Section
 4.3 of the revised manuscript). The performance of all algorithms
@@
 \textcolor{blue}{
-As for the computational complexity of the algorithms, the main bottleneck for OSC and
+For the simulation studies, we only considered $K \leq 4$ as larger $K$ will, 
+in general, require larger values of $n$ to achieve comparable estimation accuracy as lower values of $K$. 
+Indeed, consider for example estimating the entries of the edge probabilities matrix $P$. 
+Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of $P$ when $K = 4$ is considerably simpler than estimation of $P$ when $K = 6$. 
+We have performed some cursory numerical experiments that suggest that OSC does behave similarly to the $K \leq 4$ cases, given sufficient $n$, which is consistent with the theory. 
+}  
+\textcolor{blue}{
+As for the computational complexity of the algorithms, 
+the main bottleneck for OSC and
 SSC-ASE is the construction of the affinity matrix $B$. 
@@
 In practice the choice of parameter $K$ does not affect
-runtimes too much. For the simulation studies we only considered $K
-\leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
-comparable estimation accuracy as $K \leq 4$.  Indeed, consider for
-example estimating the entries of the edge probabilities matrix
-$P$. Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of
-$P$ when $K = 4$ is considerably simpler than estimation of $P$ when
-$K = 6$. Cursory numerical experiments suggest that OSC and
-SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
+runtimes too much in our simulation studies, 
+largely because SSC-A takes up the bulk of the runtime. 
 }

### 2022-02-09 referee-response.tex
changes based on meeting today

@@
   $\lambda_{kk}$ is a vector in $\mathbb{R}^{n_k}$ whose elements
-  are i.i.d $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
+  are i.i.d. $\mathrm{Beta}(1,2)$ random variables and $\lambda_{k\ell}$
   (for $k \not = \ell$) is a
-  vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d $\mathrm{Beta}(2,1)$
-  random variables. This result in an average within-community edge
-  probability of $1/3$ and a between-community edge
-  probability of $2/3$. The results for this simulation setting are
+  vector in $\mathbb{R}^{n_k}$ whose elements are i.i.d. $\mathrm{Beta}(2,1)$
+  random variables. This result in an average within-community popularity 
+  parameter of $1/3$ and a between-community popularity
+  parameter of $2/3$. The results for this simulation setting are
   very similar to those presented in the original manuscript and is also
@@
   As we discussed in the response to comment
-  $14$, we have added a third simulation to address the issue of disassortativity (see Section
+  $16$, we have added a third simulation to address the issue of disassortativity (see Section
   4.3 of the revised manuscript). The performance of all algorithms
@@
   \textcolor{blue}{
-  As for the computational complexity of the algorithms, the main bottleneck for OSC and
+  For the simulation studies, we only considered $K \leq 4$ as larger $K$ will, 
+  in general, require larger values of $n$ to achieve comparable estimation accuracy as lower values of $K$. 
+  Indeed, consider for example estimating the entries of the edge probabilities matrix $P$. 
+  Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of $P$ when $K = 4$ is considerably simpler than estimation of $P$ when $K = 6$. 
+  We have performed some cursory numerical experiments that suggest that OSC does behave similarly to the $K \leq 4$ cases, given sufficient $n$, which is consistent with the theory. 
+  }\\
+  \textcolor{blue}{
+  As for the computational complexity of the algorithms, 
+  the main bottleneck for OSC and
   SSC-ASE is the construction of the affinity matrix $B$. 
@@
   In practice the choice of parameter $K$ does not affect
-  runtimes too much. For the simulation studies we only considered $K
-  \leq 4$ as values of $K \geq 5$ will in general, require larger values of $n$ to achieve 
-  comparable estimation accuracy as $K \leq 4$.  Indeed, consider for
-  example estimating the entries of the edge probabilities matrix
-  $P$. Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of
-  $P$ when $K = 4$ is considerably simpler than estimation of $P$ when
-  $K = 6$. Cursory numerical experiments suggest that OSC and
-  SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
+  runtimes too much in our simulation studies, 
+  largely because SSC-A takes up the bulk of the runtime. 
   }
@@
   \textcolor{blue}{
-  More specifically, under ideal circumstances, we would be able to
-  circumvent the issue of selecting the hyperparameter $\vartheta$ by
-  using the theoretical value that ensures, with high probability, the rows of $\hat{V}$ satisfy the subspace
+  More specifically, under ideal circumstances, we can set the hyperparameter $\vartheta$ to
+  the theoretical values that ensure, with high probability, the rows of $\hat{V}$ satisfy the subspace
   detection property (SDP). However, as other researchers had pointed out (see
@@
   \textcolor{blue}{
-  Finally we note that in our simulations the proportion of mislabeled
-  vertices does decreases as $n$ increases. 
-  For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered
-  vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
-  $n = 4096$; this corresponds to an error rate of around $16\%$ and
-  $2\%$, respectively.
+  Finally we emphasize that in our simulations the proportion of mislabeled
+  vertices does decreases as $n$ increases. For instance, in section 4.1, if $K = 2$ then the average number of mis-clustered vertices for SSC-ASE is around 20 when $n = 128$ and around 100 when
+  $n = 4096$; this correspond to an error rate of around $16\%$ and
+  $2\%$, respectively. In other words SSC-ASE also works quite well for $K =
+  2$ as the proportion of mis-clustered vertices decreases to $0$ as $n$
+  increases; it is only when we compared the number of mis-clustered
+  vertices that $K = 2$ performs worse compared to $K \geq 3$. 
   }

### 2022-02-11 referee-response.Rmd
Premerge

--- 2022-02-09 referee-response.Rmd
@@
 deteriorate significantly as the graphs become sparser. Nevertheless
-it appears that the accuracy of OSC is always comparable and/or better
-than the remaining algorithms for all levels of the sparsity
+the accuracy of OSC is always comparable and/or better than the remaining algorithms for all levels of the sparsity
 considered.
@@
 $P$ when $K = 4$ is considerably simpler than estimation of $P$ when
-$K = 6$. Cursory numerical experiments suggest that OSC and
+$K = 8$. Cursory numerical experiments suggest that OSC and
 SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 

### 2022-02-11 referee-response.Rmd

--- 2022-02-09 referee-response.Rmd
@@
 For the simulation studies, we only considered $K \leq 4$ as larger $K$ will, 
-in general, require larger values of $n$ to achieve comparable estimation accuracy as lower values of $K$. 
-Indeed, consider for example estimating the entries of the edge probabilities matrix $P$. 
-Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of $P$ when $K = 4$ is considerably simpler than estimation of $P$ when $K = 6$. 
-We have performed some cursory numerical experiments that suggest that OSC does behave similarly to the $K \leq 4$ cases, given sufficient $n$, which is consistent with the theory. 
+in general, require larger values of $n$ to achieve comparable estimation accuracy as lower values of $K$ and SSC-A 
+(where SSC is applied to the rows of the adjacency matrix) takes a very long time to run when $n$ is large  (see the discussion about computational complexity below). 
+Our theoretical results are limit results and work for any choice of $K$ as long as $n$ is large.
+The reason for needing large $n$ when $K$ increases is that the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of $P$ when $K$ is say $4$ is considerably simpler than estimation of $P$ when $K$ is say $8$.
+We have performed some cursory numerical experiments and found that OSC and SSC-ASE for $K = 8$ and $n = 4096$ behave similarly to $K \leq 4$ and $n = 1024$.
 }  
@@
 SSC-A (applying SSC on the rows of $A$) as proposed in \cite{noroozi2019estimation} involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$.
-In practice the choice of parameter $K$ does not affect
-runtimes too much in our simulation studies, 
+In practice the choice of parameter $K$ does not affect runtimes too much in our simulation studies, 
 largely because SSC-A takes up the bulk of the runtime. 

### 2022-02-11 referee-response.Rmd
Minor changes

--- 2022-02-09 referee-response.Rmd
@@
 \textcolor{blue}{
-We thank the referee for the feedback. 
-The concerns addressed here are answered below. 
-In particular, we made considerable effort to address the referees' concerns about the simulation studies. 
+We thank the referee for the feedback. The referee's comments are
+answered below. 
+In particular, we made considerable effort to address the referee's concerns about the simulation studies. 
 }
@@
 \textcolor{blue}{
-For the simulation studies, we only considered $K \leq 4$ as larger $K$ will, 
-in general, require larger values of $n$ to achieve comparable estimation accuracy as lower values of $K$. 
-Indeed, consider for example estimating the entries of the edge probabilities matrix $P$. 
-Then the rank of $P$ is equal to $K^2$ and hence, for a given $n$, estimation of $P$ when $K = 4$ is considerably simpler than estimation of $P$ when $K = 6$. 
-We have performed some cursory numerical experiments that suggest that OSC does behave similarly to the $K \leq 4$ cases, given sufficient $n$, which is consistent with the theory. 
+For the simulation studies, we only considered $K \leq 4$ as larger
+$K$ will, in general, require larger values of $n$ to achieve
+comparable estimation accuracy as lower values of $K$. The running
+time for a simulation studies when $n$ is large then becomes quite
+prohibitive, and this is due mainly to the running-time of the SSC-A
+algorithm (see the discussion about computationaly
+complexity); the running time of the SSC-ASE and OSC is still
+reasonable when we increases $n$. We believe that our current simulation
+studies (including the new section 4.3) is sufficiently
+comprehensive to illustrate our theoretical results results; indeed
+our theoretical results are large-sample results and is valid for any
+choice of $K$. The only reason for needing large
+$n$ when $K$ increases is that the rank of $P$ is equal to $K^2$ and
+hence, for a given $n$, estimation of $P$ when $K$ is say $4$ is
+considerably simpler than estimation of $P$ when $K$ is say $8$. 
 }  
 \textcolor{blue}{
-As for the computational complexity of the algorithms, 
+For the computational complexity of the algorithms, 
 the main bottleneck for OSC and
@@
 the theoretical values that ensure, with high probability, the rows of $\hat{V}$ satisfy the subspace
-detection property (SDP). However, as other researchers had pointed out (see
-e.g., \citep{sdp_sufficiency,liu_ssc}, the fact that SDP is satisfied
+detection property (SDP). However, as other researchers had pointed
+out that, even in the context of sparse subspace clustering where the points lie
+exactly on a union of $K$ subspaces, the fact that SDP is satisfied
 does not guarantee that the affinity matrix $B$ represents a graph
 with $K$ connected components but rather a graph with {\em at least}
-$K$ connected component. Indeed, we found in our simulations that
+$K$ connected component (see e.g., \citep{liu_ssc,sdp_sufficiency,}).
+Indeed, we found in our simulations that
 setting $\vartheta$ to the theoretical values that guarantees 

### 2022-02-11 summary_minh_edited.tex
minor update referencing example 1 in proof of theorem 1
Premerge
minor typos

--- 2022-02-06 summary_minh_edited.tex
@@
 For the general case, we can extend $\tilde{P} = X \Pi X^\top$ to larger $K$. 
+For a more concrete example of this, refer to Example~1. 
 We once again consider $\tilde{P}$ as defined in
@@
 We first define the following matrices
+
 \begin{gather}
@@
 \end{gather}
+
 It is then straightforward to verify that
+
 \begin{gather*}
@@
 In our methods, the data that are observed for estimation is the adjacency matrix, $A \sim \PABM(\{\lambda^{(k \ell)}\}_K, \rho_n)$, along with an assumed number of communities, $K$. 
-To motivate our methods, we also consider community detection and parameter estimation in the case where we know the edge probability matrix, $P$, beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
+To motivate our methods, we first consider community detection and parameter estimation in the case where we know the edge probability matrix $P$ beforehand, noting that community memberships and popularity parameters are not immediately discernible from $P$ itself.
 After establishing methods for community detection and parameter estimation from $P$, we use the consistency property of the ASE \citep{doi:10.1080/01621459.2012.699795,rubindelanchy2017statistical} to demonstrate that the same methods work for $A$ almost surely as $n \to \infty$.
@@
 In practice, due to presence of noise, SSC is often done by solving the LASSO problems
-
 \begin{equation} \label{eq:ssc}
@@
 \end{equation}
-
 for some sparsity parameter \(\vartheta > 0\). 
@@
 such that \(P = Y I_{p, q} Y^\top\). Then it is generally the case that
-we  \( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
+\( Y = \tilde{\Pi} XU Q \) for some indefinite orthogonal matrix
 $Q \in \mathbb{O}(p,q)$. 

### 2022-02-21 simulations-and-data-analysis/cluster-sim-balanced.R
reorganizing some files

@@ -64,7 +64,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
 
 gc()
 
-readr::write_csv(clustering.df, 'clustering-k.csv')
+readr::write_csv(clustering.df, 'simulation-results/cluster-sim-balanced.csv')
 
 clustering.df %>%
   dplyr::group_by(n, K) %>%
@@ -87,6 +87,7 @@ clustering.df %>%
   theme_bw() + 
   theme(text = element_text(size = 10)) + 
   scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
+  scale_y_log10() + 
   labs(y = 'community detection error count', 
        colour = NULL, shape = NULL) +
   geom_line(aes(x = n, y = med.err * n,

### 2022-02-21 simulations-and-data-analysis/cluster-sim-disassortative.R

@@ -62,7 +62,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
 
 gc()
 
-readr::write_csv(clustering.df, 'clustering-k-disassortative.csv')
+readr::write_csv(clustering.df, 'simulation-results/clustering-k-disassortative.csv')
 
 clustering.df %>%
   dplyr::group_by(n, K) %>%

### 2022-02-21 simulations-and-data-analysis/cluster-sim-imbalanced.R

@@ -64,7 +64,7 @@ clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
 
 gc()
 
-# readr::write_csv(clustering.df, 'clustering-k-imbalanced.csv')
+readr::write_csv(clustering.df, 'simulation-results/cluster-sim-imbalanced.csv')
 
 clustering.df %>%
   dplyr::group_by(n, K) %>%

### 2022-02-21 simulations-and-data-analysis/param-est-disassortative.R

@@ -46,7 +46,7 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   return(out)
 }
 
-readr::write_csv(rmse.df, 'rmse-k-disassortative.csv')
+readr::write_csv(rmse.df, 'simulation-results/param-est-disassortative.csv')
 
 rmse.df %>% 
   na.omit() %>% 

### 2022-02-21 simulations-and-data-analysis/param-est-sim-balanced.R

@@ -45,7 +45,7 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   return(out)
 }
 
-readr::write_csv(rmse.df, 'rmse-k.csv')
+readr::write_csv(rmse.df, 'simulation-results/param-est-sim-balanced.csv')
 
 rmse.df %>% 
   na.omit() %>% 

### 2022-02-21 simulations-and-data-analysis/param-est-sim-imbalanced.R

@@ -46,7 +46,7 @@ rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
   return(out)
 }
 
-readr::write_csv(rmse.df, 'rmse-k-imbalanced.csv')
+readr::write_csv(rmse.df, 'simulation-results/param-estim-sim-imbalanced.csv')
 
 rmse.df %>% 
   na.omit() %>% 

### 2022-02-22 summary_minh_edited.tex
minor adjustments for submission

@@
 
-\item[Code:] The code is not uploaded here but can be found online at \\
+\item[Code:] The code for simulations and data anlyses can be found in code-and-results.zip. 
+This file also includes the simulation results in csv format (please note that the simulations can take a long time to complete). 
+The code for data analyses requires the files found in data.zip. 
+All of these files can also be found at\\
 \url{https://github.com/johneverettkoo/pabm-grdpg}.
+
+\item[Sparsity:] Additional supplemental materials regarding simulations for the effect of the sparsity parameter can be found in sparsity-sim.zip.
 

### 2022-03-23 fls-lightning-talk/slides.Rmd
lightning talk slides

--- 2021-11-27 s675/blockmodels-and-rdpg.Rmd
@@
 ---
+title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
+subtitle: Future Leaders Summit 2022 Lightning Presentation
+author: John Koo, Indiana University
+date: 'April 2022'
 output: 
@@
     includes:
-      in_header: 
-        - page_headers.tex
+      in_header: page_headers.tex
 header-includes:
@@
 - \usepackage{comment}
+- \usepackage{graphicx}
+- \setbeamertemplate{itemize items}[circle]
 ---
@@
 
+source('~/dev/pabm-grdpg/functions.R')
 set.seed(314159)
-source('~/dev/pabm-grdpg/functions.R')
 ```
 
-## {.plain}
+## Contributors
 
-\center
+::: columns
 
-\LARGE
+:::: {.column width='33%'}
 
-\textcolor{darkred}{Block Models\\and\\(Generalized) Random Dot Product Graphs}
+```{r, out.width = '100px'}
+knitr::include_graphics('john-koo.jpeg')
+```
 
-\normalsize
+John Koo,  
+PhD Student in Statistical Science,  
+Indiana University
 
-STAT-S 675
+::::
 
-Fall 2021
+:::: {.column width='33%'}
 
-# Recall From Last Time ...
+```{r, out.width = '100px'}
+knitr::include_graphics('minh-tang.jpg')
+```
+
+Minh Tang,  
+Assistant Professor of Statistics,  
+NC State University
+
+::::
+
+:::: {.column width='33%'}
+
+```{r, out.width = '100px'}
+knitr::include_graphics('michael-trosset.jpg')
+```
+
+Michael Trosset,  
+Professor of Statistics,  
+Indiana University
+
+::::
+
+:::
+
+## Community Detection for Networks
 
@@
 \newcommand{\PABM}{\text{PABM}}
-\newcommand{\argmin}{\mathrm{argmin}}
 
-* Let $G = (V, E)$ be an undirected and hollow graph with $|V| = n$ 
-and adjacency matrix $A$
-  * $A \in \mathbb{R}^{n \times n}$ is symmetric with zero diagonals
-* Suppose $G \sim F(\theta)$
-  * What kind of $F(\theta)$ make sense here?
-  * Given $F$ and observed $G$, how can we estimate $\theta$?
+```{r out.width = '50%', fig.height = 3, fig.width = 4}
+Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
+P <- Pz$P
+z <- Pz$clustering
+A <- draw.graph(P)
+qgraph::qgraph(A, groups = factor(z), legend = FALSE)
+```
 
-# Recall From Last Time ...
+How can we cluster the nodes of a network?
+
+Statistical inference (parametric approach):
+
+1. Define a generative model for graph: 
+$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
+2. Develop a method for obtaining estimators: 
+$f(G) = (\hat{\vec{z}}, \hat{\vec{\theta}})$.
+3. Identify asymptotic properties of estimators: 
+$(\hat{\vec{z}}, \hat{\vec{\theta}}) \to (\vec{z}, \vec{\theta})$.
+
+## Overview
+
+<style type="text/css">
+.caption {
+    font-size: x-small;
+}
+</style>
+
+1. Block Models and the Popularity Adjusted Block Model
+
+2. Generalized Random Dot Product Graphs
+
+3. Connecting the PABM to the GRDPG
+
+4. Community Detection for the PABM
+
+# Block Models
+
+## Bernoulli Graphs
 
@@
 
-:::: {.column width=55%}
+:::: {.column width=62.5%}
 
+Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.
+
+$G$ is described by adjacency matrix $A$ such that
 $A_{ij} = \begin{cases} 
@@
 
-For estimation, we need to impose some structure on $P$.
-
 ::::
 
-:::: {.column width=45%}
+:::: {.column width=37.5%}
 
-**Example 1**: If every entry $P_{ij} = \theta \in (0, 1)$, 
-then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
+**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
 For this model, $A_{ij} \iid \Bernoulli(\theta)$.
@@
 n <- 2 ** 5
-p <- 1 / 5
+p <- 1 / log(n)
 P <- matrix(p, nrow = n, ncol = n)
@@
 
-# Recall From Last Time ...
+## Block Models
 
-Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
+Suppose each vertex $v_1, ..., v_n$ has labels $z_1, ..., z_n \in \{1, ..., K\}$,  
 and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
@@
 
-# Recall From Last Time ...
+## Block Models
 
@@
 
-# Recall From Last Time ...
+## Popularity Adjusted Block Model
+
+**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
+
+Let each vertex $i \in [n]$ have $K$ popularity parameters
+$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
+Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
+e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
+
+**Lemma** (Noroozi, Rimal, and Pensky, 2020):
+
+$A$ is sampled from a PABM if $P$ can be described as:
+
+1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
+between communities $k$ and $l$. 
+2. Organize popularity parameters as vectors 
+$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
+such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
+of the $i$^th^ vertex of community $k$ towards community $l$. 
+3. Each block can be decomposed as 
+$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
+
+**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.
+
+# Generalized Random Dot Product Graphs
+
+## (Generalized) Random Dot Product Graph Model
 
@@
 
-# Recall From Last Time ...
+If latent vectors $X_1, ..., X_n \iid F$, then we write 
+$(A, X) \sim \RDPG(F, n)$ or $(A, X) \sim \GRDPG_{p, q}(F, n)$.
+
+## (Generalized) Random Dot Product Graph Model
+
+### Recovery/Estimation
+
+Want to estimate $X$ from $A$, or alternatively, 
+interpoint distances, inner products, or angles.
 
@@
 
-RDPG: $\max\limits_i \|\hat{x}_i - W x_i \| \stackrel{a.s.}{\to} 0$
+RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
 (Athreya et al., 2018)  
 GRDPG: 
-$\max\limits_i \|\hat{x}_i - Q x_i \| \stackrel{a.s.}{\to} 0$
+$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
 (Rubin-Delanchy et al., 2020)
 
-# Connecting Block Models to (G)RDPGs
+# Connecting the PABM to the GRDPG
+
+## Connecting Block Models to the (G)RDPG Model
 
@@
 
-# Connecting the SBM to the RDPG
+## Connecting Block Models to the (G)RDPG Model
 
@@
 
-# Connecting the DCBM to the RDPG
-
-**Example 3**: DCBM with two communities
+## Connecting Block Models to the (G)RDPG Model
 
@@
 
-* $z_1, ..., z_n \in \{1, 2\}$
-* $P_{ij} = \begin{cases} 
-p \omega_i \omega_j & z_i, z_j = 1 \\
-q \omega_i \omega_j & z_i, z_j = 2 \\
-r \omega_i \omega_j & z_i \neq z_j
-\end{cases}$
+```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
+par(mar = rep(1.75, 4))
+plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
+```
 
-* To make this an assortative DCBM, set $p q > r^2$.
-* In this example, $p = q = 1/4$, $r = 1/2$, and $\omega_i \in (0, 1)$.
+\vspace*{.5\baselineskip}
+
+```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
+par(mar = rep(1.75, 4))
+omega <- rbeta(n, 1, 1)
+dc.matrix <- omega %*% t(omega)
+P.dcbm <- P * dc.matrix
+dcbm.eigen <- eigen(P.dcbm)
+X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
+plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')
+
+```
 
@@
 
-```{r, fig.height = 4, fig.width = 4, out.width = '100%'}
-n1 <- 100
-n2 <- 100
-n <- n1 + n2
-z <- c(rep(1, n1), rep(2, n2))
-p <- 1/4
-q <- 1/4
-r <- 1/8
-omega <- rbeta(n, 2, 1)
-Omega <- tcrossprod(omega)
-P <- matrix(r, nrow = n, ncol = n)
-P[seq(n1), seq(n1)] <- p
-P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
-P <- P * Omega
-A <- draw.graph(P)
-qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
+\vspace*{0\baselineskip}
+
+```{r, out.width = '100%', fig.height = 4, fig.width = 4}
+par(mar = rep(1.75, 4))
+Pz <- generate.P.beta(n * 10)
+P <- Pz$P
+z <- Pz$clustering
+X <- embedding(P, 3, 1)
+pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
 ```
@@
 
-# Connecting the DCBM to the RDPG
+## Connecting the PABM to the GRDPG
 
-**Example 3** (cont'd): Assortative DCBM ($p q > r^2$) with $K = 2$
-
-::: columns
-
-:::: column
-
-$$P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
-\end{cases}$$
-
-```{r fig.height = 5, fig.width = 6, out.width = '100%'}
-qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
-```
-
-::::
-
-:::: column
-
-$$P = 
-\begin{bmatrix} 
-P^{(11)} & P^{(12)} \\
-P^{(21)} & P^{(22)}
-\end{bmatrix} =
-X X^\top$$
-
-$$X = \begin{bmatrix} 
-\sqrt{p} \omega_1 & 0 \\
-\vdots & \vdots \\
-\sqrt{p} \omega_{n_1} & 0 \\
-\sqrt{\frac{r^2}{p}} \omega_{n_1 + 1} & \sqrt{q - \frac{r^2}{p}} \omega_{n_1 + 1} \\
-\vdots & \vdots \\
-\sqrt{\frac{r^2}{p}} \omega_n & \sqrt{q - \frac{r^2}{p}} \omega_n
-\end{bmatrix}$$
-
-::::
-
-:::
-
-# Connecting the SBM to the RDPG
-
-**Example 3** (cont'd): If we want to perform community detection,
-
-1. Note that $A$ is a RDPG because $P = X X^\top$.
-2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
-$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
-3. Apply clustering algorithm on $\hat{X}$,  
-noting that as $n \to \infty$, the ASE approaches line segments.
-
-```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
-P.eigen = eigen(P)
-X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
-
-A.eigen <- eigen(A)
-X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
-plot(X.hat, 
-     # asp = 1, 
-     col = z * 2, 
-     xlab = NA, ylab = NA, 
-     main = 'ASE of the adjacency matrix drawn from DCBM')
-lines(rbind(c(0, 0), X[z == 1, ]), pch = 16, col = 2, lwd = 3)
-lines(rbind(c(0, 0), X[z == 2, ]), pch = 16, col = 4, lwd = 3)
-```
-
-# Assortative SBM and DCBM are RDPGs
-
-* $P$ is positive semidefinite $\implies$ RDPG
-* Number of embedding dimensions $=$ number of communities
-* Latent structure for the SBM are point masses,  
-latent structure for the DCBM are line segments
-* If the SBM or DCBM is not assortative, then $P$ is not positive semidefinite
-  * GRDPG instead of RDPG
-  * Number of embedding dimensions $=$ number of communities, 
-  except split by positive and negative eigenvalues
-  * ASE still converges to true latent positions up to multiplication by unidentifiable $Q \in \mathbb{O}(p, q)$
-  * Multiplication by $Q$ doesn't change overall latent structure
-
-# Demo
-
-# Connecting the PABM to the GRDPG
-
-**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
-
-Let each vertex $i \in [n]$ have $K$ popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
-Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
-e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
-
-**Lemma** (Noroozi, Rimal, and Pensky, 2020):
-
-$A$ is sampled from a PABM if $P$ can be described as:
-
-1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
-between communities $k$ and $l$. 
-2. Organize popularity parameters as vectors 
-$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
-such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
-of the $i$^th^ vertex of community $k$ towards community $l$. 
-3. Each block can be decomposed as 
-$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
-
-**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.
-
-# Connecting the PABM to the GRDPG
-
-**Theorem**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
+**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
 $A \sim \GRDPG_{p, q}(X U)$ with
@@
 
-# Connecting the PABM to the GRDPG
+## Connecting the PABM to the GRDPG
 
@@
 
-# Orthogonal Spectral Clustering
+# Community Detection for the PABM
+
+## Orthogonal Spectral Clustering
 
@@
 $$
+
+## Sparse Subspace Clustering
+
+**Corollary**: The ASE of $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ lies near a collection of $K$-dimensional subspaces in $K^2$ dimensions.
+
+**Algorithm**: Sparse Subspace Clustering (Elhamifar & Vidal, 2009):
+
+1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
+subject to $x_i = X^\top c$ and $c^{(i)} = 0$.  
+This is typically performed via LASSO: 
+$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i}^\top c\|_2^2 + \lambda \|c\|_1$
+2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
+3. Construct affinity matrix $B = |C| + |C^\top|$.
+
+## Sparse Subspace Clustering
+
+**Theorem** (KTT):
+
+Let
+
+* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
+$A_n \sim \BG(P_n)$;
+* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
+$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.
+
+Then
+
+* For some $\lambda > 0$ and $N < \infty$, $\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1 when $n > N$.
+
+Remarks:
+
+* For large $n$, we can identify $\lambda$ for SDP (Wang and Xu, 2016).
+* SDP does not guarantee community detection.
+
+## Simulation Results
+
+```{r, fig.height = 3, fig.width = 8, out.width = '100%'}
+clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')
+
+clustering.df %>%
+  dplyr::group_by(n, K) %>%
+  dplyr::summarise(
+    med.err = median(error.osc),
+    first.q = quantile(error.osc, .25),
+    third.q = quantile(error.osc, .75),
+    med.err.ssc = median(error.ssc.ase),
+    first.q.ssc = quantile(error.ssc.ase, .25),
+    third.q.ssc = quantile(error.ssc.ase, .75),
+    med.err.louvain = median(error.louvain),
+    first.q.louvain = quantile(error.louvain, .25),
+    third.q.louvain = quantile(error.louvain, .75),
+    med.err.ssc.A = median(error.ssc.A),
+    first.q.ssc.A = quantile(error.ssc.A, .25),
+    third.q.ssc.A = quantile(error.ssc.A, .75)
+  ) %>% 
+  dplyr::ungroup() %>% 
+  ggplot() +
+  theme_bw() + 
+  theme(legend.position = 'bottom') + 
+  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
+                labels = c(expression(2^7), 
+                           expression(2^8), 
+                           expression(2^9), 
+                           expression(2^10), 
+                           expression(2^11), 
+                           expression(2^12))) +
+  scale_y_log10() + 
+  labs(y = 'community detection error count', 
+       colour = NULL, shape = NULL) +
+  geom_line(aes(x = n, y = med.err * n,
+                colour = 'OSC')) +
+  geom_point(aes(x = n, y = med.err * n,
+                 colour = 'OSC', shape = 'OSC')) +
+  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
+                    colour = 'OSC'), width = .1) + 
+  geom_line(aes(x = n, y = med.err.ssc * n,
+                colour = 'SSC-ASE')) +
+  geom_point(aes(x = n, y = med.err.ssc * n,
+                 colour = 'SSC-ASE', shape = 'SSC-ASE')) +
+  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
+                    colour = 'SSC-ASE'), width = .1) +
+  geom_line(aes(x = n, y = med.err.ssc.A * n,
+                colour = 'SSC-A')) +
+  geom_point(aes(x = n, y = med.err.ssc.A * n,
+                 colour = 'SSC-A', shape = 'SSC-A')) +
+  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
+                    colour = 'SSC-A'), width = .1) +
+  geom_line(aes(x = n, y = med.err.louvain * n,
+                colour = 'MM-Louvain')) + 
+  geom_point(aes(x = n, y = med.err.louvain * n,
+                 colour = 'MM-Louvain', shape = 'MM-Louvain')) + 
+  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
+                    colour = 'MM-Louvain'), width = .1) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  facet_wrap(~ K, labeller = 'label_both')
+```

### 2022-03-24 fls-lightning-talk/slides.Rmd
minor tweak
updated slides per discussion at meeting

@@
 $f(G) = (\hat{\vec{z}}, \hat{\vec{\theta}})$.
-3. Identify asymptotic properties of estimators: 
+3. Describe asymptotic properties of estimators: 
 $(\hat{\vec{z}}, \hat{\vec{\theta}}) \to (\vec{z}, \vec{\theta})$.
-
-## Overview
-
-<style type="text/css">
-.caption {
-    font-size: x-small;
-}
-</style>
-
-1. Block Models and the Popularity Adjusted Block Model
-
-2. Generalized Random Dot Product Graphs
-
-3. Connecting the PABM to the GRDPG
-
-4. Community Detection for the PABM
-
-# Block Models
 
@@
 
-## Block Models
-
-Erdos-Renyi Model (1959)
-
-* $P_{ij} = \theta$ (not a block model)
-* 1 parameter $\theta$
-
-Stochastic Block Model (Lorrain and White, 1971)
-
-* $P_{ij} = \theta_{z_i z_j}$
-* $K (K + 1) / 2$ parameters $\theta_{kl}$
-
-Degree Corrected Block Model (Karrer and Newman, 2011)
-
-* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
-* $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$
-
-Popularity Adjusted Block Model (Sengupta and Chen, 2017)
-
-* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
-* $K n$ parameters $\lambda_{ik}$
-
 ## Popularity Adjusted Block Model
@@
 $\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
-Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
+Then $A \sim \PABM(\{\lambda_{ik}\}_K)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
 e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
@@
 
-**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.
+## Generalized Random Dot Product Graph
 
-# Generalized Random Dot Product Graphs
+**Def** Generalized Random Dot Product Graph  
+(Rubin-Delanchy, Cape, Tang, Priebe, 2020)
 
-## (Generalized) Random Dot Product Graph Model
-
-Random Dot Product Graph $A \sim \RDPG(X)$  
-(Young and Scheinerman, 2007)
-
-* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
-$x_i^\top x_j \in [0, 1]$
-* $A \sim \BG(X X^\top)$ where 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
-
-Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
-(Rubin-Delanchy, Cape, Tang, Priebe, 2020)
+$A \sim \GRDPG_{p, q}(X)$ iff
 
@@
 $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
-
-If latent vectors $X_1, ..., X_n \iid F$, then we write 
-$(A, X) \sim \RDPG(F, n)$ or $(A, X) \sim \GRDPG_{p, q}(F, n)$.
+  * $P(\text{edge between } v_i, v_j) = x_i^\top I_{p,q} x_j$
 
@@
 
-To embed in $\mathbb{R}^d$,
+To embed in $\mathbb{R}^{p+q}$,
 
 1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
-where $\hat{\Lambda} \in \mathbb{R}^{d \times d}$ and 
-$\hat{V} \in \mathbb{R}^{n \times d}$.  
-For RDPG, use $d$ greatest eigenvalues; 
-for GRDPG, use $p$ most positive and $q$ most negative eigenvalues.
+where $\hat{\Lambda} \in \mathbb{R}^{(p+q) \times (p+q)}$ and 
+$\hat{V} \in \mathbb{R}^{n \times (p+q)}$ 
+by using the $p$ most positive and $q$ most negative eigenvalues.
 
-2. For RDPG, let $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$; 
-for GRDPG, let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.
+2. Let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.
 
@@
 
-RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
-(Athreya et al., 2018)  
-GRDPG: 
-$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
+$\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$
 (Rubin-Delanchy et al., 2020)
-
-# Connecting the PABM to the GRDPG
 
@@
 
-```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
-par(mar = rep(1.75, 4))
+```{r, out.width = '100%', fig.height = 4, fig.width = 4}
+# par(mar = rep(1.75, 4))
 plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
-```
-
-\vspace*{.5\baselineskip}
-
-```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
-par(mar = rep(1.75, 4))
-omega <- rbeta(n, 1, 1)
-dc.matrix <- omega %*% t(omega)
-P.dcbm <- P * dc.matrix
-dcbm.eigen <- eigen(P.dcbm)
-X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
-plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')
-
 ```
@@
 
-## Connecting the PABM to the GRDPG
-
-::: columns
-
-:::: column
-
-$$X = \begin{bmatrix}
-\Lambda^{(1)} & \cdots & 0 \\
-0 & \ddots & 0 \\
-0 & \cdots & \Lambda^{(K)}
-\end{bmatrix}$$
-
-::::
-
-:::: column
-
-\vspace*{2\baselineskip}
-
-$$U \in \mathbb{O}(K^2)$$
-
-::::
-
-:::
-
 $$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$
-
-**Remark 1** (orthogonality of subspaces):
-If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
-to different communities, then $y_i^\top y_j = 0$.
-
-**Remark 2** (non-uniqueness of the latent configuration):  
-If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
-the indefinite orthogonal group with signature $p, q$.
-
-**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.
-
-# Community Detection for the PABM
 
@@
 
-## Sparse Subspace Clustering
+## Simulation Results
 
-**Corollary**: The ASE of $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ lies near a collection of $K$-dimensional subspaces in $K^2$ dimensions.
+Simulation setup:
 
-**Algorithm**: Sparse Subspace Clustering (Elhamifar & Vidal, 2009):
-
-1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
-subject to $x_i = X^\top c$ and $c^{(i)} = 0$.  
-This is typically performed via LASSO: 
-$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i}^\top c\|_2^2 + \lambda \|c\|_1$
-2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
-3. Construct affinity matrix $B = |C| + |C^\top|$.
-
-## Sparse Subspace Clustering
-
-**Theorem** (KTT):
-
-Let
-
-* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
-$A_n \sim \BG(P_n)$;
-* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
-$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.
-
-Then
-
-* For some $\lambda > 0$ and $N < \infty$, $\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1 when $n > N$.
-
-Remarks:
-
-* For large $n$, we can identify $\lambda$ for SDP (Wang and Xu, 2016).
-* SDP does not guarantee community detection.
-
-## Simulation Results
+1. $z_1, ..., z_n \iid \Cat(1/K, ..., 1/K)$
+2. $\lambda_{ik} \iid \Betadist(a_{ik}, b_{ik})$  
+$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$ 
+$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
+3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
+4. $A \sim \BG(P)$
 
@@
 ```
+
+## Conclusion
+
+1. The PABM is a recently developed flexible block model that can be used to describe graphs with community structure.
+
+2. The GRDPG, which can describe all block models, motivates a spectral approach to statistical inference on graphs.
+
+3.
+Under the GRDPG framework, the PABM with $K$ communities can be induced by a latent configuration in $\mathbb{R}^{K^2}$ consisting of $K$ $K$-dimensional subspaces that are orthogonal to each other.
+
+4. The latent configuration of the PABM under the GRDPG framework leads to an intuitive method for community detection with nice theoretical asymptotic properties.

### 2022-03-28 fls-lightning-talk/slides.Rmd
updated script and minor tweaks to slides
changes based on feedback

@@
 
-How can we cluster the nodes of a network?
-
-Statistical inference (parametric approach):
-
-1. Define a generative model for graph: 
-$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
-2. Develop a method for obtaining estimators: 
-$f(G) = (\hat{\vec{z}}, \hat{\vec{\theta}})$.
-3. Describe asymptotic properties of estimators: 
-$(\hat{\vec{z}}, \hat{\vec{\theta}}) \to (\vec{z}, \vec{\theta})$.
+How might we cluster the nodes of a network?
 
@@
 </style>
-
-::: columns
-
-:::: {.column width=62.5%}
 
@@
 
-::::
-
-:::: {.column width=37.5%}
-
-**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
-For this model, $A_{ij} \iid \Bernoulli(\theta)$.
-
-```{r, fig.height = 2, fig.width = 2}
-n <- 2 ** 5
-p <- 1 / log(n)
-P <- matrix(p, nrow = n, ncol = n)
-A <- draw.graph(P)
-qgraph::qgraph(A, vsize = 4)
-```
-
-::::
-
-:::
-
 ## Block Models
@@
 
-**Example 2**: Stochastic Block Model with two communities
+**Example 1**: Stochastic Block Model with $K=2$ communities
 
@@
 \end{cases}$
-
-* To make this an assortative SBM, set $p q > r^2$.
-* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.
 
@@
 $\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
-Then $A \sim \PABM(\{\lambda_{ik}\}_K)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
-e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.
+Then $A \sim \PABM(\{\lambda_{ik}\}_K)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
 
@@
 
-$A \sim \GRDPG_{p, q}(X)$ iff
+Let $I_{p,q} = \blockdiag(I_p, -I_q)$ and suppose that $x_1, \ldots, x_n \in \mathbb{R}^{p+q}$ are such that $x_i^\top I_{p,q} x_j \in [0,1]$.  
+Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
 
-* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
-$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
-* $A \sim \BG(X I_{p, q} X^\top)$ where 
-$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
-  * $P(\text{edge between } v_i, v_j) = x_i^\top I_{p,q} x_j$
+Adjacency Spectral Embedding (Sussman et al., 2012) estimates $x_1, ..., x_n \in \mathbb{R}^{p+q}$ from $A$:
 
-## (Generalized) Random Dot Product Graph Model
+1. Let $\hat{\Lambda}$ be the diagonal matrix that contains the absolute values of the $p$ most positive and the $q$ most negative eigenvalues.
+2. Let $\hat{V}$ be the matrix whose columns are the corresponding eigenvectors.
+3. Compute $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
 
-### Recovery/Estimation
-
-Want to estimate $X$ from $A$, or alternatively, 
-interpoint distances, inner products, or angles.
-
-### Adjacency Spectral Embedding
-
-To embed in $\mathbb{R}^{p+q}$,
-
-1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
-where $\hat{\Lambda} \in \mathbb{R}^{(p+q) \times (p+q)}$ and 
-$\hat{V} \in \mathbb{R}^{n \times (p+q)}$ 
-by using the $p$ most positive and $q$ most negative eigenvalues.
-
-2. Let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.
-
-\vspace*{.5\baselineskip}
-
-$\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$
-(Rubin-Delanchy et al., 2020)
+Theorem: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$
 
@@
 
-**Example 2** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$
+**Example 1** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$
 
@@
 
-**Example 2** (cont'd): If we want to perform community detection,
+**Example 1** (cont'd): If we want to perform community detection,
 
@@
 
-$$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$
+$$A \sim \PABM(\{\lambda_{ik}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$
 
@@
 ```
-
-## Conclusion
-
-1. The PABM is a recently developed flexible block model that can be used to describe graphs with community structure.
-
-2. The GRDPG, which can describe all block models, motivates a spectral approach to statistical inference on graphs.
-
-3.
-Under the GRDPG framework, the PABM with $K$ communities can be induced by a latent configuration in $\mathbb{R}^{K^2}$ consisting of $K$ $K$-dimensional subspaces that are orthogonal to each other.
-
-4. The latent configuration of the PABM under the GRDPG framework leads to an intuitive method for community detection with nice theoretical asymptotic properties.

### 2022-03-28 fls-lightning-talk/slides.Rmd
updated slides per discussion

@@
 
+1. Define a probability model with communities that might generate the graph (e.g., popularity adjusted block model).
+2. Develop estimators for the parameters of the probability model, including the community labels.
+3. Describe the properties of the estimators (e.g., consistency).
+
 ## Bernoulli Graphs
@@
 
-Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.
+Let $G$ be an undirected and unweighted graph with $n$ vertices.
 
@@
 $A_{ij} = \begin{cases} 
-1 & \exists \text{ edge between } i \text{ and } j \\
-0 & \text{else}
+1 & \text{an edge connects vertices } i \text{ and } j \\
+0 & \text{otherwise}
 \end{cases}$
 
-$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.
+$A_{ji} = A_{ij}$ and $A_{ii} = 0$.
 
@@
 
-1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
-vertices.
+1. $P$ is an $n \times n$ matrix of edge probabilities between pairs of vertices.
 2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.
@@
 
-**Example 1**: Stochastic Block Model with $K=2$ communities
+**Example 1**: Stochastic Block Model with $K=2$ communities.
 
@@
 
-* $z_1, ..., z_n \in \{1, 2\}$
-* $P_{ij} = \begin{cases} 
+$P_{ij} = \begin{cases} 
 p & z_i = z_j = 1 \\
@@
 
-```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
+```{r, fig.height = 3, fig.width = 4, out.width = '100%', fig.cap = 'Two community SBM with $p=1/2$, $q=1/4$, $r=1/8$'}
 n1 <- 2 ** 5
@@
 
-**Lemma** (Noroozi, Rimal, and Pensky, 2020):
+**Def** (Noroozi, Rimal, and Pensky, 2020):
 
@@
 Let $I_{p,q} = \blockdiag(I_p, -I_q)$ and suppose that $x_1, \ldots, x_n \in \mathbb{R}^{p+q}$ are such that $x_i^\top I_{p,q} x_j \in [0,1]$.  
-Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
+Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
 
@@
 
-Theorem: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$
+**Theorem**: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$ 
+as $n \to \infty$
 
-## Connecting Block Models to the (G)RDPG Model
+## Connecting Block Models to the GRDPG Model
 
-All Bernoulli Graphs are RDPG (if $P$ is 
-positive semidefinite)  
-or GRDPG (in general).
+All Bernoulli Graphs are GRDPGs.
 
-**Example 1** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$
+**Example 1** (cont'd): SBM with $K = 2$.
 
@@
 \end{bmatrix} =
-X X^\top$$
+X I_{2,0} X^\top$$
 
@@
 
-## Connecting Block Models to the (G)RDPG Model
+## Connecting Block Models to the GRDPG Model
 
-**Example 1** (cont'd): If we want to perform community detection,
+**Example 1** (cont'd): To perform community detection,
 
@@
 $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
-3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
-noting that as $n \to \infty$, the ASE approaches point masses.
+3. Apply a clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
+noting that $\hat{X}$ approaches point masses as $n \to \infty$.
 
@@
 
-## Connecting Block Models to the (G)RDPG Model
+## Connecting Block Models to the GRDPG Model
 
@@
 
-* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
-* $U \in \mathbb{O}(K^2)$
-* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
-composed of $\{\lambda^{(kl)}\}_K$ 
-with each block corresponding to a community.
+* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
+* $U$ is an orthogonal matrix;
+* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
+composed of popularity vectors with each block corresponding to a community.
 
@@
 
-$$A \sim \PABM(\{\lambda_{ik}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$
+$$A \sim \PABM(\{\lambda_{ik}\}_K) \text{ iff } A \sim \GRDPG_{p, q}(X U)$$
 
@@
 **Theorem** (KTT): 
-Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
+Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC. 
 Then $\forall$ pairs $(i, j)$ belonging to different communities 
@@
 
-$$
-\max_{i, j} \hat{B}^{(ij)}_n = 
-O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
-$$
+$\max_{i, j} \hat{B}_{ij} = 
+O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$ 
+as $n \to \infty$.
 
@@
 
-Simulation setup:
+We compared a variety of algorithms for community detection on randomly generated PABMs:
 
-1. $z_1, ..., z_n \iid \Cat(1/K, ..., 1/K)$
-2. $\lambda_{ik} \iid \Betadist(a_{ik}, b_{ik})$  
-$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$ 
-$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
-3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
-4. $A \sim \BG(P)$
+* Modularity Maximization (Sengupta and Chen) using the Louvain algorithm;
+* Orthogonal Spectral Clustering (KTT);
+* Sparse Subspace Clustering on the ASE (KTT);
+* Sparse Subspace Clustering on the columns of $A$ (Noorozi, Rimal, Pensky).
 
@@
 ```
+
+## Simulation Setup
+
+1. $z_1, ..., z_n \iid \Cat(1/K, ..., 1/K)$
+2. $\lambda_{ik} \iid \Betadist(a_{ik}, b_{ik})$  
+    * $a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$   
+    * $b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
+3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
+4. $A \sim \BG(P)$

### 2022-03-29 fls-lightning-talk/slides.Rmd
numbered items on slide 4 are one line each now
figured out how to add images to title slide

@@
 ---
-title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
-subtitle: Future Leaders Summit 2022 Lightning Presentation
-author: John Koo, Indiana University
-date: 'April 2022'
 output: 
   beamer_presentation:
+    keep_tex: true
     fig_crop: no
@@
     includes:
-      in_header: page_headers.tex
+      in_header: 
+        - page_headers.tex
 header-includes:
@@
 
-## Contributors
+## {.plain}
+
+\center
+
+\LARGE
+
+\textcolor{darkred}{Popularity Adjusted Block Models are Generalized Random Dot Product Graphs}
+
+\normalsize
+
+Future Leaders Summit
+
+April 2022
 
@@
 
-```{r, out.width = '100px'}
+```{r, out.width = '75px'}
 knitr::include_graphics('john-koo.jpeg')
@@
 
-```{r, out.width = '100px'}
+```{r, out.width = '75px'}
 knitr::include_graphics('minh-tang.jpg')
@@
 
-```{r, out.width = '100px'}
+```{r, out.width = '75px'}
 knitr::include_graphics('michael-trosset.jpg')
@@
 
-1. $P$ is an $n \times n$ matrix of edge probabilities between pairs of vertices.
+1. $P$ is a matrix of edge probabilities between pairs of vertices.
 2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.
@@
 
-```{r, fig.height = 3, fig.width = 4, out.width = '100%', fig.cap = 'Two community SBM with $p=1/2$, $q=1/4$, $r=1/8$'}
+```{r, fig.height = 3, fig.width = 4, out.width = '100%', fig.cap = 'SBM with $p=1/2$, $q=1/4$, $r=1/8$'}
 n1 <- 2 ** 5
@@
 
-We compared a variety of algorithms for community detection on randomly generated PABMs:
+We compare four algorithms for community detection on randomly generated PABMs:
 
@@
 * Orthogonal Spectral Clustering (KTT);
-* Sparse Subspace Clustering on the ASE (KTT);
-* Sparse Subspace Clustering on the columns of $A$ (Noorozi, Rimal, Pensky).
+* Sparse Subspace Clustering on the columns of $A$ (Noorozi, Rimal, Pensky); 
+* Sparse Subspace Clustering on the ASE (KTT).
 
@@
 
+## Additional Slides
+
 ## Simulation Setup
@@
 4. $A \sim \BG(P)$
+
+## Future Work

### 2022-03-29 fls-lightning-talk/slides.tex
figured out how to add images to title slide

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usecolortheme{beaver}
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\iffalse
https://tex.stackexchange.com/questions/392324/how-to-exclude-total-slide-number-from-beamer-slide
\fi
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{footline}{\quad\hfill\insertframenumber\strut\quad}
\usefonttheme[onlymath]{serif}
\setbeamertemplate{itemize items}[circle]
\author[John Koo]{John Koo}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\begin{frame}[plain]{}
\protect\hypertarget{section}{}
\center

\LARGE

\textcolor{darkred}{Popularity Adjusted Block Models are Generalized Random Dot Product Graphs}

\normalsize

Future Leaders Summit

April 2022

\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=75px]{john-koo} \end{center}

John Koo,\\
PhD Student in Statistical Science,\\
Indiana University
\end{column}

\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=75px]{minh-tang} \end{center}

Minh Tang,\\
Assistant Professor of Statistics,\\
NC State University
\end{column}

\begin{column}{0.33\textwidth}
\begin{center}\includegraphics[width=75px]{michael-trosset} \end{center}

Michael Trosset,\\
Professor of Statistics,\\
Indiana University
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Community Detection for Networks}
\protect\hypertarget{community-detection-for-networks}{}
\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-4-1} \end{center}

How might we cluster the nodes of a network?

\begin{enumerate}
\tightlist
\item
  Define a probability model with communities that might generate the
  graph (e.g., popularity adjusted block model).
\item
  Develop estimators for the parameters of the probability model,
  including the community labels.
\item
  Describe the properties of the estimators (e.g., consistency).
\end{enumerate}
\end{frame}

\begin{frame}{Bernoulli Graphs}
\protect\hypertarget{bernoulli-graphs}{}
Let \(G\) be an undirected and unweighted graph with \(n\) vertices.

\(G\) is described by adjacency matrix \(A\) such that
\(A_{ij} = \begin{cases} 1 & \text{an edge connects vertices } i \text{ and } j \\ 0 & \text{otherwise} \end{cases}\)

\(A_{ji} = A_{ij}\) and \(A_{ii} = 0\).

\vspace*{1\baselineskip}

\(A \sim \text{BernoulliGraph}(P)\) iff:

\begin{enumerate}
\tightlist
\item
  \(P\) is a matrix of edge probabilities between pairs of vertices.
\item
  \(A_{ij} \stackrel{\text{ind}}{\sim}\text{Bernoulli}(P_{ij})\) for
  each \(i < j\).
\end{enumerate}
\end{frame}

\begin{frame}{Block Models}
\protect\hypertarget{block-models}{}
Suppose each vertex \(v_1, ..., v_n\) has labels
\(z_1, ..., z_n \in \{1, ..., K\}\),\\
and each \(P_{ij}\) depends on labels \(z_i\) and \(z_j\).\\
Then \(A \sim \text{BernoulliGraph}(P)\) is a \emph{block model}.

\textbf{Example 1}: Stochastic Block Model with \(K=2\) communities.

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\(P_{ij} = \begin{cases} p & z_i = z_j = 1 \\ q & z_i = z_j = 2 \\ r & z_i \neq z_j \end{cases}\)
\end{column}

\begin{column}{0.48\textwidth}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1}

}

\caption{SBM with $p=1/2$, $q=1/4$, $r=1/8$}\label{unnamed-chunk-5}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Popularity Adjusted Block Model}
\protect\hypertarget{popularity-adjusted-block-model}{}
\textbf{Def} Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex \(i \in [n]\) have \(K\) popularity parameters
\(\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]\). Then
\(A \sim \text{PABM}(\{\lambda_{ik}\}_K)\) if each
\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\).

\textbf{Def} (Noroozi, Rimal, and Pensky, 2020):

\(A\) is sampled from a PABM if \(P\) can be described as:

\begin{enumerate}
\tightlist
\item
  Let each \(P^{(kl)}\) denote the \(n_k \times n_l\) matrix of edge
  probabilities between communities \(k\) and \(l\).
\item
  Organize popularity parameters as vectors
  \(\lambda^{(kl)} \in \mathbb{R}^{n_k}\) such that
  \(\lambda^{(kl)}_i = \lambda_{k_i l}\) is the popularity parameter of
  the \(i\)\textsuperscript{th} vertex of community \(k\) towards
  community \(l\).
\item
  Each block can be decomposed as
  \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\).
\end{enumerate}
\end{frame}

\begin{frame}{Generalized Random Dot Product Graph}
\protect\hypertarget{generalized-random-dot-product-graph}{}
\textbf{Def} Generalized Random Dot Product Graph\\
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

Let \(I_{p,q} = \text{blockdiag}(I_p, -I_q)\) and suppose that
\(x_1, \ldots, x_n \in \mathbb{R}^{p+q}\) are such that
\(x_i^\top I_{p,q} x_j \in [0,1]\).\\
Then \(A \sim \text{GRDPG}_{p, q}(X)\) iff
\(A \sim \text{BernoulliGraph}(X I_{p,q} X^\top)\), where
\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).

Adjacency Spectral Embedding (Sussman et al., 2012) estimates
\(x_1, ..., x_n \in \mathbb{R}^{p+q}\) from \(A\):

\begin{enumerate}
\tightlist
\item
  Let \(\hat{\Lambda}\) be the diagonal matrix that contains the
  absolute values of the \(p\) most positive and the \(q\) most negative
  eigenvalues.
\item
  Let \(\hat{V}\) be the matrix whose columns are the corresponding
  eigenvectors.
\item
  Compute \(\hat{X} = \hat{V} \hat{\Lambda}^{1/2}\).
\end{enumerate}

\textbf{Theorem}:
\(\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)\)
as \(n \to \infty\)
\end{frame}

\begin{frame}{Connecting Block Models to the GRDPG Model}
\protect\hypertarget{connecting-block-models-to-the-grdpg-model}{}
All Bernoulli Graphs are GRDPGs.

\textbf{Example 1} (cont'd): SBM with \(K = 2\).

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\[P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}\]

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}
\end{column}

\begin{column}{0.48\textwidth}
\[P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X I_{2,0} X^\top\]

\[X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}\]
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Connecting Block Models to the GRDPG Model}
\protect\hypertarget{connecting-block-models-to-the-grdpg-model-1}{}
\textbf{Example 1} (cont'd): To perform community detection,

\begin{enumerate}
\tightlist
\item
  Note that \(A\) is a RDPG because \(P = X X^\top\).
\item
  Compute the ASE \(A \approx \hat{X} \hat{X}^\top\) with
  \(\hat{X} = \hat{V} \hat{\Lambda}^{1/2}\).
\item
  Apply a clustering algorithm (e.g., \(K\)-means) to \(\hat{X}\),\\
  noting that \(\hat{X}\) approaches point masses as \(n \to \infty\).
\end{enumerate}

\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-7-1} \end{center}
\end{frame}

\begin{frame}{Connecting Block Models to the GRDPG Model}
\protect\hypertarget{connecting-block-models-to-the-grdpg-model-2}{}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\end{column}

\begin{column}{0.48\textwidth}
\vspace*{0\baselineskip}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-9-1} \end{center}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Connecting the PABM to the GRDPG}
\protect\hypertarget{connecting-the-pabm-to-the-grdpg}{}
\textbf{Theorem} (KTT): \(A \sim \text{PABM}(\{\lambda^{(kl)}\}_K)\) is
equivalent to \(A \sim \text{GRDPG}_{p, q}(X U)\) with

\begin{itemize}
\tightlist
\item
  \(p = K (K + 1) / 2\), \(q = K (K - 1) / 2\);
\item
  \(U\) is an orthogonal matrix;
\item
  \(X \in \mathbb{R}^{n \times K^2}\) is a block diagonal matrix
  composed of popularity vectors with each block corresponding to a
  community.
\end{itemize}

\[X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}\]

\[\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}\]

\[A \sim \text{PABM}(\{\lambda_{ik}\}_K) \text{ iff } A \sim \text{GRDPG}_{p, q}(X U)\]
\end{frame}

\begin{frame}{Orthogonal Spectral Clustering}
\protect\hypertarget{orthogonal-spectral-clustering}{}
\textbf{Theorem} (KTT): If \(P = V \Lambda V^\top\) and
\(B = n V V^\top\),\\
then \(B_{ij} = 0\) if \(z_i \neq z_j\).

\textbf{Algorithm}: Orthogonal Spectral Clustering:

\begin{enumerate}
\tightlist
\item
  Let \(V\) be the eigenvectors of \(A\) corresponding to the
  \(K (K+1)/2\) most positive and \(K (K-1) / 2\) most negative
  eigenvalues.
\item
  Compute \(B = |n V V^\top|\) applying \(|\cdot|\) entry-wise.
\item
  Construct graph \(G\) using \(B\) as its similarity matrix.
\item
  Partition \(G\) into \(K\) disconnected subgraphs.
\end{enumerate}

\textbf{Theorem} (KTT): Let \(\hat{B}\) with entries \(\hat{B}_{ij}\) be
the affinity matrix from OSC. Then \(\forall\) pairs \((i, j)\)
belonging to different communities and sparsity factor satisfying
\(n \rho_n = \omega\big((\log n)^{4c}\big)\),

\(\max_{i, j} \hat{B}_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\)
as \(n \to \infty\).
\end{frame}

\begin{frame}{Simulation Results}
\protect\hypertarget{simulation-results}{}
We compare four algorithms for community detection on randomly generated
PABMs:

\begin{itemize}
\tightlist
\item
  Modularity Maximization (Sengupta and Chen) using the Louvain
  algorithm;
\item
  Orthogonal Spectral Clustering (KTT);
\item
  Sparse Subspace Clustering on the columns of \(A\) (Noorozi, Rimal,
  Pensky);
\item
  Sparse Subspace Clustering on the ASE (KTT).
\end{itemize}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-10-1} \end{center}
\end{frame}

\begin{frame}{Additional Slides}
\protect\hypertarget{additional-slides}{}
\end{frame}

\begin{frame}{Simulation Setup}
\protect\hypertarget{simulation-setup}{}
\begin{enumerate}
\tightlist
\item
  \(z_1, ..., z_n \stackrel{\text{iid}}{\sim}\text{Categorical}(1/K, ..., 1/K)\)
\item
  \(\lambda_{ik} \stackrel{\text{iid}}{\sim}\text{Beta}(a_{ik}, b_{ik})\)

  \begin{itemize}
  \tightlist
  \item
    \(a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}\)
  \item
    \(b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}\)
  \end{itemize}
\item
  \(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\)
\item
  \(A \sim \text{BernoulliGraph}(P)\)
\end{enumerate}
\end{frame}

\begin{frame}{Future Work}
\protect\hypertarget{future-work}{}
\end{frame}

\end{document}

### 2022-03-30 fls-lightning-talk/slides.Rmd
updated notes/script

@@
 
-**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
+**Theorem** (KTT): $A \sim \PABM(\{\lambda_{ik}\}_K)$ is equivalent to 
 $A \sim \GRDPG_{p, q}(X U)$ with
@@
 and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,  
-
 $\max_{i, j} \hat{B}_{ij} = 
@@
 * Orthogonal Spectral Clustering (KTT);
-* Sparse Subspace Clustering on the columns of $A$ (Noorozi, Rimal, Pensky); 
+* Sparse Subspace Clustering on the columns of $A$  
+(Noorozi, Rimal, Pensky); 
 * Sparse Subspace Clustering on the ASE (KTT).

### 2022-03-30 fls-lightning-talk/slides.tex
updated notes/script

@@
 \protect\hypertarget{connecting-the-pabm-to-the-grdpg}{}
-\textbf{Theorem} (KTT): \(A \sim \text{PABM}(\{\lambda^{(kl)}\}_K)\) is
+\textbf{Theorem} (KTT): \(A \sim \text{PABM}(\{\lambda_{ik}\}_K)\) is
 equivalent to \(A \sim \text{GRDPG}_{p, q}(X U)\) with
@@
 belonging to different communities and sparsity factor satisfying
-\(n \rho_n = \omega\big((\log n)^{4c}\big)\),
-
+\(n \rho_n = \omega\big((\log n)^{4c}\big)\),\\
 \(\max_{i, j} \hat{B}_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\)
@@
 \item
-  Sparse Subspace Clustering on the columns of \(A\) (Noorozi, Rimal,
-  Pensky);
+  Sparse Subspace Clustering on the columns of \(A\)\\
+  (Noorozi, Rimal, Pensky);
 \item

### 2022-04-11 fls-lightning-talk/slides.Rmd
added paper info

@@
 
-```{r, out.width = '75px'}
+```{r, out.width = '67px'}
 knitr::include_graphics('john-koo.jpeg')
@@
 
-```{r, out.width = '75px'}
+```{r, out.width = '67px'}
 knitr::include_graphics('minh-tang.jpg')
@@
 
-```{r, out.width = '75px'}
+```{r, out.width = '67px'}
 knitr::include_graphics('michael-trosset.jpg')
@@
 
+**Corollary**: OSC results in zero clustering error as $n \to \infty$, with probability 1.
+
 ## Simulation Results
@@
 
-## Additional Slides
+## Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
 
-## Simulation Setup
+Submitted to *Journal of Computational and Graphical Statistics* (under review)
 
-1. $z_1, ..., z_n \iid \Cat(1/K, ..., 1/K)$
-2. $\lambda_{ik} \iid \Betadist(a_{ik}, b_{ik})$
-    * $a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$
-    * $b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
-3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
-4. $A \sim \BG(P)$
+arXiv preprint: https://arxiv.org/abs/2109.04010
 
-## Future Work
+GitHub repository: https://github.com/johneverettkoo/pabm-grdpg
+
+R package: https://github.com/johneverettkoo/osc

### 2022-04-11 fls-lightning-talk/slides.tex
added paper info

@@
 \begin{column}{0.33\textwidth}
-\begin{center}\includegraphics[width=75px]{john-koo} \end{center}
+\begin{center}\includegraphics[width=67px]{john-koo} \end{center}
 
@@
 \begin{column}{0.33\textwidth}
-\begin{center}\includegraphics[width=75px]{minh-tang} \end{center}
+\begin{center}\includegraphics[width=67px]{minh-tang} \end{center}
 
@@
 \begin{column}{0.33\textwidth}
-\begin{center}\includegraphics[width=75px]{michael-trosset} \end{center}
+\begin{center}\includegraphics[width=67px]{michael-trosset} \end{center}
 
@@
 as \(n \to \infty\).
+
+\textbf{Corollary}: OSC results in zero clustering error as
+\(n \to \infty\), with probability 1.
 \end{frame}
@@
 
-\begin{frame}{Additional Slides}
-\protect\hypertarget{additional-slides}{}
-\end{frame}
+\begin{frame}{Popularity Adjusted Block Models are Generalized Random
+Dot Product Graphs}
+\protect\hypertarget{popularity-adjusted-block-models-are-generalized-random-dot-product-graphs}{}
+Submitted to \emph{Journal of Computational and Graphical Statistics}
+(under review)
 
-\begin{frame}{Simulation Setup}
-\protect\hypertarget{simulation-setup}{}
-\begin{enumerate}
-\tightlist
-\item
-  \(z_1, ..., z_n \stackrel{\text{iid}}{\sim}\text{Categorical}(1/K, ..., 1/K)\)
-\item
-  \(\lambda_{ik} \stackrel{\text{iid}}{\sim}\text{Beta}(a_{ik}, b_{ik})\)
+arXiv preprint: \url{https://arxiv.org/abs/2109.04010}
 
-  \begin{itemize}
-  \tightlist
-  \item
-    \(a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}\)
-  \item
-    \(b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}\)
-  \end{itemize}
-\item
-  \(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\)
-\item
-  \(A \sim \text{BernoulliGraph}(P)\)
-\end{enumerate}
-\end{frame}
+GitHub repository: \url{https://github.com/johneverettkoo/pabm-grdpg}
 
-\begin{frame}{Future Work}
-\protect\hypertarget{future-work}{}
+R package: \url{https://github.com/johneverettkoo/osc}
 \end{frame}

### 2022-05-21 clustering-sim-k-imbalanaced.R
cleaning things up

@@ -1,131 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(igraph)
-library(ggplot2)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg/codes_and_data')
-source('EPalgosim.R')
-source('functions.R')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, 
-             generate.P.beta,
-             normalized.laplacian,
-             cluster.pabm, ssc,
-             .from = 'functions.R')
-
-mod.max <- function(A) {
-  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
-  clustering <- cluster_louvain(Ag)
-  return(membership(clustering))
-}
-
-cluster.acc <- function(yhat, yobs) {
-  table(yhat, yobs) %>% 
-    as.matrix() %>% 
-    apply(1, max) %>% 
-    sum() %>% 
-    magrittr::divide_by(length(yhat)) %>% 
-    return()
-}
-
-cluster.sg <- function(A) {
-  b.can = EPalgo(A,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  Q.DC.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A, b=b.can[,i])   # fit PABM
-    Q.DC.can[i] = Q.DC(A, b=b.can[,i])   # fit DCBM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  b.PA = b.can[,foo1]   # community assignment that maximises Q.PA
-  return(b.PA)
-}
-
-# simulation parameters
-K.vec <- c(2, 3, 4)
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
[…]

### 2022-05-21 clustering-sim-k.R

@@ -1,144 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-library(igraph)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg/codes_and_data')
-source('EPalgosim.R')
-source('functions.R')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, 
-             generate.P.beta,
-             normalized.laplacian,
-             cluster.pabm, ssc,
-             .from = 'functions.R')
-
-mod.max <- function(A) {
-  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
-  clustering <- cluster_louvain(Ag)
-  return(membership(clustering))
-}
-
-cluster.sg <- function(A) {
-  b.can = EPalgo(A,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  Q.DC.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A, b=b.can[,i])   # fit PABM
-    Q.DC.can[i] = Q.DC(A, b=b.can[,i])   # fit DCBM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  b.PA = b.can[,foo1]   # community assignment that maximises Q.PA
-  return(b.PA)
-}
-
-cluster.acc <- function(yhat, yobs) {
-  table(yhat, yobs) %>% 
-    as.matrix() %>% 
-    apply(1, max) %>% 
-    sum() %>% 
-    magrittr::divide_by(length(yhat)) %>% 
-    return()
-}
-
-# simulation parameters
-K.vec <- c(2, 3, 4)
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
[…]

### 2022-05-21 clustering-sim-ssc-imbalanced.R

@@ -1,151 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-library(igraph)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg/codes_and_data')
-source('EPalgosim.R')
-source('functions.R')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, 
-             generate.P.beta,
-             normalized.laplacian,
-             cluster.pabm, ssc, ssc2,
-             .from = 'functions.R')
-
-mod.max <- function(A) {
-  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
-  clustering <- cluster_louvain(Ag)
-  return(membership(clustering))
-}
-
-cluster.sg <- function(A) {
-  b.can = EPalgo(A,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  Q.DC.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A, b=b.can[,i])   # fit PABM
-    Q.DC.can[i] = Q.DC(A, b=b.can[,i])   # fit DCBM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  b.PA = b.can[,foo1]   # community assignment that maximises Q.PA
-  return(b.PA)
-}
-
-cluster.acc <- function(yhat, yobs) {
-  table(yhat, yobs) %>% 
-    as.matrix() %>% 
-    apply(1, max) %>% 
-    sum() %>% 
-    magrittr::divide_by(length(yhat)) %>% 
-    return()
-}
-
-# simulation parameters
-K.vec <- c(2, 3, 4)
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
[…]

### 2022-05-21 clustering-sim-ssc.R

@@ -1,151 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-library(igraph)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg/codes_and_data')
-source('EPalgosim.R')
-source('functions.R')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, 
-             generate.P.beta,
-             normalized.laplacian,
-             cluster.pabm, ssc, ssc2,
-             .from = 'functions.R')
-
-mod.max <- function(A) {
-  Ag <- graph_from_adjacency_matrix(A, mode = 'undirected')
-  clustering <- cluster_louvain(Ag)
-  return(membership(clustering))
-}
-
-cluster.sg <- function(A) {
-  b.can = EPalgo(A,eps=0) # EP algorithm (no perturbation)
-  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
-  Q.DC.can = rep(NA, ncol(b.can))	# array to store Q values
-  for (i in 1:ncol(b.can)){
-    #check if any cluster is empty
-    foo = rep(NA, 2)
-    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
-    if (min(foo)==0) {stop('Empty groups are not allowed')} 
-    Q.PA.can[i] = Q.PA(A, b=b.can[,i])   # fit PABM
-    Q.DC.can[i] = Q.DC(A, b=b.can[,i])   # fit DCBM
-  } # end of i for loop
-  foo1 = order(-Q.PA.can)[1] 
-  b.PA = b.can[,foo1]   # community assignment that maximises Q.PA
-  return(b.PA)
-}
-
-cluster.acc <- function(yhat, yobs) {
-  table(yhat, yobs) %>% 
-    as.matrix() %>% 
-    apply(1, max) %>% 
-    sum() %>% 
-    magrittr::divide_by(length(yhat)) %>% 
-    return()
-}
-
-# simulation parameters
-K.vec <- c(2, 3, 4)
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
[…]

### 2022-05-21 clustering-sim.R

@@ -1,74 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, draw.pabm.beta.2,
-             normalized.laplacian,
-             cluster.pabm, ssc,
-             .from = 'functions.R')
-# library(ggplot2)
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores())
-
-# simulation parameters
-K <- 2
-p <- K * (K + 1) / 2
-q <- K * (K - 1) / 2
-alpha <- .5
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
-sparsity <- .01
-n.vec <- c(64, 128, 256, 512, 1024, 2048)
-iter <- 100
-set.seed(314159)
-
-# clustering simulation
-clustering.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-  foreach(i = seq(iter), 
-          .combine = dplyr::bind_rows, 
-          .errorhandling = 'remove') %dopar% {
-    Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
-    A <- Az$A
-    z <- Az$z
-    clustering <- cluster.pabm(A, K)
-    error <- mean(clustering == z)
-    if (error > .5) error <- 1 - error
-    clustering.ssc <- ssc(A, K, sparsity)
-    error.ssc <- mean(clustering.ssc == z)
-    if (error.ssc > .5) error.ssc <- 1 - error.ssc
-    dplyr::tibble(n = n, error = error, error.ssc = error.ssc) %>% 
-      return()
-  } %>% 
-    return()
-}
-
-clustering.df %>%
-  dplyr::group_by(n) %>%
-  dplyr::summarise(med.err = median(error),
-                   first.q = quantile(error, .25),
-                   third.q = quantile(error, .75),
-                   med.err.ssc = median(error.ssc),
-                   first.q.ssc = quantile(error.ssc, .25),
-                   third.q.ssc = quantile(error.ssc, .75)) %>%
[…]

### 2022-05-21 density-sim.R

@@ -1,52 +1,0 @@
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, draw.pabm.beta.2,
-             normalized.laplacian,
-             cluster.pabm, ssc, 
-             embedding, 
-             .from = 'functions.R')
-library(ggplot2)
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores())
-
-# simulation parameters
-K <- 2
-p <- K * (K + 1) / 2
-q <- K * (K - 1) / 2
-alpha <- .5
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
-sparsity <- .01
-# n.vec <- c(64, 128, 256, 512, 1024, 2048)
-n.vec <- c(512, 1024, 2048, 3072, 4096)
-iter <- 100
-subsample <- 1e3
-set.seed(314159)
-
-# clustering simulation
-densities.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-  Az <- draw.pabm.beta.2(n, a1, b1, a2, b2, alpha)
-  A <- Az$A
-  z <- Az$z
-  A.eigen <- eigen(A, symmetric = TRUE)
-  V <- A.eigen$vectors[, c(1:3, n)]
-  B <- V %*% t(V)
-  between.clust <- B[z == 1, z == 2]
-  between.clust <- between.clust[upper.tri(between.clust)]
-  between.clust <- sample(between.clust, subsample)
-  dplyr::tibble(n = n, inner.prods = between.clust) %>% 
-    return()
-}
-
-ggplot(densities.df) + 
-  geom_density(aes(x = inner.prods * n, colour = factor(n)))
-
-readr::write_csv(densities.df, 'densities.csv')

### 2022-05-21 disassortative-test.R

@@ -1,113 +1,0 @@
-# simulation study for clustering algorithms
-# compares OSC, SSC-ASE, SSC-A, and Louvain (stand-in for MM)
-# simulations are for balanced community sizes
-# configuration as-is parallelizes simulations across 10 threads
-# which uses ~20 GB memory at its peak
-# modify line 28 to increase/decrease parallelization (and memory use)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-setwd('~/dev/pabm-grdpg')
-source('simulations-and-data-analysis/functions.R')
-
-# simulation parameters
-K.vec <- c(2, 3, 4)
-a1 <- 1
-b1 <- 2
-a2 <- 2
-b2 <- 1
-sparsity <- 1e-2
-n.vec <- c(128, 256, 512, 1024, 2048)
-iter <- 10
-n.vec <- rev(n.vec)
-K.vec <- rev(K.vec)
-
-doMC::registerDoMC(10)
-
-# clustering simulation
-clustering.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
-  reorder.mat <- gtools::permutations(K, K)
-  out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-    print(paste('K =', K, ', n =', n))
-    foreach(i = seq(iter), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
-      Pz <- generate.P.beta(n, K, a1, a2, b1, b2)
-      P <- Pz$P
-      z <- Pz$clustering
-      A <- draw.graph(P)
-      clustering.osc <- cluster.pabm(A, K, use.all = TRUE, normalize = FALSE)
-      error.osc <- 1 - cluster.acc(clustering.osc, z, reorder.mat)
-      clustering.ssc <- ssc(A, K, K / n, normalize = TRUE, scale = FALSE)
-      error.ssc <- 1 - cluster.acc(clustering.ssc, z, reorder.mat)
-      clustering.louvain <- mod.max(A)
-      error.louvain <- 1 - cluster.acc(clustering.louvain, z)
-      clustering.ssc.A <- ssc2(A, K, sparsity, 
-                               normalize = TRUE, parallel = FALSE)
-      error.ssc.A <- 1 - cluster.acc(clustering.ssc.A, z)
-      out <- dplyr::tibble(K = K, n = n,
[…]

### 2022-05-21 rmse-sim-k-imbalanced.R

@@ -1,72 +1,0 @@
-# simulation using proposed parameter estimation method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packgaes, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`)
-library(ggplot2)
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, I.pq, generate.P.beta,
-             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
-             .from = 'functions.R')
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores() / 1.5)
-
-# simulation params
-K.vec <- c(4, 3, 2)
-a1 <- b2 <- 2
-a2 <- b1 <- 1
-n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-n.vec <- rev(n.vec)
-iter <- 50
-set.seed(314159)
-
-# simulation
-rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
-  p <- K * (K + 1) / 2
-  q <- K * (K - 1) / 2
-  Ipq <- I.pq(p, q)
-  
-  out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-    print(paste('n =', n, 'K =', K))
-    plyr::ldply(seq(iter), function(i) {
-      P.z <- generate.P.beta(n, K, a1, a2, b1, b2, unbalanced = TRUE)
-      P <- P.z$P
-      z <- P.z$clustering
-      A <- draw.graph(P)
-      rmse <- lambda.rmse(P, A, z)
-      rmse.mle <- lambda.rmse.mle(P, A, z)
-      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>% 
-        return()
-    }, .parallel = TRUE) %>% 
-      return()
-  }
-  gc()
-  return(out)
-}
-
-rmse.df %>%
-  na.omit() %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
[…]

### 2022-05-21 rmse-sim-k.R

@@ -1,72 +1,0 @@
-# simulation using proposed parameter estimation method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packgaes, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`)
-library(ggplot2)
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, I.pq, generate.P.beta,
-             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
-             .from = 'functions.R')
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores() - 2)
-
-# simulation params
-K.vec <- c(4, 3, 2)
-a1 <- b2 <- 2
-a2 <- b1 <- 1
-n.vec <- c(128, 256, 512, 1024, 2048, 4096)
-n.vec <- rev(n.vec)
-iter <- 50
-set.seed(314159)
-
-# simulation
-rmse.df <- foreach(K = K.vec, .combine = dplyr::bind_rows) %do% {
-  p <- K * (K + 1) / 2
-  q <- K * (K - 1) / 2
-  Ipq <- I.pq(p, q)
-
-  out <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-    print(paste('n =', n, 'K =', K))
-    plyr::ldply(seq(iter), function(i) {
-      P.z <- generate.P.beta(n, K)
-      P <- P.z$P
-      z <- P.z$clustering
-      A <- draw.graph(P)
-      rmse <- lambda.rmse(P, A, z)
-      rmse.mle <- lambda.rmse.mle(P, A, z)
-      dplyr::tibble(K = K, n = n, rmse = rmse, rmse.mle = rmse.mle) %>%
-        return()
-    }, .parallel = TRUE) %>%
-      return()
-  }
-  gc()
-  return(out)
-}
-
-rmse.df %>%
-  na.omit() %>%
-  dplyr::group_by(K, n) %>%
-  dplyr::summarise(median.rmse = median(rmse),
-                   q1.rmse = quantile(rmse, .25),
-                   q3.rmse = quantile(rmse, .75),
-                   median.rmse.mle = median(rmse.mle),
-                   q1.rmse.mle = quantile(rmse.mle, .25),
-                   q3.rmse.mle = quantile(rmse.mle, .75)) %>%
-  dplyr::ungroup() %>%
[…]

### 2022-05-21 rmse-sim.R

@@ -1,66 +1,0 @@
-# simulation using proposed parameter estimation method
-# run on IU RED (not a good idea to attempt on laptop)
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`)
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, I.pq, 
-             estimate.lambda.block, lambda.rmse, lambda.rmse.mle,
-             .from = 'functions.R')
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores() - 2)
-
-# simulation params
-K <- 2
-p <- K * (K + 1) / 2
-q <- K * (K - 1) / 2
-alpha <- .5
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
-n.vec <- c(64, 128, 256, 512, 1024, 2048)
-iter <- 100
-Ipq <- I.pq(p, q)
-set.seed(314159)
-
-# simulation
-rmse.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-  plyr::ldply(seq(iter), function(i) {
-    z <- sample(c(1, 2), n, prob = c(alpha, 1 - alpha), replace = TRUE)
-    z <- sort(z)
-    
-    n1 <- sum(z == 1)
-    n2 <- sum(z == 2)
-    
-    lambda11 <- rbeta(n1, a1, b1)
-    lambda22 <- rbeta(n2, a1, b1)
-    lambda12 <- rbeta(n1, a2, b2)
-    lambda21 <- rbeta(n2, a2, b2)
-    
-    X <- cbind(c(lambda11, rep(0, n2)),
-               c(lambda12, rep(0, n2)),
-               c(rep(0, n1), lambda21),
-               c(rep(0, n1), lambda22))
-    Y <- cbind(c(lambda11, rep(0, n2)),
-               c(rep(0, n1), lambda21),
-               c(lambda12, rep(0, n2)),
-               c(rep(0, n1), lambda22))
-    P <- X %*% t(Y)
-    A <- draw.graph(P)
-    # lambda.matrix <- cbind(c(lambda11, lambda21),
-    #                        c(lambda12, lambda22))
-    Zhat <- embedding(A, p, q)
-    Phat <- Zhat %*% Ipq %*% t(Zhat)
-    rmse <- lambda.rmse(P, Phat, z)
[…]

### 2022-05-21 mod-max-sim.R
cleaning up

@@ -1,95 +1,0 @@
-# simulation using proposed clustering method
-# run on IU RED (not a good idea to attempt on laptop)
-# this code is mostly from scripts provided by sengupta & chen
-
-# packages, functions, etc.
-import::from(magrittr, `%>%`)
-import::from(foreach, foreach, `%do%`, `%dopar%`)
-library(mclust)
-library(ggplot2)
-library(igraph)
-library(clue)
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-setwd('~/dev/pabm-grdpg')
-import::here(embedding, draw.graph, 
-             generate.P.beta,
-             normalized.laplacian,
-             cluster.pabm, ssc,
-             .from = 'functions.R')
-setwd('codes_and_data')
-source('functions.R')
-source('EPalgosim.R')
-setwd('..')
-
-mod.max.clust <- function(A) {
-  # code copy-pasted from sengupta & chen
-  Eps <- (0:19)/20
-  b.can <- NULL
-  for (ieps in 1:length(Eps)){
-    b.can<-c(b.can,list(EPalgo(A,eps=Eps[ieps]))) # eps=0 for unperturbed version
-  }
-  
-  b.can.unpert <- b.can[[1]]     # matrix of candidate assignments for eps = 0
-  n.can.unpert <- dim(b.can.unpert)[2]  # total number of candidate assignments
-  Q.PA.unpert = rep(NA, n.can.unpert)  # array to store modularity values
-  
-  ##### modularity calculation #####
-  ptm<-proc.time()
-  for (ican in 1:n.can.unpert){
-    Q.PA.unpert[ican] = Q.PA(A, b=b.can.unpert[,ican])
-  }
-  index.PA<-which(Q.PA.unpert==max(Q.PA.unpert))
-  b.PA.unpert = b.can.unpert[,index.PA]
-  
-  return(b.PA.unpert)
-}
-
-# parallel backend
-doMC::registerDoMC(parallel::detectCores())
-
-# simulation parameters
-K <- 2
-a1 <- 2
-b1 <- 1
-a2 <- 1
-b2 <- 2
-sparsity <- .01
-n.vec <- c(64, 128, 256, 512, 1024, 2048)
-iter <- 100
-set.seed(314159)
-
-# clustering simulation
-clustering.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
-  print(paste('K =', K, ', n =', n))
-  foreach(i = seq(iter), 
[…]

### 2022-05-21 pabm-scratch.R

@@ -1,417 +1,0 @@
-import::from(magrittr, `%>%`)
-library(plot.matrix)
-import::from(foreach, `%do%`)
-# library(CVXR)
-
-source('http://pages.iu.edu/~mtrosset/Courses/675/stress.r')
-
-doMC::registerDoMC(parallel::detectCores())
-
-embedding <- function(A, p = NULL, q = NULL,
-                      eps = 1e-6) {
-  eigen.A <- eigen(A, symmetric = TRUE)
-  if (is.null(p) | is.null(q)) {
-    keep <- (abs(eigen.A$values) > eps)
-  } else {
-    keep <- c(seq(p), seq(n, n - q + 1))
-  }
-  
-  U <- eigen.A$vectors[, keep]
-  S <- diag(sqrt(abs(eigen.A$values[keep])))
-  return(U %*% S)
-}
-
-draw.graph <- function(P) {
-  A <- apply(P, 1:2, function(p) rbinom(1, 1, p))
-  A[lower.tri(A)] <- 0
-  diag(A) <- 0
-  A <- A + t(A)
-  return(A)
-}
-
-n1 <- 2 ** 5
-n2 <- n1
-n <- n1 + n2
-
-z <- c(rep(1, n1), 
-       rep(2, n2))
-
-Ipq <- diag(c(1, 1, 1, -1))
-
-lambda11 <- rbeta(n1, 2, 1)
-lambda22 <- rbeta(n2, 2, 1)
-lambda12 <- rbeta(n1, 1, 2)
-lambda21 <- rbeta(n2, 1, 2)
-
-X <- cbind(c(lambda11, rep(0, n2)),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(rep(0, n1), lambda22))
-Y <- cbind(c(lambda11, rep(0, n2)),
-           c(rep(0, n1), lambda21),
-           c(lambda12, rep(0, n2)),
-           c(rep(0, n1), lambda22))
-P <- X %*% t(Y)
-A <- draw.graph(P)
-
-Z <- embedding(P)
-Zhat <- embedding(A, 3, 1)
-pairs(Z, col = z, asp = 1)
-
-P - Z %*% Ipq %*% t(Z)
-
-XU <- cbind(c(lambda11, rep(0, n2)),
-            c(rep(0, n1), lambda22),
-            c(lambda12 / sqrt(2), lambda21 / sqrt(2)),
[…]

### 2022-05-21 plotting-temp.R

@@ -1,102 +1,0 @@
-clustering.df %>%
-  dplyr::group_by(n, K) %>%
-  dplyr::summarise(
-    med.err = median(error),
-    first.q = quantile(error, .25),
-    third.q = quantile(error, .75),
-    med.err.ssc = median(error.ssc),
-    first.q.ssc = quantile(error.ssc, .25),
-    third.q.ssc = quantile(error.ssc, .75),
-    med.err.ep = median(error.ep, na.rm = TRUE),
-    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
-    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
-    med.err.louvain = median(error.louvain),
-    first.q.louvain = quantile(error.louvain, .25),
-    third.q.louvain = quantile(error.louvain, .75)
-  ) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::inner_join(
-    ssc.df %>% 
-      dplyr::group_by(n, K) %>% 
-      dplyr::summarise(med.err.ssc.A = median(error.ssc2),
-                       first.q.ssc.A = quantile(error.ssc2, .25),
-                       third.q.ssc.A = quantile(error.ssc2, .75)) %>% 
-      dplyr::ungroup()
-  ) %>% 
-  ggplot() +
-  theme_bw() + 
-  theme(text = element_text(size = 20),
-        legend.position = 'bottom') + 
-  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
-                labels = c(expression(2^7), 
-                           expression(2^8), 
-                           expression(2^9), 
-                           expression(2^10), 
-                           expression(2^11), 
-                           expression(2^12))) +
-  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
-  scale_y_log10() +
-  labs(y = 'error count', 
-       colour = NULL, shape = NULL, size = NULL) +
-  geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
-  geom_point(aes(x = n, y = med.err * n,
-                 colour = 'OSC', shape = 'OSC'), size = 3) +
-  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
-  geom_line(aes(x = n, y = med.err.ssc * n,
[…]

### 2022-07-12 jsm/speed-presentation/slides.Rmd
slides for jsm

--- 2022-04-11 fls-lightning-talk/slides.Rmd
@@
 
-Future Leaders Summit
+JSM Speed Presentation
 
-April 2022
+August 2022
 
@@
 ```{r out.width = '50%', fig.height = 3, fig.width = 4}
-Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
+n <- 64
+Pz <- generate.P.beta(n, 4, 4, 1, 1, 4)
 P <- Pz$P
@@
 
-How might we cluster the nodes of a network?
+**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
 
-1. Define a probability model with communities that might generate the graph (e.g., popularity adjusted block model).
-2. Develop estimators for the parameters of the probability model, including the community labels.
-3. Describe the properties of the estimators (e.g., consistency).
-
-## Bernoulli Graphs
-
-<style type="text/css">
-.caption {
-    font-size: x-small;
-}
-</style>
-
-Let $G$ be an undirected and unweighted graph with $n$ vertices.
-
-$G$ is described by adjacency matrix $A$ such that
-$A_{ij} = \begin{cases} 
-1 & \text{an edge connects vertices } i \text{ and } j \\
-0 & \text{otherwise}
-\end{cases}$
-
-$A_{ji} = A_{ij}$ and $A_{ii} = 0$.
-
-\vspace*{1\baselineskip}
-
-$A \sim \BG(P)$ iff:
-
-1. $P$ is a matrix of edge probabilities between pairs of vertices.
-2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.
-
-## Block Models
-
-Suppose each vertex $v_1, ..., v_n$ has labels $z_1, ..., z_n \in \{1, ..., K\}$,  
-and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
-Then $A \sim \BG(P)$ is a *block model*.
-
-**Example 1**: Stochastic Block Model with $K=2$ communities.
-
-::: columns
-
-:::: column
-
-$P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
-\end{cases}$
-
-::::
-
-:::: column
-
-```{r, fig.height = 3, fig.width = 4, out.width = '100%', fig.cap = 'SBM with $p=1/2$, $q=1/4$, $r=1/8$'}
-n1 <- 2 ** 5
-n2 <- 2 ** 5
-n <- n1 + n2
-z <- c(rep(1, n1), rep(2, n2))
-p <- 1/2
-q <- 1/4
-r <- 1/8
-P <- matrix(r, nrow = n, ncol = n)
-P[seq(n1), seq(n1)] <- p
-P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
-A <- draw.graph(P)
-qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
-```
-
-::::
-
-:::
+Let each vertex $i \in [n]$ have $K$ popularity parameters
+$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$.  
+Then $A \sim \BG(P)$ is a PABM if each 
+$P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
 
@@
 
-## Connecting Block Models to the GRDPG Model
-
-All Bernoulli Graphs are GRDPGs.
-
-**Example 1** (cont'd): SBM with $K = 2$.
+## Connecting Block Models to the (G)RDPG Model
 
@@
 
-$$P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
-\end{cases}$$
-
-```{r fig.height = 3, fig.width = 6, out.width = '100%'}
-qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
-```
-
-::::
-
-:::: column
-
-$$P = 
-\begin{bmatrix} 
-P^{(11)} & P^{(12)} \\
-P^{(21)} & P^{(22)}
-\end{bmatrix} =
-X I_{2,0} X^\top$$
-
-$$X = \begin{bmatrix} 
-\sqrt{p} & 0 \\
-\vdots & \vdots \\
-\sqrt{p} & 0 \\
-\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
-\vdots & \vdots \\
-\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
-\end{bmatrix}$$
-
-::::
-
-:::
-
-## Connecting Block Models to the GRDPG Model
-
-**Example 1** (cont'd): To perform community detection,
-
-1. Note that $A$ is a RDPG because $P = X X^\top$.
-2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
-$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
-3. Apply a clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
-noting that $\hat{X}$ approaches point masses as $n \to \infty$.
-
-```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
+```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
+n1 <- n / 2
+n2 <- n / 2
+z <- c(rep(1, n1), rep(2, n2))
+p <- 1/2
+q <- 1/4
+r <- 1/8
+P <- matrix(r, nrow = n, ncol = n)
+P[seq(n1), seq(n1)] <- p
+P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
 P.eigen = eigen(P)
@@
 X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
-plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA, 
-     main = 'ASE of the adjacency matrix drawn from SBM')
-points(X, pch = 16, col = z * 2)
+par(mar = rep(1.75, 4))
+plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
 ```
 
-## Connecting Block Models to the GRDPG Model
+\vspace*{.5\baselineskip}
 
-::: columns
+```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
+par(mar = rep(1.75, 4))
+omega <- rbeta(n, 1, 1)
+dc.matrix <- omega %*% t(omega)
+P.dcbm <- P * dc.matrix
+dcbm.eigen <- eigen(P.dcbm)
+X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
+plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')
 
-:::: column
-
-```{r, out.width = '100%', fig.height = 4, fig.width = 4}
-# par(mar = rep(1.75, 4))
-plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
 ```
@@
 
-Submitted to *Journal of Computational and Graphical Statistics* (under review)
+Published in *Journal of Computational and Graphical Statistics*
 

### 2022-07-12 jsm/speed-presentation/slides.tex
slides for jsm

--- 2022-04-11 fls-lightning-talk/slides.tex
@@
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \author[John Koo]{John Koo}
-\ifluatex
+\usepackage{setspace}
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{natbib}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\setcitestyle{numbers,square,comma}
+\usepackage{verbatim}
+\usepackage{amsthm}
+\usepackage{comment}
+\usepackage{graphicx}
+\setbeamertemplate{itemize items}[circle]
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
@@
 
-How might we cluster the nodes of a network?
+\textbf{Def} Popularity Adjusted Block Model (Sengupta and Chen, 2017):
 
-\begin{enumerate}
-\tightlist
-\item
-  Define a probability model with communities that might generate the
-  graph (e.g., popularity adjusted block model).
-\item
-  Develop estimators for the parameters of the probability model,
-  including the community labels.
-\item
-  Describe the properties of the estimators (e.g., consistency).
-\end{enumerate}
-\end{frame}
-
-\begin{frame}{Bernoulli Graphs}
-\protect\hypertarget{bernoulli-graphs}{}
-Let \(G\) be an undirected and unweighted graph with \(n\) vertices.
-
-\(G\) is described by adjacency matrix \(A\) such that
-\(A_{ij} = \begin{cases} 1 & \text{an edge connects vertices } i \text{ and } j \\ 0 & \text{otherwise} \end{cases}\)
-
-\(A_{ji} = A_{ij}\) and \(A_{ii} = 0\).
-
-\vspace*{1\baselineskip}
-
-\(A \sim \text{BernoulliGraph}(P)\) iff:
-
-\begin{enumerate}
-\tightlist
-\item
-  \(P\) is a matrix of edge probabilities between pairs of vertices.
-\item
-  \(A_{ij} \stackrel{\text{ind}}{\sim}\text{Bernoulli}(P_{ij})\) for
-  each \(i < j\).
-\end{enumerate}
-\end{frame}
-
-\begin{frame}{Block Models}
-\protect\hypertarget{block-models}{}
-Suppose each vertex \(v_1, ..., v_n\) has labels
-\(z_1, ..., z_n \in \{1, ..., K\}\),\\
-and each \(P_{ij}\) depends on labels \(z_i\) and \(z_j\).\\
-Then \(A \sim \text{BernoulliGraph}(P)\) is a \emph{block model}.
-
-\textbf{Example 1}: Stochastic Block Model with \(K=2\) communities.
-
-\begin{columns}[T]
-\begin{column}{0.48\textwidth}
-\(P_{ij} = \begin{cases} p & z_i = z_j = 1 \\ q & z_i = z_j = 2 \\ r & z_i \neq z_j \end{cases}\)
-\end{column}
-
-\begin{column}{0.48\textwidth}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1}
-
-}
-
-\caption{SBM with $p=1/2$, $q=1/4$, $r=1/8$}\label{unnamed-chunk-5}
-\end{figure}
-\end{column}
-\end{columns}
+Let each vertex \(i \in [n]\) have \(K\) popularity parameters
+\(\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]\).\\
+Then \(A \sim \text{BernoulliGraph}(P)\) is a PABM if each
+\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\)
 \end{frame}
@@
 
-\begin{frame}{Connecting Block Models to the GRDPG Model}
+\begin{frame}{Connecting Block Models to the (G)RDPG Model}
 \protect\hypertarget{connecting-block-models-to-the-grdpg-model}{}
-All Bernoulli Graphs are GRDPGs.
-
-\textbf{Example 1} (cont'd): SBM with \(K = 2\).
-
 \begin{columns}[T]
 \begin{column}{0.48\textwidth}
-\[P_{ij} = \begin{cases} 
-p & z_i = z_j = 1 \\
-q & z_i = z_j = 2 \\
-r & z_i \neq z_j
-\end{cases}\]
+\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1} \end{center}
 
-\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}
-\end{column}
+\vspace*{.5\baselineskip}
 
-\begin{column}{0.48\textwidth}
-\[P = 
-\begin{bmatrix} 
-P^{(11)} & P^{(12)} \\
-P^{(21)} & P^{(22)}
-\end{bmatrix} =
-X I_{2,0} X^\top\]
-
-\[X = \begin{bmatrix} 
-\sqrt{p} & 0 \\
-\vdots & \vdots \\
-\sqrt{p} & 0 \\
-\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
-\vdots & \vdots \\
-\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
-\end{bmatrix}\]
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Connecting Block Models to the GRDPG Model}
-\protect\hypertarget{connecting-block-models-to-the-grdpg-model-1}{}
-\textbf{Example 1} (cont'd): To perform community detection,
-
-\begin{enumerate}
-\tightlist
-\item
-  Note that \(A\) is a RDPG because \(P = X X^\top\).
-\item
-  Compute the ASE \(A \approx \hat{X} \hat{X}^\top\) with
-  \(\hat{X} = \hat{V} \hat{\Lambda}^{1/2}\).
-\item
-  Apply a clustering algorithm (e.g., \(K\)-means) to \(\hat{X}\),\\
-  noting that \(\hat{X}\) approaches point masses as \(n \to \infty\).
-\end{enumerate}
-
-\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-7-1} \end{center}
-\end{frame}
-
-\begin{frame}{Connecting Block Models to the GRDPG Model}
-\protect\hypertarget{connecting-block-models-to-the-grdpg-model-2}{}
-\begin{columns}[T]
-\begin{column}{0.48\textwidth}
-\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-8-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}
 \end{column}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-7-1} \end{center}
 \end{column}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-8-1} \end{center}
 \end{frame}
@@
 \protect\hypertarget{popularity-adjusted-block-models-are-generalized-random-dot-product-graphs}{}
-Submitted to \emph{Journal of Computational and Graphical Statistics}
-(under review)
+Published in \emph{Journal of Computational and Graphical Statistics}
 

### 2022-07-12 pabm-grdpg.tex
slides for jsm

--- 2022-02-22 summary_minh_edited.tex
@@
 
-The classical Stochastic Block Model (SBM) of \citet{doi:10.1080/0022250X.1971.9989788} specifies that each edge probability $P_{ij}$ depends only on the labels $z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_,z_j}$.
+The classical Stochastic Block Model (SBM) of \citet{doi:10.1080/0022250X.1971.9989788} specifies that each edge probability $P_{ij}$ depends only on the labels $z_i$ and $z_j$, i.e., $P_{ij} = \omega_{z_i,z_j}$.
 Subsequent researchers have weakened this assumption.
-The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_,z_j}$.
+The Degree-Corrected Block Model (DCBM) of \citet{Karrer_2011} assigns an additional parameter $\theta_i$ to each vertex and sets $P_{ij} = \theta_i \theta_j \omega_{z_i,z_j}$.
 The Popularity Adjusted Block Model (PABM) of \citet{307cbeb9b1be48299388437423d94bf1} generalizes the DCBM, allowing heterogeneity of edge probabilities within and between communities while still maintaining distinct community structure.
@@
 
+\section*{Acknowledgements}
+
+This work was partially supported by the Naval Engineering Education Consortium (NEEC), Office of Naval Research (ONR) Award Number N00174-19-1-0011.
+
 \bigskip

### 2022-07-14 jsm/speed-presentation/poster/poster.Rmd
finished poster

---
title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
author:
  - name: John Koo
    affil: 1
  - name: Minh Tang
    affil: 2
  - name: Michael W. Trosset
    affil: 1
affiliation:
  - num: 1
    address: Department of Statistics, Indiana University
  - num: 2
    address: Department of Statistics, North Carolina State University
title_textsize: "56pt"
author_textsize: "32pt"
affiliation_textsize: "8pt"
sectitle_textsize: "36pt"
sectitle2_textsize: "36pt"
body_textsize: "32pt"
column_numbers: 4
poster_height: "20.5in"
poster_width: "36.5in"
column_margins: "0.1in"
primary_colour: "#990000"
secondary_colour: "#000000"
accent_colour: "#990000"
titlebox_borderwidth: "0.1cm"
titlebox_bordercol: "#000000"
title_textcol: "#ffffff"
author_textcol: "#ffffff"
sectitle2_textcol: "#990000"
sectitle_borderwidth: "0mm"
font_family: "Times New Roman"
logoleft_name: poster_files/trident_medium.png
output: 
  posterdown::posterdown_html:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
import::from(magrittr, `%>%`)
library(ggplot2)
library(ggraph)
source('~/dev/pabm-grdpg/functions.R')
```

# Block Models

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newtheorem{theorem}{Theorem}

A *block model* is a random graph model in which each vertex $v_i$ has a community label $z_i \in [K]$ and the edge probability between each pair of vertices is determined in part by the pair of labels.

$A$ is an adjacency matrix of a block model if and only if each $A_{ij} \indep \Bernoulli(P_{ij})$ and $P_{ij} = f(z_i, z_j, \cdot)$.

## Stochastic Block Model

Each edge probability is fixed for each pair of communities, i.e., $P_{ij} = \theta_{z_i, z_j}$.

## Degree Corrected Block Model

Each edge probability is determined by the community edge probability as well as each vertex's degree factor, i.e., $P_{ij} = \theta_{z_i, z_j} \omega_i \omega_j$.

## Popularity Adjusted Block Model

Each vertex has a popularity parameter for each community that determines its affinity toward that community, i.e., $P_{ij} = \lambda_{i, z_j} \lambda_{j, z_i}$.

# Generalized Random Dot Product Graphs

The *generalized random dot product graph* is a random graph model in which each vertex $v_i$ has a corresponding hidden vector $x_i \in \mathbb{R}^{p+q}$ and each edge probability is the indefinite inner product of the corresponding pair of hidden vectors, i.e., $P_{ij} = x_i^\top I_{p, q} x_j$, $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0 \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$

## Adjacency Spectral Embedding

Approximate $A$ by spectral decomposition $A \approx V_{p, q} \Lambda_{p, q} V_{p, q}^\top$. 
The subscript $p, q$ denotes the $p$ most positive and $q$ most negative eigenvalues and corresponding eigenvectors. 
Each $\hat{x}_i$, the $i^{th}$ row of $\hat{X} = V_{p, q} |\Lambda_{p, q}|^{1/2},$ estimates the relative position of its corresponding latent vector $x_i$, up to an indefinite orthogonal transformation.

**Theorem**: $\max_i \| \hat{x}_i - Q_n x_i \| = O_P \Big( \frac{\log^c n}{n^{1/2}} \Big)$ for some $Q_n \in \mathbb{O}(p, q)$.

```{r, cache = TRUE}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)

data.matrix <- cbind(c(x1, x2), c(y1, y2))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

Xhat <- embedding(A, 2, 0)
Xhat[, 1] <- -Xhat[, 1]
```

```{r, fig.height = 2.2, fig.width = 6, out.width = '100%'}
latent.config.plot <- ggplot() + 
  geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, title = 'Latent Vectors') + 
  scale_colour_brewer(palette = 'Set1')

graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'GRDPG') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z))) + 
  # coord_fixed() + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, title = 'ASE') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(latent.config.plot, graph.plot, ase.plot, nrow = 1)
```

# Connecting Block Models to the GRDPG

It has been previously shown that the SBM and DCBM are GRDPGs in which the communities lie on point masses and line segments respectively.

**Theorem** (KTT): The PABM is GRDPG in which the communities lie on mutually orthogonal $K$-dimensional subspaces.

```{r, cache = TRUE, fig.width = 5, fig.height = 2, out.width = '100%'}
n <- 64
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

sbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'Stochastic Block Model') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

sbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 2], y = Xhat[, 1],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 2], y = X[, 1], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Point Masses') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(sbm.graph.plot, sbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 5, fig.height = 2, out.width = '100%'}
n <- 128
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
omega <- rbeta(n, 2, 1)
P <- P * (omega %*% t(omega))
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
Xhat <- -Xhat
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

dcbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'Degree Corrected Block Model') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

dcbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Rays') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(dcbm.graph.plot, dcbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 5, fig.height = 2, out.width = '100%'}
n <- 64
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
A <- draw.graph(P)
Xhat <- embedding(A, 3, 1)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

pabm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'Popularity Adjusted Block Model') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

pabm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Subspaces (Projected)') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(pabm.graph.plot, pabm.ase.plot, nrow = 1)
```

# Orthogonal Spectral Clustering

1. Decompose $A \approx V_{p,q} \Lambda_{p,q} V_{p,q}^\top$, $p = \frac{K (K+1)}{2}$ and $q = \frac{K (K-1)}{2}$.
2. Compute $B = |n V_{p,q} V_{p,q}^\top|$.
3. Construct graph $G'$ from $B$.
4. Partition $G'$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $B$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big(\log^{4c} n\big)$,  
$\max_{i, j} B_{ij} = O_P \Big( \frac{\log^c n}{\sqrt{n \rho_n}} \Big)$.

## Simulation Study

```{r, fig.height = 2, fig.width = 5, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'bottom',
        legend.margin = margin(0, 0, 0, 0)) + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(expression(2^7), 
                           expression(2^8), 
                           expression(2^9), 
                           expression(2^10), 
                           expression(2^11), 
                           expression(2^12))) +
  scale_y_log10() + 
  labs(y = 'error count', 
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'OSC')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'OSC', shape = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC-ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'SSC-ASE', shape = 'SSC-ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'SSC-A')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'SSC-A', shape = 'SSC-A')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'SSC-A'), width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'MM-Louvain')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'MM-Louvain', shape = 'MM-Louvain')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

# Conclusion

Exploiting the geometry of PABMs (and other block models) results in consistent community detection algorithms.

### 2022-07-14 jsm/speed-presentation/slides.Rmd
finished poster

@@
 $\max_{i, j} \hat{B}_{ij} = 
-O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$ 
-as $n \to \infty$.
+O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$.
 
@@
 
-## Simulation Results
+## Simulation Study
 

### 2022-07-14 jsm/speed-presentation/poster/poster.Rmd
minor tweaks

@@
 
-1. Decompose $A \approx V_{p,q} \Lambda_{p,q} V_{p,q}^\top$, $p = \frac{K (K+1)}{2}$ and $q = \frac{K (K-1)}{2}$.
+1. Eigendecompose $A \approx V_{p,q} \Lambda_{p,q} V_{p,q}^\top$ with $p = \frac{K (K+1)}{2}$ and $q = \frac{K (K-1)}{2}$.
 2. Compute $B = |n V_{p,q} V_{p,q}^\top|$.
@@
 Let $B$ be the affinity matrix from OSC. 
-Then $\forall$ pairs $(i, j)$ belonging to different communities 
-and sparsity factor satisfying $n \rho_n = \omega\big(\log^{4c} n\big)$,  
-$\max_{i, j} B_{ij} = O_P \Big( \frac{\log^c n}{\sqrt{n \rho_n}} \Big)$.
+Then $\forall$ $(i, j)$ belonging to different communities and sparsity factor satisfying $n \rho_n = \omega\big(\log^{4c} n\big),$ $\max_{i, j} B_{ij} = O_P \Big( \frac{\log^c n}{\sqrt{n \rho_n}} \Big)$.
 
@@
 
-Exploiting the geometry of PABMs (and other block models) results in consistent community detection algorithms.
+Exploiting the geometry of PABMs (and other block models) results in intuitive and consistent community detection algorithms.

### 2022-07-14 jsm/speed-presentation/slides.tex
finished poster
minor tweaks

@@
 \textbf{Theorem}:
-\(\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)\)
-as \(n \to \infty\)
+\(\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)\).
 \end{frame}
 
-\begin{frame}{Connecting Block Models to the (G)RDPG Model}
+\begin{frame}{Connecting Block Models to the GRDPG Model}
 \protect\hypertarget{connecting-block-models-to-the-grdpg-model}{}
@@
 \(n \rho_n = \omega\big((\log n)^{4c}\big)\),\\
-\(\max_{i, j} \hat{B}_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\)
-as \(n \to \infty\).
+\(\max_{i, j} \hat{B}_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\).
 
@@
 
-\begin{frame}{Simulation Results}
-\protect\hypertarget{simulation-results}{}
+\begin{frame}{Simulation Study}
+\protect\hypertarget{simulation-study}{}
 We compare four algorithms for community detection on randomly generated

### 2022-07-19 jsm/speed-presentation/poster/poster.Rmd
changes to poster based on meeting

@@
 
-## Popularity Adjusted Block Model
+## Popularity Adjusted Block Model (Sengupta and Chen 2018)
 
@@
 
-**Theorem**: $\max_i \| \hat{x}_i - Q_n x_i \| = O_P \Big( \frac{\log^c n}{n^{1/2}} \Big)$ for some $Q_n \in \mathbb{O}(p, q)$.
+**Theorem** (Rubin-Delanchy et al. 2022): $\max_i \| \hat{x}_i - Q_n x_i \| = O_P \Big( \frac{\log^c n}{n^{1/2}} \Big)$ for some $Q_n \in \mathbb{O}(p, q)$.
 
@@
   labs(colour = NULL, 
-       title = 'Stochastic Block Model') + 
+       title = 'SBM') + 
   # theme_graph() + 
@@
   labs(colour = NULL, 
-       title = 'Degree Corrected Block Model') + 
+       title = 'DCBM') + 
   # theme_graph() + 
@@
   labs(colour = NULL, 
-       title = 'Popularity Adjusted Block Model') + 
+       title = 'PABM') + 
   # theme_graph() + 
@@
 
-1. Eigendecompose $A \approx V_{p,q} \Lambda_{p,q} V_{p,q}^\top$ with $p = \frac{K (K+1)}{2}$ and $q = \frac{K (K-1)}{2}$.
-2. Compute $B = |n V_{p,q} V_{p,q}^\top|$.
-3. Construct graph $G'$ from $B$.
-4. Partition $G'$ into $K$ disconnected subgraphs.
+1. Set $p = \frac{K (K+1)}{2}$, $q = \frac{K (K-1)}{2}$ and decompose $A \approx V_{p,q} \Lambda_{p,q} V_{p,q}^\top$.
+2. Construct graph $G'$ from $B = |n V_{p,q} V_{p,q}^\top|$.
+3. Partition $G'$ into $K$ disconnected subgraphs.
 
 **Theorem** (KTT): 
-Let $B$ be the affinity matrix from OSC. 
-Then $\forall$ $(i, j)$ belonging to different communities and sparsity factor satisfying $n \rho_n = \omega\big(\log^{4c} n\big),$ $\max_{i, j} B_{ij} = O_P \Big( \frac{\log^c n}{\sqrt{n \rho_n}} \Big)$.
+$\max\limits_{i, j} B_{ij} = O_P \Big( \frac{\log^c n}{\sqrt{n \rho_n}} \Big)$ for each $(i, j)$ in different communities.
+
+**Corollary**: OSC results in zero community detection error with probability 1 as $n \to \infty$.
 
@@
 
-```{r, fig.height = 2, fig.width = 5, out.width = '100%'}
+```{r, fig.height = 2.25, fig.width = 5, out.width = '100%'}
 clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')

### 2022-07-19 jsm/speed-presentation/slides.Rmd
minor tweaks
updated slides based on meeting

@@
 library(ggplot2)
+library(ggraph)
 import::from(magrittr, `%>%`)
@@
 
-Michael Trosset,  
+Michael W. Trosset,  
 Professor of Statistics,  
@@
 
-## Popularity Adjusted Block Model
-
-**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):
-
-Let each vertex $i \in [n]$ have $K$ popularity parameters
-$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
-Then $A \sim \PABM(\{\lambda_{ik}\}_K)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
-
-**Def** (Noroozi, Rimal, and Pensky, 2020):
-
-$A$ is sampled from a PABM if $P$ can be described as:
-
-1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
-between communities $k$ and $l$. 
-2. Organize popularity parameters as vectors 
-$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
-such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
-of the $i$^th^ vertex of community $k$ towards community $l$. 
-3. Each block can be decomposed as 
-$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
-
-## Generalized Random Dot Product Graph
-
-**Def** Generalized Random Dot Product Graph  
-(Rubin-Delanchy, Cape, Tang, Priebe, 2020)
-
-Let $I_{p,q} = \blockdiag(I_p, -I_q)$ and suppose that $x_1, \ldots, x_n \in \mathbb{R}^{p+q}$ are such that $x_i^\top I_{p,q} x_j \in [0,1]$.  
-Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
-
-Adjacency Spectral Embedding (Sussman et al., 2012) estimates $x_1, ..., x_n \in \mathbb{R}^{p+q}$ from $A$:
-
-1. Let $\hat{\Lambda}$ be the diagonal matrix that contains the absolute values of the $p$ most positive and the $q$ most negative eigenvalues.
-2. Let $\hat{V}$ be the matrix whose columns are the corresponding eigenvectors.
-3. Compute $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
-
-**Theorem**: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$ 
-as $n \to \infty$
-
-## Connecting Block Models to the (G)RDPG Model
+## Connecting Block Models to the GRDPG
 
@@
 
-:::: column
+:::: {.column width='67%'}
 
-```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
+```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
+n <- 64
 n1 <- n / 2
@@
 p <- 1/2
-q <- 1/4
+q <- 1/2
 r <- 1/8
@@
 P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
-P.eigen = eigen(P)
-X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
+X <- embedding(P, 2, 0)
+A <- draw.graph(P)
+Xhat <- embedding(A, 2, 0)
+A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)
 
-A.eigen <- eigen(A)
-X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
-par(mar = rep(1.75, 4))
-plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
+sbm.graph.plot <- ggraph(A, layout = 'stress') + 
+  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
+  geom_node_point(aes(colour = factor(z)),
+                  show.legend = FALSE) + 
+  labs(colour = NULL, 
+       title = 'SBM') + 
+  # theme_graph() + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+sbm.ase.plot <- ggplot() + 
+  geom_point(aes(x = Xhat[, 2], y = Xhat[, 1],
+                 colour = factor(z)),
+             shape = 4) + 
+  geom_point(aes(x = X[, 2], y = X[, 1], 
+                 colour = factor(z))) + 
+  # coord_fixed() +
+  theme_bw() + 
+  theme(legend.position = 'none') + 
+  labs(x = NULL, y = NULL, colour = NULL, 
+       title = 'Point Masses') + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(sbm.graph.plot, sbm.ase.plot, nrow = 1)
 ```
 
-\vspace*{.5\baselineskip}
+```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
+n <- 128
+n1 <- n / 2
+n2 <- n / 2
+z <- c(rep(1, n1), rep(2, n2))
+p <- 1/2
+q <- 1/2
+r <- 1/8
+P <- matrix(r, nrow = n, ncol = n)
+P[seq(n1), seq(n1)] <- p
+P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
+omega <- rbeta(n, 2, 1)
+P <- P * (omega %*% t(omega))
+X <- embedding(P, 2, 0)
+A <- draw.graph(P)
+Xhat <- embedding(A, 2, 0)
+# Xhat <- -Xhat
+A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)
 
-```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
-par(mar = rep(1.75, 4))
-omega <- rbeta(n, 1, 1)
-dc.matrix <- omega %*% t(omega)
-P.dcbm <- P * dc.matrix
-dcbm.eigen <- eigen(P.dcbm)
-X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
-plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')
+dcbm.graph.plot <- ggraph(A, layout = 'stress') + 
+  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
+  geom_node_point(aes(colour = factor(z)),
+                  show.legend = FALSE) + 
+  labs(colour = NULL, 
+       title = 'DCBM') + 
+  # theme_graph() + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
 
+dcbm.ase.plot <- ggplot() + 
+  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
+                 colour = factor(z)),
+             shape = 4) + 
+  geom_point(aes(x = X[, 1], y = X[, 2], 
+                 colour = factor(z))) + 
+  # coord_fixed() +
+  theme_bw() + 
+  theme(legend.position = 'none') + 
+  labs(x = NULL, y = NULL, colour = NULL, 
+       title = 'Rays') + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(dcbm.graph.plot, dcbm.ase.plot, nrow = 1)
+```
+
+```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
+n <- 64
+Pz <- generate.P.beta(n)
+P <- Pz$P
+z <- Pz$clustering
+X <- embedding(P, 3, 1)
+A <- draw.graph(P)
+Xhat <- embedding(A, 3, 1)
+Xhat[, 1] <- -Xhat[, 1]
+A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)
+
+pabm.graph.plot <- ggraph(A, layout = 'stress') + 
+  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
+  geom_node_point(aes(colour = factor(z)),
+                  show.legend = FALSE) + 
+  labs(colour = NULL, 
+       title = 'PABM') + 
+  # theme_graph() + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+pabm.ase.plot <- ggplot() + 
+  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
+                 colour = factor(z)),
+             shape = 4) + 
+  geom_point(aes(x = X[, 1], y = X[, 2], 
+                 colour = factor(z))) + 
+  # coord_fixed() +
+  theme_bw() + 
+  theme(legend.position = 'none') + 
+  labs(x = NULL, y = NULL, colour = NULL, 
+       title = 'Subspaces (Projected)') + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(pabm.graph.plot, pabm.ase.plot, nrow = 1)
 ```
@@
 
-:::: column
+:::: {.column width='33%'}
 
-\vspace*{0\baselineskip}
+&nbsp;
 
-```{r, out.width = '100%', fig.height = 4, fig.width = 4}
-par(mar = rep(1.75, 4))
-Pz <- generate.P.beta(n * 10)
-P <- Pz$P
-z <- Pz$clustering
-X <- embedding(P, 3, 1)
-pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
-```
+&nbsp;
+
+* K-means clustering
+* Gaussian mixture models
+
+&nbsp;
+
+* K-means with cosine similarity
+* GMM on angles
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+* ???
 
@@
 :::
-
-## Connecting the PABM to the GRDPG
-
-**Theorem** (KTT): $A \sim \PABM(\{\lambda_{ik}\}_K)$ is equivalent to 
-$A \sim \GRDPG_{p, q}(X U)$ with
-
-* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
-* $U$ is an orthogonal matrix;
-* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
-composed of popularity vectors with each block corresponding to a community.
-
-$$X = \begin{bmatrix}
-\Lambda^{(1)} & \cdots & 0 \\
-0 & \ddots & 0 \\
-0 & \cdots & \Lambda^{(K)}
-\end{bmatrix} 
-\in \mathbb{R}^{n \times K^2}$$
-
-$$\Lambda^{(k)} = \begin{bmatrix} 
-\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
-\end{bmatrix} 
-\in \mathbb{R}^{n_k \times K}$$
-
-$$A \sim \PABM(\{\lambda_{ik}\}_K) \text{ iff } A \sim \GRDPG_{p, q}(X U)$$
 
@@
 **Theorem** (KTT): 
-Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC. 
-Then $\forall$ pairs $(i, j)$ belonging to different communities 
-and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,  
-$\max_{i, j} \hat{B}_{ij} = 
-O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$.
+$\forall$ pairs $(i, j)$ belonging to different communities, 
+$\max\limits_{i, j} B_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$.
 
@@
 ## Simulation Study
-
-We compare four algorithms for community detection on randomly generated PABMs:
-
-* Modularity Maximization (Sengupta and Chen) using the Louvain algorithm;
-* Orthogonal Spectral Clustering (KTT);
-* Sparse Subspace Clustering on the columns of $A$  
-(Noorozi, Rimal, Pensky); 
-* Sparse Subspace Clustering on the ASE (KTT).
 

### 2022-07-19 jsm/speed-presentation/slides.tex
updated slides based on meeting

@@
 
-Michael Trosset,\\
+Michael W. Trosset,\\
 Professor of Statistics,\\
@@
 
-\begin{frame}{Popularity Adjusted Block Model}
-\protect\hypertarget{popularity-adjusted-block-model}{}
-\textbf{Def} Popularity Adjusted Block Model (Sengupta and Chen, 2017):
+\begin{frame}{Connecting Block Models to the GRDPG}
+\protect\hypertarget{connecting-block-models-to-the-grdpg}{}
+\begin{columns}[T]
+\begin{column}{0.67\textwidth}
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1} \end{center}
 
-Let each vertex \(i \in [n]\) have \(K\) popularity parameters
-\(\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]\). Then
-\(A \sim \text{PABM}(\{\lambda_{ik}\}_K)\) if each
-\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\).
-
-\textbf{Def} (Noroozi, Rimal, and Pensky, 2020):
-
-\(A\) is sampled from a PABM if \(P\) can be described as:
-
-\begin{enumerate}
-\tightlist
-\item
-  Let each \(P^{(kl)}\) denote the \(n_k \times n_l\) matrix of edge
-  probabilities between communities \(k\) and \(l\).
-\item
-  Organize popularity parameters as vectors
-  \(\lambda^{(kl)} \in \mathbb{R}^{n_k}\) such that
-  \(\lambda^{(kl)}_i = \lambda_{k_i l}\) is the popularity parameter of
-  the \(i\)\textsuperscript{th} vertex of community \(k\) towards
-  community \(l\).
-\item
-  Each block can be decomposed as
-  \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\).
-\end{enumerate}
-\end{frame}
-
-\begin{frame}{Generalized Random Dot Product Graph}
-\protect\hypertarget{generalized-random-dot-product-graph}{}
-\textbf{Def} Generalized Random Dot Product Graph\\
-(Rubin-Delanchy, Cape, Tang, Priebe, 2020)
-
-Let \(I_{p,q} = \text{blockdiag}(I_p, -I_q)\) and suppose that
-\(x_1, \ldots, x_n \in \mathbb{R}^{p+q}\) are such that
-\(x_i^\top I_{p,q} x_j \in [0,1]\).\\
-Then \(A \sim \text{GRDPG}_{p, q}(X)\) iff
-\(A \sim \text{BernoulliGraph}(X I_{p,q} X^\top)\), where
-\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).
-
-Adjacency Spectral Embedding (Sussman et al., 2012) estimates
-\(x_1, ..., x_n \in \mathbb{R}^{p+q}\) from \(A\):
-
-\begin{enumerate}
-\tightlist
-\item
-  Let \(\hat{\Lambda}\) be the diagonal matrix that contains the
-  absolute values of the \(p\) most positive and the \(q\) most negative
-  eigenvalues.
-\item
-  Let \(\hat{V}\) be the matrix whose columns are the corresponding
-  eigenvectors.
-\item
-  Compute \(\hat{X} = \hat{V} \hat{\Lambda}^{1/2}\).
-\end{enumerate}
-
-\textbf{Theorem}:
-\(\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)\).
-\end{frame}
-
-\begin{frame}{Connecting Block Models to the GRDPG Model}
-\protect\hypertarget{connecting-block-models-to-the-grdpg-model}{}
-\begin{columns}[T]
-\begin{column}{0.48\textwidth}
-\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-5-1} \end{center}
-
-\vspace*{.5\baselineskip}
-
-\begin{center}\includegraphics[width=0.5\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}
-\end{column}
-
-\begin{column}{0.48\textwidth}
-\vspace*{0\baselineskip}
+\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-6-1} \end{center}
 
@@
 \end{column}
-\end{columns}
-\end{frame}
 
-\begin{frame}{Connecting the PABM to the GRDPG}
-\protect\hypertarget{connecting-the-pabm-to-the-grdpg}{}
-\textbf{Theorem} (KTT): \(A \sim \text{PABM}(\{\lambda_{ik}\}_K)\) is
-equivalent to \(A \sim \text{GRDPG}_{p, q}(X U)\) with
+\begin{column}{0.33\textwidth}
+~
+
+~
 
@@
 \item
-  \(p = K (K + 1) / 2\), \(q = K (K - 1) / 2\);
+  K-means clustering
 \item
-  \(U\) is an orthogonal matrix;
-\item
-  \(X \in \mathbb{R}^{n \times K^2}\) is a block diagonal matrix
-  composed of popularity vectors with each block corresponding to a
-  community.
+  Gaussian mixture models
 \end{itemize}
 
-\[X = \begin{bmatrix}
-\Lambda^{(1)} & \cdots & 0 \\
-0 & \ddots & 0 \\
-0 & \cdots & \Lambda^{(K)}
-\end{bmatrix} 
-\in \mathbb{R}^{n \times K^2}\]
+~
 
-\[\Lambda^{(k)} = \begin{bmatrix} 
-\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
-\end{bmatrix} 
-\in \mathbb{R}^{n_k \times K}\]
+\begin{itemize}
+\tightlist
+\item
+  K-means with cosine similarity
+\item
+  GMM on angles
+\end{itemize}
 
-\[A \sim \text{PABM}(\{\lambda_{ik}\}_K) \text{ iff } A \sim \text{GRDPG}_{p, q}(X U)\]
+~
+
+~
+
+~
+
+\begin{itemize}
+\tightlist
+\item
+  ???
+\end{itemize}
+\end{column}
+\end{columns}
 \end{frame}
@@
 
-\textbf{Theorem} (KTT): Let \(\hat{B}\) with entries \(\hat{B}_{ij}\) be
-the affinity matrix from OSC. Then \(\forall\) pairs \((i, j)\)
-belonging to different communities and sparsity factor satisfying
-\(n \rho_n = \omega\big((\log n)^{4c}\big)\),\\
-\(\max_{i, j} \hat{B}_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\).
+\textbf{Theorem} (KTT): \(\forall\) pairs \((i, j)\) belonging to
+different communities,
+\(\max\limits_{i, j} B_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\).
 
@@
 \protect\hypertarget{simulation-study}{}
-We compare four algorithms for community detection on randomly generated
-PABMs:
-
-\begin{itemize}
-\tightlist
-\item
-  Modularity Maximization (Sengupta and Chen) using the Louvain
-  algorithm;
-\item
-  Orthogonal Spectral Clustering (KTT);
-\item
-  Sparse Subspace Clustering on the columns of \(A\)\\
-  (Noorozi, Rimal, Pensky);
-\item
-  Sparse Subspace Clustering on the ASE (KTT).
-\end{itemize}
-
 \begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/unnamed-chunk-8-1} \end{center}

### 2022-07-19 jsm/speed-presentation/slides.tex
minor change

@@
 
-\begin{frame}{Popularity Adjusted Block Models are Generalized Random
-Dot Product Graphs}
-\protect\hypertarget{popularity-adjusted-block-models-are-generalized-random-dot-product-graphs}{}
+\begin{frame}{Thank you}
+\protect\hypertarget{thank-you}{}
 Published in \emph{Journal of Computational and Graphical Statistics}

### 2022-07-19 jsm/speed-presentation/slides.Rmd
minor change
changed simulation results slide

@@
 
-```{r, fig.height = 3, fig.width = 8, out.width = '100%'}
+```{r, fig.height = 3, fig.width = 5, out.width = '100%'}
 clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')
@@
   theme_bw() + 
-  theme(legend.position = 'bottom') + 
+  theme(legend.position = 'top',
+        legend.direction = 'vertical',
+        legend.justification = 0) + 
   scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
@@
   scale_y_log10() + 
-  labs(y = 'community detection error count', 
+  labs(y = 'error count', 
        colour = NULL, shape = NULL) +
   geom_line(aes(x = n, y = med.err * n,
-                colour = 'OSC')) +
+                colour = 'Orthogonal Spectral Clustering')) +
   geom_point(aes(x = n, y = med.err * n,
-                 colour = 'OSC', shape = 'OSC')) +
+                 colour = 'Orthogonal Spectral Clustering', 
+                 shape = 'Orthogonal Spectral Clustering')) +
   geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
-                    colour = 'OSC'), width = .1) + 
+                    colour = 'Orthogonal Spectral Clustering'), width = .1) + 
   geom_line(aes(x = n, y = med.err.ssc * n,
-                colour = 'SSC-ASE')) +
+                colour = 'Sparse Subspace Clustering on ASE')) +
   geom_point(aes(x = n, y = med.err.ssc * n,
-                 colour = 'SSC-ASE', shape = 'SSC-ASE')) +
+                 colour = 'Sparse Subspace Clustering on ASE', 
+                 shape = 'Sparse Subspace Clustering on ASE')) +
   geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
-                    colour = 'SSC-ASE'), width = .1) +
+                    colour = 'Sparse Subspace Clustering on ASE'), width = .1) +
   geom_line(aes(x = n, y = med.err.ssc.A * n,
-                colour = 'SSC-A')) +
+                colour = 'Sparse Subspace Clustering on Adj. Matrix')) +
   geom_point(aes(x = n, y = med.err.ssc.A * n,
-                 colour = 'SSC-A', shape = 'SSC-A')) +
+                 colour = 'Sparse Subspace Clustering on Adj. Matrix', 
+                 shape = 'Sparse Subspace Clustering on Adj. Matrix')) +
   geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
-                    colour = 'SSC-A'), width = .1) +
+                    colour = 'Sparse Subspace Clustering on Adj. Matrix'), 
+                width = .1) +
   geom_line(aes(x = n, y = med.err.louvain * n,
-                colour = 'MM-Louvain')) + 
+                colour = 'Modularity Maximization')) + 
   geom_point(aes(x = n, y = med.err.louvain * n,
-                 colour = 'MM-Louvain', shape = 'MM-Louvain')) + 
+                 colour = 'Modularity Maximization', 
+                 shape = 'Modularity Maximization')) + 
   geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
-                    colour = 'MM-Louvain'), width = .1) + 
+                    colour = 'Modularity Maximization'), width = .1) + 
   scale_colour_brewer(palette = 'Set1') + 
@@
 
-## Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
+## Thank you
 

### 2022-07-20 jsm/speed-presentation/poster/poster.Rmd
fixed minor layout issues

@@
 
-```{r, fig.height = 2.2, fig.width = 6, out.width = '100%'}
+```{r, fig.height = 2.3, fig.width = 6, out.width = '100%'}
 latent.config.plot <- ggplot() + 
@@
 
-```{r, cache = TRUE, fig.width = 5, fig.height = 2, out.width = '100%'}
+```{r, cache = TRUE, fig.width = 5, fig.height = 2.05, out.width = '100%'}
 n <- 64

### 2022-07-20 jsm/speed-presentation/poster/poster.Rmd
trying to include tex file

@@
     self_contained: true
+    keep_tex: true
 ---

### 2022-08-03 jsm/speed-presentation/poster/poster.Rmd
minor adjustments based on discussion

@@
 
-**Corollary**: OSC results in zero community detection error with probability 1 as $n \to \infty$.
+**Corollary**: OSC results in zero clustering error almost surely as $n \to \infty$.
 
@@
   scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
-                labels = c(expression(2^7), 
-                           expression(2^8), 
-                           expression(2^9), 
-                           expression(2^10), 
-                           expression(2^11), 
-                           expression(2^12))) +
+                labels = c(7, 8, 9, 10, 11, 12)) +
   scale_y_log10() + 
   labs(y = 'error count', 
+       x = expression(log[2](n)), 
        colour = NULL, shape = NULL) +

### 2022-08-03 jsm/speed-presentation/slides.Rmd
minor adjustments based on discussion

@@
 John Koo,  
-PhD Student in Statistical Science,  
+PhD Student in Statistics,  
 Indiana University
@@
 $\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$.
+
 Then $A \sim \BG(P)$ is a PABM if each 
-$P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
+$P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.
 
@@
 **Theorem** (KTT): 
-If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
-then $B_{ij} = 0$ if $z_i \neq z_j$.
+If $P = V \Lambda V^\top$, $B = n V V^\top$, and $z_i \neq z_j$,  
+then $B_{ij} = 0$.
 
@@
 positive and $K (K-1) / 2$ most negative eigenvalues.
-2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
+2. Compute $B = |n V V^\top|$, applying $|\cdot|$ entry-wise.
 3. Construct graph $G$ using $B$ as its similarity matrix.
@@
 **Theorem** (KTT): 
-$\forall$ pairs $(i, j)$ belonging to different communities, 
+For all pairs $(i, j)$ belonging to different communities, 
 $\max\limits_{i, j} B_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$.
 
-**Corollary**: OSC results in zero clustering error as $n \to \infty$, with probability 1.
+**Corollary**: OSC results in zero clustering error almost surely as $n \to \infty$.
 
@@
   scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
-                labels = c(expression(2^7), 
-                           expression(2^8), 
-                           expression(2^9), 
-                           expression(2^10), 
-                           expression(2^11), 
-                           expression(2^12))) +
+                labels = c(7, 8, 9, 10, 11, 12)) +
   scale_y_log10() + 
   labs(y = 'error count', 
+       x = expression(log[2] (n)),
        colour = NULL, shape = NULL) +
@@
 
-R package: https://github.com/johneverettkoo/osc
+R package (WIP): https://github.com/johneverettkoo/osc

### 2022-08-03 jsm/speed-presentation/slides.tex
minor adjustments based on discussion

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \newif\ifbibliography
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 John Koo,\\
-PhD Student in Statistical Science,\\
+PhD Student in Statistics,\\
 Indiana University
@@
 Let each vertex \(i \in [n]\) have \(K\) popularity parameters
-\(\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]\).\\
+\(\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]\).
+
 Then \(A \sim \text{BernoulliGraph}(P)\) is a PABM if each
-\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\)
+\(P_{ij} = \lambda_{i z_j} \lambda_{j z_i}\).
 \end{frame}
@@
 \protect\hypertarget{orthogonal-spectral-clustering}{}
-\textbf{Theorem} (KTT): If \(P = V \Lambda V^\top\) and
-\(B = n V V^\top\),\\
-then \(B_{ij} = 0\) if \(z_i \neq z_j\).
+\textbf{Theorem} (KTT): If \(P = V \Lambda V^\top\), \(B = n V V^\top\),
+and \(z_i \neq z_j\),\\
+then \(B_{ij} = 0\).
 
@@
 \item
-  Compute \(B = |n V V^\top|\) applying \(|\cdot|\) entry-wise.
+  Compute \(B = |n V V^\top|\), applying \(|\cdot|\) entry-wise.
 \item
@@
 
-\textbf{Theorem} (KTT): \(\forall\) pairs \((i, j)\) belonging to
-different communities,
+\textbf{Theorem} (KTT): For all pairs \((i, j)\) belonging to different
+communities,
 \(\max\limits_{i, j} B_{ij} = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)\).
 
-\textbf{Corollary}: OSC results in zero clustering error as
-\(n \to \infty\), with probability 1.
+\textbf{Corollary}: OSC results in zero clustering error almost surely
+as \(n \to \infty\).
 \end{frame}
@@
 
-R package: \url{https://github.com/johneverettkoo/osc}
+R package (WIP): \url{https://github.com/johneverettkoo/osc}
 \end{frame}
